### 2019-06-05 write-up/fda-recurrent.tex
Updated bibtex and expanded introduction

\documentclass[11pt]{amsart}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
\usepackage[stretch=10]{microtype}
%\usepackage{hyperref}
\def\struckint{\mathop{%
\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
 {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
\mathpalette
{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
\halign\bgroup\hfill$}
{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
}\limits}
\usepackage{natbib}
\usepackage{color}
\usepackage{booktabs,caption,fixltx2e}
\usepackage[flushleft]{threeparttable}

\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
%\usepackage[stretch=10]{microtype}
%\usepackage{hyperref}
%\def\struckint{\mathop{%
%\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
% {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
%\mathpalette
%{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
%\halign\bgroup\hfill$}
%{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
%}\limits}

\newcommand{\sam}[1]{{\color{blue}{#1}}}
\newcommand{\walt}[1]{\textcolor{red}{[WD:\ #1]}}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{-0.25in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
%\setlength{\evensidemargin}{.25in}
%\setlength{\oddsidemargin}{.25in}

\usepackage{setspace}
\doublespacing

\usepackage{soul,color, mathtools}

\def\pr{\mathop{\text{pr}}\nolimits}

\def\Bka{{\it Biometrika}}
\def\AIC{\textsc{aic}}
\def\T{{ \mathrm{\scriptscriptstyle T} }}
\def\v{{\varepsilon}}
\def\partitionsn{\mathop{\mathcal{P}_{[n]}}\nolimits}
\def\partitionsN{\mathop{\mathcal{P}_{\infty}}\nolimits}
\def\Dmn{\mathop{{D}_{m,n}}\nolimits}
\def\symmetricn{\mathop{\mathcal{S}_{n}}\nolimits}
\def\PE{\mathop{\rm Pitman\mbox{-}Ewens}\nolimits}
\def\per{\mathop{\rm per}\nolimits}
\def\U{\mathop{\mathcal{U}_{}}\nolimits}
\def\Nb{\mathop{\mathbb{N}_{}}\nolimits}
\def\Nbb{\mathop{\mathbf{N}_{}}\nolimits}
\def\nbb{\mathop{\mathbf{n}_{}}\nolimits}
\def\Xbb{\mathop{\mathbf{X}_{}}\nolimits}
\def\fin{\mathop{\text{fin}}\nolimits}
\def\pr{\mathop{\text{pr}}\nolimits}
\def\E{\mathcal{E}}
\def\G{\mathcal{G}}
\def\deg{\text{deg}_{}}
\def\equalinlaw{=_{\mathcal{D}}}
\def\Fk{\mathop{\mathcal{F}_k^{\downarrow}}\nolimits}
\def\F{\mathop{\mathcal{F}^{\downarrow}}\nolimits}
\def\size{\mathop{\text{size}}\nolimits}
\def\Ycong{\mathop{\EY}\nolimits}
\def\EIsig{\mathop{\mathcal{E}_{\mathcal{I}}^{\sigma}}\nolimits}
\def\EY{\mathop{\mathcal{E}_{Y}}\nolimits}
\def\finp{\mathop{\fin(\mathcal{P})}\nolimits}
\def\En{\mathop{\mathfrak{E}_{[n]}}\nolimits}
\def\EN{\mathop{\mathfrak{E}_{\Nb}}\nolimits}
\def\logit{\mathop{\text{logit}_{}}\nolimits}
\def\EI{\mathop{\mathcal{E}_{\mathcal{I}}}\nolimits}
\def\ES{\mathop{\mathfrak{E}_S}\nolimits}
\def\Ifk{\mathop{{I}_{}}\nolimits}
%\def\mathcal{I}{\mathop{\mathcal{I}_{}}\nolimits}
\def\Ycong{\mathop{\EY}\nolimits}
\def\EIsig{\mathop{\mathcal{E}_{\mathcal{I}}^{\sigma}}\nolimits}
\def\EY{\mathop{\mathcal{E}_{Y}}\nolimits}
\def\finp{\mathop{\fin(\mathcal{P})}\nolimits}
\def\En{\mathop{\mathfrak{E}_{[n]}}\nolimits}
\def\EN{\mathop{\mathfrak{E}_{\Nb}}\nolimits}
\def\logit{\mathop{\text{logit}_{}}\nolimits}
\def\EI{\mathop{\mathcal{E}_{\mathcal{I}}}\nolimits}
\def\ES{\mathop{\mathfrak{E}_S}\nolimits}

\DeclareRobustCommand{\citeext}[1]{\citeauthor{#1} (\citeyear{#1})}
\DeclareRobustCommand{\citeint}[1]{(\citeauthor{#1}, \citeyear{#1})}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{defn}[thm]{Definition}
\newtheorem{example}[thm]{Example}
\newtheorem{meas}[thm]{Measurement}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{assumption}[thm]{Assumption}
\newtheorem{rmk}[thm]{Remark}%\endlocaldefs

% change default numbering for enumerate environment to be in parentheses

\makeatletter

\def\S{\mathcal{S}}
\def\indep{\mathrel{\rlap{$\perp$}\kern1.6pt\mathord{\perp}}}
\def\R{\mathcal{R}}
\def\H{\mathcal{H}}
\def\E{\mathbb{E}}
\def\Y{{\bf Y}}
\def\Cov{\text{Cov}}
\def\one{{\bf 1}}
\def\diag{\text{diag}}
\def\given{\, | \,}
\def\Given{\, \big | \,}
\def\Nat{\mathbb{N}}
\def\Real{\mathbb{R}}
\def\bft{{\bf t}}
\def\bfx{{\bf x}}
\def\bfp{{\bf p}}
\def\bfT{{\bf T}}
\def\dotminussym#1#2{%
  \setbox0=\hbox{$\m@th#1-$}%
  \kern.5\wd0%
  \hbox to 0pt{\hss\hbox{$\m@th#1-$}\hss}%
  \raise.6\ht0\hbox to 0pt{\hss$\m@th#1.$\hss}%
  \kern.5\wd0}
\newcommand{\dotminus}{\mathbin{\mathpalette\dotminussym{}}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\mathchardef\mhyphen="2D

% display breaks

\begin{document}

\title{Recurrent event analysis in the presence of functional
  covariates via random subsampling}
\author{Walter Dempsey}
\address {Department of Statistics, 
  Harvard University, One Oxford Street
   Cambridge, MA  02138, USA}
 \email{wdempsey@fas.harvard.edu}

\date{\today}

\begin{abstract}
Digital monitoring studies collect real-time high frequency data via
mobile sensors in the subjects' natural environment.  
This data can be used to model the impact of changes in physiology on
recurrent event outcomes such as smoking, drug use, alcohol use, or
self-identified moments of suicide ideation. 
Likelihood calculations for the recurrent event analysis, however,
become computationally prohibitive in this setting.  
% Unlike classic joint models, event risk is a function of
% recent sensor data, not simply current sensor values.  
Motivated by this, a random subsampling framework is proposed for
computationally efficient estimation of the effect of time-varying
covariates on the recurrent event process. A subsampling-unbiased
estimator for the cumulative hazard enters into an approximation of
log-likelihood. The proposed estimator then maximizes this
approximation. The estimator has two sources of variation: the first
due to the recurrent event model and the second due to subsampling.
The latter can be reduced by increasing the sampling rate; however,
this leads to increased computational costs. 
Simulations demonstrate the method and efficiency-computation
tradeoff. We then illustrate the approach using data from a digital
monitoring study of suicide ideation.
\end{abstract}

\keywords{recurrent events; probabilistic subsampling; estimating equations;
high frequency time series; suicidal ideation}

\maketitle

\section{Introduction}

Advancement in mobile technology has led to the rapid integration of
mobile and wearable sensors into behavioral health~\citep{Freeetal2013}.
Take HeartSteps, for example, a mobile health (mHealth) study designed
to increase physical activity in sedentary adults
\citep{KlasnjaHS2019}. Here, a Jawbone sensor is used to monitor step
count every minute of the participant's study day.
Of interest in many mHealth studies is the relation of such
real-time high frequency sensor data to an adverse, recurrent event
process. In the smoking cessation mHealth study~\citep{}, for
example, the relation between a time-varying sensor-based measure of
physiological stress and the smoking lapse process is of interest.
In a suicidal ideation mHealth study~\citep{}, the relation of
electrodermal activity (EDA) and accelerometer with self-identified
moments of suicidal ideation is of interest.

The goal of this paper is two-fold: (1) to discuss the appropriate
choice of a statistical model for joint high-frequency sensor data and
the recurrent event data, and (2) to construct a robust,
easy-to-implement method for parameter estimation and inference. 
For (1), we discuss an important issue regarding measurement-error
models when paired with recurrent event outcomes.
For (2), we introduce a random subsampling procedure that has
several benefits.  First, the resulting inference is unbiased;
however, there is a computation-efficiency trade-off. In particular, a
higher sampling rate can decrease variance at the cost of increased
computation.  We show via simulations that the benefits of incredibly
high sampling rates is often negligible, as the contribution to the
variation is small in relation to to variation in the underlying
stochastic processes. Second, the subsampling procedure allows
inference to not depend on choice of dynamics for the health process.
Third, optimal estimating equations are derived that depend on the
instantaneous rate of the recurrent event process, which increase
increase efficiency of the subsampling procedure. Finally,
implementation can leverage existing software for functional data
analysis, leading to fast adoption by domain scientists. We show how
to implement methods via standard mixed effects software.

\section{Related prior work}

\section{Recurrent event process and associated high frequency data}

Suppose $n$ subjects are independently sampled with observed event
times~$\bfT_{i} = \{ T_{i,1}, \ldots, T_{i,k_i}\}$ over some observation
window $[0, \tau_i]$ for each subject $i = 1,\cdots, n$.  Assume the
event times are ordered, i.e., $T_{i,j} < T_{i,j^\prime}$ for $j <
j^\prime$. The window length, $\tau_i$, is the censoring time and is
assumed independent of the event process. Let~$N_{i} (t)$ denote the
associated counting process of $\bfT_{i}$; that is, $N_i (t) =
\sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.  
In this section, we assume a single-dimensional health process~$\bfx_i
= \{ x_i (s) \}_{0 < s < \tau_i}$ for each participant is measured at
a dense grid of time points.  Accelerometer, for example, is measured
at a rate of 32Hz (i.e., $32$ times per second). Electrodermal
activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e.,
4 times per second).  Given the high frequency nature of the sensor 
measurements, this paper assumes the process is measured continuously.

Let~$H_{i,t}^{NX} = H_{i,t}^{N} \otimes H_{i,t}^{X}$ be the $\sigma$-field
generated by all past values~$(N_i (s), x_i (s))_{0 \leq s \leq t}$.
The instantaneous risk of an event at time~$t$ is assumed to depend on
the health process, time-in-study, and the event history through a
fully parametric conditional hazard function:
\begin{equation}
\label{eq:hazard}
h_i \left( t \Given H_{i,t}^{NX} ; \theta \right) =
\lim_{\delta \to 0} \delta^{-1} \pr \left( N_i(t+\delta) - N_i(t)
  \given H_{i,t}^{NX}; \theta \right).
\end{equation}
where~$\theta$ is the parameter vector. Equation~\eqref{eq:hazard}
states the hazard may depend on past values of the event history and
covariates.  For high frequency physiological data, we assume that 
current risk is a linear functional of the health process over some
recent window of time; that is, 
\begin{equation}
\label{eq:hazardlinear}
h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) = 
h_0 (t) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
  + \int_{t-\Delta}^t x_i (s) \beta(s) ds  \right)
\end{equation}
where~$\Delta$ is an unknown window-length, and $g_t( H_{i,t}^N ) \in
\mathbb{R}^p$ is a $p$-length feature vector summarizing the
event-history and time-in-study information. The final
term~$\int_{t-\Delta}^t x_i(s) \beta(s) ds$ reflects the unknown
linear functional form of the impact of the time-varying covariate on
current risk.

An alternative to~\eqref{eq:hazardlinear} would be to construct
features from the sensor data history~$f_t ( H_{i,t}^{X}) \in
\mathbb{R}^q$ and incorporated these features in the place of the
final term. Our current approach builds linear features of~$H_{i,t}^X$
directly from the integrated history, avoiding the feature
construction problem -- a highly nontrivial issue for high frequency
time-series data.  The main caveat is the additional
parameter~$\Delta$; however, we show that so long as the
estimated~$\hat \Delta$ exceeds~$\Delta$, then resulting estimation is
unbiased albeit at a loss of efficiency.  Moreover, sensitivity
analysis can be peformed to determine how choice of $\hat \Delta$
affects inference.  One limitation of the approach presented here is
that only fully parametric hazard models may be fit to the data.
However, a spline model for the log baseline hazard affords sufficient
model flexibility.

\subsection{Measurement-error models with event processes}

One potential criticism for~\eqref{eq:hazard} is that the health
process may be measured with error.  
A common mathematical strategy for joint models is to consider an
unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i
\given \eta_i$. Take~$\eta_i$ to be a zero-mean Gaussian process with 
\begin{equation}\label{eq:jm}
x_i(t) = \eta_i (t) + \epsilon_i (t),\quad \text{ and } \log h_i (t
\given \eta ) = \log h_0 (t) + g_t \left( H_{i,t}^N \right)^{\prime}
\alpha + \int_{t-\Delta}^t \eta_i (s) \beta (s) ds
\end{equation}
where~$\epsilon (t)$ is a white-noise measurement error. Thus,
$\eta_i (t)$ is the ``true and unobserved value of the longitudinal
outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional hazard
function given~$\bfx_i$ is given by
\[
\pr \left ( T_{i,j} = t \Given \H_{i,\tau_i}^X, \{ T_{i,j^\prime} < t \}_{j^\prime <
  j}, T_{i,j} \geq t \right) = E \left( h_i (t \given \eta) \exp
\left( \int_{T_{i,j-1}}^t h_i (\eta (s)) ds \right) \given \bfx_i
\right).
\]
Therefore, it appears that the conditional hazard function depends 
not only on past $x$-values, but also on future $x$-values.
Therefore, the processes~$\bfx$ and~$\bfT$ do not satisfy
\emph{independent evolution}~\citep{DempseyPMCC2}.
This is quite unnatural, as~\eqref{eq:jm} suggests current risk
depends on future values of the sensor process.
To alleviate this issue, we employ the strategy of taking~$\epsilon
(t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero Gaussian process.
Note, the proposed inferential procedure will not rely on this
modeling assumption being correct; instead, we will use the random
subsampling mechanisms to address model misspecification.

\subsection{Likelihood calculation}

The component of the log-likelihood related to the event process
is given by
\[
  L_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} 
  \left[ \log ( h_i (T_{i,j}) ) - H_i (T_{i,j}, T_{i,j-1}; \theta)
  \right] - H_{i} (\tau_i, T_{i,k_i}; \theta) \right) + \log \left(
  \pr ( \bfx_i ; \psi ) \right)
\]
where~$T_{i,0} = 0$, $T_{i, k_i + 1} = \tau_i$, and
\[
H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \int_{T_{i,j-1}}^{T_{i,j}} h_{i} (t; \theta) dt.
\]
Solving the associated score equations~$U_n (\theta) = 0$ yields the
maximum likelihood estimator~$\hat \theta$, where
\[
U_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} 
  \left[ \frac{h^{(1)}_i (T_{i,j})}{h_i (T_{i,j})} - H^{(1)}_i
    (T_{i,j}, T_{i,j-1}; \theta) \right]  - H^{(1)}_{i} (\tau_i,
  T_{i,k_i}; \theta) \right) 
\]
and~$h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (T_{i,j}, T_{i,j-1}; \theta)$
are the derivatives of~$h_i^{(1)} (T_{i,j}; \theta)$ and~$H_i^{(1)}
(T_{i,j}, T_{i,j-1}; \theta)$ with repsect to $\theta$.

In classical joint models~\citep{Henderson2000, Tsiatis2004},
time-varying covariates~$x_i (t)$ are observed only intermittently at
appointment times, which may or may not be
administrative~\citep{DempseyPMCC}.  In this setting, it is not feasible
to compute the cumulative hazard functions~$H_{i} (T_{i,j}, T_{i,j-1};
\theta)$.  In our current setting, we assume~$x_{i} (t)$ is observed
continuously.  While this relieves the issue of intermittent
observations, a joint model requires a flexible class of longitudinal
models for the sensor stream data that adequately capture their
dynamics.  Moreover, the risk model now depends on an integrated past
history of the time-varying covariate leads to severe increase in
computational complexity.

\subsection{Probabilistic subsampling framework}

To avoid both the statistical modeling challenge and the computational
burden, we employ a point-process subsampling design to obtain
unbiased estimates of the cumulative hazards for each subject. The
subsampling procedure treats the collected sensor data as a set of
\emph{potential observations}. 
Suppose covariate information is sampled at times drawn from an
independent inhomogeneous Poisson point process with known
intensity~$\pi_i (t)$. At a subsampled time~$t$, the \emph{windowed
  covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq \Delta}$ is
observed. Optimal choice of~$\pi_i (t)$ is beyond the scope of this
paper; however, the literature suggests setting proportional to the
hazard function~$h_i (t; \theta)$.

An estimator is design-unbiased if its expectation is equal to that
parameter under the probability distribution induced by the sampling
design~\citep{Cassel1977}. Let~$D_i \subset [0,t_i]$ denote the set of
subsampled points.  Note, by design, this set is distinct from the set
of event times, i.e.,~$\bfT_i \cap D_i = \emptyset$.  Under
subsampling via $\pi_i (t)$, then we may compute the Horvitz-Thompson
estimator 
\[
\hat H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \sum_{t \in D \cap [T_{i,j-1},
  T_{i,j}]} \frac{ \lambda^{(1)} ( t \given \{ x(t - s) \}_{0 < s <
    \Delta}; \theta ) }{ \pi (s) }
\]
Let~$N = \{ T_{i,j} \}_{i =1, j=1}^{n,k_i}$ denote the complete set of
event times.  Then an alternative design-unbiased estimator of the
cumulative hazards is given by
\begin{equation}
\label{eq:WPest}
\hat H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \sum_{t \in (N \cup D) \cap [T_{i,j-1},
  T_{i,j}]} \frac{ \lambda^{(1)} ( t \given \{ x(t - s) \}_{0 < s <
    \Delta}; \theta ) }{ \pi (s) + \lambda ( t \given \{ x(t - s) \}_{0 < s <
    \Delta}; \theta) }
\end{equation}
Equation~\eqref{eq:WPest} is the estimator suggested
by~\cite{Waagepetersen2008}.  This estimator depends on the
superposition of the event and subsampling processes.  This estimator
has been shown to be more efficient than the Horvitz-Thompson
estimator; therefore, we restrict our attention to~\eqref{eq:WPest}
for the remainder of this paper. Let
\[
w (t; \theta) = \frac{\pi (t)}{\pi (t) + \lambda (t \given \{
  x(t - s) \}_{0 < s < \Delta}; \theta)}
\]
Then the resulting estimating equations are given by
\begin{equation}
\label{eq:approxscore}
\hat{U}_n = \sum_{t \in N} w(t; \theta) \frac{\lambda^{(1)} (t \given \{
  x(t - s) \}_{0 < s < \Delta}; \theta)}{ \lambda ( t \given \{
  x(t - s) \}_{0 < s < \Delta}; \theta)}  - 
\sum_{t \in D} w(t; \theta) \frac{\lambda^{(1)} (t \given \{
  x(t - s) \}_{0 < s < \Delta}; \theta)}{ \pi (t) }
\end{equation}
Equation~\eqref{eq:approxscore} are approximate score functions built
via plug-in of the design-unbiased estimator of the cumulative hazard.

We note that this approach avoids modeling the sensor stream process,
which would be required if one were to consider a standard joint model
approach.  Sensor streams such as accelerometer and skin-conductance
relate to complex dynamical systems.  The subsampling
approach avoids this difficult modeling task entirely.

\section{Functional principal components within event-history
  analysis}

Take $\{ X (s) \}_{0 < s < \Delta}$. Let~$\psi_{j} (s)$ is an
orthonormal basis for $L^2 [ 0,1 ]$.  Then

\[
X(s) - \mu(s) 
\]

A simpler alternative is the Horvitz-Thompson estimator where the
denominator only depends on

\[
L(\theta) = \sum_{s \in N} \log \lambda \left( \{ x(s-t) \}_{0 < t
    \leq \Delta} \right) 
\]

\section{Theoretical analysis}

\begin{assumption}[Event process assumptions]

\end{assumption}

\begin{assumption}[Covariance assumptions]

\end{assumption}

\begin{assumption}[Functional data assumptions]

\end{assumption}

\section{Simulations}

\subsection{}

\section{Case Study: }

\bibliographystyle{plainnat}
\bibliography{si-fda-refs}

\end{document}

### 2019-06-06 write-up/fda-recurrent.tex
Notes added deriving the score equations

@@
 
-The component of the log-likelihood related to the event process
-is given by
+For the sake of notational simplicity, we leave the dependency of the
+conditional hazard function on~$H_{i,t}^{NX}$ implicit, and write~$h_i
+(t; \theta)$. The component of the log-likelihood related to the event
+process is then given by
 \[
   L_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} 
-  \left[ \log ( h_i (T_{i,j}) ) - H_i (T_{i,j}, T_{i,j-1}; \theta)
-  \right] - H_{i} (\tau_i, T_{i,k_i}; \theta) \right) + \log \left(
-  \pr ( \bfx_i ; \psi ) \right)
+  \left[ \log \left( h_i \left( T_{i,j}; \theta
+      \right) \right) - H_i \left( T_{i,j}, T_{i,j-1}; \theta \right) 
+  \right] - H_{i} \left( \tau_i, T_{i,k_i}; \theta \right) \right) 
 \]
@@
 U_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} 
-  \left[ \frac{h^{(1)}_i (T_{i,j})}{h_i (T_{i,j})} - H^{(1)}_i
+  \left[ \frac{h^{(1)}_i (T_{i,j}; \theta)}{h_i (T_{i,j}; \theta)} - H^{(1)}_i
     (T_{i,j}, T_{i,j-1}; \theta) \right]  - H^{(1)}_{i} (\tau_i,
@@
 are the derivatives of~$h_i^{(1)} (T_{i,j}; \theta)$ and~$H_i^{(1)}
-(T_{i,j}, T_{i,j-1}; \theta)$ with repsect to $\theta$.
+(T_{i,j}, T_{i,j-1}; \theta)$ with respect to $\theta$.
 
@@
 time-varying covariates~$x_i (t)$ are observed only intermittently at
-appointment times, which may or may not be
-administrative~\citep{DempseyPMCC}.  In this setting, it is not feasible
-to compute the cumulative hazard functions~$H_{i} (T_{i,j}, T_{i,j-1};
-\theta)$.  In our current setting, we assume~$x_{i} (t)$ is observed
-continuously.  While this relieves the issue of intermittent
-observations, a joint model requires a flexible class of longitudinal
-models for the sensor stream data that adequately capture their
-dynamics.  Moreover, the risk model now depends on an integrated past
-history of the time-varying covariate leads to severe increase in
+appointment times. In our current setting, maximizing the likelihood
+is computationally prohibitive since for any~$\theta$ we must compute
+the cumulative hazard functions~$H_{i} (T_{i,j}, T_{i,j-1}; \theta)$
+which require integration of~$h_i(t;\theta)$ given
+by~\eqref{eq:hazardlinear} which itself depends on the
+integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$.  
+That is, the risk model now depends on an integrated past history of
+the time-varying covariate which leads to severe increase in
 computational complexity.
@@
 \[
-\hat H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \sum_{t \in D \cap [T_{i,j-1},
+\hat H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \sum_{t \in D_i \cap [T_{i,j-1},
   T_{i,j}]} \frac{ \lambda^{(1)} ( t \given \{ x(t - s) \}_{0 < s <
@@
 
-\section{Functional principal components within event-history
-  analysis}
+\section{Longitudinal functional principal components within
+  event-history analysis}
+
+Probabilistic subsampling converts the single sensor stream~$\bar x_i$
+into a sequence of functions observed repeatedly at sampled
+times~$D_i$ and event times~$N_i$. Such a data structure is commonly
+referred to as \emph{longitudinal functional data}~\ref{Xiao2013,
+  GoldSmith2015}.
+Given the large increase in longitudinal functional data in recent
+years, corresponding analysis has received much recent
+attention~\citep{Morris2003, MorrisCarroll2006,
+  Baladandayuthapani2008, Di2009, Greven2010, Staicu2010,
+  ChenMuller2012, LiGuan2014}.
+In this paper, we combine work by~\cite{Park2018}
+and~\cite{Goldsmith2011} to construct a computationally efficient
+penalized functional method for solving the estimation equations~$\hat
+U_n (\theta)$.
+
+In our setting, probabilistic subsampling leads to observation
+times~$t \in N \cup D$ at which
+
+The marginal covariance~\cite{Park2018} allows a single 
+\begin{itemize}
+\item {\bf Step 1}: Estimate~$\hat \mu(s,t)$ via sandwich smoother for
+  all~$s \in [0,\Delta]$ and~$t \in \cup_{i=1}^n \left( D_i \cup N_i
+  \right)$. Define~$\tilde X (s,t) = X(s,t) - \hat \mu(s,t)$ to be a
+  mean-zero process for each~$t$.
+\item {\bf Step 2}: Compute the pooled sample covariance
+\[
+\hat \Xi (s , s^\prime) = 
+\left( \sum_{i=1}^n \left | N_i \cup D_i \right |\right)^{-1}
+\left( \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \tilde X (s,t) \tilde X
+  (s^\prime, t) \right).
+\]
+Compute at a fine set of grid points. This is an estimate of 
+\[
+\Xi (s, s^\prime) = \int_{\mathcal{T}} c( (s,T), (s^\prime, T) ) g(T) dT
+\]
+where~$g(T)$ is a superposition of the event process and the
+subsampling process.
+\item {\bf Step 3}: Take spectral decomposition~$\{ \hat \phi_k (s),
+  \hat \lambda_k \}_{k \geq 1}$. Choose~$K_x \geq K_b$ large enough.
+\item {\bf Step 4}: Compute
+\[
+\hat c_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (s,t) \hat \phi_k (s) ds
+\]
+Then
+\[
+X(s,t) \approx \hat \mu(s,t) + \sum_{k=1}^{K_x} \hat c_{i,k} (t)
+\phi_k (s).
+\]
+\item {\bf Step 5}: Plugging both the spline and KL decomposition we
+  have
+\begin{align*}
+\int_{t-\Delta}^t X(s,t) \beta(s) ds 
+  &= \int_{t-\Delta}^t \left[ \hat \mu(s,t) + \sum_{k=1}^{K_x}
+    c_{i,t,k} \phi_k (s) \right] \times \left[ \sum_{k^\prime=1}^{K_b}
+    \beta_k \psi_k (s) \right] \\
+  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\phi, \psi} ] \beta
+\end{align*}
+where~$M_{i,t}$ is a vector with~$i$th entry equal
+to~$\int_{t-\Delta}^t \hat \mu(s,t) \psi_k (s) ds$, $C_{i,t} =
+(c_{i,t,1}, \ldots, c_{i,t,K_x}) \in \mathbb{R}^{K_x \times 1}$,
+$\beta = (\beta_1, \ldots, \beta_{K_b}) \in \mathbb{R}^{K_b \times 1}$
+and~$J_{\phi, \psi} \in \mathbb{R}^{K_x \times K_b}$ where the~$(i,j)$
+entry is given by
+\[
+\int_{t-\Delta}^t \phi_{i} (s) \psi_{j} (s) ds.
+\]
+\end{itemize}
+
+We re-write the hazard function as:
+\[
+h_i (t; \theta) = \exp \left( Z_{i,t}^\prime, \alpha + [M_{i,t} + C_{i,t}^\prime
+  J_{\phi, \psi} ] \beta \right) = \exp \left( W_{i,t}^\prime \theta \right),
+\]
+where~$W_{i,t}^\prime = [ Z_{i,t}^\prime (M_{i,t}^\prime +
+C_{i,t}^\prime J_{\phi, \psi}) ]$ and~$\theta = (\alpha,
+\beta)$. Then, the score function is:
+\[
+\hat U_n (\theta) = \sum_{i=1}^n \left \{  \sum_{t \in N_i}
+  \frac{h_i^{(1)} (t; \theta)}{h_i (t; \theta)} - \hat H_i^{(1)}
+  (\tau_i, 0; \theta) 
+\right \}
+\]
+where
+\[
+\hat H_i^{(1)} (\tau_i; \theta) = \sum_{u \in N_i \cup D_i} \frac{
+  h^{(1)}_i (u; \theta)}{\pi_i (u) + h_i (u; \theta)}.
+\]
+Note that
+\[
+\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
+  \theta)}{h_i (t; \theta)}.
+\]
+Therefore, these terms have very simple forms!!! 
+\[
+\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}
+\]
+On the other hand,
+\[
+\frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) = 
+\frac{\exp \left( W_{i,t}^\prime \theta \right)}{\pi_i (t) + \exp
+  \left( W_{i,t}^\prime \theta \right)} W_{i,t}
+\]
+Therefore,
+\begin{align*}
+\hat U_n (\theta) 
+  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
+    \sum_{u \in N_i \cup D_i} \frac{\exp \left( W_{i,u}^\prime \theta
+    \right)}{\pi_i (t) + \exp \left( W_{i,u}^\prime \theta \right)}
+    W_{i,u} \right \} \\
+  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
+    + \exp \left( W_{i,u}^\prime \theta \right)} W_{i,t} -
+    \sum_{u \in D_i} \frac{\exp \left( W_{i,u}^\prime \theta
+    \right)}{\pi_i (t) + \exp \left( W_{i,u}^\prime \theta \right)}
+    W_{i,u} \right \} \\
+  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
+    \sum_{u \in D_i} (1 - w_i (u; \theta)) W_{i,u} \right \} \\
+\end{align*}
+
+\section{Prediction}
+We wish to estimate the risk at non-sampled points.
+
+For prediction, we use sparse FPCA (Yao 2005) 
+\begin{itemize}
+\item Take $\hat c_{k} (s,t)$ for~$t \in N \cup D$ and~$s \in
+  [0,\Delta]$
+\item 
+\end{itemize}
 
@@
 
-A simpler alternative is the Horvitz-Thompson estimator where the
-denominator only depends on
-
 \[
-L(\theta) = \sum_{s \in N} \log \lambda \left( \{ x(s-t) \}_{0 < t
-    \leq \Delta} \right) 
+- b^\prime D b
 \]
@@
 
+\begin{thm}[No reconstruction error]
+
+\end{thm}
+
+\begin{thm}[Reconstruction error but~$\hat \Delta > \Delta$]
+\label{eq:biaseddown}
+\end{thm}
+
+Theorem~\ref{eq:}
+
+\begin{thm}[When~$\hat \Delta < \Delta$]
+
+\end{thm}
+
 \section{Simulations}
 
-\subsection{}
+\subsection{Unbiased estimation}
+
+\subsection{Unbiased estimation}
 

### 2019-06-10 visualization/weight_calculator_per_participant.R
Updated comments in weight calculation file

@@ -6,7 +6,7 @@ args=(commandArgs(TRUE))
 ## Then cycle through each element of the list and evaluate the expressions.
 if(length(args)==0){
   print("No arguments supplied.")
-  ##supply default values
+  ## Supply default participant id
   id = 1040
 }else{
   id = as.numeric(args[[1]])
@@ -46,18 +46,12 @@ event_times_output = foreach(bp_obs=1:nrow(id_bp), .combine = 'cbind') %dopar%
 event_numerator_output = matrix(event_times_output[1,], ncol = length(sequence), byrow = TRUE)
 event_denominator_output = matrix(event_times_output[2,], ncol = length(sequence), byrow = TRUE)
 
-#outcome = colSums(numerator_output, na.rm = TRUE)/colSums(denominator_output, na.rm = TRUE)
-# plot(sequence, outcome-mean(outcome) , type = "l", 
-#      ylab = "FSF EDA (mean-centered)", xlab = "Time (minutes) prior to event time",
-#      main = paste("Participant",id))
-
-num.iters = 5000
+num.iters = 5000 # Number of non-event samples per participant
 nonevent_times_output = foreach(bp_obs=1:num.iters, .combine = 'cbind', .options.RNG=197311) %dopar% 
   nonevent_kernel_apply(eda, id_bp, sequence)
 
 nonevent_numerator_output = matrix(nonevent_times_output[1,], ncol = length(sequence), byrow = TRUE)
 nonevent_denominator_output = matrix(nonevent_times_output[2,], ncol = length(sequence), byrow = TRUE)
-#outcome = colSums(nonevent_numerator_output,na.rm = TRUE)/colSums(nonevent_denominator_output, na.rm=TRUE)
 
 full_data = list("event_numerator_output"=event_numerator_output, 
                  "event_denominator_output"=event_denominator_output,

### 2019-06-10 visualization/plot_generator.R
Fixed dir links

@@ -2,7 +2,7 @@ sequence = seq(-30,10, by = 0.25) # sequence along which we compute
 event_total_numerator = event_total_denominator = rep(NA, length(sequence))
 nonevent_total_numerator = nonevent_total_denominator = rep(NA, length(sequence))
 id_counter = NA
-setwd("/Volumes/murphy_lab/users/wdempsey/fda-recurrent-events/data")
+setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
 
 for(id in 1001:1091) {
   print(paste("On participant ",id))

### 2019-06-10 visualization/weight_calculator_per_participant.R

@@ -58,6 +58,6 @@ full_data = list("event_numerator_output"=event_numerator_output,
                  "nonevent_numerator_output"=nonevent_numerator_output,
                  "nonevent_numerator_output"=nonevent_numerator_output)
 
-setwd("~/Documents/github/fda-recurrentevents/data_visualization/data")
+setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
 saveRDS(object = full_data, file = paste("kerneloutput_id_",id,".RDS", sep=""))
 

### 2019-06-10 visualization/plot_generator.R
Fixing plot generator: Removed plot_obs

@@ -1,4 +1,5 @@
 sequence = seq(-30,10, by = 0.25) # sequence along which we compute
+plot_obs = sequence <= 0
 event_total_numerator = event_total_denominator = rep(NA, length(sequence))
 nonevent_total_numerator = nonevent_total_denominator = rep(NA, length(sequence))
 id_counter = NA
@@ -33,7 +34,7 @@ for(id in 1001:1091) {
     #mtext("Time until event", side = 1, line = 2, cex = 0.75)
     mtext(paste("Participant", id), side = 3, line = -1)
     axis(side =1, cex.axis = 0.75); axis(side=2,  cex.axis = 0.75)
-    lines(sequence[plot_obs], nonevent_mean-mean(nonevent_mean), col = "red")
+    lines(sequence, nonevent_mean-mean(nonevent_mean), col = "red")
     
     max.obs = max(c(event_mean, nonevent_mean), na.rm = TRUE); min.obs = min(c(event_mean, nonevent_mean), na.rm = TRUE)
     plot(sequence, event_mean, 
@@ -43,7 +44,7 @@ for(id in 1001:1091) {
     mtext("EDA", side = 2, line = 2, cex = 0.75)
     mtext("Time until event", side = 1, line = 2, cex = 0.75)
     axis(side =1, cex.axis = 0.75); axis(side=2,  cex.axis = 0.75)
-    lines(sequence[plot_obs], nonevent_mean, col = "red")
+    lines(sequence, nonevent_mean, col = "red")
     dev.off()
   }
 }

### 2019-06-10 methods/estimation_functions.R
Built sampling protocol with input = 2 samples per hour (i.e., rate = 0.5 per hour)

@@ -28,3 +28,26 @@ nonevent_approximate_eda_apply <- function(iter, sequence, sampled_times, eda) {
   output = unlist(lapply(sequence, approximate_eda(minute_diff[keep_obs],eda$EDA_FeatureScaled_Filtered[keep_obs])))
   return(output)
 }
+
+generate_noneventtimes <- function(datetime_ts, sampling_rate, max.iters) {
+  sampled_times = c()
+  current_time = datetime_ts[1]
+  for(iter in 1:max.iters) {
+    random.next.time <- current_time + rexp(1,rate = sampling_rate)*60*60
+    if(random.next.time > max(datetime_ts)) {break}
+    diff = datetime_ts-random.next.time
+    if (seconds(min(abs(diff))) < 1) {
+      ## If the sampled time is < 1 minute from
+      ## sensor times, then keep as a sampled time
+      sampled_obs = which(abs(diff) == min(abs(diff)))
+      sampled_times = c(sampled_times, datetime_ts[sampled_obs[1]])
+      current_time = datetime_ts[sampled_obs[1]]
+    } else{
+      ## Else, set current.time to min time with diff > 0
+      print("ELSE")
+      new_obs = min(which(diff > 0))
+      current_time = datetime_ts[new_obs[1]]
+    }
+  }
+  return(sampled_times)
+}

### 2019-06-10 methods/generate_dataset.R

@@ -1,4 +1,6 @@
 source('./estimation_functions.R')
+source('./inputs.R')
+require(lubridate)
 
 setwd("/Users/walterdempsey/Harvard University/R21_Study - EDA")
 buttonpress = readRDS("../R21_Study - tags/button_presses.RDS")
@@ -28,7 +30,8 @@ for (id in 1001:1091) {
     print(paste("No EDA file with id", id))
     break
   }
-  
+  datetime_ts = as_datetime(eda$ts/1000)
+  sampled_times = generate_noneventtimes(datetime_ts, sampling_rate, max.iters = 20000)
   sampled_times = sample(eda$ts, size = num.iters, replace = FALSE)
   output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_eda_apply(iter, sequence, id_bp, eda)
   output_nonevent = foreach(iter=1:num.iters, .combine=rbind) %dopar% nonevent_approximate_eda_apply(iter, sequence, sampled_times, eda)

### 2019-06-10 methods/generate_dataset.R
Fixed timestamp 1000 issue (i.e., some are in milliseconds)

@@ -32,14 +32,13 @@ for (id in 1001:1091) {
   }
   datetime_ts = as_datetime(eda$ts/1000)
   sampled_times = generate_noneventtimes(datetime_ts, sampling_rate, max.iters = 20000)
-  sampled_times = sample(eda$ts, size = num.iters, replace = FALSE)
   output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_eda_apply(iter, sequence, id_bp, eda)
-  output_nonevent = foreach(iter=1:num.iters, .combine=rbind) %dopar% nonevent_approximate_eda_apply(iter, sequence, sampled_times, eda)
+  output_nonevent = foreach(iter=1:length(sampled_times), .combine=rbind) %dopar% nonevent_approximate_eda_apply(iter, sequence, sampled_times, eda)
   
   ## Define time as since minimum time in EDA or ID_BP
   base_time = min(eda$ts, id_bp$ts_ms)
   event_times_mins_since_base = (id_bp$ts_ms-base_time)/1000/60
-  nonevent_times_mins_since_base = (sampled_times-base_time)/1000/60
+  nonevent_times_mins_since_base = (sampled_times*1000-base_time)/1000/60
   
   event_temp = cbind(id, event_times_mins_since_base, output_event)
   nonevent_temp = cbind(id, nonevent_times_mins_since_base, output_nonevent)

### 2019-06-10 methods/generate_dataset.R
Fixed working directory issue

@@ -2,12 +2,12 @@ source('./estimation_functions.R')
 source('./inputs.R')
 require(lubridate)
 
-setwd("/Users/walterdempsey/Harvard University/R21_Study - EDA")
-buttonpress = readRDS("../R21_Study - tags/button_presses.RDS")
+setwd("/n/murphy_lab/users/wdempsey/data-for-fda/data/")
+buttonpress = readRDS("/R21_Study - tags/button_presses.RDS")
 
 library(doParallel)
 set.seed(87514)
-sequence = seq(-30,0,1/60); num.iters = 1000; 
+sequence = seq(-30,0,1/60);
 event_complete = nonevent_complete = matrix(nrow = 0, ncol = 2+length(sequence))
 
 cl = Sys.getenv("SLURM_NTASKS_PER_NODE")

### 2019-06-10 methods/generate_dataset.R
Added ./ to point to right directory for BP data

@@ -3,7 +3,7 @@ source('./inputs.R')
 require(lubridate)
 
 setwd("/n/murphy_lab/users/wdempsey/data-for-fda/data/")
-buttonpress = readRDS("/R21_Study - tags/button_presses.RDS")
+buttonpress = readRDS("./R21_Study - tags/button_presses.RDS")
 
 library(doParallel)
 set.seed(87514)

### 2019-06-10 methods/generate_dataset.R
Fixed eda df location issue

@@ -23,13 +23,14 @@ for (id in 1001:1091) {
   
   id_bp = subset(buttonpress, ID == id)
   
-  file_name = paste(id,"_EDA.rds", sep = "")
+  file_name = paste("./R21_Study - EDA/",id,"_EDA.rds", sep = "")
   if(file.exists(file_name)) {
     eda = readRDS(paste(id,"_EDA.rds", sep = ""))
   } else{
     print(paste("No EDA file with id", id))
     break
   }
+  print("Made it to sampling nonevent times")
   datetime_ts = as_datetime(eda$ts/1000)
   sampled_times = generate_noneventtimes(datetime_ts, sampling_rate, max.iters = 20000)
   output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_eda_apply(iter, sequence, id_bp, eda)

### 2019-06-10 methods/generate_dataset.R
Fixed eda = readRDS line

@@ -25,7 +25,7 @@ for (id in 1001:1091) {
   
   file_name = paste("./R21_Study - EDA/",id,"_EDA.rds", sep = "")
   if(file.exists(file_name)) {
-    eda = readRDS(paste(id,"_EDA.rds", sep = ""))
+    eda = readRDS(file_name)
   } else{
     print(paste("No EDA file with id", id))
     break

### 2019-06-10 methods/inputs.R
Inputs inserted

@@ -1,0 +1,36 @@
+# Directory for EDA participant data
+eda_dir <- "/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/R21_Study - EDA" 
+
+# Directory for ACC participant data
+acc_dir <- NA # Currently not linked in.
+
+# Directory for BP data
+button_press_dir <- "/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/R21_Study - tags"
+
+# List of all participant ids
+participant_ids <- seq(1001,1091)
+
+# Feature Generator Function
+feature_generator <- function(timestamp, id, bp_data) {
+  # Generates Features from BP data
+  # Inputs: Timestamp, participant id, button press data
+  # Output: Feature vector
+  # Currently an indicator of BP in past hour
+  
+  # Take subset of bp data with participant ID == id
+  bp_data_participant_subset <- subset(bp_data, ID==id)
+  # Compute time since timestamp given
+  temp <- (timestamp - bp_data_participant_subset$ts)/(60*60)
+  # Return indicator of <1hr since last BP
+  return(as.numeric(any(temp[temp > 0.0] < 1.0)))
+}
+
+## User-specified indicator of whether partial
+## pooling should be used. If 1, fit hierarchical model; 
+## if 0, fit population model.
+partial_pooling_indicator <- 0
+
+
+## User-specified sampling rate (per hour)
+## for the non-event times
+sampling_rate = 2 # 2 times per hour (i.e., every 30-minutes)

### 2019-06-10 methods/mean_cov_estimation.R
Adding sandwich mu and pooled Sigma estimators

@@ -1,0 +1,101 @@
+setwd("/Volumes/murphy_lab/users/wdempsey/fda-recurrent-events/methods/")
+event_complete = readRDS("./event_complete.RDS")
+sequence = seq(-30,0,by =1/60)
+plot(sequence, event_complete[3,3:1803], type = "l")
+
+library(refund)
+full_obs = apply(X = event_complete, MARGIN = 1, FUN = function(x){!any(is.na(x))})
+Y = event_complete[full_obs,3:1803]
+x = event_complete[full_obs,2]
+z = sequence
+est <- fbps(Y,list(x=x,z=z))
+obs = 100
+plot(sequence,est$Yhat[obs,]- mean(est$Yhat[obs,]))
+lines(sequence, Y[obs,]-mean(Y[obs,]))
+
+## COMPUTE ESTIMATED MARGINAL COVARIANCE 
+
+inflation = 100
+residual = inflation*(Y-est$Yhat)
+## TAKE PAIRS OF ROWS and CALCULATE THE MSE
+Sigma = matrix(nrow = ncol(residual), ncol = ncol(residual))
+for (i in 1:ncol(residual)) {
+  for (j in 1:ncol(residual)) {
+    Sigma[i,j] = sum(residual[,i]*residual[,j])/nrow(residual)
+  }
+}
+
+Sigma_est <- fbps(Sigma)
+Sigma_est$Yhat[1:10,1:10]
+eig_Sigma_est = eigen(Sigma_est$Yhat)
+
+eig_vectors = eig_Sigma_est$vectors
+eig_values = eig_Sigma_est$values
+
+eig_values[eig_values<0] = 0
+
+(cumsum(eig_values)/sum(eig_values))[1:35]
+K = 35
+phi_vectors = eig_vectors[,1:K]
+coef = residual%*%phi_vectors
+plot(sequence, phi_vectors%*%coef[obs,]/inflation + est$Yhat[obs,], type = "l")
+lines(sequence, Y[obs,], col = "red", lwd = 2)
+
+## SAVE THE MEAN, COEFFICIENT MATRIX, AND FIRST K EIGEN VECTORS
+saveRDS(object = est$Yhat, file = "event_means.RDS")
+saveRDS(object = coef/inflation, file = "event_coef_matrix.RDS")
+saveRDS(object = phi_vectors, file = "event_eigen_vectors.RDS")
+saveRDS(object = x, file = "event_complete_case_times.RDS")
[…]

### 2019-06-10 methods/estimation_functions.R
Minor typos in each were fixed and not pushed; now pushing to fix that! :)

@@ -22,8 +22,8 @@ approximate_eda_apply <- function(iter, sequence, id_bp, eda) {
 
 nonevent_approximate_eda_apply <- function(iter, sequence, sampled_times, eda) {
   current_ts = sampled_times[iter]
-  diff = eda$ts - current_ts
-  minute_diff = diff/1000/60 
+  diff = eda$ts/1000 - current_ts
+  minute_diff = diff/60 
   keep_obs = (-30 < minute_diff) & (minute_diff <= 0)
   output = unlist(lapply(sequence, approximate_eda(minute_diff[keep_obs],eda$EDA_FeatureScaled_Filtered[keep_obs])))
   return(output)

### 2019-06-10 visualization/plot_generator.R

@@ -54,6 +54,14 @@ all_data = list("event_total_numerator" = event_total_numerator,
                 "nonevent_total_numerator" = nonevent_total_numerator,
                 "nonevent_total_denominator" = nonevent_total_denominator)
 saveRDS(object = all_data, file = "alldata.RDS")
+## If you don't want to re-run top, you can skip here and 
+## just pull in the computed data set, remove commenting below
+# all_data = readRDS("alldata.RDS")
+# sequence = seq(-30,10, by = 0.25) # sequence along which we compute
+# event_total_numerator = all_data$event_total_numerator
+# event_total_denominator = all_data$event_total_denominator
+# nonevent_total_numerator = all_data$nonevent_total_numerator
+# nonevent_total_denominator = all_data$nonevent_total_denominator
 
 event_avg_mean = nonevent_avg_mean = rep(0,length(sequence))
 num_samples = nrow(event_total_numerator)-1
@@ -77,7 +85,7 @@ par(mfrow = c(1,1), mar = c(3,3,1,1)+0.1)
 plot(sequence[plot_obs], event_avg[plot_obs], 
      type = "l", axes = FALSE, ylab = "", xlab = "",
      ylim = c(min.obs, max.obs))
-mtext("EDA (mean-centered)", side = 2, line = 2, cex = 0.75)
+mtext("EDA", side = 2, line = 2, cex = 0.75)
 #mtext("Time until event", side = 1, line = 2, cex = 0.75)
 axis(side =1, cex.axis = 0.75); axis(side=2,  cex.axis = 0.75, ylim = c(min.obs, max.obs))
 lines(sequence, nonevent_avg, col = "red")

### 2019-06-11 methods/gendata_functions.R
Added comments for functions including inputs and outputs

@@ -1,9 +1,14 @@
 approximate_eda <- function(x, y) {
+  ## Input: x = minute_diff from current_time; y = corresponding EDA observations
+  ## Output: Function that takes in time-until-event and outputs closest observed EDA
   inside_fn <-function(time_until_event) {
+    ## Input: Time until event = s
+    ## Output: Finds the observations within 1 second of t-s; 
+    ## Output: outputs mean of these if multiple min(gap) 
     gap = abs(x-time_until_event)
     which_obs = which(gap == min(gap))
-    if( gap[which_obs] < 1/60 ) {
-      return (y[which_obs])  
+    if( min(gap) < 1/60 ) {
+      return (mean(y[which_obs]))  
     } else {
       return (NA)
     }
@@ -12,6 +17,8 @@ approximate_eda <- function(x, y) {
 }
 
 approximate_eda_apply <- function(iter, sequence, id_bp, eda) {
+  ## Input: Current iter, sequence of times until event, set of event times, EDA data
+  ## Output: Outputs corresponding EDA at time-until-event
   current_ts = id_bp$ts_ms[iter]
   diff = eda$ts - current_ts
   minute_diff = diff/1000/60 
@@ -21,6 +28,8 @@ approximate_eda_apply <- function(iter, sequence, id_bp, eda) {
 }
 
 nonevent_approximate_eda_apply <- function(iter, sequence, sampled_times, eda) {
+  ## Input: Current iter, sequence of times until event, set of sampled times, EDA data
+  ## Output: Outputs corresponding EDA at time-until-event
   current_ts = sampled_times[iter]
   diff = eda$ts/1000 - current_ts
   minute_diff = diff/60 
@@ -30,9 +39,15 @@ nonevent_approximate_eda_apply <- function(iter, sequence, sampled_times, eda) {
 }
 
 generate_noneventtimes <- function(datetime_ts, sampling_rate, max.iters) {
+  ## Code to generate sequence of sampled times
+  ## When sample is not near a point, we move forward to next time
+  ## This is because of gaps in coverage; luckily exponentials are memoryless
+  ## Input: Sequence of datetimes, sampling rate, and max number of samples
+  ## Output: Sequence of random sampled datetimes
   sampled_times = c()
   current_time = datetime_ts[1]
   for(iter in 1:max.iters) {
+    print(iter)
     random.next.time <- current_time + rexp(1,rate = sampling_rate)*60*60
[…]

### 2019-06-11 methods/generate_dataset.R

@@ -1,8 +1,17 @@
-source('./estimation_functions.R')
+## Code to generate the dataframes that will be used
+## to construct functional covariates 
+## Requires subsampling non-event times 
+## and the pooled covariance estimator.
+
+## Inputs: RDS files for participant level eda and button_press
+## Outputs: event_complete.RDS and nonevent_complete.RDS
+
+source('./gendata_functions.R')
 source('./inputs.R')
 require(lubridate)
 
-setwd("/n/murphy_lab/users/wdempsey/data-for-fda/data/")
+#setwd("/n/murphy_lab/users/wdempsey/data-for-fda/data/")
+setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
 buttonpress = readRDS("./R21_Study - tags/button_presses.RDS")
 
 library(doParallel)

### 2019-06-14 write-up/fda-recurrent.tex
Minor typos in each were fixed and not pushed; now pushing to fix that! :)
Filling out sections now

@@
 
-\title{Recurrent event analysis in the presence of functional
-  covariates via random subsampling}
+\title{Recurrent event analysis in the presence of real-time high
+  frequency data via random subsampling}
 \author{Walter Dempsey}
@@
 Motivated by this, a random subsampling framework is proposed for
-computationally efficient estimation of the effect of time-varying
-covariates on the recurrent event process. A subsampling-unbiased
-estimator for the cumulative hazard enters into an approximation of
-log-likelihood. The proposed estimator then maximizes this
-approximation. The estimator has two sources of variation: the first
-due to the recurrent event model and the second due to subsampling.
-The latter can be reduced by increasing the sampling rate; however,
-this leads to increased computational costs. 
+computationally efficient, approximate likelihood-based estimation.
+A subsampling-unbiased estimator for the derivative of the cumulative
+hazard enters into an approximation of log-likelihood. The estimator
+has two sources of variation: the first due to the recurrent event
+model and the second due to subsampling. The latter can be reduced by
+increasing the sampling rate; however, this leads to increased
+computational costs.  The approximate score equations are equivalent
+to logistic regression score equations, allowing for standard,
+``off-the-shelf'' software to be used in fitting these models.
 Simulations demonstrate the method and efficiency-computation
@@
 \keywords{recurrent events; probabilistic subsampling; estimating equations;
-high frequency time series; suicidal ideation}
+high frequency time series; logistic regression}
 
@@
 real-time high frequency sensor data to an adverse, recurrent event
-process. In the smoking cessation mHealth study~\citep{}, for
+process. In a smoking cessation mHealth study~\citep{Sense2Stop}, for
 example, the relation between a time-varying sensor-based measure of
-physiological stress and the smoking lapse process is of interest.
-In a suicidal ideation mHealth study~\citep{}, the relation of
+physiological stress and smoking lapse is of scientific interest.
+In a suicidal ideation mHealth study~\citep{Kleiman2018}, the relation of
 electrodermal activity (EDA) and accelerometer with self-identified
-moments of suicidal ideation is of interest.
+moments of suicidal ideation is of scientific interest.
 
 The goal of this paper is two-fold: (1) to discuss the appropriate
-choice of a statistical model for joint high-frequency sensor data and
+choice of statistical model for joint high-frequency sensor data and
 the recurrent event data, and (2) to construct a robust,
@@
 variation is small in relation to to variation in the underlying
-stochastic processes. Second, the subsampling procedure allows
-inference to not depend on choice of dynamics for the health process.
-Third, optimal estimating equations are derived that depend on the
-instantaneous rate of the recurrent event process, which increase
-increase efficiency of the subsampling procedure. Finally,
-implementation can leverage existing software for functional data
-analysis, leading to fast adoption by domain scientists. We show how
-to implement methods via standard mixed effects software.
-
-\section{Related prior work}
+stochastic processes. Second, optimal estimating equations are derived
+that depend on the instantaneous rate of the recurrent event process,
+which increases efficiency of the subsampling procedure. 
+Finally, implementation can leverage existing, standard software for
+functional data analysis and logistic regression, leading to fast
+adoption by domain scientists.
 
@@
 times~$\bfT_{i} = \{ T_{i,1}, \ldots, T_{i,k_i}\}$ over some observation
-window $[0, \tau_i]$ for each subject $i = 1,\cdots, n$.  Assume the
+window $[0, \tau_i]$ for each subject $i = 1,\ldots, n$.  Assume the
 event times are ordered, i.e., $T_{i,j} < T_{i,j^\prime}$ for $j <
@@
 h_i \left( t \Given H_{i,t}^{NX} ; \theta \right) =
-\lim_{\delta \to 0} \delta^{-1} \pr \left( N_i(t+\delta) - N_i(t)
-  \given H_{i,t}^{NX}; \theta \right).
+\lim_{\delta \to 0} \delta^{-1} \pr \left( N_i(t+\delta) - N_i(t) > 0
+  \given H_{i,t}^{NX}; \theta \right), 
 \end{equation}
@@
 covariates.  For high frequency physiological data, we assume that 
-current risk is a linear functional of the health process over some
-recent window of time; that is, 
+current risk is log-additive and depends on a linear functional of the
+health process over some recent window of time and some pre-specified
+features of the counting process; that is,  
 \begin{equation}
@@
 $\eta_i (t)$ is the ``true and unobserved value of the longitudinal
-outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional hazard
+outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional intensity
 function given~$\bfx_i$ is given by
@@
 (t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero Gaussian process.
-Note, the proposed inferential procedure will not rely on this
-modeling assumption being correct; instead, we will use the random
-subsampling mechanisms to address model misspecification.
+The proposed inferential procedure only relies on the Gaussian
+assumption when~$\bfx$ is observed with missingness (see
+Section~\ref{section:missingdata}).
 
@@
   L_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} 
-  \left[ \log \left( h_i \left( T_{i,j}; \theta
-      \right) \right) - H_i \left( T_{i,j}, T_{i,j-1}; \theta \right) 
-  \right] - H_{i} \left( \tau_i, T_{i,k_i}; \theta \right) \right) 
+    \log \left( h_i \left( T_{i,j}; \theta
+      \right) \right) - H_{i} \left( \tau_i; \theta \right) \right) 
 \]
-where~$T_{i,0} = 0$, $T_{i, k_i + 1} = \tau_i$, and
+where~$H_{i} (\tau_i ; \theta) = \int_{0}^{\tau_i} h_{i} (t; \theta)
+dt$. Solving the associated score equations~$U_n (\theta) = {\bf 0}$
+yields the maximum likelihood estimator~$\hat \theta$, where
 \[
-H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \int_{T_{i,j-1}}^{T_{i,j}} h_{i} (t; \theta) dt.
+U_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} \frac{h^{(1)}_i
+    (T_{i,j}; \theta)}{h_i (T_{i,j}; \theta)} - H^{(1)}_{i} (\tau_i;
+  \theta) \right)  
 \]
-Solving the associated score equations~$U_n (\theta) = 0$ yields the
-maximum likelihood estimator~$\hat \theta$, where
-\[
-U_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} 
-  \left[ \frac{h^{(1)}_i (T_{i,j}; \theta)}{h_i (T_{i,j}; \theta)} - H^{(1)}_i
-    (T_{i,j}, T_{i,j-1}; \theta) \right]  - H^{(1)}_{i} (\tau_i,
-  T_{i,k_i}; \theta) \right) 
-\]
-and~$h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (T_{i,j}, T_{i,j-1}; \theta)$
-are the derivatives of~$h_i^{(1)} (T_{i,j}; \theta)$ and~$H_i^{(1)}
-(T_{i,j}, T_{i,j-1}; \theta)$ with respect to $\theta$.
+where~$h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (\tau_{i}; \theta)$
+are derivatives with respect to $\theta$.
 
@@
 is computationally prohibitive since for any~$\theta$ we must compute
-the cumulative hazard functions~$H_{i} (T_{i,j}, T_{i,j-1}; \theta)$
-which require integration of~$h_i(t;\theta)$ given
-by~\eqref{eq:hazardlinear} which itself depends on the
-integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$.  
-That is, the risk model now depends on an integrated past history of
-the time-varying covariate which leads to severe increase in
-computational complexity.
+the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require
+integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which
+itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s)
+ds$.  That is, the risk model now depends on an integrated past
+history of the time-varying covariate which leads to severe increase
+in computational complexity.
 
@@
 
-To avoid both the statistical modeling challenge and the computational
-burden, we employ a point-process subsampling design to obtain
-unbiased estimates of the cumulative hazards for each subject. The
-subsampling procedure treats the collected sensor data as a set of
-\emph{potential observations}. 
-Suppose covariate information is sampled at times drawn from an
-independent inhomogeneous Poisson point process with known
-intensity~$\pi_i (t)$. At a subsampled time~$t$, the \emph{windowed
-  covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq \Delta}$ is
+To solve the computational challenge we employ a point-process
+subsampling design to obtain unbiased estimates of the derivative of
+the cumulative hazards for each subject. The subsampling procedure
+treats the collected sensor data as a set of \emph{potential
+  observations}. Suppose covariate information is sampled at times
+drawn from an independent inhomogeneous Poisson point process with
+known intensity~$\pi_i (t)$. At a subsampled time~$t$, the
+\emph{windowed covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq
+  \Delta}$ and counting process features~$g_t (H_{i,t}^N)$ are
 observed. Optimal choice of~$\pi_i (t)$ is beyond the scope of this
@@
 of event times, i.e.,~$\bfT_i \cap D_i = \emptyset$.  Under
-subsampling via $\pi_i (t)$, then we may compute the Horvitz-Thompson
-estimator 
+subsampling via $\pi_i (t)$, one may compute a Horvitz-Thompson
+estimator of the derivative of the cumulative hazard:
 \[
-\hat H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \sum_{t \in D_i \cap [T_{i,j-1},
-  T_{i,j}]} \frac{ \lambda^{(1)} ( t \given \{ x(t - s) \}_{0 < s <
-    \Delta}; \theta ) }{ \pi (s) }
+\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in D_i} \frac{
+  h^{(1)}_i (u; \theta) }{ \pi_i (s) }
 \]
-Let~$N = \{ T_{i,j} \}_{i =1, j=1}^{n,k_i}$ denote the complete set of
-event times.  Then an alternative design-unbiased estimator of the
+An alternative design-unbiased estimator of the derivative of the
 cumulative hazards is given by
@@
 \label{eq:WPest}
-\hat H_{i} (T_{i,j}, T_{i,j-1}; \theta) = \sum_{t \in (N \cup D) \cap [T_{i,j-1},
-  T_{i,j}]} \frac{ \lambda^{(1)} ( t \given \{ x(t - s) \}_{0 < s <
-    \Delta}; \theta ) }{ \pi (s) + \lambda ( t \given \{ x(t - s) \}_{0 < s <
-    \Delta}; \theta) }
+\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in (\bfT_i \cup D_i)}
+\frac{ h^{(1)}_i ( t; \theta ) }{ \pi_i (s) + h_i ( t ; \theta) }
 \end{equation}
@@
 by~\cite{Waagepetersen2008}.  This estimator depends on the
-superposition of the event and subsampling processes.  This estimator
-has been shown to be more efficient than the Horvitz-Thompson
-estimator; therefore, we restrict our attention to~\eqref{eq:WPest}
-for the remainder of this paper. Let
+superposition of the event and subsampling processes.
+Proposition~\ref{prop:optimal} shows the estimator for~$\theta$
+associated with using~\eqref{eq:WPest} is the most efficient within a
+suitable class of estimators for the derivative of the cumulative
+hazard function (including the Horvitz-Thompson estimator).
+Therefore, we restrict our attention to~\eqref{eq:WPest} for the
+remainder of this paper. Letting
 \[
-w (t; \theta) = \frac{\pi (t)}{\pi (t) + \lambda (t \given \{
-  x(t - s) \}_{0 < s < \Delta}; \theta)}
+w_i (t; \theta) = \frac{\pi_i (t)}{\pi_i (t) + h_i (t ; \theta)},
 \]
-Then the resulting estimating equations are given by
+the resulting estimating equations can be written as
 \begin{equation}
 \label{eq:approxscore}
-\hat{U}_n = \sum_{t \in N} w(t; \theta) \frac{\lambda^{(1)} (t \given \{
-  x(t - s) \}_{0 < s < \Delta}; \theta)}{ \lambda ( t \given \{
-  x(t - s) \}_{0 < s < \Delta}; \theta)}  - 
-\sum_{t \in D} w(t; \theta) \frac{\lambda^{(1)} (t \given \{
-  x(t - s) \}_{0 < s < \Delta}; \theta)}{ \pi (t) }
+\hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
+  \frac{h^{(1)} (u; \theta)}{ h ( t; \theta)}  - \sum_{u \in D_i} w_i(u;
+  \theta) \frac{h_i^{(1)} (t; \theta)}{ \pi_i (t) } \right].
 \end{equation}
 Equation~\eqref{eq:approxscore} are approximate score functions built
-via plug-in of the design-unbiased estimator of the cumulative hazard.
-
-We note that this approach avoids modeling the sensor stream process,
-which would be required if one were to consider a standard joint model
-approach.  Sensor streams such as accelerometer and skin-conductance
-relate to complex dynamical systems.  The subsampling
-approach avoids this difficult modeling task entirely.
+via plug-in of the design-unbiased estimator of the derivative of the
+cumulative hazard given in~\eqref{eq:WPest}.
 
@@
 
-Probabilistic subsampling converts the single sensor stream~$\bar x_i$
+Probabilistic subsampling converts the single sensor stream~$\bfx_i$
 into a sequence of functions observed repeatedly at sampled
-times~$D_i$ and event times~$N_i$. Such a data structure is commonly
-referred to as \emph{longitudinal functional data}~\ref{Xiao2013,
-  GoldSmith2015}.
+times~$D_i$ and event times~$\bfT_i$. Such a data structure is
+commonly referred to as \emph{longitudinal functional
+  data}~\citep{Xiao2013, GoldSmith2015}.
 Given the large increase in longitudinal functional data in recent
@@
   ChenMuller2012, LiGuan2014}.
-In this paper, we combine work by~\cite{Park2018}
-and~\cite{Goldsmith2011} to construct a computationally efficient
-penalized functional method for solving the estimation equations~$\hat
-U_n (\theta)$.
+Here, we combine work by~\cite{Park2018} and~\cite{Goldsmith2011} to
+construct a computationally efficient penalized functional method for
+solving the estimation equations~$\hat U_n (\theta)$.
 
-In our setting, probabilistic subsampling leads to observation
-times~$t \in N \cup D$ at which
+We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0
+\leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup D_i$.
+We use the sandiwch smoother~\citep{Xiao2013} to estimate the
+mean~$\mu(s,t) = \E [ X(t,s)]$.
+Alternative bivariate smoothers exist, such as the kernel-based local
+linear smoother~\citep{Hastie2009} bivariate tensor product
+splines~\cite{Wood2006} and the bivar
+The sandwich smoother was chosen for its computational efficiency and
+estimation accuracy.
+We then define~$\tilde X(s,t) = X(s,t) - \hat \mu(s,t)$ to be the
+mean-zero process at each time~$t \in $.
 
-The marginal covariance~\cite{Park2018} allows a single 
+Next,
+
+Use of the marginal covariance~\cite{Park2018} reduces allows us to avoid
+computationally prohibitive computations.
+
 \begin{itemize}
@@
 \]
+\item {\bf Step 6}: The approximate score equations
+\begin{equation}
+\label{eq:weighted_esteq}
+\hat U_n (\theta) = \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
+    \sum_{u \in D_i} (1 - w_i (u; \theta)) W_{i,u} \right \} 
+\end{equation}
+where
+\begin{equation}
+\label{eq:waage_weight}
+w_i (t; \theta) = \frac{h_i(t;\theta)}{\pi_i (t) + h_i (t;\theta)}.
+\end{equation}
+Under weights~\eqref{eq:waage_weight} and log-linear intensity
+function,~\eqref{eq:weighted_esteq} is the logistic score function for
+the binary variable~$Y_i(u)$ for $u \in N_i \cup D_i$ and~$i \in [n]$
+where~$Y_i(u) = 1[u \in N_i]$. To see this, we rewrite $\hat U_n
+(\theta)$ as 
+\[
+  - \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+    \left[ {\bf 1} [t \in D_i]  - w_{i} (t; \theta) \right] W_{i,t} 
+= \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+    \left[ {\bf 1} [t \in D_i]  - \frac{1}{1 + \exp\left( - (
+          \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
+    \right] \tilde W_{i,t}.
+\]
+where~$\tilde W_{i,t} = - W_{i,t}$.
+This is equivalent to
+\[
+\sum_{i=1}^n y_i \log p (x_i)
+\]
 \end{itemize}
+
+\begin{thm}[Asymptotic theory]
+For large~$n$, the estimate~$\hat \theta_n$ is approximately normal
+\[
+\sqrt(n) (\hat \theta_n - \theta) \to N(0, \Xi (\theta)^{-1})
+\]
+where
+\[
+  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \left[ \frac{h^{(1)}(s;
+      \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
+  ds.
+\]
+and~$\tau$ is a random 
+\end{thm}
+An estimator for~$\Xi(\theta)$ is
+\[
+  \hat \Xi (\theta) = n^{-1} \sum_{i=1}^n \sum_{u \in N_i \cup D_i} 
+  w_i(u;\theta) (1-w_i(u;\theta)) \left[ \frac{h_i^{(1)}(u;
+      \theta)}{h_i (u; \theta)} \right] \times  \left [ \frac{h_i^{(1)} (u;\theta)}{h_i(u; \theta)} \right]^\top
+\]
+For the Cox model, the sampling-unbiased estimator for~$\hat
+\Xi(\theta)$ is equivalent to the Fisher information for the
+previously described logistic regression model.
+This implies that subsampling from an inhomogeneous Poisson process,
+standard logistic regression software can be used to fit the recurrent
+event model by specifying an offset equal to~$\log (\pi_i (s))$.
+
+\begin{prop}[Optimality of proposed approach]
+\label{prop:optimal}
+If the event process is an inhomogeneous Poisson point process with
+intensity~$h(t; \theta)$ and subsampling occurs via an independent,
+inhomogeneous Poisson point process with intensity~$\pi (t)$,
+then~$\hat U_n (\theta)$ are optimal estimating functions in a class
+of weighted estimating functions given by~\eqref{eq:weighted_esteq}
+with replacing~\eqref{} by any weight function~$w_i (t; \theta)$. This
+class includes the Horvitz-Thompson estimator under~$w(s; \theta)
+= 1$.
+\end{prop}
+
+Proposition~\ref{} follows immediately from~\cite{Rathbun2013}.
+This ensures the only loss of statistical efficiency is related to the 
+subsampling procedure and not the associated methodology.
+
+\subsection{Computation versus statistical efficiency tradeoff}
+
+Under~$\pi(t) = c \times h(t; \theta)$ for $c>0$,
+$w(t; \theta) = c/(c+1)$.  Then
+\[
+\Xi (\theta) = \frac{c}{c+1} \int_0^\tau \frac{ h^{(1)} (t; \theta)
+  h^{(1)} (t; \theta)^\top}{h (t; \theta)} dt = \frac{c}{c+1} \Sigma (\theta)
+\]
+Therefore the relative efficiency compared to the asymptotic variance
+of the maximum likelihood estimator is~$1 + 1/c$.
+Suppose we have~$\pi (t) = c \times \max_{t \in (0,\infty)} h(t;
+\theta)$, then~$w(t;\theta) \leq c/(c+1)$.  this implies
+\[
+\Xi (\theta) \leq \frac{c}{c+1} \Sigma(\theta)
+\]
+which again leads to an upper bound on efficiency of~$c/(1+c)$.
+Recall that the sampling rate is on the order of multiple measurements
+per second.  Suppose the risk is bounded above by $3$ (i.e., $3$ 
+events per hour). Then, we can sample~$30$ times per hour and have
+efficiency of $0.91$; for a 4Hz sensor, this reduces the number of
+samples from~$4 \times 60 \times 60 = 14,400$ per hour to on average
+$30$.  Even if the computational complexity is linear in the number of
+samples, we get $480$ times reduction in the data size at the cost of
+a $0.91$ statistical efficiency. 
+If we sample~$300$ times per hour, then the efficiency loss is only
+$0.99$, with a $48$ times reduction in data size.
+
+\subsection{Time-varying coefficient and prediction}
+
+Replace~$\beta(s)$ by~$\beta_t (s)$ and assume it's smooth in~$t$.
+
+THIS IS WHERE PREDICTION IS IMPORTANT!!
+
+We wish to estimate the risk at non-sampled points.
+
+For prediction, we use sparse FPCA (Yao 2005) 
+\begin{itemize}
+\item Take $\hat c_{k} (s,t)$ for~$t \in N \cup D$ and~$s \in
+  [0,\Delta]$
+\end{itemize}
+
+Take $\{ X (s) \}_{0 < s < \Delta}$. Let~$\psi_{j} (s)$ is an
+orthonormal basis for $L^2 [ 0,1 ]$.  Then
+
+\[
+X(s) - \mu(s) 
+\]
+
+\[
+- b^\prime D b
+\]
+
+\section{Theoretical analysis}
+
+\begin{assumption}[Event process assumptions]
+
+\end{assumption}
+
+\begin{assumption}[Covariance assumptions]
+
+\end{assumption}
+
+\begin{assumption}[Functional data assumptions]
+
+\end{assumption}
+
+\begin{thm}[No reconstruction error]
+
+\end{thm}
+
+\begin{thm}[Reconstruction error but~$\hat \Delta > \Delta$]
+\label{eq:biaseddown}
+\end{thm}
+
+Theorem~\ref{eq:}
+
+\begin{thm}[When~$\hat \Delta < \Delta$]
+
+\end{thm}
+
+\section{Extensions}
+
+\subsection{Re-using estimates under additional sampling}
+Inverse weights can be used to build new pooled covariance.
+Weights are set to
+\[
+\frac{\pi_{\text{new}} (t) + \hat \lambda (t; \theta)}{\pi_{\text{old}} (t) + \hat \lambda (t; \theta)}
+\]
+
+\subsection{Missing data}
+\label{section:missingdata}
+
+\subsection{Multilevel models}
+
+\subsection{Multiple sensors and availability}
+
+\section{Simulations}
+
+\subsection{Unbiased estimation}
+
+\subsection{Impact of $\Delta$}
+
+\section{Case Study: }
+
+\bibliographystyle{plainnat}
+\bibliography{si-fda-refs}
+
+\appendix
+
+\section{Derivation of the design-unbiased score equations}
 
@@
 
-\section{Prediction}
-We wish to estimate the risk at non-sampled points.
-
-For prediction, we use sparse FPCA (Yao 2005) 
-\begin{itemize}
-\item Take $\hat c_{k} (s,t)$ for~$t \in N \cup D$ and~$s \in
-  [0,\Delta]$
-\item 
-\end{itemize}
-
-Take $\{ X (s) \}_{0 < s < \Delta}$. Let~$\psi_{j} (s)$ is an
-orthonormal basis for $L^2 [ 0,1 ]$.  Then
-
-\[
-X(s) - \mu(s) 
-\]
-
-\[
-- b^\prime D b
-\]
-
-\section{Theoretical analysis}
-
-\begin{assumption}[Event process assumptions]
-
-\end{assumption}
-
-\begin{assumption}[Covariance assumptions]
-
-\end{assumption}
-
-\begin{assumption}[Functional data assumptions]
-
-\end{assumption}
-
-\begin{thm}[No reconstruction error]
-
-\end{thm}
-
-\begin{thm}[Reconstruction error but~$\hat \Delta > \Delta$]
-\label{eq:biaseddown}
-\end{thm}
-
-Theorem~\ref{eq:}
-
-\begin{thm}[When~$\hat \Delta < \Delta$]
-
-\end{thm}
-
-\section{Simulations}
-
-\subsection{Unbiased estimation}
-
-\subsection{Unbiased estimation}
-
-\section{Case Study: }
-
-\bibliographystyle{plainnat}
-\bibliography{si-fda-refs}
-
 \end{document}

### 2019-07-02 write-up/fda-recurrent.tex
Fixing up paper: working on proofs of technical results

@@
 
-\newcommand{\sam}[1]{{\color{blue}{#1}}}
-\newcommand{\walt}[1]{\textcolor{red}{[WD:\ #1]}}
+%\newcommand{\sam}[1]{{\color{blue}{#1}}}
+\newcommand{\walt}[1]{\textcolor{blue}{[WD:\ #1]}}
 
@@
 
-\usepackage{soul,color, mathtools}
-
+\usepackage{soul,color, mathtools, fixmath}
+\usepackage{multirow}
 \def\pr{\mathop{\text{pr}}\nolimits}
@@
 however, there is a computation-efficiency trade-off. In particular, a
-higher sampling rate can decrease variance at the cost of increased
-computation.  We show via simulations that the benefits of incredibly
-high sampling rates is often negligible, as the contribution to the
-variation is small in relation to to variation in the underlying
-stochastic processes. Second, optimal estimating equations are derived
-that depend on the instantaneous rate of the recurrent event process,
-which increases efficiency of the subsampling procedure. 
-Finally, implementation can leverage existing, standard software for
-functional data analysis and logistic regression, leading to fast
-adoption by domain scientists.
+higher sampling rate can decrease estimator variance at the cost of
+increased computation.  We show via simulations that the benefits of
+incredibly high sampling rates is often negligible, as the
+contribution to the variation is small in relation to to variation in
+the underlying stochastic processes. Second, derived estimating
+equations are optimal, implying loss of statistical efficiency is only
+due to subsampling procedure and not derived methodology.  Finally,
+implementation can leverage existing, standard software for functional
+data analysis and logistic regression, leading to fast adoption by
+domain scientists.
 
@@
 event times are ordered, i.e., $T_{i,j} < T_{i,j^\prime}$ for $j <
-j^\prime$. The window length, $\tau_i$, is the censoring time and is
-assumed independent of the event process. Let~$N_{i} (t)$ denote the
-associated counting process of $\bfT_{i}$; that is, $N_i (t) =
-\sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.  
+j^\prime$. The observation window length, $\tau_i$, is the censoring
+time and is assumed independent of the event process. Let~$N_{i} (t)$
+denote the associated counting process of $\bfT_{i}$; that is, $N_i
+(t) = \sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.  
 In this section, we assume a single-dimensional health process~$\bfx_i
@@
 activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e.,
-4 times per second).  Given the high frequency nature of the sensor 
-measurements, this paper assumes the process is measured continuously.
+4 times per second).  Given the high frequency nature of sensor data,
+this paper assumes the process is measured continuously.
 
@@
 time-series data.  The main caveat is the additional
-parameter~$\Delta$; however, we show that so long as the
-estimated~$\hat \Delta$ exceeds~$\Delta$, then resulting estimation is
-unbiased albeit at a loss of efficiency.  Moreover, sensitivity
-analysis can be peformed to determine how choice of $\hat \Delta$
-affects inference.  One limitation of the approach presented here is
-that only fully parametric hazard models may be fit to the data.
-However, a spline model for the log baseline hazard affords sufficient
-model flexibility.
+parameter~$\Delta$; however, as long as the estimated~$\hat \Delta$
+exceeds~$\Delta$, then resulting estimation is unbiased albeit at a
+loss of efficiency.  Moreover, sensitivity analysis can be peformed to
+determine how choice of $\hat \Delta$ affects inference.  One
+limitation of the approach presented here is that only fully
+parametric hazard models may be fit to the data. However, a spline
+model for the log baseline hazard affords sufficient model
+flexibility.
 
 \subsection{Measurement-error models with event processes}
+\label{section:memproblems}
 
@@
 $\eta_i (t)$ is the ``true and unobserved value of the longitudinal
-outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional intensity
-function given~$\bfx_i$ is given by
+outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional
+density function of the event process~$\bfT_i$ given~$\bfx_i$ is 
 \[
-\pr \left ( T_{i,j} = t \Given \H_{i,\tau_i}^X, \{ T_{i,j^\prime} < t \}_{j^\prime <
-  j}, T_{i,j} \geq t \right) = E \left( h_i (t \given \eta) \exp
-\left( \int_{T_{i,j-1}}^t h_i (\eta (s)) ds \right) \given \bfx_i
-\right).
+\pr \left ( \bfT_i = \bft \Given \H_{i,\tau_i}^X \right) = E \left(
+  \prod_{j=1}^{k_i} h_i (t_j \given \eta) \exp \left(
+    \int_{0}^{\tau_i} h_i (s \given \eta ) ds \right) \given \bfx_i \right).
 \]
-Therefore, it appears that the conditional hazard function depends 
-not only on past $x$-values, but also on future $x$-values.
-Therefore, the processes~$\bfx$ and~$\bfT$ do not satisfy
+Therefore, the conditional density function depends not only on past
+$x$-values, but also on future $x$-values.
+That is, the processes~$\bfx$ and~$\bfT$ do not satisfy
 \emph{independent evolution}~\citep{DempseyPMCC2}. 
-This is quite unnatural, as~\eqref{eq:jm} suggests current risk
-depends on future values of the sensor process.
-To alleviate this issue, we employ the strategy of taking~$\epsilon
-(t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero Gaussian process.
-The proposed inferential procedure only relies on the Gaussian
-assumption when~$\bfx$ is observed with missingness (see
+This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous
+risk of an event at time~$t$ depends on future values of the sensor
+process. To alleviate this issue, we employ the strategy of
+taking~$\epsilon (t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero
+Gaussian process. The proposed inferential procedure only relies on
+the Gaussian assumption when~$\bfx$ is observed with missingness (see 
 Section~\ref{section:missingdata}).
@@
 observed. Optimal choice of~$\pi_i (t)$ is beyond the scope of this
-paper; however, the literature suggests setting proportional to the
-hazard function~$h_i (t; \theta)$.
+paper; however, simulation studies have suggested setting proportional
+to the hazard function~$h_i (t; \theta)$.
 
@@
 remainder of this paper. Letting
-\[
+\begin{equation}
+\label{eq:waage_weights}
 w_i (t; \theta) = \frac{\pi_i (t)}{\pi_i (t) + h_i (t ; \theta)},
-\]
+\end{equation}
 the resulting estimating equations can be written as
@@
 
+\subsection{Estimation of the windowed covariate history}
+
 We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0
 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup D_i$.
-We use the sandiwch smoother~\citep{Xiao2013} to estimate the
+We use the sandwich smoother~\citep{Xiao2013} to estimate the
 mean~$\mu(s,t) = \E [ X(t,s)]$.
@@
 linear smoother~\citep{Hastie2009} bivariate tensor product
-splines~\cite{Wood2006} and the bivar
+splines~\citep{Wood2006} and the bivariate penalized spline
+smoother~\citep{MarxEilers2005}.
 The sandwich smoother was chosen for its computational efficiency and
@@
 We then define~$\tilde X(s,t) = X(s,t) - \hat \mu(s,t)$ to be the
-mean-zero process at each time~$t \in $.
+mean-zero process at each time~$t \in \bfT_i \cup D_i$.
 
-Next,
+As in~\cite{Park2018}, define the \emph{marginal covariance} by
+\[
+\Sigma (s, s^\prime) = \int_{0}^\tau c( (s,T), (s^\prime, T) ) g(T) dt.
+\]
+for~$0 \leq s,s^\prime \leq \Delta$, where~$c((s,t), (s^\prime,t))$ is
+the covariance function of the windowed covariate history~$X(t,\cdot)$
+and $g(T)$ is the joint sampling density of event and subsampled
+timepoints. 
+Estimation of~$\Sigma$ occurs in two steps. 
+First, the pooled sample covariance is calculated at a set of grid
+points:
+\[
+\tilde \Sigma (s_r , s_{r^\prime}) = \left( \sum_{i=1}^n \left | N_i 
+    \cup D_i \right |\right)^{-1} \left( \sum_{i=1}^n \sum_{t \in N_i
+    \cup D_i} \tilde X (s_r,t) \tilde X (s_{r^\prime}, t) \right).
+\]
+Due to our concern over independent evolution as discussed in
+section~\ref{section:memproblems}, we do not assume that each
+observation is observed with white noise and therefore the diagonal
+elements of~$\hat \Sigma$ are not inflated. Second, the
+estimator~$\hat \Sigma$ is further smoothed again using the sandwich 
+smoother~\citep{Xiao2013}. 
+Note~\cite{Park2018} smooth the off-diagonal elements, while here we
+smooth the entire pooled sample covariance matrix. All negative
+eigenvalues are set to zero to ensure positive semi-definiteness.
+The result is used as an estimator~$\hat \Sigma$ for the pooled
+covariance $\Sigma$.
 
-Use of the marginal covariance~\cite{Park2018} reduces allows us to avoid
-computationally prohibitive computations.
+Next, we take the spectral decomposition of the estimated covariance
+function; let~$\{ \hat \psi_k (s),\hat \lambda_k \}_{k \geq 1}$ be the
+resulting sequence of eigenfunctions and eigenvalues. 
+The key benefit of the marginal covariance approach is that it allows
+us to compute a single, time-invariant basis expansion; this reduces
+the computational burden by avoiding the three dimensional covariance
+function (i.e., covariance depends on~$t$) and associated spectral
+decomposition in methods considered by \cite{ChenMuller2012}.
+Using the Karhunen-Lo{\`e}ve decomposition, we can represent~$X(t,s)$
+by
+\[
+X(s,t) = \hat \mu(s,t) + \sum_{k=1}^{\infty} \hat c_{i,k} (t) \hat
+\psi_k (s) \approx \hat \mu(s,t) + {\bf c}_{i} (t)^\top \mathbold{\hat
+  \psi} (s)
+\]
+where $\hat c_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (s,t) \hat
+\psi_k (s) ds$, ${\bf c}_i (t) = (c_{i,1} (t), \ldots, c_{i,K_x}
+(t))^\top$, $\mathbold{\hat \psi} (s) = (\hat \psi_1 (s), \ldots, \hat
+\psi_{K_x} (s))^\top$, and~$K_x < \infty$ is the truncation level of
+the infinite expansion. Following~\cite{Goldsmith2011}, we set~$K_x$
+to satisfy identifiability constraints (see Section~\ref{section:beta}
+for details).
 
+\subsection{Estimation of~$\beta(s)$}
+\label{section:beta}
+
+The next step of our method is modeling~$\beta(s)$. Here, we leverage
+ideas from the penalized spline literature~\citep{Ruppert2003,
+  Wood2006book}. Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots,
+\phi_{K_b} (x) \}$ be a spline basis such that~$\beta(s) =
+\sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$
+where~${\bf b} = [b_1, \ldots, b_{K_b}]^{\top}$. Thus, the integral
+in~\eqref{eq:hazardlinear} can be restated as 
+\begin{align*}
+\int_{t-\Delta}^t X(s,t) \beta(s) ds 
+  &= \int_{t-\Delta}^t \left[ \hat \mu(s,t) + {\bf c} (t)^\top
+    \mathbold{\hat \psi} (s) \right] \times \left[
+    \mathbold{\phi} (s) {\bf b} \right] \\
+  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\psi, \phi} ] {\bf b}
+\end{align*}
+where~$M_{i} (t) = (M_{i,1} (t), \ldots, M_{i,K_b} (t))$, $M_{i,j} (t)
+= \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\psi, \phi}$ is a
+$K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal
+to~$\int_{0}^\Delta \psi_k (s) \phi_l (s)
+ds$~\citep{RamsaySilverman2005}.
+
+Given the basis for~$\beta(t)$, the model depends on choice of
+both~$K_b$ and~$K_x$.  We follow~\cite{Ruppert2002} by choosing $K_b$
+large enough to prevent undersmoothing and~$K_x \geq K_b$ to satisfy
+identifability constraints.
+In this paper, we follow the simple rule of thumb and set~$K_b = K_x =
+35$.
+As long as the choices of~$K_x$ and $K_b$ are large enough, their
+impact on estimation is typically negligible. 
+We formalize this statement in Theorem~\ref{thm:truncation}.
+Note the choice of~$K_b$ large may lead to overfitting the data.
+We therefore follow~\cite{GoldSmith2015} and induce smoothing by
+assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2 I)$.
+Below, we will exploit a connection between~\eqref{eq:approxscore}
+and score equations for a logistic regression model.  Before moving
+on, we introduce notation:
+\begin{equation*}
+h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) \approx
+\exp \left( Z_{i,t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
+  + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\phi, \psi} {\bf b} \right)
+= \exp \left( W_{i,t}^\top \theta \right),
+\end{equation*}
+where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{i,t}^\top
+\gamma) =: h_0 (t)$ is the parametrized baseline intensity function.
+
+\subsection{Connection to logistic score functions}
+
+We next establish a connection between the above approximate score 
+equations and the score equations for a logistic regression model.
+All proofs are provided in the supplementary materials.
+We can then exploit this connection to allow the model to be fit
+robustly using standard mixed effects software~\citep{Ruppert2002,
+  McCulloch2001}.
+
+\begin{lemma} \normalfont
+\label{lemma:logistic}
+Under weights~\eqref{eq:waage_weights} and the log-linear intensity
+function~\eqref{eq:hazardlinear}, the approximate score
+function~\eqref{eq:approxscore} is equivalent to 
+\[
+\sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i } \left[ {\bf 1}[t \in D_i]
+  - \frac{1}{1 + \exp \left[- \left( \tilde{W}_{i,t}^\top \theta +
+        \log \pi_i (t) \right) \right]} \right] \tilde{W}_{i,t} 
+\]
+where~$\tilde W_{i,t} = -W_{i,t}$. This is the score function for
+logistic regression with binary response~$Y_i(t)$ for $t \in N_i \cup
+D_i$ and~$i \in [n]$ where~$Y_i(t) = 1[t \in N_i]$, offset~$\log \pi_i
+(t)$, and covariates~$\tilde W_{i,t}$. 
+\end{lemma}
+
+In this section, we make the following strong assumption.
+
+\begin{assumption} \normalfont
+\label{assumption:truncation}
+Both the functional covariate~$\{ X(t,s) \}_{0\leq s \leq \Delta}$ and
+the functional coefficient~$\beta(t)$ sit in the span of their
+respective truncated basis expansions. That is, there
+exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$
+and~$X(s,t) = \hat \mu (s,t) + {\bf c}_i (t)^\top \mathbold{\hat \psi}
+(s)$.
+\end{assumption}
+
+Assumption~\ref{assumption:truncation} allows us to ignore the impact
+of truncation and present Lemma~\ref{lemma:simpleasym} which gives a
+simple asymptotic theory in this particular case.
+
+\begin{lemma} \normalfont
+\label{lemma:simpleasym}
+Under Assumption~\ref{assumption:truncation} and~$\Delta$ known, for
+large~$n$ the estimate~$\hat \theta_n$ is approximately normal
+\[
+\sqrt{n} (\hat \theta - \theta) \to N(0, \Xi (\theta)^{-1})
+\]
+where
+\[
+  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \left[ \frac{h^{(1)}(s;
+      \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
+  ds.
+\]
+and~$\tau$ is the random censoring time of the event process.
+\end{lemma}
+An estimator for~$\Xi(\theta)$ is
+\begin{equation}
+\label{eq:fisher}
+  \hat \Xi (\theta) = n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i} 
+  w_i(t;\theta) (1-w_i(t;\theta)) \left[ \frac{h_i^{(1)}(t;
+      \theta)}{h_i (t; \theta)} \right] \times  \left [
+    \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top
+\end{equation}
+For the log-linear intensity model, the sampling-unbiased estimator
+for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the
+previously described logistic regression model.
+This implies that subsampling from an inhomogeneous Poisson process,
+standard logistic regression software can be used to fit the recurrent
+event model by specifying an offset equal to~$\log \pi_i (t)$.
+Not only this, Lemma~\ref{prop:optimal} shows
+weights~\eqref{eq:waage_weights} are optimal within a particular class
+of weighted estimating equations.
+
+\begin{prop} \normalfont
+\label{prop:optimal}
+If the event process is an inhomogeneous Poisson point process with
+intensity~$h(t; \theta)$ and subsampling occurs via an independent,
+inhomogeneous Poisson point process with intensity~$\pi (t)$,
+then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most
+efficient) in the class of weighted estimating functions given
+by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any
+weight function~$w_i (t; \theta)$. This class includes the
+Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
+\end{prop}
+
+\noindent Proposition~\ref{prop:optimal} ensures the only loss of
+statistical efficiency is due to subsampling and not using a
+suboptimal estimation procedure given subsampling.
+
+\subsection{Computation versus statistical efficiency tradeoff}
+
+Under assumption~\ref{assumption:truncation}, we next consider the
+statistical efficiency of our proposed estimator when compared to
+complete-data maximum likelihood estimation.
+While subsampling introduces additional variation, it may
+significantly reduce the overall computational burden. It is this
+trade-off that we next make precise.
+In particular, we consider the following choice of subsampling
+rate,~$\pi(t) = c \times h(t; \theta)$ for $c>0$. That is, the
+subsampling rate is proportional to the intensity function with
+time-independent constant~$c > 0$.
+Under this subsampling rate, the weight
+function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.  
+Under Lemma~\ref{lemma:simpleasym}, 
+\[
+\Xi (\theta) = \frac{c}{c+1} \int_0^\tau \frac{ h^{(1)} (t; \theta)
+  h^{(1)} (t; \theta)^\top}{h (t; \theta)} dt = \frac{c}{c+1} \Sigma (\theta)
+\]
+where~$\Sigma(\theta)$ is the Fisher information of the complete-data 
+maximum likelihood estimator.
+Therefore the relative efficiency is~$c/(1+c)$. For an upper bound~$H
+= \max_{t \in (0,\tau)} h(t;\theta)$, if we set~$\pi (t) = c \times
+H$, then the relative efficiency can be lower bounded by~$c / (c+1)$.
+
+Recall sensor measurements occur on the order of multiple times per
+second.  Suppose the intensity rate is bounded above by $3$ and the
+unit time scale was hours. If we can subsample the data at~$30$
+times per hour, then we have a lower bound on the efficiency of
+$0.909$. For a 4Hz sensor, this reduces the number of samples per hour
+from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
+While the computational complexity of logistic regression is linear in
+the number of samples, we get $480$ times reduction in the data size
+at the cost of a $0.909$ statistical efficiency. If we sample~$300$
+times per hour, then the efficiency loss is only $0.999$, with a $48$
+times reduction in data size. Table~\ref{tab:compvseff} provides
+additional examples for a 4Hz sensor.  The data reduction depends on
+the sensing rate; however, the lower bound on statistical efficiency
+does not.  This is because the subsampling rate only depends on the
+upper bound of the intensity function. That is, if the events are rare
+then the amount of data we need to subsample is greatly reduced with
+no impact to statistical efficiency.
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{l r r r r r | c}
+\multirow{2}{2.5cm}{Subsampling constant ($c$)} 
+  & \multicolumn{5}{c}{Upper bound on intensity rate per
+    hour} 
+  & \multirow{2}{2cm}{Statistical efficiency}\\ \cline{2-6} 
+& 0.5 & 1 & 3 & 5 & 10 \\ \hline
+1 & 28800.0 & 14400.0 & 4800.0 & 2880.0 & 1440.0 & 0.500 \\
+5 & 5760.0 & 2880.0 & 960.0 & 576.0 & 288.0 & 0.833 \\
+10 & 2880.0 & 1440.0 & 480.0 & 288.0 & 144.0 & 0.909 \\
+100 & 288.0 & 144.0 & 48.0 & 28.8 & 14.4 & 0.990 \\
+500 & 57.6 & 28.8 & 9.6 & 5.8 & 2.9 & 0.999 \\ \hline
+\end{tabular}
+\caption{Data reduction (total number of measurements divided by
+  expected number of subsampled measurements) given 4Hz sensor,
+  subsampling constant~$c$ and an upper bound on the intensity
+  rate per hour~$H$. The subsampling rate is set at~$c \times H$;
+  note, lower bound on statistical efficiency only depends on the
+  subsampling rate.}   
+\label{tab:compvseff}
+\end{table}
+
+\section{Theoretical analysis}
+
+\begin{assumption}[Event process assumptions]
 \begin{itemize}
+\item The subsampling rate~$\pi_i (t) \in [L_\pi, U_\pi]$ 
+\item There exists a nonnegative matrix
+\item There exists 
+\end{itemize}
+\end{assumption}
+
+\begin{assumption}[Covariance assumptions]
+\begin{itemize}
+\item 
+\end{itemize}
+\end{assumption}
+
+\begin{assumption}[Functional data assumptions]
+M1 -- M4 in 
+\end{assumption}
+
+\begin{thm}[No reconstruction error]
+\label{thm:truncation}
+Under regularity conditions A-E in Andersen et al. 1993, 
+\[
+\frac{n (\hat \theta - \theta) \Xi (\theta) (\hat \theta - \theta) -
+  K_n+1}{\sqrt{2 (K_n + 1) }} \to N(0,1).
+\]
+
+\end{thm}
+
+\begin{thm}[Reconstruction error OR $\hat \Delta > \Delta$]
+\label{thm:biaseddown}
+
+\end{thm}
+
+Theorem~\ref{thm:biaseddown} shows that if we keep~$K_x = K_b = 35$
+and therefore have reconstruction error, we can expect a downward bias
+in parameter estimates.
+Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias
+in the parameter estimates.  For the former, however, we can simply
+let~$K_x = K_b$ grow as a function of data size; for the latter, we
+can see if~$\hat \beta(s)$ is statistically significant for~$s$
+near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the
+condition holds.
+
+\section{Extensions}
+
+In this section, we demonstrate the flexibility of our approach by
+exploring extensions in several important directions to ensure these
+methods are robust for practical use.
+This section will continue to leverage the connection to generalized
+functional linear models provided by Lemma~\ref{lemma:logistic}.
+
+\subsection{Missing data}
+\label{section:missingdata}
+
+A core concern in mHealth is that sensor data can be missing due to
+sensor wearing issue.
+To see this, for the suicidal ideation case study, there are XX event
+times across the whole dataset.  Of these, XX event times had complete 
+data for the prior thirty minutes, XX had missingness less than 10\%
+(i.e., sufficient to ), and XX had missingness above 30\%.
+
+The issue here is that~$c_{i,k} (t)$ cannot be estimated if~$X(s,t)$
+is not observed for all~$s \in [0,\Delta]$. 
+Moreover, we want standard erros to reflect the uncertainty
+in these coefficients when missing data is prevalent.
+\cite{Goldsmith2015} suggest using best linear unbiased predictors
+(BLUP) or posterior modes in the mixed effects model to
+estimate~$c_{i,k} (t)$; however, one criticism would be when there is
+substantial variability in these estimates.  To deal with this,
+\cite{Crainiceanu2010, Goldsmith2010} take a full Bayesian analysis.
+\cite{Yao2005} introduced PACE as a frequentist analysis that deals
+with this issue.
+However,~\cite{Reimherr2018} show that for sparse, irregular
+longitudinal, the imputation model should not ignore the outcome~$Y_i
+(t)$.
+
+Here we present an extension of~\cite{Reimherr2018} to our setting,
+leverage Lemma~\ref{lemma:logistic} and the marginal covariance
+estimation procedure to build a multiple imputation approach.
+\[
+\E [ X_i (s,t) \given Y_i(t) = y, \bfx ] = \mu_y (s,t) + {\bf a}_{i,t}^\top
+{\bf B}_i (\bfx_i (t) - \mathbold{\mu}_i )
+\]
+where~$\mu_y (t) = \E [ X_i (t) \given Y_i (t) = y]$, and
+\[
+{\bf a}_i^\top = ( \Sigma(s, )
+\]
+This generalizes the procedure of~\cite{Reimerr2018} to the
+longitudinal data setting.
+This is possible due to LEmma~\ref{} and our 
+Here, we ignore the issue that~$X(t,s)$ is
+
+The estimate is then given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat
+\theta^{(m)}$; the covariance matrix can be estimated by
+\[
+M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
+\sum_{m=1}^M \hat \theta - \bar \theta) (\hat \theta - \bar \theta)
+\]
+where~$\hat \Xi^{(m)}$ is the estimator~\eqref{eq:fisher} for
+the~$m$th imputated dataset.
+
+\subsection{Multilevel models}
+
+\subsection{Multiple sensors and availability}
+
+\subsection{Re-using estimates under additional sampling}
+
+If~$\pi$ is constant then we can continue to use the same pooled
+covariance.
+
+\section{Simulations}
+
+\subsection{Unbiased estimation}
+
+\subsection{Impact of $\Delta$}
+
+\section{Case Study: }
+
+\bibliographystyle{plainnat}
+\bibliography{si-fda-refs}
+
+\appendix
+
+\section{Derivation of the design-unbiased score equations}
+
+\begin{proof}[Proof of Lemma~\ref{lemma:logistic}]
+Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
+  \theta)}{h_i (t; \theta)}$. Therefore, under the log-linear
+intensity function, $\frac{d}{d\theta} \log h_i (t; \theta) =
+W_{i,t}$. On the other hand,
+\[
+\frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) = 
+\frac{\exp \left( W_{i,t}^\prime \theta \right)}{\pi_i (t) + \exp
+  \left( W_{i,t}^\prime \theta \right)} W_{i,t}
+\]
+Therefore,
+\begin{align*}
+\hat U_n (\theta) 
+  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
+    \sum_{t \in N_i \cup D_i} \frac{\exp \left( W_{i,t}^\prime \theta
+    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\prime \theta \right)}
+    W_{i,t} \right \} \\
+  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
+    + \exp \left( W_{i,t}^\prime \theta \right)} W_{i,t} -
+    \sum_{t \in D_i} \frac{\exp \left( W_{i,t}^\prime \theta
+    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\prime \theta \right)}
+    W_{i,t} \right \} \\
+  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
+    \sum_{t \in D_i} (1 - w_i (t; \theta)) W_{i,t} \right \} \\
+  &= - \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \left[ {\bf 1} [t \in
+    D_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
+  &= \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+    \left[ {\bf 1} [t \in D_i]  - \frac{1}{1 + \exp\left( - (
+          \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
+    \right] \tilde W_{i,t}.
+\end{align*}
+where~$\tilde W_{i,t} = - W_{i,t}$. This is exactly the score equation
+for logistic regression with offset~$\log \pi_i (t)$.
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
+Lemma~\ref{lemma:simpleasym} will be proven under regularity
+conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
+following assumptions related to the subsampling rate~$\pi_i (t)$.
+For simplicty we assume there exists~$\tau < \infty$ and all censoring
+times satisfy~$\tau_i \leq \tau$. Define $R_i (t)$ to be the at risk
+indicator for participant~$i$.  That is, 
+\[
+R_i (t) = \left \{ \begin{array}{c c} 1 & t \in (0,\tau_i) \\ 0
+                                        & t > \tau_i\end{array} \right .
+\]
+We define the joint counting process
+\[
+M_i (t) = N_i (t) - \int_0^t (h_i (s; \theta) + \pi_i (s)) R_i (s) ds. 
+\]
+where~$N_i (t)$ is the counting process with jumps at all~$t \in
+\bfT_i \cup D_i$.
+
+Asymptotic consistency is guaranteed~\cite[Theorem VI.1.1]{Andersen1993} if
+\begin{equation}
+\label{eq:andersen1}
+n^{-1} \hat U_n (\theta_0 ) \overset{P}{\to} 0,
+\end{equation}
+and 
+\[
+n^{-1} \frac{d}{d \theta^{\top}} \hat U_n (\theta) \big |_{\theta =
+  \theta_0} = \Xi (\theta),
+\]
+and
+\[
+\lim_{n \to \infty} P \left(  and all \theta \in \Theta_0 \right) = 1
+\]
+To prove~\eqref{eq:andersen1}, we decompose~$\hat U_n (\theta_0)$ into
+three terms
+\[
+n^{-1} \hat U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
+  \tilde U_n (\theta) - U_n (\theta) \right) + n^{-1} \left(\hat U_n
+  (\theta) - \tilde U_n (\theta) \right)
+\]
+where~$\tilde U_n (\theta_0)$ are the logistic score equations with
+$\int_0^\infty X_i(t-s) \beta(s) ds$ replaced by the
+approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$. 
+\cite{Andersen1993}[Theorem VI.1.1] show that~$n^{-1} U_n (\theta_0)
+\to 0$. The second term
+\begin{align*}
+n^{-1} \left( \hat U_n (\theta_0) - U_n (\theta_0) \right) =
+  \frac{1}{n} \sum_{i=1}^n \int_0^{\tau} \frac{h^{(1)} (s;
+  \theta_0)}{\pi_i (s) + h (s;\theta_0)} d M_i (s). 
+\end{align*}
+Lenglart's inequality\walt{Need to make precise} implies
+the second term converges in probability to zero.  This statement
+requires the following two conditions:
+\begin{enumerate}
+\item The subsampling rate is both lower and upper bounded for all
+  ``at risk times''; that is,~$0 < L < \pi_i (t) < U < \infty$ for
+  all~$i=1,2,\ldots$ and~$t \in [0,\tau]$. \walt{Could we just bound
+    $h_i (t; \theta) + \pi_i (t)$?}
+\item There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
+\[
+n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
+\theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
+      \theta)}{h_i (t; \theta)} \right] \times  \left [
+    \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
+  dt \overset{P}{\to} \Xi (\theta).
+\] 
+\end{enumerate}
+The third term satisfies
+\begin{align*}
+n^{-1} \left \| \hat U_n (\theta_0) - \tilde U_n (\theta_0) \right \|
+  &\leq M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+    \left \| g \left( W_{i,t}^\top \theta + \log (\pi_i (t)
+  ) \right) - g \left( \tilde W_{i,t}^\top \theta + \log (\pi_i (t) 
+  ) \right) \right \| \\
+&= M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+    \left \| g^\prime \left( W_{i,t}^\top \theta + \log (\pi_i (t) )
+  \right) \right\| \times \left \| W_{i,t}^\top \theta_0 - \tilde W_{i,t}^\top
+  \tilde \theta_0  \right \| \\
+&= \frac{M}{4} \| N_i \cup D_i \| \left ( \sum_{j=p_n+1}^\infty
+  \lambda_{j}^2 \sum_{j=q_n+1}^\infty b_j^2 \right)^{1/2}
+\end{align*}
+where~$M = \sup \| \tilde W_{i,t} \|$,~$g(x) = 1/(1+\exp(-x))$ is the
+expit function, and the final line is by Cauchy-Schwarz inequality.
+
+\end{proof}
+
+\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
+Consider the weighted estimating equations given by~\eqref{eq:approxscore}.
+Then
+\end{proof}
+
+\end{document}
+
 \item {\bf Step 1}: Estimate~$\hat \mu(s,t)$ via sandwich smoother for
@@
 Under weights~\eqref{eq:waage_weight} and log-linear intensity
-function,~\eqref{eq:weighted_esteq} is the logistic score function for
-the binary variable~$Y_i(u)$ for $u \in N_i \cup D_i$ and~$i \in [n]$
-where~$Y_i(u) = 1[u \in N_i]$. To see this, we rewrite $\hat U_n
-(\theta)$ as 
-\[
-  - \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
-    \left[ {\bf 1} [t \in D_i]  - w_{i} (t; \theta) \right] W_{i,t} 
-= \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
-    \left[ {\bf 1} [t \in D_i]  - \frac{1}{1 + \exp\left( - (
-          \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
-    \right] \tilde W_{i,t}.
-\]
-where~$\tilde W_{i,t} = - W_{i,t}$.
-This is equivalent to
-\[
-\sum_{i=1}^n y_i \log p (x_i)
-\]
+function,~\eqref{eq:weighted_esteq} is the 
 \end{itemize}
-
-\begin{thm}[Asymptotic theory]
-For large~$n$, the estimate~$\hat \theta_n$ is approximately normal
-\[
-\sqrt(n) (\hat \theta_n - \theta) \to N(0, \Xi (\theta)^{-1})
-\]
-where
-\[
-  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \left[ \frac{h^{(1)}(s;
-      \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
-  ds.
-\]
-and~$\tau$ is a random 
-\end{thm}
-An estimator for~$\Xi(\theta)$ is
-\[
-  \hat \Xi (\theta) = n^{-1} \sum_{i=1}^n \sum_{u \in N_i \cup D_i} 
-  w_i(u;\theta) (1-w_i(u;\theta)) \left[ \frac{h_i^{(1)}(u;
-      \theta)}{h_i (u; \theta)} \right] \times  \left [ \frac{h_i^{(1)} (u;\theta)}{h_i(u; \theta)} \right]^\top
-\]
-For the Cox model, the sampling-unbiased estimator for~$\hat
-\Xi(\theta)$ is equivalent to the Fisher information for the
-previously described logistic regression model.
-This implies that subsampling from an inhomogeneous Poisson process,
-standard logistic regression software can be used to fit the recurrent
-event model by specifying an offset equal to~$\log (\pi_i (s))$.
-
-\begin{prop}[Optimality of proposed approach]
-\label{prop:optimal}
-If the event process is an inhomogeneous Poisson point process with
-intensity~$h(t; \theta)$ and subsampling occurs via an independent,
-inhomogeneous Poisson point process with intensity~$\pi (t)$,
-then~$\hat U_n (\theta)$ are optimal estimating functions in a class
-of weighted estimating functions given by~\eqref{eq:weighted_esteq}
-with replacing~\eqref{} by any weight function~$w_i (t; \theta)$. This
-class includes the Horvitz-Thompson estimator under~$w(s; \theta)
-= 1$.
-\end{prop}
-
-Proposition~\ref{} follows immediately from~\cite{Rathbun2013}.
-This ensures the only loss of statistical efficiency is related to the 
-subsampling procedure and not the associated methodology.
-
-\subsection{Computation versus statistical efficiency tradeoff}
-
-Under~$\pi(t) = c \times h(t; \theta)$ for $c>0$,
-$w(t; \theta) = c/(c+1)$.  Then
-\[
-\Xi (\theta) = \frac{c}{c+1} \int_0^\tau \frac{ h^{(1)} (t; \theta)
-  h^{(1)} (t; \theta)^\top}{h (t; \theta)} dt = \frac{c}{c+1} \Sigma (\theta)
-\]
-Therefore the relative efficiency compared to the asymptotic variance
-of the maximum likelihood estimator is~$1 + 1/c$.
-Suppose we have~$\pi (t) = c \times \max_{t \in (0,\infty)} h(t;
-\theta)$, then~$w(t;\theta) \leq c/(c+1)$.  this implies
-\[
-\Xi (\theta) \leq \frac{c}{c+1} \Sigma(\theta)
-\]
-which again leads to an upper bound on efficiency of~$c/(1+c)$.
-Recall that the sampling rate is on the order of multiple measurements
-per second.  Suppose the risk is bounded above by $3$ (i.e., $3$ 
-events per hour). Then, we can sample~$30$ times per hour and have
-efficiency of $0.91$; for a 4Hz sensor, this reduces the number of
-samples from~$4 \times 60 \times 60 = 14,400$ per hour to on average
-$30$.  Even if the computational complexity is linear in the number of
-samples, we get $480$ times reduction in the data size at the cost of
-a $0.91$ statistical efficiency. 
-If we sample~$300$ times per hour, then the efficiency loss is only
-$0.99$, with a $48$ times reduction in data size.
-
-\subsection{Time-varying coefficient and prediction}
-
-Replace~$\beta(s)$ by~$\beta_t (s)$ and assume it's smooth in~$t$.
-
-THIS IS WHERE PREDICTION IS IMPORTANT!!
-
-We wish to estimate the risk at non-sampled points.
-
-For prediction, we use sparse FPCA (Yao 2005) 
-\begin{itemize}
-\item Take $\hat c_{k} (s,t)$ for~$t \in N \cup D$ and~$s \in
-  [0,\Delta]$
-\end{itemize}
-
-Take $\{ X (s) \}_{0 < s < \Delta}$. Let~$\psi_{j} (s)$ is an
-orthonormal basis for $L^2 [ 0,1 ]$.  Then
-
-\[
-X(s) - \mu(s) 
-\]
-
-\[
-- b^\prime D b
-\]
-
-\section{Theoretical analysis}
-
-\begin{assumption}[Event process assumptions]
-
-\end{assumption}
-
-\begin{assumption}[Covariance assumptions]
-
-\end{assumption}
-
-\begin{assumption}[Functional data assumptions]
-
-\end{assumption}
-
-\begin{thm}[No reconstruction error]
-
-\end{thm}
-
-\begin{thm}[Reconstruction error but~$\hat \Delta > \Delta$]
-\label{eq:biaseddown}
-\end{thm}
-
-Theorem~\ref{eq:}
-
-\begin{thm}[When~$\hat \Delta < \Delta$]
-
-\end{thm}
-
-\section{Extensions}
-
-\subsection{Re-using estimates under additional sampling}
-Inverse weights can be used to build new pooled covariance.
-Weights are set to
-\[
-\frac{\pi_{\text{new}} (t) + \hat \lambda (t; \theta)}{\pi_{\text{old}} (t) + \hat \lambda (t; \theta)}
-\]
-
-\subsection{Missing data}
-\label{section:missingdata}
-
-\subsection{Multilevel models}
-
-\subsection{Multiple sensors and availability}
-
-\section{Simulations}
-
-\subsection{Unbiased estimation}
-
-\subsection{Impact of $\Delta$}
-
-\section{Case Study: }
-
-\bibliographystyle{plainnat}
-\bibliography{si-fda-refs}
-
-\appendix
-
-\section{Derivation of the design-unbiased score equations}
-
-We re-write the hazard function as:
-\[
-h_i (t; \theta) = \exp \left( Z_{i,t}^\prime, \alpha + [M_{i,t} + C_{i,t}^\prime
-  J_{\phi, \psi} ] \beta \right) = \exp \left( W_{i,t}^\prime \theta \right),
-\]
-where~$W_{i,t}^\prime = [ Z_{i,t}^\prime (M_{i,t}^\prime +
-C_{i,t}^\prime J_{\phi, \psi}) ]$ and~$\theta = (\alpha,
-\beta)$. Then, the score function is:
-\[
-\hat U_n (\theta) = \sum_{i=1}^n \left \{  \sum_{t \in N_i}
-  \frac{h_i^{(1)} (t; \theta)}{h_i (t; \theta)} - \hat H_i^{(1)}
-  (\tau_i, 0; \theta) 
-\right \}
-\]
-where
-\[
-\hat H_i^{(1)} (\tau_i; \theta) = \sum_{u \in N_i \cup D_i} \frac{
-  h^{(1)}_i (u; \theta)}{\pi_i (u) + h_i (u; \theta)}.
-\]
-Note that
-\[
-\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
-  \theta)}{h_i (t; \theta)}.
-\]
-Therefore, these terms have very simple forms!!! 
-\[
-\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}
-\]
-On the other hand,
-\[
-\frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) = 
-\frac{\exp \left( W_{i,t}^\prime \theta \right)}{\pi_i (t) + \exp
-  \left( W_{i,t}^\prime \theta \right)} W_{i,t}
-\]
-Therefore,
-\begin{align*}
-\hat U_n (\theta) 
-  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
-    \sum_{u \in N_i \cup D_i} \frac{\exp \left( W_{i,u}^\prime \theta
-    \right)}{\pi_i (t) + \exp \left( W_{i,u}^\prime \theta \right)}
-    W_{i,u} \right \} \\
-  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
-    + \exp \left( W_{i,u}^\prime \theta \right)} W_{i,t} -
-    \sum_{u \in D_i} \frac{\exp \left( W_{i,u}^\prime \theta
-    \right)}{\pi_i (t) + \exp \left( W_{i,u}^\prime \theta \right)}
-    W_{i,u} \right \} \\
-  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
-    \sum_{u \in D_i} (1 - w_i (u; \theta)) W_{i,u} \right \} \\
-\end{align*}
-
-\end{document}

### 2019-07-03 methods/mean_cov_estimation.R
Fixed adding a row so now 4:1804 is the sensor data

@@ -13,14 +13,17 @@
 ## hatX(t,s) = coefficients%*%eigenvectors + mu(t,s) 
 ## And seeing how close hatX(t,s) is to X(t,s).
 
-setwd("/Volumes/murphy_lab/users/wdempsey/fda-recurrent-events/methods/")
-event_complete = readRDS("./event_complete.RDS")
+setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
+event_complete = readRDS("./event_complete_2019-07-01.RDS")
 sequence = seq(-30,0,by =1/60)
-plot(sequence, event_complete[3,3:1803], type = "l")
+plot(sequence, event_complete[3,4:1804], type = "l")
 
 library(refund)
 full_obs = apply(X = event_complete, MARGIN = 1, FUN = function(x){!any(is.na(x))})
-Y = event_complete[full_obs,3:1803]
+greater_than_ten_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[4:1804])) > 0.1})
+greater_than_thirty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[4:1804])) > 0.3})
+greater_than_fifty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[4:1804])) > 0.5})
+Y = event_complete[full_obs,4:1804]
 x = event_complete[full_obs,2]
 z = sequence
 est <- fbps(Y,list(x=x,z=z))

### 2019-07-03 methods/mean_cov_estimation.R
Fixed how saving works

@@ -15,15 +15,15 @@
 
 setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
 event_complete = readRDS("./event_complete_2019-07-01.RDS")
-sequence = seq(-30,0,by =1/60)
-plot(sequence, event_complete[3,4:1804], type = "l")
+sequence = seq(-30,0,by =1/60); sensor_obs = 4:1804
+plot(sequence, event_complete[3,sensor_obs], type = "l")
 
 library(refund)
 full_obs = apply(X = event_complete, MARGIN = 1, FUN = function(x){!any(is.na(x))})
-greater_than_ten_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[4:1804])) > 0.1})
-greater_than_thirty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[4:1804])) > 0.3})
-greater_than_fifty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[4:1804])) > 0.5})
-Y = event_complete[full_obs,4:1804]
+greater_than_ten_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.1})
+greater_than_thirty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.3})
+greater_than_fifty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.5})
+Y = event_complete[full_obs,sensor_obs]
 x = event_complete[full_obs,2]
 z = sequence
 est <- fbps(Y,list(x=x,z=z))
@@ -60,35 +60,35 @@ plot(sequence, phi_vectors%*%coef[obs,]/inflation + est$Yhat[obs,], type = "l")
 lines(sequence, Y[obs,], col = "red", lwd = 2)
 
 ## SAVE THE MEAN, COEFFICIENT MATRIX, AND FIRST K EIGEN VECTORS
-saveRDS(object = est$Yhat, file = "event_means.RDS")
-saveRDS(object = coef/inflation, file = "event_coef_matrix.RDS")
-saveRDS(object = phi_vectors, file = "event_eigen_vectors.RDS")
[…]

### 2019-07-03 write-up/fda-recurrent.tex

@@
-\documentclass[11pt]{amsart}
+\documentclass[12pt]{amsart}
 \usepackage{amsmath}
@@
 \usepackage{multirow}
+\usepackage{enumitem}
 \def\pr{\mathop{\text{pr}}\nolimits}
@@
 
-\title{Recurrent event analysis in the presence of real-time high
-  frequency data via random subsampling}
+\title[Recurrent event analysis with functional covariates via random
+subsampling]{Recurrent event analysis in the presence of real-time
+  high frequency data via random subsampling}
 \author{Walter Dempsey}
@@
   Wood2006book}. Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots,
-\phi_{K_b} (x) \}$ be a spline basis such that~$\beta(s) =
+\phi_{K_b} (x) \}$ be a spline basis and assume that~$\beta(s) =
 \sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$
@@
 \int_{t-\Delta}^t X(s,t) \beta(s) ds 
-  &= \int_{t-\Delta}^t \left[ \hat \mu(s,t) + {\bf c} (t)^\top
+  &\approx \int_{t-\Delta}^t \left[ \hat \mu(s,t) + {\bf c} (t)^\top
     \mathbold{\hat \psi} (s) \right] \times \left[
     \mathbold{\phi} (s) {\bf b} \right] \\
-  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\psi, \phi} ] {\bf b}
+  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\hat \psi, \phi} ] {\bf b}
 \end{align*}
-where~$M_{i} (t) = (M_{i,1} (t), \ldots, M_{i,K_b} (t))$, $M_{i,j} (t)
-= \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\psi, \phi}$ is a
-$K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal
-to~$\int_{0}^\Delta \psi_k (s) \phi_l (s)
+where~$M_{i,t} = (M_{i,1,t}, \ldots, M_{i,K_b,t})$, $M_{i,j,t}
+= \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\hat \psi,
+  \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th
+entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s)
 ds$~\citep{RamsaySilverman2005}.
@@
 identifability constraints.
-In this paper, we follow the simple rule of thumb and set~$K_b = K_x =
-35$.
-As long as the choices of~$K_x$ and $K_b$ are large enough, their
-impact on estimation is typically negligible. 
-We formalize this statement in Theorem~\ref{thm:truncation}.
-Note the choice of~$K_b$ large may lead to overfitting the data.
-We therefore follow~\cite{GoldSmith2015} and induce smoothing by
-assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2 I)$.
+While our theoretical analysis considers truncation levels that depend
+on~$n$, in practice, we follow the simple rule of thumb and set~$K_b 
+= K_x = 35$. As long as the choices of~$K_x$ and $K_b$ are large
+enough, their impact on estimation is typically negligible. 
+%We formalize this statement in Theorem~\ref{thm:truncation}.
 Below, we will exploit a connection between~\eqref{eq:approxscore}
@@
 on, we introduce notation:
-\begin{equation*}
+\begin{equation}
+\label{eq:approx_hazard}
 h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) \approx
 \exp \left( Z_{i,t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\phi, \psi} {\bf b} \right)
+  + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\hat \psi, \phi} {\bf b} \right)
 = \exp \left( W_{i,t}^\top \theta \right),
-\end{equation*}
-where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{i,t}^\top
-\gamma) =: h_0 (t)$ is the parametrized baseline intensity function.
+\end{equation}
+where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp (
+Z_{i,t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline
+intensity function. We write~$\tilde U_n (\theta)$ to denote
+the approximate score function when substituting
+in~\eqref{eq:approx_hazard} for~\eqref{eq:hazardlinear}.
 
@@
 We next establish a connection between the above approximate score 
-equations and the score equations for a logistic regression model.
-All proofs are provided in the supplementary materials.
-We can then exploit this connection to allow the model to be fit
-robustly using standard mixed effects software~\citep{Ruppert2002,
-  McCulloch2001}.
+equations~$\tilde U_n (\theta)$ and the score equations for a logistic
+regression model. We can then exploit this connection to allow the
+model to be fit robustly using standard mixed effects
+software~\citep{Ruppert2002, McCulloch2001}.
 
@@
 Under weights~\eqref{eq:waage_weights} and the log-linear intensity
-function~\eqref{eq:hazardlinear}, the approximate score
-function~\eqref{eq:approxscore} is equivalent to 
+function~\eqref{eq:approx_hazard}, the approximate score
+function~$\tilde U_n (\theta)$ is equivalent to 
 \[
@@
 
-In this section, we make the following strong assumption.
-
-\begin{assumption} \normalfont
+\subsection{Theoretical analysis}
+
+Our theoretical analysis requires the several assumptions, some
+regarding the subsampling procedure, some for the event process, and
+some for the functional data. We state these assumptions and then our
+main theorems.  
+We start by assuming there exists a~$\tau < \infty$ such that all
+individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
+Moreover, define $R_i (t)$ to be the at risk indicator for
+participant~$i$.  That is,   
+\[
+R_i (t) = \left \{ \begin{array}{c c} 1 & t \in (0,\tau_i) \\ 0
+                                        & t > \tau_i\end{array} \right .
+\]
+Lemma~\ref{lemma:simpleasym} will be proven under regularity
+conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
+following additional assumptions:
+
+\begin{assumption}[Event process assumptions] 
+\label{assumption:events}\normalfont
+We assume the following holds:
+\begin{enumerate}[label=(E.\arabic*)]
+\item\label{E1} The subsampling rate is both lower and upper
+  bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U <
+  \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$. 
+\item\label{E2} There exists a nonnegative definite matrix~$\Xi
+  (\theta)$ such that
+  \[
+    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
+    \theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
+        \theta)}{h_i (t; \theta)} \right] \times  \left [
+      \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
+    dt \overset{P}{\to} \Xi (\theta).
+  \] 
+\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for
+  all~$(i,j,t)$.
+\item\label{E4} For all~$j,k$ 
+\[
+n^{-1} \sum_{i=1}^n \int_0^\tau \left | \frac{d^2}{d\theta_j
+    d\theta_k} h_i (t;\theta_0) \right|^2 R_i (t) dt \overset{P}{\to}
+C < \infty
+\]
+as $n \to \infty$.
+\end{enumerate}
+\end{assumption}
+
+We also require several assumptions due to the truncation of the
+Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
+\begin{assumption}[Functional assumptions~\citep{Park2018}] \normalfont
 \label{assumption:truncation}
-Both the functional covariate~$\{ X(t,s) \}_{0\leq s \leq \Delta}$ and
-the functional coefficient~$\beta(t)$ sit in the span of their
-respective truncated basis expansions. That is, there
-exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$
-and~$X(s,t) = \hat \mu (s,t) + {\bf c}_i (t)^\top \mathbold{\hat \psi}
-(s)$.
+The following assumptions are standard in prior work on longitudinal
+functional data analysis~\citep{Park2018, Yao2005, ChenMuller2012}:
+\begin{enumerate}[label=(A.\arabic*)]
+\item\label{A1} $X = \{ X(s, T) : (s,T) \in \mathcal{S} \times
+  \mathcal{T} \}$ is a square integrable element of the $L^2 (
+  \mathcal{S} \times \mathcal{T})$
+\item\label{A2} The sampling density~$g(T)$ is continuous
+  and~$\sup |g(T)| < \infty$
+\item\label{A3} $\E[X(s,T) X(s^\prime, T) X(s,T^\prime)
+  X(s^\prime, T^\prime) ] < \infty$ for
+  each $s,s^\prime [0,\Delta]$ and~$0 < T, T^\prime < \tau$.
+\item\label{A4} $\E[\|X(\cdot, T)\|^4] < \infty$ for each~$0< T
+  < \tau$.
+\end{enumerate}
 \end{assumption}
-
-Assumption~\ref{assumption:truncation} allows us to ignore the impact
-of truncation and present Lemma~\ref{lemma:simpleasym} which gives a
-simple asymptotic theory in this particular case.
-
+Assumption~\ref{A3} is guaranteed for by~\ref{E1} and~\ref{E3}, but is
+stated here to emphasize its importance for the marginal covariance
+procedure of~\cite{Park2018}.
+Finally, for simplicity, we assume that there exists~$b^\star$
+such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the
+true function~$\beta(t)$ sits in the span of the spline basis
+expansion. 
+With the necessary assumptions stated, we can now state
+Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
 \begin{lemma} \normalfont
 \label{lemma:simpleasym}
-Under Assumption~\ref{assumption:truncation} and~$\Delta$ known, for
-large~$n$ the estimate~$\hat \theta_n$ is approximately normal
+Under Assumption~\ref{assumption:events},
+Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for
+large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
 \[
@@
 
-\section{Theoretical analysis}
-
-\begin{assumption}[Event process assumptions]
-\begin{itemize}
-\item The subsampling rate~$\pi_i (t) \in [L_\pi, U_\pi]$ 
-\item There exists a nonnegative matrix
-\item There exists 
-\end{itemize}
-\end{assumption}
-
-\begin{assumption}[Covariance assumptions]
-\begin{itemize}
-\item 
-\end{itemize}
-\end{assumption}
-
-\begin{assumption}[Functional data assumptions]
-M1 -- M4 in 
-\end{assumption}
-
-\begin{thm}[No reconstruction error]
-\label{thm:truncation}
-Under regularity conditions A-E in Andersen et al. 1993, 
-\[
-\frac{n (\hat \theta - \theta) \Xi (\theta) (\hat \theta - \theta) -
-  K_n+1}{\sqrt{2 (K_n + 1) }} \to N(0,1).
-\]
-
-\end{thm}
-
-\begin{thm}[Reconstruction error OR $\hat \Delta > \Delta$]
-\label{thm:biaseddown}
-
-\end{thm}
-
-Theorem~\ref{thm:biaseddown} shows that if we keep~$K_x = K_b = 35$
-and therefore have reconstruction error, we can expect a downward bias
-in parameter estimates.
-Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias
-in the parameter estimates.  For the former, however, we can simply
-let~$K_x = K_b$ grow as a function of data size; for the latter, we
-can see if~$\hat \beta(s)$ is statistically significant for~$s$
-near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the
-condition holds.
+\subsection{Penalized functional regression models}
+
+Recall theoretical results were proven under the assumption that there
+exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
+To make this assumption plausible, we set~$K_b$ large enough (but less
+than~$K_x$ to ensure identifiability) such that the spline basis
+expansion is sufficiently expressive. However, in practice, such a
+choice of~$K_b$ may lead to overfitting the data. We therefore, in the
+remainder of this paper, follow~\cite{GoldSmith2015} and induce
+smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2
+I)$. Incorporating penalization is simple due to the connection
+between with the score functions for logistic regression.
+
+\subsection{Confidence intervals for~$\beta(t)$}
+
+Due to the connection with generalized mixed effects models, we can
+leverage existing inferential machinery to obtain variance-covariance
+estimates of model parameters. 
+That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional
+matrix obtained by plugging in the REML estimates of variance
+components into the formula for the variance of~$\hat b$, then the
+standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0)
+= \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 )
+  \hat \Sigma_{bb} \phi(t_0)^\top}$.  Then the approximate 95\%
+confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96
+\sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+
+We acknowledge two important limitations of confidence intervals
+obtained via this approach. First, penalization may lead to confidence
+intervals that perform poorly in regions where~$\hat \beta(t)$ is
+oversmoothed. Second, we ignore the variability inherent in the
+longitudinal functional principal component analysis; that is, our
+estimates ignore the variability in estimation of eigenfunctions~$\hat
+\psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
+Joint modeling could be considered as in~\cite{Crainiceanu2010},
+however, this is beyond the scope of this article.
 
@@
 sensor wearing issue.
-To see this, for the suicidal ideation case study, there are XX event
-times across the whole dataset.  Of these, XX event times had complete 
-data for the prior thirty minutes, XX had missingness less than 10\%
-(i.e., sufficient to ), and XX had missingness above 30\%.
-
-The issue here is that~$c_{i,k} (t)$ cannot be estimated if~$X(s,t)$
-is not observed for all~$s \in [0,\Delta]$. 
-Moreover, we want standard erros to reflect the uncertainty
+To see this, for the suicidal ideation case study, there are 2139
+self-identified moments of distress across all 91 participants.  Of
+these, 1289 event times had complete data for the prior thirty minutes,
+1984 had fraction of missing data on a fine grid less than 30\%, and
+1998 had fraction of missing data on a fine grid less than 10\%.
+
+Missing data is a critical issue because~$c_{i,k} (t)$ cannot be
+estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$. 
+Moreover, we want standard errors to reflect the uncertainty
 in these coefficients when missing data is prevalent.
@@
 longitudinal data setting.
-This is possible due to LEmma~\ref{} and our 
+This is possible due to  and our 
 Here, we ignore the issue that~$X(t,s)$ is
@@
 
+\subsection{Coverage properties}
+
 \subsection{Impact of $\Delta$}
+
+Theorem~\ref{thm:biaseddown} shows that if we keep~$K_x = K_b = 35$
+and therefore have reconstruction error, we can expect a downward bias
+in parameter estimates.
+Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias
+in the parameter estimates.  For the former, however, we can simply
+let~$K_x = K_b$ grow as a function of data size; for the latter, we
+can see if~$\hat \beta(s)$ is statistically significant for~$s$
+near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the
+condition holds.
 
@@
 Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
-  \theta)}{h_i (t; \theta)}$. Therefore, under the log-linear
-intensity function, $\frac{d}{d\theta} \log h_i (t; \theta) =
-W_{i,t}$. On the other hand,
+  \theta)}{h_i (t; \theta)}$. Under the log-linear
+intensity function given by~\eqref{eq:approx_hazard},
+$\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}$ and 
 \[
 \frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) = 
-\frac{\exp \left( W_{i,t}^\prime \theta \right)}{\pi_i (t) + \exp
-  \left( W_{i,t}^\prime \theta \right)} W_{i,t}
+\frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp
+  \left( W_{i,t}^\top \theta \right)} W_{i,t}
 \]
@@
 \begin{align*}
-\hat U_n (\theta) 
+\tilde U_n (\theta) 
   &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
-    \sum_{t \in N_i \cup D_i} \frac{\exp \left( W_{i,t}^\prime \theta
-    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\prime \theta \right)}
+    \sum_{t \in N_i \cup D_i} \frac{\exp \left( W_{i,t}^\top \theta
+    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}
     W_{i,t} \right \} \\
   &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
-    + \exp \left( W_{i,t}^\prime \theta \right)} W_{i,t} -
-    \sum_{t \in D_i} \frac{\exp \left( W_{i,t}^\prime \theta
-    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\prime \theta \right)}
+    + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} -
+    \sum_{t \in D_i} \frac{\exp \left( W_{i,t}^\top \theta
+    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}
     W_{i,t} \right \} \\
@@
 following assumptions related to the subsampling rate~$\pi_i (t)$.
-For simplicty we assume there exists~$\tau < \infty$ and all censoring
-times satisfy~$\tau_i \leq \tau$. Define $R_i (t)$ to be the at risk
-indicator for participant~$i$.  That is, 
-\[
-R_i (t) = \left \{ \begin{array}{c c} 1 & t \in (0,\tau_i) \\ 0
-                                        & t > \tau_i\end{array} \right .
-\]
 We define the joint counting process
@@
 \label{eq:andersen1}
-n^{-1} \hat U_n (\theta_0 ) \overset{P}{\to} 0,
+n^{-1} \tilde U_n (\theta_0 ) \overset{P}{\to} 0,
 \end{equation}
 and 
-\[
-n^{-1} \frac{d}{d \theta^{\top}} \hat U_n (\theta) \big |_{\theta =
+\begin{equation}
+\label{eq:andersen2}
+n^{-1} \frac{d}{d \theta^{\top}} \tilde U_n (\theta) \big |_{\theta =
   \theta_0} = \Xi (\theta),
-\]
+\end{equation}
 and
-\[
-\lim_{n \to \infty} P \left(  and all \theta \in \Theta_0 \right) = 1
-\]
+\begin{equation}
+\label{eq:andersen3}
+\lim_{n \to \infty} P \left( \left | n^{-1} \frac{d^2}{d \theta_j d
+      \theta_k} \tilde U_n (\theta) \right | < M \text{ for all }
+  j,k \text{ and all } \theta \in \Theta_0 \right) = 1
+\end{equation}
 To prove~\eqref{eq:andersen1}, we decompose~$\hat U_n (\theta_0)$ into
@@
 \[
-n^{-1} \hat U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
-  \tilde U_n (\theta) - U_n (\theta) \right) + n^{-1} \left(\hat U_n
-  (\theta) - \tilde U_n (\theta) \right)
-\]
-where~$\tilde U_n (\theta_0)$ are the logistic score equations with
-$\int_0^\infty X_i(t-s) \beta(s) ds$ replaced by the
-approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$. 
-\cite{Andersen1993}[Theorem VI.1.1] show that~$n^{-1} U_n (\theta_0)
+n^{-1} \tilde U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
+  \hat U_n (\theta_0) - U_n (\theta) \right) + n^{-1} \left(\tilde U_n
+  (\theta_0) - \hat U_n (\theta_0) \right)
+\]
+where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations
+with $\int_0^\infty X_i(t-s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$
+with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$. 
+\cite{Andersen1993} (Theorem VI.1.1) show that~$n^{-1} U_n (\theta_0)
 \to 0$. The second term
@@
 \end{align*}
-Lenglart's inequality\walt{Need to make precise} implies
-the second term converges in probability to zero.  This statement
-requires the following two conditions:
-\begin{enumerate}
-\item The subsampling rate is both lower and upper bounded for all
-  ``at risk times''; that is,~$0 < L < \pi_i (t) < U < \infty$ for
-  all~$i=1,2,\ldots$ and~$t \in [0,\tau]$. \walt{Could we just bound
-    $h_i (t; \theta) + \pi_i (t)$?}
-\item There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
-\[
-n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
-\theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
-      \theta)}{h_i (t; \theta)} \right] \times  \left [
-    \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
-  dt \overset{P}{\to} \Xi (\theta).
-\] 
-\end{enumerate}
+where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}.
+Lenglart's inequality \walt{check} implies the second term converges
+in probability to zero under Assumptions~\ref{A1} and~\ref{A2}.
+
 The third term satisfies
 \begin{align*}
-n^{-1} \left \| \hat U_n (\theta_0) - \tilde U_n (\theta_0) \right \|
+  &n^{-1} \left \| \hat U_n (\theta_0) - \tilde U_n (\theta_0) \right \| \\
+  \leq & M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \bigg \| g
+         \left( W_{i,t}^\top \theta  + \sum_{k=K_x+1}^\infty \hat
+         c_{i,k} (t) \int_{t-\Delta}^t \hat \psi_k (s) \phi (s)^\top b
+         + \log (\pi_i (t) ) \right) - g \left( \tilde W_{i,t}^\top
+         \theta + \log (\pi_i (t)) \right) \bigg \| \\
 &\leq M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
-    \left \| g \left( W_{i,t}^\top \theta + \log (\pi_i (t)
-  ) \right) - g \left( \tilde W_{i,t}^\top \theta + \log (\pi_i (t) 
-  ) \right) \right \| \\
-&= M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
-    \left \| g^\prime \left( W_{i,t}^\top \theta + \log (\pi_i (t) )
-  \right) \right\| \times \left \| W_{i,t}^\top \theta_0 - \tilde W_{i,t}^\top
-  \tilde \theta_0  \right \| \\
-&= \frac{M}{4} \| N_i \cup D_i \| \left ( \sum_{j=p_n+1}^\infty
-  \lambda_{j}^2 \sum_{j=q_n+1}^\infty b_j^2 \right)^{1/2}
+    \sup_{x \in \mathbb{R}} \left \| g^\prime \left( x \right)
+  \right\| \times \left \| \sum_{k=K_x+1}^\infty \hat c_{i,k} (t)
+  \sum_{l=1}^{K_b} \int_{t-\Delta}^t \hat \psi_k (s) \phi_l (s) b_l
+  \right \| \\ 
+&= \frac{M}{4} n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+  \times \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t 
+  \left( X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
+  \phi_l (s) b_l\right \| \\
+&\to \frac{M}{4} \int_0^\tau \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t 
+  \left( X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
+  \phi_l (s) b_l\right \| g(T) dT
 \end{align*}
 where~$M = \sup \| \tilde W_{i,t} \|$,~$g(x) = 1/(1+\exp(-x))$ is the
-expit function, and the final line is by Cauchy-Schwarz inequality.
-
+expit function. 
+The second inequality is due to the Taylor remainder theorem.
+The third is due to~$g^\prime (x) = e^x/(1+e^x)^2 \leq e^0/(1+e^0)^2 =
+1/4$, reordering the summation, and rewriting the error term in terms
+of the difference between~$X(s,t)$ and the approximation with
+truncation level~$K_x$. Letting~$n \to \infty$ allows us to re-write
+the outside sums in terms of the integrated difference where the
+integral is with respect to the sampling distribution~$g(T)$. Finally,
+under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
+\[
+X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \overset{p}{\to} 0
+\]
+as~$K_x \to \infty$, which implies the third term goes to zero as the
+truncation level goes to infinity.
 \end{proof}

### 2019-07-03 methods/mean_cov_estimation.R
Made plots for presentation

@@ -18,7 +18,7 @@ event_complete = readRDS("./event_complete_2019-07-01.RDS")
 sequence = seq(-30,0,by =1/60); sensor_obs = 4:1804
 plot(sequence, event_complete[3,sensor_obs], type = "l")
 
-library(refund)
+library(refund); library(lubridate)
 full_obs = apply(X = event_complete, MARGIN = 1, FUN = function(x){!any(is.na(x))})
 greater_than_ten_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.1})
 greater_than_thirty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.3})
@@ -56,8 +56,15 @@ eig_values[eig_values<0] = 0
 K = 35
 phi_vectors = eig_vectors[,1:K]
 coef = residual%*%phi_vectors
-plot(sequence, phi_vectors%*%coef[obs,]/inflation + est$Yhat[obs,], type = "l")
+## To show the impact of K
+obs = 20
+# par(mar = c(4,3,2,1)+ 0.1)
+# png(filename = "~/Dropbox/Presentations/06-21-19-SAA/realdataplot_K10.png",
+#     width = 480, height = 480, units = "px", pointsize = 12)
+plot(sequence, phi_vectors%*%coef[obs,]/inflation + est$Yhat[obs,], type = "l", axes = FALSE, xlab = "Time until button press", ylab = "")
+axis(side = 1); axis(side = 2, labels = FALSE)
 lines(sequence, Y[obs,], col = "red", lwd = 2)
+# dev.off()
 
 ## SAVE THE MEAN, COEFFICIENT MATRIX, AND FIRST K EIGEN VECTORS
 saveRDS(object = est$Yhat, file = paste("event_means_",today(),".RDS", sep=""))

### 2019-07-08 methods/loglik/likelihood_computation.R
Likelihood computation now is simple logistic

@@ -18,29 +18,19 @@
 ## denominator =  c^prime %*% (mle_output$var_subsampl + mle_output$var_event ) %*% c 
 ## rel eff = numerator / denominator
 
-## Source the underlying functions
-source("./likelihood_functions.R")
-
-library(doParallel)
-cl = Sys.getenv("SLURM_NTASKS_PER_NODE")
-cl
-
-registerDoParallel(cores = (cl))
-
-# Shows the number of Parallel Workers to be used
-getDoParWorkers()
-print("Parallel setup complete")
 ## GENERATE SPLINES
 library(splines)
 sequence <- seq(-30,0, by = 1/60)
-# knots <- seq(-36, 6, 2)  # 10 => 10-4 = 6 Basis splines
-# x <- seq(-30, 0, by = 1/60)
-# bb <- splineDesign(knots, x = x, outer.ok = FALSE)
-knots <- seq(0, 30, 1)  # 10 => 10-4 = 6 Basis splines
-x <- seq(0, 30, by = 1/60)
-bb <- splineDesign(knots, x = x, outer.ok = TRUE)
+# knots <- seq(-40, 10, 1)  # 10 => 10-4 = 6 Basis splines
+knots <- c(-32,seq(-30,0,1), 2)  # 10 => 10-4 = 6 Basis splines
+x <- seq(-30, 0, by = 1/60)
+bb <- splineDesign(knots, x = x, outer.ok = FALSE, ord = 2)
+# knots <- seq(0, 30, 1)  # 10 => 10-4 = 6 Basis splines
+# x <- seq(0, 30, by = 1/60)
+# bb <- splineDesign(knots, x = x, outer.ok = TRUE)
 print("Generated Splines")
 
+library(lubridate)
 setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
 # setwd("/n/murphy_lab/users/wdempsey/data-for-fda/data/")
 ## PULL IN EVENT RDS FILES
@@ -79,44 +69,28 @@ print("RDS Files Readin correctly")
 
 Y = c(rep(1,nrow(event_model.matrix)), rep(0, nrow(nonevent_model.matrix)))
 
-model.matrix = rbind(event_model.matrix, nonevent_model.matrix)
+model.matrix = rbind(event_model.matrix[,-1], nonevent_model.matrix[,-1])
+hour.info = c(event_model.matrix[,1], nonevent_model.matrix[,1])
 
[…]

### 2019-07-08 write-up/fda-recurrent.tex
Adding multiple imputation section

@@
 
-A core concern in mHealth is that sensor data can be missing due to
-sensor wearing issue.
-To see this, for the suicidal ideation case study, there are 2139
-self-identified moments of distress across all 91 participants.  Of
-these, 1289 event times had complete data for the prior thirty minutes,
-1984 had fraction of missing data on a fine grid less than 30\%, and
-1998 had fraction of missing data on a fine grid less than 10\%.
+A core concern in mHealth is that sensor data can often be missing for
+intervals of time due to sensor wearing issue (e.g., sensor band
+becomes loose, turns off accidentally, data packets lost, or battery
+dies). In the suicidal ideation case study, for example, there are
+2139 self-identified moments of distress across all 91 participants.
+Of these, 1289 event times had complete data for the prior thirty
+minutes, 1984 had fraction of missing data on a fine grid less than
+30\%, and 1998 had fraction of missing data on a fine grid less than
+10\%.
 
@@
 in these coefficients when missing data is prevalent.
-\cite{Goldsmith2015} suggest using best linear unbiased predictors
+\cite{Goldsmith2011} suggest using best linear unbiased predictors
 (BLUP) or posterior modes in the mixed effects model to
@@
 \cite{Crainiceanu2010, Goldsmith2010} take a full Bayesian analysis.
-\cite{Yao2005} introduced PACE as a frequentist analysis that deals
-with this issue.
-However,~\cite{Reimherr2018} show that for sparse, irregular
-longitudinal, the imputation model should not ignore the outcome~$Y_i
-(t)$.
-
-Here we present an extension of~\cite{Reimherr2018} to our setting,
-leverage Lemma~\ref{lemma:logistic} and the marginal covariance
-estimation procedure to build a multiple imputation approach.
-\[
-\E [ X_i (s,t) \given Y_i(t) = y, \bfx ] = \mu_y (s,t) + {\bf a}_{i,t}^\top
-{\bf B}_i (\bfx_i (t) - \mathbold{\mu}_i )
-\]
-where~$\mu_y (t) = \E [ X_i (t) \given Y_i (t) = y]$, and
-\[
-{\bf a}_i^\top = ( \Sigma(s, )
-\]
-This generalizes the procedure of~\cite{Reimerr2018} to the
-longitudinal data setting.
-This is possible due to  and our 
-Here, we ignore the issue that~$X(t,s)$ is
-
-The estimate is then given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat
-\theta^{(m)}$; the covariance matrix can be estimated by
+\cite{Yao2005} introduced PACE as an alternative frequentist method.
+However,~\cite{Petrovich2018} shows that for sparse, irregular
+longitudinal, the imputation model should not ignore the outcome
+variable~$Y_i (t)$.
+
+Here we present an extension of~\cite{Petrovich2018} to our setting by
+leveraging Lemma~\ref{lemma:logistic} and the marginal covariance
+estimation procedure to construct a multiple imputation procedure.
+Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at
+times~$\{ s_{i,r} \}_{r=1}^{k_{it}} \}$ in~$[0,\Delta]$. Then
+\begin{align}
+\label{eq:impute}
+\E [ X_i (s,t) \given Y_i(t) = y, \bfx_i (t) ] 
+  &= \mu_y (s,t) + {\bf a}_{i,t}^\top (s) {\bf B}_{i,t} (\bfx_i (t) -
+    \mathbold{\mu}_i (t) ) \\
+\text{Cov} \left( X_i (s,t), X_i (s^\prime, t) \given Y_i (t) = y,
+  \bfx_i (t) \right) 
+  &= \Sigma_{t} (s, s^\prime) - 
+    {\bf a}_{i,t} (s)^\top {\bf B}_{i,t} {\bf a}_{i,t} (s^\prime)
+\end{align}
+where we have
+\[
+{\bf a}_{i,t} (s)^\top = \left( \begin{array}{c} \Sigma_t (s_{i,1}, s) \\ \vdots \\
+                       \Sigma_t (s_{i,k_{it}}, s) \end{array} \right); \quad
+{\bf B}_{i,t}^{-1} = \left( 
+  \begin{array}{c c c} 
+    \Sigma_t(s_{i,1}, s_{i,1}) & \Sigma_t(s_{i,1}, s_{i,2}) & \cdots \\
+    \Sigma_t(s_{i,1}, s_{i,2}) & \ddots & \vdots \\
+    \vdots & \cdots & \Sigma_t (s_{i,k_{it}}, s_{i,k_{it}}) 
+  \end{array} \right),
+\]
+$\mathbold{\mu}_i (t) = \E [ \bfx_i (t) \given Y_i (t) = y] = \{
+\mu_y (s_{j}, t) \}_{j=1}^r$ and~$\mu_y (s,t)$ is the mean of~$X(t,s)$
+from group~$y$, and~$\Sigma_t ( s^\prime, s)$ is the covariance
+between~$X(s^\prime,t)$ and $X(s,t)$ for $s^\prime, s \in [0, \Delta]$ 
+and~$t \in \mathbb{R}_+$.
+These covariances can be estimated using the methods considered by
+\cite{ChenMuller2012}.  Our imputation procedure instead uses the
+marginal covariance~$\Xi$ as a proxy for the true covariance
+function~$\Sigma_t(\cdot, \cdot)$.  We found this to work well in
+simulations as~$\Sigma_t$ was relatively constant across~$t$; however,
+this imputation strategy may be inappropriate when the covariance
+structure changes significantly with time~$t$.
+
+The imputation strategy generalizes the procedure introduced
+by~\cite{Petrovich2018} to the longitudinal functional data
+setting.  This is only possible due to the connection established with
+logistic regression by Lemma~\ref{lemma:logistic}.
+We consider a multiple imputation estimate with~$M$ imputations;
+the likelihood estimate is given by~$\bar \theta = M^{-1}
+\sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate
+from the~$m$th imputated dataset; the covariance matrix can be
+estimated by 
 \[
 M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
-\sum_{m=1}^M \hat \theta - \bar \theta) (\hat \theta - \bar \theta)
+\sum_{m=1}^M (\hat \theta^{(m)} - \bar \theta) (\hat \theta^{(m)} -
+\bar \theta)
 \]
@@
 
-\subsection{Re-using estimates under additional sampling}
-
-If~$\pi$ is constant then we can continue to use the same pooled
-covariance.
-
-\section{Simulations}
-
-\subsection{Unbiased estimation}
+\section{Simulation study}
+
+We next assess the proposed methodology via a simulation study.
+Simulation studies of joint high frequency sensor and recurrent
+event data can be quite difficult to construct.  Here, we use the
+imputation strategy introduced in Section~\ref{section:missingdata} to
+circumvent some of the issues.  
+In this section, we assume each individual is observed over five
+12-hour days.
+
+First, we generate a sequence of event times via a marginal hazard
+rate that ignores sensor data.
+
+We generate 5 days of 12 hours with range of marginal rates (1 per
+12 hours, 1 per 6 hours, 1 per hour).  We then sample non-events at
+rates or (1 per 12/6/1/0.5 hours).
+We assume that~$Y_t = 0$ implies the curve is flat with mean 0.
+We assume that~$Y_t = 1$ implies three curves (A) the curve jumps
+around 10 minutes prior to the event and remains elevated, (B) the
+curve goes down then up and down smoothly, (C) 
+These curves are given in Figure~\ref{} for three choices of
+parameters that indicate varying degrees of departure from the~$Y_t =
+0$ case.
+
+Using~\eqref{eq:}
+We define~$X(t,s)$ at the grid of observerations per secondw ith
+\[
+  \Sigma_t (s, s^\prime) = 
+  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}} 
+  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
+  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
+\]
+where~$K_v$ is the modified Bessel function of the second kind, and
+we choose~$\nu = 5/2$ and $\sigma^2 = 1$. 
+We choose three versions of~$\rho_t$: (A) constant across time
+at $0.5$, (B) constant within each day but linearly decreasing
+across days from~$0.7$ to~$0.3$, and (C) constant within each day but
+linearly increasing across days from~$0.3$ to~$0.7$.
+
+These curves are evaluate
+
+We estimate with~$K_b = K_x = 35$ under both the logistic, penalized
+logistic, and multi-level penalized logistic models.
+
+\subsection{Mean-integrated squared error (MISE)}
+
+Since we are primarily interested in accuracy of the final
+estimates~$\hat \beta(s)$, we report the mean integrated  squared
+error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{1000}
+\sum_{j=1}^{1000} \int_{-\infty}^t ( \hat \beta_j (t) - \beta(s))^2
+ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.  We define the MISE in
+this manner to account for settings where~$\Delta$ is unknown.
+For~$\Delta$ known, the MISE simplifies to integrals from~$t-\Delta$
+to~$t$.
+Table~\ref{} show the MISE of~$\hat \beta (s)$ decomposed into the
+variance and squared biase
 
 \subsection{Coverage properties}
+
+In each (A)-(C) we parametrize the curves and check the type I error
+for various sizes of the
 
@@
 
-\section{Case Study: }
+\section{A worked example: Adolescent psychiatric inpatient mHealth study}
+
+During a XX month period in 2019, 100 psychiatric inpatients admitted
+for suicidal risk to Franciscan Children's Hospital
+
+hospitals with histologically verified liver cirrhosis were recu
+those presenting to a psychiatric hospital with suicide ideation
+and/or a recent suicide attempt) during a high risk time period (i.e.,
+post-hospitalization)
+
+Currently, only 51 patients for whom the initial biopsy could be
+reevaluated using more restrictive criteria were available for
+analysis.
+The goal of the study was to ask whether sensor information can used
+to identify distress early prior in the process of escalation by
+monitoring physiological correlates of distress -- in particular,
+electrodermal activity (EDA), a measure of skin conductance, and
+``gACC'', a coordinate-free feature built from accelerometer data 
+that measures physical movement. 
+The scientific hypothesis to be tested is whether a spike in
+EDA and/or gACC
+
+The goal of the sutyd
+
+EDA and ACC sensor data along with buttonpress timestamps were
+graciously provided by Evan Kleiman and his study team {\tt
+https://kleimanlab.org}.
+At the end of the study period, the mortality rate was 292/488, or
+approximately 60\%.
+
+The focus here is on two EDA and the  prothrombin index, a composite
+blood coagulation index
+related to liver function, measured initially at three-month intervals
+and subsequently at roughly twelve-month intervals.
+The individual prothrombin trajectories are highly variable,
+both in forward and in reverse time,
+which tends to obscure patterns and trends.
+In Figure~\ref{fig:mean_traj}a the mean trajectory is plotted against time from recruitment
+for two patient groups placebo/prednisone and censored/not censored.
+Naturally, only those patients who are still alive are included in the average
+for that time.
+Figure~\ref{fig:mean_traj}b shows the same plots in reverse alignment.
+While there are certain similarities in the two plots,
+the differences in temporal trends are rather striking.
+In particular, prothrombin levels in the six months prior to censoring are fairly stable,
+which is in marked contrast with levels in the six months prior to failure,
+as seen in the lower pair of curves.
+
+Inspection of the graphs for uncensored patients in the right panel of Figure~\ref{fig:mean_traj} suggests
+beginning with the simplest revival model in which the sequences
+for distinct patients are independent Gaussian with moments
+
+\subsection{Effect of electrodermal activity on prognosis}
+Over a period of 5~years and one month following recruitment,
+patient~$u$ had eight appointments with prothrombin values as follows:
 

### 2019-07-26 methods/generate_data/gendata_functions.R
Fixed code to now fully run with ACC and rm files when not necessary due to memory limits

@@ -23,7 +23,7 @@ approximate_eda_apply <- function(iter, sequence, id_bp, eda) {
   diff = eda$ts - current_ts
   minute_diff = diff/1000/60 
   keep_obs = (-30 < minute_diff) & (minute_diff <= 0)
-  output = unlist(lapply(sequence, approximate_eda(minute_diff[keep_obs],eda$EDA_HighLowPass[keep_obs])))
+  output = unlist(lapply(sequence, approximate_eda(minute_diff[keep_obs],eda$EDA_FeatureScaledFiltered_Day[keep_obs])))
   return(output)
 }
 
@@ -45,7 +45,7 @@ nonevent_approximate_eda_apply <- function(iter, sequence, sampled_times, eda) {
   diff = eda$ts/1000 - current_ts
   minute_diff = diff/60 
   keep_obs = (-30 < minute_diff) & (minute_diff <= 0)
-  output = unlist(lapply(sequence, approximate_eda(minute_diff[keep_obs],eda$EDA_HighLowPass[keep_obs])))
+  output = unlist(lapply(sequence, approximate_eda(minute_diff[keep_obs],eda$EDA_FeatureScaledFiltered_Day[keep_obs])))
   return(output)
 }
 

### 2019-07-26 methods/generate_data/generate_dataset.R

@@ -40,18 +40,24 @@ for (id in 1001:1091) {
   eda_file_name = paste("./R21_Study - EDA/",id,"_EDA.rds", sep = "")
   acc_file_name = paste("./R21_Study - ACC/",id,"_acc.rds", sep = "")
   if(!(file.exists(eda_file_name) & file.exists(acc_file_name))) {
-    print(paste("No EDA file with id", id))
+    print(paste("No EDA and/or ACC file with id", id))
   } else{
     eda = readRDS(eda_file_name)
-    acc = readRDS(acc_file_name); max_acc = 127 * 0.5773503 * 3
-    acc$g = sqrt((acc$acc_x/max_acc)^2 + (acc$acc_y/max_acc)^2 + (acc$acc_z/max_acc)^2)
     print("Made it to sampling nonevent times")
     datetime_ts = as_datetime(eda$ts/1000)
     sampled_times = generate_noneventtimes(datetime_ts, sampling_rate, max.iters = 20000)
+    
+    print("And now onto EDA calculations")
     eda_output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_eda_apply(iter, sequence, id_bp, eda)
-    acc_output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_acc_apply(iter, sequence, id_bp, acc)
     eda_output_nonevent = foreach(iter=1:length(sampled_times), .combine=rbind) %dopar% nonevent_approximate_eda_apply(iter, sequence, sampled_times, eda)
+    rm("eda")
+    
+    print("And now onto ACC calculations")
+    acc = readRDS(acc_file_name); max_acc = 127 * 0.5773503 * 3
+    acc$g = sqrt((acc$acc_x/max_acc)^2 + (acc$acc_y/max_acc)^2 + (acc$acc_z/max_acc)^2)
+    acc_output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_acc_apply(iter, sequence, id_bp, acc)
     acc_output_nonevent = foreach(iter=1:length(sampled_times), .combine=rbind) %dopar% nonevent_approximate_acc_apply(iter, sequence, sampled_times, acc)
+    rm("acc")
     
     ## Define time as since minimum time in EDA or ID_BP
     base_time = min(eda$ts, id_bp$ts_ms, acc$ts)
@@ -73,9 +79,11 @@ for (id in 1001:1091) {
[…]

### 2019-07-26 methods/generate_data/generate_dataset.R
Don't rm() cause then can't compute base_time

@@ -50,14 +50,12 @@ for (id in 1001:1091) {
     print("And now onto EDA calculations")
     eda_output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_eda_apply(iter, sequence, id_bp, eda)
     eda_output_nonevent = foreach(iter=1:length(sampled_times), .combine=rbind) %dopar% nonevent_approximate_eda_apply(iter, sequence, sampled_times, eda)
-    rm("eda")
     
     print("And now onto ACC calculations")
     acc = readRDS(acc_file_name); max_acc = 127 * 0.5773503 * 3
     acc$g = sqrt((acc$acc_x/max_acc)^2 + (acc$acc_y/max_acc)^2 + (acc$acc_z/max_acc)^2)
     acc_output_event = foreach(iter=1:nrow(id_bp), .combine=rbind) %dopar% approximate_acc_apply(iter, sequence, id_bp, acc)
     acc_output_nonevent = foreach(iter=1:length(sampled_times), .combine=rbind) %dopar% nonevent_approximate_acc_apply(iter, sequence, sampled_times, acc)
-    rm("acc")
     
     ## Define time as since minimum time in EDA or ID_BP
     base_time = min(eda$ts, id_bp$ts_ms, acc$ts)

### 2019-07-27 methods/mean_cov_estimation.R
Mean Cov Now cycles over quad/lin and acc/eda, could convert into 4 or event 8 different runs on the system. Write that on plnae but run for now since it'll work

@@ -14,113 +14,195 @@
 ## And seeing how close hatX(t,s) is to X(t,s).
 
 setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
-event_complete = readRDS("./event_complete_2019-07-01.RDS")
-sequence = seq(-30,0,by =1/60); sensor_obs = 4:1804
-plot(sequence, event_complete[3,sensor_obs], type = "l")
-
-library(refund); library(lubridate)
-full_obs = apply(X = event_complete, MARGIN = 1, FUN = function(x){!any(is.na(x))})
-greater_than_ten_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.1})
-greater_than_thirty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.3})
-greater_than_fifty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.5})
-Y = event_complete[full_obs,sensor_obs]
-x = event_complete[full_obs,2]
-z = sequence
-est <- fbps(Y,list(x=x,z=z))
-obs = 100
-plot(sequence,est$Yhat[obs,]- mean(est$Yhat[obs,]))
-lines(sequence, Y[obs,]-mean(Y[obs,]))
-
-## COMPUTE ESTIMATED MARGINAL COVARIANCE 
-
-inflation = 100
-residual = inflation*(Y-est$Yhat)
-## TAKE PAIRS OF ROWS and CALCULATE THE MSE
-Sigma = matrix(nrow = ncol(residual), ncol = ncol(residual))
-for (i in 1:ncol(residual)) {
-  for (j in 1:ncol(residual)) {
-    Sigma[i,j] = sum(residual[,i]*residual[,j])/nrow(residual)
+set_of_types = c("eda", "acc")
+for (type in set_of_types){
+  event_complete = readRDS(paste(type, "_event_complete_HLP_2019-07-27.RDS", sep = ""))
+  sequence = seq(-30,0,by =1/60); sensor_obs = 4:1804
+  # plot(sequence, event_complete[3,sensor_obs], type = "l")
+  
+  library(refund); library(lubridate)
[…]

### 2019-07-27 write-up/fda-recurrent.tex
Latest version of write-up, fixed some typos

@@
 \hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in D_i} \frac{
-  h^{(1)}_i (u; \theta) }{ \pi_i (s) }
+  h^{(1)}_i (u; \theta) }{ \pi_i (u) }
 \]
@@
 \hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in (\bfT_i \cup D_i)}
-\frac{ h^{(1)}_i ( t; \theta ) }{ \pi_i (s) + h_i ( t ; \theta) }
+\frac{ h^{(1)}_i ( u; \theta ) }{ \pi_i (u) + h_i ( u ; \theta) }
 \end{equation}
@@
 \hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
-  \frac{h^{(1)} (u; \theta)}{ h ( t; \theta)}  - \sum_{u \in D_i} w_i(u;
-  \theta) \frac{h_i^{(1)} (t; \theta)}{ \pi_i (t) } \right].
+  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in D_i} w_i(u;
+  \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
 \end{equation}
@@
 
-Our theoretical analysis requires the several assumptions, some
-regarding the subsampling procedure, some for the event process, and
-some for the functional data. We state these assumptions and then our
-main theorems.  
+Our theoretical analysis requires the several
+assumptions, some regarding the subsampling procedure, some for the
+event process, and some for the functional data. We state these
+assumptions and then our main theorems.  
 We start by assuming there exists a~$\tau < \infty$ such that all
@@
   X(s^\prime, T^\prime) ] < \infty$ for
-  each $s,s^\prime [0,\Delta]$ and~$0 < T, T^\prime < \tau$.
+  each $s,s^\prime \in [0,\Delta]$ and~$0 < T, T^\prime < \tau$.
 \item\label{A4} $\E[\|X(\cdot, T)\|^4] < \infty$ for each~$0< T
@@
 \label{section:missingdata}
+\walt{Question: How to perform multiple imputation for~$X(\cdot,t)$
+  and $X(\cdot, t^\prime)$ for~$|t - t^\prime| <\Delta$? That is, how
+  do you perform multiple imputation with this overlapping structure?}
 
@@
 substantial variability in these estimates.  To deal with this,
-\cite{Crainiceanu2010, Goldsmith2010} take a full Bayesian analysis.
+\cite{Crainiceanu2010} take a full Bayesian analysis.
 \cite{Yao2005} introduced PACE as an alternative frequentist method.
@@
 These covariances can be estimated using the methods considered by
-\cite{ChenMuller2012}.  Our imputation procedure instead uses the
-marginal covariance~$\Xi$ as a proxy for the true covariance
-function~$\Sigma_t(\cdot, \cdot)$.  We found this to work well in
-simulations as~$\Sigma_t$ was relatively constant across~$t$; however,
-this imputation strategy may be inappropriate when the covariance
-structure changes significantly with time~$t$.
+\cite{ChenMuller2012}.  
+% Our imputation procedure instead uses the
+% marginal covariance~$\Xi$ as a proxy for the true covariance
+% function~$\Sigma_t(\cdot, \cdot)$.  
+% We found this to work well in simulations as~$\Sigma_t$ was relatively
+% constant across~$t$; however, this imputation strategy may be
+% inappropriate when the covariance structure changes significantly with
+% time~$t$.
 
@@
 
+\newpage
+
 \subsection{Multilevel models}
@@
 imputation strategy introduced in Section~\ref{section:missingdata} to
-circumvent some of the issues.  
-In this section, we assume each individual is observed over five
-12-hour days.
-
-First, we generate a sequence of event times via a marginal hazard
-rate that ignores sensor data.
-
-We generate 5 days of 12 hours with range of marginal rates (1 per
-12 hours, 1 per 6 hours, 1 per hour).  We then sample non-events at
-rates or (1 per 12/6/1/0.5 hours).
+circumvent some of these issues.
+
+Here, we assume each individual is observed over five 12-hour days.  
+We then generate a sequence of event times via a marginal hazard
+rate that ignores sensor data.  We assume the marginal rate is
+constant and set to 1 per 12/6/1 hour(s). We then sample non-events at
+rates of 1 per 12/6/1/0.5 hours.
+This sampling strategy yields (non-)event times~$t$ along with
+labels~$Y_t \in \{0,1\}$. 
+We then use the imputation strategy given by
+equation~\eqref{eq:impute} to generate the curves~$\{ X(s,t) \}$.
+For simplicity, we do not deal with 
 We assume that~$Y_t = 0$ implies the curve is flat with mean 0.
@@
 0$ case.
+The associated~$\beta$ depends on the covariance~$\Sigma_t (s,t)$.
+In particular, $\beta = \Sigma_t^{-1} (\mu_{1,t} - \mu_{0,t})$ is
+guaranteed by Lemma~\ref{} and Theorem 1 of \cite{Petrovich2018}.
+For~$\beta(s)$ non time-varying implies the covariance term
 
@@
 
-These curves are evaluate
+to estimate the impact of strong X dependence for close X's then we
+will GENERATE SEQUENTIALLY TO FILL IN TO ACCOUNT FOR OVERLAP!
+HOW SHOULD WE HANDLE COVARIANCE BETWEEN X's outside the range?
+WE WILL APPROXIMATE?
 

### 2019-10-25 write-up/fda-recurrent.tex
Fixing up paper and the repo

@@
 
-\title[Recurrent event analysis with functional covariates via random
-subsampling]{Recurrent event analysis in the presence of real-time
-  high frequency data via random subsampling}
+\title[Recurrent event analysis with functional covariates via random subsampling]{Recurrent event analysis in the presence of real-time high frequency data via random subsampling}
 \author{Walter Dempsey}
-\address {Department of Statistics, 
-  Harvard University, One Oxford Street
-   Cambridge, MA  02138, USA}
- \email{wdempsey@fas.harvard.edu}
+\address {Department of Biostatistics, University of Michigan, 1415 Washington Heights, Ann Arbor, MI 48109, USA}
+ \email{wdem@umich.edu}
 
@@
 \begin{abstract}
-Digital monitoring studies collect real-time high frequency data via
-mobile sensors in the subjects' natural environment.  
-This data can be used to model the impact of changes in physiology on
-recurrent event outcomes such as smoking, drug use, alcohol use, or
-self-identified moments of suicide ideation. 
-Likelihood calculations for the recurrent event analysis, however,
-become computationally prohibitive in this setting.  
-% Unlike classic joint models, event risk is a function of
-% recent sensor data, not simply current sensor values.  
-Motivated by this, a random subsampling framework is proposed for
-computationally efficient, approximate likelihood-based estimation.
-A subsampling-unbiased estimator for the derivative of the cumulative
-hazard enters into an approximation of log-likelihood. The estimator
-has two sources of variation: the first due to the recurrent event
-model and the second due to subsampling. The latter can be reduced by
-increasing the sampling rate; however, this leads to increased
-computational costs.  The approximate score equations are equivalent
-to logistic regression score equations, allowing for standard,
+Digital monitoring studies collect real-time high frequency data via mobile sensors in the subjects' natural environment.
+This data can be used to model the impact of changes in physiology on recurrent event outcomes such as smoking, drug use, alcohol use, or self-identified moments of suicide ideation.
+Likelihood calculations for the recurrent event analysis, however, become computationally prohibitive in this setting.
+Motivated by this, a random subsampling framework is proposed for computationally efficient, approximate likelihood-based estimation.
+A subsampling-unbiased estimator for the derivative of the cumulative hazard enters into an approximation of log-likelihood.
+The estimator has two sources of variation: the first due to the recurrent event model and the second due to subsampling.
+The latter can be reduced by increasing the sampling rate; however, this leads to increased computational costs.
+The approximate score equations are equivalent to logistic regression score equations, allowing for standard,
 ``off-the-shelf'' software to be used in fitting these models.
-Simulations demonstrate the method and efficiency-computation
-tradeoff. We then illustrate the approach using data from a digital
-monitoring study of suicide ideation.
+Simulations demonstrate the method and efficiency-computation trade-off.
+We then illustrate the approach using data from a digital monitoring study of suicidal ideation.
 \end{abstract}
 
-\keywords{recurrent events; probabilistic subsampling; estimating equations;
-high frequency time series; logistic regression}
+\keywords{recurrent events; probabilistic subsampling; estimating equations; high frequency time series; logistic regression}
 
@@
 
-Advancement in mobile technology has led to the rapid integration of
-mobile and wearable sensors into behavioral health~\citep{Freeetal2013}.
-Take HeartSteps, for example, a mobile health (mHealth) study designed
-to increase physical activity in sedentary adults
-\citep{KlasnjaHS2019}. Here, a Jawbone sensor is used to monitor step
-count every minute of the participant's study day.
-Of interest in many mHealth studies is the relation of such
-real-time high frequency sensor data to an adverse, recurrent event
-process. In a smoking cessation mHealth study~\citep{Sense2Stop}, for
-example, the relation between a time-varying sensor-based measure of
-physiological stress and smoking lapse is of scientific interest.
-In a suicidal ideation mHealth study~\citep{Kleiman2018}, the relation of
-electrodermal activity (EDA) and accelerometer with self-identified
-moments of suicidal ideation is of scientific interest.
-
-The goal of this paper is two-fold: (1) to discuss the appropriate
-choice of statistical model for joint high-frequency sensor data and
-the recurrent event data, and (2) to construct a robust,
-easy-to-implement method for parameter estimation and inference. 
-For (1), we discuss an important issue regarding measurement-error
-models when paired with recurrent event outcomes.
-For (2), we introduce a random subsampling procedure that has
-several benefits.  First, the resulting inference is unbiased;
-however, there is a computation-efficiency trade-off. In particular, a
-higher sampling rate can decrease estimator variance at the cost of
-increased computation.  We show via simulations that the benefits of
-incredibly high sampling rates is often negligible, as the
-contribution to the variation is small in relation to to variation in
-the underlying stochastic processes. Second, derived estimating
-equations are optimal, implying loss of statistical efficiency is only
-due to subsampling procedure and not derived methodology.  Finally,
-implementation can leverage existing, standard software for functional
-data analysis and logistic regression, leading to fast adoption by
-domain scientists.
+Advancement in mobile technology has led to the rapid integration of mobile and wearable sensors into behavioral health~\citep{Freeetal2013}.
+Take HeartSteps, for example, a mobile health (mHealth) study designed to increase physical activity in sedentary adults \citep{KlasnjaHS2019}.
+Here, a Jawbone sensor is used to monitor step count every minute of the participant's study day.
+Of interest in many mHealth studies is the relation of such real-time high frequency sensor data to an adverse, recurrent event process.
+In a smoking cessation mHealth study~\citep{Sense2Stop}, for example, the relation between a time-varying sensor-based measure of physiological stress and smoking lapse is of scientific interest.
+In a suicidal ideation mHealth study~\citep{Kleiman2018}, the relation of electrodermal activity (EDA) and accelerometer with self-identified moments of suicidal ideation is of scientific interest.
+
+The goal of this paper is two-fold: (1) to discuss the appropriate choice of statistical model for joint high-frequency sensor data and the recurrent event data, and (2) to construct a simple, easy-to-implement method for parameter estimation and inference.
+For (1), we discuss an important issue regarding measurement-error models when paired with recurrent event outcomes.
+For (2), we introduce a random subsampling procedure that has several benefits.
+First, the resulting inference is unbiased; however, there is a computation-efficiency trade-off.
+In particular, a
+higher sampling rate can decrease estimator variance at the cost of increased computation.
+We show via simulations that the benefits of incredibly high sampling rates is often negligible, as the contribution to the variation is small in relation to to variation in the underlying stochastic processes.
+Second, derived estimating equations are optimal, implying loss of statistical efficiency is only due to subsampling procedure and not derived methodology.
+Finally,
+implementation can leverage existing, standard software for functional data analysis and logistic regression, leading to fast adoption by domain scientists.
 
@@
 
-Suppose $n$ subjects are independently sampled with observed event
-times~$\bfT_{i} = \{ T_{i,1}, \ldots, T_{i,k_i}\}$ over some observation
-window $[0, \tau_i]$ for each subject $i = 1,\ldots, n$.  Assume the
-event times are ordered, i.e., $T_{i,j} < T_{i,j^\prime}$ for $j <
-j^\prime$. The observation window length, $\tau_i$, is the censoring
-time and is assumed independent of the event process. Let~$N_{i} (t)$
-denote the associated counting process of $\bfT_{i}$; that is, $N_i
+Suppose $n$ subjects are independently sampled with observed event times~$\bfT_{i} = \{ T_{i,1}, \ldots, T_{i,k_i}\}$ over some observation window $[0, \tau_i]$ for each subject $i = 1,\ldots, n$.
+Assume the event times are ordered, i.e., $T_{i,j} < T_{i,j^\prime}$ for $j < j^\prime$.
+The observation window length, $\tau_i$, is the censoring time and is assumed independent of the event process.
+Let~$N_{i} (t)$ denote the associated counting process of $\bfT_{i}$; that is, $N_i
 (t) = \sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.
-In this section, we assume a single-dimensional health process~$\bfx_i
-= \{ x_i (s) \}_{0 < s < \tau_i}$ for each participant is measured at
-a dense grid of time points.  Accelerometer, for example, is measured
-at a rate of 32Hz (i.e., $32$ times per second). Electrodermal
-activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e.,
-4 times per second).  Given the high frequency nature of sensor data,
-this paper assumes the process is measured continuously.
-
-Let~$H_{i,t}^{NX} = H_{i,t}^{N} \otimes H_{i,t}^{X}$ be the $\sigma$-field
-generated by all past values~$(N_i (s), x_i (s))_{0 \leq s \leq t}$.
-The instantaneous risk of an event at time~$t$ is assumed to depend on
-the health process, time-in-study, and the event history through a
-fully parametric conditional hazard function:
+In this section, we assume a single-dimensional health process~$\bfx_i = \{ x_i (s) \}_{0 < s < \tau_i}$ for each participant is measured at a dense grid of time points.
+Accelerometer, for example, is measured at a rate of 32Hz (i.e., $32$ times per second).
+Electrodermal activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e.,
+4 times per second).  Given the high frequency nature of sensor data, this paper assumes the process is measured continuously.
+
+Let~$H_{i,t}^{NX} = H_{i,t}^{N} \otimes H_{i,t}^{X}$ be the $\sigma$-field generated by all past values~$(N_i (s), x_i (s))_{0 \leq s \leq t}$.
+The instantaneous risk of an event at time~$t$ is assumed to depend on the health process, time-in-study, and the event history through a fully parametric conditional hazard function:
 \begin{equation}

### 2019-10-25 write-up/fda-recurrent.tex
Updated paper

@@
 \def\bfT{{\bf T}}
+\def\bfD{{\bf D}}
 \def\dotminussym#1#2{%
@@
 The observation window length, $\tau_i$, is the censoring time and is assumed independent of the event process.
-Let~$N_{i} (t)$ denote the associated counting process of $\bfT_{i}$; that is, $N_i
-(t) = \sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.
+Let~$N_{i} (t)$ denote the associated counting process of $\bfT_{i}$; that is, $N_i (t) = \sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.
 In this section, we assume a single-dimensional health process~$\bfx_i = \{ x_i (s) \}_{0 < s < \tau_i}$ for each participant is measured at a dense grid of time points.
@@
 Let~$H_{i,t}^{NX} = H_{i,t}^{N} \otimes H_{i,t}^{X}$ be the $\sigma$-field generated by all past values~$(N_i (s), x_i (s))_{0 \leq s \leq t}$.
-The instantaneous risk of an event at time~$t$ is assumed to depend on the health process, time-in-study, and the event history through a fully parametric conditional hazard function:
+In this paper, the instantaneous risk of an event at time~$t$ is assumed to depend on the health process, time-in-study, and the event history through a fully parametric conditional hazard function:
 \begin{equation}
@@
 \lim_{\delta \to 0} \delta^{-1} \pr \left( N_i(t+\delta) - N_i(t) > 0
-  \given H_{i,t}^{NX}; \theta \right),
-\end{equation}
-where~$\theta$ is the parameter vector. Equation~\eqref{eq:hazard}
-states the hazard may depend on past values of the event history and
-covariates.  For high frequency physiological data, we assume that
-current risk is log-additive and depends on a linear functional of the
-health process over some recent window of time and some pre-specified
-features of the counting process; that is,
+  \given H_{i,t}^{NX} \right),
+\end{equation}
+where~$\theta$ is the parameter vector.
+Equation~\eqref{eq:hazard} states the hazard may depend on past values of the event history and covariates.
+For high frequency physiological data, we assume that current risk is log-additive and depends on a linear functional of the health process over some recent window of time and some pre-specified features of the counting process; that is,
 \begin{equation}
@@
 h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) =
-h_0 (t) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
+h_0 (t; \gamma) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
   + \int_{t-\Delta}^t x_i (s) \beta(s) ds  \right)
 \end{equation}
-where~$\Delta$ is an unknown window-length, and $g_t( H_{i,t}^N ) \in
-\mathbb{R}^p$ is a $p$-length feature vector summarizing the
-event-history and time-in-study information. The final
-term~$\int_{t-\Delta}^t x_i(s) \beta(s) ds$ reflects the unknown
-linear functional form of the impact of the time-varying covariate on
-current risk.
-
-An alternative to~\eqref{eq:hazardlinear} would be to construct
-features from the sensor data history~$f_t ( H_{i,t}^{X}) \in
-\mathbb{R}^q$ and incorporated these features in the place of the
-final term. Our current approach builds linear features of~$H_{i,t}^X$
-directly from the integrated history, avoiding the feature
-construction problem -- a highly nontrivial issue for high frequency
-time-series data.  The main caveat is the additional
-parameter~$\Delta$; however, as long as the estimated~$\hat \Delta$
-exceeds~$\Delta$, then resulting estimation is unbiased albeit at a
-loss of efficiency.  Moreover, sensitivity analysis can be peformed to
-determine how choice of $\hat \Delta$ affects inference.  One
-limitation of the approach presented here is that only fully
-parametric hazard models may be fit to the data. However, a spline
-model for the log baseline hazard affords sufficient model
-flexibility.
+where~$h_0(t;\gamma)$ is a parametrized baseline hazard function,~$\Delta$ is an unknown window-length, and $g_t( H_{i,t}^N ) \in \mathbb{R}^p$ is a $p$-length feature vector summarizing the event-history and time-in-study information.
+The final term~$\int_{t-\Delta}^t x_i(s) \beta(s) ds$ reflects the unknown linear functional form of the impact of the time-varying covariate on current risk.
+
+An alternative to~\eqref{eq:hazardlinear} would be to construct features from the sensor data history~$f_t ( H_{i,t}^{X}) \in \mathbb{R}^q$ and incorporated these features in the place of the
+final term.
+Our current approach builds linear features of~$H_{i,t}^X$ directly from the integrated history, avoiding the feature construction problem -- a highly nontrivial issue for high frequency time-series data.
+The main caveat is the additional parameter~$\Delta$; however, as long as the estimated~$\hat \Delta$ exceeds~$\Delta$, then resulting estimation is unbiased albeit at a loss of efficiency.
+Moreover, sensitivity analysis can be performed to determine how choice of $\hat \Delta$ affects inference.
+One limitation of the approach presented here is that only fully parametric hazard models may be fit to the data.
+However, a spline model for the log baseline hazard affords sufficient model flexibility.
 
@@
 
-One potential criticism for~\eqref{eq:hazard} is that the health
-process may be measured with error.
-A common mathematical strategy for joint models is to consider an
-unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i
-\given \eta_i$. Take~$\eta_i$ to be a zero-mean Gaussian process with
+One potential criticism for~\eqref{eq:hazard} is that the health process may be measured with error.
+A common mathematical strategy for joint models is to consider an unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i \given \eta_i$, i.e., the two processes are conditionally independent given the latent trajectory.
+For example, take~$\eta_i$ to be a zero-mean Gaussian process with
 \begin{equation}\label{eq:jm}
@@
 \end{equation}
-where~$\epsilon (t)$ is a white-noise measurement error. Thus,
-$\eta_i (t)$ is the ``true and unobserved value of the longitudinal
-outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional
-density function of the event process~$\bfT_i$ given~$\bfx_i$ is
-\[
-\pr \left ( \bfT_i = \bft \Given \H_{i,\tau_i}^X \right) = E \left(
-  \prod_{j=1}^{k_i} h_i (t_j \given \eta) \exp \left(
-    \int_{0}^{\tau_i} h_i (s \given \eta ) ds \right) \given \bfx_i \right).
-\]
-Therefore, the conditional density function depends not only on past
-$x$-values, but also on future $x$-values.
-That is, the processes~$\bfx$ and~$\bfT$ do not satisfy
-\emph{independent evolution}~\citep{DempseyPMCC2}.
-This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous
-risk of an event at time~$t$ depends on future values of the sensor
-process. To alleviate this issue, we employ the strategy of
-taking~$\epsilon (t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero
-Gaussian process. The proposed inferential procedure only relies on
-the Gaussian assumption when~$\bfx$ is observed with missingness (see
-Section~\ref{section:missingdata}).
+where~$\epsilon_i(t)$ is a white-noise measurement error term.
+Thus, $\eta_i (t)$ is the ``true and unobserved value of the longitudinal outcome'' \citep[Sec.
+2.1, pp.3]{Rizopoulos2010}.
+The conditional survival function of the $j$th event~$T_j$ given~$\bfx_i$ and all prior events~$\bfT_{-j} := \{ T_1, \ldots, T_{j-1} \}$ is
+\[
+\pr \left ( T_j > t+s \Given \H_{i,\tau_i}^X, \bfT_{-j} = \bft_{-j}, T_j > t \right) = E \left( - \exp \left( \int_{t}^{t+s} h_i (u \given \eta ) du \right) \Given \bfx_i, \bfT_{-j} = \bft_{-j} \right),
+\]
+where the expectation is a Gaussian integral, albeit infinite-dimensional.
+For most choices of covariance structure, if the white-noise error term is non-zero then the above calculation will show the conditional survival function depends not only on past $x$-values, but also on future $x$-values, i.e.,~$\bfx$ and~$\bfT$ do not satisfy \emph{independent evolution}~\citep{DempseyPMCC2}.
+This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process.
+To ensure independent evolution, we employ the strategy of setting~$\epsilon (t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero Gaussian process.
+The proposed inferential procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed (see Section~\ref{section:missingdata}).
 
@@
 
-For the sake of notational simplicity, we leave the dependency of the
-conditional hazard function on~$H_{i,t}^{NX}$ implicit, and write~$h_i
-(t; \theta)$. The component of the log-likelihood related to the event
-process is then given by
+For the sake of notational simplicity, we leave the dependency of the conditional hazard function on~$H_{i,t}^{NX}$ implicit, and write~$h_i (t; \theta)$.
+The component of the log-likelihood related to the event process is then given by
 \[
@@
 \]
-where~$H_{i} (\tau_i ; \theta) = \int_{0}^{\tau_i} h_{i} (t; \theta)
-dt$. Solving the associated score equations~$U_n (\theta) = {\bf 0}$
-yields the maximum likelihood estimator~$\hat \theta$, where
+where~$H_{i} (\tau_i ; \theta) = \int_{0}^{\tau_i} h_{i} (t; \theta) dt$ is the cumulative hazard function.
+Solving the associated score equations~$U_n (\theta) = {\bf 0}$ yields the maximum likelihood estimator~$\hat \theta$, where
 \[
@@
 \]
-where~$h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (\tau_{i}; \theta)$
-are derivatives with respect to $\theta$.
-
-In classical joint models~\citep{Henderson2000, Tsiatis2004},
-time-varying covariates~$x_i (t)$ are observed only intermittently at
-appointment times. In our current setting, maximizing the likelihood
-is computationally prohibitive since for any~$\theta$ we must compute
-the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require
-integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which
-itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s)
-ds$.  That is, the risk model now depends on an integrated past
-history of the time-varying covariate which leads to severe increase
-in computational complexity.
+where~$h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (\tau_{i}; \theta)$ are derivatives with respect to $\theta$.
+
+In classical joint models~\citep{Henderson2000, Tsiatis2004}, time-varying covariates~$x_i (t)$ are observed only intermittently at appointment times.
+In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s)
+ds$.  That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
 
@@
 
-To solve the computational challenge we employ a point-process
-subsampling design to obtain unbiased estimates of the derivative of
-the cumulative hazards for each subject. The subsampling procedure
-treats the collected sensor data as a set of \emph{potential
-  observations}. Suppose covariate information is sampled at times
-drawn from an independent inhomogeneous Poisson point process with
-known intensity~$\pi_i (t)$. At a subsampled time~$t$, the
-\emph{windowed covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq
-  \Delta}$ and counting process features~$g_t (H_{i,t}^N)$ are
-observed. Optimal choice of~$\pi_i (t)$ is beyond the scope of this
-paper; however, simulation studies have suggested setting proportional
-to the hazard function~$h_i (t; \theta)$.
-
-An estimator is design-unbiased if its expectation is equal to that
-parameter under the probability distribution induced by the sampling
-design~\citep{Cassel1977}. Let~$D_i \subset [0,t_i]$ denote the set of
-subsampled points.  Note, by design, this set is distinct from the set
-of event times, i.e.,~$\bfT_i \cap D_i = \emptyset$.  Under
-subsampling via $\pi_i (t)$, one may compute a Horvitz-Thompson
-estimator of the derivative of the cumulative hazard:
-\[
-\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in D_i} \frac{
+To solve the computational challenge we employ a point-process subsampling design to obtain unbiased estimates of the derivative of the cumulative hazards for each subject.
+The subsampling procedure treats the collected sensor data as a set of \emph{potential observations}.
+Suppose covariate information is sampled at times drawn from an independent inhomogeneous Poisson point process with known intensity~$\pi_i (t)$.
+At a subsampled time~$t$, the \emph{windowed covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq \Delta}$ and counting process features~$g_t (H_{i,t}^N)$ are observed.
+Optimal choice of~$\pi_i (t)$ is beyond the scope of this paper; however, simulation studies have suggested setting proportional to the hazard function~$h_i (t; \theta)$.
+
+An estimator is design-unbiased if its expectation is equal to that parameter under the probability distribution induced by the sampling design~\citep{Cassel1977}.
+Let~$\bfD_i \subset [0,t_i]$ denote the set of subsampled points.
+Note, by construction, this set is distinct from the set of event times, i.e.,~$\bfT_i \cap \bfD_i = \emptyset$.
+Under subsampling via $\pi_i (t)$, one may compute a Horvitz-Thompson estimator of the derivative of the cumulative hazard:
+\[
+\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in \bfD_i} \frac{
   h^{(1)}_i (u; \theta) }{ \pi_i (u) }
@@
 \label{eq:WPest}
-\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in (\bfT_i \cup D_i)}
+\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in (\bfT_i \cup \bfD_i)}
 \frac{ h^{(1)}_i ( u; \theta ) }{ \pi_i (u) + h_i ( u ; \theta) }
 \end{equation}
-Equation~\eqref{eq:WPest} is the estimator suggested
-by~\cite{Waagepetersen2008}.  This estimator depends on the
-superposition of the event and subsampling processes.
-Proposition~\ref{prop:optimal} shows the estimator for~$\theta$
-associated with using~\eqref{eq:WPest} is the most efficient within a
-suitable class of estimators for the derivative of the cumulative
-hazard function (including the Horvitz-Thompson estimator).
-Therefore, we restrict our attention to~\eqref{eq:WPest} for the
-remainder of this paper. Letting
+Equation~\eqref{eq:WPest} is the estimator suggested by~\cite{Waagepetersen2008}.
+This estimator depends on the superposition of the event and subsampling processes.
+Proposition~\ref{prop:optimal} shows the estimator for~$\theta$ associated with using~\eqref{eq:WPest} is the most efficient within a suitable class of estimators for the derivative of the cumulative hazard function (including the Horvitz-Thompson estimator).
+Therefore, we restrict our attention to~\eqref{eq:WPest} for the remainder of this paper.
+Letting
 \begin{equation}
@@
 \end{equation}
-Equation~\eqref{eq:approxscore} are approximate score functions built
-via plug-in of the design-unbiased estimator of the derivative of the
-cumulative hazard given in~\eqref{eq:WPest}.
+Equation~\eqref{eq:approxscore} are approximate score functions built via plug-in of the design-unbiased estimator of the derivative of the cumulative hazard given in~\eqref{eq:WPest}.
+
+\begin{rmk}[Connection to design-based estimation]
+The approximate score equations given by~\eqref{eq:approxscore} arise in design-based inference of point processes.
+Design-based inference is common for spatial point processes~\citep{Waagepetersen2008} where the spatial varying covariate is observed at a random sample of locations.
+It is common in mobile health where ecological momentary assessments~\citep{Rathbun2012,Rathbun2016} are used to randomly sample individuals at various time-points to assess their emotional state.
+In the current setting, we leverage these ideas to form a subsampling protocol that can substantially reduce computationally complexity.
+Therefore, the purpose is quite different.
+Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.
+\end{rmk}
 
@@
 
-Probabilistic subsampling converts the single sensor stream~$\bfx_i$
-into a sequence of functions observed repeatedly at sampled
-times~$D_i$ and event times~$\bfT_i$. Such a data structure is
-commonly referred to as \emph{longitudinal functional
-  data}~\citep{Xiao2013, GoldSmith2015}.
-Given the large increase in longitudinal functional data in recent
-years, corresponding analysis has received much recent
-attention~\citep{Morris2003, MorrisCarroll2006,
-  Baladandayuthapani2008, Di2009, Greven2010, Staicu2010,
-  ChenMuller2012, LiGuan2014}.
-Here, we combine work by~\cite{Park2018} and~\cite{Goldsmith2011} to
-construct a computationally efficient penalized functional method for
-solving the estimation equations~$\hat U_n (\theta)$.
+Probabilistic subsampling converts the single sensor stream~$\bfx_i$ into a sequence of functions observed repeatedly at sampled times~$\bfD_i$ and event times~$\bfT_i$.
+Such a data structure is commonly referred to as \emph{longitudinal functional data}~\citep{Xiao2013, GoldSmith2015}.
+Given the large increase in longitudinal functional data in recent years, corresponding analysis has received much recent attention~\citep{Morris2003, MorrisCarroll2006, Baladandayuthapani2008, Di2009, Greven2010, Staicu2010, ChenMuller2012, LiGuan2014}.
+Here, we combine work by~\cite{Park2018} and~\cite{Goldsmith2011} to construct a computationally efficient penalized functional method for solving the estimation equations~$\hat U_n (\theta)$.
 
@@
 
-We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0
-\leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup D_i$.
-We use the sandwich smoother~\citep{Xiao2013} to estimate the
-mean~$\mu(s,t) = \E [ X(t,s)]$.
-Alternative bivariate smoothers exist, such as the kernel-based local
-linear smoother~\citep{Hastie2009} bivariate tensor product
-splines~\citep{Wood2006} and the bivariate penalized spline
-smoother~\citep{MarxEilers2005}.
-The sandwich smoother was chosen for its computational efficiency and
-estimation accuracy.
-We then define~$\tilde X(s,t) = X(s,t) - \hat \mu(s,t)$ to be the
-mean-zero process at each time~$t \in \bfT_i \cup D_i$.
+We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
+We use the sandwich smoother~\citep{Xiao2013} to estimate the mean~$\mu_y(s,t) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
+Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009} bivariate tensor product splines~\citep{Wood2006} and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
+The sandwich smoother was chosen for its computational efficiency and estimation accuracy.
+We then define~$\tilde X(s,t) = X(s,t) - \hat \mu_y(s,t)$ to be the mean-zero process at each time~$t \in \bfT_i \cup D_i$.
 
@@
 \[
-\Sigma (s, s^\prime) = \int_{0}^\tau c( (s,T), (s^\prime, T) ) g(T) dt.
-\]
-for~$0 \leq s,s^\prime \leq \Delta$, where~$c((s,t), (s^\prime,t))$ is
-the covariance function of the windowed covariate history~$X(t,\cdot)$
-and $g(T)$ is the joint sampling density of event and subsampled
-timepoints.
-Estimation of~$\Sigma$ occurs in two steps.
-First, the pooled sample covariance is calculated at a set of grid
-points:
-\[
-\tilde \Sigma (s_r , s_{r^\prime}) = \left( \sum_{i=1}^n \left | N_i
-    \cup D_i \right |\right)^{-1} \left( \sum_{i=1}^n \sum_{t \in N_i
-    \cup D_i} \tilde X (s_r,t) \tilde X (s_{r^\prime}, t) \right).
-\]
-Due to our concern over independent evolution as discussed in
-section~\ref{section:memproblems}, we do not assume that each
-observation is observed with white noise and therefore the diagonal
-elements of~$\hat \Sigma$ are not inflated. Second, the
-estimator~$\hat \Sigma$ is further smoothed again using the sandwich
-smoother~\citep{Xiao2013}.
-Note~\cite{Park2018} smooth the off-diagonal elements, while here we
-smooth the entire pooled sample covariance matrix. All negative
-eigenvalues are set to zero to ensure positive semi-definiteness.
-The result is used as an estimator~$\hat \Sigma$ for the pooled
-covariance $\Sigma$.
-
-Next, we take the spectral decomposition of the estimated covariance
-function; let~$\{ \hat \psi_k (s),\hat \lambda_k \}_{k \geq 1}$ be the
-resulting sequence of eigenfunctions and eigenvalues.
-The key benefit of the marginal covariance approach is that it allows
-us to compute a single, time-invariant basis expansion; this reduces
-the computational burden by avoiding the three dimensional covariance
-function (i.e., covariance depends on~$t$) and associated spectral
-decomposition in methods considered by \cite{ChenMuller2012}.
-Using the Karhunen-Lo{\`e}ve decomposition, we can represent~$X(t,s)$
-by
-\[
-X(s,t) = \hat \mu(s,t) + \sum_{k=1}^{\infty} \hat c_{i,k} (t) \hat
-\psi_k (s) \approx \hat \mu(s,t) + {\bf c}_{i} (t)^\top \mathbold{\hat
-  \psi} (s)
-\]
-where $\hat c_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (s,t) \hat
-\psi_k (s) ds$, ${\bf c}_i (t) = (c_{i,1} (t), \ldots, c_{i,K_x}
-(t))^\top$, $\mathbold{\hat \psi} (s) = (\hat \psi_1 (s), \ldots, \hat
-\psi_{K_x} (s))^\top$, and~$K_x < \infty$ is the truncation level of
-the infinite expansion. Following~\cite{Goldsmith2011}, we set~$K_x$
-to satisfy identifiability constraints (see Section~\ref{section:beta}
-for details).
+\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (s,T), (s^\prime, T) ) g_y(T) dt.
+\]
+for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((s,t), (s^\prime,t))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $g_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
+Estimation of~$\Sigma_y$ occurs in two steps.
+For simplicity, we present the steps for subsampled times (i.e., $y=0$) but the steps are the same for event times as well.
+First, the pooled sample covariance is calculated at a set of grid points:
+\[
+\tilde \Sigma_0 (s_r , s_{r^\prime}) = \left( \sum_{i=1}^n \left | \bfD_i \right |\right)^{-1} \left( \sum_{i=1}^n \sum_{t \in \bfD_i} \tilde X (s_r,t) \tilde X (s_{r^\prime}, t) \right).
+\]
+Due to our concern over independent evolution as discussed in section~\ref{section:memproblems}, we do not assume that each observation is observed with white noise and therefore the diagonal
+elements of~$\hat \Sigma_0$ are not inflated.
+Second, the estimator~$\hat \Sigma$ is further smoothed again using the sandwich smoother~\citep{Xiao2013}.
+Note~\cite{Park2018} smooth the off-diagonal elements, while here we smooth the entire pooled sample covariance matrix.
+All negative eigenvalues are set to zero to ensure positive semi-definiteness.
+The result is used as an estimator~$\hat \Sigma_0$ for the pooled covariance $\Sigma_0$.
+
+Next, we take the spectral decomposition of the estimated covariance function; let $\{ \hat \psi^{(0)}_k (s),\hat \lambda^{(0)}_k \}_{k \geq 1}$ be the resulting sequence of eigenfunctions and eigenvalues.
+The key benefit of the marginal covariance approach is that it allows us to compute a single, time-invariant basis expansion; this reduces the computational burden by avoiding the three dimensional covariance function (i.e., covariance depends on~$t$) and associated spectral decomposition in methods considered by \cite{ChenMuller2012}.
+Using the Karhunen-Lo{\`e}ve decomposition, we can represent~$X(t,s)$ for $t \in \bfT_i \cap \bfD_i$ by
+\[
+X(s,t) = \hat \mu_y (s,t) + \sum_{k=1}^{\infty} \hat c^{(y)}_{i,k} (t) \hat
+\psi^{(y)}_k (s) \approx \hat \mu_y (s,t) + {\bf c}^{(y)}_{i} (t)^\top \mathbold{\hat
+  \psi}^{(y)} (s)
+\]
+where $\hat c^{(y)}_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (s,t) \hat \psi^{(y)}_k (s) ds$, ${\bf c}^{(y)}_i (t) = (c^{(y)}_{i,1} (t), \ldots, c^{(y)}_{i,K_x} (t))^\top$, $\mathbold{\hat \psi}^{(y)} (s) = (\hat \psi^{(y)}_1 (s), \ldots, \hat \psi^{(y)}_{K_x} (s))^\top$, and~$K_x < \infty$ is the truncation level of the infinite expansion.
+Following~\cite{Goldsmith2011}, we set~$K_x$ to satisfy identifiability constraints (see Section~\ref{section:beta} for details).
+In subsequent sections, we leave the dependence on $y$ (i.e., whether $t \in \bfT_i$ or $\in \bfD_i$) implicit unless required for notational simplicity.
 
@@
 
-The next step of our method is modeling~$\beta(s)$. Here, we leverage
-ideas from the penalized spline literature~\citep{Ruppert2003,
-  Wood2006book}. Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots,
-\phi_{K_b} (x) \}$ be a spline basis and assume that~$\beta(s) =
-\sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$
-where~${\bf b} = [b_1, \ldots, b_{K_b}]^{\top}$. Thus, the integral
-in~\eqref{eq:hazardlinear} can be restated as
+The next step of our method is modeling~$\beta(s)$.
+Here, we leverage ideas from the penalized spline literature~\citep{Ruppert2003, Wood2006book}.
+Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots, \phi_{K_b} (x) \}$ be a spline basis and assume that~$\beta(s) = \sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$ where~${\bf b} = [b_1, \ldots, b_{K_b}]^{\top}$.
+Thus, the integral in~\eqref{eq:hazardlinear} can be restated as
 \begin{align*}
@@
     \mathbold{\phi} (s) {\bf b} \right] \\
-  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\hat \psi, \phi} ] {\bf b}
+  &= [ ( M_{i,t} )^\top + (C_{i,t})^\top J_{\hat \psi, \phi} ] {\bf b}
 \end{align*}
-where~$M_{i,t} = (M_{i,1,t}, \ldots, M_{i,K_b,t})$, $M_{i,j,t}
-= \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\hat \psi,
-  \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th
-entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s)
-ds$~\citep{RamsaySilverman2005}.
-
-Given the basis for~$\beta(t)$, the model depends on choice of
-both~$K_b$ and~$K_x$.  We follow~\cite{Ruppert2002} by choosing $K_b$
-large enough to prevent undersmoothing and~$K_x \geq K_b$ to satisfy
-identifability constraints.
-While our theoretical analysis considers truncation levels that depend
-on~$n$, in practice, we follow the simple rule of thumb and set~$K_b
-= K_x = 35$. As long as the choices of~$K_x$ and $K_b$ are large
-enough, their impact on estimation is typically negligible.
-%We formalize this statement in Theorem~\ref{thm:truncation}.
-Below, we will exploit a connection between~\eqref{eq:approxscore}
-and score equations for a logistic regression model.  Before moving
-on, we introduce notation:
+where~$M_{i,t} = (M_{i,1,t}, \ldots, M_{i,K_b,t})$, $M_{i,j,t} = \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\hat \psi, \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s) ds$~\citep{RamsaySilverman2005}.
+
+Given the basis for~$\beta(t)$, the model depends on choice of both~$K_b$ and~$K_x$.
+We follow~\cite{Ruppert2002} by choosing $K_b$ large enough to prevent under-smoothing and~$K_x \geq K_b$ to satisfy identifiability constraints.
+While our theoretical analysis considers truncation levels that depend on~$n$, in practice, we follow the simple rule of thumb and set~$K_b = K_x = 35$.
+As long as the choices of~$K_x$ and $K_b$ are large enough, their impact on estimation is typically negligible.
+Below, we will exploit a connection between~\eqref{eq:approxscore} and score equations for a logistic regression model.
+Before moving on, we introduce some additional notation.
+First, let the baseline hazard $h_0(t; \gamma) := \exp(Z_{i,t}^\top \gamma)$ where~$Z_{i,t} \in \mathbb{R}^{q}$ parametrizes some piecewise-constant baseline   be :
 \begin{equation}
@@
 \end{equation}
-where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp (
-Z_{i,t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline
-intensity function. We write~$\tilde U_n (\theta)$ to denote
-the approximate score function when substituting
-in~\eqref{eq:approx_hazard} for~\eqref{eq:hazardlinear}.
+where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{i,t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline intensity function.
+We write~$\tilde U_n (\theta)$ to denote the approximate score function when substituting in~\eqref{eq:approx_hazard} for~\eqref{eq:hazardlinear}.
 

### 2019-10-29 write-up/fda-recurrent.tex
Update fda-recurrent.tex
Updating the tex file to get the math right.

@@
 Before moving on, we introduce some additional notation.
-First, let the baseline hazard $h_0(t; \gamma) := \exp(Z_{i,t}^\top \gamma)$ where~$Z_{i,t} \in \mathbb{R}^{q}$ parametrizes some piecewise-constant baseline   be :
+Define
 \begin{equation}
@@
 h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) \approx
-\exp \left( Z_{i,t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
+\exp \left( Z_{t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
   + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\hat \psi, \phi} {\bf b} \right)
@@
 \end{equation}
-where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{i,t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline intensity function.
+where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline intensity function.
 We write~$\tilde U_n (\theta)$ to denote the approximate score function when substituting in~\eqref{eq:approx_hazard} for~\eqref{eq:hazardlinear}.
@@
 \label{lemma:logistic}
-Under weights~\eqref{eq:waage_weights} and the log-linear intensity
-function~\eqref{eq:approx_hazard}, the approximate score
-function~$\tilde U_n (\theta)$ is equivalent to
+Under weights~\eqref{eq:waage_weights} and the log-linear intensity function~\eqref{eq:approx_hazard}, the approximate score function~$\tilde U_n (\theta)$ is equivalent to
 \[
@@
 \]
-where~$\tilde W_{i,t} = -W_{i,t}$. This is the score function for
-logistic regression with binary response~$Y_i(t)$ for $t \in N_i \cup
-D_i$ and~$i \in [n]$ where~$Y_i(t) = 1[t \in N_i]$, offset~$\log \pi_i
-(t)$, and covariates~$\tilde W_{i,t}$.
+where~$\tilde W_{i,t} = -W_{i,t}$.
+This is the score function for logistic regression with binary response~$Y_i(t)$ for $t \in \bfT_i \cup \bfD_i$ and~$i \in [n]$ where~$Y_i(t) = 1[t \in \bfT_i]$, offset~$\log \pi_i (t)$, and covariates~$\tilde W_{i,t}$.
 \end{lemma}
 
+The connection to logistic regression allows our methodology to leverage ``off-the-shelf'' software.
+The main complication is pre-processing of the functional data; however, these additional steps can also be taken care of via existing software.
+Therefore, the entire data analytic pipeline is easy-to-implement and require minimal additional effort by the end-user.
+Before demonstrating the methodology via simulation in Section~\ref{section:simstudy}) and a worked example in Section~\ref{section:example}, we detail a theoretical analysis of our proposal.
+
 \subsection{Theoretical analysis}
 
-Our theoretical analysis requires the several
-assumptions, some regarding the subsampling procedure, some for the
-event process, and some for the functional data. We state these
-assumptions and then our main theorems.
-We start by assuming there exists a~$\tau < \infty$ such that all
-individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
-Moreover, define $R_i (t)$ to be the at risk indicator for
-participant~$i$.  That is,
-\[
-R_i (t) = \left \{ \begin{array}{c c} 1 & t \in (0,\tau_i) \\ 0
-                                        & t > \tau_i\end{array} \right .
-\]
-Lemma~\ref{lemma:simpleasym} will be proven under regularity
-conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
-following additional assumptions:
+Our theoretical analysis requires assumptions regarding the subsampling procedure, the event process, and the functional data.
+We state these assumptions and then our main theorems.
+We start by assuming there exists a~$\tau < \infty$ such that all individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
+Moreover, define $R_i (t)$ to be the at risk indicator for participant~$i$, i.e., $R_i (t) = {\bf 1} [t \in (0,\tau_i)]$.
+Asymptotic theory provided in Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp.
+420--421]{Andersen1993} along with the following additional assumptions:
 
@@
 
-We also require several assumptions due to the truncation of the
-Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
+We also require several assumptions due to the truncation of the Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
 \begin{assumption}[Functional assumptions~\citep{Park2018}] \normalfont
 \label{assumption:truncation}
-The following assumptions are standard in prior work on longitudinal
-functional data analysis~\citep{Park2018, Yao2005, ChenMuller2012}:
+The following assumptions are standard in prior work on longitudinal functional data analysis~\citep{Park2018, Yao2005, ChenMuller2012}:
 \begin{enumerate}[label=(A.\arabic*)]
@@
   \mathcal{S} \times \mathcal{T})$
-\item\label{A2} The sampling density~$g(T)$ is continuous
-  and~$\sup |g(T)| < \infty$
+\item\label{A2} The subsampling and conditional intensity rate functions~$g_y(T)$ are continuous
+  and~$\sup |g_y(T)| < \infty$.
 \item\label{A3} $\E[X(s,T) X(s^\prime, T) X(s,T^\prime)
@@
 \end{assumption}
-Assumption~\ref{A3} is guaranteed for by~\ref{E1} and~\ref{E3}, but is
-stated here to emphasize its importance for the marginal covariance
-procedure of~\cite{Park2018}.
-Finally, for simplicity, we assume that there exists~$b^\star$
-such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the
-true function~$\beta(t)$ sits in the span of the spline basis
-expansion.
-With the necessary assumptions stated, we can now state
-Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
+Assumption~\ref{A3} is guaranteed by~\ref{E1} and~\ref{E3}, but is stated here to emphasize its importance for the marginal covariance procedure of~\cite{Park2018}.
+Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the
+true function~$\beta(t)$ sits in the span of the spline basis expansion. With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
 \begin{lemma} \normalfont
@@
 
-\section{Extensions}
-
-In this section, we demonstrate the flexibility of our approach by
-exploring extensions in several important directions to ensure these
-methods are robust for practical use.
-This section will continue to leverage the connection to generalized
-functional linear models provided by Lemma~\ref{lemma:logistic}.
-
-\subsection{Missing data}
-\label{section:missingdata}
-\walt{Question: How to perform multiple imputation for~$X(\cdot,t)$
-  and $X(\cdot, t^\prime)$ for~$|t - t^\prime| <\Delta$? That is, how
-  do you perform multiple imputation with this overlapping structure?}
-
-A core concern in mHealth is that sensor data can often be missing for
-intervals of time due to sensor wearing issue (e.g., sensor band
-becomes loose, turns off accidentally, data packets lost, or battery
-dies). In the suicidal ideation case study, for example, there are
-2139 self-identified moments of distress across all 91 participants.
-Of these, 1289 event times had complete data for the prior thirty
-minutes, 1984 had fraction of missing data on a fine grid less than
-30\%, and 1998 had fraction of missing data on a fine grid less than
-10\%.
-
-Missing data is a critical issue because~$c_{i,k} (t)$ cannot be
-estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$.
-Moreover, we want standard errors to reflect the uncertainty
-in these coefficients when missing data is prevalent.
-\cite{Goldsmith2011} suggest using best linear unbiased predictors
-(BLUP) or posterior modes in the mixed effects model to
-estimate~$c_{i,k} (t)$; however, one criticism would be when there is
-substantial variability in these estimates.  To deal with this,
-\cite{Crainiceanu2010} take a full Bayesian analysis.
-\cite{Yao2005} introduced PACE as an alternative frequentist method.
-However,~\cite{Petrovich2018} shows that for sparse, irregular
-longitudinal, the imputation model should not ignore the outcome
-variable~$Y_i (t)$.
-
-Here we present an extension of~\cite{Petrovich2018} to our setting by
-leveraging Lemma~\ref{lemma:logistic} and the marginal covariance
-estimation procedure to construct a multiple imputation procedure.
-Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at
-times~$\{ s_{i,r} \}_{r=1}^{k_{it}} \}$ in~$[0,\Delta]$. Then
-\begin{align}
-\label{eq:impute}
-\E [ X_i (s,t) \given Y_i(t) = y, \bfx_i (t) ]
-  &= \mu_y (s,t) + {\bf a}_{i,t}^\top (s) {\bf B}_{i,t} (\bfx_i (t) -
-    \mathbold{\mu}_i (t) ) \\
-\text{Cov} \left( X_i (s,t), X_i (s^\prime, t) \given Y_i (t) = y,
-  \bfx_i (t) \right)
-  &= \Sigma_{t} (s, s^\prime) -
-    {\bf a}_{i,t} (s)^\top {\bf B}_{i,t} {\bf a}_{i,t} (s^\prime)
-\end{align}
-where we have
-\[
-{\bf a}_{i,t} (s)^\top = \left( \begin{array}{c} \Sigma_t (s_{i,1}, s) \\ \vdots \\
-                       \Sigma_t (s_{i,k_{it}}, s) \end{array} \right); \quad
-{\bf B}_{i,t}^{-1} = \left(
-  \begin{array}{c c c}
-    \Sigma_t(s_{i,1}, s_{i,1}) & \Sigma_t(s_{i,1}, s_{i,2}) & \cdots \\
-    \Sigma_t(s_{i,1}, s_{i,2}) & \ddots & \vdots \\
-    \vdots & \cdots & \Sigma_t (s_{i,k_{it}}, s_{i,k_{it}})
-  \end{array} \right),
-\]
-$\mathbold{\mu}_i (t) = \E [ \bfx_i (t) \given Y_i (t) = y] = \{
-\mu_y (s_{j}, t) \}_{j=1}^r$ and~$\mu_y (s,t)$ is the mean of~$X(t,s)$
-from group~$y$, and~$\Sigma_t ( s^\prime, s)$ is the covariance
-between~$X(s^\prime,t)$ and $X(s,t)$ for $s^\prime, s \in [0, \Delta]$
-and~$t \in \mathbb{R}_+$.
-These covariances can be estimated using the methods considered by
-\cite{ChenMuller2012}.
-% Our imputation procedure instead uses the
-% marginal covariance~$\Xi$ as a proxy for the true covariance
-% function~$\Sigma_t(\cdot, \cdot)$.
-% We found this to work well in simulations as~$\Sigma_t$ was relatively
-% constant across~$t$; however, this imputation strategy may be
-% inappropriate when the covariance structure changes significantly with
-% time~$t$.
-
-The imputation strategy generalizes the procedure introduced
-by~\cite{Petrovich2018} to the longitudinal functional data
-setting.  This is only possible due to the connection established with
-logistic regression by Lemma~\ref{lemma:logistic}.
-We consider a multiple imputation estimate with~$M$ imputations;
-the likelihood estimate is given by~$\bar \theta = M^{-1}
-\sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate
-from the~$m$th imputated dataset; the covariance matrix can be
-estimated by
-\[
-M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
-\sum_{m=1}^M (\hat \theta^{(m)} - \bar \theta) (\hat \theta^{(m)} -
-\bar \theta)
-\]
-where~$\hat \Xi^{(m)}$ is the estimator~\eqref{eq:fisher} for
-the~$m$th imputated dataset.
-
-\newpage
-
-\subsection{Multilevel models}
-
-\subsection{Multiple sensors and availability}
-
-\section{Simulation study}
+\section{Simulation study} \label{section:simstudy}
 
@@
 
-\section{A worked example: Adolescent psychiatric inpatient mHealth study}
+\section{Extensions}
+
+In this section, we demonstrate the flexibility of our approach by exploring extensions in several important directions to ensure these methods are robust for practical use.
+This section will continue to leverage the connection to generalized functional linear models provided by Lemma~\ref{lemma:logistic}.
+
+\subsection{Missing data}
+\label{section:missingdata}
+\walt{Question: How to perform multiple imputation for~$X(\cdot,t)$
+  and $X(\cdot, t^\prime)$ for~$|t - t^\prime| <\Delta$? That is, how
+  do you perform multiple imputation with this overlapping structure?}
+
+A core concern in mHealth is that sensor data can often be missing for
+intervals of time due to sensor wearing issue (e.g., sensor band
+becomes loose, turns off accidentally, data packets lost, or battery
+dies). In the suicidal ideation case study, for example, there are
+2139 self-identified moments of distress across all 91 participants.
+Of these, 1289 event times had complete data for the prior thirty
+minutes, 1984 had fraction of missing data on a fine grid less than
+30\%, and 1998 had fraction of missing data on a fine grid less than
+10\%.
+
+Missing data is a critical issue because~$c_{i,k} (t)$ cannot be
+estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$.
+Moreover, we want standard errors to reflect the uncertainty
+in these coefficients when missing data is prevalent.
+\cite{Goldsmith2011} suggest using best linear unbiased predictors
+(BLUP) or posterior modes in the mixed effects model to
+estimate~$c_{i,k} (t)$; however, one criticism would be when there is
+substantial variability in these estimates.  To deal with this,
+\cite{Crainiceanu2010} take a full Bayesian analysis.
+\cite{Yao2005} introduced PACE as an alternative frequentist method.
+However,~\cite{Petrovich2018} shows that for sparse, irregular
+longitudinal, the imputation model should not ignore the outcome
+variable~$Y_i (t)$.
+
+Here we present an extension of~\cite{Petrovich2018} to our setting by
+leveraging Lemma~\ref{lemma:logistic} and the marginal covariance
+estimation procedure to construct a multiple imputation procedure.
+Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at
+times~$\{ s_{i,r} \}_{r=1}^{k_{it}} \}$ in~$[0,\Delta]$. Then
+\begin{align}
+\label{eq:impute}
+\E [ X_i (s,t) \given Y_i(t) = y, \bfx_i (t) ]
+  &= \mu_y (s,t) + {\bf a}_{i,t}^\top (s) {\bf B}_{i,t} (\bfx_i (t) -
+    \mathbold{\mu}_i (t) ) \\
+\text{Cov} \left( X_i (s,t), X_i (s^\prime, t) \given Y_i (t) = y,
+  \bfx_i (t) \right)
+  &= \Sigma_{t} (s, s^\prime) -
+    {\bf a}_{i,t} (s)^\top {\bf B}_{i,t} {\bf a}_{i,t} (s^\prime)
+\end{align}
+where we have
+\[
+{\bf a}_{i,t} (s)^\top = \left( \begin{array}{c} \Sigma_t (s_{i,1}, s) \\ \vdots \\
+                       \Sigma_t (s_{i,k_{it}}, s) \end{array} \right); \quad
+{\bf B}_{i,t}^{-1} = \left(
+  \begin{array}{c c c}
+    \Sigma_t(s_{i,1}, s_{i,1}) & \Sigma_t(s_{i,1}, s_{i,2}) & \cdots \\
+    \Sigma_t(s_{i,1}, s_{i,2}) & \ddots & \vdots \\
+    \vdots & \cdots & \Sigma_t (s_{i,k_{it}}, s_{i,k_{it}})
+  \end{array} \right),
+\]
+$\mathbold{\mu}_i (t) = \E [ \bfx_i (t) \given Y_i (t) = y] = \{
+\mu_y (s_{j}, t) \}_{j=1}^r$ and~$\mu_y (s,t)$ is the mean of~$X(t,s)$
+from group~$y$, and~$\Sigma_t ( s^\prime, s)$ is the covariance
+between~$X(s^\prime,t)$ and $X(s,t)$ for $s^\prime, s \in [0, \Delta]$
+and~$t \in \mathbb{R}_+$.
+These covariances can be estimated using the methods considered by
+\cite{ChenMuller2012}.
+% Our imputation procedure instead uses the
+% marginal covariance~$\Xi$ as a proxy for the true covariance
+% function~$\Sigma_t(\cdot, \cdot)$.
+% We found this to work well in simulations as~$\Sigma_t$ was relatively
+% constant across~$t$; however, this imputation strategy may be
+% inappropriate when the covariance structure changes significantly with
+% time~$t$.
+
+The imputation strategy generalizes the procedure introduced
+by~\cite{Petrovich2018} to the longitudinal functional data
+setting.  This is only possible due to the connection established with
+logistic regression by Lemma~\ref{lemma:logistic}.
+We consider a multiple imputation estimate with~$M$ imputations;
+the likelihood estimate is given by~$\bar \theta = M^{-1}
+\sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate
+from the~$m$th imputated dataset; the covariance matrix can be
+estimated by
+\[
+M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
+\sum_{m=1}^M (\hat \theta^{(m)} - \bar \theta) (\hat \theta^{(m)} -
+\bar \theta)
+\]
+where~$\hat \Xi^{(m)}$ is the estimator~\eqref{eq:fisher} for
+the~$m$th imputated dataset.
+
+\newpage
+
+\subsection{Multilevel models}
+
+\subsection{Multiple sensors and availability}
+
+\section{A worked example: Adolescent psychiatric inpatient mHealth study} \label{section:example}
 
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:logistic}]
-Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
-  \theta)}{h_i (t; \theta)}$. Under the log-linear
-intensity function given by~\eqref{eq:approx_hazard},
-$\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}$ and
+Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t; \theta)}{h_i (t; \theta)}$.
+Under the log-linear intensity function given by~\eqref{eq:approx_hazard}, $\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}$ and
 \[
@@
 \tilde U_n (\theta)
-  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
-    \sum_{t \in N_i \cup D_i} \frac{\exp \left( W_{i,t}^\top \theta
-    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}
-    W_{i,t} \right \} \\
-  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
-    + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} -
-    \sum_{t \in D_i} \frac{\exp \left( W_{i,t}^\top \theta
-    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}
-    W_{i,t} \right \} \\
-  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
+  &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} W_{i,t} - \sum_{t \in \bfT_i \cup \bfD_i} \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}  W_{i,t} \right \} \\
+  &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} \frac{\pi_i (t)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} - \sum_{t \in \bfD_i} \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} \right \} \\
+  &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} w_{i} (t; \theta) W_{i,t} -
     \sum_{t \in D_i} (1 - w_i (t; \theta)) W_{i,t} \right \} \\
-  &= - \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \left[ {\bf 1} [t \in
-    D_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
-  &= \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
-    \left[ {\bf 1} [t \in D_i]  - \frac{1}{1 + \exp\left( - (
+  &= - \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \left[ {\bf 1} [t \in \bfD_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
+  &= \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
+    \left[ {\bf 1} [t \in \bfD_i]  - \frac{1}{1 + \exp\left( - (
           \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
-Lemma~\ref{lemma:simpleasym} will be proven under regularity
-conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
-following assumptions related to the subsampling rate~$\pi_i (t)$.
+Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
 We define the joint counting process
@@
 \]
-where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations
-with $\int_0^\infty X_i(t-s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$
-with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
+where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t-s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
 \cite{Andersen1993} (Theorem VI.1.1) show that~$n^{-1} U_n (\theta_0)
@@
 \end{align*}
-where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}.
-Lenglart's inequality \walt{check} implies the second term converges
-in probability to zero under Assumptions~\ref{A1} and~\ref{A2}.
+where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}. Lenglart's inequality implies the second term converges in probability to zero under Assumptions~\ref{E1} and~\ref{E2}.
 
@@
   &n^{-1} \left \| \hat U_n (\theta_0) - \tilde U_n (\theta_0) \right \| \\
-  \leq & M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \bigg \| g
+  \leq & M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \bigg \| g
          \left( W_{i,t}^\top \theta  + \sum_{k=K_x+1}^\infty \hat
@@
          \theta + \log (\pi_i (t)) \right) \bigg \| \\
-&\leq M n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+&\leq M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
     \sup_{x \in \mathbb{R}} \left \| g^\prime \left( x \right)
@@
   \right \| \\
-&= \frac{M}{4} n^{-1} \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
+&= \frac{M}{4} n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
   \times \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
@@
   \left( X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
-  \phi_l (s) b_l\right \| g(T) dT
+  \phi_l (s) b_l\right \| (g_0(T) + g_1(T)) dT
 \end{align*}
@@
 truncation level goes to infinity.
+
+The same argument can be used in conjunction under assumptions~\ref{E1} and~\ref{E4} to prove~\eqref{eq:andersen2}.
 \end{proof}

### 2019-12-02 write-up/fda-recurrent.tex
Making all paragraphs without jumps
updating the tex file

@@
-\documentclass[12pt]{amsart}
+\documentclass[11pt]{amsart}
 \usepackage{amsmath}
@@
 The latter can be reduced by increasing the sampling rate; however, this leads to increased computational costs.
-The approximate score equations are equivalent to logistic regression score equations, allowing for standard,
-``off-the-shelf'' software to be used in fitting these models.
+The approximate score equations are equivalent to logistic regression score equations, allowing for standard, ``off-the-shelf'' software to be used in fitting these models.
 Simulations demonstrate the method and efficiency-computation trade-off.
@@
 First, the resulting inference is unbiased; however, there is a computation-efficiency trade-off.
-In particular, a
-higher sampling rate can decrease estimator variance at the cost of increased computation.
+In particular, a higher sampling rate can decrease estimator variance at the cost of increased computation.
 We show via simulations that the benefits of incredibly high sampling rates is often negligible, as the contribution to the variation is small in relation to to variation in the underlying stochastic processes.
 Second, derived estimating equations are optimal, implying loss of statistical efficiency is only due to subsampling procedure and not derived methodology.
-Finally,
-implementation can leverage existing, standard software for functional data analysis and logistic regression, leading to fast adoption by domain scientists.
+Finally, implementation can leverage existing, standard software for functional data analysis and logistic regression, leading to fast adoption by domain scientists.
 
@@
 Accelerometer, for example, is measured at a rate of 32Hz (i.e., $32$ times per second).
-Electrodermal activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e.,
-4 times per second).  Given the high frequency nature of sensor data, this paper assumes the process is measured continuously.
+Electrodermal activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e., 4 times per second).
+Given the high frequency nature of sensor data, this paper assumes the process is measured continuously.
 
@@
 
-An alternative to~\eqref{eq:hazardlinear} would be to construct features from the sensor data history~$f_t ( H_{i,t}^{X}) \in \mathbb{R}^q$ and incorporated these features in the place of the
-final term.
+An alternative to~\eqref{eq:hazardlinear} would be to construct features from the sensor data history~$f_t ( H_{i,t}^{X}) \in \mathbb{R}^q$ and incorporated these features in the place of the final term.
 Our current approach builds linear features of~$H_{i,t}^X$ directly from the integrated history, avoiding the feature construction problem -- a highly nontrivial issue for high frequency time-series data.
@@
 In classical joint models~\citep{Henderson2000, Tsiatis2004}, time-varying covariates~$x_i (t)$ are observed only intermittently at appointment times.
-In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s)
-ds$.  That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
+In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$.
+That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
 
@@
 \]
-An alternative design-unbiased estimator of the derivative of the
-cumulative hazards is given by
+An alternative design-unbiased estimator of the derivative of the cumulative hazards is given by
 \begin{equation}
@@
 \]
-Due to our concern over independent evolution as discussed in section~\ref{section:memproblems}, we do not assume that each observation is observed with white noise and therefore the diagonal
-elements of~$\hat \Sigma_0$ are not inflated.
+Due to our concern over independent evolution as discussed in section~\ref{section:memproblems}, we do not assume that each observation is observed with white noise and therefore the diagonal elements of~$\hat \Sigma_0$ are not inflated.
 Second, the estimator~$\hat \Sigma$ is further smoothed again using the sandwich smoother~\citep{Xiao2013}.
@@
 
-We next establish a connection between the above approximate score
-equations~$\tilde U_n (\theta)$ and the score equations for a logistic
-regression model. We can then exploit this connection to allow the
-model to be fit robustly using standard mixed effects
-software~\citep{Ruppert2002, McCulloch2001}.
+We next establish a connection between the above approximate score equations~$\tilde U_n (\theta)$ and the score equations for a logistic regression model.
+We can then exploit this connection to allow the model to be fit robustly using standard mixed effects software~\citep{Ruppert2002, McCulloch2001}.
 
@@
 \begin{enumerate}[label=(E.\arabic*)]
-\item\label{E1} The subsampling rate is both lower and upper
-  bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U <
-  \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$.
-\item\label{E2} There exists a nonnegative definite matrix~$\Xi
-  (\theta)$ such that
+\item\label{E1} The subsampling rate is both lower and upper bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$.
+\item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
   \[
@@
   \]
-\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for
-  all~$(i,j,t)$.
+\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for all~$(i,j,t)$.
 \item\label{E4} For all~$j,k$
@@
 Assumption~\ref{A3} is guaranteed by~\ref{E1} and~\ref{E3}, but is stated here to emphasize its importance for the marginal covariance procedure of~\cite{Park2018}.
-Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the
-true function~$\beta(t)$ sits in the span of the spline basis expansion. With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
+Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
+With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
 \begin{lemma} \normalfont
 \label{lemma:simpleasym}
-Under Assumption~\ref{assumption:events},
-Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for
-large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
+Under Assumption~\ref{assumption:events}, Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
 \[
@@
 \end{equation}
-For the log-linear intensity model, the sampling-unbiased estimator
-for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the
-previously described logistic regression model.
-This implies that subsampling from an inhomogeneous Poisson process,
-standard logistic regression software can be used to fit the recurrent
-event model by specifying an offset equal to~$\log \pi_i (t)$.
-Not only this, Lemma~\ref{prop:optimal} shows
-weights~\eqref{eq:waage_weights} are optimal within a particular class
-of weighted estimating equations.
+For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
+This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
+Not only this, Lemma~\ref{prop:optimal} shows weights~\eqref{eq:waage_weights} are optimal within a particular class of weighted estimating equations.
 
@@
 \label{prop:optimal}
-If the event process is an inhomogeneous Poisson point process with
-intensity~$h(t; \theta)$ and subsampling occurs via an independent,
-inhomogeneous Poisson point process with intensity~$\pi (t)$,
-then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most
-efficient) in the class of weighted estimating functions given
-by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any
-weight function~$w_i (t; \theta)$. This class includes the
-Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
+If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
+This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
 \end{prop}
 
-\noindent Proposition~\ref{prop:optimal} ensures the only loss of
-statistical efficiency is due to subsampling and not using a
-suboptimal estimation procedure given subsampling.
+\noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
 
@@
 
-Under assumption~\ref{assumption:truncation}, we next consider the
-statistical efficiency of our proposed estimator when compared to
-complete-data maximum likelihood estimation.
-While subsampling introduces additional variation, it may
-significantly reduce the overall computational burden. It is this
-trade-off that we next make precise.
-In particular, we consider the following choice of subsampling
-rate,~$\pi(t) = c \times h(t; \theta)$ for $c>0$. That is, the
-subsampling rate is proportional to the intensity function with
-time-independent constant~$c > 0$.
-Under this subsampling rate, the weight
-function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
+Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
+While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
+It is this trade-off that we next make precise.
+In particular, we consider the following choice of subsampling rate,~$\pi(t) = c \times h(t; \theta)$ for $c>0$.
+That is, the subsampling rate is proportional to the intensity function with time-independent constant~$c > 0$.
+Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
 Under Lemma~\ref{lemma:simpleasym},
@@
 \]
-where~$\Sigma(\theta)$ is the Fisher information of the complete-data
-maximum likelihood estimator.
-Therefore the relative efficiency is~$c/(1+c)$. For an upper bound~$H
-= \max_{t \in (0,\tau)} h(t;\theta)$, if we set~$\pi (t) = c \times
-H$, then the relative efficiency can be lower bounded by~$c / (c+1)$.
-
-Recall sensor measurements occur on the order of multiple times per
-second.  Suppose the intensity rate is bounded above by $3$ and the
-unit time scale was hours. If we can subsample the data at~$30$
-times per hour, then we have a lower bound on the efficiency of
-$0.909$. For a 4Hz sensor, this reduces the number of samples per hour
-from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
-While the computational complexity of logistic regression is linear in
-the number of samples, we get $480$ times reduction in the data size
-at the cost of a $0.909$ statistical efficiency. If we sample~$300$
-times per hour, then the efficiency loss is only $0.999$, with a $48$
-times reduction in data size. Table~\ref{tab:compvseff} provides
-additional examples for a 4Hz sensor.  The data reduction depends on
-the sensing rate; however, the lower bound on statistical efficiency
-does not.  This is because the subsampling rate only depends on the
-upper bound of the intensity function. That is, if the events are rare
-then the amount of data we need to subsample is greatly reduced with
-no impact to statistical efficiency.
+where~$\Sigma(\theta)$ is the Fisher information of the complete-data maximum likelihood estimator.
+Therefore the relative efficiency is~$c/(1+c)$.
+For an upper bound~$H = \max_{t \in (0,\tau)} h(t;\theta)$, if we set~$\pi (t) = c \times H$, then the relative efficiency can be lower bounded by~$c / (c+1)$.
+
+Recall sensor measurements occur on the order of multiple times per second.
+Suppose the intensity rate is bounded above by $3$ and the unit time scale was hours.
+If we can subsample the data at~$30$ times per hour, then we have a lower bound on the efficiency of $0.909$.
+For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
+While the computational complexity of logistic regression is linear in the number of samples, we get $480$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
+If we sample~$300$ times per hour, then the efficiency loss is only $0.999$, with a $48$ times reduction in data size.
+Table~\ref{tab:compvseff} provides additional examples for a 4Hz sensor.
+The data reduction depends on the sensing rate; however, the lower bound on statistical efficiency does not.
+This is because the subsampling rate only depends on the upper bound of the intensity function.
+That is, if the events are rare then the amount of data we need to subsample is greatly reduced with no impact to statistical efficiency.
 
@@
 \end{tabular}
-\caption{Data reduction (total number of measurements divided by
-  expected number of subsampled measurements) given 4Hz sensor,
-  subsampling constant~$c$ and an upper bound on the intensity
-  rate per hour~$H$. The subsampling rate is set at~$c \times H$;
-  note, lower bound on statistical efficiency only depends on the
-  subsampling rate.}
+\caption{Data reduction (total number of measurements divided by expected number of subsampled measurements) given 4Hz sensor, subsampling constant~$c$ and an upper bound on the intensity rate per hour~$H$.
+The subsampling rate is set at~$c \times H$; note, lower bound on statistical efficiency only depends on the subsampling rate.}
 \label{tab:compvseff}
@@
 
-Recall theoretical results were proven under the assumption that there
-exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
-To make this assumption plausible, we set~$K_b$ large enough (but less
-than~$K_x$ to ensure identifiability) such that the spline basis
-expansion is sufficiently expressive. However, in practice, such a
-choice of~$K_b$ may lead to overfitting the data. We therefore, in the
-remainder of this paper, follow~\cite{GoldSmith2015} and induce
-smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2
-I)$. Incorporating penalization is simple due to the connection
-between with the score functions for logistic regression.
+Recall theoretical results were proven under the assumption that there exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
+To make this assumption plausible, we set~$K_b$ large enough (but less than~$K_x$ to ensure identifiability) such that the spline basis expansion is sufficiently expressive.
+However, in practice, such a choice of~$K_b$ may lead to overfitting the data.
+We therefore, in the remainder of this paper, follow~\cite{GoldSmith2015} and induce smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2 I)$.
+Incorporating penalization is simple due to the connection between with the score functions for logistic regression.
 
@@
 
-Due to the connection with generalized mixed effects models, we can
-leverage existing inferential machinery to obtain variance-covariance
-estimates of model parameters.
-That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional
-matrix obtained by plugging in the REML estimates of variance
-components into the formula for the variance of~$\hat b$, then the
-standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0)
-= \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 )
-  \hat \Sigma_{bb} \phi(t_0)^\top}$.  Then the approximate 95\%
-confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96
-\sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-
-We acknowledge two important limitations of confidence intervals
-obtained via this approach. First, penalization may lead to confidence
-intervals that perform poorly in regions where~$\hat \beta(t)$ is
-oversmoothed. Second, we ignore the variability inherent in the
-longitudinal functional principal component analysis; that is, our
-estimates ignore the variability in estimation of eigenfunctions~$\hat
-\psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
-Joint modeling could be considered as in~\cite{Crainiceanu2010},
-however, this is beyond the scope of this article.
+Due to the connection with generalized mixed effects models, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
+That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional matrix obtained by plugging in the REML estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+
+We acknowledge two important limitations of confidence intervals obtained via this approach.
+First, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is oversmoothed.
+Second, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
+Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
 
@@
 We next assess the proposed methodology via a simulation study.
-Simulation studies of joint high frequency sensor and recurrent
-event data can be quite difficult to construct.  Here, we use the
-imputation strategy introduced in Section~\ref{section:missingdata} to
-circumvent some of these issues.
-
-Here, we assume each individual is observed over five 12-hour days.
-We then generate a sequence of event times via a marginal hazard
-rate that ignores sensor data.  We assume the marginal rate is
-constant and set to 1 per 12/6/1 hour(s). We then sample non-events at
-rates of 1 per 12/6/1/0.5 hours.
-This sampling strategy yields (non-)event times~$t$ along with
-labels~$Y_t \in \{0,1\}$.
-We then use the imputation strategy given by
-equation~\eqref{eq:impute} to generate the curves~$\{ X(s,t) \}$.
-For simplicity, we do not deal with
-We assume that~$Y_t = 0$ implies the curve is flat with mean 0.
-We assume that~$Y_t = 1$ implies three curves (A) the curve jumps
-around 10 minutes prior to the event and remains elevated, (B) the
-curve goes down then up and down smoothly, (C)
-These curves are given in Figure~\ref{} for three choices of
-parameters that indicate varying degrees of departure from the~$Y_t =
-0$ case.
+Simulation studies of joint high frequency sensor and recurrent event data can be quite difficult to construct.
+Here, we use the imputation strategy introduced in Section~\ref{section:missingdata} to circumvent some of these issues.
+
+Here, we assume each individual is observed over five 12-hour days. We then generate a sequence of event times via a marginal hazard rate that ignores sensor data.  We assume the marginal rate is
+constant and set to 1 per 12/6/1 hour(s).
+We then sample non-events at rates of 1 per 12/6/1/0.5 hours.
+This sampling strategy yields (non-)event times~$t$ along with labels~$Y_t \in \{0,1\}$.
+We then use the imputation strategy given by equation~\eqref{eq:impute} to generate the curves~$\{ X(s,t) \}$.
+For simplicity, we do not deal with We assume that~$Y_t = 0$ implies the curve is flat with mean 0.
+We assume that~$Y_t = 1$ implies three curves (A) the curve jumps around 10 minutes prior to the event and remains elevated, (B) the curve goes down then up and down smoothly, (C) These curves are given in Figure~XX for three choices of parameters that indicate varying degrees of departure from the~$Y_t = 0$ case.
 The associated~$\beta$ depends on the covariance~$\Sigma_t (s,t)$.
-In particular, $\beta = \Sigma_t^{-1} (\mu_{1,t} - \mu_{0,t})$ is
-guaranteed by Lemma~\ref{} and Theorem 1 of \cite{Petrovich2018}.
-For~$\beta(s)$ non time-varying implies the covariance term
-
-Using~\eqref{eq:}
-We define~$X(t,s)$ at the grid of observerations per secondw ith
+In particular, $\beta = \Sigma_t^{-1} (\mu_{1,t} - \mu_{0,t})$ is guaranteed by Lemma~XX and Theorem XX of~\cite{Petrovich2018}.
+For~$\beta(s)$ non time-varying implies the covariance term.
+We define~$X(t,s)$ at the grid of observerations per second with
 \[
@@
 
-Theorem~\ref{thm:biaseddown} shows that if we keep~$K_x = K_b = 35$
-and therefore have reconstruction error, we can expect a downward bias
-in parameter estimates.
-Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias
-in the parameter estimates.  For the former, however, we can simply
-let~$K_x = K_b$ grow as a function of data size; for the latter, we
-can see if~$\hat \beta(s)$ is statistically significant for~$s$
-near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the
-condition holds.
+Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates.
+Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.
+For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$.
+If so, we can set~$\Delta$ larger to ensure the condition holds.
 
@@
 
-A core concern in mHealth is that sensor data can often be missing for
-intervals of time due to sensor wearing issue (e.g., sensor band
-becomes loose, turns off accidentally, data packets lost, or battery
-dies). In the suicidal ideation case study, for example, there are
-2139 self-identified moments of distress across all 91 participants.
-Of these, 1289 event times had complete data for the prior thirty
-minutes, 1984 had fraction of missing data on a fine grid less than
-30\%, and 1998 had fraction of missing data on a fine grid less than
-10\%.
-
-Missing data is a critical issue because~$c_{i,k} (t)$ cannot be
-estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$.
-Moreover, we want standard errors to reflect the uncertainty
-in these coefficients when missing data is prevalent.
-\cite{Goldsmith2011} suggest using best linear unbiased predictors
-(BLUP) or posterior modes in the mixed effects model to
-estimate~$c_{i,k} (t)$; however, one criticism would be when there is
-substantial variability in these estimates.  To deal with this,
-\cite{Crainiceanu2010} take a full Bayesian analysis.
+A core concern in mHealth is that sensor data can often be missing for intervals of time due to sensor wearing issue (e.g., sensor band becomes loose, turns off accidentally, data packets lost, or battery dies).
+In the suicidal ideation case study, for example, there are 2139 self-identified moments of distress across all 91 participants.
+Of these, 1289 event times had complete data for the prior thirty minutes, 1984 had fraction of missing data on a fine grid less than 30\%, and 1998 had fraction of missing data on a fine grid less than 10\%.
+
+Missing data is a critical issue because~$c_{i,k} (t)$ cannot be estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$.
+Moreover, we want standard errors to reflect the uncertainty in these coefficients when missing data is prevalent.
+\cite{Goldsmith2011} suggest using best linear unbiased predictors (BLUP) or posterior modes in the mixed effects model to estimate~$c_{i,k} (t)$; however, one criticism would be when there is substantial variability in these estimates.
+To deal with this, \cite{Crainiceanu2010} take a full Bayesian analysis.
 \cite{Yao2005} introduced PACE as an alternative frequentist method.
-However,~\cite{Petrovich2018} shows that for sparse, irregular
-longitudinal, the imputation model should not ignore the outcome
-variable~$Y_i (t)$.
-
-Here we present an extension of~\cite{Petrovich2018} to our setting by
-leveraging Lemma~\ref{lemma:logistic} and the marginal covariance
-estimation procedure to construct a multiple imputation procedure.
-Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at
-times~$\{ s_{i,r} \}_{r=1}^{k_{it}} \}$ in~$[0,\Delta]$. Then
+However,~\cite{Petrovich2018} shows that for sparse, irregular longitudinal, the imputation model should not ignore the outcome variable~$Y_i (t)$.
+
+Here we present an extension of~\cite{Petrovich2018} to our setting by leveraging Lemma~\ref{lemma:logistic} and the marginal covariance estimation procedure to construct a multiple imputation procedure.
+Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at times~$\{ s_{i,r} \}_{r=1}^{k_{it}} \}$ in~$[0,\Delta]$.
+Then
 \begin{align}
@@
 
-The imputation strategy generalizes the procedure introduced
-by~\cite{Petrovich2018} to the longitudinal functional data
-setting.  This is only possible due to the connection established with
-logistic regression by Lemma~\ref{lemma:logistic}.
-We consider a multiple imputation estimate with~$M$ imputations;
-the likelihood estimate is given by~$\bar \theta = M^{-1}
-\sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate
-from the~$m$th imputated dataset; the covariance matrix can be
-estimated by
+The imputation strategy generalizes the procedure introduced by~\cite{Petrovich2018} to the longitudinal functional data setting.
+This is only possible due to the connection established with logistic regression by Lemma~\ref{lemma:logistic}.
+We could consider a multiple imputation estimate with~$M$ imputations; the likelihood estimate is given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate from the~$m$th imputated dataset; the covariance matrix can be estimated by
 \[
@@
 the~$m$th imputated dataset.
+
+\subsection{Bootstrap and multiple imputation under uncongenial }
+
+Positives: (1) our imputation model is ``approximately'' congenial; (2) our imputatoin model is simple and easy to use.
+Negatives: (1) clearly not congenial under many events and/or high sampling rates. (2) difficult when we overlap due to over-sampling.
+
+Congeniality matters because we want to ensure good coverage
+Q: Can we derive a variance that provides coverage
+We know our imputation models are uncongenial albeit simple to implement.
+Impact should be small if events are rare as impute only in ranges where 1,0 labels give you everything.
+To be careful, we investigate alternatives and assess impact.
+
+Simple version compared to the von Hippel's boot MI proposal.
+\begin{itemize}
+\item Bootstrap $B$ times
+\item For $b = 1, \ldots, B$, impute $M$ times
+\item Let~$\hat \theta_b = M^{-1} \sum_m \hat \theta_{b,m}$
+\item Compute $\text{Var}_{BootMI} = (B-1)^{-1} \sum_{b=1}^B (\hat \theta_b - \hat \theta_{BM})^2$ where $\hat \theta_{BM} = B^{-1} \sum_{b=1}^B \hat \theta_b$.
+\end{itemize}
 

### 2019-12-02 methods/mean_cov_estimation.R
Updated the R files from the harvard system

@@ -13,119 +13,123 @@
 ## hatX(t,s) = coefficients%*%eigenvectors + mu(t,s) 
 ## And seeing how close hatX(t,s) is to X(t,s).
 
-setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
+setwd("/n/murphy_lab/users/wdempsey/data-for-fda/data/")
+#setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
+set_of_types = c("eda", "acc")
+# for (type in set_of_types){
+#   event_complete = readRDS(paste(type, "_event_complete_HLP_2019-07-27.RDS", sep = ""))
+#   sequence = seq(-30,0,by =1/60); sensor_obs = 4:1804
+#   # plot(sequence, event_complete[3,sensor_obs], type = "l")
+#   
+#   library(refund); library(lubridate)
+#   full_obs = apply(X = event_complete, MARGIN = 1, FUN = function(x){!any(is.na(x))})
+#   # greater_than_ten_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.1})
+#   # greater_than_thirty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.3})
+#   # greater_than_fifty_percent_missing = apply(X = event_complete, MARGIN = 1, FUN = function(x){mean(!is.na(x[sensor_obs])) > 0.5})
+#   Y = event_complete[full_obs,sensor_obs]
+#   x = event_complete[full_obs,2]
+#   z = sequence
+#   est <- fbps(Y,list(x=x,z=z))
+#   # obs = 100
+#   # plot(sequence,est$Yhat[obs,]- mean(est$Yhat[obs,]))
+#   # lines(sequence, Y[obs,]-mean(Y[obs,]))
+#   
+#   ## COMPUTE ESTIMATED MARGINAL COVARIANCE 
+#   
+#   inflation = 100
+#   residual = inflation*(Y-est$Yhat)
+#   ## TAKE PAIRS OF ROWS and CALCULATE THE MSE
+#   print(paste("Made it to event, linear Sigma calc for", type))
+#   Sigma = matrix(nrow = ncol(residual), ncol = ncol(residual))
+#   for (i in 1:ncol(residual)) {
[…]

### 2019-12-17 write-up/fda-recurrent.tex
Update fda-recurrent.tex

--- 2019-10-29 write-up/fda-recurrent.tex
@@
 The sandwich smoother was chosen for its computational efficiency and estimation accuracy.
-We then define~$\tilde X(s,t) = X(s,t) - \hat \mu_y(s,t)$ to be the mean-zero process at each time~$t \in \bfT_i \cup D_i$.
+We then define~$\tilde X(s,t) = X(s,t) - \hat \mu_y(s,t)$ to be the mean-zero process at each time~$t \in \bfT_i \cup \bfD_i$.
 
@@
 \[
-\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (s,T), (s^\prime, T) ) g_y(T) dt.
-\]
-for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((s,t), (s^\prime,t))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $g_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
+\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (s,T), (s^\prime, T) ) f_y(T) dt.
+\]
+for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((s,t), (s^\prime,t))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $f_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
 Estimation of~$\Sigma_y$ occurs in two steps.
@@
 
-The connection to logistic regression allows our methodology to leverage ``off-the-shelf'' software.
+This connection established by Lemma~\ref{lemma:logistic} between our proposed methodology and logistic regression allows us to leverage ``off-the-shelf'' software.
 The main complication is pre-processing of the functional data; however, these additional steps can also be taken care of via existing software.
-Therefore, the entire data analytic pipeline is easy-to-implement and require minimal additional effort by the end-user.
-Before demonstrating the methodology via simulation in Section~\ref{section:simstudy}) and a worked example in Section~\ref{section:example}, we detail a theoretical analysis of our proposal.
+Therefore, the entire data analytic pipeline is easy-to-implement and requires minimal additional effort by the end-user.
+To see this, we briefly review the proposed inference procedure.
+
+\begin{rmk}[Inference procedure review] \normalfont
+Given observed recurrent event and high frequency data~$\{ \bfT_i, \bfx_i \}_{i=1}^n$,
+\begin{enumerate}
+\item For each $i \in [n]$, sample non-event times as a time-inhomogeneous Poisson point process with intensity according to~$\pi_i (t)$
+\item Estimate mean~$\mu_y (t,s)$ for $0 \leq s \leq \Delta$ at all event times~$t \in \cup_{i=1}^n \bfT_i$ and sampled non-event times~$t \in \cup_{i=1}^n \bfD_i$.
+\item Compute marginal covariance across event times,~$\Sigma_1$, and non-event times and~$\Sigma_0$.
+\item Compute eigendecomposition~$\{ \hat \psi_k^{(y)}, \hat \lambda_k^{()} \}$ of marginal covariance~$\Sigma_y$
+\item Use the eigendecomposition to construct~$W_{i,t}$ for all $i \in [n]$ and $t \in \bfD_i \cup \bfT_i$
+\item \label{point:log} Perform logistic regression with binary outcome~$\{ \Y_i (t) \}$  and offset of~$\log \pi_i (t)$.
+\end{enumerate}
+\end{rmk}
+
+\noindent Before demonstrating the methodology via simulation in Section~\ref{section:simstudy} and a worked example in Section~\ref{section:example}, we provide a theoretical analysis of our current proposal.
 
@@
 We start by assuming there exists a~$\tau < \infty$ such that all individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
-Moreover, define $R_i (t)$ to be the at risk indicator for participant~$i$, i.e., $R_i (t) = {\bf 1} [t \in (0,\tau_i)]$.
+Moreover, define $R_i (t)$ to be the at-risk indicator for participant~$i$, i.e., $R_i (t) = {\bf 1} [t \in (0,\tau_i)]$.
 Asymptotic theory provided in Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp.
@@
 \begin{enumerate}[label=(E.\arabic*)]
-\item\label{E1} The subsampling rate is both lower and upper
-  bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U <
-  \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$.
-\item\label{E2} There exists a nonnegative definite matrix~$\Xi
-  (\theta)$ such that
+\item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0, \tau]$ such that $R_i (t) = 1$.
+\item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
   \[
@@
 
-We also require several assumptions due to the truncation of the Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
+We also require several assumptions due to the truncation of the Karhunen-Lo{\`e}ve decomposition that represents~$X(s,t)$.
 \begin{assumption}[Functional assumptions~\citep{Park2018}] \normalfont
@@
 \begin{enumerate}[label=(A.\arabic*)]
-\item\label{A1} $X = \{ X(s, T) : (s,T) \in \mathcal{S} \times
+\item\label{A1} $X = \{ X(s, t) : (s,t) \in \mathcal{S} \times
   \mathcal{T} \}$ is a square integrable element of the $L^2 (
   \mathcal{S} \times \mathcal{T})$
-\item\label{A2} The subsampling and conditional intensity rate functions~$g_y(T)$ are continuous
-  and~$\sup |g_y(T)| < \infty$.
-\item\label{A3} $\E[X(s,T) X(s^\prime, T) X(s,T^\prime)
-  X(s^\prime, T^\prime) ] < \infty$ for
-  each $s,s^\prime \in [0,\Delta]$ and~$0 < T, T^\prime < \tau$.
-\item\label{A4} $\E[\|X(\cdot, T)\|^4] < \infty$ for each~$0< T
-  < \tau$.
+\item\label{A2} The subsampling and conditional intensity rate functions~$f_y(T)$ are continuous and~$\sup |f_y(T)| < \infty$.
+\item\label{A3} $\E[X(s,t) X(s^\prime, t) X(s,t^\prime)
+  X(s^\prime, t^\prime) ] < \infty$ for
+  each $s,s^\prime \in [0,\Delta]$ and~$0 < t, t^\prime < \tau$.
+\item\label{A4} $\E[\|X(\cdot, t)\|^4] < \infty$ for each~$0< t < \tau$.
 \end{enumerate}
 \end{assumption}
-Assumption~\ref{A3} is guaranteed by~\ref{E1} and~\ref{E3}, but is stated here to emphasize its importance for the marginal covariance procedure of~\cite{Park2018}.
-Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the
-true function~$\beta(t)$ sits in the span of the spline basis expansion. With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
+Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
+With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
 \begin{lemma} \normalfont
@@
 \end{equation}
-For the log-linear intensity model, the sampling-unbiased estimator
-for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the
-previously described logistic regression model.
-This implies that subsampling from an inhomogeneous Poisson process,
-standard logistic regression software can be used to fit the recurrent
-event model by specifying an offset equal to~$\log \pi_i (t)$.
-Not only this, Lemma~\ref{prop:optimal} shows
-weights~\eqref{eq:waage_weights} are optimal within a particular class
-of weighted estimating equations.
-
-\begin{prop} \normalfont
+For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
+This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
+Not only this, Lemma~\ref{prop:optimal} shows weights~\eqref{eq:waage_weights} are optimal within a particular class of weighted estimating equations.
+
+\begin{lemma} \normalfont
 \label{prop:optimal}
-If the event process is an inhomogeneous Poisson point process with
-intensity~$h(t; \theta)$ and subsampling occurs via an independent,
-inhomogeneous Poisson point process with intensity~$\pi (t)$,
-then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most
-efficient) in the class of weighted estimating functions given
-by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any
-weight function~$w_i (t; \theta)$. This class includes the
-Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
-\end{prop}
-
-\noindent Proposition~\ref{prop:optimal} ensures the only loss of
-statistical efficiency is due to subsampling and not using a
-suboptimal estimation procedure given subsampling.
+If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
+This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
+\end{lemma}
+
+\noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
+% This result follows from the probability-based sampling design literature~\cite{Rathbun2012}.
 
@@
 
-Under assumption~\ref{assumption:truncation}, we next consider the
-statistical efficiency of our proposed estimator when compared to
-complete-data maximum likelihood estimation.
-While subsampling introduces additional variation, it may
-significantly reduce the overall computational burden. It is this
-trade-off that we next make precise.
-In particular, we consider the following choice of subsampling
-rate,~$\pi(t) = c \times h(t; \theta)$ for $c>0$. That is, the
-subsampling rate is proportional to the intensity function with
-time-independent constant~$c > 0$.
-Under this subsampling rate, the weight
-function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
+Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
+While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
+It is this trade-off that we next make precise.
+To start, we consider the subsampling rate~$\pi(t) = c \times h(t; \theta)$ for $c>0$, i.e., the subsampling rate proportional to the intensity function with time-independent constant~$c > 0$.
+Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
 Under Lemma~\ref{lemma:simpleasym},
@@
 \]
-where~$\Sigma(\theta)$ is the Fisher information of the complete-data
-maximum likelihood estimator.
-Therefore the relative efficiency is~$c/(1+c)$. For an upper bound~$H
-= \max_{t \in (0,\tau)} h(t;\theta)$, if we set~$\pi (t) = c \times
-H$, then the relative efficiency can be lower bounded by~$c / (c+1)$.
-
-Recall sensor measurements occur on the order of multiple times per
-second.  Suppose the intensity rate is bounded above by $3$ and the
-unit time scale was hours. If we can subsample the data at~$30$
-times per hour, then we have a lower bound on the efficiency of
-$0.909$. For a 4Hz sensor, this reduces the number of samples per hour
-from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
-While the computational complexity of logistic regression is linear in
-the number of samples, we get $480$ times reduction in the data size
-at the cost of a $0.909$ statistical efficiency. If we sample~$300$
-times per hour, then the efficiency loss is only $0.999$, with a $48$
-times reduction in data size. Table~\ref{tab:compvseff} provides
-additional examples for a 4Hz sensor.  The data reduction depends on
-the sensing rate; however, the lower bound on statistical efficiency
-does not.  This is because the subsampling rate only depends on the
-upper bound of the intensity function. That is, if the events are rare
-then the amount of data we need to subsample is greatly reduced with
-no impact to statistical efficiency.
+where~$\Sigma(\theta)$ is the Fisher information of the complete-data maximum likelihood estimator.
+Therefore the relative efficiency is~$c/(1+c)$.
+For an upper bound~$H = \max_{t \in (0,\tau)} h(t;\theta)$, if we set~$\pi (t) = c \times H$, then the relative efficiency can be lower bounded by~$c / (c+1)$.
+
+Recall sensor measurements occur on the order of multiple times per second.
+Suppose the intensity rate is bounded above by $3$ and the unit time scale was hours.
+If we can subsample the data at~$30$ times per hour, then we have a lower bound on the efficiency of $0.909$.
+For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
+While the computational complexity of logistic regression is linear in the number of samples, we get $480$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
+If we sample~$300$ times per hour, then the efficiency loss is only $0.999$, with a $48$ times reduction in data size.
+Table~\ref{tab:compvseff} provides additional examples for a 4Hz sensor.
+The data reduction depends on the sensor measurement rate; however, the lower bound on statistical efficiency does not.
+This is because the subsampling rate only depends on the upper bound of the intensity function.
+That is, if the events are rare then the amount of data we need to subsample is greatly reduced with no impact to statistical efficiency.
 
@@
 
-Recall theoretical results were proven under the assumption that there
-exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
-To make this assumption plausible, we set~$K_b$ large enough (but less
-than~$K_x$ to ensure identifiability) such that the spline basis
-expansion is sufficiently expressive. However, in practice, such a
-choice of~$K_b$ may lead to overfitting the data. We therefore, in the
-remainder of this paper, follow~\cite{GoldSmith2015} and induce
-smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2
-I)$. Incorporating penalization is simple due to the connection
-between with the score functions for logistic regression.
+Recall theoretical results were proven under the assumption that there exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
+To make this assumption plausible, we set~$K_b$ large enough (but less than~$K_x$ to ensure identifiability) to ensure the spline basis expansion is sufficiently expressive.
+However, in practice, such a choice of~$K_b$ may lead to overfitting the data.
+Following~\cite{Goldsmith2011}, we choose the polynomial spline model and set~$\phi(t) = b_0 + b_1 t + b_2 t^2 + \sum_{j=3}^{K_b} (t-\kappa_j)^3$ where~$\{ \kappa_j \}_{j=3}^{K_b}$ are the chosen knots and assume~$\{ b_j \}_{j=3}^{K_b} \sim N(0, \sigma^2 I)$ to induce smoothness on the spline model.
+Combining the penalized spline formulation with Lemma~\ref{lemma:logistic} establishes a connection between our approximate score equations and solving a generalized mixed effects logistic regression with offset.
 
@@
 
-Due to the connection with generalized mixed effects models, we can
-leverage existing inferential machinery to obtain variance-covariance
-estimates of model parameters.
-That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional
-matrix obtained by plugging in the REML estimates of variance
-components into the formula for the variance of~$\hat b$, then the
-standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0)
-= \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 )
-  \hat \Sigma_{bb} \phi(t_0)^\top}$.  Then the approximate 95\%
-confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96
-\sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-
-We acknowledge two important limitations of confidence intervals
-obtained via this approach. First, penalization may lead to confidence
-intervals that perform poorly in regions where~$\hat \beta(t)$ is
-oversmoothed. Second, we ignore the variability inherent in the
-longitudinal functional principal component analysis; that is, our
-estimates ignore the variability in estimation of eigenfunctions~$\hat
-\psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
-Joint modeling could be considered as in~\cite{Crainiceanu2010},
-however, this is beyond the scope of this article.
+Due to the connection with generalized mixed effects models, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
+That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ dimensional matrix obtained by plugging in the estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+
+We acknowledge two important limitations of confidence intervals obtained via this approach.
+First, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is oversmoothed.
+Second, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
+Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
 
@@
 We next assess the proposed methodology via a simulation study.
-Simulation studies of joint high frequency sensor and recurrent
-event data can be quite difficult to construct.  Here, we use the
-imputation strategy introduced in Section~\ref{section:missingdata} to
-circumvent some of these issues.
-
-Here, we assume each individual is observed over five 12-hour days.
-We then generate a sequence of event times via a marginal hazard
-rate that ignores sensor data.  We assume the marginal rate is
-constant and set to 1 per 12/6/1 hour(s). We then sample non-events at
-rates of 1 per 12/6/1/0.5 hours.
-This sampling strategy yields (non-)event times~$t$ along with
-labels~$Y_t \in \{0,1\}$.
-We then use the imputation strategy given by
-equation~\eqref{eq:impute} to generate the curves~$\{ X(s,t) \}$.
-For simplicity, we do not deal with
-We assume that~$Y_t = 0$ implies the curve is flat with mean 0.
-We assume that~$Y_t = 1$ implies three curves (A) the curve jumps
-around 10 minutes prior to the event and remains elevated, (B) the
-curve goes down then up and down smoothly, (C)
-These curves are given in Figure~\ref{} for three choices of
-parameters that indicate varying degrees of departure from the~$Y_t =
-0$ case.
-The associated~$\beta$ depends on the covariance~$\Sigma_t (s,t)$.
-In particular, $\beta = \Sigma_t^{-1} (\mu_{1,t} - \mu_{0,t})$ is
-guaranteed by Lemma~\ref{} and Theorem 1 of \cite{Petrovich2018}.
-For~$\beta(s)$ non time-varying implies the covariance term
-
-Using~\eqref{eq:}
-We define~$X(t,s)$ at the grid of observerations per secondw ith
-\[
-  \Sigma_t (s, s^\prime) =
+Here, we assume each individual is observed over five days each defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day.
+We define~$X(t)$ at the grid of observerations as a mean-zero Gaussian process with covariance
+\[
+  \Sigma (t, t^\prime) =
   \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
@@
 \]
-where~$K_v$ is the modified Bessel function of the second kind, and
-we choose~$\nu = 5/2$ and $\sigma^2 = 1$.
-We choose three versions of~$\rho_t$: (A) constant across time
-at $0.5$, (B) constant within each day but linearly decreasing
-across days from~$0.7$ to~$0.3$, and (C) constant within each day but
-linearly increasing across days from~$0.3$ to~$0.7$.
-
-to estimate the impact of strong X dependence for close X's then we
-will GENERATE SEQUENTIALLY TO FILL IN TO ACCOUNT FOR OVERLAP!
-HOW SHOULD WE HANDLE COVARIANCE BETWEEN X's outside the range?
-WE WILL APPROXIMATE?
-
-We estimate with~$K_b = K_x = 35$ under both the logistic, penalized
-logistic, and multi-level penalized logistic models.
-
-\subsection{Mean-integrated squared error (MISE)}
-
-Since we are primarily interested in accuracy of the final
-estimates~$\hat \beta(s)$, we report the mean integrated  squared
-error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{1000}
-\sum_{j=1}^{1000} \int_{-\infty}^t ( \hat \beta_j (t) - \beta(s))^2
-ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.  We define the MISE in
-this manner to account for settings where~$\Delta$ is unknown.
-For~$\Delta$ known, the MISE simplifies to integrals from~$t-\Delta$
-to~$t$.
-Table~\ref{} show the MISE of~$\hat \beta (s)$ decomposed into the
-variance and squared biase
+where~$K_v$ is the modified Bessel function of the second kind.
+We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.5$ as well as set~$K_b = K_x = 35$.
+For simplicity, we assume~$\Sigma$ is known in computation of the the eigendecomposisions.
+Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; theta)$.
+To mimic our real data, we set
+\[
+h(t; \theta) = \exp \left( \theta_0 + \int_{0}^{\Delta} X(t-s) \beta(s) ds \right).
+\]
+We set~$\Delta = 1/(12*2)$ to mimic a $30$-minute window for a $12$-hour day.
+We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day.
+We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$
+
+We sample non-events at rates of 1 per 12/6/1/0.5 hours.
+
+Since we are primarily interested in accuracy of the final estimates~$\hat \beta(s)$, we report the mean integrated squared error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{100} \sum_{j=1}^{100} \int_{0}^{\Delta^\prime} ( \hat \beta_j (s) - \beta(s))^2 ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
+We define the MISE in this manner to account for settings where~$\Delta$ is unknown.
+Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biase
+
+\begin{table}[!th]
+\begin{tabular}{c | c c c c c}
+Sampling rate & 0.5 & 1 & 5 & 25 & 50 \\
+Variance
+Squared Bias
+MISE
+\end{tabular}
+\caption{XXX}
+\label{tab:mise}
+\end{table}
 
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
-Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
+% Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
 We define the joint counting process
@@
 \]
-where~$N_i (t)$ is the counting process with jumps at all~$t \in
-\bfT_i \cup D_i$.
-
-Asymptotic consistency is guaranteed~\cite[Theorem VI.1.1]{Andersen1993} if
+where~$N_i (t)$ is the counting process with jumps at all~$t \in \bfT_i \cup \bfD_i$. Asymptotic consistency is guaranteed~\cite[Theorem VI.1.1]{Andersen1993} if
 \begin{equation}
@@
 \end{equation}
-To prove~\eqref{eq:andersen1}, we decompose~$\hat U_n (\theta_0)$ into
+To prove~\eqref{eq:andersen1}, we decompose~$\tilde U_n (\theta_0)$ into
 three terms
@@
 \end{align*}
-where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}. Lenglart's inequality implies the second term converges in probability to zero under Assumptions~\ref{E1} and~\ref{E2}.
-
+where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}.
+Lenglart's inequality implies the second term converges in probability to zero under Assumptions~\ref{E1} and~\ref{E2}.
 The third term satisfies
@@
   \leq & M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \bigg \| g
-         \left( W_{i,t}^\top \theta  + \sum_{k=K_x+1}^\infty \hat
+         \left( \tilde W_{i,t}^\top \theta  + \sum_{k=K_x+1}^\infty \hat
          c_{i,k} (t) \int_{t-\Delta}^t \hat \psi_k (s) \phi (s)^\top b
@@
   \left( X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
-  \phi_l (s) b_l\right \| (g_0(T) + g_1(T)) dT
+  \phi_l (s) b_l\right \| (f_0(t) + f_1(t)) dt
 \end{align*}
@@
 the outside sums in terms of the integrated difference where the
-integral is with respect to the sampling distribution~$g(T)$. Finally,
-under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
+integral is with respect to the joint event and sampling distributions~$(f_0(t) + f_1 (t))$. Finally, under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
 \[
@@
 \]
-as~$K_x \to \infty$, which implies the third term goes to zero as the
-truncation level goes to infinity.
-
+as~$K_x \to \infty$, which implies the third term goes to zero as the truncation level goes to infinity.
 The same argument can be used in conjunction under assumptions~\ref{E1} and~\ref{E4} to prove~\eqref{eq:andersen2}.
+
+To prove~\eqref{eq:andersen3},
+\[
+\left | n^{-1}  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau H_{in} (t) dN_i^
+\]
 \end{proof}
@@
 \[
-\Xi (s, s^\prime) = \int_{\mathcal{T}} c( (s,T), (s^\prime, T) ) g(T) dT
-\]
-where~$g(T)$ is a superposition of the event process and the
+\Xi (s, s^\prime) = \int_{\mathcal{T}} c( (s,T), (s^\prime, T) ) f(T) dT
+\]
+where~$f(T)$ is a superposition of the event process and the
 subsampling process.

### 2019-12-17 write-up/fda-recurrent.tex

@@
 \item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0, \tau]$ such that $R_i (t) = 1$.
+\item\label{E1} The subsampling rate is both lower and upper bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$.
 \item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
@@
 With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
+
+Assumption~\ref{A3} is guaranteed by~\ref{E1} and~\ref{E3}, but is stated here to emphasize its importance for the marginal covariance procedure of~\cite{Park2018}.
+Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
+With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
+
 \begin{lemma} \normalfont
 \label{lemma:simpleasym}
-Under Assumption~\ref{assumption:events},
-Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for
-large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
+Under Assumption~\ref{assumption:events}, Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
 \[
@@
 To start, we consider the subsampling rate~$\pi(t) = c \times h(t; \theta)$ for $c>0$, i.e., the subsampling rate proportional to the intensity function with time-independent constant~$c > 0$.
+Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
+Under Lemma~\ref{lemma:simpleasym},
+
+For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
+This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
+Not only this, Lemma~\ref{prop:optimal} shows weights~\eqref{eq:waage_weights} are optimal within a particular class of weighted estimating equations.
+
+\begin{prop} \normalfont
+\label{prop:optimal}
+If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
+This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
+\end{prop}
+
+\noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
+
+\subsection{Computation versus statistical efficiency tradeoff}
+
+Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
+While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
+It is this trade-off that we next make precise.
+In particular, we consider the following choice of subsampling rate,~$\pi(t) = c \times h(t; \theta)$ for $c>0$.
+That is, the subsampling rate is proportional to the intensity function with time-independent constant~$c > 0$.
 Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
@@
 The data reduction depends on the sensor measurement rate; however, the lower bound on statistical efficiency does not.
+This is because the subsampling rate only depends on the upper bound of the intensity function.
+That is, if the events are rare then the amount of data we need to subsample is greatly reduced with no impact to statistical efficiency.
+
+Recall sensor measurements occur on the order of multiple times per second.
+Suppose the intensity rate is bounded above by $3$ and the unit time scale was hours.
+If we can subsample the data at~$30$ times per hour, then we have a lower bound on the efficiency of $0.909$.
+For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
+While the computational complexity of logistic regression is linear in the number of samples, we get $480$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
+If we sample~$300$ times per hour, then the efficiency loss is only $0.999$, with a $48$ times reduction in data size.
+Table~\ref{tab:compvseff} provides additional examples for a 4Hz sensor.
+The data reduction depends on the sensing rate; however, the lower bound on statistical efficiency does not.
 This is because the subsampling rate only depends on the upper bound of the intensity function.
@@
 That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ dimensional matrix obtained by plugging in the estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+
+Recall theoretical results were proven under the assumption that there exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
+To make this assumption plausible, we set~$K_b$ large enough (but less than~$K_x$ to ensure identifiability) such that the spline basis expansion is sufficiently expressive.
+However, in practice, such a choice of~$K_b$ may lead to overfitting the data.
+We therefore, in the remainder of this paper, follow~\cite{GoldSmith2015} and induce smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2 I)$.
+Incorporating penalization is simple due to the connection between with the score functions for logistic regression.
+
+\subsection{Confidence intervals for~$\beta(t)$}
+
+Due to the connection with generalized mixed effects models, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
+That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional matrix obtained by plugging in the REML estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
 Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.

### 2019-12-17 simulations/basic_simulation.R
Adding simulation files

@@ -1,0 +1,103 @@
+## Source the 
+source("./basic_simulation_functions.R")
+
+### Matern covariance parameters
+nu = 1/2 # controls smoothness
+sigma_sq = 1 # controls overall variance
+rho = 0.3 # controls correlation
+constant = sigma_sq / (gamma(nu) * 2^(nu-1))
+
+### User-day information
+### Each user day is t \in [0,1]
+### We use M dense equally spaced timepoints
+M = 1000 # Num of time points
+times = seq(0,1, length.out = M) # Equally spaced timepoints in [0,1]
+gap = diff(times)[1] # What is assumed gap times
+abs_diff = abs(outer(times, times, "-")) # Absolute distance 
+x = sqrt(2 * nu) * abs_diff/ rho 
+Cov_X = constant * x^nu * besselK(x, nu) # Matern Covariance
+diag(Cov_X) = 1
+C = chol(Cov_X) # Cholesky decomposition
+
+## Generate corresponding event times
+## Simulation assumes we want events near
+## (A) Want Up then Down pattern trends -> beta_1t
+## (B) Overall high mean in the window -> beta_2t
+# beta_1t = 30*c(rep(-1, 11), rep(1, 22), rep(-1, 11))
+# beta_1t = 50*sin(1:44/44*2*pi-pi/2)
+# beta_1t = 20*exp(0.1*1:44)/mean(exp(0.1*1:44))
+beta_1t = 100*sin(1:44/44*2*pi-pi/2)
+# beta_2t = 50*rep(1, 44)
+# plot(times[1:44],beta_1t+beta_2t)
+
+## Collect coefficients for first 35
+## Take covariance in the 44 timeslots
+Sigma = Cov_X[1:length(beta_1t),1:length(beta_1t)]
+eig_Sigma = eigen(Sigma)
+K_x = 35  # Pick first 35 eigen-vectors
+cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
+
+set.seed("131232")
+dataset = generate_complete_data(N = 1000, Cov_X, C, times, eig_Sigma, beta_1t)
+rm(abs_diff, C, Cov_X, x)
+
+dataset =data.frame(dataset)
+names(dataset)[1] = "userday"
+names(dataset)[2] = "Y"
+
+agg_results = aggregate(Y~userday,dataset, sum)
+table(agg_results[,2])
+
+## Construct the J matrix
+K_b = 35
[…]

### 2019-12-17 simulations/basic_simulation_functions.R

@@ -1,0 +1,80 @@
+generate_userday <- function(Cov_X, C, times, eig_Sigma, beta_1t) {
+  ## Generate X(t) for t \in [0,1]
+  X = t(C)%*%rnorm(M)  # Use Chol Decomp and Random Normals
+  # plot(times, X)
+  
+  ## Compute the integrals at each time t
+  result = rep(0,0)
+  event_nodes = (length(beta_1t)+1):length(times)
+  event_times = times[event_nodes]
+  for (t in event_nodes) {
+    temp_X = X[(t+1-length(beta_1t)):t]
+    temp_Xb = beta_1t*temp_X
+    term2 = (temp_Xb[1] + temp_Xb[length(temp_Xb)])/2 * gap + sum(temp_Xb[2:(length(temp_Xb)-1)]*gap)
+    result = c(result, term2)
+  }
+  
+  ## Calculate the logit based probabilities
+  probs = expit(logit(10/length(times))+result)
+  # plot(event_times, probs)
+  Y = sapply(probs,function(p) rbinom(n=1,size = 1,p))
+  # abline(v = event_times[Y==1])
+  events = event_times[Y==1]
+  # plot(event_times, probs, type = 'l')
+  
+  ## Given time stamp I want to generate the summary
+  nonevents = event_times[rbinom(n = length(event_times), size = 1, prob = 10/1000)==1]
+  nonevents = nonevents[!is.element(nonevents, events)]
+  
+  data = rep(0,0)
+  # print("Events at times")
+  # print(events)
+  if(length(events) > 0) {
+    for (t in events) {
+      t_loc = which(is.element(times,t))
+      X_Delta = X[(t_loc+1-length(beta_1t)):(t_loc)] # Past window of X
+      C_x = rep(0, K_x)  
+      for (i in 1:length(C_x)) {
+        C_x[i] = eig_Sigma$vectors[,i]%*%X_Delta
+      }
+      row_data = c(1,C_x)
+      data = rbind(data, row_data)
+    }
+  }
+  
+  # print("Nonevents at times")
+  # print(nonevents)
+  if(length(nonevents) > 0) {
+    for (t in nonevents) {
+      t_loc = which(is.element(times,t))
+      X_Delta = X[(t_loc+1-length(beta_1t)):t_loc] # Past window of X
+      C_x = rep(0, K_x)  
[…]

### 2020-01-27 write-up/fda-recurrent.tex
tex updating!
Adding changes from SPH paper
Edited simulations and paper.

@@
 \begin{enumerate}[label=(E.\arabic*)]
-\item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0, \tau]$ such that $R_i (t) = 1$.
-\item\label{E1} The subsampling rate is both lower and upper bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$.
+\item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$ such that $R_i (t) = 1$
 \item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
@@
 We next assess the proposed methodology via a simulation study.
-Here, we assume each individual is observed over five days each defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day.
-We define~$X(t)$ at the grid of observerations as a mean-zero Gaussian process with covariance
+Here, we assume each individual is observed over five days where each day is defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day.
+We define~$X(t)$ at the grid of observations as a mean-zero Gaussian process with covariance
 \[
@@
 where~$K_v$ is the modified Bessel function of the second kind.
-We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.5$ as well as set~$K_b = K_x = 35$.
-For simplicity, we assume~$\Sigma$ is known in computation of the the eigendecomposisions.
-Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; theta)$.
+We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.3$ as well as set~$K_b = K_x = 35$.
+For simplicity, we assume~$\Sigma$ is known in computation of the eigendecompositions.
+Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; \theta)$.
 To mimic our real data, we set
@@
 \begin{tabular}{c | c c c c c}
-Sampling rate & 0.5 & 1 & 5 & 25 & 50 \\
-Variance
-Squared Bias
-MISE
+Sampling rate & 0.5 & 1 & 5 & 25 & 50 \\ \hline
+Variance & (10,12) & - & - & - & - \\
+Squared Bias & - & - & - & - & -  \\
+MISE & - & - & - & - & -  \\ \hline
+Type I error & - & - & - & - & -  \\ \hline
+\end{tabular}
+\caption{Mean-integrated squared error and type I error as a function of sampling rate for $\beta(s)$ given by (1) and (2) respectively.}
+\label{tab:mise}
+\end{table}
+
+\subsection{Coverage properties}
+
+We simulate the system $1000$ times under $\beta \equiv 0$ and check the type I error.
+
+\subsection{Impact of $\Delta$}
+
+Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates.
+Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.
+For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$.
+If so, we can set~$\Delta$ larger to ensure the condition holds.
+In each simulation, we use $\Delta \in (18,21,24,27,30)$.
+
+\begin{table}[!th]
+\begin{tabular}{c c c c c c}
+& \multicolumn{5}{c}{Sampling rate} \\ \cline{2-6}
+$\Delta=$ & 0.5 & 1 & 5 & 25 & 50 \\ \hline
+18  & - & - & - & - & -  \\
+21 & - & - & - & - & -  \\
+24 & - & - & - & - & -  \\
+27 & - & - & - & - & -  \\
+30 & - & - & - & - & -  \\ \hline
 \end{tabular}
@@
 \end{table}
-
-\subsection{Coverage properties}
-
-In each (A)-(C) we parametrize the curves and check the type I error
-for various sizes of the
-
-\subsection{Impact of $\Delta$}
-
-Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates.
-Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.
-For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$.
-If so, we can set~$\Delta$ larger to ensure the condition holds.
 
@@
 
+\subsection{Online estimation}
+
 \section{A worked example: Adolescent psychiatric inpatient mHealth study} \label{section:example}
 
-During a XX month period in 2019, 100 psychiatric inpatients admitted
-for suicidal risk to Franciscan Children's Hospital
-
-hospitals with histologically verified liver cirrhosis were recu
+During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital  were enrolled in
+ with histologically verified liver cirrhosis were recu
 those presenting to a psychiatric hospital with suicide ideation
@@
 analysis.
-The goal of the study was to ask whether sensor information can used
-to identify distress early prior in the process of escalation by
-monitoring physiological correlates of distress -- in particular,
-electrodermal activity (EDA), a measure of skin conductance, and
-``gACC'', a coordinate-free feature built from accelerometer data
-that measures physical movement.
-The scientific hypothesis to be tested is whether a spike in
-EDA and/or gACC
+One of the goals of the study was to assess the association between sensor-based covariate information can used to identify distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement.
+The scientific hypothesis to be tested is whether a spike in EDA and/or gACC
 
@@
 for distinct patients are independent Gaussian with moments
+
+{\bf comparison over $\Delta$ and using imputation}
 
@@
 \[
-\left | n^{-1}  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau H_{in} (t) dN_i^
+\left | n^{-1}  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau H_{in} (t) dN_i
 \]

### 2020-01-27 simulations/basic_simulation.R

@@ -37,7 +37,7 @@ eig_Sigma = eigen(Sigma)
 K_x = 35  # Pick first 35 eigen-vectors
 cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
 
-set.seed("131232")
+set.seed("134163")
 dataset = generate_complete_data(N = 1000, Cov_X, C, times, eig_Sigma, beta_1t)
 rm(abs_diff, C, Cov_X, x)
 
@@ -78,7 +78,7 @@ K <- 100
 lambdapath <- round(exp(seq(log(lambda_max), log(lambda_max*epsilon), 
                             length.out = K)), digits = 10)
 
-set.seed("7918731")
+set.seed("97139817")
 ridge.fit.cv <- cv.glmnet(w, dataset$Y, alpha = 0, intercept = TRUE, 
                           penalty.factor = p.fac, standardize = FALSE,
                           lambda = lambdapath, nfolds = 20,
@@ -97,7 +97,8 @@ plot(max(local_times)-local_times, betaHat.net, col = "blue", type = "l",
 axis(side = 1); axis(side = 2, labels = FALSE)
 # lines(local_times, betaHat.net.star, col = "red")
 lines(local_times, beta_1t)
-mtext("Time until event", side = 1, line = 1)
+mtext("Time until event", side = 1, line = 2)
 mtext(expression(paste(beta, "(s)")),side = 2, line = 1)
 
+mise_calc(beta_1t, betaHat.net)
 

### 2020-01-27 simulations/basic_simulation_functions.R

@@ -74,7 +74,8 @@ expit <- function(x) {exp(x)/(1+exp(x))}
 
 logit <- function(p) {log(p/(1-p))}
 
-mise_calc <- function(beta_1t, betaHat.net.list) {
-  temp_Xb = (beta_1t - betaHat.net.list[i] - mean(betaHat.net.list))^2
+mise_calc <- function(beta_1t, betaHat.net) {
+  temp_Xb = (beta_1t - betaHat.net)^2
   term2 = (temp_Xb[1] + temp_Xb[length(temp_Xb)])/2 * gap + sum(temp_Xb[2:(length(temp_Xb)-1)]*gap) 
+  return(term2)
 }

### 2020-02-26 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 Simulations demonstrate the method and efficiency-computation trade-off.
-We then illustrate the approach using data from a digital monitoring study of suicidal ideation.
+We end by illustrating our approach using data from a digital monitoring study of suicidal ideation.
 \end{abstract}
@@
 In particular, a higher sampling rate can decrease estimator variance at the cost of increased computation.
-We show via simulations that the benefits of incredibly high sampling rates is often negligible, as the contribution to the variation is small in relation to to variation in the underlying stochastic processes.
-Second, derived estimating equations are optimal, implying loss of statistical efficiency is only due to subsampling procedure and not derived methodology.
+We show via simulations that the benefits of incredibly high sampling rates is often negligible, as the contribution to the variation is small relative to the variation in the underlying stochastic processes.
+Second, derived estimating equations are optimal, implying loss of statistical efficiency is only due to subsampling procedure and not the derived methodology.
 Finally, implementation can leverage existing, standard software for functional data analysis and logistic regression, leading to fast adoption by domain scientists.
@@
 where~$\theta$ is the parameter vector.
-Equation~\eqref{eq:hazard} states the hazard may depend on past values of the event history and covariates.
 For high frequency physiological data, we assume that current risk is log-additive and depends on a linear functional of the health process over some recent window of time and some pre-specified features of the counting process; that is,
@@
 This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process.
-To ensure independent evolution, we employ the strategy of setting~$\epsilon (t) \equiv 0$ and treating~$\bfx_i$ as a mean-zero Gaussian process.
+To ensure independent evolution, we employ the strategy of setting~$\epsilon (t) \equiv 0$.and treating~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
 The proposed inferential procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed (see Section~\ref{section:missingdata}).
@@
     (T_{i,j}; \theta)}{h_i (T_{i,j}; \theta)} - H^{(1)}_{i} (\tau_i;
-  \theta) \right)
-\]
-where~$h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (\tau_{i}; \theta)$ are derivatives with respect to $\theta$.
+  \theta) \right),
+\]
+with $h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (\tau_{i}; \theta)$ are derivatives with respect to $\theta$.
 
@@
 \end{equation}
-the resulting estimating equations can be written as
+the resulting approximate estimating equations can be re-written as
 \begin{equation}
@@
 \hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
-  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in D_i} w_i(u;
+  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
   \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
 \end{equation}
-Equation~\eqref{eq:approxscore} are approximate score functions built via plug-in of the design-unbiased estimator of the derivative of the cumulative hazard given in~\eqref{eq:WPest}.
+Equation~\eqref{eq:approxscore} represents the approximate score functions built via plug-in of the design-unbiased estimator of the derivative of the cumulative hazard given in~\eqref{eq:WPest}.
 
@@
 We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
-We use the sandwich smoother~\citep{Xiao2013} to estimate the mean~$\mu_y(s,t) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
+We use the sandwich smoother~\citep{Xiao2013} to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
 Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009} bivariate tensor product splines~\citep{Wood2006} and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
 The sandwich smoother was chosen for its computational efficiency and estimation accuracy.
-We then define~$\tilde X(s,t) = X(s,t) - \hat \mu_y(s,t)$ to be the mean-zero process at each time~$t \in \bfT_i \cup \bfD_i$.
+We then define~$\tilde X(t,s) = X(t,s) - \hat \mu_y(t,s)$ to be the mean-zero process at each time~$t \in \bfT_i \cup \bfD_i$.
 
@@
 \[
-\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (s,T), (s^\prime, T) ) f_y(T) dt.
-\]
-for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((s,t), (s^\prime,t))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $f_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
+\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (T,s), (T,s^\prime) ) f_y(T) dt.
+\]
+for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((t,s), (t,s^\prime))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $f_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
 Estimation of~$\Sigma_y$ occurs in two steps.
@@
 \[
-\tilde \Sigma_0 (s_r , s_{r^\prime}) = \left( \sum_{i=1}^n \left | \bfD_i \right |\right)^{-1} \left( \sum_{i=1}^n \sum_{t \in \bfD_i} \tilde X (s_r,t) \tilde X (s_{r^\prime}, t) \right).
+\tilde \Sigma_0 (s_r , s_{r^\prime}) = \left( \sum_{i=1}^n \left | \bfD_i \right |\right)^{-1} \left( \sum_{i=1}^n \sum_{t \in \bfD_i} \tilde X (t,s_r) \tilde X (t,s_{r^\prime}) \right).
 \]
@@
 \[
-X(s,t) = \hat \mu_y (s,t) + \sum_{k=1}^{\infty} \hat c^{(y)}_{i,k} (t) \hat
-\psi^{(y)}_k (s) \approx \hat \mu_y (s,t) + {\bf c}^{(y)}_{i} (t)^\top \mathbold{\hat
+X(t,s) = \hat \mu_y (t,s) + \sum_{k=1}^{\infty} \hat c^{(y)}_{i,k} (t) \hat
+\psi^{(y)}_k (s) \approx \hat \mu_y (t,s) + {\bf c}^{(y)}_{i} (t)^\top \mathbold{\hat
   \psi}^{(y)} (s)

### 2020-02-26 write-up/fda-recurrent.tex
Edited up to the theory, need to finish proofs and then move to sims and real data one last time

@@
 \]
-where $\hat c^{(y)}_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (s,t) \hat \psi^{(y)}_k (s) ds$, ${\bf c}^{(y)}_i (t) = (c^{(y)}_{i,1} (t), \ldots, c^{(y)}_{i,K_x} (t))^\top$, $\mathbold{\hat \psi}^{(y)} (s) = (\hat \psi^{(y)}_1 (s), \ldots, \hat \psi^{(y)}_{K_x} (s))^\top$, and~$K_x < \infty$ is the truncation level of the infinite expansion.
+where $\hat c^{(y)}_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (t,s) \hat \psi^{(y)}_k (s) ds$, ${\bf c}^{(y)}_i (t) = (c^{(y)}_{i,1} (t), \ldots, c^{(y)}_{i,K_x} (t))^\top$, $\mathbold{\hat \psi}^{(y)} (s) = (\hat \psi^{(y)}_1 (s), \ldots, \hat \psi^{(y)}_{K_x} (s))^\top$, and~$K_x < \infty$ is the truncation level of the infinite expansion.
 Following~\cite{Goldsmith2011}, we set~$K_x$ to satisfy identifiability constraints (see Section~\ref{section:beta} for details).
@@
 \begin{align*}
-\int_{t-\Delta}^t X(s,t) \beta(s) ds
-  &\approx \int_{t-\Delta}^t \left[ \hat \mu(s,t) + {\bf c} (t)^\top
+\int_{t-\Delta}^t X(t,s) \beta(s) ds
+  &\approx \int_{t-\Delta}^t \left[ \hat \mu(t,s) + {\bf c} (t)^\top
     \mathbold{\hat \psi} (s) \right] \times \left[
-    \mathbold{\phi} (s) {\bf b} \right] \\
-  &= [ ( M_{i,t} )^\top + (C_{i,t})^\top J_{\hat \psi, \phi} ] {\bf b}
+    \mathbold{\phi} (s) {\bf b} \right] ds \\
+  &= [ M_{i,t}^\top + {\bf c} (t)^\top J_{\hat \psi, \phi} ] {\bf b}
 \end{align*}
-where~$M_{i,t} = (M_{i,1,t}, \ldots, M_{i,K_b,t})$, $M_{i,j,t} = \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\hat \psi, \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s) ds$~\citep{RamsaySilverman2005}.
+where~$M_{t} = (M_{1,t}, \ldots, M_{K_b,t})$, $M_{j,t} = \int_{t-\Delta}^t \hat \mu (t,s) \phi_j (s)$, and~$J_{\hat \psi, \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s) ds$~\citep{RamsaySilverman2005}.
 
@@
 \[
-\sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i } \left[ {\bf 1}[t \in D_i]
+\sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i } \left[ {\bf 1}[t \in D_i]
   - \frac{1}{1 + \exp \left[- \left( \tilde{W}_{i,t}^\top \theta +
@@
 \item For each $i \in [n]$, sample non-event times as a time-inhomogeneous Poisson point process with intensity according to~$\pi_i (t)$
-\item Estimate mean~$\mu_y (t,s)$ for $0 \leq s \leq \Delta$ at all event times~$t \in \cup_{i=1}^n \bfT_i$ and sampled non-event times~$t \in \cup_{i=1}^n \bfD_i$.
-\item Compute marginal covariance across event times,~$\Sigma_1$, and non-event times and~$\Sigma_0$.
-\item Compute eigendecomposition~$\{ \hat \psi_k^{(y)}, \hat \lambda_k^{()} \}$ of marginal covariance~$\Sigma_y$
+\item \label{p2} Estimate mean~$\mu_y (t,s)$ for $0 \leq s \leq \Delta$ at all event times~$t \in \cup_{i=1}^n \bfT_i$ and sampled non-event times~$t \in \cup_{i=1}^n \bfD_i$.
+\item \label{p3} Compute marginal covariance across event times,~$\Sigma_1$, and non-event times,~$\Sigma_0$.
+\item \label{p4} Compute eigendecomposition~$\{ \hat \psi_k^{(y)}, \hat \lambda_k^{()} \}$ of marginal covariance~$\Sigma_y$
 \item Use the eigendecomposition to construct~$W_{i,t}$ for all $i \in [n]$ and $t \in \bfD_i \cup \bfT_i$
@@
 
-We also require several assumptions due to the truncation of the Karhunen-Lo{\`e}ve decomposition that represents~$X(s,t)$.
+We also require several assumptions due to the truncation of the Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
 \begin{assumption}[Functional assumptions~\citep{Park2018}] \normalfont
@@
 \begin{enumerate}[label=(A.\arabic*)]
-\item\label{A1} $X = \{ X(s, t) : (s,t) \in \mathcal{S} \times
-  \mathcal{T} \}$ is a square integrable element of the $L^2 (
-  \mathcal{S} \times \mathcal{T})$
+\item\label{A1} $X = \{ X(t, s) : (t,s) \in \mathcal{T} \times
+  \mathcal{S} \}$ is a square integrable element of the $L^2 (
+  \mathcal{T} \times \mathcal{S})$
 \item\label{A2} The subsampling and conditional intensity rate functions~$f_y(T)$ are continuous and~$\sup |f_y(T)| < \infty$.
-\item\label{A3} $\E[X(s,t) X(s^\prime, t) X(s,t^\prime)
-  X(s^\prime, t^\prime) ] < \infty$ for
-  each $s,s^\prime \in [0,\Delta]$ and~$0 < t, t^\prime < \tau$.
-\item\label{A4} $\E[\|X(\cdot, t)\|^4] < \infty$ for each~$0< t < \tau$.
+\item\label{A3} {\color{red}$\E[X(t,s) X(t,s^\prime) X(t^\prime,s)
+  X(s^\prime, t^\prime) ] < \infty$ for each $s,s^\prime \in [0,\Delta]$ and~$0 < t, t^\prime < \tau$.}
+\item\label{A4} $\E[\|X(t,\cdot)\|^4] < \infty$ for each~$0< t < \tau$.
 \end{enumerate}
 \end{assumption}
-Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
-With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
-
-Assumption~\ref{A3} is guaranteed by~\ref{E1} and~\ref{E3}, but is stated here to emphasize its importance for the marginal covariance procedure of~\cite{Park2018}.
 Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.

### 2020-02-27 write-up/fda-recurrent.tex
small update, switching comps
Updated paper, finish CLT and the do optimal

@@
 \usepackage{multirow}
-\usepackage{enumitem}
+\usepackage[shortlabels]{enumitem}
 \def\pr{\mathop{\text{pr}}\nolimits}
@@
 Given observed recurrent event and high frequency data~$\{ \bfT_i, \bfx_i \}_{i=1}^n$,
-\begin{enumerate}
+\begin{enumerate}[(a)]
 \item For each $i \in [n]$, sample non-event times as a time-inhomogeneous Poisson point process with intensity according to~$\pi_i (t)$
@@
   \[
-    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
-    \theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
-        \theta)}{h_i (t; \theta)} \right] \times  \left [
-      \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
-    dt \overset{P}{\to} \Xi (\theta).
+    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t; \theta) \times \left[ \frac{h_i^{(1)}(t; \theta) h_i^{(1)} (t;\theta)^\top}{h_i (t; \theta)} \right] \times   R_i (t) dt \overset{P}{\to} \Xi (\theta).
   \]
@@
 \item\label{A2} The subsampling and conditional intensity rate functions~$f_y(T)$ are continuous and~$\sup |f_y(T)| < \infty$.
-\item\label{A3} {\color{red}$\E[X(t,s) X(t,s^\prime) X(t^\prime,s)
-  X(s^\prime, t^\prime) ] < \infty$ for each $s,s^\prime \in [0,\Delta]$ and~$0 < t, t^\prime < \tau$.}
+\item\label{A3} $\E[X(t,s) X(t,s^\prime) X(t^\prime,s) X(t^\prime, s^\prime) ] < \infty$ for each $s,s^\prime \in [0,\Delta]$ and~$0 < t, t^\prime < \tau$.
 \item\label{A4} $\E[\|X(t,\cdot)\|^4] < \infty$ for each~$0< t < \tau$.
@@
 \[
-  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \left[ \frac{h^{(1)}(s;
+  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \times \left[ \frac{h^{(1)}(s;
       \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
@@
 
-\subsection{Computation versus statistical efficiency tradeoff}
-
-Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
+% \subsection{Computation versus statistical efficiency tradeoff}
+
+% Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
 While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
@@
 
-For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
+% For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
 This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
@@
 
-\begin{prop} \normalfont
-\label{prop:optimal}
-If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
+% \begin{prop} \normalfont
+% \label{prop:optimal}
+% If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
 This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
-\end{prop}
-
-\noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
+% \end{prop}
+
+% \noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
 
@@
 That is, the subsampling rate is proportional to the intensity function with time-independent constant~$c > 0$.
-Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
+Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c/ (c+1)$.
 Under Lemma~\ref{lemma:simpleasym},
@@
 
-Recall sensor measurements occur on the order of multiple times per second.
-Suppose the intensity rate is bounded above by $3$ and the unit time scale was hours.
-If we can subsample the data at~$30$ times per hour, then we have a lower bound on the efficiency of $0.909$.
-For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
-While the computational complexity of logistic regression is linear in the number of samples, we get $480$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
-If we sample~$300$ times per hour, then the efficiency loss is only $0.999$, with a $48$ times reduction in data size.
-Table~\ref{tab:compvseff} provides additional examples for a 4Hz sensor.
-The data reduction depends on the sensor measurement rate; however, the lower bound on statistical efficiency does not.
-This is because the subsampling rate only depends on the upper bound of the intensity function.
-That is, if the events are rare then the amount of data we need to subsample is greatly reduced with no impact to statistical efficiency.
-
-Recall sensor measurements occur on the order of multiple times per second.
-Suppose the intensity rate is bounded above by $3$ and the unit time scale was hours.
-If we can subsample the data at~$30$ times per hour, then we have a lower bound on the efficiency of $0.909$.
-For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $30$.
-While the computational complexity of logistic regression is linear in the number of samples, we get $480$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
-If we sample~$300$ times per hour, then the efficiency loss is only $0.999$, with a $48$ times reduction in data size.
-Table~\ref{tab:compvseff} provides additional examples for a 4Hz sensor.
-The data reduction depends on the sensing rate; however, the lower bound on statistical efficiency does not.
-This is because the subsampling rate only depends on the upper bound of the intensity function.
-That is, if the events are rare then the amount of data we need to subsample is greatly reduced with no impact to statistical efficiency.
+Sensor measurements occur multiple times per second.
+If the intensity rate is bounded above by $1$ and the unit time scale was hours.
+If we can subsample the data at~$10$ times per hour, then we have a lower bound on the efficiency of $0.909$.
+For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $10$.
+While the computational complexity of logistic regression is linear in the number of samples, we get $1440$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
+If we sample~$100$ times per hour, then the efficiency loss is only $0.999$, with a $144$ times reduction in data size.
+Table~\ref{tab:compvseff} provides additional examples for a 4Hz and 32Hz sensor rate respectively.
+The data reduction depends on this rate; however, the lower bound on statistical efficiency does not because the subsampling rate only depends on the upper bound of the intensity function.
+In particular, if the events are rare then subsampling rate can be greatly reduced with no impact to statistical efficiency.
 
@@
 \centering
-\begin{tabular}{l r r r r r | c}
+\begin{tabular}{l l r r r r r | c}
+\multirow{2}{2.5cm}{Sensor rate} &
 \multirow{2}{2.5cm}{Subsampling constant ($c$)}
@@
     hour}
-  & \multirow{2}{2cm}{Statistical efficiency}\\ \cline{2-6}
-& 0.5 & 1 & 3 & 5 & 10 \\ \hline
-1 & 28800.0 & 14400.0 & 4800.0 & 2880.0 & 1440.0 & 0.500 \\
-5 & 5760.0 & 2880.0 & 960.0 & 576.0 & 288.0 & 0.833 \\
-10 & 2880.0 & 1440.0 & 480.0 & 288.0 & 144.0 & 0.909 \\
-100 & 288.0 & 144.0 & 48.0 & 28.8 & 14.4 & 0.990 \\
-500 & 57.6 & 28.8 & 9.6 & 5.8 & 2.9 & 0.999 \\ \hline
+  & \multirow{2}{2cm}{Statistical efficiency}\\ \cline{3-7}
+& & 0.5 & 1 & 3 & 5 & 10 \\ \hline
+% 1 & 28800.0 & 14400.0 & 4800.0 & 2880.0 & 1440.0 & 0.500 \\
+\multirow{3}{*}{4Hz (EDA)}
+& 5 & 5760 & 2880 & 960 & 576 & 288 & 0.833 \\
+& 10 & 2880 & 1440 & 480 & 288 & 144 & 0.909 \\
+& 100 & 288 & 144 & 48 & 29 & 14 & 0.990 \\ \hline
+\multirow{3}{*}{32Hz (ACC)}
+& 5 & 46080 & 23040 & 7680 & 4608 & 2304 & 0.833 \\
+& 10 & 23040 & 11520 & 3840 & 2304 & 1152 & 0.909 \\
+& 100 & 2304 & 1152 & 384 & 230 & 115 & 0.990 \\ \hline
+% 500 & 57.6 & 28.8 & 9.6 & 5.8 & 2.9 & 0.999 \\ \hline
 \end{tabular}
-\caption{Data reduction (total number of measurements divided by expected number of subsampled measurements) given 4Hz sensor, subsampling constant~$c$ and an upper bound on the intensity rate per hour~$H$.
-The subsampling rate is set at~$c \times H$; note, lower bound on statistical efficiency only depends on the subsampling rate.}
+\caption{Data reduction (total \# of measurements divided by expected number of subsampled measurements) given sensor rate, subsampling constant and an upper bound on the intensity rate.}
 \label{tab:compvseff}
@@
 That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ dimensional matrix obtained by plugging in the estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-
-Recall theoretical results were proven under the assumption that there exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
-To make this assumption plausible, we set~$K_b$ large enough (but less than~$K_x$ to ensure identifiability) such that the spline basis expansion is sufficiently expressive.
-However, in practice, such a choice of~$K_b$ may lead to overfitting the data.
-We therefore, in the remainder of this paper, follow~\cite{GoldSmith2015} and induce smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2 I)$.
-Incorporating penalization is simple due to the connection between with the score functions for logistic regression.
-
-\subsection{Confidence intervals for~$\beta(t)$}
-
-Due to the connection with generalized mixed effects models, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
-That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional matrix obtained by plugging in the REML estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
 Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
@@
 
-In this section, we demonstrate the flexibility of our approach by exploring extensions in several important directions to ensure these methods are robust for practical use.
+In this section, we demonstrate the flexibility of our approach by exploring extensions in several important directions to ensure these methods are robust for practical use in mobile health.
 This section will continue to leverage the connection to generalized functional linear models provided by Lemma~\ref{lemma:logistic}.
@@
 
-\newpage
-
 \subsection{Multilevel models}
 
-\subsection{Multiple sensors and availability}
-
-\subsection{Online estimation}
+\subsection{Multiple sensors}
 
@@
 
+\newpage
 \appendix
@@
 \]
-where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t-s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
+where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t,s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
 \cite{Andersen1993} (Theorem VI.1.1) show that~$n^{-1} U_n (\theta_0)
@@
   \times \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
-  \left( X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
+  \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
   \phi_l (s) b_l\right \| \\
 &\to \frac{M}{4} \int_0^\tau \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
-  \left( X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
+  \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
   \phi_l (s) b_l\right \| (f_0(t) + f_1(t)) dt
@@
 The second inequality is due to the Taylor remainder theorem.
-The third is due to~$g^\prime (x) = e^x/(1+e^x)^2 \leq e^0/(1+e^0)^2 =
-1/4$, reordering the summation, and rewriting the error term in terms
-of the difference between~$X(s,t)$ and the approximation with
-truncation level~$K_x$. Letting~$n \to \infty$ allows us to re-write
-the outside sums in terms of the integrated difference where the
-integral is with respect to the joint event and sampling distributions~$(f_0(t) + f_1 (t))$. Finally, under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
-\[
-X(s,t) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \overset{p}{\to} 0
+The third is due to~$g^\prime (x) = e^x/(1+e^x)^2 \leq e^0/(1+e^0)^2 = 1/4$, reordering the summation, and rewriting the error term in terms of the difference between~$X(t,s)$ and the approximation with truncation level~$K_x$.
+Letting~$n \to \infty$ allows us to re-write the outside sums in terms of the integrated difference where the integral is with respect to the joint event and sampling distributions~$(f_0(t) + f_1 (t))$.
+Finally, under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
+\[
+X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \overset{P}{\to} 0
 \]

### 2020-02-27 write-up/fda-recurrent.tex
Fixing up consistency and asymptotic nromality proofs

@@
 
-To prove~\eqref{eq:andersen3},
-\[
-\left | n^{-1}  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau H_{in} (t) dN_i
-\]
+To prove~\eqref{eq:andersen3}, we start with equation~\eqref{eq:approxscore}, which we rewrite here for completeness:
+\[
+\hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
+  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
+  \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
+\]
+Let~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ be the event and subsampling counting processes with jumps at $t \in \bfT_i$ and $\bfD_i$ respectively.  Then
+\begin{equation}
+\label{eqapp:upperbound}
+\left | n^{-1} \frac{d^2}{d \theta_j d \theta_k} \tilde U_n (\theta)  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau w_i (t;\theta) \times H_{in} (t) R_i (t) dN_i^{(e)} (t) +
+\frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{w_i (t;\theta)}{\pi_i (t)} \times G_{in} (t) R_i (t) dN_i^{(e)} (t)
+\end{equation}
+where $H_{in}(t)$ and $G_{in}(t)$ are from Condition VI.1.1(E) in~\cite[pp.
+421]{Andersen1993}.
+The first term converges in probability to a finite quantity by~\cite{Andersen1993}.
+Since~$w_i(t;\theta)/\pi_i (t) = (\pi_i(t)+h_i(t;\theta))^{-1}$, the second term which simplifies to
+\[
+\frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{G_{in} (t)}{\pi_i (t)+h_i(t;\theta)} R_i(t) dN_i^{(e)} (t)
+\leq
+\frac{1}{L \cdot n} \sum_{i=1}^n \int_0^\tau G_{in} (t) R_i (t) dN_i^{(e)} (t)
+\]
+by Assumption~\ref{E1} (i.e., the subsampling rate is lower bounded by $L$). The right hand side is the optional variation of the local square integrable martingale
+$$
+\frac{1}{L \cdot n} \sum_{i=1}^n \int_0^t G^{1/2}_{in} (s) dN_i^{(e)} (s)
+$$
+This martingale has predicable variation process
+$$
+\frac{1}{L \cdot n} \sum_{i=1}^n \int_0^t G_{in} (s) \pi_i (s) R_i (s) ds
+$$
+Since both processes have the same limits, conditions VI.1.1(E) in~\cite{Andersen1993} and Assumption~\ref{E1} imply that the second term on the right hand side of~\eqref{eqapp:upperbound} also converges to a finite quantity as $n\ to \infty$ for every $K_x$ which completes the proof.
+
+This proves~$\hat \theta_n \overset{P}{\to} \theta_0$ as $n \to \infty$ and $K_x \to \infty$.
+Note that the argument holds for any choice of weight functions~$w_i (t;\theta)$; implying convergence holds more broadly for all approximate score equations in~\eqref{eq:approxscore}.
+\end{proof}
+
+\begin{proof}[Proof of Asymptotic Normality]
+Consider the weighted estimating equations given by~\eqref{eq:approxscore}. We have shown that afjlakfjlakfj
+Then
 \end{proof}

### 2020-02-27 write-up/fda-recurrent.tex
Doing asymp normality now

@@
 three terms
-\[
+\begin{equation}
+\label{eq:decomp}
 n^{-1} \tilde U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
@@
   (\theta_0) - \hat U_n (\theta_0) \right)
-\]
+\end{equation}
 where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t,s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
@@
 \begin{proof}[Proof of Asymptotic Normality]
-Consider the weighted estimating equations given by~\eqref{eq:approxscore}. We have shown that afjlakfjlakfj
-Then
+Consider a Taylor-expansion of $\tilde U_n (\theta)$ centered at $\theta_0$ when $\hat \theta_n \in \Theta_0$.  We can write as
+\begin{align*}
+\tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
+0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + (\hat \theta_n - \theta_0)^2 \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
+\end{align*}
+where~$\theta^\star$ s on the line segment between $\hat \theta_n$ and $\theta_0$.
+
+Now $\tilde U_n (\theta_0 )$ can be decomposed into three terms as in~\eqref{eq:decomp}. Here, we assume $K_x \to \infty$ so the only two terms are
+$$
+U_n (\theta_0) = \sum_{i=1}^n
+$$ and
+$$
+n^{-1} (\hat U_n - U_n (\theta_0) ) = \frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s)
+$$
 \end{proof}

### 2020-02-28 write-up/fda-recurrent.tex
Realized not saving freq caused issues
Finished inner part

--- 2020-02-27 write-up/fda-recurrent.tex
@@
 % Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
-We define the joint counting process
+Define the joint counting process
 \[
@@
 $$
-Since both processes have the same limits, conditions VI.1.1(E) in~\cite{Andersen1993} and Assumption~\ref{E1} imply that the second term on the right hand side of~\eqref{eqapp:upperbound} also converges to a finite quantity as $n\ to \infty$ for every $K_x$ which completes the proof.
+Since both processes have the same limits, conditions VI.1.1(E) in~\cite{Andersen1993} and Assumption~\ref{E1} imply that the second term on the right hand side of~\eqref{eqapp:upperbound} also converges to a finite quantity as $n \to \infty$ for every $K_x$ which completes the proof.
 
@@
 \begin{proof}[Proof of Asymptotic Normality]
-Consider the weighted estimating equations given by~\eqref{eq:approxscore}. We have shown that afjlakfjlakfj
-Then
+Consider a Taylor-expansion of $\tilde U_n (\theta)$ centered at $\theta_0$ when $\hat \theta_n \in \Theta_0$.  We can write as
+\begin{align*}
+\tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
+0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
+n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
+\end{align*}
+where~$\theta^\star$ is on the line segment between $\hat \theta_n$ and $\theta_0$. First, the term
+$$
+\frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
+\overset{P}{\to} \Xi(\theta_0)
+$$
+under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}.
+To see this, note the first term converges to $\Xi(\theta_0)$ while the second term is bounded in probability by $p M \| \theta^\star - \theta_0 \|$ for some finite constant~$M$ independent of~$\theta^\star$ where $\| \cdot \|$ is the Euclidean norm.
+This goes to zero since $\| \theta^\star - \theta_0 \| \to 0$.
+
+Now $\tilde U_n (\theta_0 )$ can be decomposed into three terms as in~\eqref{eq:decomp}. Here, we assume $K_x \to \infty$ so the final term is negligible.  This leaves two terms
+$$
+U_n (\theta_0) = \sum_{i=1}^n
+$$ and
+$$
+n^{-1} (\hat U_n (\theta_0) - U_n (\theta_0) ) = \frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s)
+$$
 \end{proof}

### 2020-02-28 write-up/fda-recurrent.tex
Finished proof of asymp normality

@@
 \tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
-0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
-n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
+0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
+n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
 \end{align*}
@@
 $$
-\frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
+\frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
 \overset{P}{\to} \Xi(\theta_0)
@@
 To see this, note the first term converges to $\Xi(\theta_0)$ while the second term is bounded in probability by $p M \| \theta^\star - \theta_0 \|$ for some finite constant~$M$ independent of~$\theta^\star$ where $\| \cdot \|$ is the Euclidean norm.
-This goes to zero since $\| \theta^\star - \theta_0 \| \to 0$.
-
-Now $\tilde U_n (\theta_0 )$ can be decomposed into three terms as in~\eqref{eq:decomp}. Here, we assume $K_x \to \infty$ so the final term is negligible.  This leaves two terms
+The second term therefore goes to zero since $\| \theta^\star - \theta_0 \| \to 0$ as $n \to \infty$.
+
+We now consider $n^{-1/2} \tilde U_n (\theta_0 )$.
+Recall $\tilde U_n (\theta_0)$ can be decomposed into three terms as in~\eqref{eq:decomp}.
+Here, we assume $K_x \to \infty$ so the final term is negligible.
+This leaves two terms
 $$
-U_n (\theta_0) = \sum_{i=1}^n
-$$ and
+U_n (\theta_0) = \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(u; \theta)}{h (u;\theta)} dM_i^{(e)} (t)
 $$
-n^{-1} (\hat U_n (\theta_0) - U_n (\theta_0) ) = \frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s)
+where~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ are the orthogonal event and subsampling counting processes with jumps at times $t \in \bfT_i$ and $t \in \bfD_i$ respectively.
+and
+\begin{align*}
+(\hat U_n (\theta_0) - U_n (\theta_0) )
+&= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s) \\
+&= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM^{(e)}_i (s) +
+\sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i^{(s)} (s) \\
+&= \sum_{i=1}^n \int_0^\tau (1-w(s;\theta_0)) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s) +
+\sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} dM_i^{(s)} (s)
+\end{align*}
+Taking difference with $U_n(\theta_0)$, the first term becomes
+\[
+\sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s)
+\]
+
+These are two orthogonal square integrable martingales with quadratic variation given by
+\[
+\sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds
+\]
+and
+\[
+\sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds
+\]
+respectively.  Using the definition $w(s;\theta_0) = \pi_i(s) / (\pi_i(s)+h_i(s;\theta_0))$, we have
+\begin{align*}
+&\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds \\
+&+
+\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds \\
+=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ 1 + \frac{h(s;\theta_0)}{\pi_i (s)} \right] ds \\
+=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ \frac{\pi_i (s) + h(s;\theta_0)}{\pi_i (s)} \right] ds \\
+=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} ds
+\to \Xi (\theta_0).
+\end{align*}
+Returning to the Taylor-expansion, under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}, we can apply Rebolledo's martingale central limit theorem so that
 $$
+\sqrt{n} (\hat \theta_n - \theta_0) \overset{D}{\to} N(0, \Xi(\theta_0) ^{-1} \times \Xi(\theta_0) \times \Xi(\theta_0) ^{-1} )
+$$
+which completes the proof.
 \end{proof}
@@
 \begin{proof}[Proof of Proposition~\ref{prop:optimal}]
-Consider the weighted estimating equations given by~\eqref{eq:approxscore}.
-Then
+The above proof establishes asymptotic normality for all
+The main additional assumption is
+
+The argument
+
 \end{proof}

### 2020-03-21 write-up/fda-recurrent.tex
update sims and tex to match

@@
 \begin{tabular}{c | c c c c c}
-Sampling rate & 0.5 & 1 & 5 & 25 & 50 \\ \hline
+& \multicolumn{5}{c}{Sampling rate} \\ \cline{2-6}
+ & 0.5 & 1 & 5 & 25 & 50 \\ \hline
+Subsampling variance & (10,12) & - & - & - & - \\
 Variance & (10,12) & - & - & - & - \\
@@
 Type I error & - & - & - & - & -  \\ \hline
+Avg runtime & - & - & - & - & - \\
 \end{tabular}

### 2020-03-21 simulations/basic_simulation.R

@@ -37,7 +37,8 @@ eig_Sigma = eigen(Sigma)
 K_x = 35  # Pick first 35 eigen-vectors
 cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
 
-set.seed("134163")
+# set.seed("134163")
+set.seed("131312")
 base_num_events = 5
 base_rate = logit(5/length(times))
 sampling_rate = 10/1000
@@ -82,12 +83,18 @@ lambdapath <- round(exp(seq(log(lambda_max), log(lambda_max*epsilon),
                             length.out = K)), digits = 10)
 
 set.seed("97139817")
+# Start the clock!
+ptm <- proc.time()
+
 ridge.fit.cv <- cv.glmnet(w, dataset$Y, alpha = 0, intercept = TRUE, 
                           penalty.factor = p.fac, standardize = FALSE,
                           lambda = lambdapath, nfolds = 20,
                           family = "binomial")
+# Stop the clock
+proc.time() - ptm
+
 ridge.fit.lambda <- ridge.fit.cv$lambda.min
-plot(ridge.fit.cv)
+# plot(ridge.fit.cv)
 
 # Extract coefficient values for lambda.1se (without intercept)
 ridge.coef <- (coef(ridge.fit.cv, s = ridge.fit.lambda))[-1]
@@ -104,4 +111,3 @@ mtext("Time until event", side = 1, line = 2)
 mtext(expression(paste(beta, "(s)")),side = 2, line = 1)
 
 mise_calc(beta_1t, betaHat.net)
-

### 2020-05-19 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 
+\section{Penalized mixed effects logistic model}
+
+Here we derive a penalized adaptive Gaussian quadrature (PADQ) to fit the   penalized logistic mixed-effects model.
+Here, for clarity, we write $Y_{ij} \ in \{ 0, 1 \}$ to be the sequence of binary indicators for participant $i = 1,\ldots,n$ and $j$ indexes the event and subsampled times.
+We write $W_{ij}$ to denote the time-varying covariate.
+Then the score equations for the penalized logistic regression with offset $\log(\pi_{ij})$ is written as
+$$
+\sum_{i=1}^n \sum_{j=1}^{k_i} \left( y_{ij} - \frac{1}{1+\exp (- W_{ij} \theta + \log (\pi_{ij}) )} \right) - \frac{1}{2 \sigma^2} \sum_{k=3}^{K_b} \beta_k^2 - \frac{1}{2 \sigma^2} (K_b - 3).
+$$
+We derive the iteratively re-weighted least squares algorithm next.
+
+\subsection{Penalized adaptive Gaussian quadrature}
+
+In standard logistic mixed-effects model, the likelihood is given by
+$$
+\sum_{i=1}^n \int \left( \frac{\exp(W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij})}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{y_{ij}} \left( \frac{1}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{1-y_{ij}} \phi (b_i; \sigma_b^2) db_i
+$$
+
 \end{document}

### 2020-05-20 write-up/fda-recurrent.tex
Update fda-recurrent.tex

--- 2020-03-21 write-up/fda-recurrent.tex
@@
 
+\def\W{{\bf W}}
 \def\Bka{{\it Biometrika}}
@@
 
+\section{Penalized logistic mixed-effects model}
+
+Here we derive a penalized adaptive Gaussian quadrature (PADQ) to fit the   penalized logistic mixed-effects model.
+Here, for clarity, we write $Y_{ij} \in \{ 0, 1 \}$ to be the sequence of binary indicators for participant $i = 1,\ldots,n$ and $j$ indexes the event and subsampled times.
+We write $W_{ij}$ to denote the time-varying covariate.
+Then the score equations for the penalized logistic regression with offset $\log(\pi_{ij})$ is written as
+$$
+\sum_{i=1}^n \sum_{j=1}^{k_i} \left( y_{ij} - \frac{1}{1+\exp (- W_{ij} \theta - \log (\pi_{ij}) )} \right) W_{ij} - \frac{1}{\sigma^2} \sum_{k=3}^{K_b} \beta_k
+$$
+Let $B$ be the matrix that extracts the entries in $\theta$ corresponding to $\{ \beta_k\}_{k=3}^{K_b}$ (i.e., $B \theta = ({\bf 0}, \beta_3, \ldots, \beta_{K_b},{\bf 0})$).
+Let $\W_i$ be a $k_i \times p$ matrix with the $j$th row equal to $W_{ij}$.
+Let $\mu(\W_i; \theta, \pi_i)$ be the vector of length $k_i$ with the $j$th entry equal to $\mu(W_{ij}; \theta, \pi_{ij}) = \frac{1}{1+\exp(-W \theta - \log(\pi))}$.
+Finally, let $Y_i = (y_{i1},\ldots, y_{i k_i}$ and $E_i (\theta) = Y_i - \mu(\W_i; \theta, \pi_i)$.
+Then we can succinctly write the above as $\sum_{i=1}^n \W_i^\top E_i (\theta)- \frac{1}{\sigma^2} B \theta$.
+
+We derive the iteratively re-weighted least squares algorithm next. First, take the derivative with respect to $\theta$ to construct the Hessian
+$$
+- \left[ \sum_{i=1}^n \sum_{j=1}^{k_i} \frac{W_{ij} W_{ij}^\top}{(1+\exp (- W_{ij} \theta - \log (\pi_{ij}) ))^2} \exp(-W_{ij} \theta - \log (\pi_{ij})) + \frac{B}{\sigma^2} \right].
+$$
+Define the matrix $V_i (\theta)$ be the diagonal matrix with the $(j,j)$th entry equal to $\mu(W_{ij}; \theta, \pi_{ij}) \cdot (1-\mu(W_{ij}; \theta, \pi_{ij})$.
+Then we can express the Hessian more succinctly as $- \left[ \sum_{i=1}^n \W_i^\top V_i (\theta) \W_i + \frac{B}{\sigma^2} \right]$.
+Combining these, starting at an initial estimate, the Newton-Rhapson update given current parameter value $\hat \theta_k$ is given by
+$$
+\hat \theta_{k+1} = \hat \theta_{k} + \left[ \sum_{i=1}^n \W_i^\top V_i (\hat \theta_k) \W_i + \frac{B}{\sigma^2} \right]^{-1} \left(\sum_{i=1}^n \W_i^\top E_i (\hat \theta_k) - \frac{1}{\sigma^2} B \hat \theta_k \right)
+$$
+
+To learn $\sigma^2$, we use the connection to the mixed-effects literature.
+Fixing $\theta$, the log-likelihood components related to $\sigma^2$ is given by $-\frac{1}{2\sigma^2} \theta^\top B \theta - \frac{B {\bf 1}}{2} \log (\sigma^2)$.
+Differentiating and setting equal to zero yields the estimator
+$\hat \sigma^2 = \frac{\theta^\top B \theta}{B {\bf 1}}$.
+The complete algorithm is then alternating between Netwon-Rhapson for fixed $\sigma^2$ to estimate $\theta$ and then estimating $\sigma^2$ for fixed $\theta$ using the above formula.
+
+\subsection{Estimation of random-effects via Newton-Rhapson}
+
+In standard logistic mixed-effects model, with $b_i \sim N(0, \Psi)$ and $b_i \in \mathbb{R}^q$, then the likelihood is given by
+$$
+\begin{aligned}
+& \prod_{i=1}^n \int \prod_{j=1}^{k_i} \left( \frac{\exp(W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij})}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{y_{ij}} \left( \frac{1}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{1-y_{ij}} \phi (b_i; \sigma_b^2) db_i \\
+\propto & \prod_{i=1}^n \int \exp \left[ \sum_{j=1}^{k_i} \left( y_{ij} \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) - \log \left( 1 + \exp \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) \right) \right) - b_i^\top \Psi^{-1} b_i /2 \right] db_i \\
+\propto& \prod_{i=1}^n \int \exp \left[ g(\theta, b_i, \pi_{ij}, \Psi) \right] db_i,
+\end{aligned}
+$$
+where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$. Differentiating the function $g(\theta, b_i, \pi_{ij}, \Psi)$ with respect to $b_i$ yields
+$$
+\begin{aligned}
+\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i} &= \sum_{j=1}^{k_i} \left( y_{ij} Z_{ij} - \frac{Z_{ij}}{1+\exp(-W_{ij}^\top \theta - Z_{ij}^\top b_i - \log \pi_{ij}}\right) - \Psi^{-1} b_i \\
+\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i^2} &= - \sum_{j=1}^{k_i} \frac{Z_{ij}Z_{ij}^\top}{(W_{ij}^\top \theta + Z_{ij}^\top b_i + \log \pi_{ij})^2} - \Psi^{-1}
+\end{aligned}
+$$
+Redefining terms from the prior section, we can express the above as
+$$
+\begin{aligned}
+\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
+\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
+\end{aligned}
+$$
+
+For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
+$$
+\hat b_i^{(k+1)} = \hat b_i^{(k)} + \left( \sum_{j=1}^{k_i} \frac{Z_{ij}Z_{ij}^\top}{(W_{ij}^\top \theta + Z_{ij}^\top \hat b_i^{(k)} + \log \pi_{ij})^2} + \Psi^{-1} \right)^{-1} \left( \sum_{j=1}^{k_i} \left( y_{ij} Z_{ij} - \frac{Z_{ij}}{W_{ij}^\top \theta + Z_{ij}^\top \hat b_i^{(k)} + \log \pi_{ij}}\right) - \Psi^{-1} \hat b_i^{(k)} \right)
+$$
+Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
+We employ adaptive gaussian quadrature (cites) as it has been shown that for infrequent events the Laplace approximation tends to underestimate effects (cites).
+To this end, we
+
+We write this as $\prod_{i=1}^n l_{ij} (\theta, b_i, \pi_{ij}) $
+
 \end{document}
-
-\item {\bf Step 1}: Estimate~$\hat \mu(s,t)$ via sandwich smoother for
-  all~$s \in [0,\Delta]$ and~$t \in \cup_{i=1}^n \left( D_i \cup N_i
-  \right)$. Define~$\tilde X (s,t) = X(s,t) - \hat \mu(s,t)$ to be a
-  mean-zero process for each~$t$.
-\item {\bf Step 2}: Compute the pooled sample covariance
-\[
-\hat \Xi (s , s^\prime) =
-\left( \sum_{i=1}^n \left | N_i \cup D_i \right |\right)^{-1}
-\left( \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \tilde X (s,t) \tilde X
-  (s^\prime, t) \right).
-\]
-Compute at a fine set of grid points. This is an estimate of
-\[
-\Xi (s, s^\prime) = \int_{\mathcal{T}} c( (s,T), (s^\prime, T) ) f(T) dT
-\]
-where~$f(T)$ is a superposition of the event process and the
-subsampling process.
-\item {\bf Step 3}: Take spectral decomposition~$\{ \hat \phi_k (s),
-  \hat \lambda_k \}_{k \geq 1}$. Choose~$K_x \geq K_b$ large enough.
-\item {\bf Step 4}: Compute
-\[
-\hat c_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (s,t) \hat \phi_k (s) ds
-\]
-Then
-\[
-X(s,t) \approx \hat \mu(s,t) + \sum_{k=1}^{K_x} \hat c_{i,k} (t)
-\phi_k (s).
-\]
-\item {\bf Step 5}: Plugging both the spline and KL decomposition we
-  have
-\begin{align*}
-\int_{t-\Delta}^t X(s,t) \beta(s) ds
-  &= \int_{t-\Delta}^t \left[ \hat \mu(s,t) + \sum_{k=1}^{K_x}
-    c_{i,t,k} \phi_k (s) \right] \times \left[ \sum_{k^\prime=1}^{K_b}
-    \beta_k \psi_k (s) \right] \\
-  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\phi, \psi} ] \beta
-\end{align*}
-where~$M_{i,t}$ is a vector with~$i$th entry equal
-to~$\int_{t-\Delta}^t \hat \mu(s,t) \psi_k (s) ds$, $C_{i,t} =
-(c_{i,t,1}, \ldots, c_{i,t,K_x}) \in \mathbb{R}^{K_x \times 1}$,
-$\beta = (\beta_1, \ldots, \beta_{K_b}) \in \mathbb{R}^{K_b \times 1}$
-and~$J_{\phi, \psi} \in \mathbb{R}^{K_x \times K_b}$ where the~$(i,j)$
-entry is given by
-\[
-\int_{t-\Delta}^t \phi_{i} (s) \psi_{j} (s) ds.
-\]
-\item {\bf Step 6}: The approximate score equations
-\begin{equation}
-\label{eq:weighted_esteq}
-\hat U_n (\theta) = \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
-    \sum_{u \in D_i} (1 - w_i (u; \theta)) W_{i,u} \right \}
-\end{equation}
-where
-\begin{equation}
-\label{eq:waage_weight}
-w_i (t; \theta) = \frac{h_i(t;\theta)}{\pi_i (t) + h_i (t;\theta)}.
-\end{equation}
-Under weights~\eqref{eq:waage_weight} and log-linear intensity
-function,~\eqref{eq:weighted_esteq} is the
-\end{itemize}

### 2020-05-21 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 
+\def\k{{\bf k}}
+\def\z{{\bf z}}
 \def\W{{\bf W}}
@@
 $$
-where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$. Differentiating the function $g(\theta, b_i, \pi_{ij}, \Psi)$ with respect to $b_i$ yields
+where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$.
+Differentiating the function $g(\theta, b_i, \pi_{ij}, \Psi)$ with respect to $b_i$, the connection to the prior section can be used.
+Re-defining terms appropriately allows us to express the derivatives as
 $$
 \begin{aligned}
-\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i} &= \sum_{j=1}^{k_i} \left( y_{ij} Z_{ij} - \frac{Z_{ij}}{1+\exp(-W_{ij}^\top \theta - Z_{ij}^\top b_i - \log \pi_{ij}}\right) - \Psi^{-1} b_i \\
-\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i^2} &= - \sum_{j=1}^{k_i} \frac{Z_{ij}Z_{ij}^\top}{(W_{ij}^\top \theta + Z_{ij}^\top b_i + \log \pi_{ij})^2} - \Psi^{-1}
+\frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
+\frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
 \end{aligned}
 $$
-Redefining terms from the prior section, we can express the above as
+For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
+$$
+\hat b_i^{(k+1)} = \hat b_i^{(k)} + \left[ Z_i V_i (\theta, \hat b_i^{(k)}) Z_i + \Psi^{-1} \right]^{-1}  \left( Z_i E_i (\theta, \hat b_i^{(k)}) - \Psi^{-1} \hat b_i^{(k)} \right)
+$$
+
+Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
+Here we employ adaptive gaussian quadrature (cites) as it has been shown that for infrequent events the Laplace approximation tends to underestimate effects (cites).
+Let $R_i^\top R_i = Z_i^\top V_i (\theta, b_i) Z_i + \Psi^{-1}$, $z_j$ and $w_j$ for $j=1,\ldots, N_{GQ}$ be the weights for the one-dimensional Gaussian quadrature rule.
+Let $\z_\k = (z_{k_1},\ldots, z_{k_q})$, $\tilde b_{i \k} = \hat b_i + R_i^{-1} \z_\k$, and $W_{\k} = \exp ( \| \z_\k \|^2 ) \prod_{l=1}^{q} w_{k_l}$.
+Then the AGQ rule is given by
 $$
 \begin{aligned}
-\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
-\frac{\partial g(\theta, b_i, \pi_{ij}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
+&\int \exp \left[ g(\theta, b_i, \pi_{i}, \Psi) \right] db_i \\
+=& \int | R_i |^{-1} \exp \left[ g(\theta, \hat b_i + R_i^{-1} \z, \pi_{i}, \Psi) + \| \z \|^2 / 2 \right] \exp ( - \| \z \|^2 / 2 ) d\z \\
+=& (2 \pi)^{q/2} | R_i |^{-1} \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k}
 \end{aligned}
 $$
-
-For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
-$$
-\hat b_i^{(k+1)} = \hat b_i^{(k)} + \left( \sum_{j=1}^{k_i} \frac{Z_{ij}Z_{ij}^\top}{(W_{ij}^\top \theta + Z_{ij}^\top \hat b_i^{(k)} + \log \pi_{ij})^2} + \Psi^{-1} \right)^{-1} \left( \sum_{j=1}^{k_i} \left( y_{ij} Z_{ij} - \frac{Z_{ij}}{W_{ij}^\top \theta + Z_{ij}^\top \hat b_i^{(k)} + \log \pi_{ij}}\right) - \Psi^{-1} \hat b_i^{(k)} \right)
-$$
-Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
-We employ adaptive gaussian quadrature (cites) as it has been shown that for infrequent events the Laplace approximation tends to underestimate effects (cites).
-To this end, we
-
-We write this as $\prod_{i=1}^n l_{ij} (\theta, b_i, \pi_{ij}) $
+The score equations
+$$
+\sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}}
+\left( \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right)^{-1}
+\left( \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi)
+$$
+Define $\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi)$ to be the weights.  Then we can express score and Hessian as:
+$$
+\begin{aligned}
+&\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \\
+&\sum_{\k}
+\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial^2}{\partial \theta \partial \theta^\top} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) +
+\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top \\
+&- \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top
+\end{aligned}
+$$
+
+When $N_{GQ} = 1$, the AGQ approximation is equivalent to a Laplacian approximation.
 

### 2020-06-22 write-up/fda-recurrent.tex
Updating paper

@@
 \label{prop:optimal}
+\textcolor{red}{
 If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
 This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
+WD: Consider larger class and show the proof doesn't extend to non PP models?}
 \end{lemma}

### 2020-06-22 write-up/fda-recurrent.tex
Fixing writeup

@@
 
+\subsection{Multiple sensors}
+\label{section:multiplesensors}
+
 \subsection{Missing data}
@@
 
-\subsection{Bootstrap and multiple imputation under uncongenial }
+\subsubsection{Bootstrap and multiple imputation under uncongenial }
 
@@
 
-\subsection{Multiple sensors}
-
 \section{A worked example: Adolescent psychiatric inpatient mHealth study} \label{section:example}
 
-During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital  were enrolled in
- with histologically verified liver cirrhosis were recu
-those presenting to a psychiatric hospital with suicide ideation
-and/or a recent suicide attempt) during a high risk time period (i.e.,
-post-hospitalization)
-
-Currently, only 51 patients for whom the initial biopsy could be
-reevaluated using more restrictive criteria were available for
-analysis.
-One of the goals of the study was to assess the association between sensor-based covariate information can used to identify distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement.
+During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital were enrolled in an observational study.
+Study data were graciously provided by Evan Kleiman and his study team ({\tt https://kleimanlab.org}).
+Each study participant wore an Empatica E4~\cite{empaticae3}, a medical-grade wearable device that offers real-time physiological data.
+One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified
+
+can used to identify distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement.
 The scientific hypothesis to be tested is whether a spike in EDA and/or gACC
 
-The goal of the sutyd
-
-EDA and ACC sensor data along with buttonpress timestamps were
-graciously provided by Evan Kleiman and his study team {\tt
-https://kleimanlab.org}.
-At the end of the study period, the mortality rate was 292/488, or
+To avoid, we focus on the study window 9AM to 8PM each study day.
+
+At the end of the study period, the base rate of button pressing  rate was 292/488, or
 approximately 60\%.
+
+\begin{figure}[!th]
+\caption{Button press distribution across study days (include number of people).}
+\end{figure}
 
@@
 and subsequently at roughly twelve-month intervals.
-The individual prothrombin trajectories are highly variable,
-both in forward and in reverse time,
-which tends to obscure patterns and trends.
-In Figure~\ref{fig:mean_traj}a the mean trajectory is plotted against time from recruitment
-for two patient groups placebo/prednisone and censored/not censored.
+The individual EDA and AI trajectories are highly variable, both in forward and when considering reverse time from event timepoints, which tends to obscure patterns and trends.
+In Figure~XX, the mean trajectory is plotted against time from recruitment for two patient groups placebo/prednisone and censored/not censored.
 Naturally, only those patients who are still alive are included in the average
 for that time.
-Figure~\ref{fig:mean_traj}b shows the same plots in reverse alignment.
+Figure XX hows the same plots in reverse alignment.
 While there are certain similarities in the two plots,
@@
 
-Inspection of the graphs for uncensored patients in the right panel of Figure~\ref{fig:mean_traj} suggests
-beginning with the simplest revival model in which the sequences
-for distinct patients are independent Gaussian with moments
-
-{\bf comparison over $\Delta$ and using imputation}
-
-\subsection{Effect of electrodermal activity on prognosis}
+\begin{figure}[!th]
+\caption{EDA and ACC plots for average at non-event and event times.}
+\end{figure}
+
+Inspection of the graphs for uncensored patients in the right panel of Figure XX suggests beginning with the simplest revival model in which the sequences
+for distinct patients are independent Gaussian with moments {\bf comparison over $\Delta$ and using imputation}
+
+\begin{figure}[!th]
+\caption{$\beta (t)$ for ACC and EDA with 95\% CI.}
+\end{figure}
+
+\subsection{Accounting for missing data and heterogeneity}
+
+\subsection{Effect of electrodermal activity and activity index on moments of distress}
+
 Over a period of 5~years and one month following recruitment,

### 2020-06-25 write-up/fda-recurrent.tex
putting sim to run on servers

@@
 
+\begin{itemize}
+\item Sample every 0.5 hours; then subsample at each subrate. Estimates are given by $\hat \beta_{i,0.5} (t)$ and $\hat \beta_{i,1} (t)$.
+\item Bias is average over the estimates $\bar \hat \beta_{0.5} (t) = 100^{-1} \sum_{i=1}^100 \hat \beta_{i, 0.5} (t)$ compared to truth.
+\item Take squared difference integrated to get variance.
+\item To get the subsampling variance take $1/1000 \sum_{i=1}^1000 (\hat \beta_{i,0.5} (t) - \hat \beta_{i,1} (t))^2$.
+\item Normalize everything by squared function!
+\end{itemize}
+
 We sample non-events at rates of 1 per 12/6/1/0.5 hours.
@@
 
+\begin{equation}
+\label{eq:multihazardlinear}
+h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) =
+h_0 (t; \gamma) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
+  + \int_{t-\Delta}^t x_{1i} (s) \beta_1(s) ds + \int_{t-\Delta}^t x_{2i} (s) \beta_2(s) ds  \right)
+\end{equation}
+
 \subsection{Missing data}
@@
 \subsection{Multilevel models}
+
+\begin{equation}
+\label{eq:multilevel}
+h_i \left( t \given  H_{i,t}^{NX} ; \theta, b_i \right) \approx
+\exp \left( Z_{t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
+  + \left[ M_{i,t}^\top + C_{i,t}^\top J_{\hat \psi, \phi}\right] (\beta + b_i) \right)
+= \exp \left( W_{i,t}^\top \theta  + Z_{i,t}^\top b_i \right),
+\end{equation}
 
@@
 
+\textcolor{red}{Assume fixed $\lambda$.}
+
 In standard logistic mixed-effects model, with $b_i \sim N(0, \Psi)$ and $b_i \in \mathbb{R}^q$, then the likelihood is given by

### 2020-06-25 simulations/basic_simulation.R

@@ -38,11 +38,11 @@ K_x = 35  # Pick first 35 eigen-vectors
 cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
 
 # set.seed("134163")
-set.seed("131312")
+# set.seed("131312")
 base_num_events = 5
 base_rate = logit(5/length(times))
 sampling_rate = 10/1000
-dataset = generate_complete_data(N = 1000, Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate, base_rate)
+dataset = generate_complete_data(N = 4000, Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate, base_rate)
 rm(abs_diff, C, Cov_X, x)
 
 dataset =data.frame(dataset)
@@ -82,7 +82,7 @@ K <- 100
 lambdapath <- round(exp(seq(log(lambda_max), log(lambda_max*epsilon), 
                             length.out = K)), digits = 10)
 
-set.seed("97139817")
+# set.seed("97139817")
 # Start the clock!
 ptm <- proc.time()
 

### 2020-06-25 simulations/basic_simulation_functions.R

@@ -60,7 +60,7 @@ generate_userday <- function(Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate,
 generate_complete_data <- function(N = 500, Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate, base_rate) {
   full_data = rep(0,0)
   for (i in 1:N) {
-    print(paste("On userday", i))
+    # print(paste("On userday", i))
     userday_data = cbind(i, generate_userday(Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate, base_rate))
     full_data = rbind(full_data, userday_data)
   }

### 2020-06-25 simulations/basic_simulation.R
Making sims ready for servers

@@ -1,5 +1,6 @@
-## Source the functions
+## Source the functions and necessary packages
 source("./basic_simulation_functions.R")
+require("glmnet")
 
 ### Matern covariance parameters
 nu = 1/2 # controls smoothness
@@ -37,12 +38,12 @@ eig_Sigma = eigen(Sigma)
 K_x = 35  # Pick first 35 eigen-vectors
 cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
 
-# set.seed("134163")
 # set.seed("131312")
+id = runif(1, min = 0, max = 100000)
 base_num_events = 5
 base_rate = logit(5/length(times))
-sampling_rate = 10/1000
-dataset = generate_complete_data(N = 4000, Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate, base_rate)
+max_sampling_rate = 12/1000 # Max sampling rate is 1 obs every 1/2 hour
+dataset = generate_complete_data(N = 500, Cov_X, C, times, eig_Sigma, beta_1t, max_sampling_rate, base_rate)
 rm(abs_diff, C, Cov_X, x)
 
 dataset =data.frame(dataset)
@@ -50,64 +51,18 @@ names(dataset)[1] = "userday"
 names(dataset)[2] = "Y"
 
 agg_results = aggregate(Y~userday,dataset, sum)
-table(agg_results[,2])
+agg_summary = c(mean(agg_results[,2]),var(agg_results[,2])) # Report the mean and variance of number of events per day 
+write.table(agg_summary, file = "aggregate_summary.csv", append = TRUE)
 
-## Construct the J matrix
-K_b = 35
-local_times = times[1:44]
-num=K_b-3
-qtiles <- seq(0, local_times[length(local_times)], length = num + 2)[-c(1, num + 2)]
-knots <- quantile(local_times, qtiles)
-## Basis = bs(t, kb)
-Basis = cbind(1, local_times, local_times^2, sapply(knots, function(k) ((local_times - k > 0) * (local_times - k)) ^ 2))
-Psi = t(eig_Sigma$vectors[,1:K_x])
-Phi = Basis
-J = Psi%*%Phi
-model.matrix = dataset[,3:ncol(dataset)]
-new.model.matrix = as.matrix(model.matrix)%*%J*gap
-w = cbind(new.model.matrix)
-X=as.matrix(w[,1:3])
-Z=as.matrix(w[,4:dim(w)[2]])
[…]

### 2020-06-25 simulations/basic_simulation_functions.R

@@ -67,7 +67,6 @@ generate_complete_data <- function(N = 500, Cov_X, C, times, eig_Sigma, beta_1t,
   return(full_data)
 }
 
-
 expit <- function(x) {exp(x)/(1+exp(x))}
 
 logit <- function(p) {log(p/(1-p))}
@@ -76,4 +75,73 @@ mise_calc <- function(beta_1t, betaHat.net) {
   temp_Xb = (beta_1t - betaHat.net)^2
   term2 = (temp_Xb[1] + temp_Xb[length(temp_Xb)])/2 * gap + sum(temp_Xb[2:(length(temp_Xb)-1)]*gap) 
   return(term2)
+}
+
+subsample_dataset <- function(dataset, thinning_rate) {
+  keep_obs = rbinom(nrow(dataset), size = 1, prob = thinning_rate*(1-dataset$Y) + dataset$Y)
+  return(dataset[keep_obs,])
+}
+
+construct_J <- function(times, eig_Sigma, dataset) {
+  ## Construct the J matrix
+  K_b = 35
+  local_times = times[1:44]
+  num=K_b-3
+  qtiles <- seq(0, local_times[length(local_times)], length = num + 2)[-c(1, num + 2)]
+  knots <- quantile(local_times, qtiles)
+  ## Basis = bs(t, kb)
+  Basis = cbind(1, local_times, local_times^2, sapply(knots, function(k) ((local_times - k > 0) * (local_times - k)) ^ 2))
+  Psi = t(eig_Sigma$vectors[,1:K_x])
+  Phi = Basis
+  J = Psi%*%Phi
+  model.matrix = dataset[,3:ncol(dataset)]
+  new.model.matrix = as.matrix(model.matrix)%*%J*gap
+  w = cbind(new.model.matrix)
+  return(w)
+}
+
+runglmnet <- function(sampling_rate, dataset, w, epsilon = 0.0001) {
+  n.tmp = length(dataset$Y)
+  p.tmp = ncol(w)
+  subsample_offset = rep(log(sampling_rate),nrow(dataset))
+  p.fac = rep(1, ncol(w))
+  p.fac[1:4] = 0 #no penalty on the first 4 variables
+  lambda_max <- 1/n.tmp
+  epsilon <- .0001
+  K <- 100
+  lambdapath <- round(exp(seq(log(lambda_max), log(lambda_max*epsilon), 
+                              length.out = K)), digits = 10)
+  
+  # set.seed("97139817")
+  # Start the clock!
+  ptm <- proc.time()
[…]

### 2020-08-01 write-up/fda-recurrent.tex
Updated paper for simulation results

@@
 \]
-We set~$\Delta = 1/(12*2)$ to mimic a $30$-minute window for a $12$-hour day.
+We set~$\Delta$ to mimic a 30-minute window for a $12$-hour day.
 We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day.
-We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$
-
+We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$.
+
+Given each simulated user-day, we then
 \begin{itemize}
@@
 
-We sample non-events at rates of 1 per 12/6/1/0.5 hours.
-
-Since we are primarily interested in accuracy of the final estimates~$\hat \beta(s)$, we report the mean integrated squared error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{100} \sum_{j=1}^{100} \int_{0}^{\Delta^\prime} ( \hat \beta_j (s) - \beta(s))^2 ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
+We sample non-events at rates of 1 per 4/2/1/0.5 hours.
+
+Since we are primarily interested in accuracy of the final estimates~$\hat \beta(s)$, we report the mean integrated squared error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{1000} \sum_{j=1}^{1000} \int_{0}^{\Delta^\prime} ( \hat \beta_j (s) - \beta(s))^2 ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
 We define the MISE in this manner to account for settings where~$\Delta$ is unknown.
@@
 \begin{table}[!th]
-\begin{tabular}{c | c c c c c}
-& \multicolumn{5}{c}{Sampling rate} \\ \cline{2-6}
- & 0.5 & 1 & 5 & 25 & 50 \\ \hline
-Subsampling variance & (10,12) & - & - & - & - \\
-Variance & (10,12) & - & - & - & - \\
-Squared Bias & - & - & - & - & -  \\
-MISE & - & - & - & - & -  \\ \hline
-Type I error & - & - & - & - & -  \\ \hline
-Avg runtime & - & - & - & - & - \\
+\begin{tabular}{c | p{2cm} p{2cm} p{2cm} p{2cm}}
+& \multicolumn{4}{c}{Sampling rate} \\ \cline{2-5}
+ & 0.5 & 1 & 2 & 4 \\ \hline
+Subsampling variance & - & $9.0 \times 10^{-7}$ & $2.9 \times 10^{-6}$ & $5.8 \times 10^{-6}$  \\
+Variance & $7.9 \times 10^{-3}$ & $9.2 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.30 \times 10^{-2}$ \\
+Squared Bias & $9.8 \times 10^{-2}$ & $9.5 \times 10^{-2}$ & $9.3 \times 10^{-2}$ & $9.2 \times 10^{-2}$   \\
+MISE & $1.1 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ \\ \hline
+Avg. runtime (secs) & 317 & 177 & 104 & 68 \\ \hline
+Subsampling variance & - & $2.9 \times 10^{-6}$ & $8.9 \times 10^{-6}$ & $2.0 \times 10^{-5}$  \\
+Variance & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.1 \times 10^{-2}$ \\
+Squared Bias & $6.9 \times 10^{-2}$ & $6.5 \times 10^{-2}$ & $6.0 \times 10^{-2}$ & $5.4 \times 10^{-2}$   \\
+MISE & $7.9 \times 10^{-2}$ & $7.9 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
+Avg. runtime (secs) & 36 & 23 & 15 & 12 \\ \hline
 \end{tabular}
@@
 \end{table}
-
-\subsection{Coverage properties}
-
-We simulate the system $1000$ times under $\beta \equiv 0$ and check the type I error.
 
@@
 patient~$u$ had eight appointments with prothrombin values as follows:
+
+\section{Discussion}
+We abide by the comment of Robert Gentleman on big data: (1) make the data as small as possible as fast as possible.
+Statistics works!
+With subsampling and proper re-weighting the analysis is converted to well-known territory.
+(2) Make the software and tools as fast and as maintanable as possible.
+This is exactly teh benefit of subsampling.
+we convert to penalized logistic regression on a smaller but still relatively efficient dataset.
+This means we can use all the existing tools!
+
+Extensions to online and sampling over and over as we improve estimators.
 

### 2020-08-01 simulations/simulation_summary.R

@@ -1,0 +1,70 @@
+source("./basic_simulation_functions.R")
+df = data.frame(read.table("./output_csvs/results2.csv", sep = ",", header = T))
+# df = data.frame(read.table("./output_csvs/results_sim22.csv", sep = ",", header = T))
+df = df[order(df$ids, df$rates),]
+
+M = 1000 # Num of time points
+times = seq(0,1, length.out = M) # Equally spaced timepoints in [0,1]
+gap = diff(times)[1] # What is assumed gap times
+beta_truth = 100*sin(1:44/44*2*pi-pi/2)
+# beta_truth = 20*exp(0.1*1:44)/mean(exp(0.1*1:44))
+true_variance = mise_calc(beta_truth, beta_truth*0)
+
+unique_rates = unique(df$rates)
+unique_rates = sort(unique_rates, decreasing = T)
+betameans = matrix(nrow = length(unique_rates), ncol = length(6:ncol(df)))
+for (i in 1:4) {
+  betameans[i,] = colMeans(df[df$rates == unique_rates[i],6:ncol(df)], na.rm = TRUE)
+}
+
+## Simple visualization to show that means are same across subsamples
+## This is a good thing!  Means there is no bias due to subsampling 
+## The bias inherent in even most data case is due to small n and penalization
+par(mar = c(3,3,2,1)+0.1)
+plot(1:44, beta_truth, type = "l", axes = F, xlab = "Time", ylab = "")
+axis(side = 1, labels = F)
+axis(side = 2, labels = F)
+for(i in 1:4) {
+  lines(as.numeric(betameans[i,]), type = "l", col = "darkgrey")
+}
+
+
+
+sqbias = vector(length = length(unique_rates))
+for(i in 1:length(unique_rates)) {
+  sqbias[i] = mise_calc(beta_1t = beta_truth, betaHat.net = betameans[i,])
+}
+
+avg_runtime = rep(0,length(unique_rates))
+for(i in 1:length(unique_rates)) {
+  avg_runtime[i] = mean(df[df$rates == unique_rates[i],]$runtime)
+}
+
+
+variance = rep(0,length(unique_rates))
+mise = rep(0,length(unique_rates))
+for(i in 1:length(unique_rates)) {
+  subsetdf = df[df$rates == unique_rates[i],6:ncol(df)]
+  for (row in 1:nrow(subsetdf)) {
[…]

### 2020-08-05 write-up/fda-recurrent.tex
updated paper

@@
 \subsection{Estimation of the windowed covariate history}
-
+\label{sec:margcov}
 We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
@@
 \subsection{Connection to logistic score functions}
+\label{eq:logistication}
 
@@
 
-Given each simulated user-day, we then
-\begin{itemize}
-\item Sample every 0.5 hours; then subsample at each subrate. Estimates are given by $\hat \beta_{i,0.5} (t)$ and $\hat \beta_{i,1} (t)$.
-\item Bias is average over the estimates $\bar \hat \beta_{0.5} (t) = 100^{-1} \sum_{i=1}^100 \hat \beta_{i, 0.5} (t)$ compared to truth.
-\item Take squared difference integrated to get variance.
-\item To get the subsampling variance take $1/1000 \sum_{i=1}^1000 (\hat \beta_{i,0.5} (t) - \hat \beta_{i,1} (t))^2$.
-\item Normalize everything by squared function!
-\end{itemize}
-
-We sample non-events at rates of 1 per 4/2/1/0.5 hours.
-
-Since we are primarily interested in accuracy of the final estimates~$\hat \beta(s)$, we report the mean integrated squared error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{1000} \sum_{j=1}^{1000} \int_{0}^{\Delta^\prime} ( \hat \beta_j (s) - \beta(s))^2 ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
-We define the MISE in this manner to account for settings where~$\Delta$ is unknown.
-Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biase
+We generate $1000$ datasets, each consisting of $500$ user-days.
+For a given simulated user-day, we randomly sample non-event times using a Poisson process with rate of a halfhour.
+We use the proposed methodology to construct the estimate $\hat \beta_{i,0.5} (t)$ for the $i$th simulated dataset; we then subsample the sampled non-event times with thinning probability $1/2$, $1/4$ and $1/8$.
+This results in randomly sampled non-event times given by a Poisson process with rates of an hour, two hours, and four hours.
+We can construct the corresponding estimates: $\hat \beta_{i,1} (t)$, $\hat \beta_{i,2} (t)$, and $\hat \beta_{i,4} (t)$ respectively.
+Subsampling allows us to compare the variance due to subsampling as compared to the variance due to sampling fewer non-event times.
+
+Since we are primarily interested in accuracy, we report the mean integrated squared error (MISE) defined as~$\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \beta(s))^2 ds$ for $j=0.5,1,2,4$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
+The MISE is defined in this manner to account for settings where~$\Delta$ is unknown.
+Next, let $\bar \beta_{j} (t) = \frac{1}{1000} \sum_{i=1}^{1000} \hat \beta_{i,j} (t)$ denote the average estimate for $j=0.5,1,2,4$.
+Then the squared bias is given by $\int_{0}^{\infty} ( \bar \beta_{j} (s) - \beta (s))^2 ds$ and the variance is given by $\frac{1}{1000} \sum_{j=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \bar \beta_j(s))^2 ds$.
+The subsampling variance is defined as $\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \hat \beta_{i,0.5} (s))^2 ds$.
+Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biased as well as the subsampling variance.
+To allow fair comparisons across the two choices of $\beta(s)$, all reported numbers are scaled by the integrated square of the true function $\int_{0}^{\infty} \beta(s)^2 ds$.
+The average runtime (in seconds) is also reported.
+
+Table~\ref{tab:mise} demonstrates that the variance does increase as the sampling rate decreases.
+However, the rate of increase in the MISE is quite low relative.
+In the second case study, for example, the MISE increases by 6\% while the run time is 3 times faster.
+In the first case study, the MISE remains roughly constant but the run time is 4.7 times faster.
+This highlights the efficiency-computation trade-off.
+Extrapolating to a complete data analysis, we can see that the run time will continue to increase significantly with minimal improvement in the mean integrated squared error.
 
@@
 \end{tabular}
-\caption{Mean-integrated squared error and type I error as a function of sampling rate for $\beta(s)$ given by (1) and (2) respectively.}
+\caption{Mean-integrated squared error, varaince, squared bias, and subsampling variance for $\beta(s)$ given by (1) and (2) respectively.}
 \label{tab:mise}
@@
 
-In this section, we demonstrate the flexibility of our approach by exploring extensions in several important directions to ensure these methods are robust for practical use in mobile health.
+In this section, we demonstrate the flexibility of the proposed approach by exploring extensions in several important directions to ensure these methods are robust for practical use with high frequency data.
 This section will continue to leverage the connection to generalized functional linear models provided by Lemma~\ref{lemma:logistic}.
 
-\subsection{Multiple sensors}
+\subsection{Multivariate extensions}
 \label{section:multiplesensors}
 
+In this section, we extend our model to the case of multiple functional regressors.
+That is, suppose $L$ health processes, i.e., $\bfx_i = \{ \bfx_i (t) = (x_{i,1} (t), \ldots, x_{i,L} (t)) \}_{0 < s < \tau_i}$, for each participant is measured at a dense grid of time points.
+In the suicidal ideation case study, for example, accelerometer is measured at a rate of 32Hz while electrodermal activity (EDA) is measured at a rate of 4Hz.
+A multivariate extension of our model~\eqref{eq:hazard} is given by
 \begin{equation}
@@
 h_0 (t; \gamma) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + \int_{t-\Delta}^t x_{1i} (s) \beta_1(s) ds + \int_{t-\Delta}^t x_{2i} (s) \beta_2(s) ds  \right)
-\end{equation}
+  + \int_{t-\Delta}^t x_{i,1} (s) \beta_1(s) ds + \cdots + \int_{t-\Delta}^t x_{i,L} (s) \beta_2(s) ds  \right)
+\end{equation}
+The approach given in Section~\ref{eq:logistication} extends naturally to the multivariate functional setting.
+For each functional regressor, we estimate the pooled sample covariance $\Sigma_{y,l}$ for $y \in \{0,1\}$ and $l=1,\ldots,L$ as in Section~\ref{sec:margcov}.
+Let $\sum_{k=1}^\infty \hat \lambda^{(y)}_{k,l} \hat \psi^{(y)}_{k,l} (s) \hat \psi_{k,l}(t)$ be the spectral decomposition of $\hat \Sigma_{y,l}$.
+Then $x_{i,j}(t)$ is approximated using a truncated Karhunen-Loeve
+
+, the coefficient is expressed using separate marginal covaraiance
+$$
+\int_0^\Delta x_{il} \beta(t) =
+$$
+
+Note that we express each coefficient function in the same spline basis; however, different basis expansions are used for each health process as defined by the marginal covariance calculation in Section~\ref{sec:margcov}.
 
@@
 \label{section:missingdata}
-\walt{Question: How to perform multiple imputation for~$X(\cdot,t)$
-  and $X(\cdot, t^\prime)$ for~$|t - t^\prime| <\Delta$? That is, how
-  do you perform multiple imputation with this overlapping structure?}
-
-A core concern in mHealth is that sensor data can often be missing for intervals of time due to sensor wearing issue (e.g., sensor band becomes loose, turns off accidentally, data packets lost, or battery dies).
+% \walt{Question: How to perform multiple imputation for~$X(\cdot,t)$
+%   and $X(\cdot, t^\prime)$ for~$|t - t^\prime| <\Delta$? That is, how
+%   do you perform multiple imputation with this overlapping structure?}
+
+A core concern in mHealth is that sensor data can often be missing for intervals of time due to sensor wearing issues (e.g., sensor band becomes loose, turns off accidentally, data packets lost, or battery dies).
 In the suicidal ideation case study, for example, there are 2139 self-identified moments of distress across all 91 participants.
@@
 Missing data is a critical issue because~$c_{i,k} (t)$ cannot be estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$.
-Moreover, we want standard errors to reflect the uncertainty in these coefficients when missing data is prevalent.
+Moreover, standard errors should reflect the uncertainty in these coefficients when missing data is prevalent.
 \cite{Goldsmith2011} suggest using best linear unbiased predictors (BLUP) or posterior modes in the mixed effects model to estimate~$c_{i,k} (t)$; however, one criticism would be when there is substantial variability in these estimates.
@@
 
+The above procedure is
+
 Positives: (1) our imputation model is ``approximately'' congenial; (2) our imputatoin model is simple and easy to use.
@@
 To be careful, we investigate alternatives and assess impact.
+
+Recommendation from~\cite{Bartlett2020}.
 

### 2020-08-12 write-up/fda-recurrent.tex
updated

@@
 Then $x_{i,j}(t)$ is approximated using a truncated Karhunen-Loeve
-
-, the coefficient is expressed using separate marginal covaraiance
-$$
-\int_0^\Delta x_{il} \beta(t) =
-$$
-
+$$
+\int_{t-\Delta}^t X_{l} (t,s) \beta(t) = \left[M_{l,t} + {\bf c}_l (t)^\top J_{\hat \psi_l^{(y)}, \phi} \right] {\bf b}_l
+$$
 Note that we express each coefficient function in the same spline basis; however, different basis expansions are used for each health process as defined by the marginal covariance calculation in Section~\ref{sec:margcov}.
@@
 Here we present an extension of~\cite{Petrovich2018} to our setting by leveraging Lemma~\ref{lemma:logistic} and the marginal covariance estimation procedure to construct a multiple imputation procedure.
-Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at times~$\{ s_{i,r} \}_{r=1}^{k_{it}} \}$ in~$[0,\Delta]$.
+Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at times~$\{ s_{i,r} \}_{r=1}^{k_{it}}$ in~$[0,\Delta]$.
 Then
@@
 The imputation strategy generalizes the procedure introduced by~\cite{Petrovich2018} to the longitudinal functional data setting.
-This is only possible due to the connection established with logistic regression by Lemma~\ref{lemma:logistic}.
-We could consider a multiple imputation estimate with~$M$ imputations; the likelihood estimate is given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate from the~$m$th imputated dataset; the covariance matrix can be estimated by
+This is possible due to the connection established with logistic regression by Lemma~\ref{lemma:logistic}.
+We could then consider a multiple imputation estimate with~$M$ imputations; the likelihood estimate is given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate from the~$m$th imputated dataset; the covariance matrix can be estimated by
 \[
@@
 
-\subsubsection{Bootstrap and multiple imputation under uncongenial }
-
-The above procedure is
-
-Positives: (1) our imputation model is ``approximately'' congenial; (2) our imputatoin model is simple and easy to use.
-Negatives: (1) clearly not congenial under many events and/or high sampling rates. (2) difficult when we overlap due to over-sampling.
-
-Congeniality matters because we want to ensure good coverage
-Q: Can we derive a variance that provides coverage
-We know our imputation models are uncongenial albeit simple to implement.
-Impact should be small if events are rare as impute only in ranges where 1,0 labels give you everything.
-To be careful, we investigate alternatives and assess impact.
-
-Recommendation from~\cite{Bartlett2020}.
-
-Simple version compared to the von Hippel's boot MI proposal.
-\begin{itemize}
-\item Bootstrap $B$ times
-\item For $b = 1, \ldots, B$, impute $M$ times
-\item Let~$\hat \theta_b = M^{-1} \sum_m \hat \theta_{b,m}$
-\item Compute $\text{Var}_{BootMI} = (B-1)^{-1} \sum_{b=1}^B (\hat \theta_b - \hat \theta_{BM})^2$ where $\hat \theta_{BM} = B^{-1} \sum_{b=1}^B \hat \theta_b$.
-\end{itemize}
+\subsubsection{Multiple imputation under uncongeniality}
+
+Multiple imputation yields valid frequentist inferences when the imputation and analysis procedure are congenial~\citep{Meng1994};  the above procedure was derived for function-on-scalar multiple imputation for binary outcomes, which ignored the joint nature of recurrent event analysis in the presence of high frequency sensor data.
+Consider, for example, two non-events sampled five minutes apart.
+If there is missing data in the overlapping part of the temporal windows, then the above procedure imputes different values for each record.
+Moreover, the imputatoin does not depend on whether one of the two were an event rather than non-event.
+
+The main positive of the above procedure is that the above imputation framework is simple and approximately congenial when events are rare and the sampling rate is low as well as under the null of $\beta (s) \equiv 0$.
+The main negative is that the model is clearly uncongenial under many events and/or high sampling rates.
+Again, congeniality ensures goo frequentist coverage properties.
+A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
+
+To address this, we use the recommendation from~\cite{Bartlett2020} and consider a method that first bootstraps a sample from the randomly sampled dataset and then apply multiple imputation to each bootstrap sample.
+This general approach was originally proposed by~\cite{Shao1994} and~\cite{Little2002}.
+We suppose $B$ bootstraps and $M$ imputations per bootstrap; let $\hat \theta_{b,m}$ denote the estimator for the $m$th imputation of the $b$th bootstrap.
+The point estimator is given by $(B)^{-1} \sum_{b=1}^B \hat \theta_{b}$ where $\hat \theta_b = M^{-1} \sum_{m=1}^M \hat \theta_{b,m}$.
+To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b )^2$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B M \cdot (\hat \theta_b - \hat \theta)^2$ respectively.
+Then the estimator of the variance of $\hat \theta$ is given by $\left( \frac{B+1}{BM} \right) MSB - MSW/M$.
+
+The bootstrap followed by multiple imputation procedure has been studied extensively by~\cite{Bartlett2020} and is robust to uncongeniality.
+The main disadvantage of this approach is its considerable computational intensity.
+Recall likelihood calculations were computationally prohibitive by themselves, so combining with bootstrap and MI would further increase this large-scale computation.
+The random subsampling framework thus simplifies handling of missing data via connections to function-on-scalar multiple imputation by~\cite{Petrovich2018} as well as to bootstrap to handle uncongeniality by~\cite{Bartlett2020}.
 
@@
 
+In this section, we extend our model to the case of multilevel models with  functional regressors.
+Multilevel models are critical in mobile health where we expect high degree of individual variation.
+In this case, we suppose that there is individual variation $b_i \sim N(0, \sigma_{b}^2)$ such that the multilevel extension of our model~\eqref{eq:hazard} is given by
 \begin{equation}
@@
 \end{equation}
+where $Z_Pi,t$
+
+Ignoring the computational time of bootstrap sampling, the computational time of $200$ boostraps and $2$ imputations per boostrap leads to $35$ hours compared to $7$ hours.
+Again, this highlights the substantial benefits of
 
@@
 Each study participant wore an Empatica E4~\cite{empaticae3}, a medical-grade wearable device that offers real-time physiological data.
-One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified
+One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress.
 
@@
 
+\section{Optimal weights}
+
 \end{document}

### 2020-08-12 write-up/fda-recurrent.tex
about to edit missing data a bit shorter

@@
 The random subsampling framework thus simplifies handling of missing data via connections to function-on-scalar multiple imputation by~\cite{Petrovich2018} as well as to bootstrap to handle uncongeniality by~\cite{Bartlett2020}.
+Ignoring the computational time of bootstrap sampling, the computational time for the first choise in the simulation study with $B=200$ bootstraps and $M=2$ imputations per bootstrap leads to $35$ hours for a sampling rate of $0.5$ compared to $7$ hours for a sampling rate of $4$.
+This is simply to highlight the substantial benefits of the subsampling framework.
 
@@
 \end{equation}
-where $Z_Pi,t$
-
-Ignoring the computational time of bootstrap sampling, the computational time of $200$ boostraps and $2$ imputations per boostrap leads to $35$ hours compared to $7$ hours.
-Again, this highlights the substantial benefits of
+where $Z_{i,t} = M_{i,t} + C_{i,t} J_{\hat \psi, \phi}$.
+Lemma~\ref{lemma:logistic} implies that the random subsampling framework applied to equation~\eqref{eq:multilevel} leads to a penalized logistic mixed-effects model.
+As far as the authors are aware, the combination of mixed-effects and $L_2$-penalization on a subset of parameters has not been addressed in existing packages.
+To address this, we derive a penalized adaptive Gaussian quadarture (PADQ) for estimation.
+For conciseness, this derivation is presented in Appendix~\ref{app:penlogit}.
 
@@
 \section{Penalized logistic mixed-effects model}
+\label{app:penlogit}
 

### 2020-08-12 write-up/fda-recurrent.tex
tex

--- 2020-08-12 write-up/fda-recurrent.tex
@@
 
-To address this, we use the recommendation from~\cite{Bartlett2020} and consider a method that first bootstraps a sample from the randomly sampled dataset and then apply multiple imputation to each bootstrap sample.
-This general approach was originally proposed by~\cite{Shao1994} and~\cite{Little2002}.
-We suppose $B$ bootstraps and $M$ imputations per bootstrap; let $\hat \theta_{b,m}$ denote the estimator for the $m$th imputation of the $b$th bootstrap.
-The point estimator is given by $(B)^{-1} \sum_{b=1}^B \hat \theta_{b}$ where $\hat \theta_b = M^{-1} \sum_{m=1}^M \hat \theta_{b,m}$.
-To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b )^2$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B M \cdot (\hat \theta_b - \hat \theta)^2$ respectively.
-Then the estimator of the variance of $\hat \theta$ is given by $\left( \frac{B+1}{BM} \right) MSB - MSW/M$.
-
-The bootstrap followed by multiple imputation procedure has been studied extensively by~\cite{Bartlett2020} and is robust to uncongeniality.
-The main disadvantage of this approach is its considerable computational intensity.
-Recall likelihood calculations were computationally prohibitive by themselves, so combining with bootstrap and MI would further increase this large-scale computation.
-The random subsampling framework thus simplifies handling of missing data via connections to function-on-scalar multiple imputation by~\cite{Petrovich2018} as well as to bootstrap to handle uncongeniality by~\cite{Bartlett2020}.
+Here we highlight the recommendation from~\cite{Bartlett2020}, and employ a boostrap
+
+Here we
+Simple version compared to the von Hippel's boot MI proposal.
+\begin{itemize}
+\item Bootstrap $B$ times
+\item For $b = 1, \ldots, B$, impute $M$ times
+\item Let~$\hat \theta_b = M^{-1} \sum_m \hat \theta_{b,m}$
+\item Compute $\text{Var}_{BootMI} = (B-1)^{-1} \sum_{b=1}^B (\hat \theta_b - \hat \theta_{BM})^2$ where $\hat \theta_{BM} = B^{-1} \sum_{b=1}^B \hat \theta_b$.
+\end{itemize}
+
+\cite{vonHippel2019}
+
+To be careful, we investigate alternatives and assess impact.
+
+Reapply simulation and study only the coverage 95\% confidence interval under the null hypothesis.
+
+For a small loss in statistical efficiency, random sampling reduces the computation time considerably, which permits imputations closer to congeniality and less computation when performing bootstrap and multiple imputation.
 
@@
 \end{equation}
-where $Z_Pi,t$
-
-Ignoring the computational time of bootstrap sampling, the computational time of $200$ boostraps and $2$ imputations per boostrap leads to $35$ hours compared to $7$ hours.
-Again, this highlights the substantial benefits of
 
@@
 Each study participant wore an Empatica E4~\cite{empaticae3}, a medical-grade wearable device that offers real-time physiological data.
-One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress.
+One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified
 

### 2020-08-14 simulations/delta_incorrect/sim_delta.R
Delta incorrect sims constructed

@@ -1,0 +1,83 @@
+## Source the functions and necessary packages
+source("./sim_delta_functions.R")
+if(!require("glmnet")){install.packages('glmnet')} else{library('glmnet')}
+
+### Matern covariance parameters
+nu = 1/2 # controls smoothness
+sigma_sq = 1 # controls overall variance
+rho = 0.3 # controls correlation
+constant = sigma_sq / (gamma(nu) * 2^(nu-1))
+
+### User-day information
+### Each user day is t \in [0,1]
+### We use M dense equally spaced timepoints
+M = 1000 # Num of time points
+times = seq(0,1, length.out = M) # Equally spaced timepoints in [0,1]
+gap = diff(times)[1] # What is assumed gap times
+abs_diff = abs(outer(times, times, "-")) # Absolute distance
+x = sqrt(2 * nu) * abs_diff/ rho
+Cov_X = constant * x^nu * besselK(x, nu) # Matern Covariance
+diag(Cov_X) = 1
+C = chol(Cov_X) # Cholesky decomposition
+
+## Generate corresponding event times
+## Simulation assumes we want events near
+## (A) Want Up then Down pattern trends -> beta_1t
+## (B) Overall high mean in the window -> beta_2t
+# beta_1t = 30*c(rep(-1, 11), rep(1, 22), rep(-1, 11))
+# beta_1t = 50*sin(1:44/44*2*pi-pi/2)
+# beta_1t = 20*exp(0.1*1:44)/mean(exp(0.1*1:44))
+beta_1t = 100*sin(1:44/44*2*pi-pi/2)
+# beta_2t = 50*rep(1, 44)
+# plot(times[1:44],beta_1t+beta_2t)
+
+# test if there is at least one argument: if not, return an error
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args)==0) {
+  print("Using default value!")
+  window_length = length(beta_1t)
+} else if (length(args)==1) {
+  # default output file
+  window_length = args[1]
+}
+
+## Collect coefficients for first 35
+## Take covariance in the 44 timeslots
+## Change window length to be potentially incorrect
+Sigma = Cov_X[1:window_length,1:window_length]
+eig_Sigma = eigen(Sigma)
+K_x = min(35, window_length)  # Pick first 35 eigen-vectors or window_length
+cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
+
+# set.seed("131312")
[…]

### 2020-08-14 simulations/delta_incorrect/sim_delta_functions.R

@@ -1,0 +1,156 @@
+generate_userday <- function(Cov_X, C, times, eig_Sigma, beta_1t, sampling_rate, base_rate, window_length = NA) {
+  if (is.na(window_length)) { window_length = length(beta_1t) }
+  ## Generate X(t) for t \in [0,1]
+  X = t(C)%*%rnorm(M)  # Use Chol Decomp and Random Normals
+  # plot(times, X)
+  
+  ## Compute the integrals at each time t
+  result = rep(0,0)
+  event_nodes = (length(beta_1t)+1):length(times)
+  event_times = times[event_nodes]
+  for (t in event_nodes) {
+    temp_X = X[(t+1-length(beta_1t)):t]
+    temp_Xb = beta_1t*temp_X
+    term2 = (temp_Xb[1] + temp_Xb[length(temp_Xb)])/2 * gap + sum(temp_Xb[2:(length(temp_Xb)-1)]*gap)
+    result = c(result, term2)
+  }
+  
+  ## Calculate the logit based probabilities
+  probs = expit(base_rate+result)
+  # plot(event_times, probs)
+  Y = sapply(probs,function(p) rbinom(n=1,size = 1,p))
+  # abline(v = event_times[Y==1])
+  events = event_times[Y==1]
+  # plot(event_times, probs, type = 'l')
+  
+  ## Given time stamp I want to generate the summary
+  nonevents = event_times[rbinom(n = length(event_times), size = 1, prob = sampling_rate)==1]
+  nonevents = nonevents[!is.element(nonevents, events)]
+  
+  data = rep(0,0)
+  ## Construct event times
+  if(length(events) > 0) {
+    for (t in events) {
+      t_loc = which(is.element(times,t))
+      min_loc = max(t_loc+1-window_length, 1)
+      X_Delta = X[(min_loc):(t_loc)] # Past window of X
+      if(length(X_Delta) < window_length) {
+        X_Delta = c(rep(0,window_length - length(X_Delta)), X_Delta)
+      }
+      C_x = rep(0, K_x)  
+      for (i in 1:length(C_x)) {
+        C_x[i] = eig_Sigma$vectors[,i]%*%X_Delta
+      }
+      row_data = c(1,C_x)
+      data = rbind(data, row_data)
+    }
+  }
+  
+  ## Construct non-event times
+  if(length(nonevents) > 0) {
[…]

### 2020-08-14 write-up/fda-recurrent.tex
Update fda-recurrent.tex

--- 2020-08-12 write-up/fda-recurrent.tex
@@
 \usepackage[flushleft]{threeparttable}
-
 \usepackage{amsmath}
@@
 %}\limits}
-
 %\newcommand{\sam}[1]{{\color{blue}{#1}}}
 \newcommand{\walt}[1]{\textcolor{blue}{[WD:\ #1]}}
-
 \setlength{\textwidth}{6.5in}
@@
 %\setlength{\oddsidemargin}{.25in}
-
 \usepackage{setspace}
 \doublespacing
-
 \usepackage{soul,color, mathtools, fixmath}
@@
 \def\pr{\mathop{\text{pr}}\nolimits}
-
 \def\k{{\bf k}}
@@
 \def\ES{\mathop{\mathfrak{E}_S}\nolimits}
-
 \DeclareRobustCommand{\citeext}[1]{\citeauthor{#1} (\citeyear{#1})}
 \DeclareRobustCommand{\citeint}[1]{(\citeauthor{#1}, \citeyear{#1})}
-
 \newtheorem{thm}{Theorem}[section]
@@
 \newtheorem{rmk}[thm]{Remark}%\endlocaldefs
-
 % change default numbering for enumerate environment to be in parentheses
-
 \makeatletter
-
 \def\S{\mathcal{S}}
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
-
 \mathchardef\mhyphen="2D
-
 % display breaks
@@
 
-A core concern in mHealth is that sensor data can often be missing for intervals of time due to sensor wearing issues (e.g., sensor band becomes loose, turns off accidentally, data packets lost, or battery dies).
+Sensor data can often be missing for intervals of time due to sensor wearing issues.
 In the suicidal ideation case study, for example, there are 2139 self-identified moments of distress across all 91 participants.
@@
 Moreover, standard errors should reflect the uncertainty in these coefficients when missing data is prevalent.
-\cite{Goldsmith2011} suggest using best linear unbiased predictors (BLUP) or posterior modes in the mixed effects model to estimate~$c_{i,k} (t)$; however, one criticism would be when there is substantial variability in these estimates.
+\cite{Goldsmith2011} suggest using best linear unbiased predictors (BLUP) or posterior modes in the mixed effects model to estimate~$c_{i,k} (t)$; however, this is ineffective when there is substantial variability in these estimates.
 To deal with this, \cite{Crainiceanu2010} take a full Bayesian analysis.
 \cite{Yao2005} introduced PACE as an alternative frequentist method.
-However,~\cite{Petrovich2018} shows that for sparse, irregular longitudinal, the imputation model should not ignore the outcome variable~$Y_i (t)$.
+\cite{Petrovich2018} shows that for sparse, irregular longitudinal, the imputation model should not ignore the outcome variable~$Y_i (t)$.
 
@@
 and~$t \in \mathbb{R}_+$.
-These covariances can be estimated using the methods considered by
-\cite{ChenMuller2012}.
-% Our imputation procedure instead uses the
-% marginal covariance~$\Xi$ as a proxy for the true covariance
-% function~$\Sigma_t(\cdot, \cdot)$.
-% We found this to work well in simulations as~$\Sigma_t$ was relatively
-% constant across~$t$; however, this imputation strategy may be
-% inappropriate when the covariance structure changes significantly with
-% time~$t$.
-
-The imputation strategy generalizes the procedure introduced by~\cite{Petrovich2018} to the longitudinal functional data setting.
-This is possible due to the connection established with logistic regression by Lemma~\ref{lemma:logistic}.
-We could then consider a multiple imputation estimate with~$M$ imputations; the likelihood estimate is given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate from the~$m$th imputated dataset; the covariance matrix can be estimated by
-\[
-M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
-\sum_{m=1}^M (\hat \theta^{(m)} - \bar \theta) (\hat \theta^{(m)} -
-\bar \theta)
-\]
-where~$\hat \Xi^{(m)}$ is the estimator~\eqref{eq:fisher} for
-the~$m$th imputated dataset.
 
@@
 Multiple imputation yields valid frequentist inferences when the imputation and analysis procedure are congenial~\citep{Meng1994};  the above procedure was derived for function-on-scalar multiple imputation for binary outcomes, which ignored the joint nature of recurrent event analysis in the presence of high frequency sensor data.
-Consider, for example, two non-events sampled five minutes apart.
-If there is missing data in the overlapping part of the temporal windows, then the above procedure imputes different values for each record.
-Moreover, the imputatoin does not depend on whether one of the two were an event rather than non-event.
-
-The main positive of the above procedure is that the above imputation framework is simple and approximately congenial when events are rare and the sampling rate is low as well as under the null of $\beta (s) \equiv 0$.
-The main negative is that the model is clearly uncongenial under many events and/or high sampling rates.
-Again, congeniality ensures goo frequentist coverage properties.
+The main advantage of the above imputation framework is its simplicity and approximate congeniality when events are rare and the sampling rate is low.
+The main disadvantage is that the above framework is uncongenial under many events and/or high sampling rates.
+Again, congeniality ensures good frequentist coverage properties.
 A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
@@
 
-\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
-The above proof establishes asymptotic normality for all
-The main additional assumption is
-
-The argument
-
-\end{proof}
-
 \section{Penalized logistic mixed-effects model}
@@
 
+\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
+The above proof establishes asymptotic normality for all
+The main additional assumption is
+
+The argument
+
+\end{proof}
+
 \end{document}

### 2020-08-14 simulations/delta_incorrect/sim_delta.R
Fixed to work due to bug in construct J and rscript

@@ -32,14 +32,16 @@ beta_1t = 100*sin(1:44/44*2*pi-pi/2)
 # plot(times[1:44],beta_1t+beta_2t)
 
 # test if there is at least one argument: if not, return an error
+print("Made it to window length")
 args <- commandArgs(trailingOnly = TRUE)
 if (length(args)==0) {
   print("Using default value!")
   window_length = length(beta_1t)
 } else if (length(args)==1) {
   # default output file
-  window_length = args[1]
+  window_length = as.numeric(args[1])
 }
+print(paste("Window length is", window_length))
 
 ## Collect coefficients for first 35
 ## Take covariance in the 44 timeslots
@@ -71,7 +73,7 @@ col_names = as.vector(c("ids", "runtime", "rates", "mean_count", "var_count", pa
 for(rates in thinning_rates) {
   print(paste("On rate", rates*max_sampling_rate))
   subdataset = subsample_dataset(dataset, thinning_rate = rates)
-  intermediate_step = construct_J(times, eig_Sigma, subdataset)
+  intermediate_step = construct_J(times, eig_Sigma, subdataset, window_length)
   output = runglmnet(max_sampling_rate*rates, subdataset, intermediate_step$w, intermediate_step$Basis, epsilon = 0.0001)
   results = matrix(c(id, output$runtime, rates, agg_summary, output$beta), nrow = 1)
   colnames(results) = col_names

### 2020-08-18 write-up/fda-recurrent.tex
updated delta tex
Updated all until final section!

@@
 \usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
+\usepackage{multirow}
 %\usepackage[stretch=10]{microtype}
@@
 %\newcommand{\sam}[1]{{\color{blue}{#1}}}
-\newcommand{\walt}[1]{\textcolor{blue}{[WD:\ #1]}}
-\setlength{\textwidth}{6.5in}
-\setlength{\textheight}{8.5in}
-\setlength{\topmargin}{-0.25in}
-\setlength{\evensidemargin}{0in}
-\setlength{\oddsidemargin}{0in}
+% \newcommand{\walt}[1]{\textcolor{blue}{[WD:\ #1]}}
+% \setlength{\textwidth}{6.5in}
+% \setlength{\textheight}{8.5in}
+% \setlength{\topmargin}{-0.25in}
+% \setlength{\evensidemargin}{0in}
+% \setlength{\oddsidemargin}{0in}
 %\setlength{\evensidemargin}{.25in}
 %\setlength{\oddsidemargin}{.25in}
+\usepackage[margin=1in]{geometry}
 \usepackage{setspace}
 \doublespacing
+
 \usepackage{soul,color, mathtools, fixmath}
@@
 
-Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates.
-Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.
-For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$.
-If so, we can set~$\Delta$ larger to ensure the condition holds.
-In each simulation, we use $\Delta \in (18,21,24,27,30)$.
+A concern with the proposed approach is the selection of window-length, $\Delta$.
+Here we investigate the impact of misspecification of the window length for $\beta(t) = \beta_1 \cdot \sin \left( 2 \pi s / \Delta - \pi /2 \right)$ and the true window length ($\Delta^\star$) is set to 32-minutes.
+See Appendix~\ref{app:delta} for a similar discussion for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
+As in the previous simulation, we generate 1000 datasets per condition each consisting of 500 user-days.
+For each simulation, we analyze the data using window lengths $\Delta \in (26,29,32,35,37)$.
+
+When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$; however, we incur a penalty in finite samples, especially for settings where the function is far from zero near $t = \Delta$.
+We find the MISE increases the absolute error $|\Delta - \Delta^\star|$ increases, while the variance decreases as a function of $\Delta$.
+While the MISE increased for $\Delta < \Delta^\star$, the pointwise estimation error remains low for $t < \Delta$.
+This does not hold for $\Delta > \Delta^\star$, where instead we see parameter attenuation, i.e., a bias towards zero in the estimates at each $0 < t< \Delta$.
+To capture this, we define a partial MISE as $\frac{\Delta}{\tilde \Delta} \int_0^{\tilde \Delta} ( \hat \beta_{i,j} (s) - \hat \beta (s) )^2 ds$ where $\tilde \Delta = \min(\Delta, \Delta^\star)$, which is the MISE on the subset $0<t< \min (\Delta, \Delta^\star)$ and scaled for comparative purposes.
 
 \begin{table}[!th]
-\begin{tabular}{c c c c c c}
-& \multicolumn{5}{c}{Sampling rate} \\ \cline{2-6}
-$\Delta=$ & 0.5 & 1 & 5 & 25 & 50 \\ \hline
-18  & - & - & - & - & -  \\
-21 & - & - & - & - & -  \\
-24 & - & - & - & - & -  \\
-27 & - & - & - & - & -  \\
-30 & - & - & - & - & -  \\ \hline
+\begin{tabular}{c | c c c c c}
+& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
+$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
+\multirow{3}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
+ & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\
+  & P-MISE & 0.225 & 0.229 & 0.233 & 0.239 \\ \hline
+\multirow{3}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
+ & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\
+  & P-MISE & 0.107 & 0.109 & 0.111 & 0.117 \\ \hline
+\multirow{3}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
+ & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\
+  & P-MISE & 0.105 & 0.103 & 0.103 & 0.105 \\\hline
+\multirow{3}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
+ & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\
+  & P-MISE & 0.273 & 0.270 & 0.269, & 0.269 \\\hline
+\multirow{3}{*}{37} & MISE & 0.589 & 0.588 & 0.589 & 0.591 \\
+ & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\
+  & P-MISE & 0.473 & 0.470 & 0.469 & 0.470 \\\hline
 \end{tabular}
-\caption{XXX}
+\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
 \label{tab:mise}
@@
 Study data were graciously provided by Evan Kleiman and his study team ({\tt https://kleimanlab.org}).
-Each study participant wore an Empatica E4~\cite{empaticae3}, a medical-grade wearable device that offers real-time physiological data.
+Each study participant wore an Empatica E4~\citep{empaticae3}, a medical-grade wearable device that offers real-time physiological data.
+On each user-day, participants were asked to self-identify moments of suicidal distress.
+At these times, the participant pressed a button on the Empatica E4 device and the moment was recorded.
 One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress.
 
-can used to identify distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement.
+The
+distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement.
 The scientific hypothesis to be tested is whether a spike in EDA and/or gACC
@@
 
-At the end of the study period, the base rate of button pressing  rate was 292/488, or
-approximately 60\%.
+At the end of the study period, the base rate of button pressing  rate was 292/488, or approximately 60\%.
 
@@
 
+\section{Additional simulations and case study details}
+
+\subsection{Impact of $\Delta$}
+\label{app:delta}
+
+Here we investigate the impact of misspecification of the window length for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
+To do this, we generate 1000 datasets per condition each consisting of 500 user-days.
+For each simulation, we analyze the data using window lengths $\Delta \in (26,29,32,35,37)$.
+
+When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$ and therefore the true function is in the function space; however, in finite samples, we will incur a penalty, especially for settings where the function is far from zero near $t = \Delta$.
+
+we can expect a downward bias in parameter estimates.
+Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.
+For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$.
+If so, we can set~$\Delta$ larger to ensure the condition holds.
+
+\begin{table}[!th]
+\begin{tabular}{c | c c c c c}
+& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
+$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
+\multirow{2}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
+ & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\ \hline
+\multirow{2}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
+ & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\ \hline
+\multirow{2}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
+ & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\ \hline
+\multirow{2}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
+ & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\ \hline
+\multirow{2}{*}{37} & MISE & 0.589 & 0.588 & 0.589, & 0.591 \\
+ & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\ \hline
+\end{tabular}
+\caption{Mean-integrated squared error as a function of $\Delta^\star$ and sampling rate when true $\Delta = 30$ minutes.  Demonstrates the }
+\label{tab:mise}
+\end{table}
+
 \end{document}

### 2020-08-19 write-up/fda-recurrent.tex
Getting closer: need to get sims finished and the application done

@@
 On each user-day, participants were asked to self-identify moments of suicidal distress.
-At these times, the participant pressed a button on the Empatica E4 device and the moment was recorded.
+At these times, the participant was asked to press a button on the Empatica E4 device.
+The timestamp of the button press was recorded.
 One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress.
-
-The
-distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement.
-The scientific hypothesis to be tested is whether a spike in EDA and/or gACC
-
-To avoid, we focus on the study window 9AM to 8PM each study day.
-
-At the end of the study period, the base rate of button pressing  rate was 292/488, or approximately 60\%.
+In particular, the scientific question is whether there are early indicators of escalating distress by monitoring physiological correlates.
+
+A key concern is whether all moments of suicidal distress are recorded.
+To ensure this, clinical staff interviewed participants in the evening who were then asked to review their button press activity.
+Any events that were identified as incorrect button press activity were removed.
+An additional concern is that events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
+Due to (B), we define each study day to begin at 9AM and end at 8PM.
+Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
+At the end of the study period, the average number of button presses per day was XX with a standard deviation of XX.
+Figure~\ref{fig:buttons} presents the distribution of button presses as a function of day in study.
 
@@
 \caption{Button press distribution across study days (include number of people).}
+\label{fig:buttons}
 \end{figure}
 
-The focus here is on two EDA and the  prothrombin index, a composite
-blood coagulation index
-related to liver function, measured initially at three-month intervals
-and subsequently at roughly twelve-month intervals.
-The individual EDA and AI trajectories are highly variable, both in forward and when considering reverse time from event timepoints, which tends to obscure patterns and trends.
-In Figure~XX, the mean trajectory is plotted against time from recruitment for two patient groups placebo/prednisone and censored/not censored.
-Naturally, only those patients who are still alive are included in the average
-for that time.
-Figure XX hows the same plots in reverse alignment.
-While there are certain similarities in the two plots,
-the differences in temporal trends are rather striking.
-In particular, prothrombin levels in the six months prior to censoring are fairly stable,
+The focus here is on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
+The EDA measures are cleaned to  Due to high between study-day variation in EDA which may be due to external factors, we analyze EDA standardized per study-day.
+
+The individual EDA and gACC trajectories are highly variable which tends to obscure patterns and trends.
+In Figure~XX, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps and randomly sampled non-event timestamps.
+While there are certain similarities in the two plots, the differences in temporal trends are rather striking.
+In particular, we see sharp increases in EDA in the XX minutes prior to button presses,
 which is in marked contrast with levels in the six months prior to failure,
@@
 \caption{EDA and ACC plots for average at non-event and event times.}
+\label{fig:means}
 \end{figure}
 
-Inspection of the graphs for uncensored patients in the right panel of Figure XX suggests beginning with the simplest revival model in which the sequences
-for distinct patients are independent Gaussian with moments {\bf comparison over $\Delta$ and using imputation}
+Inspection of the graphs suggests setting $\Delta = XX$ for EDA and $XX$ for gACC and  Gaussian with moments {\bf comparison over $\Delta$ and using imputation} where the baseline rate is assumed to depend on
 
@@
 \end{document}
+
+1. Start simulation for appendix (delta wrong, beta2)
+2. Button Press Figure
+3. Compute wear minutes
+4. Standardize EDA
+5. Subsample EDA and gACC only at wear times between 9am and 8pm
+6. EDA/ACC figures
+7. Estimate with complete data
+8. Estimate with bootMI; make 1 plot with CIs for both compl and bootMI
+9. Estimate with heterogeneity
+10. Present

### 2020-08-21 write-up/fda-recurrent.tex
Updated writing, now allday casestudy

@@
 
-When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$ and therefore the true function is in the function space; however, in finite samples, we will incur a penalty, especially for settings where the function is far from zero near $t = \Delta$.
-
-we can expect a downward bias in parameter estimates.
-Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.
-For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$.
-If so, we can set~$\Delta$ larger to ensure the condition holds.
+When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$; however, we incur a penalty in finite samples, especially for settings where the function is far from zero near $t = \Delta$.
+We find the MISE increases and variance slightly decreases as $\Delta$ increases.
+While the MISE increased for $\Delta < \Delta^\star$, the pointwise estimation error remains low for $t < \Delta$.
+This does not hold for $\Delta > \Delta^\star$, where instead we see parameter attenuation, i.e., a bias towards zero in the estimates at each $0 < t< \Delta$, which is captured by the partial MISE.
 
@@
 $\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
-\multirow{2}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
- & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\ \hline
-\multirow{2}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
- & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\ \hline
-\multirow{2}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
- & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\ \hline
-\multirow{2}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
- & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\ \hline
-\multirow{2}{*}{37} & MISE & 0.589 & 0.588 & 0.589, & 0.591 \\
- & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\ \hline
+\multirow{3}{*}{26} & MISE & $5.9 \times 10^{-2}$ & $6.1 \times 10^{-2}$ & $6.6 \times 10^{-2}$ & $7.4 \times 10^{-2}$ \\
+ & Variance & $1.3 \times 10^{-2}$ & $1.6 \times 10^{-2}$ & $12.4 \times 10^{-2}$ & $3.6 \times 10^{-2}$ \\
+  & P-MISE & $7.1 \times 10^{-2}$ & $7.3 \times 10^{-2}$ & $7.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ \\ \hline
+\multirow{3}{*}{29} & MISE & $6.9 \times 10^{-2}$ & $7.0 \times 10^{-2}$ & $7.1 \times 10^{-2}$ & $7.8 \times 10^{-2}$ \\
+ & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.4 \times 10^{-2}$ \\
+  & P-MISE & $7.5 \times 10^{-2}$ & $7.5 \times 10^{-2}$ & $7.7 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
+\multirow{3}{*}{32} & MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\
+ & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $1.9 \times 10^{-2}$ & $3.3 \times 10^{-2}$ \\
+  & P-MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\ \hline
+\multirow{3}{*}{35} & MISE & $9.0 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $9.2 \times 10^{-2}$ \\
+ & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.2 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
+  & P-MISE & $8.5 \times 10^{-2}$ & $8.2 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
+\multirow{3}{*}{37} & MISE & $10.3 \times 10^{-2}$ & $10.1 \times 10^{-2}$ & $9.9 \times 10^{-2}$ & $10.1 \times 10^{-2}$ \\
+ & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
+  & P-MISE & $9.4 \times 10^{-2}$ & $9.2 \times 10^{-2}$ & $9.0 \times 10^{-2}$ & $9.0 \times 10^{-2}$ \\ \hline
 \end{tabular}
-\caption{Mean-integrated squared error as a function of $\Delta^\star$ and sampling rate when true $\Delta = 30$ minutes.  Demonstrates the }
-\label{tab:mise}
+\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
+\label{tab:appmise}
 \end{table}
@@
 
-1. Start simulation for appendix (delta wrong, beta2)
-2. Button Press Figure
-3. Compute wear minutes
+1. Start simulation for appendix (delta wrong, beta2): CHECK
+2. Button Press Figure:
+3. Compute wear minutes:
 4. Standardize EDA

### 2020-08-21 write-up/fda-recurrent.tex
That boing noise on linux is uber annoying...anyways another update to case study

@@
 \usepackage{multirow}
-%\usepackage[stretch=10]{microtype}
-%\usepackage{hyperref}
-%\def\struckint{\mathop{%
-%\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
-% {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
-%\mathpalette
-%{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
-%\halign\bgroup\hfill$}
-%{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
-%}\limits}
-%\newcommand{\sam}[1]{{\color{blue}{#1}}}
-% \newcommand{\walt}[1]{\textcolor{blue}{[WD:\ #1]}}
-% \setlength{\textwidth}{6.5in}
-% \setlength{\textheight}{8.5in}
-% \setlength{\topmargin}{-0.25in}
-% \setlength{\evensidemargin}{0in}
-% \setlength{\oddsidemargin}{0in}
-%\setlength{\evensidemargin}{.25in}
-%\setlength{\oddsidemargin}{.25in}
+\usepackage{floatrow}
+% Table float box with bottom caption, box width adjusted to content
+\newfloatcommand{capbtabbox}{table}[][\FBwidth]
+
 \usepackage[margin=1in]{geometry}
@@
 To address this, we derive a penalized adaptive Gaussian quadarture (PADQ) for estimation.
-For conciseness, this derivation is presented in Appendix~\ref{app:penlogit}.
+For conciseness, the algorithm is presented in Appendix~\ref{app:penlogit}.
 
@@
 
-During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital were enrolled in an observational study.
+During an eight month period in 2018, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital were enrolled in an observational study.
 Study data were graciously provided by Evan Kleiman and his study team ({\tt https://kleimanlab.org}).
@@
 Any events that were identified as incorrect button press activity were removed.
-An additional concern is that events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
+At the end of the study period, the average number of button presses per day was $1.91$ with a standard deviation of $2.23$.
+Figure~\ref{fig:buttons} presents the distribution of button presses as a function of hour in the day.
+This demonstrates low button-press counts prior to 7AM and a sharp drop off by 11PM.
+This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
 Due to (B), we define each study day to begin at 9AM and end at 8PM.
 Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
-At the end of the study period, the average number of button presses per day was XX with a standard deviation of XX.
-Figure~\ref{fig:buttons} presents the distribution of button presses as a function of day in study.
+
+\vspace{0.25cm}
+\begin{minipage}{\textwidth}
+  \begin{minipage}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{figs/buttonpresses.png}
+    \captionof{figure}{Button-presses per hour within the user-day}
+    \label{fig:buttons}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.49\textwidth}
+    \centering
+    \centering
+    \includegraphics[width=1.0\textwidth]{figs/recurrent_events_plot.png}
+    \captionof{figure}{User button-presses versus time in study}
+    \label{fig:buttons2}
+    \end{minipage}
+\end{minipage}
+\vspace{0.25cm}
+
+Figure~\ref{fig:buttons2} presents the raw button presses as a function of time in study for each user, showing the potential heterogeneity in button press rates between users and study days.
+To assess whether there is between user or between study day variation, Table~\ref{tab:anova} presents a two-way ANOVA decomposition of the button press counts as a function of participant and day in study.
+The ANOVA decomposition demonstrates high variation with day in study compared to across users in terms of button presses.
+
+\begin{table}[!th]
+\begin{tabular}{ccccccc}\hline
+      & DF & Sum Sq & Mean Sq & F value & Pr($>$F) \\ \hline
+        Participant & 88 & 13 & 13.31 & 2.71 & 0.10 \\
+        Day in Study & 75 & 54.71 & 54.71 & $8.8 \times 10^{-4}$ \\
+        Residuals & 989 & 4.91 & - \\ \hline
+      \end{tabular}  \captionof{table}{ANOVA decomposition of daily button press counts}
+      \label{tab:anova}
+\end{table}
+
+The focus here is on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
+Electrodermal activity can be significantly impacted by external factors (e.g., room temperature).
+To account for the high between user-day variation, we analyze EDA standardized per study-day.
+The individual EDA and gACC trajectories are highly variable which tends to obscure patterns and trends.
+In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps and randomly sampled non-event timestamps.
+While there are certain similarities in the two plots, the differences in temporal trends are rather striking.
+In particular, we see sharp increases in EDA in the XX minutes prior to button presses, which is in marked contrast with levels in the six months prior to failure, as seen in the lower pair of curves.
 
 \begin{figure}[!th]
-\caption{Button press distribution across study days (include number of people).}
-\label{fig:buttons}
+\rule{3cm}{3cm}%
+\caption{EDA and ACC plots for average at non-event and event times.}
+\label{fig:mean_edacc}
 \end{figure}
 
-The focus here is on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
-The EDA measures are cleaned to  Due to high between study-day variation in EDA which may be due to external factors, we analyze EDA standardized per study-day.
-
-The individual EDA and gACC trajectories are highly variable which tends to obscure patterns and trends.
-In Figure~XX, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps and randomly sampled non-event timestamps.
-While there are certain similarities in the two plots, the differences in temporal trends are rather striking.
-In particular, we see sharp increases in EDA in the XX minutes prior to button presses,
-which is in marked contrast with levels in the six months prior to failure,
-as seen in the lower pair of curves.
+Inspection of the graphs suggests setting $\Delta = XX$ for EDA and $XX$ for gACC and  Gaussian with moments {\bf comparison over $\Delta$ and using imputation} where the baseline rate is assumed to depend on
 
 \begin{figure}[!th]
-\caption{EDA and ACC plots for average at non-event and event times.}
-\label{fig:means}
-\end{figure}
-
-Inspection of the graphs suggests setting $\Delta = XX$ for EDA and $XX$ for gACC and  Gaussian with moments {\bf comparison over $\Delta$ and using imputation} where the baseline rate is assumed to depend on
-
-\begin{figure}[!th]
-\caption{$\beta (t)$ for ACC and EDA with 95\% CI.}
+\rule{3cm}{3cm}%
+\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is complete-case analysis, while dotted line is for }
+\label{}
 \end{figure}
@@
 1. Start simulation for appendix (delta wrong, beta2): CHECK
-2. Button Press Figure:
+2. Button Press Figure: Check
 3. Compute wear minutes:
-4. Standardize EDA
+4. Standardize EDA:
 5. Subsample EDA and gACC only at wear times between 9am and 8pm

### 2021-02-22 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 In classical joint models~\citep{Henderson2000, Tsiatis2004}, time-varying covariates~$x_i (t)$ are observed only intermittently at appointment times.
-In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$.
+In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$ that is a function of the unknown functional parameter $\beta(\cdot)$.
 That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
@@
 Sensor measurements occur multiple times per second.
-If the intensity rate is bounded above by $1$ and the unit time scale was hours.
-If we can subsample the data at~$10$ times per hour, then we have a lower bound on the efficiency of $0.909$.
+Suppose the intensity rate is bounded above by $1$ and the unit time scale was hours.
+If we then  subsample the data at a rate of~$10$ times per hour, then we have a lower bound on the efficiency of $0.909$.
 For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $10$.
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{figs/buttonpresses.png}
+%    \includegraphics[width=0.9\textwidth]{figs/buttonpresses.png}
     \captionof{figure}{Button-presses per hour within the user-day}
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figs/recurrent_events_plot.png}
+%    \includegraphics[width=1.0\textwidth]{figs/recurrent_events_plot.png}
     \captionof{figure}{User button-presses versus time in study}

### 2021-11-05 write-up/fda-recurrent.tex
updating paper

@@
 This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process.
-To ensure independent evolution, we employ the strategy of setting~$\epsilon (t) \equiv 0$.and treating~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
-The proposed inferential procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed (see Section~\ref{section:missingdata}).
+To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
+Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed (see Section~\ref{section:missingdata}).
 
@@
 At a subsampled time~$t$, the \emph{windowed covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq \Delta}$ and counting process features~$g_t (H_{i,t}^N)$ are observed.
-Optimal choice of~$\pi_i (t)$ is beyond the scope of this paper; however, simulation studies have suggested setting proportional to the hazard function~$h_i (t; \theta)$.
+Optimal choice of~$\pi_i (t)$ is beyond the scope of this paper; however, simulation studies have suggested setting the subsampling rate proportional to the hazard function~$h_i (t; \theta)$.
 
 An estimator is design-unbiased if its expectation is equal to that parameter under the probability distribution induced by the sampling design~\citep{Cassel1977}.
-Let~$\bfD_i \subset [0,t_i]$ denote the set of subsampled points.
-Note, by construction, this set is distinct from the set of event times, i.e.,~$\bfT_i \cap \bfD_i = \emptyset$.
-Under subsampling via $\pi_i (t)$, one may compute a Horvitz-Thompson estimator of the derivative of the cumulative hazard:
-\[
-\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in \bfD_i} \frac{
-  h^{(1)}_i (u; \theta) }{ \pi_i (u) }
-\]
+Let~$\bfD_i \subset [0,t_i]$ denote the random set of subsampled points.
+Note, by construction, this random set is distinct from the set of event times with probabilty one, i.e.,~$\pr( \bfT_i \cap \bfD_i = \emptyset) = 1$.
+Under subsampling via $\pi_i (t)$, one may compute a Horvitz-Thompson estimator of the derivative of the cumulative hazard $\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in \bfD_i} h^{(1)}_i (u; \theta)/\pi_i (u)$.
 An alternative design-unbiased estimator of the derivative of the cumulative hazards is given by
@@
 We use the sandwich smoother~\citep{Xiao2013} to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
-Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009} bivariate tensor product splines~\citep{Wood2006} and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
+Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009}, bivariate tensor product splines~\citep{Wood2006}, and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
 The sandwich smoother was chosen for its computational efficiency and estimation accuracy.
@@
 If we then  subsample the data at a rate of~$10$ times per hour, then we have a lower bound on the efficiency of $0.909$.
-For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $10$.
+For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $10$ per hour.
 While the computational complexity of logistic regression is linear in the number of samples, we get $1440$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.

### 2021-11-05 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 h_0 (t; \gamma) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + \int_{t-\Delta}^t x_{i,1} (s) \beta_1(s) ds + \cdots + \int_{t-\Delta}^t x_{i,L} (s) \beta_2(s) ds  \right)
+  + \int_{t-\Delta_1}^t x_{i,1} (s) \beta_1(s) ds + \cdots + \int_{t-\Delta_L}^t x_{i,L} (s) \beta_L(s) ds  \right)
 \end{equation}
@@
 Let $\sum_{k=1}^\infty \hat \lambda^{(y)}_{k,l} \hat \psi^{(y)}_{k,l} (s) \hat \psi_{k,l}(t)$ be the spectral decomposition of $\hat \Sigma_{y,l}$.
-Then $x_{i,j}(t)$ is approximated using a truncated Karhunen-Loeve
-$$
-\int_{t-\Delta}^t X_{l} (t,s) \beta(t) = \left[M_{l,t} + {\bf c}_l (t)^\top J_{\hat \psi_l^{(y)}, \phi} \right] {\bf b}_l
-$$
-Note that we express each coefficient function in the same spline basis; however, different basis expansions are used for each health process as defined by the marginal covariance calculation in Section~\ref{sec:margcov}.
+Then $x_{i,j}(t)$ is approximated using a truncated Karhunen-Loeve $\int_{t-\Delta_l}^t X_{l} (t,s) \beta_l(t) = \left[M_{l,t} + {\bf c}_l (t)^\top J_{\hat \psi_l^{(y)}, \phi_l} \right] {\bf b}_l$.
+% Note that we express each coefficient function in the same spline basis; however, different basis expansions are used for each health process as defined by the marginal covariance calculation in Section~\ref{sec:margcov}.
 
@@
 and~$t \in \mathbb{R}_+$.
+To account for overlap between $\Delta$-windows for $t \in \bfT_i \cup \bfD_i$ - i.e., avoid imputing values at a particular time to different values -- imputation is performed sequentially over this set.
+%$\{ (t,t') : t, t^\prime \in \bfT_i \cup \bfD_i \& |t - t'| < \Delta \}$
 
@@
 
-Multiple imputation yields valid frequentist inferences when the imputation and analysis procedure are congenial~\citep{Meng1994};  the above procedure was derived for function-on-scalar multiple imputation for binary outcomes, which ignored the joint nature of recurrent event analysis in the presence of high frequency sensor data.
+Multiple imputation yields valid frequentist inferences when the imputation and analysis procedure are congenial~\citep{Meng1994};  the above procedure is derived for function-on-scalar multiple imputation for binary outcomes, which ignored the joint nature of recurrent event analysis in the presence of high frequency sensor data.
 The main advantage of the above imputation framework is its simplicity and approximate congeniality when events are rare and the sampling rate is low.
@@
 
-To address this, we use the recommendation from~\cite{Bartlett2020} and consider a method that first bootstraps a sample from the randomly sampled dataset and then apply multiple imputation to each bootstrap sample.
+To address this, we use the recommendation from~\cite{Bartlett2020} and consider a method that first bootstraps a sample from the dataset and then apply multiple imputation to each bootstrap sample.
 This general approach was originally proposed by~\cite{Shao1994} and~\cite{Little2002}.
@@
 The random subsampling framework thus simplifies handling of missing data via connections to function-on-scalar multiple imputation by~\cite{Petrovich2018} as well as to bootstrap to handle uncongeniality by~\cite{Bartlett2020}.
-Ignoring the computational time of bootstrap sampling, the computational time for the first choise in the simulation study with $B=200$ bootstraps and $M=2$ imputations per bootstrap leads to $35$ hours for a sampling rate of $0.5$ compared to $7$ hours for a sampling rate of $4$.
-This is simply to highlight the substantial benefits of the subsampling framework.
+Ignoring the computational time of bootstrap sampling, the computational time for the first choise in the simulation study with $B=200$ bootstraps and $M=2$ imputations per bootstrap leads to $35$ hours for a sampling rate of $0.5$ compared to $7$ hours for a sampling rate of $4$, which highlights the benefits of the proposed framework.
 
@@
 
-In this section, we extend our model to the case of multilevel models with  functional regressors.
-Multilevel models are critical in mobile health where we expect high degree of individual variation.
-In this case, we suppose that there is individual variation $b_i \sim N(0, \sigma_{b}^2)$ such that the multilevel extension of our model~\eqref{eq:hazard} is given by
+The approach can be extended to multilevel models with functional regressors, which are critical in mobile health where a high degree of individual variation is often observed.
+Let $b_i \sim N(0, \sigma_{b}^2)$, then the multilevel extension of~\eqref{eq:hazard} is
 \begin{equation}

### 2021-11-08 preprocessing/acc_preprocessing/acc_measure_allFiles.R
acc and eda preprocessing

@@ -1,0 +1,36 @@
+# Code to compute AI for files #
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the AI data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+library(tidyverse)
+library(lubridate)
+source('./acc_measure_single.R')
+
+
+## ------------------ Compute AI for all observations ------------------ ##
+setwd("..")
+
+epch = 10 # integer, the length of the time window (in seconds) to compute AI
+n_obs = 91  # the largest number of observations (files)
+acc_dir = "../R21_Study - ACC" # the directory where the accelerometer data is
+for (i in 1:n_obs){
+  id = 1000 + i 
+  input_name = paste0(acc_dir, "/", id, "_acc.rds")
+  output_name = paste0("./", id, "_AI.rds")
+  
+  if (file.exists(input_name)){
+    acc = readRDS(input_name)
+    acc$time = as_datetime(acc$ts/1000) # transform timestamp from numeric to required data type
+    rate = round(1000 / (acc$ts[2] - acc$ts[1]))
+    acc_ai = acc_AI(acc, rate, epch)
+    saveRDS(acc_ai, file = output_name)
+  }
+}
+
+
+
+
+
+

### 2021-11-08 preprocessing/acc_preprocessing/acc_measure_single.R

@@ -1,0 +1,72 @@
+# Function to compute AI #
+
+## -------------------- Note -------------------- ##
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the AI data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+## Function name:
+##    acc_AI: compute AI for a single obs.
+
+## Install the required package to compute Activity Index from Accelerometer data:
+install.packages("devtools")
+devtools::install_github("javybai/ActivityIndex")
+
+## The function:
+acc_AI = function(data, rate, epch){
+  # Input:
+  #   data: accelerometry data for one person with variables 
+  #         time: record time in timestamp data type ("POSIXct", "POSIXt")
+  #         acc_x, acc_y, acc_z: position variables
+  #         Device: device name
+  #   rate: the rate of the accelerometry data
+  #   epch: integer, the length of the time window (in seconds) to compute AI
+  # Output:
+  #   measure: dataframe for AI, including Priod (which continuous period the record is in), Device,
+  #                                        timestamp (position of the record in that period) and AI
+  
+  require(lubridate) # arrange timestamp
+  require("ActivityIndex")
+  require(tidyverse)
+  
+  acc_all = data.frame(transmute(data, time, Device=E4_serial, acc_x, acc_y, acc_z))
+  devices = unique(acc_all$Device)
+  measure = data.frame(Period=NULL, Device=NULL, timestamp=NULL, AI=NULL)
+  
+  # process the data from different devices seperately due to the timestamp overlap 
+  for (d in 1:length(devices)){ 
+    acc = filter(acc_all, Device == devices[d]) %>%
+      arrange(time)
+    
+    # view loss of record larger than 1/32 (about 0.032) second as discontinuity.
+    cutpoint = c(0, which(diff(acc$time) > 0.032), dim(acc)[1])
+    
+    # process the data in each consecutive period, 
+    # because the function "computeActivityIndex" does not recognize time
+    for (i in 1:(length(cutpoint) - 1)){
+      # drop a few data on the tail of the period to make sure 
+      # the number of data entries is a multiple of rate*epch,
+      # as rate*epch entries in the acc data will be used to compute an AI entry
[…]

### 2021-11-08 preprocessing/acc_preprocessing/bp_ai.R

@@ -1,0 +1,94 @@
+## Code for the mean AI plots 30 minuets backward from the button press
+
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the scaled AI data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+library(tidyverse)
+library(lubridate)
+
+
+setwd('..')
+
+# read in the button press data and transform the data type of the timestamp
+bp = readRDS('../R21_Study - tags/button_presses.RDS')
+bp = transmute(bp, ID, time=as_datetime(ts))
+
+
+#### Match button presses and the previous AI data ####
+# data frame to store the matched bp and AI
+bp_ai = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, 
+                   Device=NULL, AI_num=NULL, AI_time=NULL, AI=NULL)
+
+n_obs = 91 # the largest number of observations (files)
+for (i in 1:n_obs){
+  id = 1000 + i
+  input_name = paste0('./', id, '_AI.rds')
+  
+  ## check the progress of the loop and save temporary result ##
+  if (i%%10 == 0){
+    print(i)
+    saveRDS(bp_ai, file = './methods/summary_data/bp_ai.rds')
+  }
+  
+  ## match bp and AI for this observation
+  if (file.exists(input_name)){
+    sub_bp = filter(bp, ID==id)
+    ai = readRDS(input_name)
+    device = unique(ai$Device)
+    
+    # process the data from different devices seperately due to the timestamp overlap 
+    for (d in 1:length(device)){
+      sub_ai =  filter(ai, Device==device[d])
+      for (j in 1:dim(sub_bp)[1]){
+        bp_time = sub_bp$time[j]
+        # AI_time in [BP_time-30min, BP_time)
+        ai_per_device_bp = filter(sub_ai, timestamp<bp_time, timestamp>=(bp_time-30*60))
+        if (dim(ai_per_device_bp)[1] > 0){
+          sub_bp_ai = transmute(ai_per_device_bp, ID=rep(id,n()), 
+                                BP_num=rep(j,n()), BP_time=rep(bp_time, n()), 
+                                Device=rep(device[d],n()), AI_num=1:n(), AI_time=timestamp, AI)
[…]

### 2021-11-08 preprocessing/acc_preprocessing/summary_and_plots_AI.R

@@ -1,0 +1,219 @@
+# Summary and plots on AI and pause time
+
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the scaled AI data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+library(tidyverse)
+library(lubridate)
+
+setwd('..')
+
+## ------------------ Summary for AI and Pause time------------------ ##
+n_obs = 91  # the largest number of observations (files)
+epch = 10   # the length of the time window (in seconds) used to compute AI
+
+# summary of AI for each person
+summary_AI = data.frame(matrix(0, nrow = n_obs, ncol = 14))
+colnames(summary_AI) = c('ID', 'AI_mean', 'AI_sd', 'AI_max', 'AI_min', 
+                         'AI.25', 'AI.50', 'AI.75', 'AI.95', 'AI.99', 
+                         'Days', 'Continuous Periods', 
+                         'Pause_max(hour)', 'Pause_min(hour)')
+# pause and period (after the pause) length in seconds
+all_pause = data.frame(ID=integer(), pause_num=integer(), pause=double(), period=double()) 
+
+
+none_ind = 0
+for (i in 1:n_obs){
+  id = 1000 + i 
+  input_name = paste0("./", id, "_AI.rds")
+  
+  ## check the progress of the loop##
+  if (i%%10 == 0){
+    print(i)
+  }
+  
+  if (file.exists(input_name)){
+    acc_ai = readRDS(input_name)
+    summary_AI[i, 'ID'] = id
+    summary_AI[i, 'AI_mean'] = mean(acc_ai$AI)
+    summary_AI[i, 'AI_sd'] = sd(acc_ai$AI)
+    summary_AI[i, 'AI_max'] = max(acc_ai$AI)
+    summary_AI[i, 'AI_min'] = min(acc_ai$AI)
+    summary_AI[i, 'AI_max'] = max(acc_ai$AI)
+    summary_AI[i, c('AI.25', 'AI.50', 'AI.75', 'AI.95', 'AI.99')] = 
+      quantile(acc_ai$AI, probs = c(0.25, 0.5, 0.75, 0.95, 0.99))
+    summary_AI[i, 'Days'] = length(unique(as.Date(acc_ai$timestamp)))
+    summary_AI[i, 'Continuous Periods'] = length(unique(acc_ai$Period))
+    
+    # pause information
[…]

### 2021-11-08 preprocessing/eda_preprocessing/bp_eda.R

@@ -1,0 +1,95 @@
+## Code for the mean EDA plots backward from the button press
+
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the scaled EDA data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+library(tidyverse)
+library(lubridate)
+
+
+setwd('..')
+
+
+# read in the button press data and transform the data type of the timestamp
+bp = readRDS('../R21_Study - tags/button_presses.RDS')
+bp = transmute(bp, ID, time=as_datetime(ts))
+
+
+#### Match button presses and the previous EDA (scaled) data ####
+bp_eda = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, Device=NULL, 
+                    EDA_time=NULL, EDA=NULL, EDA_num=NULL)
+
+n_obs = 91 # the largest number of observations (files)
+for (i in 1:n_obs){
+  id = 1000 + i
+  input_name = paste0('./', id, '_EDA_scaled.rds')
+  
+  ## check the progress of the loop and save temporary result ##
+  if (i%%10 == 0){
+    print(i)
+    saveRDS(bp_eda, file = './methods/summary_data/bp_eda.rds')
+  }
+  
+  ## match bp and EDA for this observation
+  ## observations 1010 and 1012 should be deleted since they do not have ACC data
+  if (file.exists(input_name) & i!=10 & i!=12){ 
+    sub_bp = filter(bp, ID==id)
+    eda = readRDS(input_name)
+    for (j in 1:dim(sub_bp)[1]){
+      bp_time = sub_bp$time[j]
+      # EDA_time in [BP_time-30min, BP_time)
+      sub_eda = filter(eda, timestamp<bp_time, timestamp>=(bp_time-30*60))
+      if (dim(sub_eda)[1] > 0){
+        sub_bp_eda = sub_eda %>%
+          transmute(ID=rep(id,n()), BP_num=rep(j,n()), BP_time=rep(bp_time, n()), 
+                    Device, EDA_time=timestamp, EDA=EDA_scaled) %>%
+          group_by(Device) %>%
+          mutate(EDA_num = 1:n()) %>%
+          ungroup() %>%
+          arrange(Device, EDA_num)
+        bp_eda = rbind(bp_eda, sub_bp_eda)
+      }
+    }
[…]

### 2021-11-08 preprocessing/eda_preprocessing/eda_scale_allFiles.R

@@ -1,0 +1,49 @@
+# Code to scale EDA for files #
+
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the scaled EDA data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+#### For each user:
+####     EDA_scaled = (EDA - mean_by_date) / std_by_date
+
+library(tidyverse)
+library(lubridate)
+
+setwd('..')
+
+## ------------------ Scale EDA for all observations ------------------ ##
+
+n_obs = 91  # the largest number of observations (files)
+eda_dir = "../R21_Study - EDA" # the directory where the raw EDA data is
+for (i in 1:n_obs){
+  id = 1000 + i 
+  input_name = paste0(eda_dir, "/", id, "_EDA.rds")
+  output_name = paste0("./", id, "_EDA_scaled.rds")
+  
+  if (file.exists(input_name)){
+    eda_raw = readRDS(input_name)
+    eda_scaled = mutate(eda_raw, time=as_datetime(ts/1000)) %>%  # change data type of the timestamp
+      mutate(date=as_date(time)) %>%  # add date column
+      transmute(date, time, Device=E4_serial, EDA=EDA_HighLowPass) %>%
+      drop_na() %>%
+      arrange(Device, time) %>%
+      group_by(Device) %>%  # data on different devices may have time overlap
+      mutate(time_diff = c(-1, diff(time))) %>%
+      ungroup() %>%
+      filter(time_diff != 0) %>%  # drop duplicate data
+      group_by(date) %>%
+      mutate(EDA_scaled = scale(EDA)) %>%
+      ungroup() %>%
+      arrange(Device, time) %>%
+      transmute(timestamp=time, Device, EDA_scaled)
+    saveRDS(eda_scaled, file = output_name)
+  }
+}
+
+
+## drop trouble data (EDA_HighLowPass = 0) in observation 1077
+eda_scaled = readRDS('./1077_EDA_scaled.rds')
+eda_scaled = drop_na(eda_scaled)
+saveRDS(eda_scaled, file = './1077_EDA_scaled.rds')

### 2021-11-08 preprocessing/eda_preprocessing/summary_and_plots_EDA.R

@@ -1,0 +1,185 @@
+## Summary and plots on EDA and its pause time
+
+## ---- Please place this file in a folder named "methods" 
+##      under the directory where the scaled EDA data is stored. ----
+
+## Recent update: May 27, 2021, by Xinrui Wu
+
+library(tidyverse)
+library(lubridate)
+
+setwd('..')
+## ------------------ Summary for Pause time in EDA------------------ ##
+n_obs = 91  # the largest number of observations (files)
+td = 0.25 # time difference between 2 consecutive data
+
+#  pause and period (after the pause) length in seconds
+all_pause = data.frame(ID=integer(), pause_num=integer(), pause_len=double(), 
+                       period_len=double(), num_na=integer()) 
+
+for (i in 1:n_obs){
+  id = 1000 + i 
+  input_name = paste0("./", id, "_EDA.rds")
+  
+  ## check the progress of the loop and save temporary result ##
+  if (i%%10 == 0){
+    print(i)
+    saveRDS(all_pause, file = './methods/summary_data/pauses_EDA.rds')
+  }
+  
+  if (file.exists(input_name)){
+    eda = readRDS(input_name)
+    eda_hlp = eda %>%
+      transmute(ID=Participant, time = as_datetime(eda$ts/1000), EDA = EDA_HighLowPass)
+    
+    # pause information
+    time_diff = round(diff(as.numeric(eda_hlp$time)),2)
+    pause_time = which(time_diff > td)
+    n_pause = length(pause_time)
+    
+    eda_hlp = eda_hlp %>%
+      mutate(period=rep(1, dim(eda_hlp)[1]))
+    if(n_pause > 0){
+      for (i in 1:n_pause){
+        breaks = c(pause_time, dim(eda_hlp)[1])
+        eda_hlp$period[(breaks[i]+1):breaks[i+1]] = i+1
+      }
+    }
+    sub_pause = eda_hlp %>%
+      group_by(ID, period) %>%
+      summarise(period_len = as.numeric(difftime(tail(time,1),time[1], units = 'secs')), 
+                num_na = sum(is.na(EDA))) %>%
+      ungroup() %>%
+      transmute(ID, pause_num=period-1, pause_len = c(0, time_diff[which(time_diff > td)]),
[…]

### 2021-11-08 preprocessing/acc_preprocessing/bp_ai.R
Update bp_ai.R
Fixing to work in current config

@@ -8,10 +8,9 @@
 library(tidyverse)
 library(lubridate)
 
+setwd('Z:/SI_data/R21_Study - ACC - AI/')
 
-setwd('..')
-
-# read in the button press data and transform the data type of the timestamp
+# read in the button press data and transform the data type of the time stamp
 bp = readRDS('../R21_Study - tags/button_presses.RDS')
 bp = transmute(bp, ID, time=as_datetime(ts))
 

### 2021-11-08 preprocessing/acc_preprocessing/bp_ai.R
Fixing up preprocessing to get plots

@@ -17,49 +17,52 @@ bp = transmute(bp, ID, time=as_datetime(ts))
 
 #### Match button presses and the previous AI data ####
 # data frame to store the matched bp and AI
-bp_ai = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, 
-                   Device=NULL, AI_num=NULL, AI_time=NULL, AI=NULL)
-
-n_obs = 91 # the largest number of observations (files)
-for (i in 1:n_obs){
-  id = 1000 + i
-  input_name = paste0('./', id, '_AI.rds')
+if(!file.exists("../summary_data/bp_ai.rds")) {
+  bp_ai = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, 
+                     Device=NULL, AI_num=NULL, AI_time=NULL, AI=NULL)
   
-  ## check the progress of the loop and save temporary result ##
-  if (i%%10 == 0){
-    print(i)
-    saveRDS(bp_ai, file = './methods/summary_data/bp_ai.rds')
-  }
-  
-  ## match bp and AI for this observation
-  if (file.exists(input_name)){
-    sub_bp = filter(bp, ID==id)
-    ai = readRDS(input_name)
-    device = unique(ai$Device)
+  n_obs = 91 # the largest number of observations (files)
+  for (i in 1:n_obs){
+    print(paste("Current participant", i))
+    id = 1000 + i
+    input_name = paste0('./', id, '_AI.rds')
     
-    # process the data from different devices seperately due to the timestamp overlap 
-    for (d in 1:length(device)){
-      sub_ai =  filter(ai, Device==device[d])
-      for (j in 1:dim(sub_bp)[1]){
-        bp_time = sub_bp$time[j]
-        # AI_time in [BP_time-30min, BP_time)
-        ai_per_device_bp = filter(sub_ai, timestamp<bp_time, timestamp>=(bp_time-30*60))
-        if (dim(ai_per_device_bp)[1] > 0){
-          sub_bp_ai = transmute(ai_per_device_bp, ID=rep(id,n()), 
-                                BP_num=rep(j,n()), BP_time=rep(bp_time, n()), 
-                                Device=rep(device[d],n()), AI_num=1:n(), AI_time=timestamp, AI)
-          bp_ai = rbind(bp_ai, sub_bp_ai)
[…]

### 2021-11-08 preprocessing/acc_preprocessing/bp_ai.R
updating 30 min plots for paper!

@@ -1,10 +1,4 @@
 ## Code for the mean AI plots 30 minuets backward from the button press
-
-## ---- Please place this file in a folder named "methods" 
-##      under the directory where the scaled AI data is stored. ----
-
-## Recent update: May 27, 2021, by Xinrui Wu
-
 library(tidyverse)
 library(lubridate)
 

### 2021-11-08 preprocessing/eda_preprocessing/bp_eda.R

@@ -1,15 +1,9 @@
 ## Code for the mean EDA plots backward from the button press
-
-## ---- Please place this file in a folder named "methods" 
-##      under the directory where the scaled EDA data is stored. ----
-
-## Recent update: May 27, 2021, by Xinrui Wu
-
 library(tidyverse)
 library(lubridate)
 
 
-setwd('..')
+setwd('Z:/SI_data/R21_Study - EDA - scaled/')
 
 
 # read in the button press data and transform the data type of the timestamp
@@ -18,47 +12,52 @@ bp = transmute(bp, ID, time=as_datetime(ts))
 
 
 #### Match button presses and the previous EDA (scaled) data ####
-bp_eda = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, Device=NULL, 
-                    EDA_time=NULL, EDA=NULL, EDA_num=NULL)
-
-n_obs = 91 # the largest number of observations (files)
-for (i in 1:n_obs){
-  id = 1000 + i
-  input_name = paste0('./', id, '_EDA_scaled.rds')
+if(!file.exists("../summary_data/bp_eda.rds")) {
+  bp_eda = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, Device=NULL, 
+                      EDA_time=NULL, EDA=NULL, EDA_num=NULL)
   
-  ## check the progress of the loop and save temporary result ##
-  if (i%%10 == 0){
-    print(i)
-    saveRDS(bp_eda, file = './methods/summary_data/bp_eda.rds')
-  }
-  
-  ## match bp and EDA for this observation
-  ## observations 1010 and 1012 should be deleted since they do not have ACC data
-  if (file.exists(input_name) & i!=10 & i!=12){ 
-    sub_bp = filter(bp, ID==id)
-    eda = readRDS(input_name)
-    for (j in 1:dim(sub_bp)[1]){
-      bp_time = sub_bp$time[j]
-      # EDA_time in [BP_time-30min, BP_time)
-      sub_eda = filter(eda, timestamp<bp_time, timestamp>=(bp_time-30*60))
-      if (dim(sub_eda)[1] > 0){
-        sub_bp_eda = sub_eda %>%
-          transmute(ID=rep(id,n()), BP_num=rep(j,n()), BP_time=rep(bp_time, n()), 
-                    Device, EDA_time=timestamp, EDA=EDA_scaled) %>%
-          group_by(Device) %>%
[…]

### 2021-11-08 preprocessing/eda_preprocessing/bp_eda.R
putting in plots

@@ -50,9 +50,9 @@ if(!file.exists("../summary_data/bp_eda.rds")) {
       }
     }
   }
-  saveRDS(bp_eda, file = './methods/summary_data/bp_eda.rds')
+  saveRDS(bp_eda, file = '../summary_data/bp_eda.rds')
 } else {
-  bp_eda = readRDS(file = './methods/summary_data/bp_eda.rds')
+  bp_eda = readRDS(file = '../summary_data/bp_eda.rds')
 }
 
 
@@ -62,9 +62,8 @@ bp_eda_summary = bp_eda %>%
   summarise(EDA_entries = n())
 max(bp_eda_summary$EDA_entries) # should be no larger than 7200 = 30*60*4
 
-
 #### Draw the mean plot and the boxplot ####
-bp_eda = bp_eda_read %>%
+bp_eda = bp_eda %>%
   group_by(ID, BP_num, Device) %>%
   mutate(time_to_bp = max(EDA_num)+1-EDA_num) %>%
   ungroup() %>%
@@ -75,20 +74,26 @@ eda_mean_plot =  bp_eda %>%
   ungroup()
 
 
+png("../figures/smoothed_eda.png",
+    width = 480, height = 480, units = "px", pointsize = 12)
 # mean plot
+eda_mean_plot$time_to_bp = eda_mean_plot$time_to_bp/4/60
 plot(x=eda_mean_plot$time_to_bp, y=eda_mean_plot$EDA_mean, axes=F, cex = 0.25, 
      main = "Mean of EDA across observations", ylab = 'mean(EDA)', 
-     xlab = "time before a button press / 0.25 secs")
+     xlab = "Time before Button Press (in minutes)")
 axis(side = 1)
 axis(side = 2)
+smoothed_mean = loess(EDA_mean ~ time_to_bp, data = eda_mean_plot, span = 1/3)
+lines(eda_mean_plot$time_to_bp, smoothed_mean$fitted, col= "red", lwd = 2)
+dev.off()
 
 
-# boxplot for time points in the first minute before a button press
-sub_bp_eda = filter(bp_eda, ID!=1009, time_to_bp<=240)  # 1009 has trouble raw EDA, at about 240
-boxplot(EDA~time_to_bp, data=sub_bp_eda, cex=0.08, axes = F, 
-        main = "Boxplots of EDA across observations", ylab = 'EDA',
-        xlab = "time before a button press / 0.25 secs")
-axis(side = 1)
-axis(side = 2)
[…]

### 2021-11-09 preprocessing/eda_preprocessing/bp_eda.R
updating the cleaning of EDA a bit
Going to max at 20 cause HL pass doesnt seem to work well

@@ -62,6 +62,8 @@ bp_eda_summary = bp_eda %>%
   summarise(EDA_entries = n())
 max(bp_eda_summary$EDA_entries) # should be no larger than 7200 = 30*60*4
 
+bp_eda = subset(bp_eda, abs(EDA) < 5)
+
 #### Draw the mean plot and the boxplot ####
 bp_eda = bp_eda %>%
   group_by(ID, BP_num, Device) %>%
@@ -75,10 +77,10 @@ eda_mean_plot =  bp_eda %>%
 
 
 png("../figures/smoothed_eda.png",
-    width = 480, height = 480, units = "px", pointsize = 12)
+    width = 800, height = 600, units = "px", pointsize = 18)
 # mean plot
 eda_mean_plot$time_to_bp = eda_mean_plot$time_to_bp/4/60
-plot(x=eda_mean_plot$time_to_bp, y=eda_mean_plot$EDA_mean, axes=F, cex = 0.25, 
+plot(x=eda_mean_plot$time_to_bp, y=eda_mean_plot$EDA_mean, axes=F, cex = 1, 
      main = "Mean of EDA across observations", ylab = 'mean(EDA)', 
      xlab = "Time before Button Press (in minutes)")
 axis(side = 1)

### 2021-11-09 preprocessing/eda_preprocessing/eda_scale_allFiles.R

@@ -11,7 +11,7 @@
 library(tidyverse)
 library(lubridate)
 
-setwd('..')
+setwd('Z:/SI_data/R21_Study - EDA - scaled/')
 
 ## ------------------ Scale EDA for all observations ------------------ ##
 
@@ -24,6 +24,7 @@ for (i in 1:n_obs){
   
   if (file.exists(input_name)){
     eda_raw = readRDS(input_name)
+    eda_raw = subset(eda_raw, EDA_HighLowPass < 20)
     eda_scaled = mutate(eda_raw, time=as_datetime(ts/1000)) %>%  # change data type of the timestamp
       mutate(date=as_date(time)) %>%  # add date column
       transmute(date, time, Device=E4_serial, EDA=EDA_HighLowPass) %>%

### 2021-11-12 preprocessing/eda_preprocessing/bp_eda.R
updating to save raw, scalemean, and scale

@@ -64,8 +64,6 @@ bp_eda_summary = bp_eda %>%
   summarise(EDA_entries = n())
 max(bp_eda_summary$EDA_entries) # should be no larger than 7200 = 30*60*4
 
-bp_eda = subset(bp_eda, abs(EDA) < 5)
-
 #### Draw the mean plot and the boxplot ####
 bp_eda = bp_eda %>%
   group_by(ID, BP_num, Device) %>%
@@ -73,6 +71,7 @@ bp_eda = bp_eda %>%
   ungroup() %>%
   arrange(time_to_bp)
 eda_mean_plot =  bp_eda %>%
+  filter(!is.na(EDA)) %>%
   group_by(time_to_bp) %>%
   summarize(EDA_mean = mean(EDA), EDA_sd = sd(EDA), num_bp = n()) %>%
   ungroup()
@@ -88,6 +87,7 @@ plot(x=eda_mean_plot$time_to_bp, y=eda_mean_plot$EDA_mean, axes=F, cex = 1,
 axis(side = 1)
 axis(side = 2)
 smoothed_mean = loess(EDA_mean ~ time_to_bp, data = eda_mean_plot, span = 1/3)
+smoothed_sd = loess(EDA_sd ~ time_to_bp, data = eda_mean_plot, span = 1/3)
 lines(eda_mean_plot$time_to_bp, smoothed_mean$fitted, col= "red", lwd = 2)
 dev.off()
 

### 2021-11-12 preprocessing/eda_preprocessing/eda_scale_allFiles.R

@@ -6,12 +6,12 @@
 ## Recent update: May 27, 2021, by Xinrui Wu
 
 #### For each user:
-####     EDA_scaled = (EDA - mean_by_date) / std_by_date
+####     EDA_scaled = (EDA - mean_by_date)
 
 library(tidyverse)
 library(lubridate)
 
-setwd('Z:/SI_data/R21_Study - EDA - scaled2/')
+setwd('Z:/SI_data/R21_Study - EDA - scaled3/')
 
 ## ------------------ Scale EDA for all observations ------------------ ##
 
@@ -36,10 +36,12 @@ for (i in 1:n_obs){
       ungroup() %>%
       filter(time_diff != 0) %>%  # drop duplicate data
       group_by(date) %>%
-      mutate(EDA_scaled = scale(EDA)) %>%
+      mutate(EDA_raw = EDA) %>%
+      mutate(EDA_scaled = scale(EDA, scale = FALSE)) %>%
+      mutate(EDA_scaled2 = scale(EDA)) %>%
       ungroup() %>%
       arrange(Device, time) %>%
-      transmute(timestamp=time, Device, EDA_scaled)
+      transmute(timestamp=time, Device, EDA_raw, EDA_scaled, EDA_scaled2)
     saveRDS(eda_scaled, file = output_name)
   }
 }

### 2021-11-12 preprocessing/eda_preprocessing/bp_eda.R
WEnt back to scaled cause it was cleanest

@@ -3,9 +3,9 @@ library(tidyverse)
 library(lubridate)
 
 ## WINDOWS
-setwd('Z:/SI_data/R21_Study - EDA - scaled2/')
-## LINUS
-setwd("/mnt/turbo/SI_data/R21_Study - EDA - scaled2/")
+setwd('Z:/SI_data/R21_Study - EDA - scaled/')
+## LINUX
+setwd("/mnt/turbo/SI_data/R21_Study - EDA - scaled/")
 
 
 # read in the button press data and transform the data type of the timestamp
@@ -14,7 +14,7 @@ bp = transmute(bp, ID, time=as_datetime(ts))
 
 
 #### Match button presses and the previous EDA (scaled) data ####
-if(!file.exists("../summary_data/bp_eda2.rds")) {
+if(!file.exists("../summary_data/bp_eda.rds")) {
   bp_eda = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, Device=NULL, 
                       EDA_time=NULL, EDA=NULL, EDA_num=NULL)
   
@@ -42,7 +42,7 @@ if(!file.exists("../summary_data/bp_eda2.rds")) {
         if (dim(sub_eda)[1] > 0){
           sub_bp_eda = sub_eda %>%
             transmute(ID=rep(id,n()), BP_num=rep(j,n()), BP_time=rep(bp_time, n()), 
-                      Device, EDA_time=timestamp, EDA=EDA_scaled) %>%
+                      Device, EDA_time=timestamp, EDA=EDA) %>%
             group_by(Device) %>%
             mutate(EDA_num = 1:n()) %>%
             ungroup() %>%
@@ -52,9 +52,9 @@ if(!file.exists("../summary_data/bp_eda2.rds")) {
       }
     }
   }
-  saveRDS(bp_eda, file = '../summary_data/bp_eda2.rds')
+  saveRDS(bp_eda, file = '../summary_data/bp_eda.rds')
 } else {
-  bp_eda = readRDS(file = '../summary_data/bp_eda2.rds')
+  bp_eda = readRDS(file = '../summary_data/bp_eda.rds')
 }
 
 
@@ -73,21 +73,25 @@ bp_eda = bp_eda %>%
 eda_mean_plot =  bp_eda %>%
   filter(!is.na(EDA)) %>%
   group_by(time_to_bp) %>%
[…]

### 2021-11-12 preprocessing/eda_preprocessing/eda_scale_allFiles.R

@@ -36,12 +36,10 @@ for (i in 1:n_obs){
       ungroup() %>%
       filter(time_diff != 0) %>%  # drop duplicate data
       group_by(date) %>%
-      mutate(EDA_raw = EDA) %>%
-      mutate(EDA_scaled = scale(EDA, scale = FALSE)) %>%
-      mutate(EDA_scaled2 = scale(EDA)) %>%
+      mutate(EDA = scale(EDA)) %>%
       ungroup() %>%
       arrange(Device, time) %>%
-      transmute(timestamp=time, Device, EDA_raw, EDA_scaled, EDA_scaled2)
+      transmute(timestamp=time, Device, EDA)
     saveRDS(eda_scaled, file = output_name)
   }
 }

### 2021-11-12 preprocessing/eda_preprocessing/bp_eda.R
Deleting visualizations which are now part of processing code

@@ -79,15 +79,14 @@ eda_mean_plot =  bp_eda %>%
   ungroup()
 
 smoothed_mean = loess(EDA_mean ~ time_to_bp, data = eda_mean_plot, span = 1/3)
+eda_mean_plot$time_to_bp = eda_mean_plot$time_to_bp/4/60
 
 png("../figures/smoothed_eda.png",
     width = 480, height = 480, units = "px", pointsize = 12)
-
 par(mfrow = c(1,1), mar = c(5,4,1,1)+0.1)
 # mean plot
-eda_mean_plot$time_to_bp = eda_mean_plot$time_to_bp/4/60
 plot(x=eda_mean_plot$time_to_bp, y=eda_mean_plot$EDA_mean, axes=F, cex = 0.5,
-     ylab = 'Avg. Scaled EDA', 
+     ylab = 'Average Scaled EDA', 
      xlab = "Time before Button Press (in minutes)",
 )
 axis(side = 1)

### 2021-11-12 visualization/functions.R
Re do the visualization folder

@@ -1,34 +1,0 @@
-kernel_outcome <- function(eda, current_bp_ts, window_length = 3, bandwidth = 20/60) {
-  # Sets up the dataset and current time-stamp to compare to
-  # Choice of window_length is mostly to speed up computation
-  # Choice of bandwidth
-  ## NOTE: If all NAs near 
-  inside_fn <- function(time_until_event) {
-    # Computes numerator and denominator of the kernel estimates
-    diff = eda$ts - current_bp_ts
-    minute_diff = diff/1000/60 
-    good_obs = (time_until_event - window_length < minute_diff) & (minute_diff <= time_until_event + window_length) & !is.na(eda$EDA_filtered)
-    weights = dnorm(abs(minute_diff[good_obs]-time_until_event)^2, sd = bandwidth)
-    return(list("numerator_weight"=sum((weights*eda$EDA_FeatureScaled_Filtered[good_obs]), na.rm = TRUE), 
-                "denominator_weight"=sum(weights)))
-  }
-  return(inside_fn)
-}
-
-event_kernel_apply <- function(bp_obs, eda, id_bp, current_bp_ts, sequence) {
-  current_bp_ts = id_bp$ts_ms[bp_obs]
-  temp_output <- matrix(unlist(lapply(sequence, kernel_outcome(eda,current_bp_ts))), nrow = 2)
-  return(temp_output)
-}
-
-nonevent_kernel_apply <- function(eda, id_bp, sequence) {
-  ## Throw away if within 30 minutes of button press
-  current_ts = sample(eda$ts, size = 1)
-  if(min(abs(current_ts-id_bp$ts_ms)/1000/60) < 30){bad_choice = TRUE} else{bad_choice=FALSE}
-  while(bad_choice) {
-    current_ts = sample(eda$ts, size = 1)
-    if(min(abs(current_ts-id_bp$ts_ms)/1000/60) < 30){bad_choice = TRUE} else{bad_choice=FALSE}
-  }
-  temp_output <- matrix(unlist(lapply(sequence, kernel_outcome(eda,current_ts))), nrow = 2)
-  return(temp_output)
-}

### 2021-11-12 visualization/plot_generator.R

@@ -1,92 +1,0 @@
-sequence = seq(-30,10, by = 0.25) # sequence along which we compute
-plot_obs = sequence <= 0
-event_total_numerator = event_total_denominator = rep(NA, length(sequence))
-nonevent_total_numerator = nonevent_total_denominator = rep(NA, length(sequence))
-id_counter = NA
-setwd("/Volumes/murphy_lab/users/wdempsey/data-for-fda/data/")
-
-for(id in 1001:1091) {
-  print(paste("On participant ",id))
-  rdsfile = paste("kerneloutput_id_",id,".RDS",sep="")
-  if(!file.exists(rdsfile)) {
-    print("No data so we did nothing!")
-  } else {
-    temp = readRDS(file = rdsfile)
-    
-    id_counter = c(id_counter, id)
-    event_total_numerator = rbind(event_total_numerator,colSums(temp$event_numerator_output, na.rm = TRUE))
-    event_total_denominator = rbind(event_total_denominator,colSums(temp$event_denominator_output, na.rm = TRUE))
-    nonevent_total_numerator = rbind(nonevent_total_numerator,colSums(temp$nonevent_numerator_output, na.rm = TRUE))
-    nonevent_total_denominator = rbind(nonevent_total_denominator,colSums(temp$nonevent_denominator_output, na.rm = TRUE))
-    
-    event_mean = colSums(temp$event_numerator_output, na.rm = TRUE)/colSums(temp$event_denominator_output, na.rm = TRUE)
-    nonevent_mean = colSums(temp$nonevent_numerator_output, na.rm = TRUE)/colSums(temp$nonevent_denominator_output, na.rm = TRUE)
-    
-    print(paste("Number of events is", nrow(temp$event_numerator_output)))
-    
-    filename = paste("~/Documents/github/fda-recurrentevents/data_visualization/figs/participant_",id,"_meancurves.png", sep="")
-    png(filename, width = 480, height = 480, units = "px", pointsize = 12)
-    par(mfrow = c(2,1), mar = c(3,3,1,1)+0.1)
-    plot(sequence, event_mean-mean(event_mean, na.rm = TRUE), 
-         type = "l", xlim = c(min(sequence), max(sequence)), 
-         axes = FALSE, ylab = "", xlab = "")
-    mtext("EDA (mean-centered)", side = 2, line = 2, cex = 0.75)
[…]

### 2021-11-12 visualization/weight_calculator_per_participant.R

@@ -1,63 +1,0 @@
-#First read in the arguments listed at the command line
-args=(commandArgs(TRUE))
-
-##args is now a list of character vectors
-## First check to see if arguments are passed.
-## Then cycle through each element of the list and evaluate the expressions.
-if(length(args)==0){
-  print("No arguments supplied.")
-  ## Supply default participant id
-  id = 1040
-}else{
-  id = as.numeric(args[[1]])
-}
-
-print(paste("Current participant is", id))
-source('./functions.R')
-
-setwd("/Users/walterdempsey/Harvard University/R21_Study - EDA")
-buttonpress = readRDS("../R21_Study - tags/button_presses.RDS")
-
-file_name = paste(id,"_EDA.rds", sep = "")
-if(file.exists(file_name)) {
-  eda = readRDS(paste(id,"_EDA.rds", sep = ""))
-} else{
-  print(paste("No EDA file with id", id))
-  break
-}
-
-id_bp = subset(buttonpress, ID == id)
-library(doParallel)
-library(doRNG)
-
-cl = Sys.getenv("SLURM_NTASKS_PER_NODE")
-cl
-
-registerDoParallel(cores = (cl))
-
-# Shows the number of Parallel Workers to be used
-getDoParWorkers()
-
-sequence = seq(-30,10, by = 0.25) # sequence along which we compute
-
-event_times_output = foreach(bp_obs=1:nrow(id_bp), .combine = 'cbind') %dopar% 
-  event_kernel_apply(bp_obs, eda, id_bp, current_bp_ts, sequence)
-
-event_numerator_output = matrix(event_times_output[1,], ncol = length(sequence), byrow = TRUE)
-event_denominator_output = matrix(event_times_output[2,], ncol = length(sequence), byrow = TRUE)
-
-num.iters = 5000 # Number of non-event samples per participant
-nonevent_times_output = foreach(bp_obs=1:num.iters, .combine = 'cbind', .options.RNG=197311) %dopar% 
-  nonevent_kernel_apply(eda, id_bp, sequence)
-
-nonevent_numerator_output = matrix(nonevent_times_output[1,], ncol = length(sequence), byrow = TRUE)
-nonevent_denominator_output = matrix(nonevent_times_output[2,], ncol = length(sequence), byrow = TRUE)
-
-full_data = list("event_numerator_output"=event_numerator_output, 
-                 "event_denominator_output"=event_denominator_output,
[…]

### 2021-11-12 preprocessing/acc_preprocessing/summary_and_plots_AI.R
Making preprocessing for acc

@@ -1,14 +1,9 @@
 # Summary and plots on AI and pause time
 
-## ---- Please place this file in a folder named "methods" 
-##      under the directory where the scaled AI data is stored. ----
-
-## Recent update: May 27, 2021, by Xinrui Wu
-
 library(tidyverse)
 library(lubridate)
 
-setwd('..')
+setwd('/mnt/turbo/SI_data/R21_Study - ACC - AI/')
 
 ## ------------------ Summary for AI and Pause time------------------ ##
 n_obs = 91  # the largest number of observations (files)
@@ -27,6 +22,7 @@ all_pause = data.frame(ID=integer(), pause_num=integer(), pause=double(), period
 none_ind = 0
 for (i in 1:n_obs){
   id = 1000 + i 
+  print(paste("At Patient", id))
   input_name = paste0("./", id, "_AI.rds")
   
   ## check the progress of the loop##
@@ -48,7 +44,8 @@ for (i in 1:n_obs){
     summary_AI[i, 'Continuous Periods'] = length(unique(acc_ai$Period))
     
     # pause information
-    time_diff = round(diff(as.numeric(acc_ai$timestamp)))
+    timestamps = acc_ai$timestamp[order(acc_ai$timestamp)]
+    time_diff = round(diff(as.numeric(timestamps)))
     cum_time = cumsum(time_diff)[(c(which(time_diff > epch), length(time_diff)+1)-1)] - 
       c(0, cumsum(time_diff)[which(time_diff > epch)])
     n_pause = length(which(time_diff > epch))
@@ -64,9 +61,7 @@ for (i in 1:n_obs){
       summary_AI[i, 'Pause_max(hour)'] = 0
       summary_AI[i, 'Pause_min(hour)'] = 0
     }
-  }
-  
-  else{
+  } else{
     none_ind = c(none_ind, i)
   }
 }
@@ -75,8 +70,8 @@ for (i in 1:n_obs){
 none_ind = none_ind[-1]
 if(length(none_ind) > 0) {summary_AI = summary_AI[-none_ind,]}
 
-saveRDS(summary_AI, file = './methods/summary_data/summary_AI.rds')
-saveRDS(all_pause, file = './methods/summary_data/pauses_AI.rds')
+saveRDS(summary_AI, file = '../summary_data/summary_AI.rds')
+saveRDS(all_pause, file = '../summary_data/pauses_AI.rds')
 
 # summary of pause time for each observation
 summary_pause = all_pause %>%
[…]

### 2021-11-12 write-up/fda-recurrent.tex
putting in plots
Updating paper
ADding plots  for appendix (to show heterogeneity)

@@
 \usepackage{floatrow}
+\usepackage{subcaption}
 % Table float box with bottom caption, box width adjusted to content
@@
 
-The focus here is on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
+Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
 Electrodermal activity can be significantly impacted by external factors (e.g., room temperature).
-To account for the high between user-day variation, we analyze EDA standardized per study-day.
+To account for the high between user-day variation, we analyze EDA standardized per study-day and  device.
 The individual EDA and gACC trajectories are highly variable which tends to obscure patterns and trends.
-In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps and randomly sampled non-event timestamps.
-While there are certain similarities in the two plots, the differences in temporal trends are rather striking.
-In particular, we see sharp increases in EDA in the XX minutes prior to button presses, which is in marked contrast with levels in the six months prior to failure, as seen in the lower pair of curves.
+In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 10 minutes prior to button presses.
 
 \begin{figure}[!th]
-\rule{3cm}{3cm}%
-\caption{EDA and ACC plots for average at non-event and event times.}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/smoothed_eda.png}
+  \caption{Electrodermal activity (EDA)}
+  \label{fig:mean_eda}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/smoothed_ai.png}
+  \caption{Activity Index (AI)
+}  \label{fig:mean_acc}
+\end{subfigure}
+\caption{Average scaled EDA and AI in the 30 minutes prior to button presses}
 \label{fig:mean_edacc}

### 2021-11-12 preprocessing/acc_preprocessing/summary_and_plots_AI.R

@@ -130,25 +130,16 @@ saveRDS(num_devices, file = '../summary_data/num_devices_AI.rds')
 
 table(num_devices$num_device)
 
-## ------------------ Plots ------------------ ##
-## summary of AI ##
-#### Histogram of AI mean ####
-hist(summary_AI$AI_mean, breaks = 25, axes = F, xlab = '', ylab = '', 
-     main = 'Histogram of mean(AI) across Observations')
-axis(side = 1)
-axis(side = 2)
-
-
 #### AI quantiles for each obs. ####
-par(xpd=T, mar=c(5,5,5,7))
-ymax = max(summary_AI$AI_max)
-ymin = min(summary_AI$AI_min)
-plot(x = (summary_AI$ID - 1000), y = summary_AI$AI_mean, type = 'l', axes = F, ylim = c(ymin-50, ymax+50), 
-     xlab = 'Observations', ylab = 'AI', main = 'Summary of AI acorss Observations')
+png("~/Documents/github/fda-recurrentevents/figures/ai_summary.png",
+    width = 720, height = 480, units = "px", pointsize = 18)
+par(mar=c(4,4,1,7))
+ymax = max(summary_AI$AI.99)
+ymin = min(summary_AI$AI.25)
+plot(x = (summary_AI$ID - 1000), y = summary_AI$AI_mean, type = 'l', axes = F, ylim = c(ymin*0.8, ymax*1.2), 
+     xlab = '', ylab = '')
 axis(side = 1)
 axis(side = 2)
-points(x = (summary_AI$ID - 1000), y = summary_AI$AI_max, type = 'l', lty = 1, col = 'grey')
-points(x = (summary_AI$ID - 1000), y = summary_AI$AI_min, type = 'l', lty = 1, col = 'grey')
 points(x = (summary_AI$ID - 1000), y = summary_AI$AI.25, type = 'l', lty = 3, col = 'blue')
 points(x = (summary_AI$ID - 1000), y = summary_AI$AI.50, type = 'l', lty = 3, col = 'red')
 points(x = (summary_AI$ID - 1000), y = summary_AI$AI.75, type = 'l', lty = 2, col = 'blue')
@@ -158,36 +149,24 @@ text(x = (summary_AI$ID - 1000), y = rep(ymax+20, dim(summary_AI)[1]),
      labels = summary_AI$Days, col = "dimgrey", cex = .5)
 text(x = (summary_AI$ID - 1000), y = rep(ymax+45, dim(summary_AI)[1]), 
[…]

### 2021-11-12 preprocessing/eda_preprocessing/summary_and_plots_EDA.R

@@ -1,17 +1,19 @@
 ## Summary and plots on EDA and its pause time
-
-## ---- Please place this file in a folder named "methods" 
-##      under the directory where the scaled EDA data is stored. ----
-
-## Recent update: May 27, 2021, by Xinrui Wu
-
 library(tidyverse)
 library(lubridate)
 
-setwd('..')
+setwd('/mnt/turbo/SI_data/R21_Study - EDA - scaled/')
+
 ## ------------------ Summary for Pause time in EDA------------------ ##
 n_obs = 91  # the largest number of observations (files)
 td = 0.25 # time difference between 2 consecutive data
+
+# summary of AI for each person
+summary_EDA = data.frame(matrix(0, nrow = n_obs, ncol = 14))
+colnames(summary_EDA) = c('ID', 'EDA_mean', 'EDA_sd', 'EDA_max', 'EDA_min', 
+                         'EDA.25', 'EDA.50', 'EDA.75', 'EDA.95', 'EDA.99', 
+                         'Days', 'Continuous Periods', 
+                         'Pause_max(hour)', 'Pause_min(hour)')
 
 #  pause and period (after the pause) length in seconds
 all_pause = data.frame(ID=integer(), pause_num=integer(), pause_len=double(), 
@@ -19,44 +21,50 @@ all_pause = data.frame(ID=integer(), pause_num=integer(), pause_len=double(),
 
 for (i in 1:n_obs){
   id = 1000 + i 
-  input_name = paste0("./", id, "_EDA.rds")
+  print(paste("At patient ", id))
+  input_name = paste0("./", id, "_EDA_scaled.rds")
   
   ## check the progress of the loop and save temporary result ##
   if (i%%10 == 0){
     print(i)
-    saveRDS(all_pause, file = './methods/summary_data/pauses_EDA.rds')
+    saveRDS(all_pause, file = '../summary_data/pauses_EDA.rds')
   }
   
   if (file.exists(input_name)){
     eda = readRDS(input_name)
-    eda_hlp = eda %>%
-      transmute(ID=Participant, time = as_datetime(eda$ts/1000), EDA = EDA_HighLowPass)
-    
+    summary_EDA[i, 'ID'] = id
+    summary_EDA[i, 'EDA_mean'] = mean(eda$EDA_scaled)
[…]

### 2021-11-12 preprocessing/eda_preprocessing/summary_and_plots_EDA.R
Issue with eda cleaning for appendix. must deal with at home

@@ -18,7 +18,7 @@ colnames(summary_EDA) = c('ID', 'EDA_mean', 'EDA_sd', 'EDA_max', 'EDA_min',
 #  pause and period (after the pause) length in seconds
 all_pause = data.frame(ID=integer(), pause_num=integer(), pause_len=double(), 
                        period_len=double(), num_na=integer()) 
-
+none_ind = 0
 for (i in 1:n_obs){
   id = 1000 + i 
   print(paste("At patient ", id))
@@ -39,7 +39,7 @@ for (i in 1:n_obs){
     summary_EDA[i, 'EDA_min'] = min(eda$EDA_scaled)
     summary_EDA[i, 'EDA_max'] = max(eda$EDA_scaled)
     summary_EDA[i, c('EDA.25', 'EDA.50', 'EDA.75', 'EDA.95', 'EDA.99')] = 
-      quantile(eda$EDA_scaled, probs = c(0.25, 0.5, 0.75, 0.95, 0.99))
+      quantile(eda$EDA_scaled, probs = c(0.25, 0.5, 0.75, 0.95, 0.99), na.rm = T)
     summary_EDA[i, 'Days'] = length(unique(as.Date(eda$timestamp)))
     # pause information
     timestamps = eda$timestamp[order(eda$timestamp)]
@@ -74,7 +74,7 @@ summary_pause_EDA = all_pause %>%
             num_30min_to_1day = sum(pause_num > 0 & pause_len >= 30*60 & pause_len < 60*60*24),
             num_more_than_1day = sum(pause_num > 0 & pause_len >= 60*60*24), 
             num_total = sum(pause_num > 0))
-saveRDS(summary_pause_EDA, file = './methods/summary_data/summary_pause_EDA.rds')
+saveRDS(summary_pause_EDA, file = '../summary_data/summary_pause_EDA.rds')
 
 
 ## ---------- Frequency of missing EDA data in the 30 minuetes before button presses ---------- ##
@@ -153,10 +153,6 @@ for (i in 1:n_obs){
 
 
 ## ------------------ Plots ------------------ ##
-
-## pause time ##
-hist(all_pause$pause_len)
-boxplot((pause_len/3600)~ID, data = all_pause, ylab='Pause Time (EDA) / hour')
 
 
 df = t(as.matrix(summary_pause_EDA[,2:5]))

### 2021-11-13 preprocessing/eda_preprocessing/summary_and_plots_EDA.R
Update summary_and_plots_EDA.R

@@ -2,11 +2,14 @@
 library(tidyverse)
 library(lubridate)
 
+## Windows
+setwd('Z:/SI_data/R21_Study - EDA - scaled/')
+## Linux
 setwd('/mnt/turbo/SI_data/R21_Study - EDA - scaled/')
 
 ## ------------------ Summary for Pause time in EDA------------------ ##
 n_obs = 91  # the largest number of observations (files)
-td = 0.25 # time difference between 2 consecutive data
+td = 0.50 # time difference between 2 consecutive data
 
 # summary of AI for each person
 summary_EDA = data.frame(matrix(0, nrow = n_obs, ncol = 14))
@@ -44,8 +47,10 @@ for (i in 1:n_obs){
     # pause information
     timestamps = eda$timestamp[order(eda$timestamp)]
     time_diff = round(diff(as.numeric(timestamps)),3)
-    cum_time = cumsum(time_diff)[(c(which(time_diff > td), length(time_diff)+1)-1)] - 
-      c(0, cumsum(time_diff)[which(time_diff > td)])
+    cumulative_timediff = cumsum(time_diff)
+    gap_obs = which(time_diff > td)
+    cum_time = cumulative_timediff[(c(gap_obs, length(time_diff)+1)-1)] - 
+      c(0,cumulative_timediff[gap_obs])
     n_pause = length(which(time_diff > td))
     all_pause = rbind(all_pause, data.frame(ID=id, pause_num=0, pause=0, period=cum_time[1]))
     if (n_pause > 0){
@@ -69,15 +74,15 @@ saveRDS(all_pause, file = '../summary_data/pauses_EDA.rds')
 # summary of pause time for each observation
 summary_pause_EDA = all_pause %>%
   group_by(ID) %>%
-  summarise(num_less_than_1min = sum(pause_num > 0 & pause_len < 60), 
-            num_1min_to_30min = sum(pause_num > 0 & pause_len >= 60 & pause_len < 30*60), 
-            num_30min_to_1day = sum(pause_num > 0 & pause_len >= 30*60 & pause_len < 60*60*24),
-            num_more_than_1day = sum(pause_num > 0 & pause_len >= 60*60*24), 
+  summarise(num_less_than_1min = sum(pause_num > 0 & pause < 60), 
+            num_1min_to_30min = sum(pause_num > 0 & pause >= 60 & pause < 30*60), 
+            num_30min_to_1day = sum(pause_num > 0 & pause >= 30*60 & pause < 60*60*24),
[…]

### 2021-11-13 preprocessing/eda_preprocessing/summary_and_plots_EDA.R
Update summary_and_plots_EDA.R

@@ -83,112 +83,70 @@ saveRDS(summary_pause_EDA, file = '../summary_data/summary_pause_EDA.rds')
 
 
 ## ---------- Frequency of missing EDA data in the 30 minutes before button presses ---------- ##
+time_with_miss_30min = transmute(all_pause, ID, pause_num, period) %>%
+  mutate(time_with_miss = period * (period <= 30*60 & pause_num > 0) + 
+           rep(30*60, dim(all_pause)[1]) * (period > 30*60 & pause_num > 0))
 
-time_with_miss_30min = all_pause %>%
-  mutate(time_with_miss = period_len * (period_len <= 30*60 & pause_num > 0) + 
-           rep(30*60, dim(all_pause)[1]) * (period_len > 30*60 & pause_num > 0), 
-         ave_pause=0)
+time_with_miss_30min$frac_missing = (time_with_miss_30min$time_with_miss / time_with_miss_30min$period)
+time_with_miss_30min = subset(time_with_miss_30min, !is.na(frac_missing))
 
-for (i in 1:dim(time_with_miss_30min)[1]){
-  time_with_miss = time_with_miss_30min$time_with_miss[i]
-  pause = time_with_miss_30min$pause_len[i]
-  if(time_with_miss > 0){
-    if(pause>=1800){
-      time_with_miss_30min$ave_pause[i] = (1800+(1800-time_with_miss))/2
-    }
-    else{
-      if((1800-pause)>=time_with_miss){
-        time_with_miss_30min$ave_pause[i] = pause
-      }
-      else{
-        time_with_miss_30min$ave_pause[i] = ((1800-pause)*pause + 
-                                               (pause+(1800-time_with_miss))*(time_with_miss+pause-1800)/2)/time_with_miss
-      }
-    }
-  }
-}
-
-ave_pause = time_with_miss_30min %>%
-  group_by(ID) %>%
-  summarise(miss_ave = sum(ave_pause*time_with_miss)/max(sum(time_with_miss),1), 
-            all_ave = sum(ave_pause*time_with_miss)/sum(period_len)) %>%
-  mutate(miss_ave = round(miss_ave/60,2), all_ave = round(all_ave/60,2))
-
+# note: "frequency of missing" = 
[…]

### 2021-11-13 preprocessing/eda_preprocessing/summary_and_plots_EDA.R
Building eda appendix figures

@@ -1,6 +1,7 @@
 ## Summary and plots on EDA and its pause time
 library(tidyverse)
 library(lubridate)
+library(ggplot2)
 
 ## Windows
 setwd('Z:/SI_data/R21_Study - EDA - scaled/')
@@ -102,17 +103,9 @@ prob_miss_30min_obs = time_with_miss_30min %>%
 prob_miss_30min_all = sum(time_with_miss_30min$time_with_miss) / sum(time_with_miss_30min$period)
 print(prob_miss_30min_all)
 
-library(ggplot2)
-png("~/Documents/github/fda-recurrentevents/figures/missing_data_eda.png",
-    width = 480, height = 480, units = "px", pointsize = 12)
-par(mar = c(4, 4, 1, 1), mfrow = c(1,1))
-qplot(prob_miss_30min_obs$prob, 
-      geom="histogram", xlab = "Fraction of Missing AI data in 30-minutes prior to a measurement") 
-dev.off()
-
 #### EDA quantiles for each obs. ####
 summary_EDA = summary_EDA[summary_EDA$ID > 0,]
-png("~/Documents/github/fda-recurrentevents/figures/eda_summary.png",
+png("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/eda_summary.png",
     width = 720, height = 480, units = "px", pointsize = 18)
 par(xpd=T, mar=c(4,4,1,7))
 ymax = max(summary_EDA$EDA.99)
@@ -143,10 +136,10 @@ mtext("User ID", side = 1, line = 2)
 dev.off()
 
 ## probability of having missing data in the past 30 mins for each observed time ##
-png("~/Documents/github/fda-recurrentevents/figures/missingperuser.png",
-    width = 720, height = 480, units = "px", pointsize = 18)
-par(mfrow = c(1,1),
-    mar = c(5,4,1,1) + 0.1)
-ggplot(data = prob_miss_30min_obs, aes(x = ID, y = prob)) + 
-  geom_bar(stat = "identity") + 
-  xlab("Patient ID")+ ylab("Probability of missing data in past 30 minutes")
+png("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/missing_data_eda.png",
+    width = 480, height = 480, units = "px", pointsize = 12)
+par(mar = c(4, 4, 1, 1), mfrow = c(1,1))
+qplot(prob_miss_30min_obs$prob, 
[…]

### 2021-11-13 preprocessing/acc_preprocessing/summary_and_plots_AI.R
Update summary_and_plots_AI.R

@@ -1,8 +1,11 @@
 # Summary and plots on AI and pause time
-
 library(tidyverse)
 library(lubridate)
+library(ggplot2)
 
+## Windows
+setwd('Z:/SI_data/R21_Study - ACC - AI/')
+## Linux
 setwd('/mnt/turbo/SI_data/R21_Study - ACC - AI/')
 
 ## ------------------ Summary for AI and Pause time------------------ ##
@@ -103,14 +106,6 @@ prob_miss_30min_obs = time_with_miss_30min %>%
 prob_miss_30min_all = sum(time_with_miss_30min$time_with_miss) / sum(time_with_miss_30min$period)
 print(prob_miss_30min_all)
 
-library(ggplot2)
-png("~/Documents/github/fda-recurrentevents/figures/missing_data_ai.png",
-    width = 480, height = 480, units = "px", pointsize = 12)
-par(mar = c(4, 4, 1, 1), mfrow = c(1,1))
-qplot((time_with_miss_30min$time_with_miss / time_with_miss_30min$period), 
-      geom="histogram", xlab = "Fraction of Missing AI data in 30-minutes prior to a measurement") 
-dev.off()
-
 ## ---------- Summary of devices ---------- ##
 num_devices = data.frame(ID=integer(), num_device=integer()) # number of devices for each observation
 
@@ -162,11 +157,10 @@ mtext("User ID", side = 1, line = 2)
 dev.off()
 
 ## probability of having missing data in the past 30 mins for each observed time ##
-png("~/Documents/github/fda-recurrentevents/figures/missingperuser.png",
-    width = 720, height = 480, units = "px", pointsize = 18)
-par(mfrow = c(1,1),
-    mar = c(5,4,1,1) + 0.1)
-ggplot(data = prob_miss_30min_obs, aes(x = ID, y = prob)) + 
-  geom_bar(stat = "identity") + 
-  xlab("Patient ID")+ ylab("Probability of missing data in past 30 minutes")
-
+png("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/missing_data_ai.png",
+    width = 480, height = 480, units = "px", pointsize = 12)
+par(mar = c(4, 4, 1, 1), mfrow = c(1,1))
+qplot(prob_miss_30min_obs$prob, 
+      geom="histogram", xlab = "Patient Average Fraction of Missing EDA data in 30-minutes",
+      bins = 20) 
+dev.off()

### 2021-11-13 preprocessing/acc_preprocessing/summary_and_plots_AI.R
Update summary_and_plots_AI.R

@@ -126,12 +126,14 @@ saveRDS(num_devices, file = '../summary_data/num_devices_AI.rds')
 table(num_devices$num_device)
 
 #### AI quantiles for each obs. ####
-png("~/Documents/github/fda-recurrentevents/figures/ai_summary.png",
+summary_AI = readRDS(file = '../summary_data/summary_AI.rds')
+
+png("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/ai_summary.png",
     width = 720, height = 480, units = "px", pointsize = 18)
-par(mar=c(4,4,1,7))
+par(xpd=T, mar=c(4,4,1,7))
 ymax = max(summary_AI$AI.99)
 ymin = min(summary_AI$AI.25)
-plot(x = (summary_AI$ID - 1000), y = summary_AI$AI_mean, type = 'l', axes = F, ylim = c(ymin*0.8, ymax*1.2), 
+plot(x = (summary_AI$ID - 1000), y = summary_AI$AI_mean, type = 'l', axes = F, ylim = c(ymin*0.8, ymax*1.1), 
      xlab = '', ylab = '')
 axis(side = 1)
 axis(side = 2)

### 2021-12-20 methods/mean_cov_estimation.R
Updating for complete case analysis

@@ -12,7 +12,7 @@
 ## Note: one can check if reconstruction is good by taking
 ## hatX(t,s) = coefficients%*%eigenvectors + mu(t,s) 
 ## And seeing how close hatX(t,s) is to X(t,s).
-
+  
 setwd("Z:/SI_data/")
 set_of_types = c("eda", "acc")
 for (type in set_of_types){
@@ -67,13 +67,16 @@ for (type in set_of_types){
   lines(sequence, Y[obs,], col = "black", lwd = 1, lty = 2)
   lines(sequence, optphi_vectors%*%optcoef[obs,]/inflation + est$Yhat[obs,], col = "red", lty = 2)
   legend(-10,max(Y[obs,])*0.8, c(paste("Observed", legend_name), "K=10", "K=35"), col = c("black", "red", "red"), lty = c(1,1,2), cex = 0.75, bty = "n")
-  # dev.off()
-
+  dev.off()
+  
+  ## Timestamps
+  timestamp = event_complete[full_obs,2]
   ## SAVE THE MEAN, COEFFICIENT MATRIX, AND FIRST K EIGEN VECTORS
   saveRDS(object = est$Yhat, file = paste(type, "_lin_event_means_",today(),".RDS", sep=""))
   saveRDS(object = coef/inflation, file = paste(type, "_lin_event_coef_matrix_",today(),".RDS", sep=""))
   saveRDS(object = phi_vectors, file = paste(type, "_lin_event_eigen_vectors_",today(),".RDS", sep=""))
-  saveRDS(object = x, file = paste(type, "_lin_event_complete_case_times_",today(),".RDS", sep=""))
+  saveRDS(object = x, file = paste(type, "_lin_event_complete_case_timesincebaseline_",today(),".RDS", sep=""))
+  saveRDS(object = timestamp, file = paste(type, "_lin_event_complete_case_timestamp_",today(),".RDS", sep=""))
   saveRDS(object = as.numeric(event_complete[full_obs,1]), 
           file = paste(type, "_lin_event_complete_case_ids_",today(),".RDS", sep=""))
   
@@ -135,6 +138,8 @@ for (type in set_of_types){
   x = nonevent_complete[full_obs,2]
   z = sequence
   id = nonevent_complete[full_obs,1]
+  ## Timestamps
+  timestamp = nonevent_complete[full_obs,2]
   rm("nonevent_complete") # Remove the full data to free up memory
   
[…]

### 2021-12-21 write-up/fda-recurrent.tex
Data cleaning

@@
 The point estimator is given by $(B)^{-1} \sum_{b=1}^B \hat \theta_{b}$ where $\hat \theta_b = M^{-1} \sum_{m=1}^M \hat \theta_{b,m}$.
-To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b )^2$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B M \cdot (\hat \theta_b - \hat \theta)^2$ respectively.
-Then the estimator of the variance of $\hat \theta$ is given by $\left( \frac{B+1}{BM} \right) MSB - MSW/M$.
+To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b ) (\hat \theta_{b,m} - \hat \theta_b )^\top$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B M \cdot (\hat \theta_b - \hat \theta)(\hat \theta_b - \hat \theta)^\top$ respectively.
+Then the estimator of the variance-covariance matrix of $\hat \theta$ is given by $\left( \frac{B+1}{BM} \right) MSB - MSW/M$.
 

### 2021-12-21 methods/likelihood_estimation.R

@@ -5,11 +5,11 @@ setwd('/mnt/turbo/SI_data/')
 
 ## Read in acc and eda model.matrix
 ## ACC
-acc_event_model.matrix = readRDS("acc_event_modelmatrix_2021-12-16.RDS")
-acc_nonevent_model.matrix = readRDS("acc_nonevent_modelmatrix_2021-12-16.RDS")
+acc_event_model.matrix = readRDS("acc_event_modelmatrix_2021-12-20.RDS")
+acc_nonevent_model.matrix = readRDS("acc_nonevent_modelmatrix_2021-12-20.RDS")
 ## EDA
-eda_event_model.matrix = readRDS("eda_event_modelmatrix_2021-12-16.RDS")
-eda_nonevent_model.matrix = readRDS("eda_nonevent_modelmatrix_2021-12-16.RDS")
+eda_event_model.matrix = readRDS("eda_event_modelmatrix_2021-12-20.RDS")
+eda_nonevent_model.matrix = readRDS("eda_nonevent_modelmatrix_2021-12-20.RDS")
 
 ## PULL IN PI_IDS
 log_sampling_rate = log(0.5)
@@ -19,25 +19,27 @@ print("RDS Files Reading correctly")
 if(!require("glmnet")){install.packages('glmnet')}
 library('glmnet')
 
-Y = c(rep(1,nrow(event_model.matrix)), rep(0, nrow(nonevent_model.matrix)))
+Y = c(rep(1,nrow(acc_event_model.matrix)), rep(0, nrow(acc_nonevent_model.matrix)))
 
-model.matrix = rbind(event_model.matrix[,-1], nonevent_model.matrix[,-1])
-hour.info = c(event_model.matrix[,1], nonevent_model.matrix[,1])
+model.matrix = rbind(acc_event_model.matrix[,-c(1:3)], acc_nonevent_model.matrix[,-c(1:3)])
+hour.info = c(acc_event_model.matrix[,3], acc_nonevent_model.matrix[,3])
 daytime_obs = (hour.info > 9) & (hour.info < 20)
 
 n.tmp = length(Y)
 p.tmp = ncol(model.matrix)
-subsample_offset = rep(log_sampling_rate,nrow(dataset))
+subsample_offset = rep(log_sampling_rate,nrow(model.matrix))
 p.fac = rep(1, ncol(model.matrix))
 p.fac[1:4] = 0 #no penalty on the first 4 variables
-lambda_max <- 1/n.tmp
[…]

### 2021-12-21 methods/mean_cov_estimation.R

@@ -135,7 +135,7 @@ for (type in set_of_types){
   Y = as.matrix(Y, nrow = nrow(Y), ncol = ncol(Y))
   mu = rowMeans(Y)
   quad_Y = sweep(Y,1,mu)^2
-  x = nonevent_complete[full_obs,2]
+  x = as.numeric(nonevent_complete[full_obs,3])
   z = sequence
   id = nonevent_complete[full_obs,1]
   ## Timestamps

### 2021-12-21 methods/model_matrix_construction.R

@@ -42,46 +42,50 @@ for (type in set_of_types) {
   bb = cbind(1, sequence, sequence^2, sapply(knots, function(k) ((sequence - k > 0) * (sequence - k)) ^ 2))
   print("Generated Splines")
   
-  if(!file.exists(paste(type, "_event_modelmatrix_2021-12-17.RDS", sep = ""))) {
-    event_eigen_vectors = readRDS(paste(type, "_lin_event_eigen_vectors_2021-11-22.RDS", sep = ""))
-    event_coef = readRDS(paste(type, "_lin_event_coef_matrix_2021-11-22.RDS", sep = ""))
-    event_means = readRDS(paste(type, "_lin_event_means_2021-11-22.RDS", sep = ""))
-    event_times = readRDS(paste(type, "_lin_event_complete_case_times_2021-11-22.RDS", sep = ""))
-    event_id = readRDS(paste(type, "_lin_event_complete_case_ids_2021-11-22.RDS", sep = ""))
+  if(!file.exists(paste(type, "_event_modelmatrix_2021-12-20.RDS", sep = ""))) {
+    event_eigen_vectors = readRDS(paste(type, "_lin_event_eigen_vectors_2021-12-20.RDS", sep = ""))
+    event_coef = readRDS(paste(type, "_lin_event_coef_matrix_2021-12-20.RDS", sep = ""))
+    event_means = readRDS(paste(type, "_lin_event_means_2021-12-20.RDS", sep = ""))
+    event_timesincebaseline = readRDS(paste(type, "_lin_event_complete_case_timesincebaseline_2021-12-20.RDS", sep = ""))
+    event_timestamp = readRDS(paste(type, "_lin_event_complete_case_timestamp_2021-12-20.RDS", sep = ""))
+    event_id = readRDS(paste(type, "_lin_event_complete_case_ids_2021-12-20.RDS", sep = ""))
     ## BUILD outer product of b-spline and event eigen
     event_J_coef = event_coef%*%t(event_eigen_vectors)%*%bb
     event_J_means = event_means%*%bb
-    event_model.matrix = cbind(event_id, event_times, hour(as_datetime(event_times)), event_J_coef + event_J_means)
+    event_model.matrix = cbind(event_id, event_timesincebaseline, hour(event_timestamp), event_J_coef + event_J_means)
     event_model.matrix = data.frame(event_model.matrix)
[…]

### 2021-12-21 methods/likelihood_estimation.R
all data is now clean

@@ -60,7 +60,6 @@ betaHat.net <- bb %*% ridge.coef
 plot(betaHat.net)
 
 temp_daytime = glm(Y[daytime_obs]~model.matrix[daytime_obs,], family = "binomial", offset = rep(log_sampling_rate,length(Y[daytime_obs])))
-summary(temp_daytime)
 
 W = temp_daytime$fit*(1-temp_daytime$fit)
 X = model.matrix(temp_daytime)

### 2021-12-21 methods/model_matrix_construction.R

@@ -42,7 +42,8 @@ for (type in set_of_types) {
   bb = cbind(1, sequence, sequence^2, sapply(knots, function(k) ((sequence - k > 0) * (sequence - k)) ^ 2))
   print("Generated Splines")
   
-  if(!file.exists(paste(type, "_event_modelmatrix_2021-12-20.RDS", sep = ""))) {
+  if(!file.exists(paste(type, "_event_modelmatrix_2021-12-21.RDS", sep = ""))) {
+    print(paste("Building", type, "event model matrix"))
     event_eigen_vectors = readRDS(paste(type, "_lin_event_eigen_vectors_2021-12-20.RDS", sep = ""))
     event_coef = readRDS(paste(type, "_lin_event_coef_matrix_2021-12-20.RDS", sep = ""))
     event_means = readRDS(paste(type, "_lin_event_means_2021-12-20.RDS", sep = ""))
@@ -56,15 +57,17 @@ for (type in set_of_types) {
     event_model.matrix = data.frame(event_model.matrix)
     names(event_model.matrix) = c("id", "timesincebaseline", "hour", paste("acc_X", 1:ncol(event_J_coef), sep = ""))
     
-    rm("event_means"); rm("event_coef"); rm("event_eigen_vectors"); rm("event_times")
-    rm("event_J_coef"); rm("event_J_means")
+    rm("event_means"); rm("event_coef"); rm("event_eigen_vectors"); rm("event_timestamp")
+    rm("event_timesincebaseline"); rm("event_J_coef"); rm("event_J_means"); rm("event_id")
     
     saveRDS(event_model.matrix, file = paste(type, "_event_modelmatrix_",today(), ".RDS", sep = ""))
   } else {
-    event_model.matrix = readRDS(paste(type,"_event_modelmatrix_2021-12-20.RDS",sep =""))
+    print(paste("Already built", type, "event"))
+    event_model.matrix = readRDS(paste(type,"_event_modelmatrix_2021-12-21.RDS",sep =""))
   }
   ## PULL IN NONEVENT RDS FILES
-  if(!file.exists(paste(type, "_nonevent_modelmatrix_2021-12-20.RDS", sep = ""))) {
+  if(!file.exists(paste(type, "_nonevent_modelmatrix_2021-12-21.RDS", sep = ""))) {
+    print(paste("Bulding", type, "non-event model matrix"))
[…]

### 2021-12-22 methods/likelihood_estimation.R
Update likelihood_estimation.R

@@ -5,11 +5,25 @@ setwd('/mnt/turbo/SI_data/')
 
 ## Read in acc and eda model.matrix
 ## ACC
-acc_event_model.matrix = readRDS("acc_event_modelmatrix_2021-12-20.RDS")
-acc_nonevent_model.matrix = readRDS("acc_nonevent_modelmatrix_2021-12-20.RDS")
+acc_event_model.matrix = readRDS("acc_event_modelmatrix_2021-12-21.RDS")
+acc_nonevent_model.matrix = readRDS("acc_nonevent_modelmatrix_2021-12-21.RDS")
 ## EDA
-eda_event_model.matrix = readRDS("eda_event_modelmatrix_2021-12-20.RDS")
-eda_nonevent_model.matrix = readRDS("eda_nonevent_modelmatrix_2021-12-20.RDS")
+eda_event_model.matrix = readRDS("eda_event_modelmatrix_2021-12-21.RDS")
+eda_nonevent_model.matrix = readRDS("eda_nonevent_modelmatrix_2021-12-21.RDS")
+
+## GENERATE SPLINES
+K_b = 35
+num=K_b-3
+qtiles <- seq(0, 1, length = num + 2)[-c(1, num + 2)]
+## EDA
+eda_sequence <- seq(-30,0, by = 1/60)
+eda_knots <- quantile(eda_sequence, qtiles)
+eda_bb = cbind(1, eda_sequence, eda_sequence^2, sapply(eda_knots, function(k) ((eda_sequence - k > 0) * (eda_sequence - k)) ^ 2))
+print("Generated EDA Splines")
+## ACC 
+acc_sequence <- seq(-30,0, by = 1/6)
+acc_knots <- quantile(acc_sequence, qtiles)
+acc_bb = cbind(1, acc_sequence, acc_sequence^2, sapply(acc_knots, function(k) ((acc_sequence - k > 0) * (acc_sequence - k)) ^ 2))
 
 ## PULL IN PI_IDS
 log_sampling_rate = log(0.5)
@@ -19,16 +33,16 @@ print("RDS Files Reading correctly")
 if(!require("glmnet")){install.packages('glmnet')}
 library('glmnet')
 
-Y = c(rep(1,nrow(acc_event_model.matrix)), rep(0, nrow(acc_nonevent_model.matrix)))
-
-model.matrix = rbind(acc_event_model.matrix[,-c(1:3)], acc_nonevent_model.matrix[,-c(1:3)])
+## Model per type
[…]

### 2021-12-22 methods/likelihood_estimation.R
Update likelihood_estimation.R

@@ -245,29 +245,35 @@ eda_stderr = sqrt(diag(eda_bb%*%eda_updated_Sigma%*%t(eda_bb)))
 ## 
 library(ggplot2)
 
-df_summary <- data.frame(sequence = acc_sequence+30, estimate = acc_betaHat.net,
+df_acc_summary <- data.frame(sequence = acc_sequence+30, estimate = acc_betaHat.net,
                  lowerCI = acc_betaHat.net - 1.96 * acc_stderr,
                  upperCI = acc_betaHat.net + 1.96 * acc_stderr)
 
-df_summary$sequence[which(df_summary$lowerCI > 0)]
+acc_xmax = max(df_acc_summary$sequence[which(df_acc_summary$lowerCI > 0)])
+acc_xmin = min(df_acc_summary$sequence[which(df_acc_summary$lowerCI > 0)])
 
-ggplot(df_summary, aes(x=sequence, y=estimate)) +
+png("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/acc_coef.png",
+    width = 480, height = 480, units = "px", pointsize = 16)
+ggplot(df_acc_summary, aes(x=sequence, y=estimate)) +
   geom_line(size=1, alpha=0.8) +
   geom_ribbon(aes(ymin=lowerCI, ymax=upperCI) ,fill="blue", alpha=0.2) +
-  annotate("rect",xmin=0,xmax=3,ymin=-Inf,ymax=Inf, alpha=0.1, fill="black") +
+  annotate("rect",xmin=acc_xmin,xmax=acc_xmax,ymin=-Inf,ymax=Inf, alpha=0.1, fill="black") +
   xlab("Time until Button Press") + ylab(expression(paste(beta, "(s)")))
+dev.off()
   # geom_line(data=df_tidy, aes(x=Time, y=Ratio, group=Cell), color="grey") +
 
-# png("~/Downloads/linearfit.png", width = 720, 
-# height = 480, units = "px", pointsize = 12)
-par(mar = c(4,4,1,1) + 0.1)
-plot(eda_sequence+30, eda_betaHat.net, type= "l", 
-     axes = FALSE, xlab = "Time until event (in min)", 
-     ylab = expression(paste(beta, "(s)")),
-     ylim = c(-0.0015,0.0015), lwd = 2)
-axis(side = 1); axis(side = 2)
[…]

### 2021-12-22 preprocessing/buttonpress_preprocessing.R
Create buttonpress_preprocessing.R

@@ -1,0 +1,55 @@
+rm(list=ls())
+setwd("Z:/SI_data/")
+library("lubridate")
+col_names = c("id", "start_ts", "end_ts")
+buttonpresses = readRDS("./R21_Study - tags/button_presses.RDS")
+ID_list = unique(buttonpresses$ID)
+buttonpresses$timestamp = as_datetime(buttonpresses$ts)
+buttonpresses$timesincebaseline = NA
+if(!file.exists("./R21_Study - tags/button_presses_extrainfo.RDS")) {
+  for(id in ID_list) {
+    print(paste("On participant",id))
+    if (file.exists(paste("./R21_Study - EDA - scaled3/",id,"_EDA_scaled.rds", sep = ""))) {
+      edaX <- readRDS(paste("./R21_Study - EDA - scaled3/",id,"_EDA_scaled.rds", sep = ""))
+      eda_start_ts = min(edaX$timestamp)
+      eda_end_ts = max(edaX$timestamp)
+      rm(edaX)
+    } else {
+      eda_start_ts = NA
+      eda_end_ts = NA
+    }
+    if (file.exists(paste("./R21_Study - ACC - AI/",id,"_AI.rds", sep = ""))) {
+      accX <- readRDS(paste("./R21_Study - ACC - AI/",id,"_AI.rds", sep = ""))
+      acc_start_ts = min(accX$timestamp)
+      acc_end_ts = max(accX$timestamp)
+      rm(accX)
+    } else {
+      acc_start_ts = NA
+      acc_end_ts = NA
+    }
+    start_ts = min(acc_start_ts, eda_start_ts, na.rm= T)
+    id_bps = subset(buttonpresses, ID == id)
+    buttonpresses$timesincebaseline[buttonpresses$ID == id] = id_bps$timestamp - start_ts
+  }
+  saveRDS(buttonpresses, "./R21_Study - tags/button_presses_extrainfo.RDS")
+} else {
+  buttonpresses = readRDS("./R21_Study - tags/button_presses_extrainfo.RDS")
+}
+
+## PLOT BPS Over TIME
+library(ggplot2)
+point_df = data.frame(id = buttonpresses$ID-1000,
+                      timesincebaseline = buttonpresses$timesincebaseline/60)
+
+point_df$id = as.factor(point_df$id)
+levels(point_df$id) = 1:length(unique(point_df$id))
+unique_ids = unique(point_df$id)
[…]

### 2021-12-22 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b ) (\hat \theta_{b,m} - \hat \theta_b )^\top$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B M \cdot (\hat \theta_b - \hat \theta)(\hat \theta_b - \hat \theta)^\top$ respectively.
-Then the estimator of the variance-covariance matrix of $\hat \theta$ is given by $\left( \frac{B+1}{BM} \right) MSB - MSW/M$.
-
+Then the estimator of the variance-covariance matrix of $\hat \theta$ is given by $\hat \Sigma_{B,M} = \left( \frac{B+1}{BM} \right) MSB - MSW/M$.
+We obtain the varaiance for~$\beta(t)$ by $\phi (t) \hat \Sigma_{B,M} \phi(t)^\top$.
+We follow~\cite{Bartlett2020} and construct confidence intervals based on Satterthwaite’s degrees of freedom, which here is given by
+$$
+\hat \nu = \phi (t) \hat \Sigma_{B,M} \phi(t)^\top
+$$
 The bootstrap followed by multiple imputation procedure has been studied extensively by~\cite{Bartlett2020} and is robust to uncongeniality.

### 2021-12-23 preprocessing/buttonpress_preprocessing.R
Update buttonpress_preprocessing.R

@@ -6,6 +6,7 @@ buttonpresses = readRDS("./R21_Study - tags/button_presses.RDS")
 ID_list = unique(buttonpresses$ID)
 buttonpresses$timestamp = as_datetime(buttonpresses$ts)
 buttonpresses$timesincebaseline = NA
+buttonpresses$endtime = NA
 if(!file.exists("./R21_Study - tags/button_presses_extrainfo.RDS")) {
   for(id in ID_list) {
     print(paste("On participant",id))
@@ -27,9 +28,11 @@ if(!file.exists("./R21_Study - tags/button_presses_extrainfo.RDS")) {
       acc_start_ts = NA
       acc_end_ts = NA
     }
-    start_ts = min(acc_start_ts, eda_start_ts, na.rm= T)
+    start_ts = as_datetime(min(acc_start_ts, eda_start_ts, na.rm= T))
+    end_ts = as_datetime(max(acc_end_ts, eda_end_ts, na.rm= T))
     id_bps = subset(buttonpresses, ID == id)
-    buttonpresses$timesincebaseline[buttonpresses$ID == id] = id_bps$timestamp - start_ts
+    buttonpresses$timesincebaseline[buttonpresses$ID == id] = difftime(id_bps$timestamp, start_ts, units = "hours")
+    buttonpresses$endtime[buttonpresses$ID == id] = difftime(end_ts, start_ts, units = "hours")
   }
   saveRDS(buttonpresses, "./R21_Study - tags/button_presses_extrainfo.RDS")
 } else {
@@ -39,17 +42,50 @@ if(!file.exists("./R21_Study - tags/button_presses_extrainfo.RDS")) {
 ## PLOT BPS Over TIME
 library(ggplot2)
 point_df = data.frame(id = buttonpresses$ID-1000,
-                      timesincebaseline = buttonpresses$timesincebaseline/60)
+                      timesincebaseline = buttonpresses$timesincebaseline,
+                      endtime = buttonpresses$endtime)
 
 point_df$id = as.factor(point_df$id)
 levels(point_df$id) = 1:length(unique(point_df$id))
 unique_ids = unique(point_df$id)
 png("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/recurrentevents_plot.png",
-    width = 480, height = 480, units = "px", pointsize = 16)
-ggplot(subset(point_df, timesincebaseline < 250), aes(x=timesincebaseline, y = id)) +
[…]

### 2021-12-23 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 Any events that were identified as incorrect button press activity were removed.
-At the end of the study period, the average number of button presses per day was $1.91$ with a standard deviation of $2.23$.
-Figure~\ref{fig:buttons} presents the distribution of button presses as a function of hour in the day.
-This demonstrates low button-press counts prior to 7AM and a sharp drop off by 11PM.
+At the end of the study period, the average number of button presses per day was $2.42$ with a standard deviation of $2.62$.
+Investigation of the button press data shows low button-press counts prior to 7AM and a sharp drop off by 11PM.
 This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
-Due to (B), we define each study day to begin at 9AM and end at 8PM.
-Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
+To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.
+% Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
 
 \vspace{0.25cm}
-\begin{minipage}{\textwidth}
-  \begin{minipage}[b]{0.49\textwidth}
+\begin{figure}[!th]
     \centering
-%    \includegraphics[width=0.9\textwidth]{figs/buttonpresses.png}
-    \captionof{figure}{Button-presses per hour within the user-day}
+    \includegraphics[width=0.8\textwidth]{../figures/recurrentevents_plot.png}
+    \captionof{figure}{User button-presses (red) versus time since study entry (in hours).  The black mark indicates the final sensor measurement time.}
     \label{fig:buttons}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[b]{0.49\textwidth}
-    \centering
-    \centering
-%    \includegraphics[width=1.0\textwidth]{figs/recurrent_events_plot.png}
-    \captionof{figure}{User button-presses versus time in study}
-    \label{fig:buttons2}
-    \end{minipage}
-\end{minipage}
+\end{figure}
 \vspace{0.25cm}
 
-Figure~\ref{fig:buttons2} presents the raw button presses as a function of time in study for each user, showing the potential heterogeneity in button press rates between users and study days.
+Figure~\ref{fig:buttons} visualizes button presses versus time since study entry for each user.
+Here, day 30 is assumed to censor the observation process.
+A black mark signals dropout before day 30.
+Figure~\ref{fig:buttons} shows the potential heterogeneity in button press rates between users and study days.
 To assess whether there is between user or between study day variation, Table~\ref{tab:anova} presents a two-way ANOVA decomposition of the button press counts as a function of participant and day in study.
@@
       & DF & Sum Sq & Mean Sq & F value & Pr($>$F) \\ \hline
-        Participant & 88 & 13 & 13.31 & 2.71 & 0.10 \\
-        Day in Study & 75 & 54.71 & 54.71 & $8.8 \times 10^{-4}$ \\
-        Residuals & 989 & 4.91 & - \\ \hline
+        Participant & 88 & 2675.4 & 30.4 & 8.9 & $< 2 \times 10^{-16}$ \\
+        Day in Study & 29 & 443.3 & 15.3 & 4.5 & $3.9 \times 10^{-13}$ \\
+        Residuals & 672 & 2304.2 & 3.4 \\ \hline
       \end{tabular}  \captionof{table}{ANOVA decomposition of daily button press counts}
@@
 
-\subsection{Effect of electrodermal activity and activity index on moments of distress}
-
-Over a period of 5~years and one month following recruitment,
-patient~$u$ had eight appointments with prothrombin values as follows:
+% \subsection{Effect of electrodermal activity and activity index on moments of distress}
+
+% Over a period of 5~years and one month following recruitment,
+% patient~$u$ had eight appointments with prothrombin values as follows:
 
 \section{Discussion}
-We abide by the comment of Robert Gentleman on big data: (1) make the data as small as possible as fast as possible.
-Statistics works!
+We abide by the comment of Robert Gentleman on big data: ``make data as small as possible as fast as possible.'' Statistics works!
 With subsampling and proper re-weighting the analysis is converted to well-known territory.

### 2021-12-23 methods/likelihood_estimation.R
Update likelihood_estimation.R

@@ -277,3 +277,4 @@ ggplot(df_eda_summary, aes(x=sequence, y=estimate)) +
   # annotate("rect",xmin=acc_xmin,xmax=acc_xmax,ymin=-Inf,ymax=Inf, alpha=0.1, fill="black") +
   xlab("Time until Button Press") + ylab(expression(paste(beta, "(s)")))
 dev.off()
+

### 2021-12-23 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 \label{prop:optimal}
-\textcolor{red}{
+% \textcolor{red}{
 If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
 This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
-WD: Consider larger class and show the proof doesn't extend to non PP models?}
+% WD: Consider larger class and show the proof doesn't extend to non PP models?}
 \end{lemma}
@@
 \noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
-% This result follows from the probability-based sampling design literature~\cite{Rathbun2012}.
-
-% \subsection{Computation versus statistical efficiency tradeoff}
-
-% Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
-While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
-It is this trade-off that we next make precise.
-To start, we consider the subsampling rate~$\pi(t) = c \times h(t; \theta)$ for $c>0$, i.e., the subsampling rate proportional to the intensity function with time-independent constant~$c > 0$.
-Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c / (c+1)$.
-Under Lemma~\ref{lemma:simpleasym},
-
-% For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
-This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
-Not only this, Lemma~\ref{prop:optimal} shows weights~\eqref{eq:waage_weights} are optimal within a particular class of weighted estimating equations.
-
-% \begin{prop} \normalfont
-% \label{prop:optimal}
-% If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
-This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
-% \end{prop}
-
-% \noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
 
@@
 Any events that were identified as incorrect button press activity were removed.
-At the end of the study period, the average number of button presses per day was $2.42$ with a standard deviation of $2.62$.
+At the end of the 30-day study period, the average number of button presses per day was $2.42$ with a standard deviation of $2.62$.
 Investigation of the button press data shows low button-press counts prior to 7AM and a sharp drop off by 11PM.
@@
 Figure~\ref{fig:buttons} visualizes button presses versus time since study entry for each user.
-Here, day 30 is assumed to censor the observation process.
+Day 30 is assumed to censor the observation process.
 A black mark signals dropout before day 30.
@@
 To assess whether there is between user or between study day variation, Table~\ref{tab:anova} presents a two-way ANOVA decomposition of the button press counts as a function of participant and day in study.
-The ANOVA decomposition demonstrates high variation with day in study compared to across users in terms of button presses.
+The ANOVA decomposition demonstrates high variation with day in study and across users.
 
@@
 
-Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
+Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{activity index}~\cite{10.1371/journal.pone.0160644}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
 Electrodermal activity can be significantly impacted by external factors (e.g., room temperature).
 To account for the high between user-day variation, we analyze EDA standardized per study-day and  device.
-The individual EDA and gACC trajectories are highly variable which tends to obscure patterns and trends.
-In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 10 minutes prior to button presses.
+The individual EDA and AI trajectories are highly variable which tends to obscure patterns and trends.
+In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and AI are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 10 minutes prior to button presses.
 
@@
 
-Inspection of the graphs suggests setting $\Delta = XX$ for EDA and $XX$ for gACC and  Gaussian with moments {\bf comparison over $\Delta$ and using imputation} where the baseline rate is assumed to depend on
+\subsection{Complete-case analysis}
+
+Inspection of~Figure\ref{fig:mean_eda} and Figure~\ref{fig:mean_acc} suggests setting $\Delta = 30$ minutes is adequate to capturing the proximal impact of EDA and AI
 
@@
 
+TBD.
+
 % \subsection{Effect of electrodermal activity and activity index on moments of distress}
@@
 \section{Discussion}
-We abide by the comment of Robert Gentleman on big data: ``make data as small as possible as fast as possible.'' Statistics works!
-With subsampling and proper re-weighting the analysis is converted to well-known territory.
-(2) Make the software and tools as fast and as maintanable as possible.
-This is exactly teh benefit of subsampling.
-we convert to penalized logistic regression on a smaller but still relatively efficient dataset.
-This means we can use all the existing tools!
-
-Extensions to online and sampling over and over as we improve estimators.
+In this paper, we have presented a methodology for translating a difficult functional analysis with recurrent events problem into a traditional logistic regression.
+The translation leveraged subsampling and weighting techniques, specifically the use of weights suggested by~\cite{Waagepetersen2008}, along with flexible functional data analysis methods of~\cite{Goldsmith2011} with marginal covariance methods for longitudinal functional data of~\cite{Park2018}.
+The proposed methodology abides by the comment of Robert Gentleman on big data: ``make data as small as possible as fast as possible.''  Subsampling and  weighting converts the problem to well-known territory which allowed us to leverage existing software.
+We show limited loss of efficiency when the subsampling is properly tuned to the event rates.
+Important extensions to an online sampling algorithm, optimal weighting when the Poisson point process assumption does not hold, and non-linear functional data methods are considered important future work.
 

### 2021-12-23 methods/likelihood_estimation.R
Update likelihood_estimation.R

@@ -24,7 +24,7 @@ print("Generated EDA Splines")
 acc_sequence <- seq(-30,0, by = 1/6)
 acc_knots <- quantile(acc_sequence, qtiles)
 acc_bb = cbind(1, acc_sequence, acc_sequence^2, sapply(acc_knots, function(k) ((acc_sequence - k > 0) * (acc_sequence - k)) ^ 2))
-
+print("Generated ACC Splines")
 ## PULL IN PI_IDS
 log_sampling_rate = log(0.5)
 print("RDS Files Reading correctly")
@@ -35,10 +35,11 @@ library('glmnet')
 
 ## Model per type
 acc_Y = c(rep(1,nrow(acc_event_model.matrix)), rep(0, nrow(acc_nonevent_model.matrix)))
-acc_model.matrix = rbind(acc_event_model.matrix[,-c(1:3)], acc_nonevent_model.matrix[,-c(1:3)])
-hour.info = c(acc_event_model.matrix[,3], acc_nonevent_model.matrix[,3])
+acc_model.matrix = rbind(acc_event_model.matrix[,-c(1,3)], acc_nonevent_model.matrix[,-c(1,3)])
+## HOUR WAS IN UTC BUT NEED TO BE TRANSLATED TO ETS
+## THIS IS DONE BY -5 HOURS FUNCTION
+hour.info = (c(acc_event_model.matrix[,3], acc_nonevent_model.matrix[,3]) - 5)%%24
 daytime_obs = (hour.info > 9) & (hour.info < 20)
-
 n.tmp = length(acc_Y)
 p.tmp = ncol(acc_model.matrix)
 subsample_offset = rep(log_sampling_rate,nrow(acc_model.matrix))

### 2022-07-01 write-up/fda-recurrent.tex
Updated rec event figure, writeup, and bibtex files

@@
 In classical joint models~\citep{Henderson2000, Tsiatis2004}, time-varying covariates~$x_i (t)$ are observed only intermittently at appointment times.
-In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which require integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$ that is a function of the unknown functional parameter $\beta(\cdot)$.
+In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which requires integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$ that is a function of the unknown functional parameter $\beta(\cdot)$.
 That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
@@
 We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
-We use the sandwich smoother~\citep{Xiao2013} to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
+The sandwich smoother~\citep{Xiao2013} is used to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
 Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009}, bivariate tensor product splines~\citep{Wood2006}, and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
@@
 
-This connection established by Lemma~\ref{lemma:logistic} between our proposed methodology and logistic regression allows us to leverage ``off-the-shelf'' software.
+The connection established by Lemma~\ref{lemma:logistic} between our proposed methodology and logistic regression allows us to leverage ``off-the-shelf'' software.
 The main complication is pre-processing of the functional data; however, these additional steps can also be taken care of via existing software.
@@
 and~$t \in \mathbb{R}_+$. 
-To account for overlap between $\Delta$-windows for $t \in \bfT_i \cup \bfD_i$ - i.e., avoid imputing values at a particular time to different values -- imputation is performed sequentially over this set.
+%To account for overlap between $\Delta$-windows for $t \in \bfT_i \cup \bfD_i$ - i.e., avoid imputing values at a particular time to different values -- imputation is performed sequentially over this set.
 %$\{ (t,t') : t, t^\prime \in \bfT_i \cup \bfD_i \& |t - t'| < \Delta \}$

### 2022-07-05 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 
-Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{activity index}~\cite{10.1371/journal.pone.0160644}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
+Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{activity index}~\citep{10.1371/journal.pone.0160644}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
 Electrodermal activity can be significantly impacted by external factors (e.g., room temperature).
@@
 
-Inspection of~Figure\ref{fig:mean_eda} and Figure~\ref{fig:mean_acc} suggests setting $\Delta = 30$ minutes is adequate to capturing the proximal impact of EDA and AI
+Inspection of Figures~\ref{fig:mean_eda} and~\ref{fig:mean_acc} suggest setting $\Delta = 30$ minutes is adequate to capturing the proximal impact of EDA and AI on the risk of a button press.
+To ensure minimal loss of efficiency, the subsampling rate was set to once every fifteen minutes.
+Given the daily button press rate, this ensures an average of $44$ non-events to $2.5$ events per day.
+Based on Table~\ref{tab:compvseff}, this ensures we can achieve a substantial data reduction at a minimal loss of efficiency.
+After sampling non-event times, complete-case analyses are performed, i.e., sampled times where sensors included in the model have any level of missing data are ignored.
+
+We analyzed activity index (AI) and electrodermal activity separately as given in equation~~\eqref{eq:hazardlinear} as well as jointly in a multivariate model as in equation~\eqref{eq:multihazardlinear}.
+Results did not change substantially.
+Figure~\ref{fig:eda_estimate} and~\ref{fig:acc_estimate} presents the estimates from the separate analyses with their associated 95\% confidence intervals.
+We highlight in gray the statistically significant regions.
+Here, we see that standardized EDA is not associated with increased risk of button press in our population-level model, while activity index sees a positive association in the final few minutes prior to a button press.
+In section~\ref{sec:missingdata_heterogeneity}, we investigate whether the population-level results are sensitive to missing data and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
 
 \begin{figure}[!th]
-\rule{3cm}{3cm}%
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/eda_coef.png}
+  \caption{Electrodermal activity (EDA)}
+  \label{fig:eda_estimate}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/acc_coef.png}
+  \caption{Activity Index (AI)
+}  \label{fig:acc_estimate}
+\end{subfigure}
 \caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is complete-case analysis, while dotted line is for }
-\label{}
+\label{fig:mean_edacc}
 \end{figure}

### 2022-07-19 write-up/fda-recurrent.tex
added figs to recreate at work comp

@@
 $$
-\hat \nu = \phi (t) \hat \Sigma_{B,M} \phi(t)^\top
+\hat \nu = \frac{\left[ \left( \frac{B+1}{BM} \right) MSB - \frac{MSW}{M} \right]^2}{ (B-1)^{-1} \left( \frac{B+1}{BM} \right)^2 MSB^2 + \frac{MSW^2}{BM^2(M-1)}}
 $$
@@
   \centering
-  \includegraphics[width=.8\linewidth]{../figures/eda_coef.png}
+  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint.png}
   \caption{Electrodermal activity (EDA)}
@@
   \centering
-  \includegraphics[width=.8\linewidth]{../figures/acc_coef.png}
+  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint.png}
   \caption{Activity Index (AI)
@@
 \subsection{Accounting for missing data and heterogeneity}
+\label{sec:missingdata_heterogeneity}
 

### 2022-07-19 methods/bootstrap_mi.R
Bootstrap ready for loops on GL

@@ -13,12 +13,27 @@
 ## hatX(t,s) = coefficients%*%eigenvectors + mu(t,s) 
 ## And seeing how close hatX(t,s) is to X(t,s).
 
-current_bootstrap = 1 # CONVERT TO ACCEPT THIS 
+args=(commandArgs(TRUE))
 
-num_bootstraps = 200
+print(args)
+
+##args is now a list of character vectors
+## First check to see if arguments are passed.
+## Then cycle through each element of the list and evaluate the expressions.
+if(length(args)==0){
+  print("No arguments supplied.")
+  ##supply default values
+  current_bootstrap = 1 # CONVERT TO ACCEPT THIS 
+}else{
+  current_bootstrap = as.numeric(args[[1]])
+}
+
+print(paste("Current bootstrap:", current_bootstrap))
+
+# num_bootstraps = 200
 num_imputes = 2
 library(refund); library(lubridate); library(dynr); library(Matrix)
-set.seed(1391307)
+# set.seed(1391307)
   
 ## WINDOWS
 setwd("Z:/SI_data/")
@@ -53,7 +68,9 @@ for (type in set_of_types){
   impute_Sigma = event_Sigma/inflation^2
   
   for(current_mi in 1:num_imputes) {
+    print(paste("Number of rows to impute:", length(which(!full_obs))))
     for(row in which(!full_obs)) {
+      print(paste("Current rows:", which(row == which(!full_obs))))
       current_row = bootstrap_event_complete[row,]
       current_time = current_row[3]
       data = current_row[,4:length(current_row)]
@@ -63,6 +80,7 @@ for (type in set_of_types){
       ## EXTRACT CONDITIONAL VARIANCE
       missing_spots = is.na(data)
       observed_data = data[!missing_spots]
+      print(paste("Number of missing entries:", sum(!missing_spots)))
       Sigma_11 = as.matrix(impute_Sigma[missing_spots, missing_spots], nrow = sum(missing_spots))
       Sigma_12 = as.matrix(impute_Sigma[missing_spots, !missing_spots], nrow = sum(missing_spots))
       if(all(is.na(data))) {
@@ -181,7 +199,10 @@ for (type in set_of_types){
 }
 
 ## CLEAR WORKSPACE
-rm(list=ls()) 
+rm(list=c("eig_Sigma", "eig_Sigma_est", "eig_vectors", "eigen_Sigma", "est", 
[…]

### 2022-08-03 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 \def\pr{\mathop{\text{pr}}\nolimits}
+\def\var{\mathop{\text{var}}\nolimits}
+\def\cov{\mathop{\text{cov}}\nolimits}
 \def\k{{\bf k}}
@@
 \title[Recurrent event analysis with functional covariates via random subsampling]{Recurrent event analysis in the presence of real-time high frequency data via random subsampling}
-\author{Walter Dempsey}
-\address {Department of Biostatistics, University of Michigan, 1415 Washington Heights, Ann Arbor, MI 48109, USA}
- \email{wdem@umich.edu}
+% \author{Walter Dempsey}
+% \address {Department of Biostatistics, University of Michigan, 1415 Washington Heights, Ann Arbor, MI 48109, USA}
+% \email{wdem@umich.edu}
 
@@
 In particular, a higher sampling rate can decrease estimator variance at the cost of increased computation.
-We show via simulations that the benefits of incredibly high sampling rates is often negligible, as the contribution to the variation is small relative to the variation in the underlying stochastic processes.
+We show via simulations that the benefits of very high sampling rates is often negligible, as the contribution to the variation is small relative to the variation in the underlying stochastic processes.
 Second, derived estimating equations are optimal, implying loss of statistical efficiency is only due to subsampling procedure and not the derived methodology.
@@
 where~$\theta$ is the parameter vector.
-For high frequency physiological data, we assume that current risk is log-additive and depends on a linear functional of the health process over some recent window of time and some pre-specified features of the counting process; that is,
+For high frequency physiological data, we assume that current risk is log-additive and depends on a linear functional of the health process over some recent window of time and pre-specified features of the counting process; that is,
 \begin{equation}
@@
 To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
-Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed (see Section~\ref{section:missingdata}).
+Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed; see Section~\ref{section:missingdata} for details on handling of missing data.
 
@@
 
-Probabilistic subsampling converts the single sensor stream~$\bfx_i$ into a sequence of functions observed repeatedly at sampled times~$\bfD_i$ and event times~$\bfT_i$.
+Probabilistic subsampling converts the single sensor stream~$\bfx_i$ into a sequence of functions observed repeatedly at sampled times~$\bfD_i$ and event times~$\bfT_i$ over windows of length~$\Delta$.
 Such a data structure is commonly referred to as \emph{longitudinal functional data}~\citep{Xiao2013, GoldSmith2015}.
@@
 \begin{enumerate}[label=(E.\arabic*)]
-\item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$ such that $R_i (t) = 1$
+\item\label{E1} The subsampling rate is both lower and upper bounded at all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$ such that $R_i (t) = 1$
 \item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
@@
 \end{lemma}
-An estimator for~$\Xi(\theta)$ is
+A design-unbiased estimator for~$\Xi(\theta)$ is
 \begin{equation}
@@
 Table~\ref{tab:mise} demonstrates that the variance does increase as the sampling rate decreases.
-However, the rate of increase in the MISE is quite low relative.
+However, the rate of increase in the MISE is relatively quite low.
 In the second case study, for example, the MISE increases by 6\% while the run time is 3 times faster.
@@
 Subsampling variance & - & $9.0 \times 10^{-7}$ & $2.9 \times 10^{-6}$ & $5.8 \times 10^{-6}$  \\
-Variance & $7.9 \times 10^{-3}$ & $9.2 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.30 \times 10^{-2}$ \\
+Variance & $7.9 \times 10^{-3}$ & $9.2 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ \\
 Squared Bias & $9.8 \times 10^{-2}$ & $9.5 \times 10^{-2}$ & $9.3 \times 10^{-2}$ & $9.2 \times 10^{-2}$   \\
@@
 \subsubsection{Multiple imputation under uncongeniality}
+\label{section:miuncong}
 
@@
 The point estimator is given by $(B)^{-1} \sum_{b=1}^B \hat \theta_{b}$ where $\hat \theta_b = M^{-1} \sum_{m=1}^M \hat \theta_{b,m}$.
-To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b ) (\hat \theta_{b,m} - \hat \theta_b )^\top$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B M \cdot (\hat \theta_b - \hat \theta)(\hat \theta_b - \hat \theta)^\top$ respectively.
+To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b ) (\hat \theta_{b,m} - \hat \theta_b )^\top$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B (\hat \theta_b - \hat \theta)(\hat \theta_b - \hat \theta)^\top$ respectively.
 Then the estimator of the variance-covariance matrix of $\hat \theta$ is given by $\hat \Sigma_{B,M} = \left( \frac{B+1}{BM} \right) MSB - MSW/M$.
@@
 The random subsampling framework thus simplifies handling of missing data via connections to function-on-scalar multiple imputation by~\cite{Petrovich2018} as well as to bootstrap to handle uncongeniality by~\cite{Bartlett2020}.
-Ignoring the computational time of bootstrap sampling, the computational time for the first choise in the simulation study with $B=200$ bootstraps and $M=2$ imputations per bootstrap leads to $35$ hours for a sampling rate of $0.5$ compared to $7$ hours for a sampling rate of $4$, which highlights the benefits of the proposed framework.
+Ignoring the computational time of bootstrap sampling, the computational time for the first choice in the simulation study with $B=200$ bootstraps and $M=2$ imputations per bootstrap leads to $35$ hours for a sampling rate of $0.5$ compared to $7$ hours for a sampling rate of $4$, which highlights the benefits of the proposed framework.
 
@@
 Lemma~\ref{lemma:logistic} implies that the random subsampling framework applied to equation~\eqref{eq:multilevel} leads to a penalized logistic mixed-effects model.
-As far as the authors are aware, the combination of mixed-effects and $L_2$-penalization on a subset of parameters has not been addressed in existing packages.
-To address this, we derive a penalized adaptive Gaussian quadarture (PADQ) for estimation.
-For conciseness, the algorithm is presented in Appendix~\ref{app:penlogit}.
+As far as the authors are aware, the combination of mixed-effects and $L_2$-penalization on a subset of parameters is not available in existing software.
+% An penalized adaptive Gaussian quadarture (PADQ) algorithm can be derived.  estimation is presented in Appendix~\ref{app:penlogit}.
+Given the paper's focus on ``off-the-shelf'' software implementations, multilevel models are considered important future work.
 
@@
 Results did not change substantially.
-Figure~\ref{fig:eda_estimate} and~\ref{fig:acc_estimate} presents the estimates from the separate analyses with their associated 95\% confidence intervals.
+Figure~\ref{fig:eda_estimate} and~\ref{fig:acc_estimate} presents the estimates from the joint analysis with their associated 95\% confidence intervals.
 We highlight in gray the statistically significant regions.
-Here, we see that standardized EDA is not associated with increased risk of button press in our population-level model, while activity index sees a positive association in the final few minutes prior to a button press.
-In section~\ref{sec:missingdata_heterogeneity}, we investigate whether the population-level results are sensitive to missing data and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
+Here, we see that standardized EDA is not associated with increased risk of button press, while activity index sees a positive association in the final few minutes prior to a button press.
+In section~\ref{sec:missingdata_heterogeneity}, we investigate whether the results are sensitive to missing data.
+ % and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
 
@@
 \end{subfigure}
-\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is complete-case analysis, while dotted line is for }
+\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the complete-case analysis, while highlighted region are the model-based pointwise 95\% confidence intervals.}
 \label{fig:mean_edacc}
@@
 
-\subsection{Accounting for missing data and heterogeneity}
+\subsection{Accounting for missing data}
 \label{sec:missingdata_heterogeneity}
 
-TBD.
+In Appendix~\ref{app:csmissing}, Figures~\ref{fig:missing_eda} and~\ref{fig:missing_acc} present the per-patient average fraction with missing EDA and AI in 30-minute windows respectively.
+For EDA, we have a wide range of missingness -- from $0\%$ to over $40$\% across individuals with most between $5\%$ and $30\%$ average fraction missing.
+For AI, the missingness is less pronounced -- from $0\%$ to $15\%$ across individuals with most between $0\%$ and $10\%$ average fraction missing.
+To assess the impact of missing data, we perform the imputation under uncongeniality using the boostrap and multiple imputation strategy described in Section~\ref{section:miuncong}.
+Here we see that the EDA curve remains non-significant while the region where the AI curve is significant widens slightly to be the prior 4 minutes.
+
+\begin{figure}[!th]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint_bootstrap.png}
+  \caption{Electrodermal activity (EDA)}
+  \label{fig:eda_estimate_bs}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint_bootstrap.png}
+  \caption{Activity Index (AI)
+}  \label{fig:acc_estimate_bs}
+\end{subfigure}
+\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the missing data analysis, while highlighted region are the boostrap and MI based pointwise 95\% confidence intervals.}
+\label{fig:mean_edacc}
+\end{figure}
 
@@
 The translation leveraged subsampling and weighting techniques, specifically the use of weights suggested by~\cite{Waagepetersen2008}, along with flexible functional data analysis methods of~\cite{Goldsmith2011} with marginal covariance methods for longitudinal functional data of~\cite{Park2018}.
-The proposed methodology abides by the comment of Robert Gentleman on big data: ``make data as small as possible as fast as possible.''  Subsampling and  weighting converts the problem to well-known territory which allowed us to leverage existing software.
+The proposed methodology abides by the idea that we should \emph{make data as small as possible as fast as possible}.
+Subsampling and  weighting converts the problem to well-known territory which allows us to leverage existing software.
 We show limited loss of efficiency when the subsampling is properly tuned to the event rates.
-Important extensions to an online sampling algorithm, optimal weighting when the Poisson point process assumption does not hold, and non-linear functional data methods are considered important future work.
+Important extensions to an online sampling algorithm, optimal weighting when the Poisson point process assumption does not hold, and non-linear functional data methods are all considered important future work.
 
@@
 &= \frac{M}{4} n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
-  \times \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
+  \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
   \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
@@
 
-\section{Penalized logistic mixed-effects model}
-\label{app:penlogit}
-
-Here we derive a penalized adaptive Gaussian quadrature (PADQ) to fit the   penalized logistic mixed-effects model.
+% \section{Penalized logistic mixed-effects model}
+% \label{app:penlogit}
+
+% Here we derive a penalized adaptive Gaussian quadrature (PADQ) to fit the   penalized logistic mixed-effects model.
 Here, for clarity, we write $Y_{ij} \in \{ 0, 1 \}$ to be the sequence of binary indicators for participant $i = 1,\ldots,n$ and $j$ indexes the event and subsampled times.
@@
 Then the score equations for the penalized logistic regression with offset $\log(\pi_{ij})$ is written as
-$$
-\sum_{i=1}^n \sum_{j=1}^{k_i} \left( y_{ij} - \frac{1}{1+\exp (- W_{ij} \theta - \log (\pi_{ij}) )} \right) W_{ij} - \frac{1}{\sigma^2} \sum_{k=3}^{K_b} \beta_k
-$$
-Let $B$ be the matrix that extracts the entries in $\theta$ corresponding to $\{ \beta_k\}_{k=3}^{K_b}$ (i.e., $B \theta = ({\bf 0}, \beta_3, \ldots, \beta_{K_b},{\bf 0})$).
+% $$
+% \sum_{i=1}^n \sum_{j=1}^{k_i} \left( y_{ij} - \frac{1}{1+\exp (- W_{ij} \theta - \log (\pi_{ij}) )} \right) W_{ij} - \frac{1}{\sigma^2} \sum_{k=3}^{K_b} \beta_k
+% $$
+% Let $B$ be the matrix that extracts the entries in $\theta$ corresponding to $\{ \beta_k\}_{k=3}^{K_b}$ (i.e., $B \theta = ({\bf 0}, \beta_3, \ldots, \beta_{K_b},{\bf 0})$).
 Let $\W_i$ be a $k_i \times p$ matrix with the $j$th row equal to $W_{ij}$.
@@
 
-We derive the iteratively re-weighted least squares algorithm next. First, take the derivative with respect to $\theta$ to construct the Hessian
-$$
-- \left[ \sum_{i=1}^n \sum_{j=1}^{k_i} \frac{W_{ij} W_{ij}^\top}{(1+\exp (- W_{ij} \theta - \log (\pi_{ij}) ))^2} \exp(-W_{ij} \theta - \log (\pi_{ij})) + \frac{B}{\sigma^2} \right].
-$$
-Define the matrix $V_i (\theta)$ be the diagonal matrix with the $(j,j)$th entry equal to $\mu(W_{ij}; \theta, \pi_{ij}) \cdot (1-\mu(W_{ij}; \theta, \pi_{ij})$.
+% We derive the iteratively re-weighted least squares algorithm next. First, take the derivative with respect to $\theta$ to construct the Hessian
+% $$
+% - \left[ \sum_{i=1}^n \sum_{j=1}^{k_i} \frac{W_{ij} W_{ij}^\top}{(1+\exp (- W_{ij} \theta - \log (\pi_{ij}) ))^2} \exp(-W_{ij} \theta - \log (\pi_{ij})) + \frac{B}{\sigma^2} \right].
+% $$
+% Define the matrix $V_i (\theta)$ be the diagonal matrix with the $(j,j)$th entry equal to $\mu(W_{ij}; \theta, \pi_{ij}) \cdot (1-\mu(W_{ij}; \theta, \pi_{ij})$.
 Then we can express the Hessian more succinctly as $- \left[ \sum_{i=1}^n \W_i^\top V_i (\theta) \W_i + \frac{B}{\sigma^2} \right]$.
 Combining these, starting at an initial estimate, the Newton-Rhapson update given current parameter value $\hat \theta_k$ is given by
-$$
-\hat \theta_{k+1} = \hat \theta_{k} + \left[ \sum_{i=1}^n \W_i^\top V_i (\hat \theta_k) \W_i + \frac{B}{\sigma^2} \right]^{-1} \left(\sum_{i=1}^n \W_i^\top E_i (\hat \theta_k) - \frac{1}{\sigma^2} B \hat \theta_k \right)
-$$
-
-To learn $\sigma^2$, we use the connection to the mixed-effects literature.
-Fixing $\theta$, the log-likelihood components related to $\sigma^2$ is given by $-\frac{1}{2\sigma^2} \theta^\top B \theta - \frac{B {\bf 1}}{2} \log (\sigma^2)$.
+% $$
+% \hat \theta_{k+1} = \hat \theta_{k} + \left[ \sum_{i=1}^n \W_i^\top V_i (\hat \theta_k) \W_i + \frac{B}{\sigma^2} \right]^{-1} \left(\sum_{i=1}^n \W_i^\top E_i (\hat \theta_k) - \frac{1}{\sigma^2} B \hat \theta_k \right)
+% $$
+
+% To learn $\sigma^2$, we use the connection to the mixed-effects literature.
+% Fixing $\theta$, the log-likelihood components related to $\sigma^2$ is given by $-\frac{1}{2\sigma^2} \theta^\top B \theta - \frac{B {\bf 1}}{2} \log (\sigma^2)$.
 Differentiating and setting equal to zero yields the estimator
-$\hat \sigma^2 = \frac{\theta^\top B \theta}{B {\bf 1}}$.
+% $\hat \sigma^2 = \frac{\theta^\top B \theta}{B {\bf 1}}$.
 The complete algorithm is then alternating between Netwon-Rhapson for fixed $\sigma^2$ to estimate $\theta$ and then estimating $\sigma^2$ for fixed $\theta$ using the above formula.
 
-\subsection{Estimation of random-effects via Newton-Rhapson}
-
-\textcolor{red}{Assume fixed $\lambda$.}
-
-In standard logistic mixed-effects model, with $b_i \sim N(0, \Psi)$ and $b_i \in \mathbb{R}^q$, then the likelihood is given by
-$$
-\begin{aligned}
-& \prod_{i=1}^n \int \prod_{j=1}^{k_i} \left( \frac{\exp(W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij})}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{y_{ij}} \left( \frac{1}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{1-y_{ij}} \phi (b_i; \sigma_b^2) db_i \\
-\propto & \prod_{i=1}^n \int \exp \left[ \sum_{j=1}^{k_i} \left( y_{ij} \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) - \log \left( 1 + \exp \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) \right) \right) - b_i^\top \Psi^{-1} b_i /2 \right] db_i \\
-\propto& \prod_{i=1}^n \int \exp \left[ g(\theta, b_i, \pi_{ij}, \Psi) \right] db_i,
-\end{aligned}
-$$
-where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$.
+% \subsection{Estimation of random-effects via Newton-Rhapson}
+
+% \textcolor{red}{Assume fixed $\lambda$.}
+
+% In standard logistic mixed-effects model, with $b_i \sim N(0, \Psi)$ and $b_i \in \mathbb{R}^q$, then the likelihood is given by
+% $$
+% \begin{aligned}
+% & \prod_{i=1}^n \int \prod_{j=1}^{k_i} \left( \frac{\exp(W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij})}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{y_{ij}} \left( \frac{1}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{1-y_{ij}} \phi (b_i; \sigma_b^2) db_i \\
+% \propto & \prod_{i=1}^n \int \exp \left[ \sum_{j=1}^{k_i} \left( y_{ij} \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) - \log \left( 1 + \exp \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) \right) \right) - b_i^\top \Psi^{-1} b_i /2 \right] db_i \\
+% \propto& \prod_{i=1}^n \int \exp \left[ g(\theta, b_i, \pi_{ij}, \Psi) \right] db_i,
+% \end{aligned}
+% $$
+% where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$.
 Differentiating the function $g(\theta, b_i, \pi_{ij}, \Psi)$ with respect to $b_i$, the connection to the prior section can be used.
 Re-defining terms appropriately allows us to express the derivatives as
-$$
-\begin{aligned}
-\frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
-\frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
-\end{aligned}
-$$
-For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
-$$
-\hat b_i^{(k+1)} = \hat b_i^{(k)} + \left[ Z_i V_i (\theta, \hat b_i^{(k)}) Z_i + \Psi^{-1} \right]^{-1}  \left( Z_i E_i (\theta, \hat b_i^{(k)}) - \Psi^{-1} \hat b_i^{(k)} \right)
-$$
-
-Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
+% $$
+% \begin{aligned}
+% \frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
+% \frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
+% \end{aligned}
+% $$
+% For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
+% $$
+% \hat b_i^{(k+1)} = \hat b_i^{(k)} + \left[ Z_i V_i (\theta, \hat b_i^{(k)}) Z_i + \Psi^{-1} \right]^{-1}  \left( Z_i E_i (\theta, \hat b_i^{(k)}) - \Psi^{-1} \hat b_i^{(k)} \right)
+% $$
+
+% Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
 Here we employ adaptive gaussian quadrature (cites) as it has been shown that for infrequent events the Laplace approximation tends to underestimate effects (cites).
@@
 Then the AGQ rule is given by
+% $$
+% \begin{aligned}
+% &\int \exp \left[ g(\theta, b_i, \pi_{i}, \Psi) \right] db_i \\
+% =& \int | R_i |^{-1} \exp \left[ g(\theta, \hat b_i + R_i^{-1} \z, \pi_{i}, \Psi) + \| \z \|^2 / 2 \right] \exp ( - \| \z \|^2 / 2 ) d\z \\
+% =& (2 \pi)^{q/2} | R_i |^{-1} \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k}
+% \end{aligned}
+% $$
+% The score equations
+% $$
+% \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}}
+% \left( \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right)^{-1}
+% \left( \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi)
+% $$
+% Define $\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi)$ to be the weights.  Then we can express score and Hessian as:
+% $$
+% \begin{aligned}
+% &\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \\
+% &\sum_{\k}
+% \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial^2}{\partial \theta \partial \theta^\top} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) +
+% \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top \\
+% &- \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top
+% \end{aligned}
+% $$
+
+% When $N_{GQ} = 1$, the AGQ approximation is equivalent to a Laplacian approximation.
+
+\section{Optimal weights}
+
+\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
+The proof follows closely from~\cite{Rathbun2012} and~\cite{Ferreira82}.
+Revisiting the central limit theorem above, we can see that for any weights~$w(s; \theta_0)$ that the asymptotic normality is guaranteed with the asymptotic variance $\Sigma(\theta_0; w) := \Xi(\theta_0; w)^{-1} V(\theta_0; w) \Xi(\theta_0; w)^{-1}$ where
 $$
-\begin{aligned}
-&\int \exp \left[ g(\theta, b_i, \pi_{i}, \Psi) \right] db_i \\
-=& \int | R_i |^{-1} \exp \left[ g(\theta, \hat b_i + R_i^{-1} \z, \pi_{i}, \Psi) + \| \z \|^2 / 2 \right] \exp ( - \| \z \|^2 / 2 ) d\z \\
-=& (2 \pi)^{q/2} | R_i |^{-1} \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k}
-\end{aligned}
+V(\theta_0; w) = \int_0^\tau w^2 (s; \theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} \left[ \frac{\pi (s) + h(s;\theta_0)}{\pi (s)} \right] ds
 $$
-The score equations
-$$
-\sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}}
-\left( \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right)^{-1}
-\left( \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi)
-$$
-Define $\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi)$ to be the weights.  Then we can express score and Hessian as:
-$$
-\begin{aligned}
-&\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \\
-&\sum_{\k}
-\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial^2}{\partial \theta \partial \theta^\top} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) +
-\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top \\
-&- \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top
-\end{aligned}
-$$
-
-When $N_{GQ} = 1$, the AGQ approximation is equivalent to a Laplacian approximation.
-
-\section{Optimal weights}
-
-\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
-The above proof establishes asymptotic normality for all
-The main additional assumption is
-
-The argument
+and dependence on $w$ is made explicit for ease of comprehension.
+To prove optimality of our particular weight choice we study a standardized set of estimating equations~$\Xi (\theta; w)^{-1} U_n (\theta; w)$.
+Let $w^\star$ denote the Waagepeterson weights from~\eqref{eq:waage_weights}.
+Then consider the covariance of estimators using the standardized estimating equations with weights $w$ and $w^\star$ respectively.
+Based on the asymptotic normality arguments above, the covariance will have a similar sandwich form $\Xi (\theta_0; w)^{-1} \cov (\theta_0; w,w^\star) \Xi(\theta_0; w^\star)^{-1}$ where the middle term is given by
+\begin{align*}
+&\int_0^\tau w(s;\theta_0) w^\star (s; \theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} \left[ \frac{\pi (s) + h(s;\theta_0)}{\pi (s)} \right] ds \\
+=&\int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} ds \\
+&= \Xi (\theta_0; w).
+\end{align*}
+This implies the covariance is equal to $\Xi(\theta_0; w^\star)^{-1}$.
+Optimality is defined as the difference in variances being positive semi-definite.
+This can now be easily shown by considering the stacked estimating equations variance which is positive semi-definite by definition.
+For any~$\alpha \in \mathbb{R}^{p}$,
+\begin{align*}
+0 &\leq \left( \alpha^\top - \alpha^\top \right)
+\left( \begin{array}{c c} \Sigma(\theta_0; w) & \Xi (\theta_0; w^\star)^{-1} \\ \Xi(\theta_0; w^\star)^{-1} & \Xi (\theta_0; w^\star)^{-1} \end{array} \right)
+\left( \begin{array}{c} \alpha & - \alpha \end{array} \right)\\
+&= \alpha^\top \left( \Sigma(\theta_0; w) - \Xi (\theta_0; w^\star)^{-1} \right) \alpha.
+\end{align*}
+which demonstrates optimality of the Waagepetersen weights from~\eqref{eq:waage_weights}.
+
+Uniqueness is not guaranteed by the above.  Let $w$ be another estimator with minimal variance.  Then define the estimating equation $\tilde U_n (
+\theta; w) - \tilde U_n (\theta; w^\star)$.
+By definition of the estimating equations, the difference is simply another estimation equation of the same weighted form with different weights~$w_1$.
+Thus, under the standardized set of estimating equations we have the variance is equal to
+$\Sigma (\theta; w_1) = \Sigma (\theta; w) - \Sigma(\theta; w^\star)$,
+but $\Sigma (\theta; w^\star) - \Sigma(\theta; w)$ is non-negative definite by the above argument.
+This implies~$\Sigma(\theta;w_1) = 0$ and thus the two standardized estimating equations are equivalent.
+Thus, there exists a non-singular linear transformation (specifically, $\Xi (\theta_0; w) \Xi (\theta_0; w^\star)^{-1}$) that maps the estimating equations $U(\theta;w^\star)$ to those of the other optimal estimating equations~$U(\theta; w)$.
 
@@
 
+\section{Summary of missing data for the case study}
+\label{app:csmissing}
+
+Here we present missing data information for electrodermal EDA and activity index.
+Figure~\ref{fig:missing_eda} and~\ref{fig:missing_acc} presents the patient average fractino of missing EDA and AI data in 30-minute windows respectively.
+We see that the missingness levels vary by individual and are on average higher for EDA than AI.
+
+\begin{figure}[!th]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/missing_data_eda.png}
+  \caption{Electrodermal activity (EDA)}
+  \label{fig:missing_eda}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/missing_data_ai.png}
+  \caption{Activity Index (AI)
+}  \label{fig:missing_acc}
+\end{subfigure}
+\caption{Patient average fraction missingness in 30-minute windows.}
+\label{fig:mean_edacc}
+\end{figure}
+
 \end{document}

### 2023-02-02 letters/response-ref-1-v1.tex
updating paper first, then code for revision

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-1in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right
\usepackage{amsthm, amsfonts}
\usepackage{xcolor}
\usepackage{hyperref}

\def\Y{{\bf Y}}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor
	Peter Glynn\\
	Editor, {\em Journal of Applied Probability}}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan\\
Department of Biostatistics\\
1415 Washington Heights\\
Ann Arbor, MI 48103} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\vspace{5mm}

\newpage

{\bf Response to Referees}

We appreciate all of the very helpful comments from the editor, associate editor, and the two referees which has helped improve the paper substantially. We have addressed the concerns as best as possible in this revision. We provide a further point-by-point explanation below. Any comments in {\it italics} indicate a statement from the referee report.

{\bf Response to Associate Editor}
\begin{enumerate}
\item {\it This paper was reviewed by two referees, who tend to agree in their positive assessments of the value and novelty of your work, and provide some very specific comments for further improvements.}

\vspace{5mm}
We thank the associate editor.  We have addressed the two reviewers' specific comments.  Responses to their concerns are detailed below and have significantly improved the clarify of this manuscript.
\vspace{5mm}

\item {\it Expanded literature review on existing modeling approaches. Both referees raise this point, and I agree that the contrast with existing works could be more explicit. The second referee provides some specific suggestions on relevant works.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it There appear to be some missing definitions of notation, and some transitions/equations need more details to justify them. The latter could potentially be put in the supplement to save space. The former should be added to the main manuscript to improve clarity.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it As the first reviewer, I am confused by $\varepsilon(t) = 0$ on page 4, line 39, in line with equation (3) and measurement error methodology. I suspect comment 7 from the 2nd reviewer alludes to the same thing. I suspect this may be a notation issue rather than the modeling issue, but either way, it needs to be clarified.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it Practical considerations, such as window size and smoothing parameter selection, need clarification.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it Reviewer 2 suggests adding additional details on proof steps and lemma clarifications. I can't entirely agree that the proofs should go in the main manuscript as the paper is already on the longer side. When making adjustments to this section (theoretical analysis), please keep at most its current length, and move things to supplement as needed.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it Both reviewers suggest additional simulation studies. I agree that it would be good to have a numerical benchmark of the magnitude of the computational issue that you allude to in lines 16-28 on page 5, especially for JCGS readership. I don’t think you need references for this, as the reviewer asks, as a numerical example would be more helpful and convincing in this case. However, I think you need another method for comparison, as the simulations are only done for your approach with different sampling rates. Perhaps some suggestions from the first referee could be useful. If some of these alternatives are computationally prohibitive, then doing a computational numerical comparison on even 1 rep will provide a convincing argument for why those methods are not tried.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it I also have additional comments that were not raised. A major one is the absence of code. A user-friendly software with the method’s implementation and documentation is a must for accepted JCGS papers. Please provide the code so that the results are reproducible.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it
A minor one is the absence of truncation $K_x$ in the equation on line 37, page 8. Please reread everything carefully to ensure all notations are defined and the equations are consistent with the text.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it Another minor one is your choice of smoothing the entire covariance, including the diagonals (page 8, lines 13-18). It’s unclear to me what the rationale is for changing this step compared to Park and Staicu, 2015, and how much it matters for the final performance.}

\vspace{5mm}
XXXX
\vspace{5mm}

\item {\it Expanded literature review on existing modeling approaches. Both referees raise this point, and I agree that the contrast with existing works could be more explicit. The second referee provides some specific suggestions on relevant works.}

\vspace{5mm}
XXXX
\vspace{5mm}

\end{enumerate}

\newpage
{\bf Response to Reviewer 1}
\vspace{5mm}

I am very grateful to the referee for the incredibly detailed review.  The detailed suggestions and comments helped improve the paper in numerous ways.  See below for our responses to each particular reviewer comment.

\begin{enumerate}
\item {\it The introduction it quite lacking. Please add more background literature review of prevailing/latest approaches (particularly in the area of joint modeling of longitudinal and time to event data) to the introduction.}

\vspace{5mm}
XXXXX
\vspace{5mm}

\item {\it The measurement error part can be added to the end with other exten-
sions. It does not seem to play a major role in developing the algorithm.}

\vspace{5mm}
We agree.  This point was over-emphasized in the original manuscript and so has been moved to the
\vspace{5mm}

\item {\it In equationn (4), page 4, do you mean that equation (2) and the mesurement error part $x_i (t) = \eta_i (t) + \epsilon_i (t)$ imply conditional hazard equation in (3)?  If yes, then why is $\int_{t-\Delta}^t \epsilon_i (s) \beta(s) ds = 0$?  Or are you assuming a model?  Please add more clarity.}

\vspace{5mm}
XXXXX
\vspace{5mm}

\item {\it Several notations in section 2.1 such as ${\bf x}_i, {\bf H}_{i,\tau_i}^X$ are not defined.}

\vspace{5mm}
XXXXX
\vspace{5mm}

\item {\it How did you get the equation in line 24, page 4.}

\vspace{5mm}
XXXXX
\vspace{5mm}

\item {\it In line 39, page 4, by assuming $\epsilon(t) = 0$, you assume there is no measure- ment error? Why is section 2.1 called measurement error if the error is 0? Should this be named missing data section instead?}

\vspace{5mm}
The key point raised on page 4 (no Section XX), is that if $\epsilon(t)$ is not equivalently zero, your model assumes risk of an event at time $t$ depends on both the current and future measurements of the process ${\bf x}_i$. We argue this is not sensible since current risk should not depend on future risk and thus set $\epsilon(t) = 0$.
\vspace{5mm}

\item  {\it Is $\theta$ defined anywhere prior to section 2.2?}

\vspace{5mm}
It was used in equation (1) to refer to all the parameters underlying the model to distinguish it from the functional parameter~$\beta (s)$.
We do XXXX to make this clearer.
\vspace{5mm}

\item {\it On page 5, lines 16 - 28 you claim that computing the double integral is computationally expensive. Add some references. Also, to give a greater insight to the extent of the computational cost, solve the score equation in lines 8,9 i.e. without the sub-sampling framework in the simulations (there is another comment on this for the simulation section). Reporting a rough summary of computational cost such as ” it took x min to solve equations involving n samples” here, would help the reader understand
the extent of the problem.}

\vspace{5mm}
We explain that if we were to compute it would have to be as a function of $\beta(s)$.  We implement a naive and demonstrate numerically the computational cost.  Specifically, each computation in simulation takes XXseconds, and we must compute this XX times to compute the likel
\vspace{5mm}

\item {\it Page 5, line 51, what is $t_i$?}

\vspace{5mm}
This should be $\tau_i$ which is the censoring time of the observation process for individual~$i$.
\vspace{5mm}

\item {\it Check the notation in equation (6) page 6. Some subscripts seem to be missing. Add some details on how to obtain this from the original score
equation. What is the approximation here?}

\vspace{5mm}
We have included correct subscripts in the revision.  Derivation of (6) follows from plugging in (4) into equation the score equations.  We now make this clear and provide a derivation in the supplementary materials for completeness.
\vspace{5mm}

\item{\it The remark 2.1 can be added the end of section 2.2. This along with
comment 8 will naturally lead the way your proposed method.}

\vspace{5mm}
We thank the referee for this suggestion.  We have now used Remark 2.1 and comment 8 on the computational costs to the end of section 2.2 to motivated our proposed method.
\vspace{5mm}

\item {\it Page 7, line 48, in the definition of the marginal covariance, it should be $dT$ in the integral.}

\vspace{5mm}
That is correct.  We have fixed the definition as suggested.
\vspace{5mm}

\item {\it Page 7, line 48, in the definition of the marginal covariance, what is $\tau$?}

\vspace{5mm}
The term $\tau$ refers to the censoring time of the observation process. We make this clear in Section XXX, page XX in the revision.
\vspace{5mm}

\item {\it s there a difference between the mean and covariance functions for $y = 0$ and $y = 1$ especially if points in $D_i$ and $T_i$ are spread roughly evenly on a grid even if the sets are disjoint? For example, pick a grid of equidistant points and number them from $1-10$. If all odd points are events in $T_i$ and even ones are sampled in $D_i$, won't the pooled sample covariance for both be the same}

\vspace{5mm}
The event times~$T_i$ depend heavily on the event intensity function~$h_i (t;\theta)$ and will not be guaranteed to be uniformly spread across points in $[0,\tau]$.  The subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$ and
\vspace{5mm}

\item {\it Add a plot of the mean, covariance functions for $y=0$ and $y=1$.}

\vspace{5mm}
We have added mean and covariance functions to the supplementary materials as suggested.  The plots demonstrate distinct behavior in the means for $y=0$ and $y=1$; while the covariance functions are XXXX.
\vspace{5mm}

\item {\it Page 9, line 6, isn’t the integral in (2) $\int_{t-\Delta}^t X(s) \beta(s)ds$? The whole development assumes the integral to be $\int_{t-\Delta}^t X(t−s) \beta(s)ds$. The interpretation of the coefficient function $\beta()$ would be different for two cases. Please clarify.}

\vspace{5mm}
We apologize for the confusion.  In Section 3.1, we started by defining the double-indexed $X(t,s)$ to be equal to $X(t-s)$ in order to make the notation a bit cleaner but we see that this has added confusion.  We not add a remark to make this more clear.  The reason for the notation was that writing $X(t,s)$ allows us to write the integrals over $0$ to $\Delta$ rather than having to constantly index integrals by $t$.  Moreover, we thought it helps the reader to understand we are saying ``the process $s$ units prior to time~$t$.''
\vspace{5mm}

\item {\it Page 9, equation 7, please define all of the notations before using them.}

\vspace{5mm}
We have re-written this paragraph to first define the notation and then present equation (7).
\vspace{5mm}

\item {\it Page 12, line 48, what is proposition 3.6? You mean Lemma?}

\vspace{5mm}
That is correct, we have fixed this issue.
\vspace{5mm}

\item {\it When are the assumptions satisfied or not satisfied? An example of a
situation where the functional and event process assumptions are easily satisfied will be useful.}

\vspace{5mm}
We have shown that the functional process satisfies these moment conditions under boundedenss which is
\vspace{5mm}

\item {\it All the modes of convergences in Lemma 3.5 and its proofs should be made clear.}

\vspace{5mm}
XXXxXX
\vspace{5mm}

\item {\it Outline the proof of Lemma 3.5 first and then prove the details.  It will be easier to follow.  Alternatively, we }

\vspace{5mm}
We have added a sketch of the proof with an outline as part of the supplementary details.
\vspace{5mm}

\item {\it In Lemma 3.6 state the theorem clearly.  What condition do yu show in the proof that establishes optimality}

\vspace{5mm}
We have now included the statement of how we define optimality in the main paper to make the theorem statement clear.  Proof is kept in the supplementary materials for conciseness.
\vspace{5mm}

\item {\it Combine section 3.6 and 3.7 in to a single result that establishes a confidence band for $\beta (t)$. Also, wont Lemma 3.5 lead to a confidence band for $\hat \beta (t)$? Are the two bands same?}

\vspace{5mm}
We discuss the penalization due to the potential complexity of the.  The confidence bands are based on Lemma 3.5.  We make this clear in the revision.  The issues are still thre and we keep the discussion.  We keep as 2 separate sections since they cover different issues.
\vspace{5mm}

\item The ``original'' score equation Page 5, line 9 involves
$$
H_i (\tau_i ;\theta) = \int_{t=0}^{\tau_i} h_i (t; \theta)dt = \int_0^{\tau_i} \left[ h_0 (t;\gamma) \exp \left( g_t (H_{i,t}^N)^\top \alpha + \int_{s=t-\Delta}^t x_i (s) \beta(t) ds \right) \right] dt
$$
In the subsampled version you suggest in equation 6, page 6, you essentially replace $H_i (\tau_i; \theta)$ computation with $h_i (u; \theta)$, $u \in D_i$. How much time would it take to solve the full likelihood equation that will involve the above equation (1) directly? With the basis expansion strategy, you
would will be estimating the same number of parameters. This type of approach should be added to simulations and evaluated for accuracy and time complexity.

\vspace{5mm}
XXXXX
\vspace{5mm}

\item {\it How about using a concurrent model for $h_i (t;\theta) = h_0 (t; \theta) \exp (g_t (H_{i,t}^N)^\top \alpha + x_i (t) \beta (t))$. This does not look at the past window like the suggested model but this drawback might be compensated with the constructed features $g_t (H_{i,t}^N)$ that capture the past. The concurrent approach might be much less expensive computationally. This can be added as an alternative to the simulations.}

\vspace{5mm}
XXXXX
There is no obvious way to compute bias since the model will be incorrectly specified.  We do show the computation trade-off and that computing the cumulative hazard is still computationally demanding; therefore, after sampling non-event times, the added computation of the interior integral is not very expensive and avoids model misspecification.
\vspace{5mm}

\item {\it In section 5.2.1 what is uncongeniality? Does it have a mathematical
meaning?}

\vspace{5mm}
uncongenial comes from Meng. We cite and provide a short formal definition with some intuition.
\vspace{5mm}

\item Page 23, Figure 2. EDA seems to spike at the end. Even AI seems to be
turning up. Add longer history to both the Figures.

\item {\it Doesn’t Figure 3 suggest smaller ∆?}

\vspace{5mm}
We agree.  We re-run the analysis with a window length of 15 minutes as well and move the current analysis as a sensitivity to the supplementary materials.
\vspace{5mm}

\item {\it What features were constructed for the real data analysis?}

\vspace{5mm}
As this was illustrative we only included time-of-day.
\vspace{5mm}

\end{enumerate}

\newpage

{\bf Response to Reviewer \#2}

I am incredibly grateful for this review.  I agree that the paper has evolved beyond the current title.  This was not updated in the prior submission simply to ensure continuity of presentation (i.e., didn't want to change from first submission).  Here we replace with the title ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information''.  We next address the reviewers specific comments.

\begin{enumerate}
\item {\it The paper states some facts about the SARS-CoV-2 virus that are outdated, for example “Upon infection with SARS-CoV-2, an incubation period starts and lasts approximately five days [Lauer et al., 2020] before the viral load is high enough to be detected.” (Page 5). Lauer was based on the original
strain, we see different disease dynamics with the delta strain. Perhaps to be more precise, the author could clarify that this is an example of the original SARS-CoV-2 incubation rather than definitative as there will likely be future variants that will change this further. Similarly, on page 13, the serial
interval of 7 is stated as fact – this is again based on the original strain rather than the current disease dynamics. Stating these numbers as an example rather than definitive would be preferable, I believe.}
\vspace{5mm}
\end{enumerate}

\end{letter}

\end{document}

### 2023-02-02 letters/cover letter-fdarecurrent.tex
updating paper first, then code for revision

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.75in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=0in % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

%\address{Rutgers University and University of Michigan}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professors Galin Jones and Faming Liang \\
    Co-Editors, {\em Journal of Computational and Graphical Statistics}
         }
%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan \\
Department of Biostatistics\\
M4057 SPH II \\
1415 Washington Heights \\
wdem@umich.edu} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\date\today

\opening{Dear Drs. Jones and Liang:}

Kindly find enclosed the revised version of the manuscript ID JCGS-22-236, "Recurrent event analysis in the presence of real-time high frequency data via random subsampling."   We have revised the article to address the referees’ requests.  The reports called for a major revision to include related work on modeling event data with functional data predictors as well as historical functional linear models, include empirical evidence and associated discussion of the computational benefits of our proposal, as well as clarify several technical details in the theoretical sections and provide important intuition/sketches of the proofs. The new version complies with these requests.

We believe the work presents an important advance that gives researchers a simple tool to analyze recurrent event data in the presence of high-frequency sensors data.  Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.

Thank you for your consideration. We look forward to hearing back from you.

\closing{Sincerely,}

\end{letter}

\end{document}

### 2023-02-02 write-up/fda-recurrent-supplement.tex
updating paper first, then code for revision

--- 2022-08-03 write-up/fda-recurrent.tex
@@
 \doublespacing
+
+\usepackage{xr}
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}% latexmk will find this if $recorder=0 (however, in that case, it will ignore #1 if it is a .aux or .pdf file etc and it exists!
+if it doesn't exist, it will appear in the list of dependents regardless)
+  \@addtofilelist{#1}% if you want it to appear in \listfiles, not really necessary and latexmk doesn't use this
+  \IfFileExists{#1}{}{\typeout{No file #1.}}% latexmk will find this message if #1 doesn't exist (yet)
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+%%% END HELPER CODE
+
+% put all the external documents here!
+\myexternaldocument{fda-recurrent}
 
@@
 
-\title[Recurrent event analysis with functional covariates via random subsampling]{Recurrent event analysis in the presence of real-time high frequency data via random subsampling}
+\title[Supplement to ``Recurrent event analysis with functional covariates via random subsampling'']{Supplementary Materials to ``Recurrent event analysis in the presence of real-time high frequency data via random subsampling''}
 % \author{Walter Dempsey}
@@
 \begin{abstract}
-Digital monitoring studies collect real-time high frequency data via mobile sensors in the subjects' natural environment.
-This data can be used to model the impact of changes in physiology on recurrent event outcomes such as smoking, drug use, alcohol use, or self-identified moments of suicide ideation.
-Likelihood calculations for the recurrent event analysis, however, become computationally prohibitive in this setting.
-Motivated by this, a random subsampling framework is proposed for computationally efficient, approximate likelihood-based estimation.
-A subsampling-unbiased estimator for the derivative of the cumulative hazard enters into an approximation of log-likelihood.
-The estimator has two sources of variation: the first due to the recurrent event model and the second due to subsampling.
-The latter can be reduced by increasing the sampling rate; however, this leads to increased computational costs.
-The approximate score equations are equivalent to logistic regression score equations, allowing for standard, ``off-the-shelf'' software to be used in fitting these models.
-Simulations demonstrate the method and efficiency-computation trade-off.
-We end by illustrating our approach using data from a digital monitoring study of suicidal ideation.
+This is the supplementary materials to the manuscript titled ``Recurrent event analysis in the presence of real-time high frequency data via random subsampling.''  Section~\ref{app:prooflemma31} contains the proofs for Lemmas~\ref{lemma:logistic}, Lemma~~\ref{lemma:simpleasym}.
+Section~\ref{app:optimalweights} contains that proof of Proposition~\ref{prop:optimal}.
+Section~\ref{app:extrasims} contains additional simulations and case study details.
 \end{abstract}
@@
 
-\section{Introduction}
-
-Advancement in mobile technology has led to the rapid integration of mobile and wearable sensors into behavioral health~\citep{Freeetal2013}.
-Take HeartSteps, for example, a mobile health (mHealth) study designed to increase physical activity in sedentary adults \citep{KlasnjaHS2019}.
-Here, a Jawbone sensor is used to monitor step count every minute of the participant's study day.
-Of interest in many mHealth studies is the relation of such real-time high frequency sensor data to an adverse, recurrent event process.
-In a smoking cessation mHealth study~\citep{Sense2Stop}, for example, the relation between a time-varying sensor-based measure of physiological stress and smoking lapse is of scientific interest.
-In a suicidal ideation mHealth study~\citep{Kleiman2018}, the relation of electrodermal activity (EDA) and accelerometer with self-identified moments of suicidal ideation is of scientific interest.
-
-The goal of this paper is two-fold: (1) to discuss the appropriate choice of statistical model for joint high-frequency sensor data and the recurrent event data, and (2) to construct a simple, easy-to-implement method for parameter estimation and inference.
-For (1), we discuss an important issue regarding measurement-error models when paired with recurrent event outcomes.
-For (2), we introduce a random subsampling procedure that has several benefits.
-First, the resulting inference is unbiased; however, there is a computation-efficiency trade-off.
-In particular, a higher sampling rate can decrease estimator variance at the cost of increased computation.
-We show via simulations that the benefits of very high sampling rates is often negligible, as the contribution to the variation is small relative to the variation in the underlying stochastic processes.
-Second, derived estimating equations are optimal, implying loss of statistical efficiency is only due to subsampling procedure and not the derived methodology.
-Finally, implementation can leverage existing, standard software for functional data analysis and logistic regression, leading to fast adoption by domain scientists.
-
-\section{Recurrent event process and associated high frequency data}
-
-Suppose $n$ subjects are independently sampled with observed event times~$\bfT_{i} = \{ T_{i,1}, \ldots, T_{i,k_i}\}$ over some observation window $[0, \tau_i]$ for each subject $i = 1,\ldots, n$.
-Assume the event times are ordered, i.e., $T_{i,j} < T_{i,j^\prime}$ for $j < j^\prime$.
-The observation window length, $\tau_i$, is the censoring time and is assumed independent of the event process.
-Let~$N_{i} (t)$ denote the associated counting process of $\bfT_{i}$; that is, $N_i (t) = \sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.
-In this section, we assume a single-dimensional health process~$\bfx_i = \{ x_i (s) \}_{0 < s < \tau_i}$ for each participant is measured at a dense grid of time points.
-Accelerometer, for example, is measured at a rate of 32Hz (i.e., $32$ times per second).
-Electrodermal activity (EDA), on the other hand, is measured at a rate of 4Hz (i.e., 4 times per second).
-Given the high frequency nature of sensor data, this paper assumes the process is measured continuously.
-
-Let~$H_{i,t}^{NX} = H_{i,t}^{N} \otimes H_{i,t}^{X}$ be the $\sigma$-field generated by all past values~$(N_i (s), x_i (s))_{0 \leq s \leq t}$.
-In this paper, the instantaneous risk of an event at time~$t$ is assumed to depend on the health process, time-in-study, and the event history through a fully parametric conditional hazard function:
-\begin{equation}
-\label{eq:hazard}
-h_i \left( t \Given H_{i,t}^{NX} ; \theta \right) =
-\lim_{\delta \to 0} \delta^{-1} \pr \left( N_i(t+\delta) - N_i(t) > 0
-  \given H_{i,t}^{NX} \right),
-\end{equation}
-where~$\theta$ is the parameter vector.
-For high frequency physiological data, we assume that current risk is log-additive and depends on a linear functional of the health process over some recent window of time and pre-specified features of the counting process; that is,
-\begin{equation}
-\label{eq:hazardlinear}
-h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) =
-h_0 (t; \gamma) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + \int_{t-\Delta}^t x_i (s) \beta(s) ds  \right)
-\end{equation}
-where~$h_0(t;\gamma)$ is a parametrized baseline hazard function,~$\Delta$ is an unknown window-length, and $g_t( H_{i,t}^N ) \in \mathbb{R}^p$ is a $p$-length feature vector summarizing the event-history and time-in-study information.
-The final term~$\int_{t-\Delta}^t x_i(s) \beta(s) ds$ reflects the unknown linear functional form of the impact of the time-varying covariate on current risk.
-
-An alternative to~\eqref{eq:hazardlinear} would be to construct features from the sensor data history~$f_t ( H_{i,t}^{X}) \in \mathbb{R}^q$ and incorporated these features in the place of the final term.
-Our current approach builds linear features of~$H_{i,t}^X$ directly from the integrated history, avoiding the feature construction problem -- a highly nontrivial issue for high frequency time-series data.
-The main caveat is the additional parameter~$\Delta$; however, as long as the estimated~$\hat \Delta$ exceeds~$\Delta$, then resulting estimation is unbiased albeit at a loss of efficiency.
-Moreover, sensitivity analysis can be performed to determine how choice of $\hat \Delta$ affects inference.
-One limitation of the approach presented here is that only fully parametric hazard models may be fit to the data.
-However, a spline model for the log baseline hazard affords sufficient model flexibility.
-
-\subsection{Measurement-error models with event processes}
-\label{section:memproblems}
-
-One potential criticism for~\eqref{eq:hazard} is that the health process may be measured with error.
-A common mathematical strategy for joint models is to consider an unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i \given \eta_i$, i.e., the two processes are conditionally independent given the latent trajectory.
-For example, take~$\eta_i$ to be a zero-mean Gaussian process with
-\begin{equation}\label{eq:jm}
-x_i(t) = \eta_i (t) + \epsilon_i (t),\quad \text{ and } \log h_i (t
-\given \eta ) = \log h_0 (t) + g_t \left( H_{i,t}^N \right)^{\prime}
-\alpha + \int_{t-\Delta}^t \eta_i (s) \beta (s) ds
-\end{equation}
-where~$\epsilon_i(t)$ is a white-noise measurement error term.
-Thus, $\eta_i (t)$ is the ``true and unobserved value of the longitudinal outcome'' \citep[Sec.
-2.1, pp.3]{Rizopoulos2010}.
-The conditional survival function of the $j$th event~$T_j$ given~$\bfx_i$ and all prior events~$\bfT_{-j} := \{ T_1, \ldots, T_{j-1} \}$ is
-\[
-\pr \left ( T_j > t+s \Given \H_{i,\tau_i}^X, \bfT_{-j} = \bft_{-j}, T_j > t \right) = E \left( - \exp \left( \int_{t}^{t+s} h_i (u \given \eta ) du \right) \Given \bfx_i, \bfT_{-j} = \bft_{-j} \right),
-\]
-where the expectation is a Gaussian integral, albeit infinite-dimensional.
-For most choices of covariance structure, if the white-noise error term is non-zero then the above calculation will show the conditional survival function depends not only on past $x$-values, but also on future $x$-values, i.e.,~$\bfx$ and~$\bfT$ do not satisfy \emph{independent evolution}~\citep{DempseyPMCC2}.
-This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process.
-To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
-Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed; see Section~\ref{section:missingdata} for details on handling of missing data.
-
-\subsection{Likelihood calculation}
-
-For the sake of notational simplicity, we leave the dependency of the conditional hazard function on~$H_{i,t}^{NX}$ implicit, and write~$h_i (t; \theta)$.
-The component of the log-likelihood related to the event process is then given by
-\[
-  L_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i}
-    \log \left( h_i \left( T_{i,j}; \theta
-      \right) \right) - H_{i} \left( \tau_i; \theta \right) \right)
-\]
-where~$H_{i} (\tau_i ; \theta) = \int_{0}^{\tau_i} h_{i} (t; \theta) dt$ is the cumulative hazard function.
-Solving the associated score equations~$U_n (\theta) = {\bf 0}$ yields the maximum likelihood estimator~$\hat \theta$, where
-\[
-U_n (\theta) = \sum_{i=1}^{n} \left ( \sum_{j=1}^{k_i} \frac{h^{(1)}_i
-    (T_{i,j}; \theta)}{h_i (T_{i,j}; \theta)} - H^{(1)}_{i} (\tau_i;
-  \theta) \right),
-\]
-with $h_i^{(1)} (T_{i,j}; \theta)$ and $H_i^{(1)} (\tau_{i}; \theta)$ are derivatives with respect to $\theta$.
-
-In classical joint models~\citep{Henderson2000, Tsiatis2004}, time-varying covariates~$x_i (t)$ are observed only intermittently at appointment times.
-In our current setting, maximizing the likelihood is computationally prohibitive since for any~$\theta$ we must compute the cumulative hazard functions~$H_{i} (\tau_i; \theta)$ which requires integration of~$h_i(t;\theta)$ given by~\eqref{eq:hazardlinear} which itself depends on the integral~$\int_{t-\Delta}^t x_i (s) \beta(s) ds$ that is a function of the unknown functional parameter $\beta(\cdot)$.
-That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
-
-\subsection{Probabilistic subsampling framework}
-
-To solve the computational challenge we employ a point-process subsampling design to obtain unbiased estimates of the derivative of the cumulative hazards for each subject.
-The subsampling procedure treats the collected sensor data as a set of \emph{potential observations}.
-Suppose covariate information is sampled at times drawn from an independent inhomogeneous Poisson point process with known intensity~$\pi_i (t)$.
-At a subsampled time~$t$, the \emph{windowed covariate history} $\{ x_i (t-s)\}_{0 \leq s \leq \Delta}$ and counting process features~$g_t (H_{i,t}^N)$ are observed.
-Optimal choice of~$\pi_i (t)$ is beyond the scope of this paper; however, simulation studies have suggested setting the subsampling rate proportional to the hazard function~$h_i (t; \theta)$.
-
-An estimator is design-unbiased if its expectation is equal to that parameter under the probability distribution induced by the sampling design~\citep{Cassel1977}.
-Let~$\bfD_i \subset [0,t_i]$ denote the random set of subsampled points.
-Note, by construction, this random set is distinct from the set of event times with probabilty one, i.e.,~$\pr( \bfT_i \cap \bfD_i = \emptyset) = 1$.
-Under subsampling via $\pi_i (t)$, one may compute a Horvitz-Thompson estimator of the derivative of the cumulative hazard $\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in \bfD_i} h^{(1)}_i (u; \theta)/\pi_i (u)$.
-An alternative design-unbiased estimator of the derivative of the cumulative hazards is given by
-\begin{equation}
-\label{eq:WPest}
-\hat H_{i}^{(1)} (\tau_i; \theta) = \sum_{u \in (\bfT_i \cup \bfD_i)}
-\frac{ h^{(1)}_i ( u; \theta ) }{ \pi_i (u) + h_i ( u ; \theta) }
-\end{equation}
-Equation~\eqref{eq:WPest} is the estimator suggested by~\cite{Waagepetersen2008}.
-This estimator depends on the superposition of the event and subsampling processes.
-Proposition~\ref{prop:optimal} shows the estimator for~$\theta$ associated with using~\eqref{eq:WPest} is the most efficient within a suitable class of estimators for the derivative of the cumulative hazard function (including the Horvitz-Thompson estimator).
-Therefore, we restrict our attention to~\eqref{eq:WPest} for the remainder of this paper.
-Letting
-\begin{equation}
-\label{eq:waage_weights}
-w_i (t; \theta) = \frac{\pi_i (t)}{\pi_i (t) + h_i (t ; \theta)},
-\end{equation}
-the resulting approximate estimating equations can be re-written as
-\begin{equation}
-\label{eq:approxscore}
-\hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
-  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
-  \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
-\end{equation}
-Equation~\eqref{eq:approxscore} represents the approximate score functions built via plug-in of the design-unbiased estimator of the derivative of the cumulative hazard given in~\eqref{eq:WPest}.
-
-\begin{rmk}[Connection to design-based estimation]
-The approximate score equations given by~\eqref{eq:approxscore} arise in design-based inference of point processes.
-Design-based inference is common for spatial point processes~\citep{Waagepetersen2008} where the spatial varying covariate is observed at a random sample of locations.
-It is common in mobile health where ecological momentary assessments~\citep{Rathbun2012,Rathbun2016} are used to randomly sample individuals at various time-points to assess their emotional state.
-In the current setting, we leverage these ideas to form a subsampling protocol that can substantially reduce computationally complexity.
-Therefore, the purpose is quite different.
-Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.
-\end{rmk}
-
-\section{Longitudinal functional principal components within
-  event-history analysis}
-
-Probabilistic subsampling converts the single sensor stream~$\bfx_i$ into a sequence of functions observed repeatedly at sampled times~$\bfD_i$ and event times~$\bfT_i$ over windows of length~$\Delta$.
-Such a data structure is commonly referred to as \emph{longitudinal functional data}~\citep{Xiao2013, GoldSmith2015}.
-Given the large increase in longitudinal functional data in recent years, corresponding analysis has received much recent attention~\citep{Morris2003, MorrisCarroll2006, Baladandayuthapani2008, Di2009, Greven2010, Staicu2010, ChenMuller2012, LiGuan2014}.
-Here, we combine work by~\cite{Park2018} and~\cite{Goldsmith2011} to construct a computationally efficient penalized functional method for solving the estimation equations~$\hat U_n (\theta)$.
-
-\subsection{Estimation of the windowed covariate history}
-\label{sec:margcov}
-We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
-The sandwich smoother~\citep{Xiao2013} is used to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
-Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009}, bivariate tensor product splines~\citep{Wood2006}, and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
-The sandwich smoother was chosen for its computational efficiency and estimation accuracy.
-We then define~$\tilde X(t,s) = X(t,s) - \hat \mu_y(t,s)$ to be the mean-zero process at each time~$t \in \bfT_i \cup \bfD_i$.
-
-As in~\cite{Park2018}, define the \emph{marginal covariance} by
-\[
-\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (T,s), (T,s^\prime) ) f_y(T) dt.
-\]
-for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((t,s), (t,s^\prime))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $f_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
-Estimation of~$\Sigma_y$ occurs in two steps.
-For simplicity, we present the steps for subsampled times (i.e., $y=0$) but the steps are the same for event times as well.
-First, the pooled sample covariance is calculated at a set of grid points:
-\[
-\tilde \Sigma_0 (s_r , s_{r^\prime}) = \left( \sum_{i=1}^n \left | \bfD_i \right |\right)^{-1} \left( \sum_{i=1}^n \sum_{t \in \bfD_i} \tilde X (t,s_r) \tilde X (t,s_{r^\prime}) \right).
-\]
-Due to our concern over independent evolution as discussed in section~\ref{section:memproblems}, we do not assume that each observation is observed with white noise and therefore the diagonal elements of~$\hat \Sigma_0$ are not inflated.
-Second, the estimator~$\hat \Sigma$ is further smoothed again using the sandwich smoother~\citep{Xiao2013}.
-Note~\cite{Park2018} smooth the off-diagonal elements, while here we smooth the entire pooled sample covariance matrix.
-All negative eigenvalues are set to zero to ensure positive semi-definiteness.
-The result is used as an estimator~$\hat \Sigma_0$ for the pooled covariance $\Sigma_0$.
-
-Next, we take the spectral decomposition of the estimated covariance function; let $\{ \hat \psi^{(0)}_k (s),\hat \lambda^{(0)}_k \}_{k \geq 1}$ be the resulting sequence of eigenfunctions and eigenvalues.
-The key benefit of the marginal covariance approach is that it allows us to compute a single, time-invariant basis expansion; this reduces the computational burden by avoiding the three dimensional covariance function (i.e., covariance depends on~$t$) and associated spectral decomposition in methods considered by \cite{ChenMuller2012}.
-Using the Karhunen-Lo{\`e}ve decomposition, we can represent~$X(t,s)$ for $t \in \bfT_i \cap \bfD_i$ by
-\[
-X(t,s) = \hat \mu_y (t,s) + \sum_{k=1}^{\infty} \hat c^{(y)}_{i,k} (t) \hat
-\psi^{(y)}_k (s) \approx \hat \mu_y (t,s) + {\bf c}^{(y)}_{i} (t)^\top \mathbold{\hat
-  \psi}^{(y)} (s)
-\]
-where $\hat c^{(y)}_{i,k} (t) = \int_{t-\Delta}^t \tilde X_i (t,s) \hat \psi^{(y)}_k (s) ds$, ${\bf c}^{(y)}_i (t) = (c^{(y)}_{i,1} (t), \ldots, c^{(y)}_{i,K_x} (t))^\top$, $\mathbold{\hat \psi}^{(y)} (s) = (\hat \psi^{(y)}_1 (s), \ldots, \hat \psi^{(y)}_{K_x} (s))^\top$, and~$K_x < \infty$ is the truncation level of the infinite expansion.
-Following~\cite{Goldsmith2011}, we set~$K_x$ to satisfy identifiability constraints (see Section~\ref{section:beta} for details).
-In subsequent sections, we leave the dependence on $y$ (i.e., whether $t \in \bfT_i$ or $\in \bfD_i$) implicit unless required for notational simplicity.
-
-\subsection{Estimation of~$\beta(s)$}
-\label{section:beta}
-
-The next step of our method is modeling~$\beta(s)$.
-Here, we leverage ideas from the penalized spline literature~\citep{Ruppert2003, Wood2006book}.
-Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots, \phi_{K_b} (x) \}$ be a spline basis and assume that~$\beta(s) = \sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$ where~${\bf b} = [b_1, \ldots, b_{K_b}]^{\top}$.
-Thus, the integral in~\eqref{eq:hazardlinear} can be restated as
-\begin{align*}
-\int_{t-\Delta}^t X(t,s) \beta(s) ds
-  &\approx \int_{t-\Delta}^t \left[ \hat \mu(t,s) + {\bf c} (t)^\top
-    \mathbold{\hat \psi} (s) \right] \times \left[
-    \mathbold{\phi} (s) {\bf b} \right] ds \\
-  &= [ M_{i,t}^\top + {\bf c} (t)^\top J_{\hat \psi, \phi} ] {\bf b}
-\end{align*}
-where~$M_{t} = (M_{1,t}, \ldots, M_{K_b,t})$, $M_{j,t} = \int_{t-\Delta}^t \hat \mu (t,s) \phi_j (s)$, and~$J_{\hat \psi, \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s) ds$~\citep{RamsaySilverman2005}.
-
-Given the basis for~$\beta(t)$, the model depends on choice of both~$K_b$ and~$K_x$.
-We follow~\cite{Ruppert2002} by choosing $K_b$ large enough to prevent under-smoothing and~$K_x \geq K_b$ to satisfy identifiability constraints.
-While our theoretical analysis considers truncation levels that depend on~$n$, in practice, we follow the simple rule of thumb and set~$K_b = K_x = 35$.
-As long as the choices of~$K_x$ and $K_b$ are large enough, their impact on estimation is typically negligible.
-Below, we will exploit a connection between~\eqref{eq:approxscore} and score equations for a logistic regression model.
-Before moving on, we introduce some additional notation.
-Define
-\begin{equation}
-\label{eq:approx_hazard}
-h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) \approx
-\exp \left( Z_{t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\hat \psi, \phi} {\bf b} \right)
-= \exp \left( W_{i,t}^\top \theta \right),
-\end{equation}
-where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline intensity function.
-We write~$\tilde U_n (\theta)$ to denote the approximate score function when substituting in~\eqref{eq:approx_hazard} for~\eqref{eq:hazardlinear}.
-
-\subsection{Connection to logistic score functions}
-\label{eq:logistication}
-
-We next establish a connection between the above approximate score equations~$\tilde U_n (\theta)$ and the score equations for a logistic regression model.
-We can then exploit this connection to allow the model to be fit robustly using standard mixed effects software~\citep{Ruppert2002, McCulloch2001}.
-
-\begin{lemma} \normalfont
-\label{lemma:logistic}
-Under weights~\eqref{eq:waage_weights} and the log-linear intensity function~\eqref{eq:approx_hazard}, the approximate score function~$\tilde U_n (\theta)$ is equivalent to
-\[
-\sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i } \left[ {\bf 1}[t \in D_i]
-  - \frac{1}{1 + \exp \left[- \left( \tilde{W}_{i,t}^\top \theta +
-        \log \pi_i (t) \right) \right]} \right] \tilde{W}_{i,t}
-\]
-where~$\tilde W_{i,t} = -W_{i,t}$.
-This is the score function for logistic regression with binary response~$Y_i(t)$ for $t \in \bfT_i \cup \bfD_i$ and~$i \in [n]$ where~$Y_i(t) = 1[t \in \bfT_i]$, offset~$\log \pi_i (t)$, and covariates~$\tilde W_{i,t}$.
-\end{lemma}
-
-The connection established by Lemma~\ref{lemma:logistic} between our proposed methodology and logistic regression allows us to leverage ``off-the-shelf'' software.
-The main complication is pre-processing of the functional data; however, these additional steps can also be taken care of via existing software.
-Therefore, the entire data analytic pipeline is easy-to-implement and requires minimal additional effort by the end-user.
-To see this, we briefly review the proposed inference procedure.
-
-\begin{rmk}[Inference procedure review] \normalfont
-Given observed recurrent event and high frequency data~$\{ \bfT_i, \bfx_i \}_{i=1}^n$,
-\begin{enumerate}[(a)]
-\item For each $i \in [n]$, sample non-event times as a time-inhomogeneous Poisson point process with intensity according to~$\pi_i (t)$
-\item \label{p2} Estimate mean~$\mu_y (t,s)$ for $0 \leq s \leq \Delta$ at all event times~$t \in \cup_{i=1}^n \bfT_i$ and sampled non-event times~$t \in \cup_{i=1}^n \bfD_i$.
-\item \label{p3} Compute marginal covariance across event times,~$\Sigma_1$, and non-event times,~$\Sigma_0$.
-\item \label{p4} Compute eigendecomposition~$\{ \hat \psi_k^{(y)}, \hat \lambda_k^{()} \}$ of marginal covariance~$\Sigma_y$
-\item Use the eigendecomposition to construct~$W_{i,t}$ for all $i \in [n]$ and $t \in \bfD_i \cup \bfT_i$
-\item \label{point:log} Perform logistic regression with binary outcome~$\{ \Y_i (t) \}$  and offset of~$\log \pi_i (t)$.
-\end{enumerate}
-\end{rmk}
-
-\noindent Before demonstrating the methodology via simulation in Section~\ref{section:simstudy} and a worked example in Section~\ref{section:example}, we provide a theoretical analysis of our current proposal.
-
-\subsection{Theoretical analysis}
-
-Our theoretical analysis requires assumptions regarding the subsampling procedure, the event process, and the functional data.
-We state these assumptions and then our main theorems.
-We start by assuming there exists a~$\tau < \infty$ such that all individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
-Moreover, define $R_i (t)$ to be the at-risk indicator for participant~$i$, i.e., $R_i (t) = {\bf 1} [t \in (0,\tau_i)]$.
-Asymptotic theory provided in Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp.
-420--421]{Andersen1993} along with the following additional assumptions:
-
-\begin{assumption}[Event process assumptions]
-\label{assumption:events}\normalfont
-We assume the following holds:
-\begin{enumerate}[label=(E.\arabic*)]
-\item\label{E1} The subsampling rate is both lower and upper bounded at all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$ such that $R_i (t) = 1$
-\item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
-  \[
-    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t; \theta) \times \left[ \frac{h_i^{(1)}(t; \theta) h_i^{(1)} (t;\theta)^\top}{h_i (t; \theta)} \right] \times   R_i (t) dt \overset{P}{\to} \Xi (\theta).
-  \]
-\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for all~$(i,j,t)$.
-\item\label{E4} For all~$j,k$
-\[
-n^{-1} \sum_{i=1}^n \int_0^\tau \left | \frac{d^2}{d\theta_j
-    d\theta_k} h_i (t;\theta_0) \right|^2 R_i (t) dt \overset{P}{\to}
-C < \infty
-\]
-as $n \to \infty$.
-\end{enumerate}
-\end{assumption}
-
-We also require several assumptions due to the truncation of the Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
-\begin{assumption}[Functional assumptions~\citep{Park2018}] \normalfont
-\label{assumption:truncation}
-The following assumptions are standard in prior work on longitudinal functional data analysis~\citep{Park2018, Yao2005, ChenMuller2012}:
-\begin{enumerate}[label=(A.\arabic*)]
-\item\label{A1} $X = \{ X(t, s) : (t,s) \in \mathcal{T} \times
-  \mathcal{S} \}$ is a square integrable element of the $L^2 (
-  \mathcal{T} \times \mathcal{S})$
-\item\label{A2} The subsampling and conditional intensity rate functions~$f_y(T)$ are continuous and~$\sup |f_y(T)| < \infty$.
-\item\label{A3} $\E[X(t,s) X(t,s^\prime) X(t^\prime,s) X(t^\prime, s^\prime) ] < \infty$ for each $s,s^\prime \in [0,\Delta]$ and~$0 < t, t^\prime < \tau$.
-\item\label{A4} $\E[\|X(t,\cdot)\|^4] < \infty$ for each~$0< t < \tau$.
-\end{enumerate}
-\end{assumption}
-Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
-With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
-
-\begin{lemma} \normalfont
-\label{lemma:simpleasym}
-Under Assumption~\ref{assumption:events}, Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
-\[
-\sqrt{n} (\hat \theta - \theta) \to N(0, \Xi (\theta)^{-1})
-\]
-where
-\[
-  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \times \left[ \frac{h^{(1)}(s;
-      \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
-  ds.
-\]
-and~$\tau$ is the random censoring time of the event process.
-\end{lemma}
-A design-unbiased estimator for~$\Xi(\theta)$ is
-\begin{equation}
-\label{eq:fisher}
-  \hat \Xi (\theta) = n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i}
-  w_i(t;\theta) (1-w_i(t;\theta)) \left[ \frac{h_i^{(1)}(t;
-      \theta)}{h_i (t; \theta)} \right] \times  \left [
-    \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top
-\end{equation}
-For the log-linear intensity model, the sampling-unbiased estimator for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the previously described logistic regression model.
-This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
-Not only this, Lemma~\ref{prop:optimal} shows weights~\eqref{eq:waage_weights} are optimal within a particular class of weighted estimating equations.
-
-\begin{lemma} \normalfont
-\label{prop:optimal}
-% \textcolor{red}{
-If the event process is an inhomogeneous Poisson point process with intensity~$h(t; \theta)$ and subsampling occurs via an independent, inhomogeneous Poisson point process with intensity~$\pi (t)$, then~$\hat U_n (\theta)$ are optimal estimating functions (i.e., most efficient) in the class of weighted estimating functions given by~\eqref{eq:approxscore} replacing~\eqref{eq:waage_weights} by any weight function~$w_i (t; \theta)$.
-This class includes the Horvitz-Thompson estimator under~$w(s; \theta) = 1$.
-% WD: Consider larger class and show the proof doesn't extend to non PP models?}
-\end{lemma}
-
-\noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
-
-\subsection{Computation versus statistical efficiency tradeoff}
-
-Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
-While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
-It is this trade-off that we next make precise.
-In particular, we consider the following choice of subsampling rate,~$\pi(t) = c \times h(t; \theta)$ for $c>0$.
-That is, the subsampling rate is proportional to the intensity function with time-independent constant~$c > 0$.
-Under this subsampling rate, the weight function~\eqref{eq:waage_weights} is equal to $c/ (c+1)$.
-Under Lemma~\ref{lemma:simpleasym},
-\[
-\Xi (\theta) = \frac{c}{c+1} \int_0^\tau \frac{ h^{(1)} (t; \theta)
-  h^{(1)} (t; \theta)^\top}{h (t; \theta)} dt = \frac{c}{c+1} \Sigma (\theta)
-\]
-where~$\Sigma(\theta)$ is the Fisher information of the complete-data maximum likelihood estimator.
-Therefore the relative efficiency is~$c/(1+c)$.
-For an upper bound~$H = \max_{t \in (0,\tau)} h(t;\theta)$, if we set~$\pi (t) = c \times H$, then the relative efficiency can be lower bounded by~$c / (c+1)$.
-
-Sensor measurements occur multiple times per second.
-Suppose the intensity rate is bounded above by $1$ and the unit time scale was hours.
-If we then  subsample the data at a rate of~$10$ times per hour, then we have a lower bound on the efficiency of $0.909$.
-For a 4Hz sensor, this reduces the number of samples per hour from~$4 \times 60 \times 60 = 14,400$ per hour to on average $10$ per hour.
-While the computational complexity of logistic regression is linear in the number of samples, we get $1440$ times reduction in the data size at the cost of a $0.909$ statistical efficiency.
-If we sample~$100$ times per hour, then the efficiency loss is only $0.999$, with a $144$ times reduction in data size.
-Table~\ref{tab:compvseff} provides additional examples for a 4Hz and 32Hz sensor rate respectively.
-The data reduction depends on this rate; however, the lower bound on statistical efficiency does not because the subsampling rate only depends on the upper bound of the intensity function.
-In particular, if the events are rare then subsampling rate can be greatly reduced with no impact to statistical efficiency.
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{l l r r r r r | c}
-\multirow{2}{2.5cm}{Sensor rate} &
-\multirow{2}{2.5cm}{Subsampling constant ($c$)}
-  & \multicolumn{5}{c}{Upper bound on intensity rate per
-    hour}
-  & \multirow{2}{2cm}{Statistical efficiency}\\ \cline{3-7}
-& & 0.5 & 1 & 3 & 5 & 10 \\ \hline
-% 1 & 28800.0 & 14400.0 & 4800.0 & 2880.0 & 1440.0 & 0.500 \\
-\multirow{3}{*}{4Hz (EDA)}
-& 5 & 5760 & 2880 & 960 & 576 & 288 & 0.833 \\
-& 10 & 2880 & 1440 & 480 & 288 & 144 & 0.909 \\
-& 100 & 288 & 144 & 48 & 29 & 14 & 0.990 \\ \hline
-\multirow{3}{*}{32Hz (ACC)}
-& 5 & 46080 & 23040 & 7680 & 4608 & 2304 & 0.833 \\
-& 10 & 23040 & 11520 & 3840 & 2304 & 1152 & 0.909 \\
-& 100 & 2304 & 1152 & 384 & 230 & 115 & 0.990 \\ \hline
-% 500 & 57.6 & 28.8 & 9.6 & 5.8 & 2.9 & 0.999 \\ \hline
-\end{tabular}
-\caption{Data reduction (total \# of measurements divided by expected number of subsampled measurements) given sensor rate, subsampling constant and an upper bound on the intensity rate.}
-\label{tab:compvseff}
-\end{table}
-
-\subsection{Penalized functional regression models}
-
-Recall theoretical results were proven under the assumption that there exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
-To make this assumption plausible, we set~$K_b$ large enough (but less than~$K_x$ to ensure identifiability) to ensure the spline basis expansion is sufficiently expressive.
-However, in practice, such a choice of~$K_b$ may lead to overfitting the data.
-Following~\cite{Goldsmith2011}, we choose the polynomial spline model and set~$\phi(t) = b_0 + b_1 t + b_2 t^2 + \sum_{j=3}^{K_b} (t-\kappa_j)^3$ where~$\{ \kappa_j \}_{j=3}^{K_b}$ are the chosen knots and assume~$\{ b_j \}_{j=3}^{K_b} \sim N(0, \sigma^2 I)$ to induce smoothness on the spline model.
-Combining the penalized spline formulation with Lemma~\ref{lemma:logistic} establishes a connection between our approximate score equations and solving a generalized mixed effects logistic regression with offset.
-
-\subsection{Confidence intervals for~$\beta(t)$}
-
-Due to the connection with generalized mixed effects models, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
-That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ dimensional matrix obtained by plugging in the estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-
-We acknowledge two important limitations of confidence intervals obtained via this approach.
-First, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is oversmoothed.
-Second, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
-Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
-
-\section{Simulation study} \label{section:simstudy}
-
-We next assess the proposed methodology via a simulation study.
-Here, we assume each individual is observed over five days where each day is defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day.
-We define~$X(t)$ at the grid of observations as a mean-zero Gaussian process with covariance
-\[
-  \Sigma (t, t^\prime) =
-  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
-  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
-  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
-\]
-where~$K_v$ is the modified Bessel function of the second kind.
-We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.3$ as well as set~$K_b = K_x = 35$.
-For simplicity, we assume~$\Sigma$ is known in computation of the eigendecompositions.
-Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; \theta)$.
-To mimic our real data, we set
-\[
-h(t; \theta) = \exp \left( \theta_0 + \int_{0}^{\Delta} X(t-s) \beta(s) ds \right).
-\]
-We set~$\Delta$ to mimic a 30-minute window for a $12$-hour day.
-We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day.
-We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$.
-
-We generate $1000$ datasets, each consisting of $500$ user-days.
-For a given simulated user-day, we randomly sample non-event times using a Poisson process with rate of a halfhour.
-We use the proposed methodology to construct the estimate $\hat \beta_{i,0.5} (t)$ for the $i$th simulated dataset; we then subsample the sampled non-event times with thinning probability $1/2$, $1/4$ and $1/8$.
-This results in randomly sampled non-event times given by a Poisson process with rates of an hour, two hours, and four hours.
-We can construct the corresponding estimates: $\hat \beta_{i,1} (t)$, $\hat \beta_{i,2} (t)$, and $\hat \beta_{i,4} (t)$ respectively.
-Subsampling allows us to compare the variance due to subsampling as compared to the variance due to sampling fewer non-event times.
-
-Since we are primarily interested in accuracy, we report the mean integrated squared error (MISE) defined as~$\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \beta(s))^2 ds$ for $j=0.5,1,2,4$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
-The MISE is defined in this manner to account for settings where~$\Delta$ is unknown.
-Next, let $\bar \beta_{j} (t) = \frac{1}{1000} \sum_{i=1}^{1000} \hat \beta_{i,j} (t)$ denote the average estimate for $j=0.5,1,2,4$.
-Then the squared bias is given by $\int_{0}^{\infty} ( \bar \beta_{j} (s) - \beta (s))^2 ds$ and the variance is given by $\frac{1}{1000} \sum_{j=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \bar \beta_j(s))^2 ds$.
-The subsampling variance is defined as $\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \hat \beta_{i,0.5} (s))^2 ds$.
-Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biased as well as the subsampling variance.
-To allow fair comparisons across the two choices of $\beta(s)$, all reported numbers are scaled by the integrated square of the true function $\int_{0}^{\infty} \beta(s)^2 ds$.
-The average runtime (in seconds) is also reported.
-
-Table~\ref{tab:mise} demonstrates that the variance does increase as the sampling rate decreases.
-However, the rate of increase in the MISE is relatively quite low.
-In the second case study, for example, the MISE increases by 6\% while the run time is 3 times faster.
-In the first case study, the MISE remains roughly constant but the run time is 4.7 times faster.
-This highlights the efficiency-computation trade-off.
-Extrapolating to a complete data analysis, we can see that the run time will continue to increase significantly with minimal improvement in the mean integrated squared error.
-
-\begin{table}[!th]
-\begin{tabular}{c | p{2cm} p{2cm} p{2cm} p{2cm}}
-& \multicolumn{4}{c}{Sampling rate} \\ \cline{2-5}
- & 0.5 & 1 & 2 & 4 \\ \hline
-Subsampling variance & - & $9.0 \times 10^{-7}$ & $2.9 \times 10^{-6}$ & $5.8 \times 10^{-6}$  \\
-Variance & $7.9 \times 10^{-3}$ & $9.2 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ \\
-Squared Bias & $9.8 \times 10^{-2}$ & $9.5 \times 10^{-2}$ & $9.3 \times 10^{-2}$ & $9.2 \times 10^{-2}$   \\
-MISE & $1.1 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ \\ \hline
-Avg. runtime (secs) & 317 & 177 & 104 & 68 \\ \hline
-Subsampling variance & - & $2.9 \times 10^{-6}$ & $8.9 \times 10^{-6}$ & $2.0 \times 10^{-5}$  \\
-Variance & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.1 \times 10^{-2}$ \\
-Squared Bias & $6.9 \times 10^{-2}$ & $6.5 \times 10^{-2}$ & $6.0 \times 10^{-2}$ & $5.4 \times 10^{-2}$   \\
-MISE & $7.9 \times 10^{-2}$ & $7.9 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
-Avg. runtime (secs) & 36 & 23 & 15 & 12 \\ \hline
-\end{tabular}
-\caption{Mean-integrated squared error, varaince, squared bias, and subsampling variance for $\beta(s)$ given by (1) and (2) respectively.}
-\label{tab:mise}
-\end{table}
-
-\subsection{Impact of $\Delta$}
-
-A concern with the proposed approach is the selection of window-length, $\Delta$.
-Here we investigate the impact of misspecification of the window length for $\beta(t) = \beta_1 \cdot \sin \left( 2 \pi s / \Delta - \pi /2 \right)$ and the true window length ($\Delta^\star$) is set to 32-minutes.
-See Appendix~\ref{app:delta} for a similar discussion for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
-As in the previous simulation, we generate 1000 datasets per condition each consisting of 500 user-days.
-For each simulation, we analyze the data using window lengths $\Delta \in (26,29,32,35,37)$.
-
-When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$; however, we incur a penalty in finite samples, especially for settings where the function is far from zero near $t = \Delta$.
-We find the MISE increases the absolute error $|\Delta - \Delta^\star|$ increases, while the variance decreases as a function of $\Delta$.
-While the MISE increased for $\Delta < \Delta^\star$, the pointwise estimation error remains low for $t < \Delta$.
-This does not hold for $\Delta > \Delta^\star$, where instead we see parameter attenuation, i.e., a bias towards zero in the estimates at each $0 < t< \Delta$.
-To capture this, we define a partial MISE as $\frac{\Delta}{\tilde \Delta} \int_0^{\tilde \Delta} ( \hat \beta_{i,j} (s) - \hat \beta (s) )^2 ds$ where $\tilde \Delta = \min(\Delta, \Delta^\star)$, which is the MISE on the subset $0<t< \min (\Delta, \Delta^\star)$ and scaled for comparative purposes.
-
-\begin{table}[!th]
-\begin{tabular}{c | c c c c c}
-& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
-$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
-\multirow{3}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
- & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\
-  & P-MISE & 0.225 & 0.229 & 0.233 & 0.239 \\ \hline
-\multirow{3}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
- & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\
-  & P-MISE & 0.107 & 0.109 & 0.111 & 0.117 \\ \hline
-\multirow{3}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
- & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\
-  & P-MISE & 0.105 & 0.103 & 0.103 & 0.105 \\\hline
-\multirow{3}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
- & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\
-  & P-MISE & 0.273 & 0.270 & 0.269, & 0.269 \\\hline
-\multirow{3}{*}{37} & MISE & 0.589 & 0.588 & 0.589 & 0.591 \\
- & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\
-  & P-MISE & 0.473 & 0.470 & 0.469 & 0.470 \\\hline
-\end{tabular}
-\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
-\label{tab:mise}
-\end{table}
-
-\section{Extensions}
-
-In this section, we demonstrate the flexibility of the proposed approach by exploring extensions in several important directions to ensure these methods are robust for practical use with high frequency data.
-This section will continue to leverage the connection to generalized functional linear models provided by Lemma~\ref{lemma:logistic}.
-
-\subsection{Multivariate extensions}
-\label{section:multiplesensors}
-
-In this section, we extend our model to the case of multiple functional regressors.
-That is, suppose $L$ health processes, i.e., $\bfx_i = \{ \bfx_i (t) = (x_{i,1} (t), \ldots, x_{i,L} (t)) \}_{0 < s < \tau_i}$, for each participant is measured at a dense grid of time points.
-In the suicidal ideation case study, for example, accelerometer is measured at a rate of 32Hz while electrodermal activity (EDA) is measured at a rate of 4Hz.
-A multivariate extension of our model~\eqref{eq:hazard} is given by
-\begin{equation}
-\label{eq:multihazardlinear}
-h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) =
-h_0 (t; \gamma) \exp \left( g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + \int_{t-\Delta_1}^t x_{i,1} (s) \beta_1(s) ds + \cdots + \int_{t-\Delta_L}^t x_{i,L} (s) \beta_L(s) ds  \right)
-\end{equation}
-The approach given in Section~\ref{eq:logistication} extends naturally to the multivariate functional setting.
-For each functional regressor, we estimate the pooled sample covariance $\Sigma_{y,l}$ for $y \in \{0,1\}$ and $l=1,\ldots,L$ as in Section~\ref{sec:margcov}.
-Let $\sum_{k=1}^\infty \hat \lambda^{(y)}_{k,l} \hat \psi^{(y)}_{k,l} (s) \hat \psi_{k,l}(t)$ be the spectral decomposition of $\hat \Sigma_{y,l}$.
-Then $x_{i,j}(t)$ is approximated using a truncated Karhunen-Loeve $\int_{t-\Delta_l}^t X_{l} (t,s) \beta_l(t) = \left[M_{l,t} + {\bf c}_l (t)^\top J_{\hat \psi_l^{(y)}, \phi_l} \right] {\bf b}_l$.
-% Note that we express each coefficient function in the same spline basis; however, different basis expansions are used for each health process as defined by the marginal covariance calculation in Section~\ref{sec:margcov}.
-
-\subsection{Missing data}
-\label{section:missingdata}
-% \walt{Question: How to perform multiple imputation for~$X(\cdot,t)$
-%   and $X(\cdot, t^\prime)$ for~$|t - t^\prime| <\Delta$? That is, how
-%   do you perform multiple imputation with this overlapping structure?}
-
-Sensor data can often be missing for intervals of time due to sensor wearing issues.
-In the suicidal ideation case study, for example, there are 2139 self-identified moments of distress across all 91 participants.
-Of these, 1289 event times had complete data for the prior thirty minutes, 1984 had fraction of missing data on a fine grid less than 30\%, and 1998 had fraction of missing data on a fine grid less than 10\%.
-
-Missing data is a critical issue because~$c_{i,k} (t)$ cannot be estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$.
-Moreover, standard errors should reflect the uncertainty in these coefficients when missing data is prevalent.
-\cite{Goldsmith2011} suggest using best linear unbiased predictors (BLUP) or posterior modes in the mixed effects model to estimate~$c_{i,k} (t)$; however, this is ineffective when there is substantial variability in these estimates.
-To deal with this, \cite{Crainiceanu2010} take a full Bayesian analysis.
-\cite{Yao2005} introduced PACE as an alternative frequentist method.
-\cite{Petrovich2018} shows that for sparse, irregular longitudinal, the imputation model should not ignore the outcome variable~$Y_i (t)$.
-
-Here we present an extension of~\cite{Petrovich2018} to our setting by leveraging Lemma~\ref{lemma:logistic} and the marginal covariance estimation procedure to construct a multiple imputation procedure.
-Let~$\bfx_{i} (t)$ denote incomplete sensor data at time~$t$ (i.e., at times~$\{ s_{i,r} \}_{r=1}^{k_{it}}$ in~$[0,\Delta]$.
-Then
-\begin{align}
-\label{eq:impute}
-\E [ X_i (s,t) \given Y_i(t) = y, \bfx_i (t) ]
-  &= \mu_y (s,t) + {\bf a}_{i,t}^\top (s) {\bf B}_{i,t} (\bfx_i (t) -
-    \mathbold{\mu}_i (t) ) \\
-\text{Cov} \left( X_i (s,t), X_i (s^\prime, t) \given Y_i (t) = y,
-  \bfx_i (t) \right)
-  &= \Sigma_{t} (s, s^\prime) -
-    {\bf a}_{i,t} (s)^\top {\bf B}_{i,t} {\bf a}_{i,t} (s^\prime)
-\end{align}
-where we have
-\[
-{\bf a}_{i,t} (s)^\top = \left( \begin{array}{c} \Sigma_t (s_{i,1}, s) \\ \vdots \\
-                       \Sigma_t (s_{i,k_{it}}, s) \end{array} \right); \quad
-{\bf B}_{i,t}^{-1} = \left(
-  \begin{array}{c c c}
-    \Sigma_t(s_{i,1}, s_{i,1}) & \Sigma_t(s_{i,1}, s_{i,2}) & \cdots \\
-    \Sigma_t(s_{i,1}, s_{i,2}) & \ddots & \vdots \\
-    \vdots & \cdots & \Sigma_t (s_{i,k_{it}}, s_{i,k_{it}})
-  \end{array} \right),
-\]
-$\mathbold{\mu}_i (t) = \E [ \bfx_i (t) \given Y_i (t) = y] = \{
-\mu_y (s_{j}, t) \}_{j=1}^r$ and~$\mu_y (s,t)$ is the mean of~$X(t,s)$
-from group~$y$, and~$\Sigma_t ( s^\prime, s)$ is the covariance
-between~$X(s^\prime,t)$ and $X(s,t)$ for $s^\prime, s \in [0, \Delta]$
-and~$t \in \mathbb{R}_+$. 
-%To account for overlap between $\Delta$-windows for $t \in \bfT_i \cup \bfD_i$ - i.e., avoid imputing values at a particular time to different values -- imputation is performed sequentially over this set.
-%$\{ (t,t') : t, t^\prime \in \bfT_i \cup \bfD_i \& |t - t'| < \Delta \}$
-
-\subsubsection{Multiple imputation under uncongeniality}
-\label{section:miuncong}
-
-Multiple imputation yields valid frequentist inferences when the imputation and analysis procedure are congenial~\citep{Meng1994};  the above procedure is derived for function-on-scalar multiple imputation for binary outcomes, which ignored the joint nature of recurrent event analysis in the presence of high frequency sensor data.
-The main advantage of the above imputation framework is its simplicity and approximate congeniality when events are rare and the sampling rate is low.
-The main disadvantage is that the above framework is uncongenial under many events and/or high sampling rates.
-Again, congeniality ensures good frequentist coverage properties.
-A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
-
-To address this, we use the recommendation from~\cite{Bartlett2020} and consider a method that first bootstraps a sample from the dataset and then apply multiple imputation to each bootstrap sample.
-This general approach was originally proposed by~\cite{Shao1994} and~\cite{Little2002}.
-We suppose $B$ bootstraps and $M$ imputations per bootstrap; let $\hat \theta_{b,m}$ denote the estimator for the $m$th imputation of the $b$th bootstrap.
-The point estimator is given by $(B)^{-1} \sum_{b=1}^B \hat \theta_{b}$ where $\hat \theta_b = M^{-1} \sum_{m=1}^M \hat \theta_{b,m}$.
-To construct the confidence interval, we require mean sum of squares with and between boostraps, i.e., $MSW = \frac{1}{B(M-1)} \sum_{b=1}^B \sum_{m=1}^M (\hat \theta_{b,m} - \hat \theta_b ) (\hat \theta_{b,m} - \hat \theta_b )^\top$ and $MSB = \frac{1}{B-1} \sum_{b=1}^B (\hat \theta_b - \hat \theta)(\hat \theta_b - \hat \theta)^\top$ respectively.
-Then the estimator of the variance-covariance matrix of $\hat \theta$ is given by $\hat \Sigma_{B,M} = \left( \frac{B+1}{BM} \right) MSB - MSW/M$.
-We obtain the varaiance for~$\beta(t)$ by $\phi (t) \hat \Sigma_{B,M} \phi(t)^\top$.
-We follow~\cite{Bartlett2020} and construct confidence intervals based on Satterthwaite’s degrees of freedom, which here is given by
-$$
-\hat \nu = \frac{\left[ \left( \frac{B+1}{BM} \right) MSB - \frac{MSW}{M} \right]^2}{ (B-1)^{-1} \left( \frac{B+1}{BM} \right)^2 MSB^2 + \frac{MSW^2}{BM^2(M-1)}}
-$$
-The bootstrap followed by multiple imputation procedure has been studied extensively by~\cite{Bartlett2020} and is robust to uncongeniality.
-The main disadvantage of this approach is its considerable computational intensity.
-Recall likelihood calculations were computationally prohibitive by themselves, so combining with bootstrap and MI would further increase this large-scale computation.
-The random subsampling framework thus simplifies handling of missing data via connections to function-on-scalar multiple imputation by~\cite{Petrovich2018} as well as to bootstrap to handle uncongeniality by~\cite{Bartlett2020}.
-Ignoring the computational time of bootstrap sampling, the computational time for the first choice in the simulation study with $B=200$ bootstraps and $M=2$ imputations per bootstrap leads to $35$ hours for a sampling rate of $0.5$ compared to $7$ hours for a sampling rate of $4$, which highlights the benefits of the proposed framework.
-
-\subsection{Multilevel models}
-
-The approach can be extended to multilevel models with functional regressors, which are critical in mobile health where a high degree of individual variation is often observed.
-Let $b_i \sim N(0, \sigma_{b}^2)$, then the multilevel extension of~\eqref{eq:hazard} is
-\begin{equation}
-\label{eq:multilevel}
-h_i \left( t \given  H_{i,t}^{NX} ; \theta, b_i \right) \approx
-\exp \left( Z_{t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
-  + \left[ M_{i,t}^\top + C_{i,t}^\top J_{\hat \psi, \phi}\right] (\beta + b_i) \right)
-= \exp \left( W_{i,t}^\top \theta  + Z_{i,t}^\top b_i \right),
-\end{equation}
-where $Z_{i,t} = M_{i,t} + C_{i,t} J_{\hat \psi, \phi}$.
-Lemma~\ref{lemma:logistic} implies that the random subsampling framework applied to equation~\eqref{eq:multilevel} leads to a penalized logistic mixed-effects model.
-As far as the authors are aware, the combination of mixed-effects and $L_2$-penalization on a subset of parameters is not available in existing software.
-% An penalized adaptive Gaussian quadarture (PADQ) algorithm can be derived.  estimation is presented in Appendix~\ref{app:penlogit}.
-Given the paper's focus on ``off-the-shelf'' software implementations, multilevel models are considered important future work.
-
-\section{A worked example: Adolescent psychiatric inpatient mHealth study} \label{section:example}
-
-During an eight month period in 2018, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital were enrolled in an observational study.
-Study data were graciously provided by Evan Kleiman and his study team ({\tt https://kleimanlab.org}).
-Each study participant wore an Empatica E4~\citep{empaticae3}, a medical-grade wearable device that offers real-time physiological data.
-On each user-day, participants were asked to self-identify moments of suicidal distress.
-At these times, the participant was asked to press a button on the Empatica E4 device.
-The timestamp of the button press was recorded.
-One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress.
-In particular, the scientific question is whether there are early indicators of escalating distress by monitoring physiological correlates.
-
-A key concern is whether all moments of suicidal distress are recorded.
-To ensure this, clinical staff interviewed participants in the evening who were then asked to review their button press activity.
-Any events that were identified as incorrect button press activity were removed.
-At the end of the 30-day study period, the average number of button presses per day was $2.42$ with a standard deviation of $2.62$.
-Investigation of the button press data shows low button-press counts prior to 7AM and a sharp drop off by 11PM.
-This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
-To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.
-% Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
-
-\vspace{0.25cm}
-\begin{figure}[!th]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../figures/recurrentevents_plot.png}
-    \captionof{figure}{User button-presses (red) versus time since study entry (in hours).  The black mark indicates the final sensor measurement time.}
-    \label{fig:buttons}
-\end{figure}
-\vspace{0.25cm}
-
-Figure~\ref{fig:buttons} visualizes button presses versus time since study entry for each user.
-Day 30 is assumed to censor the observation process.
-A black mark signals dropout before day 30.
-Figure~\ref{fig:buttons} shows the potential heterogeneity in button press rates between users and study days.
-To assess whether there is between user or between study day variation, Table~\ref{tab:anova} presents a two-way ANOVA decomposition of the button press counts as a function of participant and day in study.
-The ANOVA decomposition demonstrates high variation with day in study and across users.
-
-\begin{table}[!th]
-\begin{tabular}{ccccccc}\hline
-      & DF & Sum Sq & Mean Sq & F value & Pr($>$F) \\ \hline
-        Participant & 88 & 2675.4 & 30.4 & 8.9 & $< 2 \times 10^{-16}$ \\
-        Day in Study & 29 & 443.3 & 15.3 & 4.5 & $3.9 \times 10^{-13}$ \\
-        Residuals & 672 & 2304.2 & 3.4 \\ \hline
-      \end{tabular}  \captionof{table}{ANOVA decomposition of daily button press counts}
-      \label{tab:anova}
-\end{table}
-
-Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{activity index}~\citep{10.1371/journal.pone.0160644}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
-Electrodermal activity can be significantly impacted by external factors (e.g., room temperature).
-To account for the high between user-day variation, we analyze EDA standardized per study-day and  device.
-The individual EDA and AI trajectories are highly variable which tends to obscure patterns and trends.
-In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and AI are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 10 minutes prior to button presses.
-
-\begin{figure}[!th]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/smoothed_eda.png}
-  \caption{Electrodermal activity (EDA)}
-  \label{fig:mean_eda}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/smoothed_ai.png}
-  \caption{Activity Index (AI)
-}  \label{fig:mean_acc}
-\end{subfigure}
-\caption{Average scaled EDA and AI in the 30 minutes prior to button presses}
-\label{fig:mean_edacc}
-\end{figure}
-
-\subsection{Complete-case analysis}
-
-Inspection of Figures~\ref{fig:mean_eda} and~\ref{fig:mean_acc} suggest setting $\Delta = 30$ minutes is adequate to capturing the proximal impact of EDA and AI on the risk of a button press.
-To ensure minimal loss of efficiency, the subsampling rate was set to once every fifteen minutes.
-Given the daily button press rate, this ensures an average of $44$ non-events to $2.5$ events per day.
-Based on Table~\ref{tab:compvseff}, this ensures we can achieve a substantial data reduction at a minimal loss of efficiency.
-After sampling non-event times, complete-case analyses are performed, i.e., sampled times where sensors included in the model have any level of missing data are ignored.
-
-We analyzed activity index (AI) and electrodermal activity separately as given in equation~~\eqref{eq:hazardlinear} as well as jointly in a multivariate model as in equation~\eqref{eq:multihazardlinear}.
-Results did not change substantially.
-Figure~\ref{fig:eda_estimate} and~\ref{fig:acc_estimate} presents the estimates from the joint analysis with their associated 95\% confidence intervals.
-We highlight in gray the statistically significant regions.
-Here, we see that standardized EDA is not associated with increased risk of button press, while activity index sees a positive association in the final few minutes prior to a button press.
-In section~\ref{sec:missingdata_heterogeneity}, we investigate whether the results are sensitive to missing data.
- % and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
-
-\begin{figure}[!th]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint.png}
-  \caption{Electrodermal activity (EDA)}
-  \label{fig:eda_estimate}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint.png}
-  \caption{Activity Index (AI)
-}  \label{fig:acc_estimate}
-\end{subfigure}
-\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the complete-case analysis, while highlighted region are the model-based pointwise 95\% confidence intervals.}
-\label{fig:mean_edacc}
-\end{figure}
-
-\subsection{Accounting for missing data}
-\label{sec:missingdata_heterogeneity}
-
-In Appendix~\ref{app:csmissing}, Figures~\ref{fig:missing_eda} and~\ref{fig:missing_acc} present the per-patient average fraction with missing EDA and AI in 30-minute windows respectively.
-For EDA, we have a wide range of missingness -- from $0\%$ to over $40$\% across individuals with most between $5\%$ and $30\%$ average fraction missing.
-For AI, the missingness is less pronounced -- from $0\%$ to $15\%$ across individuals with most between $0\%$ and $10\%$ average fraction missing.
-To assess the impact of missing data, we perform the imputation under uncongeniality using the boostrap and multiple imputation strategy described in Section~\ref{section:miuncong}.
-Here we see that the EDA curve remains non-significant while the region where the AI curve is significant widens slightly to be the prior 4 minutes.
-
-\begin{figure}[!th]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint_bootstrap.png}
-  \caption{Electrodermal activity (EDA)}
-  \label{fig:eda_estimate_bs}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint_bootstrap.png}
-  \caption{Activity Index (AI)
-}  \label{fig:acc_estimate_bs}
-\end{subfigure}
-\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the missing data analysis, while highlighted region are the boostrap and MI based pointwise 95\% confidence intervals.}
-\label{fig:mean_edacc}
-\end{figure}
-
-% \subsection{Effect of electrodermal activity and activity index on moments of distress}
-
-% Over a period of 5~years and one month following recruitment,
-% patient~$u$ had eight appointments with prothrombin values as follows:
-
-\section{Discussion}
-In this paper, we have presented a methodology for translating a difficult functional analysis with recurrent events problem into a traditional logistic regression.
-The translation leveraged subsampling and weighting techniques, specifically the use of weights suggested by~\cite{Waagepetersen2008}, along with flexible functional data analysis methods of~\cite{Goldsmith2011} with marginal covariance methods for longitudinal functional data of~\cite{Park2018}.
-The proposed methodology abides by the idea that we should \emph{make data as small as possible as fast as possible}.
-Subsampling and  weighting converts the problem to well-known territory which allows us to leverage existing software.
-We show limited loss of efficiency when the subsampling is properly tuned to the event rates.
-Important extensions to an online sampling algorithm, optimal weighting when the Poisson point process assumption does not hold, and non-linear functional data methods are all considered important future work.
-
-\bibliographystyle{plainnat}
-\bibliography{si-fda-refs}
-
-\newpage
 \appendix
@@
 \section{Derivation of the design-unbiased score equations}
+\label{app:prooflemma31}
 
@@
 \section{Optimal weights}
+\label{app:optimalweights}
 
@@
 \left( \begin{array}{c c} \Sigma(\theta_0; w) & \Xi (\theta_0; w^\star)^{-1} \\ \Xi(\theta_0; w^\star)^{-1} & \Xi (\theta_0; w^\star)^{-1} \end{array} \right)
-\left( \begin{array}{c} \alpha & - \alpha \end{array} \right)\\
+\left( \begin{array}{c} \alpha \\ - \alpha \end{array} \right)\\
 &= \alpha^\top \left( \Sigma(\theta_0; w) - \Xi (\theta_0; w^\star)^{-1} \right) \alpha.
@@
 \section{Additional simulations and case study details}
+\label{app:extrasims}
 
@@
 
-\section{Summary of missing data for the case study}
+\subsection{Summary of missing data for the case study}
 \label{app:csmissing}
@@
 
+\newpage
+\bibliographystyle{plainnat}
+\bibliography{si-fda-refs}
+
 \end{document}
-
-1. Start simulation for appendix (delta wrong, beta2): CHECK
-2. Button Press Figure: Check
-3. Compute wear minutes:
-4. Standardize EDA:
-5. Subsample EDA and gACC only at wear times between 9am and 8pm
-6. EDA/ACC figures
-7. Estimate with complete data
-8. Estimate with bootMI; make 1 plot with CIs for both compl and bootMI
-9. Estimate with heterogeneity
-10. Present

### 2023-02-02 write-up/fda-recurrent.tex
updating paper first, then code for revision

@@
 \doublespacing
+
+\usepackage{xr}
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}% latexmk will find this if $recorder=0 (however, in that case, it will ignore #1 if it is a .aux or .pdf file etc and it exists!
+if it doesn't exist, it will appear in the list of dependents regardless)
+  \@addtofilelist{#1}% if you want it to appear in \listfiles, not really necessary and latexmk doesn't use this
+  \IfFileExists{#1}{}{\typeout{No file #1.}}% latexmk will find this message if #1 doesn't exist (yet)
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+%%% END HELPER CODE
+
+% put all the external documents here!
+\myexternaldocument{fda-recurrent-supplement}
 
@@
 
+\subsection{Related work}
+
 \section{Recurrent event process and associated high frequency data}
@@
 Here we investigate the impact of misspecification of the window length for $\beta(t) = \beta_1 \cdot \sin \left( 2 \pi s / \Delta - \pi /2 \right)$ and the true window length ($\Delta^\star$) is set to 32-minutes.
-See Appendix~\ref{app:delta} for a similar discussion for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
+See Section~\ref{app:delta} of the supplementary materials for a similar discussion for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
 As in the previous simulation, we generate 1000 datasets per condition each consisting of 500 user-days.
@@
 
+\subsection{Comparison to alternative methods} \label{section:simcomparisons}
+
 \section{Extensions}
@@
 
-In Appendix~\ref{app:csmissing}, Figures~\ref{fig:missing_eda} and~\ref{fig:missing_acc} present the per-patient average fraction with missing EDA and AI in 30-minute windows respectively.
+In Section~\ref{app:csmissing} of the supplementary materials, Figures~\ref{fig:missing_eda} and~\ref{fig:missing_acc} present the per-patient average fraction with missing EDA and AI in 30-minute windows respectively.
 For EDA, we have a wide range of missingness -- from $0\%$ to over $40$\% across individuals with most between $5\%$ and $30\%$ average fraction missing.
@@
 
-\newpage
-\appendix
-
-\section{Derivation of the design-unbiased score equations}
-
-\begin{proof}[Proof of Lemma~\ref{lemma:logistic}]
-Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t; \theta)}{h_i (t; \theta)}$.
+% \newpage
+% \appendix
+
+% \section{Derivation of the design-unbiased score equations}
+
+% \begin{proof}[Proof of Lemma~\ref{lemma:logistic}]
+% Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t; \theta)}{h_i (t; \theta)}$.
 Under the log-linear intensity function given by~\eqref{eq:approx_hazard}, $\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}$ and
-\[
-\frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) =
-\frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp
-  \left( W_{i,t}^\top \theta \right)} W_{i,t}
-\]
-Therefore,
-\begin{align*}
-\tilde U_n (\theta)
-  &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} W_{i,t} - \sum_{t \in \bfT_i \cup \bfD_i} \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}  W_{i,t} \right \} \\
-  &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} \frac{\pi_i (t)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} - \sum_{t \in \bfD_i} \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} \right \} \\
-  &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} w_{i} (t; \theta) W_{i,t} -
-    \sum_{t \in D_i} (1 - w_i (t; \theta)) W_{i,t} \right \} \\
-  &= - \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \left[ {\bf 1} [t \in \bfD_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
-  &= \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
-    \left[ {\bf 1} [t \in \bfD_i]  - \frac{1}{1 + \exp\left( - (
-          \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
-    \right] \tilde W_{i,t}.
-\end{align*}
-where~$\tilde W_{i,t} = - W_{i,t}$. This is exactly the score equation
-for logistic regression with offset~$\log \pi_i (t)$.
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
-% Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
-Define the joint counting process
-\[
-M_i (t) = N_i (t) - \int_0^t (h_i (s; \theta) + \pi_i (s)) R_i (s) ds.
-\]
-where~$N_i (t)$ is the counting process with jumps at all~$t \in \bfT_i \cup \bfD_i$. Asymptotic consistency is guaranteed~\cite[Theorem VI.1.1]{Andersen1993} if
-\begin{equation}
-\label{eq:andersen1}
-n^{-1} \tilde U_n (\theta_0 ) \overset{P}{\to} 0,
-\end{equation}
-and
-\begin{equation}
-\label{eq:andersen2}
-n^{-1} \frac{d}{d \theta^{\top}} \tilde U_n (\theta) \big |_{\theta =
-  \theta_0} = \Xi (\theta),
-\end{equation}
-and
-\begin{equation}
-\label{eq:andersen3}
-\lim_{n \to \infty} P \left( \left | n^{-1} \frac{d^2}{d \theta_j d
-      \theta_k} \tilde U_n (\theta) \right | < M \text{ for all }
-  j,k \text{ and all } \theta \in \Theta_0 \right) = 1
-\end{equation}
-To prove~\eqref{eq:andersen1}, we decompose~$\tilde U_n (\theta_0)$ into
-three terms
-\begin{equation}
-\label{eq:decomp}
-n^{-1} \tilde U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
-  \hat U_n (\theta_0) - U_n (\theta) \right) + n^{-1} \left(\tilde U_n
-  (\theta_0) - \hat U_n (\theta_0) \right)
-\end{equation}
-where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t,s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
+% \[
+% \frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) =
+% \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp
+%   \left( W_{i,t}^\top \theta \right)} W_{i,t}
+% \]
+% Therefore,
+% \begin{align*}
+% \tilde U_n (\theta)
+%   &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} W_{i,t} - \sum_{t \in \bfT_i \cup \bfD_i} \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}  W_{i,t} \right \} \\
+%   &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} \frac{\pi_i (t)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} - \sum_{t \in \bfD_i} \frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} \right \} \\
+%   &= \sum_{i=1}^n \left \{ \sum_{t \in \bfT_i} w_{i} (t; \theta) W_{i,t} -
+%     \sum_{t \in D_i} (1 - w_i (t; \theta)) W_{i,t} \right \} \\
+%   &= - \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \left[ {\bf 1} [t \in \bfD_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
+%   &= \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
+%     \left[ {\bf 1} [t \in \bfD_i]  - \frac{1}{1 + \exp\left( - (
+%           \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
+%     \right] \tilde W_{i,t}.
+% \end{align*}
+% where~$\tilde W_{i,t} = - W_{i,t}$. This is exactly the score equation
+% for logistic regression with offset~$\log \pi_i (t)$.
+% \end{proof}
+
+% \begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
+% % Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
+% Define the joint counting process
+% \[
+% M_i (t) = N_i (t) - \int_0^t (h_i (s; \theta) + \pi_i (s)) R_i (s) ds.
+% \]
+% where~$N_i (t)$ is the counting process with jumps at all~$t \in \bfT_i \cup \bfD_i$. Asymptotic consistency is guaranteed~\cite[Theorem VI.1.1]{Andersen1993} if
+% \begin{equation}
+% \label{eq:andersen1}
+% n^{-1} \tilde U_n (\theta_0 ) \overset{P}{\to} 0,
+% \end{equation}
+% and
+% \begin{equation}
+% \label{eq:andersen2}
+% n^{-1} \frac{d}{d \theta^{\top}} \tilde U_n (\theta) \big |_{\theta =
+%   \theta_0} = \Xi (\theta),
+% \end{equation}
+% and
+% \begin{equation}
+% \label{eq:andersen3}
+% \lim_{n \to \infty} P \left( \left | n^{-1} \frac{d^2}{d \theta_j d
+%       \theta_k} \tilde U_n (\theta) \right | < M \text{ for all }
+%   j,k \text{ and all } \theta \in \Theta_0 \right) = 1
+% \end{equation}
+% To prove~\eqref{eq:andersen1}, we decompose~$\tilde U_n (\theta_0)$ into
+% three terms
+% \begin{equation}
+% \label{eq:decomp}
+% n^{-1} \tilde U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
+%   \hat U_n (\theta_0) - U_n (\theta) \right) + n^{-1} \left(\tilde U_n
+%   (\theta_0) - \hat U_n (\theta_0) \right)
+% \end{equation}
+% where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t,s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
 \cite{Andersen1993} (Theorem VI.1.1) show that~$n^{-1} U_n (\theta_0)
-\to 0$. The second term
-\begin{align*}
-n^{-1} \left( \hat U_n (\theta_0) - U_n (\theta_0) \right) =
-  \frac{1}{n} \sum_{i=1}^n \int_0^{\tau} \frac{h^{(1)} (s;
-  \theta_0)}{\pi_i (s) + h (s;\theta_0)} d M_i (s).
-\end{align*}
-where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}.
+% \to 0$. The second term
+% \begin{align*}
+% n^{-1} \left( \hat U_n (\theta_0) - U_n (\theta_0) \right) =
+%   \frac{1}{n} \sum_{i=1}^n \int_0^{\tau} \frac{h^{(1)} (s;
+%   \theta_0)}{\pi_i (s) + h (s;\theta_0)} d M_i (s).
+% \end{align*}
+% where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}.
 Lenglart's inequality implies the second term converges in probability to zero under Assumptions~\ref{E1} and~\ref{E2}.
 The third term satisfies
-\begin{align*}
-  &n^{-1} \left \| \hat U_n (\theta_0) - \tilde U_n (\theta_0) \right \| \\
-  \leq & M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \bigg \| g
-         \left( \tilde W_{i,t}^\top \theta  + \sum_{k=K_x+1}^\infty \hat
-         c_{i,k} (t) \int_{t-\Delta}^t \hat \psi_k (s) \phi (s)^\top b
-         + \log (\pi_i (t) ) \right) - g \left( \tilde W_{i,t}^\top
-         \theta + \log (\pi_i (t)) \right) \bigg \| \\
-&\leq M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
-    \sup_{x \in \mathbb{R}} \left \| g^\prime \left( x \right)
-  \right\| \times \left \| \sum_{k=K_x+1}^\infty \hat c_{i,k} (t)
-  \sum_{l=1}^{K_b} \int_{t-\Delta}^t \hat \psi_k (s) \phi_l (s) b_l
-  \right \| \\
-&= \frac{M}{4} n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
-  \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
-  \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
-  \phi_l (s) b_l\right \| \\
-&\to \frac{M}{4} \int_0^\tau \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
-  \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
-  \phi_l (s) b_l\right \| (f_0(t) + f_1(t)) dt
-\end{align*}
-where~$M = \sup \| \tilde W_{i,t} \|$,~$g(x) = 1/(1+\exp(-x))$ is the
-expit function.
-The second inequality is due to the Taylor remainder theorem.
+% \begin{align*}
+%   &n^{-1} \left \| \hat U_n (\theta_0) - \tilde U_n (\theta_0) \right \| \\
+%   \leq & M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i} \bigg \| g
+%          \left( \tilde W_{i,t}^\top \theta  + \sum_{k=K_x+1}^\infty \hat
+%          c_{i,k} (t) \int_{t-\Delta}^t \hat \psi_k (s) \phi (s)^\top b
+%          + \log (\pi_i (t) ) \right) - g \left( \tilde W_{i,t}^\top
+%          \theta + \log (\pi_i (t)) \right) \bigg \| \\
+% &\leq M n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
+%     \sup_{x \in \mathbb{R}} \left \| g^\prime \left( x \right)
+%   \right\| \times \left \| \sum_{k=K_x+1}^\infty \hat c_{i,k} (t)
+%   \sum_{l=1}^{K_b} \int_{t-\Delta}^t \hat \psi_k (s) \phi_l (s) b_l
+%   \right \| \\
+% &= \frac{M}{4} n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup \bfD_i}
+%   \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
+%   \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
+%   \phi_l (s) b_l\right \| \\
+% &\to \frac{M}{4} \int_0^\tau \sum_{l=1}^{K_b} \left \| \int_{t-\Delta}^t
+%   \left( X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \right)
+%   \phi_l (s) b_l\right \| (f_0(t) + f_1(t)) dt
+% \end{align*}
+% where~$M = \sup \| \tilde W_{i,t} \|$,~$g(x) = 1/(1+\exp(-x))$ is the
+% expit function.
+% The second inequality is due to the Taylor remainder theorem.
 The third is due to~$g^\prime (x) = e^x/(1+e^x)^2 \leq e^0/(1+e^0)^2 = 1/4$, reordering the summation, and rewriting the error term in terms of the difference between~$X(t,s)$ and the approximation with truncation level~$K_x$.
@@
 Finally, under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
-\[
-X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \overset{P}{\to} 0
-\]
-as~$K_x \to \infty$, which implies the third term goes to zero as the truncation level goes to infinity.
+% \[
+% X(t,s) - \sum_{k=1}^{K_x} \hat c_{i,k} (t) \psi_k (s) \overset{P}{\to} 0
+% \]
+% as~$K_x \to \infty$, which implies the third term goes to zero as the truncation level goes to infinity.
 The same argument can be used in conjunction under assumptions~\ref{E1} and~\ref{E4} to prove~\eqref{eq:andersen2}.
 
-To prove~\eqref{eq:andersen3}, we start with equation~\eqref{eq:approxscore}, which we rewrite here for completeness:
-\[
-\hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
-  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
-  \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
-\]
-Let~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ be the event and subsampling counting processes with jumps at $t \in \bfT_i$ and $\bfD_i$ respectively.  Then
-\begin{equation}
-\label{eqapp:upperbound}
-\left | n^{-1} \frac{d^2}{d \theta_j d \theta_k} \tilde U_n (\theta)  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau w_i (t;\theta) \times H_{in} (t) R_i (t) dN_i^{(e)} (t) +
-\frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{w_i (t;\theta)}{\pi_i (t)} \times G_{in} (t) R_i (t) dN_i^{(e)} (t)
-\end{equation}
-where $H_{in}(t)$ and $G_{in}(t)$ are from Condition VI.1.1(E) in~\cite[pp.
+% To prove~\eqref{eq:andersen3}, we start with equation~\eqref{eq:approxscore}, which we rewrite here for completeness:
+% \[
+% \hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
+%   \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
+%   \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
+% \]
+% Let~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ be the event and subsampling counting processes with jumps at $t \in \bfT_i$ and $\bfD_i$ respectively.  Then
+% \begin{equation}
+% \label{eqapp:upperbound}
+% \left | n^{-1} \frac{d^2}{d \theta_j d \theta_k} \tilde U_n (\theta)  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau w_i (t;\theta) \times H_{in} (t) R_i (t) dN_i^{(e)} (t) +
+% \frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{w_i (t;\theta)}{\pi_i (t)} \times G_{in} (t) R_i (t) dN_i^{(e)} (t)
+% \end{equation}
+% where $H_{in}(t)$ and $G_{in}(t)$ are from Condition VI.1.1(E) in~\cite[pp.
 421]{Andersen1993}.
@@
 Since~$w_i(t;\theta)/\pi_i (t) = (\pi_i(t)+h_i(t;\theta))^{-1}$, the second term which simplifies to
-\[
-\frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{G_{in} (t)}{\pi_i (t)+h_i(t;\theta)} R_i(t) dN_i^{(e)} (t)
-\leq
-\frac{1}{L \cdot n} \sum_{i=1}^n \int_0^\tau G_{in} (t) R_i (t) dN_i^{(e)} (t)
-\]
-by Assumption~\ref{E1} (i.e., the subsampling rate is lower bounded by $L$). The right hand side is the optional variation of the local square integrable martingale
-$$
-\frac{1}{L \cdot n} \sum_{i=1}^n \int_0^t G^{1/2}_{in} (s) dN_i^{(e)} (s)
-$$
-This martingale has predicable variation process
-$$
-\frac{1}{L \cdot n} \sum_{i=1}^n \int_0^t G_{in} (s) \pi_i (s) R_i (s) ds
-$$
-Since both processes have the same limits, conditions VI.1.1(E) in~\cite{Andersen1993} and Assumption~\ref{E1} imply that the second term on the right hand side of~\eqref{eqapp:upperbound} also converges to a finite quantity as $n \to \infty$ for every $K_x$ which completes the proof.
-
-This proves~$\hat \theta_n \overset{P}{\to} \theta_0$ as $n \to \infty$ and $K_x \to \infty$.
+% \[
+% \frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{G_{in} (t)}{\pi_i (t)+h_i(t;\theta)} R_i(t) dN_i^{(e)} (t)
+% \leq
+% \frac{1}{L \cdot n} \sum_{i=1}^n \int_0^\tau G_{in} (t) R_i (t) dN_i^{(e)} (t)
+% \]
+% by Assumption~\ref{E1} (i.e., the subsampling rate is lower bounded by $L$). The right hand side is the optional variation of the local square integrable martingale
+% $$
+% \frac{1}{L \cdot n} \sum_{i=1}^n \int_0^t G^{1/2}_{in} (s) dN_i^{(e)} (s)
+% $$
+% This martingale has predicable variation process
+% $$
+% \frac{1}{L \cdot n} \sum_{i=1}^n \int_0^t G_{in} (s) \pi_i (s) R_i (s) ds
+% $$
+% Since both processes have the same limits, conditions VI.1.1(E) in~\cite{Andersen1993} and Assumption~\ref{E1} imply that the second term on the right hand side of~\eqref{eqapp:upperbound} also converges to a finite quantity as $n \to \infty$ for every $K_x$ which completes the proof.
+
+% This proves~$\hat \theta_n \overset{P}{\to} \theta_0$ as $n \to \infty$ and $K_x \to \infty$.
 Note that the argument holds for any choice of weight functions~$w_i (t;\theta)$; implying convergence holds more broadly for all approximate score equations in~\eqref{eq:approxscore}.
-\end{proof}
-
-\begin{proof}[Proof of Asymptotic Normality]
-Consider a Taylor-expansion of $\tilde U_n (\theta)$ centered at $\theta_0$ when $\hat \theta_n \in \Theta_0$.  We can write as
-\begin{align*}
-\tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
-0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
-n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
-\end{align*}
-where~$\theta^\star$ is on the line segment between $\hat \theta_n$ and $\theta_0$. First, the term
-$$
-\frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
-\overset{P}{\to} \Xi(\theta_0)
-$$
-under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}.
+% \end{proof}
+
+% \begin{proof}[Proof of Asymptotic Normality]
+% Consider a Taylor-expansion of $\tilde U_n (\theta)$ centered at $\theta_0$ when $\hat \theta_n \in \Theta_0$.  We can write as
+% \begin{align*}
+% \tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
+% 0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
+% n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
+% \end{align*}
+% where~$\theta^\star$ is on the line segment between $\hat \theta_n$ and $\theta_0$. First, the term
+% $$
+% \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
+% \overset{P}{\to} \Xi(\theta_0)
+% $$
+% under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}.
 To see this, note the first term converges to $\Xi(\theta_0)$ while the second term is bounded in probability by $p M \| \theta^\star - \theta_0 \|$ for some finite constant~$M$ independent of~$\theta^\star$ where $\| \cdot \|$ is the Euclidean norm.
@@
 
-We now consider $n^{-1/2} \tilde U_n (\theta_0 )$.
+% We now consider $n^{-1/2} \tilde U_n (\theta_0 )$.
 Recall $\tilde U_n (\theta_0)$ can be decomposed into three terms as in~\eqref{eq:decomp}.
@@
 This leaves two terms
-$$
-U_n (\theta_0) = \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(u; \theta)}{h (u;\theta)} dM_i^{(e)} (t)
-$$
-where~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ are the orthogonal event and subsampling counting processes with jumps at times $t \in \bfT_i$ and $t \in \bfD_i$ respectively.
-and
-\begin{align*}
-(\hat U_n (\theta_0) - U_n (\theta_0) )
-&= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s) \\
-&= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM^{(e)}_i (s) +
-\sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i^{(s)} (s) \\
-&= \sum_{i=1}^n \int_0^\tau (1-w(s;\theta_0)) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s) +
-\sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} dM_i^{(s)} (s)
-\end{align*}
-Taking difference with $U_n(\theta_0)$, the first term becomes
-\[
-\sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s)
-\]
-
-These are two orthogonal square integrable martingales with quadratic variation given by
-\[
-\sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds
-\]
-and
-\[
-\sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds
-\]
-respectively.  Using the definition $w(s;\theta_0) = \pi_i(s) / (\pi_i(s)+h_i(s;\theta_0))$, we have
-\begin{align*}
-&\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds \\
-&+
-\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds \\
-=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ 1 + \frac{h(s;\theta_0)}{\pi_i (s)} \right] ds \\
-=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ \frac{\pi_i (s) + h(s;\theta_0)}{\pi_i (s)} \right] ds \\
-=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} ds
-\to \Xi (\theta_0).
-\end{align*}
-Returning to the Taylor-expansion, under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}, we can apply Rebolledo's martingale central limit theorem so that
-$$
-\sqrt{n} (\hat \theta_n - \theta_0) \overset{D}{\to} N(0, \Xi(\theta_0) ^{-1} \times \Xi(\theta_0) \times \Xi(\theta_0) ^{-1} )
-$$
-which completes the proof.
-\end{proof}
-
-% \section{Penalized logistic mixed-effects model}
-% \label{app:penlogit}
-
-% Here we derive a penalized adaptive Gaussian quadrature (PADQ) to fit the   penalized logistic mixed-effects model.
+% $$
+% U_n (\theta_0) = \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(u; \theta)}{h (u;\theta)} dM_i^{(e)} (t)
+% $$
+% where~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ are the orthogonal event and subsampling counting processes with jumps at times $t \in \bfT_i$ and $t \in \bfD_i$ respectively.
+% and
+% \begin{align*}
+% (\hat U_n (\theta_0) - U_n (\theta_0) )
+% &= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s) \\
+% &= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM^{(e)}_i (s) +
+% \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i^{(s)} (s) \\
+% &= \sum_{i=1}^n \int_0^\tau (1-w(s;\theta_0)) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s) +
+% \sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} dM_i^{(s)} (s)
+% \end{align*}
+% Taking difference with $U_n(\theta_0)$, the first term becomes
+% \[
+% \sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s)
+% \]
+
+% These are two orthogonal square integrable martingales with quadratic variation given by
+% \[
+% \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds
+% \]
+% and
+% \[
+% \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds
+% \]
+% respectively.  Using the definition $w(s;\theta_0) = \pi_i(s) / (\pi_i(s)+h_i(s;\theta_0))$, we have
+% \begin{align*}
+% &\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds \\
+% &+
+% \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds \\
+% =& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ 1 + \frac{h(s;\theta_0)}{\pi_i (s)} \right] ds \\
+% =& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ \frac{\pi_i (s) + h(s;\theta_0)}{\pi_i (s)} \right] ds \\
+% =& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} ds
+% \to \Xi (\theta_0).
+% \end{align*}
+% Returning to the Taylor-expansion, under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}, we can apply Rebolledo's martingale central limit theorem so that
+% $$
+% \sqrt{n} (\hat \theta_n - \theta_0) \overset{D}{\to} N(0, \Xi(\theta_0) ^{-1} \times \Xi(\theta_0) \times \Xi(\theta_0) ^{-1} )
+% $$
+% which completes the proof.
+% \end{proof}
+
+% % \section{Penalized logistic mixed-effects model}
+% % \label{app:penlogit}
+
+% % Here we derive a penalized adaptive Gaussian quadrature (PADQ) to fit the   penalized logistic mixed-effects model.
 Here, for clarity, we write $Y_{ij} \in \{ 0, 1 \}$ to be the sequence of binary indicators for participant $i = 1,\ldots,n$ and $j$ indexes the event and subsampled times.
@@
 Then the score equations for the penalized logistic regression with offset $\log(\pi_{ij})$ is written as
-% $$
-% \sum_{i=1}^n \sum_{j=1}^{k_i} \left( y_{ij} - \frac{1}{1+\exp (- W_{ij} \theta - \log (\pi_{ij}) )} \right) W_{ij} - \frac{1}{\sigma^2} \sum_{k=3}^{K_b} \beta_k
-% $$
-% Let $B$ be the matrix that extracts the entries in $\theta$ corresponding to $\{ \beta_k\}_{k=3}^{K_b}$ (i.e., $B \theta = ({\bf 0}, \beta_3, \ldots, \beta_{K_b},{\bf 0})$).
+% % $$
+% % \sum_{i=1}^n \sum_{j=1}^{k_i} \left( y_{ij} - \frac{1}{1+\exp (- W_{ij} \theta - \log (\pi_{ij}) )} \right) W_{ij} - \frac{1}{\sigma^2} \sum_{k=3}^{K_b} \beta_k
+% % $$
+% % Let $B$ be the matrix that extracts the entries in $\theta$ corresponding to $\{ \beta_k\}_{k=3}^{K_b}$ (i.e., $B \theta = ({\bf 0}, \beta_3, \ldots, \beta_{K_b},{\bf 0})$).
 Let $\W_i$ be a $k_i \times p$ matrix with the $j$th row equal to $W_{ij}$.
@@
 
-% We derive the iteratively re-weighted least squares algorithm next. First, take the derivative with respect to $\theta$ to construct the Hessian
-% $$
-% - \left[ \sum_{i=1}^n \sum_{j=1}^{k_i} \frac{W_{ij} W_{ij}^\top}{(1+\exp (- W_{ij} \theta - \log (\pi_{ij}) ))^2} \exp(-W_{ij} \theta - \log (\pi_{ij})) + \frac{B}{\sigma^2} \right].
-% $$
-% Define the matrix $V_i (\theta)$ be the diagonal matrix with the $(j,j)$th entry equal to $\mu(W_{ij}; \theta, \pi_{ij}) \cdot (1-\mu(W_{ij}; \theta, \pi_{ij})$.
+% % We derive the iteratively re-weighted least squares algorithm next. First, take the derivative with respect to $\theta$ to construct the Hessian
+% % $$
+% % - \left[ \sum_{i=1}^n \sum_{j=1}^{k_i} \frac{W_{ij} W_{ij}^\top}{(1+\exp (- W_{ij} \theta - \log (\pi_{ij}) ))^2} \exp(-W_{ij} \theta - \log (\pi_{ij})) + \frac{B}{\sigma^2} \right].
+% % $$
+% % Define the matrix $V_i (\theta)$ be the diagonal matrix with the $(j,j)$th entry equal to $\mu(W_{ij}; \theta, \pi_{ij}) \cdot (1-\mu(W_{ij}; \theta, \pi_{ij})$.
 Then we can express the Hessian more succinctly as $- \left[ \sum_{i=1}^n \W_i^\top V_i (\theta) \W_i + \frac{B}{\sigma^2} \right]$.
 Combining these, starting at an initial estimate, the Newton-Rhapson update given current parameter value $\hat \theta_k$ is given by
-% $$
-% \hat \theta_{k+1} = \hat \theta_{k} + \left[ \sum_{i=1}^n \W_i^\top V_i (\hat \theta_k) \W_i + \frac{B}{\sigma^2} \right]^{-1} \left(\sum_{i=1}^n \W_i^\top E_i (\hat \theta_k) - \frac{1}{\sigma^2} B \hat \theta_k \right)
-% $$
-
-% To learn $\sigma^2$, we use the connection to the mixed-effects literature.
-% Fixing $\theta$, the log-likelihood components related to $\sigma^2$ is given by $-\frac{1}{2\sigma^2} \theta^\top B \theta - \frac{B {\bf 1}}{2} \log (\sigma^2)$.
+% % $$
+% % \hat \theta_{k+1} = \hat \theta_{k} + \left[ \sum_{i=1}^n \W_i^\top V_i (\hat \theta_k) \W_i + \frac{B}{\sigma^2} \right]^{-1} \left(\sum_{i=1}^n \W_i^\top E_i (\hat \theta_k) - \frac{1}{\sigma^2} B \hat \theta_k \right)
+% % $$
+
+% % To learn $\sigma^2$, we use the connection to the mixed-effects literature.
+% % Fixing $\theta$, the log-likelihood components related to $\sigma^2$ is given by $-\frac{1}{2\sigma^2} \theta^\top B \theta - \frac{B {\bf 1}}{2} \log (\sigma^2)$.
 Differentiating and setting equal to zero yields the estimator
-% $\hat \sigma^2 = \frac{\theta^\top B \theta}{B {\bf 1}}$.
+% % $\hat \sigma^2 = \frac{\theta^\top B \theta}{B {\bf 1}}$.
 The complete algorithm is then alternating between Netwon-Rhapson for fixed $\sigma^2$ to estimate $\theta$ and then estimating $\sigma^2$ for fixed $\theta$ using the above formula.
 
-% \subsection{Estimation of random-effects via Newton-Rhapson}
-
-% \textcolor{red}{Assume fixed $\lambda$.}
-
-% In standard logistic mixed-effects model, with $b_i \sim N(0, \Psi)$ and $b_i \in \mathbb{R}^q$, then the likelihood is given by
-% $$
-% \begin{aligned}
-% & \prod_{i=1}^n \int \prod_{j=1}^{k_i} \left( \frac{\exp(W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij})}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{y_{ij}} \left( \frac{1}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{1-y_{ij}} \phi (b_i; \sigma_b^2) db_i \\
-% \propto & \prod_{i=1}^n \int \exp \left[ \sum_{j=1}^{k_i} \left( y_{ij} \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) - \log \left( 1 + \exp \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) \right) \right) - b_i^\top \Psi^{-1} b_i /2 \right] db_i \\
-% \propto& \prod_{i=1}^n \int \exp \left[ g(\theta, b_i, \pi_{ij}, \Psi) \right] db_i,
-% \end{aligned}
-% $$
-% where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$.
+% % \subsection{Estimation of random-effects via Newton-Rhapson}
+
+% % \textcolor{red}{Assume fixed $\lambda$.}
+
+% % In standard logistic mixed-effects model, with $b_i \sim N(0, \Psi)$ and $b_i \in \mathbb{R}^q$, then the likelihood is given by
+% % $$
+% % \begin{aligned}
+% % & \prod_{i=1}^n \int \prod_{j=1}^{k_i} \left( \frac{\exp(W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij})}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{y_{ij}} \left( \frac{1}{1 + \exp(W_{ij} \beta + Z_{ij} b_i + \log \pi_{ij})} \right)^{1-y_{ij}} \phi (b_i; \sigma_b^2) db_i \\
+% % \propto & \prod_{i=1}^n \int \exp \left[ \sum_{j=1}^{k_i} \left( y_{ij} \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) - \log \left( 1 + \exp \left( W_{ij} \theta + Z_{ij} b_i + \log \pi_{ij} \right) \right) \right) - b_i^\top \Psi^{-1} b_i /2 \right] db_i \\
+% % \propto& \prod_{i=1}^n \int \exp \left[ g(\theta, b_i, \pi_{ij}, \Psi) \right] db_i,
+% % \end{aligned}
+% % $$
+% % where the second and third line ignore the constant $(2\pi)^{q/2} |\Psi|^{-1/2}$.
 Differentiating the function $g(\theta, b_i, \pi_{ij}, \Psi)$ with respect to $b_i$, the connection to the prior section can be used.
 Re-defining terms appropriately allows us to express the derivatives as
-% $$
-% \begin{aligned}
-% \frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
-% \frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
-% \end{aligned}
-% $$
-% For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
-% $$
-% \hat b_i^{(k+1)} = \hat b_i^{(k)} + \left[ Z_i V_i (\theta, \hat b_i^{(k)}) Z_i + \Psi^{-1} \right]^{-1}  \left( Z_i E_i (\theta, \hat b_i^{(k)}) - \Psi^{-1} \hat b_i^{(k)} \right)
-% $$
-
-% Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
+% % $$
+% % \begin{aligned}
+% % \frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i} &= Z_i E_i (\theta, b_i) - \Psi^{-1} b_i \\
+% % \frac{\partial g(\theta, b_i, \pi_{i}, \Psi)}{\partial b_i^2} &= - \left[ Z_i V_i (\theta, b_i) Z_i + \Psi^{-1} \right]
+% % \end{aligned}
+% % $$
+% % For fixed $\theta$, the function $g$ is stictly concave function of $b_i$ and therefore we can estimate the unique maximum $\hat b_i$ via Newton-Rhapson method as follows:
+% % $$
+% % \hat b_i^{(k+1)} = \hat b_i^{(k)} + \left[ Z_i V_i (\theta, \hat b_i^{(k)}) Z_i + \Psi^{-1} \right]^{-1}  \left( Z_i E_i (\theta, \hat b_i^{(k)}) - \Psi^{-1} \hat b_i^{(k)} \right)
+% % $$
+
+% % Fixing $\hat b_i$, we then need to estimate $\Psi$ and $\theta$.
 Here we employ adaptive gaussian quadrature (cites) as it has been shown that for infrequent events the Laplace approximation tends to underestimate effects (cites).
@@
 Then the AGQ rule is given by
+% % $$
+% % \begin{aligned}
+% % &\int \exp \left[ g(\theta, b_i, \pi_{i}, \Psi) \right] db_i \\
+% % =& \int | R_i |^{-1} \exp \left[ g(\theta, \hat b_i + R_i^{-1} \z, \pi_{i}, \Psi) + \| \z \|^2 / 2 \right] \exp ( - \| \z \|^2 / 2 ) d\z \\
+% % =& (2 \pi)^{q/2} | R_i |^{-1} \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k}
+% % \end{aligned}
+% % $$
+% % The score equations
+% % $$
+% % \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}}
+% % \left( \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right)^{-1}
+% % \left( \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi)
+% % $$
+% % Define $\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi)$ to be the weights.  Then we can express score and Hessian as:
+% % $$
+% % \begin{aligned}
+% % &\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \\
+% % &\sum_{\k}
+% % \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial^2}{\partial \theta \partial \theta^\top} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) +
+% % \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top \\
+% % &- \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top
+% % \end{aligned}
+% % $$
+
+% % When $N_{GQ} = 1$, the AGQ approximation is equivalent to a Laplacian approximation.
+
+% \section{Optimal weights}
+
+% \begin{proof}[Proof of Proposition~\ref{prop:optimal}]
+% The proof follows closely from~\cite{Rathbun2012} and~\cite{Ferreira82}.
+Revisiting the central limit theorem above, we can see that for any weights~$w(s; \theta_0)$ that the asymptotic normality is guaranteed with the asymptotic variance $\Sigma(\theta_0; w) := \Xi(\theta_0; w)^{-1} V(\theta_0; w) \Xi(\theta_0; w)^{-1}$ where
 % $$
-% \begin{aligned}
-% &\int \exp \left[ g(\theta, b_i, \pi_{i}, \Psi) \right] db_i \\
-% =& \int | R_i |^{-1} \exp \left[ g(\theta, \hat b_i + R_i^{-1} \z, \pi_{i}, \Psi) + \| \z \|^2 / 2 \right] \exp ( - \| \z \|^2 / 2 ) d\z \\
-% =& (2 \pi)^{q/2} | R_i |^{-1} \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k}
-% \end{aligned}
+% V(\theta_0; w) = \int_0^\tau w^2 (s; \theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} \left[ \frac{\pi (s) + h(s;\theta_0)}{\pi (s)} \right] ds
 % $$
-% The score equations
-% $$
-% \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}}
-% \left( \sum_{j_1 =1}^{N_{GQ}} \cdots \sum_{j_q =1}^{N_{GQ}} \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right)^{-1}
-% \left( \exp \left[ g(\theta, \tilde b_{i \k}, \pi_{i}, \Psi) \right] W_{\k} \right) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi)
-% $$
-% Define $\omega(\theta, \tilde b_{i \k}, \pi_i, \Psi)$ to be the weights.  Then we can express score and Hessian as:
-% $$
-% \begin{aligned}
-% &\sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \\
-% &\sum_{\k}
-% \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \frac{\partial^2}{\partial \theta \partial \theta^\top} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) +
-% \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \cdot \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left [\frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top \\
-% &- \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right] \left[ \sum_{\k} \omega(\theta, \tilde b_{i \k}, \pi_i, \Psi) \frac{\partial}{\partial \theta} g(\theta, \tilde b_{i \k}, \pi_i, \Psi) \right]^\top
-% \end{aligned}
-% $$
-
-% When $N_{GQ} = 1$, the AGQ approximation is equivalent to a Laplacian approximation.
-
-\section{Optimal weights}
-
-\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
-The proof follows closely from~\cite{Rathbun2012} and~\cite{Ferreira82}.
-Revisiting the central limit theorem above, we can see that for any weights~$w(s; \theta_0)$ that the asymptotic normality is guaranteed with the asymptotic variance $\Sigma(\theta_0; w) := \Xi(\theta_0; w)^{-1} V(\theta_0; w) \Xi(\theta_0; w)^{-1}$ where
-$$
-V(\theta_0; w) = \int_0^\tau w^2 (s; \theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} \left[ \frac{\pi (s) + h(s;\theta_0)}{\pi (s)} \right] ds
-$$
-and dependence on $w$ is made explicit for ease of comprehension.
-To prove optimality of our particular weight choice we study a standardized set of estimating equations~$\Xi (\theta; w)^{-1} U_n (\theta; w)$.
+% and dependence on $w$ is made explicit for ease of comprehension.
+% To prove optimality of our particular weight choice we study a standardized set of estimating equations~$\Xi (\theta; w)^{-1} U_n (\theta; w)$.
 Let $w^\star$ denote the Waagepeterson weights from~\eqref{eq:waage_weights}.
@@
 Based on the asymptotic normality arguments above, the covariance will have a similar sandwich form $\Xi (\theta_0; w)^{-1} \cov (\theta_0; w,w^\star) \Xi(\theta_0; w^\star)^{-1}$ where the middle term is given by
-\begin{align*}
-&\int_0^\tau w(s;\theta_0) w^\star (s; \theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} \left[ \frac{\pi (s) + h(s;\theta_0)}{\pi (s)} \right] ds \\
-=&\int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} ds \\
-&= \Xi (\theta_0; w).
-\end{align*}
-This implies the covariance is equal to $\Xi(\theta_0; w^\star)^{-1}$.
+% \begin{align*}
+% &\int_0^\tau w(s;\theta_0) w^\star (s; \theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} \left[ \frac{\pi (s) + h(s;\theta_0)}{\pi (s)} \right] ds \\
+% =&\int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h (s;\theta_0)} ds \\
+% &= \Xi (\theta_0; w).
+% \end{align*}
+% This implies the covariance is equal to $\Xi(\theta_0; w^\star)^{-1}$.
 Optimality is defined as the difference in variances being positive semi-definite.
@@
 For any~$\alpha \in \mathbb{R}^{p}$,
-\begin{align*}
-0 &\leq \left( \alpha^\top - \alpha^\top \right)
-\left( \begin{array}{c c} \Sigma(\theta_0; w) & \Xi (\theta_0; w^\star)^{-1} \\ \Xi(\theta_0; w^\star)^{-1} & \Xi (\theta_0; w^\star)^{-1} \end{array} \right)
-\left( \begin{array}{c} \alpha & - \alpha \end{array} \right)\\
-&= \alpha^\top \left( \Sigma(\theta_0; w) - \Xi (\theta_0; w^\star)^{-1} \right) \alpha.
-\end{align*}
-which demonstrates optimality of the Waagepetersen weights from~\eqref{eq:waage_weights}.
-
-Uniqueness is not guaranteed by the above.  Let $w$ be another estimator with minimal variance.  Then define the estimating equation $\tilde U_n (
-\theta; w) - \tilde U_n (\theta; w^\star)$.
+% \begin{align*}
+% 0 &\leq \left( \alpha^\top - \alpha^\top \right)
+% \left( \begin{array}{c c} \Sigma(\theta_0; w) & \Xi (\theta_0; w^\star)^{-1} \\ \Xi(\theta_0; w^\star)^{-1} & \Xi (\theta_0; w^\star)^{-1} \end{array} \right)
+% \left( \begin{array}{c} \alpha \\ - \alpha \end{array} \right)\\
+% &= \alpha^\top \left( \Sigma(\theta_0; w) - \Xi (\theta_0; w^\star)^{-1} \right) \alpha.
+% \end{align*}
+% which demonstrates optimality of the Waagepetersen weights from~\eqref{eq:waage_weights}.
+
+% Uniqueness is not guaranteed by the above.  Let $w$ be another estimator with minimal variance.  Then define the estimating equation $\tilde U_n (
+% \theta; w) - \tilde U_n (\theta; w^\star)$.
 By definition of the estimating equations, the difference is simply another estimation equation of the same weighted form with different weights~$w_1$.
 Thus, under the standardized set of estimating equations we have the variance is equal to
-$\Sigma (\theta; w_1) = \Sigma (\theta; w) - \Sigma(\theta; w^\star)$,
-but $\Sigma (\theta; w^\star) - \Sigma(\theta; w)$ is non-negative definite by the above argument.
+% $\Sigma (\theta; w_1) = \Sigma (\theta; w) - \Sigma(\theta; w^\star)$,
+% but $\Sigma (\theta; w^\star) - \Sigma(\theta; w)$ is non-negative definite by the above argument.
 This implies~$\Sigma(\theta;w_1) = 0$ and thus the two standardized estimating equations are equivalent.
@@
 
-\end{proof}
-
-\section{Additional simulations and case study details}
-
-\subsection{Impact of $\Delta$}
-\label{app:delta}
-
-Here we investigate the impact of misspecification of the window length for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
+% \end{proof}
+
+% \section{Additional simulations and case study details}
+
+% \subsection{Impact of $\Delta$}
+% \label{app:delta}
+
+% Here we investigate the impact of misspecification of the window length for $\beta(t) = \beta_0 \exp \left( - \beta_1 s \right)$.
 To do this, we generate 1000 datasets per condition each consisting of 500 user-days.
@@
 
-When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$; however, we incur a penalty in finite samples, especially for settings where the function is far from zero near $t = \Delta$.
+% When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$; however, we incur a penalty in finite samples, especially for settings where the function is far from zero near $t = \Delta$.
 We find the MISE increases and variance slightly decreases as $\Delta$ increases.
@@
 
-\begin{table}[!th]
-\begin{tabular}{c | c c c c c}
-& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
-$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
-\multirow{3}{*}{26} & MISE & $5.9 \times 10^{-2}$ & $6.1 \times 10^{-2}$ & $6.6 \times 10^{-2}$ & $7.4 \times 10^{-2}$ \\
- & Variance & $1.3 \times 10^{-2}$ & $1.6 \times 10^{-2}$ & $12.4 \times 10^{-2}$ & $3.6 \times 10^{-2}$ \\
-  & P-MISE & $7.1 \times 10^{-2}$ & $7.3 \times 10^{-2}$ & $7.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{29} & MISE & $6.9 \times 10^{-2}$ & $7.0 \times 10^{-2}$ & $7.1 \times 10^{-2}$ & $7.8 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.4 \times 10^{-2}$ \\
-  & P-MISE & $7.5 \times 10^{-2}$ & $7.5 \times 10^{-2}$ & $7.7 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{32} & MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $1.9 \times 10^{-2}$ & $3.3 \times 10^{-2}$ \\
-  & P-MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{35} & MISE & $9.0 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $9.2 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.2 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
-  & P-MISE & $8.5 \times 10^{-2}$ & $8.2 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{37} & MISE & $10.3 \times 10^{-2}$ & $10.1 \times 10^{-2}$ & $9.9 \times 10^{-2}$ & $10.1 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
-  & P-MISE & $9.4 \times 10^{-2}$ & $9.2 \times 10^{-2}$ & $9.0 \times 10^{-2}$ & $9.0 \times 10^{-2}$ \\ \hline
-\end{tabular}
-\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
-\label{tab:appmise}
-\end{table}
-
-\section{Summary of missing data for the case study}
-\label{app:csmissing}
-
-Here we present missing data information for electrodermal EDA and activity index.
+% \begin{table}[!th]
+% \begin{tabular}{c | c c c c c}
+% & & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
+% $\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
+% \multirow{3}{*}{26} & MISE & $5.9 \times 10^{-2}$ & $6.1 \times 10^{-2}$ & $6.6 \times 10^{-2}$ & $7.4 \times 10^{-2}$ \\
+%  & Variance & $1.3 \times 10^{-2}$ & $1.6 \times 10^{-2}$ & $12.4 \times 10^{-2}$ & $3.6 \times 10^{-2}$ \\
+%   & P-MISE & $7.1 \times 10^{-2}$ & $7.3 \times 10^{-2}$ & $7.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ \\ \hline
+% \multirow{3}{*}{29} & MISE & $6.9 \times 10^{-2}$ & $7.0 \times 10^{-2}$ & $7.1 \times 10^{-2}$ & $7.8 \times 10^{-2}$ \\
+%  & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.4 \times 10^{-2}$ \\
+%   & P-MISE & $7.5 \times 10^{-2}$ & $7.5 \times 10^{-2}$ & $7.7 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
+% \multirow{3}{*}{32} & MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\
+%  & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $1.9 \times 10^{-2}$ & $3.3 \times 10^{-2}$ \\
+%   & P-MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\ \hline
+% \multirow{3}{*}{35} & MISE & $9.0 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $9.2 \times 10^{-2}$ \\
+%  & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.2 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
+%   & P-MISE & $8.5 \times 10^{-2}$ & $8.2 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
+% \multirow{3}{*}{37} & MISE & $10.3 \times 10^{-2}$ & $10.1 \times 10^{-2}$ & $9.9 \times 10^{-2}$ & $10.1 \times 10^{-2}$ \\
+%  & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
+%   & P-MISE & $9.4 \times 10^{-2}$ & $9.2 \times 10^{-2}$ & $9.0 \times 10^{-2}$ & $9.0 \times 10^{-2}$ \\ \hline
+% \end{tabular}
+% \caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
+% \label{tab:appmise}
+% \end{table}
+
+% \section{Summary of missing data for the case study}
+% \label{app:csmissing}
+
+% Here we present missing data information for electrodermal EDA and activity index.
 Figure~\ref{fig:missing_eda} and~\ref{fig:missing_acc} presents the patient average fractino of missing EDA and AI data in 30-minute windows respectively.
@@
 
-\begin{figure}[!th]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/missing_data_eda.png}
-  \caption{Electrodermal activity (EDA)}
-  \label{fig:missing_eda}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{../figures/missing_data_ai.png}
-  \caption{Activity Index (AI)
-}  \label{fig:missing_acc}
-\end{subfigure}
-\caption{Patient average fraction missingness in 30-minute windows.}
-\label{fig:mean_edacc}
-\end{figure}
+% \begin{figure}[!th]
+% \centering
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{../figures/missing_data_eda.png}
+%   \caption{Electrodermal activity (EDA)}
+%   \label{fig:missing_eda}
+% \end{subfigure}%
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{../figures/missing_data_ai.png}
+%   \caption{Activity Index (AI)
+% }  \label{fig:missing_acc}
+% \end{subfigure}
+% \caption{Patient average fraction missingness in 30-minute windows.}
+% \label{fig:mean_edacc}
+% \end{figure}
 

### 2023-02-03 letters/response-ref-1-v1.tex
updates from home

@@
 \item {\it Page 9, line 6, isn’t the integral in (2) $\int_{t-\Delta}^t X(s) \beta(s)ds$?
-The whole development assumes the integral to be $\int_{t-\Delta}^t X(t−s) \beta(s)ds$.
+The whole development assumes the integral to be $\int_{t-\Delta}^t X(t-s) \beta(s)ds$.
 The interpretation of the coefficient function $\beta()$ would be different for two cases.
@@
 
-\item {\it Doesn’t Figure 3 suggest smaller ∆?}
+\item {\it Doesn’t Figure 3 suggest smaller $\Delta$?}
 
@@
 I am incredibly grateful for this review.
-I agree that the paper has evolved beyond the current title.
-This was not updated in the prior submission simply to ensure continuity of presentation (i.e., didn't want to change from first submission).
-Here we replace with the title ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information''.
+I am very happy that the referee found that this paper makes a nice contribution to the field of functional data; we, of course, agree.
+The application was what drove this methodology so we are also happy to hear that the referee found the application interesting.
 We next address the reviewers specific comments.
@@
 \begin{enumerate}
-\item {\it The paper states some facts about the SARS-CoV-2 virus that are outdated, for example “Upon infection with SARS-CoV-2, an incubation period starts and lasts approximately five days [Lauer et al., 2020] before the viral load is high enough to be detected.” (Page 5).
-Lauer was based on the original
-strain, we see different disease dynamics with the delta strain.
-Perhaps to be more precise, the author could clarify that this is an example of the original SARS-CoV-2 incubation rather than definitative as there will likely be future variants that will change this further.
-Similarly, on page 13, the serial
-interval of 7 is stated as fact – this is again based on the original strain rather than the current disease dynamics.
-Stating these numbers as an example rather than definitive would be preferable, I believe.}
+\item {\it There exists a good amount of recent efforts modeling event data with functional data predictor or joint modeling, so a literature review on relevant papers would be useful.
+See, for example, papers by Dr.
+Ciprian Crainiceanu from Johns Hopkins, Dr.
+Sheng Luo from Duke, and Dr.
+Jiguo Cao from Simon Fraser.}
+
+\vspace{5mm}
+XXXXXX.
+We added Section XX to describe the related work and connect our methodology to the existing literature.
 \vspace{5mm}

### 2023-02-03 write-up/letters/cover letter-fda-recurrentevents.tex

--- 2023-02-02 letters/cover letter-fdarecurrent.tex
@@
 
-Kindly find enclosed the revised version of the manuscript ID JCGS-22-236, "Recurrent event analysis in the presence of real-time high frequency data via random subsampling."   We have revised the article to address the referees’ requests.
-The reports called for a major revision to include related work on modeling event data with functional data predictors as well as historical functional linear models, include empirical evidence and associated discussion of the computational benefits of our proposal, as well as clarify several technical details in the theoretical sections and provide important intuition/sketches of the proofs.
-The new version complies with these requests.
+Kindly find enclosed the manuscript titled ``Recurrent Event Analysis in the Presence of Real-time High Frequency Data via Random Subsampling.''  This manuscript is motivated by digital monitoring studies that collect real-time high frequency data via mobile sensors and their use in the analysis of recurrent event outcomes such as smoking, drug use, alcohol use, or self-identified moments of suicide ideation.
+Likelihood calculations for the recurrent event analysis become computationally prohibitive in this setting.
+A random subsampling framework is proposed for computationally efficient, approximate likelihood-based estimation which is shown to be equivalent to penalized logistic regression estimation, allowing for standard, ``off-the-shelf'' software to be used in fitting these models.
+Simulations demonstrate the method and efficiency-computation trade-off.
+We end by illustrating our approach using data from a digital monitoring study of suicidal ideation.
 
-We believe the work presents an important advance that gives researchers a simple tool to analyze recurrent event data in the presence of high-frequency sensors data.
-Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.
+We believe the work presents an important advance in how researchers should approach recurrent event analysis in settings in the present of real-time high frequency data such as in mobile health where sensor data is prevalent.
 

### 2023-02-03 letters/response-ref-1-v1.tex
updating response file

@@
 \newpage
-{\bf Response to Reviewer 1}
+{\bf Response to Referee \#1}
 \vspace{5mm}
@@
 
-{\bf Response to Reviewer \#2}
+{\bf Response to Referee \#2}
 
@@
 \vspace{5mm}
+
+\item {\it It might be worth mentioning existing literature on historical functional linear models, e.g., papers by Dr. Damla Senturk from UCLA and collaborators.}
+
+\vspace{5mm}
+XXXXXX.
+We added Section XX to describe the related work and connect our methodology to the existing literature.
+\vspace{5mm}
+
+\item {\it As with any recent history functional linear models, the size of the window ($\Delta$) is an important parameter.
+So it would be helpful to see some efforts on developing a data-adaptive method for the selection of $\Delta$.}
+
+\vspace{5mm}
+XXXXXX
+\vspace{5mm}
+
+\item {\it While in Section 3.6 the paper discusses penalized functional regression to reduce overfitting, it is unclear if the paper actually adopts it in simulation studies and application.
+It is of interest how the smoothing parameter is selected.
+My concern is that if no penalty is employed, with more than 30 basis functions, overfitting can be severe and much wider confidence bands are often obtained}
+
+\vspace{5mm}
+XXXXXX
+\vspace{5mm}
+
+\item {\it While in Section 3.6 the paper discusses penalized functional regression to reduce overfitting, it is unclear if the paper actually adopts it in simulation studies and application.
+It is of interest how the smoothing parameter is selected.
+My concern is that if no penalty is employed, with more than 30 basis functions, overfitting can be severe and much wider confidence bands are often obtained}
+
+\vspace{5mm}
+XXXXXX
+\vspace{5mm}
+
+\item {\it In the theoretical section, it would be helpful to understand what intermediate results Assumption 3.4 will lead to.}
+
+\vspace{5mm}
+XXXXXX
+\vspace{5mm}
+
+\item {\it In the simulation study, it would make sense to me consider a coefficient function for which the magnitude is not only decaying further away from current time and also becomes 0 at $t-\Delta$.
+In addition, it might be helpful to use relative errors to quantify the accuracy of estimation of coefficient functions.}
+
+\vspace{5mm}
+XXXXXXX
+\vspace{5mm}
+
+\item {\it Section 2.1 discusses the rationale of modeling functional predictor without noises.
+Note that the proposed methodology is actually suitable for data with noises.
+First, the data application seems to suggest noises are indeed present; see Figure 2 (A).
+Second, I am not sure I fully understand why in the equation after (3), the conditional probability is conditioning on functional predictor observed up to censoring or truncation time.
+Isn’t it more sensible to condition on functional predictor observed up to t?
+Then, the conditional probability will not depend on functional predictor beyond time t.}
+
+\vspace{5mm}
+The model is under conditional independence $\eta$.
+\vspace{5mm}
+
+\item {\it In Section 3.1, the paper considers modeling the mean function separately at event times and non-event times. Is there some reasoning or justification for this?}
+
+\vspace{5mm}
+Separate models in FACE paper so we can use the ; however, it makes minimial difference.  So we cite this.
+\vspace{5mm}
+
+\item {\it Minor comments: there are typos and inconsistent naming in the paper, e.g., Proposition 3.6 should actually be Lemma 3.6, AI and ACC are used to refer to activity index.}
+
+\vspace{5mm}
+We apologize for the typos and inconsistent naming.  This has been resolved.
+\vspace{5mm}
+
 \end{enumerate}

### 2023-02-07 write-up/fda-recurrent.tex
updated related work

@@
 
+The use of smart devices and other wearables to passively monitor patients has led to a rapid increase in the number of studies with high-frequency sensor streams that can be conceptualized as studies where measurements are functions.
+The development of such applications have been accompanied by intense methodological development in regression models where some covariates are functions.
+In this article, we devleop a novel inferential approach to functional regression~\citep{James2002,JamesSilverman2005,muller2005,Ramsay2005,kokoszka2017,doi:10.1198/jasa.2009.tm08564,doi:10.1198/016214507000000527}.
+
+Recently, modeling time-to-event data with functional predictors has received a fair bit of attention.
+Functional linear Cox regression model (FLCRM) considers the association between a time-to-event and a set of functional~$\{ X(s) \}$ and scalar predictors~$Z \in \mathbb{R}^p$~\citep{https://doi.org/10.1111/biom.12748}.
+In this setting, the functional and scalar predictors are observed at baseline and the hazard function depends on a non-parametric baseline, a functional linear model~$\int_S X(s) \beta(s)ds$ and a linear model for the baseline scalar predictors~$Z^\top \gamma$.
+Since the linear assumption may be too restrictive,~\cite{doi:10.1080/10618600.2020.1853550} consider a more flexible additive functional Cox model for multivariate functional predictors in which the hazard depends on $\int  F \{s, h_{is}[X_i (s)]\} ds$ where $F$ is an unspecified bivariate twice differntiable function.
+
+In our setting, the high frequency sensor process is an outcome and must therefore be jointly modelled with the time-to-event process.
+Recently, \cite{Li2022} proposed a functional approach to joint modelling of multivariate longitudinal biomarkers and a time-to-event outcome~\cite{Tsiatis2004,Rizopoulos2010}.
+Here, as in traditional joint models, the longitudinal biomarker is observed at only a few observation times (typically $<10$) and is modelled via a functional principal components analysis.
+Shared  process  of  to identify the shared progression pattern and outcome-specific progression patterns of the outcomes.
+
+Here a design-based perspective is taken which subsamples non-event times to  avoid some of the computationally demanding aspects of joint models
+Subsampling in the context of functional data analysis for massive data has recently been investigated~\cite{https://doi.org/10.48550/arxiv.2104.03446} in which a functional L-optimality criterion is derived.
+
+The use of historical functional linear models....
+
+The main contribution of this paper is to combine advances in historical functional linear models with techniques from design-based inference to circumvent computationally challenging joint models.
+
+The combination leads to a natural penalized generalized linear model that can be estimated with off-the-shelf software.
+
 \section{Recurrent event process and associated high frequency data}

### 2023-02-10 write-up/fda-recurrent.tex
updated related work and added contributions and outline

--- 2023-02-02 write-up/fda-recurrent.tex
@@
 % display breaks
+\newenvironment{newmaterial}{\par\color{blue}}{\par}
 
@@
 
+\begin{newmaterial}
+The use of smart devices and other wearables to passively monitor patients has led to a rapid increase in the number of studies with high-frequency sensor streams that can be conceptualized as studies measurements are functions.
+The development of such applications have been accompanied by intense methodological development in regression models with functional covariates~\citep{James2002,JamesSilverman2005,muller2005,Ramsay2005,kokoszka2017,doi:10.1198/jasa.2009.tm08564,doi:10.1198/016214507000000527}.
+
+Recently, modelling time-to-event data with functional covariates has received a fair bit of attention.
+The functional linear Cox regression model (FLCRM) considers the association between a time-to-event and a set of functional~$\{ X(s) \}$ and scalar predictors~$Z \in \mathbb{R}^p$~\citep{https://doi.org/10.1111/biom.12748}.
+In this setting, the functional and scalar predictors are observed at baseline and the hazard function satisfies the proportional hazards assumption and involves a non-parametric baseline hazard with the exponential adjustment including both a functional linear model~$\int_S X(s) \beta(s)ds$ and a linear model for the baseline scalar predictors~$Z^\top \gamma$.
+Since the linear assumption may be too restrictive,~\cite{doi:10.1080/10618600.2020.1853550} consider a more flexible additive functional Cox model for multivariate functional predictors in which the hazard depends on $\int  F \{s, h_{is}[X_i (s)]\} ds$ where $F$ is an unspecified bivariate twice differentiable function.
+
+In our setting, the high frequency sensor process is an outcome and must therefore be jointly modelled with the recurrent event process.
+Recently, \cite{Li2022} proposed a functional approach to joint modelling of multivariate longitudinal biomarkers and a time-to-event outcome~\cite{Tsiatis2004,Rizopoulos2010}.
+Here, as in traditional joint models, the longitudinal biomarker is observed at only a few observation times (typically on the order of tens of observations) and is modelled via a functional principal components analysis.
+The eigenscores are the shared parameters across the longitudinal and survival submodels.
+A similar joint model was proposed in~\cite{Dong2021}.
+Inference proceeds by Monte Carlo EM which can require very large Monte Carlo sample size to ensure good performance, and na{\"i}ve implementation may not recover EM's ascent property~\cite{Caffo2005}.
+
+In this paper, we also consider joint modelling of recurrent events and a high frequency sensor process.
+In contrast to traditional longitudinal biomarkers, sensor processes are observed multiple times per second leading to thousand and/or millions of observations per individual rather.
+To address this increase in scale, a design-based perspective is taken in which we subsample non-event times to avoid some of the computationally demanding aspects of joint models.
+Subsampling in the context of functional data analysis for massive data has recently been investigated~\citep{https://doi.org/10.48550/arxiv.2104.03446} in which a functional L-optimality criterion is derived.
+Subsampling is different in our current context, as our goal is to subsample from the individual sensor processes to produce a design-unbiased estimate of the cumulative hazard function.
+Finally, traditional joint modeling often consider the hazard function to depend on the current or cumulative value of the health process~\citep{Li2022,Rizopoulos2010}.
+For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history; to do so, we use recent history functional linear models~\cite{KIM20111554}.
+We combine these models with advances in longitudinal functional data analysis~\citep{Staicu2010} and penalized functional regression~\citep{Goldsmith2015} to perform scalable inference using off-the-shelf software.
+
+\section{Main contributions and outline}
+
+The original contributions of this paper are three-fold.
+In Section~\ref{section:modeloverview}, we define our joint model using historical functional linear models, and demonstrate how techniques from design-based inference can be used to circumvent computational challenges of joint modelling.
+In Section~\ref{section:longfpc}, we discuss longitudinal functional principal components analysis and demonstrate in Lemma~\ref{lemma:logistic} that the resulting approximate score equations are equivalent to score functions for logistic regression with binary response and an offset related to the subsampling rate meaning our proposed approach can be fit using off-the-shelf software.
+An asymptotic analysis is presented in Section~\ref{section:theory}, with an accompanying discussion of the novel computational versus statistical efficiency tradeoff in Section~\ref{section:tradeoff}.
+Section~\ref{section:simstudy} presents a simulation study to demonstrate the impact of subsampling rates, the window length parameter, and a computational comparison to existing methods.
+Section~\ref{eq:jm} presents a multivariate extension as well as a method to handle missing data.
+We end by illustrating our approach using data from a digital monitoring study of suicidal ideation in Section~\ref{section:example}.
+
+\end{newmaterial}
+
 \section{Recurrent event process and associated high frequency data}
+\label{section:modeloverview}
 
@@
   event-history analysis}
+  \label{section:longfpc}
 
 Probabilistic subsampling converts the single sensor stream~$\bfx_i$ into a sequence of functions observed repeatedly at sampled times~$\bfD_i$ and event times~$\bfT_i$ over windows of length~$\Delta$.
-Such a data structure is commonly referred to as \emph{longitudinal functional data}~\citep{Xiao2013, GoldSmith2015}.
+Such a data structure is commonly referred to as \emph{longitudinal functional data}~\citep{Xiao2013,Goldsmith2015}.
 Given the large increase in longitudinal functional data in recent years, corresponding analysis has received much recent attention~\citep{Morris2003, MorrisCarroll2006, Baladandayuthapani2008, Di2009, Greven2010, Staicu2010, ChenMuller2012, LiGuan2014}.
@@
 
-The next step of our method is modeling~$\beta(s)$.
+The next step of our method is modelling~$\beta(s)$.
 Here, we leverage ideas from the penalized spline literature~\citep{Ruppert2003, Wood2006book}.
@@
 \subsection{Theoretical analysis}
+\label{section:theory}
 
@@
 \subsection{Computation versus statistical efficiency tradeoff}
+\label{section:tradeoff}
 
@@
 Second, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
-Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
+Joint modelling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
 
@@
 
-\subsection{Comparison to alternative methods} \label{section:simcomparisons}
+\subsection{Computational comparison to alternative methods} \label{section:simcomparisons}
 
 \section{Extensions}
+\label{section:extensions}
 
@@
 \caption{Average scaled EDA and AI in the 30 minutes prior to button presses}
+\label{Reduced Rank Mixed Effects Models for Spatially Correlated Hierarchical Functional Data
+}
 \label{fig:mean_edacc}
@@
 \caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the complete-case analysis, while highlighted region are the model-based pointwise 95\% confidence intervals.}
-\label{fig:mean_edacc}
+\label{fig:edacc_estimate}
 \end{figure}
@@
 \caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the missing data analysis, while highlighted region are the boostrap and MI based pointwise 95\% confidence intervals.}
-\label{fig:mean_edacc}
+\label{fig:edacc_estimate_bs}
 \end{figure}

### 2023-02-10 write-up/fda-recurrent-supplement.tex
Fixing up the paper by removing the pmcc reference to appendix as its a side point

@@
 \appendix
+
+\section{Measurement-error models with event processes}
+\label{section:memproblems}
+
+One potential criticism for~\eqref{eq:hazard} is that the health process may be measured with error.
+A common mathematical strategy for joint models is to consider an unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i \given \eta_i$, i.e., the two processes are conditionally independent given the latent trajectory.
+For example, take~$\eta_i$ to be a zero-mean Gaussian process with
+\begin{equation}\label{eq:jm}
+x_i(t) = \eta_i (t) + \epsilon_i (t),\quad \text{ and } \log h_i (t
+\given \eta ) = \log h_0 (t) + g_t \left( H_{i,t}^N \right)^{\prime}
+\alpha + \int_{t-\Delta}^t \eta_i (s) \beta (s) ds
+\end{equation}
+where~$\epsilon_i(t)$ is a white-noise measurement error term.
+Thus, $\eta_i (t)$ is the ``true and unobserved value of the longitudinal outcome'' \citep[Sec.
+2.1, pp.3]{Rizopoulos2010}.
+The conditional survival function of the $j$th event~$T_j$ given~$\bfx_i$ and all prior events~$\bfT_{-j} := \{ T_1, \ldots, T_{j-1} \}$ is
+\[
+\pr \left ( T_j > t+s \Given \H_{i,\tau_i}^X, \bfT_{-j} = \bft_{-j}, T_j > t \right) = E \left( - \exp \left( \int_{t}^{t+s} h_i (u \given \eta ) du \right) \Given \bfx_i, \bfT_{-j} = \bft_{-j} \right),
+\]
+where the expectation is a Gaussian integral, albeit infinite-dimensional.
+For most choices of covariance structure, if the white-noise error term is non-zero then the above calculation will show the conditional survival function depends not only on past $x$-values, but also on future $x$-values, i.e.,~$\bfx$ and~$\bfT$ do not satisfy \emph{independent evolution}~\citep{DempseyPMCC2}.
+This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process.
+To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
+Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed; see Section~\ref{section:missingdata} for details on handling of missing data.
 

### 2023-02-10 write-up/fda-recurrent.tex
Fixing up the paper by removing the pmcc reference to appendix as its a side point

@@
 
-\subsection{Measurement-error models with event processes}
-\label{section:memproblems}
-
-One potential criticism for~\eqref{eq:hazard} is that the health process may be measured with error.
-A common mathematical strategy for joint models is to consider an unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i \given \eta_i$, i.e., the two processes are conditionally independent given the latent trajectory.
-For example, take~$\eta_i$ to be a zero-mean Gaussian process with
-\begin{equation}\label{eq:jm}
-x_i(t) = \eta_i (t) + \epsilon_i (t),\quad \text{ and } \log h_i (t
-\given \eta ) = \log h_0 (t) + g_t \left( H_{i,t}^N \right)^{\prime}
-\alpha + \int_{t-\Delta}^t \eta_i (s) \beta (s) ds
-\end{equation}
-where~$\epsilon_i(t)$ is a white-noise measurement error term.
-Thus, $\eta_i (t)$ is the ``true and unobserved value of the longitudinal outcome'' \citep[Sec.
-2.1, pp.3]{Rizopoulos2010}.
-The conditional survival function of the $j$th event~$T_j$ given~$\bfx_i$ and all prior events~$\bfT_{-j} := \{ T_1, \ldots, T_{j-1} \}$ is
-\[
-\pr \left ( T_j > t+s \Given \H_{i,\tau_i}^X, \bfT_{-j} = \bft_{-j}, T_j > t \right) = E \left( - \exp \left( \int_{t}^{t+s} h_i (u \given \eta ) du \right) \Given \bfx_i, \bfT_{-j} = \bft_{-j} \right),
-\]
-where the expectation is a Gaussian integral, albeit infinite-dimensional.
-For most choices of covariance structure, if the white-noise error term is non-zero then the above calculation will show the conditional survival function depends not only on past $x$-values, but also on future $x$-values, i.e.,~$\bfx$ and~$\bfT$ do not satisfy \emph{independent evolution}~\citep{DempseyPMCC2}.
-This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process.
-To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
-Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed; see Section~\ref{section:missingdata} for details on handling of missing data.
-
 \subsection{Likelihood calculation}
@@
 \caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
-\label{tab:mise}
+\label{tab:mise_delta}
 \end{table}
@@
 \caption{Average scaled EDA and AI in the 30 minutes prior to button presses}
-\label{Reduced Rank Mixed Effects Models for Spatially Correlated Hierarchical Functional Data
-}
 \label{fig:mean_edacc}
@@
 % \caption{Patient average fraction missingness in 30-minute windows.}
-% \label{fig:mean_edacc}
+% \label{fig:mean_edacc2}
 % \end{figure}

### 2023-02-13 letters/response-ref-1-v1.tex
writeup and responses all in there now. making progress

@@
 \vspace{5mm}
-We thank the associate editor.  We have addressed the two reviewers' specific comments.  Responses to their concerns are detailed below and have significantly improved the clarify of this manuscript.
+We thank the associate editor.
+We have addressed the two reviewers' specific comments.
+Responses to their concerns are detailed below and have significantly improved the clarify of this manuscript.
+All major edits are highlighted in blue in order to help the referees know what has changed as part of the major revision.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+We have added a detailed discussion of related work in Section 1.1, which includes the specific suggestions on relevant works provided by the second referee; specifically, we discuss recent modeling of time-to-event data with functional covariates as well as joint modelling from a functional regression perspective.
+We contrast these approaches with our design-based approach which becomes necessary when considering high-frequency sensor processes rather than traditional longitudinal studies as the number of measurement occasions increase from tens to thousands/millions of times per individual.
+We connect and contrast our use of subsampling with the general use of subsampling in handling massive data and its recent use in massive data with functional predictors.
+As stated by the second referee, our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
+This motivates Section XX which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+We have addressed the missing notation and helped to justify the transitions/equations.  These additions are highlighted in blue.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+We have decided that the point we were trying to raise is non-essential to the current manuscript, so we have removed the discussion of \emph{independent evolution} and its implication that $\varepsilon (t) = 0$ for all $t >0$.
+The key issue we were trying to clarify is that the traditional joint model formulation in which the observed health process and time-to-event are conditionally independent given the latent process implies that the hazard function conditional on the entire observed health process depends on values after the current time.
+In the case study, for example, this implies that the expected risk of a button press given the entire sensor process will depend on future activity index and electrodermal activity.
+This holds unless we assume the measurement error is set to zero.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
+Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+We added Section XX to discuss choice of window size; specifically, we consider selection based on Akaike information criteria as part of the simulation study rather than simply the impact of incorrect choice of $\Delta$.
+Figure XX shows that the adaptive tuning of $\Delta$ based on AIC does perofrm well empirically.
+We now use this approach in selecting~$\Delta$ in the case study, and provide a plot of AIC across several choices of $\Delta$ in Figure XX in Appendix XX.
+\vspace{5mm}
+
+The smoothing parameter selection is also discussed in Section XX; given our connection to off-the-shelf penalized logistic regression, the smoothing parameter selection occurs as part of this implementation.
+Therefore, we state how these are chosen generally in the software and provide some additional details in Appendix XX.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+To accommodate the second referee's request without making the discussion too long, we have added Remark XX and XX that provide some intuition.
+We have re-written the technical proofs to first provide a sketch proof (with intuition) and then a detailed proof.
+We hope this helps the reader develop intuition on the first read while giving them some helpful guidance in order to facilitate understanding of the detailed theoretical analysis.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+There does not exist a model that can do all so measures of bias and estimation accuracy may be inappropriate.
+To handle this, we simply look at the average or the exponentially weighted where we can conceptualize it as a pre-processing step.
+Then we can apply traditional joint models.
+The issue is that the traditional jiont model cannot scale to this problem.
+Instead, we
+In Section XX, we perform the comparison in a simulation when compared to an unpenalized $K_x = 10$ linear spline model.
+WE see that both are unbiased and our glmnet takes XX minutes while joint modeling took over XX hours.
+To further this comparison, we perform a back-of-the-envelope calculation.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+We apologize for this omission.
+Upon original submission, we had a GitHub repository that included both our code and documentation.
+We did not cite the repository simply because other journals ask for anonymous submissions which means the GitHub link needs to be removed.
+Given JCGS does not require this, we have now cited the repository in the Appendix.
+The GitHub repository has been improved to include better documentation to ensure users of our proposed method can properly implement our solution on their own problems.
+Again we apologize for not correcting this before originally submitting the article.
+We think the revision now meets the standards of JCGS papers in terms of user-friendly software.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
+We apologize for this omission. We have reread everything and addressed any missing notation and equation consistency.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXX
-\vspace{5mm}
-
-\item {\it Expanded literature review on existing modeling approaches.
-Both referees raise this point, and I agree that the contrast with existing works could be more explicit.
-The second referee provides some specific suggestions on relevant works.}
-
-\vspace{5mm}
-XXXX
+We have removed the discussion of \emph{independent evolution}.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
+Therefore, we revert to the step from Park and Staicu (2018) and report the smoothing of the covariance matrix as it comes from their prior work.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXX
+We have added a detailed discussion of related work in Section 1.1, which includes the specific suggestions on relevant works provided by the second referee; specifically, we discuss recent modeling of time-to-event data with functional covariates as well as joint modelling from a functional regression perspective.
+We contrast these approaches with our design-based approach which becomes necessary when considering high-frequency sensor processes rather than traditional longitudinal studies as the number of measurement occasions increase from tens to thousands/millions of times per individual.
+We connect and contrast our use of subsampling with the general use of subsampling in handling massive data and its recent use in massive data with functional predictors.
+As stated by the second referee, our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
+This motivates Section XX which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
 \vspace{5mm}
@@
 \vspace{5mm}
-We agree.  This point was over-emphasized in the original manuscript and so has been moved to the
+We agree.
+This point was over-emphasized in the original manuscript.
+We have decided that the point we were trying to raise is non-essential to the current manuscript, so we have removed the discussion of \emph{independent evolution} and its implications.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXX
+We have now omitted this part from the current manuscript.
+The key issue we were trying to clarify is that the traditional joint model formulation in which the observed health process and time-to-event are conditionally independent given the latent process implies that the hazard function conditional on the entire observed health process depends on values after the current time.
+In the case study, for example, this implies that the expected risk of a button press given the entire sensor process will depend on future activity index and electrodermal activity.
+This holds unless we assume the measurement error is set to zero.
+In this setting,~$\epsilon_i (t) = 0$ and so $\int_{t-\Delta}^t \epsilon_i (s) \beta(s)ds = 0$ by definition.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
+Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXX
+Both of these were defined in Section 3 but we agree that the definitions were obscured in the writing, which led to confusion.  To address this, we have added a notation glossary as Appendix XX.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXX
+We believe the referee is referring to the log-likelihood as its related to the event process. DO WE NEED NO MEASUREMENT ERROR IF WE DO IT THIS WAY? OTHERWISE WE WILL HAVE THE joint likelihood issue.
 \vspace{5mm}
 
-\item {\it In line 39, page 4, by assuming $\epsilon(t) = 0$, you assume there is no measure- ment error?
+\item {\it In line 39, page 4, by assuming $\epsilon(t) = 0$, you assume there is no measurement error?
 Why is section 2.1 called measurement error if the error is 0?
@@
 
-\item {\it s there a difference between the mean and covariance functions for $y = 0$ and $y = 1$ especially if points in $D_i$ and $T_i$ are spread roughly evenly on a grid even if the sets are disjoint?
+\item {\it Is there a difference between the mean and covariance functions for $y = 0$ and $y = 1$ especially if points in $D_i$ and $T_i$ are spread roughly evenly on a grid even if the sets are disjoint?
 For example, pick a grid of equidistant points and number them from $1-10$.
@@
 The subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$ and
+
+We also now clarify that the missing data imputation from XXXX requires distinct covariance functions and thus we keep as is.
+We do think one could marginalize across the two if the mean and covariance functions aren't expected to differ greatly and state as such in Section XX (in blue).
 \vspace{5mm}

### 2023-02-13 write-up/fda-recurrent.tex
testing
writeup and responses all in there now. making progress

@@
 \begin{newmaterial}
-The use of smart devices and other wearables to passively monitor patients has led to a rapid increase in the number of studies with high-frequency sensor streams that can be conceptualized as studies measurements are functions.
+The use of wearable devices to passively monitor patients has led to a rapid increase in the number of studies with high-frequency sensors that can be conceptualized as studies where measurements are functions.
 The development of such applications have been accompanied by intense methodological development in regression models with functional covariates~\citep{James2002,JamesSilverman2005,muller2005,Ramsay2005,kokoszka2017,doi:10.1198/jasa.2009.tm08564,doi:10.1198/016214507000000527}.
 
-Recently, modelling time-to-event data with functional covariates has received a fair bit of attention.
-The functional linear Cox regression model (FLCRM) considers the association between a time-to-event and a set of functional~$\{ X(s) \}$ and scalar predictors~$Z \in \mathbb{R}^p$~\citep{https://doi.org/10.1111/biom.12748}.
-In this setting, the functional and scalar predictors are observed at baseline and the hazard function satisfies the proportional hazards assumption and involves a non-parametric baseline hazard with the exponential adjustment including both a functional linear model~$\int_S X(s) \beta(s)ds$ and a linear model for the baseline scalar predictors~$Z^\top \gamma$.
-Since the linear assumption may be too restrictive,~\cite{doi:10.1080/10618600.2020.1853550} consider a more flexible additive functional Cox model for multivariate functional predictors in which the hazard depends on $\int  F \{s, h_{is}[X_i (s)]\} ds$ where $F$ is an unspecified bivariate twice differentiable function.
-
-In our setting, the high frequency sensor process is an outcome and must therefore be jointly modelled with the recurrent event process.
-Recently, \cite{Li2022} proposed a functional approach to joint modelling of multivariate longitudinal biomarkers and a time-to-event outcome~\cite{Tsiatis2004,Rizopoulos2010}.
+Recently, modeling time-to-event data with functional covariates has received a fair bit of attention.
+The functional linear Cox regression model (FLCRM) considers the association between a time-to-event and a set of functional and scalar predictors~\citep{https://doi.org/10.1111/biom.12748}.
+In this setting, the functional and scalar predictors are observed at baseline and the hazard function satisfies the proportional hazards assumption and involves a non-parametric baseline hazard with the exponential adjustment including both a functional linear model and a linear model for the baseline scalar predictors.
+Since the linear assumption may be too restrictive,~\cite{doi:10.1080/10618600.2020.1853550} consider a more flexible additive functional Cox model for multivariate functional predictors in which the hazard depends on an unspecified bivariate twice differentiable function.
+
+In our setting, the high frequency sensor process is an outcome and must therefore be jointly modeled with the recurrent event process.
+Recently, \cite{Li2022} proposed a functional approach to joint modeling of multivariate longitudinal biomarkers and a time-to-event outcome~\citep{Tsiatis2004,Rizopoulos2010}.
 Here, as in traditional joint models, the longitudinal biomarker is observed at only a few observation times (typically on the order of tens of observations) and is modelled via a functional principal components analysis.
@@
 A similar joint model was proposed in~\cite{Dong2021}.
-Inference proceeds by Monte Carlo EM which can require very large Monte Carlo sample size to ensure good performance, and na{\"i}ve implementation may not recover EM's ascent property~\cite{Caffo2005}.
-
-In this paper, we also consider joint modelling of recurrent events and a high frequency sensor process.
-In contrast to traditional longitudinal biomarkers, sensor processes are observed multiple times per second leading to thousand and/or millions of observations per individual rather.
+Inference proceeds by Monte Carlo EM which can require very large Monte Carlo sample size to ensure good performance, and na{\"i}ve implementation may not recover EM's ascent property~\citep{Caffo2005}.
+
+In this paper, we also consider joint modeling of recurrent events and a high frequency sensor process.
+In contrast to traditional longitudinal biomarkers, sensor processes are observed multiple times per second leading to thousand and/or millions of observations per individual.
 To address this increase in scale, a design-based perspective is taken in which we subsample non-event times to avoid some of the computationally demanding aspects of joint models.
@@
 Finally, traditional joint modeling often consider the hazard function to depend on the current or cumulative value of the health process~\citep{Li2022,Rizopoulos2010}.
-For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history; to do so, we use recent history functional linear models~\cite{KIM20111554}.
+For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history and use recent history functional linear models~\cite{KIM20111554} to to account for this local dependence.
 We combine these models with advances in longitudinal functional data analysis~\citep{Staicu2010} and penalized functional regression~\citep{Goldsmith2015} to perform scalable inference using off-the-shelf software.
@@
 
-The original contributions of this paper are three-fold.
-In Section~\ref{section:modeloverview}, we define our joint model using historical functional linear models, and demonstrate how techniques from design-based inference can be used to circumvent computational challenges of joint modelling.
-In Section~\ref{section:longfpc}, we discuss longitudinal functional principal components analysis and demonstrate in Lemma~\ref{lemma:logistic} that the resulting approximate score equations are equivalent to score functions for logistic regression with binary response and an offset related to the subsampling rate meaning our proposed approach can be fit using off-the-shelf software.
+The main contributions of this paper are as follows:
+We define a joint model using historical functional linear models and demonstrate how techniques from design-based inference can be used to circumvent computational challenges of joint modeling in Section~\ref{section:modeloverview}.
+We then discuss longitudinal functional principal components analysis and demonstrate in Lemma~\ref{lemma:logistic} that the resulting approximate score equations are equivalent to score functions for logistic regression with binary response and an offset related to the subsampling rate meaning our proposed approach can be fit using off-the-shelf software.
 An asymptotic analysis is presented in Section~\ref{section:theory}, with an accompanying discussion of the novel computational versus statistical efficiency tradeoff in Section~\ref{section:tradeoff}.
-Section~\ref{section:simstudy} presents a simulation study to demonstrate the impact of subsampling rates, the window length parameter, and a computational comparison to existing methods.
-Section~\ref{eq:jm} presents a multivariate extension as well as a method to handle missing data.
+A simulation study in Section~\ref{section:simstudy} demonstrates the impact of subsampling rates, the window length parameter, and a computational comparison to existing methods.
+A multivariate extension and method to handle missing data are presented in Section~\ref{section:extensions}.
 We end by illustrating our approach using data from a digital monitoring study of suicidal ideation in Section~\ref{section:example}.
@@
 That is, the risk model now depends on an integrated past history of the time-varying covariate which leads to severe increase in computational complexity.
+
+\begin{rmk}[Connection to design-based estimation]
+The approximate score equations given by~\eqref{eq:approxscore} arise in design-based inference of point processes.
+Design-based inference is common for spatial point processes~\citep{Waagepetersen2008} where the spatial varying covariate is observed at a random sample of locations.
+It is common in mobile health where ecological momentary assessments~\citep{Rathbun2012,Rathbun2016} are used to randomly sample individuals at various time-points to assess their emotional state.
+In the current setting, we leverage these ideas to form a subsampling protocol that can substantially reduce computationally complexity.
+Therefore, the purpose is quite different.
+Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.
+\end{rmk}
+
+\begin{rmk}[Assessment of computational benefit]
+Cool we have speed ups.
+\end{rmk}
 
@@
 
-\begin{rmk}[Connection to design-based estimation]
-The approximate score equations given by~\eqref{eq:approxscore} arise in design-based inference of point processes.
-Design-based inference is common for spatial point processes~\citep{Waagepetersen2008} where the spatial varying covariate is observed at a random sample of locations.
-It is common in mobile health where ecological momentary assessments~\citep{Rathbun2012,Rathbun2016} are used to randomly sample individuals at various time-points to assess their emotional state.
-In the current setting, we leverage these ideas to form a subsampling protocol that can substantially reduce computationally complexity.
-Therefore, the purpose is quite different.
-Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.
-\end{rmk}
-
 \section{Longitudinal functional principal components within
@@
 \[
-\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (T,s), (T,s^\prime) ) f_y(T) dt.
+\Sigma_y (s, s^\prime) = \int_{0}^\tau c_y( (T,s), (T,s^\prime) ) f_y(T) dT.
 \]
@@
 
-The next step of our method is modelling~$\beta(s)$.
+The next step of our method is modeling~$\beta(s)$.
 Here, we leverage ideas from the penalized spline literature~\citep{Ruppert2003, Wood2006book}.
@@
 Second, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
-Joint modelling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
+Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
 

### 2023-02-14 letters/response-ref-1-v1.tex
updated letter.  i think we are on track.  Need to start sims by end of week

@@
 \vspace{5mm}
-We believe the referee is referring to the log-likelihood as its related to the event process. DO WE NEED NO MEASUREMENT ERROR IF WE DO IT THIS WAY? OTHERWISE WE WILL HAVE THE joint likelihood issue.
+We believe the referee is referring to the log-likelihood related to the event process.
+Conditional on the sensor process over the observation window $[0,\tau_i]$, i.e.,~$\{ x_i (t) \}_{t=0}^{\tau_i}$, we define the conditional intensity according to equation (2).
+Then the probability of the observed event times per person follows from $\exp \left( - H_i (\tau_i ; \theta) \right) \prod_{j=1}^{k_i} h \left(T_{i,j};\theta \right)$.
+Thus the log-likelihood of the even process is given by $L_n (\theta)$.
 \vspace{5mm}
@@
 \vspace{5mm}
-The key point raised on page 4 (no Section XX), is that if $\epsilon(t)$ is not equivalently zero, your model assumes risk of an event at time $t$ depends on both the current and future measurements of the process ${\bf x}_i$.
+The key point raised on page 4 (now in Appendix XX), is that if $\epsilon(t)$ is not equivalently zero, your model assumes risk of an event at time $t$ depends on both the current and future measurements of the process ${\bf x}_i$.
 We argue this is not sensible since current risk should not depend on future risk and thus set $\epsilon(t) = 0$.
@@
 It was used in equation (1) to refer to all the parameters underlying the model to distinguish it from the functional parameter~$\beta (s)$.
-We do XXXX to make this clearer.
+We have added a notation glossary in Appendix XX, and re-iterate its definition in Section 3 to make this clearer.
 \vspace{5mm}
@@
 \vspace{5mm}
+We have added Remark XX and Section XX in the simulation study to give additional details and empirical evidence.
+First, there are two areas where we have computational improvements.
+First, in the inner integral, we have
+
+Second, even if we have computed $\int_{t-\Delta} x(t-s)\beta(s) ds$
 We explain that if we were to compute it would have to be as a function of $\beta(s)$.

### 2023-02-16 letters/response-ref-1-v1.tex
have plan, now coding time for weeks and weeks

@@
 The event times~$T_i$ depend heavily on the event intensity function~$h_i (t;\theta)$ and will not be guaranteed to be uniformly spread across points in $[0,\tau]$.
-The subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$ and
+The subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$; under the assumption that $\pi_i (t) = \lambda$, we can view this as first drawing the number of non-event times from a Poisson distribution, i.e., $| D_i | \sim Poisson(\lambda)$, and then independently selecting $|D_i|$ non-event times uniformly at random over the observation window~$[0, \tau_i]$.
+Given the potential difference between the distributions, we choose to model the mean and covariance functions separately to ensure good approximation when applying th espectral decomposition of the estimated covariance function.
+\vspace{5mm}
 
-We also now clarify that the missing data imputation from XXXX requires distinct covariance functions and thus we keep as is.
-We do think one could marginalize across the two if the mean and covariance functions aren't expected to differ greatly and state as such in Section XX (in blue).
+Another reason for treating these two distinctly is that the missing data imputation from XXXX requires distinct covariance functions.
+We do think one could marginalize across the two if the mean and covariance functions are not expected to differ greatly and state as such in Section XX (in blue) where we discuss the new Figure XX of the mean and covariance functions for $y=0$ and $y=1$.
 \vspace{5mm}
@@
 In Section 3.1, we started by defining the double-indexed $X(t,s)$ to be equal to $X(t-s)$ in order to make the notation a bit cleaner but we see that this has added confusion.
-We not add a remark to make this more clear.
-The reason for the notation was that writing $X(t,s)$ allows us to write the integrals over $0$ to $\Delta$ rather than having to constantly index integrals by $t$.
+We now add a remark to clarify this change in notation.
+The reason for the change in notation was that writing $X(t,s)$ allows us to write the integrals over $0$ to $\Delta$ rather than having to constantly index integrals by $t$.
 Moreover, we thought it helps the reader to understand we are saying ``the process $s$ units prior to time~$t$.''
@@
 \vspace{5mm}
-We have shown that the functional process satisfies these moment conditions under boundedenss which is
+We have added Remark XX to provide a simple example where these assumptions are satisfied.
+Specifically, if the functional process is bounded, i.e., $|X(t,s)| < M$ and the other component is also bounded then we are guaranteed to have these assumptions hold.
+This is likely to be true in most practical settings and thus demonstrates the reasonableness of our assumptions for applied settings.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXxXX
+We have now clarified that the convergence is in law as the number of individuals goes to infinity.  Proofs are also clarified accordingly.
 \vspace{5mm}
@@
 \vspace{5mm}
-We have now included the statement of how we define optimality in the main paper to make the theorem statement clear.  Proof is kept in the supplementary materials for conciseness.
+We have now included the statement of how we define optimality in the main paper to make the theorem statement clear.
+Specifically, we view weights $w^\star$ as optimal if the difference between the asympotic variance~$V (\theta_0; w^\star)$ and the asymptotic variance under any other choice of weights~$w$, $V(\theta_0; w)$ is postivie semi-definite.
+The proof is kept in the supplementary materials for conciseness.
 \vspace{5mm}
@@
 \vspace{5mm}
-We discuss the penalization due to the potential complexity of the.
-The confidence bands are based on Lemma 3.5.
-We make this clear in the revision.
-The issues are still thre and we keep the discussion.
-We keep as 2 separate sections since they cover different issues.
+We have moved the discussion of confidence bands to follow Lemma 3.5 (now 4.5) given that it is directly a consequence of the asymptotic result.
+We then  discuss penalization in a subsequent section which is employed due to the potential complexity induced by choosing $K_b$ large enough for the assumption $\beta (t) =\phi (t) b^\star$ to hold.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXX
+We really appreciate this question which has helped clarify an important element of the paper.
+To address this question, we have added Section XX to evaluate the accuracy and time complexity.
+We recap some of our discussion here for convenience.
+\vspace{5mm}
+
+We observe each high frequency sensor process at a specific Hertz (Hz, number of observations per second) and so the time complexity will depend on this unit of measurement.
+Take EDA, for example, which is measured at 4Hz.
+Then a time complexity analysis of the entire statistical procedure must capture three elements: (1) computing the marginal covariance, (2) computing $\int_{0}^\Delta x_i (t-s) \beta(s) ds$, and (3) computing the cumulative hazard.
+Step (1) requires computing a sum $O(n E\left[|D_i| \right])$ computations under subsampling; a complete-case analysis sets $|D_i|$ equal to the maximum number of observation times.
+For a 4Hz sensor, if $\tau_i$ is measured in minutes then the total number of observations is $4 \times 60 \times \tau_i$.
+Step (2) requires computation of $\hat c_{i,k}^{(y)} (t)$ which has time-complexity $O(n E \left[ |D_i| + |T_i| \right] K_x)$.
+For a 4Hz sensor, a complete case again has approximatley $4 \times 60 \times \tau_i$ non-event times.
+Step (3) is then well approximated by the trapezoidal rule on the uniform grid of non-event times.
+This has time-complexity $O(4 \times 60 \times \tau_i)$ compared to the subsampling time-complexity of $O(E \left[ |D_i|  + |T_i| \right]$.
+Thus each step in this procedure is magnified, and this only reflects computations for a single evaluation of the log-likelihood.
+\vspace{5mm}
+
+To address the time-complexity of maximizing the log-likelihood, we use Lemma 3.1 (now 4.1).
+The complexity of logistic regression with $n$ observations and covariates of dimension $d$ is $O(nd)$; so we can view maximizing the true likelihood as subsampling with a very high rate so that the design-unbiased score functions have minimal approximation error in comparison to to directly maximizing the likelihood function.
+Under this assumption, at the final stage we have the relative complexity of $(4 \times 60 \times \tau_i) \cdot n \cdot d$ compared to the expected time-complexity under subsampling with rate $c$ per minute of $(c \times \tau_i) \cdot n \cdot d$.
+In practice, this can lead to large differences in run time; specifically, this means sampling every minute rather than every half hour will be approximately $2^5$ times as long (from 5 minutes to 2.8 hours), while sampling every second will take $2^11$ times as long (from 5 minutes to 180 hours).
+In our first example, the run time was 317 seconds (approximately 5 minutes).
+Sampling every minute or every second would increase run time to approximately 2.8 hours and 180 hours respectively.
+In our second example, the run time of 36 seconds would increase to 19 minutes and 20 hours respectively.
+In both instances, the relative efficiency gain would be negligible suggesting a huge computational increase for minimal relative information gain.
+We now make this point abundantly clear by amending Section XX to include time complexity discussion and Section XX to include the exact calculations above.
+
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXX
-There is no obvious way to compute bias since the model will be incorrectly specified.
-We do show the computation trade-off and that computing the cumulative hazard is still computationally demanding; therefore, after sampling non-event times, the added computation of the interior integral is not very expensive and avoids model misspecification.
+The above calculations hold for even the concurrent model, so that approach is still computationally expensive.
+To show this while ensuring no bias, we have added a simulation in Section XX where we show the computation trade-off and applying the concurrent model without subsampling.
+We find the approach computationally demanding with minimal information gain.
 \vspace{5mm}
@@
 turning up. Add longer history to both the Figures.
+\vspace{5mm}
+
+We have investigated this issue and there is a slight uptick in EDA but not in ACC; however, we are concerned that showing a longer history may not be helpful given our focus on a potentially shorter $\Delta$ based on the data analysis.
+Therefore, we keep the plot as is, and note that there is a spike at 30-minutes for EDA while emphasizing that most of the signal appears in the 15-minute range.
+This helps motivate our data adaptive choice of $\Delta$ in Section 7.1 which confirms that $\Delta = 15$ is a more appropriate choice.
+\vspace{5mm}
 
@@
 \vspace{5mm}
-We agree.  We re-run the analysis with a window length of 15 minutes as well and move the current analysis as a sensitivity to the supplementary materials.
+We agree.
+We have re-run the analysis with a window length of 15 minutes.
+A comparison of AIC provides empirical evidence that the smaller $\Delta$ was an appropriate choice.
+We now present $\Delta = 15$ in the main paper and move the current analysis to be a sensitivity analysis presented in the supplementary materials.
 \vspace{5mm}
@@
 \vspace{5mm}
-As this was illustrative we only included time-of-day.
+As the case study was meant to be illustrative, we only included time-of-day as a covariate.
+We now include the number of button presses in the prior 2 hours as another feature to control for the heterogeneity seen in the .
+These features are described in Section XX.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXXX.
-We added Section XX to describe the related work and connect our methodology to the existing literature.
+We have added a detailed discussion of related work in Section 1.1, which includes the specific suggestions on relevant works provided by the second referee; specifically, we discuss recent modeling of time-to-event data with functional covariates as well as joint modelling from a functional regression perspective.
+We contrast these approaches with our design-based approach which becomes necessary when considering high-frequency sensor processes rather than traditional longitudinal studies as the number of measurement occasions increase from tens to thousands/millions of times per individual.
+We connect and contrast our use of subsampling with the general use of subsampling in handling massive data and its recent use in massive data with functional predictors.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXXX.
+We now state that our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
+This motivates Section XX which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
 We added Section XX to describe the related work and connect our methodology to the existing literature.
@@
 \vspace{5mm}
-XXXXXX
+We agree.
+As with other tuning parameters, we take an AIC approach to selection of $\Delta$ in order to be data-adaptive.
+We study the performance of this criteria as part of the simulation study in Section XX and then apply it in the case study to determine the appropriate window-length.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXXX
-\vspace{5mm}
-
-\item {\it While in Section 3.6 the paper discusses penalized functional regression to reduce overfitting, it is unclear if the paper actually adopts it in simulation studies and application.
-It is of interest how the smoothing parameter is selected.
-My concern is that if no penalty is employed, with more than 30 basis functions, overfitting can be severe and much wider confidence bands are often obtained}
-
-\vspace{5mm}
-XXXXXX
+We now clarify that throughout the simulation studies and application we employ the penalization procedure to reduce overfitting.
+We have found that without careful tuning of the penalty parameter, the overfitting can be severe as stated by the referee.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXXX
+To address this and Referee \#1's comment \#X, we have added Remark XX to clarify settings where Assumption 3.4 will hold.
+Specifically, if the functional process is bounded, i.e., $|X(t,s)| < M$ and the other component is also bounded then we are guaranteed to have these assumptions hold.
+This is likely to be true in most practical settings and thus demonstrates the reasonableness of our assumptions for applied settings.
 \vspace{5mm}
@@
 \vspace{5mm}
-XXXXXXX
+By design, $\beta(s) = 0$ for $s > \Delta$.
+In practical settings we do not anticipate~$\beta(s)$ to be non-smooth and thus chose two scenarios where $\beta$ decays smoothly to 0.
+To address this, we have added a third scenario which has more rapid decay to zero; specifically, the coefficient function mimics taking a temporal difference over two average windows.
+To address the other comment, we have adjusted Table 2 to also report the relative values.
+This allows the reader to understand the MISE increases slowly and is mainly a consequence of the variance increases from reducing the subsampling rate.
 \vspace{5mm}
@@
 \vspace{5mm}
-The model is under conditional independence $\eta$.
+We have decided that the point we were trying to raise is non-essential to the current manuscript, so we have removed the discussion of \emph{independent evolution} and its implication that $\varepsilon (t) = 0$ for all $t >0$.
+The key issue we were trying to clarify is that the traditional joint model formulation in which the observed health process and time-to-event are conditionally independent given the latent process implies that the hazard function conditional on the entire observed health process depends on values after the current time.
+In the case study, for example, this implies that the expected risk of a button press given the entire sensor process will depend on future activity index and electrodermal activity.
+This holds unless we assume the measurement error is set to zero.
+The referee is correct that we will only compute conditional probabilities given the historical functional predictor; however, the likelihood factorization means that we implicitly are defining a conditional intensity given the entire functional predictor and the traditional approach has this undesirable dependence.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
+Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
 \vspace{5mm}
@@
 \vspace{5mm}
-Separate models in FACE paper so we can use the ; however, it makes minimial difference.  So we cite this.
+Since the event times~$T_i$ depend heavily on the event intensity function~$h_i (t;\theta)$ while the subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$, we chose to model the mean and covariance functions separately to ensure good approximation when applying the spectral decomposition of the estimated covariance function.
+For example
+\vspace{5mm}
+
+Another reason for treating these two distinctly is that the missing data imputation from XXXX requires distinct covariance functions.
+We do think one could marginalize across the two if the mean and covariance functions are not expected to differ greatly and state as such in Section XX (in blue) where we discuss the new Figure XX of the mean and covariance functions for $y=0$ and $y=1$.
 \vspace{5mm}

### 2023-02-16 write-up/fda-recurrent.tex
updated letter.  i think we are on track.  Need to start sims by end of week
have plan, now coding time for weeks and weeks

@@
 % display breaks
-\newenvironment{newmaterial}{\par\color{blue}}{\par}
+\newenvironment{newmaterial}{\color{blue}}{\par}
 
@@
 An estimator is design-unbiased if its expectation is equal to that parameter under the probability distribution induced by the sampling design~\citep{Cassel1977}.
-Let~$\bfD_i \subset [0,t_i]$ denote the random set of subsampled points.
+Let~$\bfD_i \subset [0,\textcolor{blue}{\tau_i}]$ denote the random set of subsampled points.
 Note, by construction, this random set is distinct from the set of event times with probabilty one, i.e.,~$\pr( \bfT_i \cap \bfD_i = \emptyset) = 1$.
@@
 \hat{U}_n (\theta) = \sum_{i=1}^n \left[ \sum_{u \in \bfT_i} w_i(u; \theta)
-  \frac{h^{(1)} (u; \theta)}{ h ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
+  \frac{h^{(1)}_i (u; \theta)}{ h_i ( u; \theta)}  - \sum_{u \in \bfD_i} w_i(u;
   \theta) \frac{h_i^{(1)} (u; \theta)}{ \pi_i (u) } \right].
@@
 This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
-Not only this, Lemma~\ref{prop:optimal} shows weights~\eqref{eq:waage_weights} are optimal within a particular class of weighted estimating equations.
-
+\begin{newmaterial}
+Based on this, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
+That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ dimensional matrix obtained by plugging in the estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
+We acknowledge an important limitations of confidence intervals obtained via this approach.
+Specifically, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
+Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
+
+Lemma~\ref{prop:optimal} shows~\eqref{eq:waage_weights} is optimal within a class of weighted estimating equations.
+The result ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
+Here, weights $w^\star$ are considered \emph{optimal} if the difference between the asympotic variance~$V (\theta_0; w^\star)$ and the asymptotic variance under any other choice of weights~$w$, $V(\theta_0; w)$ is positive semi-definite, i.e., any linear contrast has smaller asymptotic variance under weight $w^\star$ than under weight~$w$.
+\end{newmaterial}
 \begin{lemma} \normalfont
@@
 
-\noindent Proposition~\ref{prop:optimal} ensures the only loss of statistical efficiency is due to subsampling and not using a suboptimal estimation procedure given subsampling.
+\noindent
 
@@
 Combining the penalized spline formulation with Lemma~\ref{lemma:logistic} establishes a connection between our approximate score equations and solving a generalized mixed effects logistic regression with offset.
-
-\subsection{Confidence intervals for~$\beta(t)$}
-
-Due to the connection with generalized mixed effects models, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
-That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ dimensional matrix obtained by plugging in the estimates of variance components into the formula for the variance of~$\hat b$, then the standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0) = \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 ) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-Then the approximate 95\% confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96 \sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.
-
-We acknowledge two important limitations of confidence intervals obtained via this approach.
-First, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is oversmoothed.
-Second, we ignore the variability inherent in the longitudinal functional principal component analysis; that is, our estimates ignore the variability in estimation of eigenfunctions~$\hat \psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
-Joint modeling could be considered as in~\cite{Crainiceanu2010}, however, this is beyond the scope of this article.
+Given the connection with generalized mixed effects models, the inferential machinery to obtain variance-covariance estimates .
+In this context, we acknowledge another limitation.
+Specificaly, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is oversmoothed.
 

### 2023-02-16 simulations/delta_incorrect/sim_delta_functions.R
need to change to do delta in a sim not across

@@ -116,7 +116,7 @@ runglmnet <- function(sampling_rate, dataset, w, Basis, epsilon = 0.0001) {
   p.fac = rep(1, ncol(w))
   p.fac[1:4] = 0 #no penalty on the first 4 variables
   lambda_max <- 1/n.tmp
-  epsilon <- .0001
+  epsilon <- .001
   K <- 100
   lambdapath <- round(exp(seq(log(lambda_max), log(lambda_max*epsilon), 
                               length.out = K)), digits = 10)

### 2023-03-08 simulations/basic_simulation/simulation_summary.R
Can summarize basic table more efficienctly

@@ -1,11 +1,18 @@
 source("./basic_simulation_functions.R")
-setting = "sine"
+setting = "exponential"
+num_simulations = 500
+
+## BUILD THE DATASET
+df = rep(0,0)
+for (iter in 1:num_simulations) {
+  temp_df = data.frame(read.table(paste("./output_csv/",setting,"_simdelta_results_wl_44_arrayid_",iter,".csv", sep = ""), sep = ",", header = T))
+  if( nrow(temp_df)!= 5) {print(paste("Iteration", iter, "has wrong number of rows"))}
+  df = rbind(df, temp_df)
+}  
 
 if(setting == "sine") {
-  df = data.frame(read.table("./output_csv/sine_results.csv", sep = ",", header = T))
   beta_truth = 100*sin(1:44/44*2*pi-pi/2)
 } else if (setting == "exponential") {
-  df = data.frame(read.table("./output_csv/exponential_results.csv", sep = ",", header = T))
   beta_truth = 30*exp(0.2*1:44)/mean(exp(0.2*1:44))
 }
 
@@ -71,10 +78,7 @@ for (id in unique_ids) {
   }
 }
 
-table = rbind(subsample_variance/true_variance, variance/true_variance, sqbias/true_variance, mise/true_variance, avg_runtime)
+table = rbind(unique_rates, subsample_variance/true_variance, variance/true_variance, sqbias/true_variance, mise/true_variance, avg_runtime)
 
 round(table, 5)
-
-variance/variance[1]
-sqbias/sqbias[1]
 mise/mise[1]

### 2023-03-08 simulations/delta_incorrect/sim_delta.R
fixed deviance

@@ -1,4 +1,3 @@
-
 source("./sim_delta_functions.R")
 if(!require("glmnet")){install.packages('glmnet')} else{library('glmnet')}
 
@@ -49,8 +48,8 @@ for(window_length in seq_window_lengths) {
   K_x = min(35, window_length)  # Pick first 35 eigen-vectors or window_length
   cumsum(eig_Sigma$values)[K_x]/sum(eig_Sigma$values) # Explains most of the variation
   
-  arrayid = Sys.getenv("SLURM_ARRAY_TASK_ID")
-  # arrayid = 10
+  # arrayid = Sys.getenv("SLURM_ARRAY_TASK_ID")
+  arrayid = 1
   print(paste("Current ARRAY TASK ID", arrayid))
   #your array of seeds in x
   allseeds = readRDS("sim_delta_seeds.RDS")
@@ -82,7 +81,7 @@ for(window_length in seq_window_lengths) {
     subdataset = subsample_dataset(dataset, thinning_rate = rates)
     intermediate_step = construct_J(times, eig_Sigma, subdataset, window_length)
     output = runglmnet(max_sampling_rate*rates, subdataset, intermediate_step$w, intermediate_step$Basis, epsilon = 0.0001)
-    results = matrix(c(arrayid, output$runtime, output$dev, rates, agg_summary, output$beta), nrow = 1)
+    results = matrix(c(arrayid, output$runtime, min(output$dev), rates, agg_summary, output$beta), nrow = 1)
     colnames(results) = col_names
     if(!file.exists(paste("./output_csv/", setting, "_simdelta_results_wl_",window_length,"_arrayid_", arrayid, ".csv", sep = ""))) {
       write.table(results, file = paste("./output_csv/", setting, "_simdelta_results_wl_",window_length,"_arrayid_", arrayid, ".csv", sep = ""), row.names = F, col.names = T, append = T, sep = ",")

### 2023-03-08 write-up/fda-recurrent.tex
fixing table for basic simulation, onto delta and checking writeup matches

@@
 \begin{table}[!th]
-\begin{tabular}{c | p{2cm} p{2cm} p{2cm} p{2cm}}
-& \multicolumn{4}{c}{Sampling rate} \\ \cline{2-5}
- & 0.5 & 1 & 2 & 4 \\ \hline
-Subsampling variance & - & $9.0 \times 10^{-7}$ & $2.9 \times 10^{-6}$ & $5.8 \times 10^{-6}$  \\
-Variance & $7.9 \times 10^{-3}$ & $9.2 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ \\
-Squared Bias & $9.8 \times 10^{-2}$ & $9.5 \times 10^{-2}$ & $9.3 \times 10^{-2}$ & $9.2 \times 10^{-2}$   \\
-MISE & $1.1 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ \\ \hline
-Avg. runtime (secs) & 317 & 177 & 104 & 68 \\ \hline
-Subsampling variance & - & $2.9 \times 10^{-6}$ & $8.9 \times 10^{-6}$ & $2.0 \times 10^{-5}$  \\
-Variance & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.1 \times 10^{-2}$ \\
-Squared Bias & $6.9 \times 10^{-2}$ & $6.5 \times 10^{-2}$ & $6.0 \times 10^{-2}$ & $5.4 \times 10^{-2}$   \\
-MISE & $7.9 \times 10^{-2}$ & $7.9 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
-Avg. runtime (secs) & 36 & 23 & 15 & 12 \\ \hline
+\begin{tabular}{c | p{2cm} p{2cm} p{2cm} p{2cm} p{2cm}}
+& \multicolumn{4}{c}{Sampling rate (per)} \\ \cline{2-6}
+ & X1 & X2 & X3 & X4 & X5 \\ \hline
+SubS. Variance & - & $1.2 \times 10^{-5}$ & $2.0 \times 10^{-5}$ & $3.5 \times 10^{-5}$ & $8.6 \times 10^{-5}$  \\
+Variance & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.1 \times 10^{-2}$ & $2.7 \times 10^{-2}$ & $3.8 \times 10^{-2}$\\
+Squared Bias & $6.3 \times 10^{-2}$ & $6.0 \times 10^{-2}$ & $6.1 \times 10^{-2}$ & $6.1 \times 10^{-2}$ & $6.7 \times 10^{-2}$   \\
+MISE & $7.6 \times 10^{-2}$  & $7.5 \times 10^{-2}$  & $8.2 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $1.1 \times 10^{-1}$ \\
+RMISE & -  & 0.99  & 1.08 & 1.16 & 1.39 \\ \hline
+Avg. runtime (secs) & 356 & 108 & 51 & 24 & 17 \\ \hline
+SubS. variance & - & $2.7 \times 10^{-5}$ & $5.1 \times 10^{-5}$ & $5.7 \times 10^{-5}$ & $7.5 \times 10^{-5}$  \\
+Variance & $3.7 \times 10^{-2}$ & $2.7 \times 10^{-2}$ & $2.8 \times 10^{-2}$ & $3.3 \times 10^{-2}$ &  $3.8 \times 10^{-2}$\\
+Squared Bias & $3.5 \times 10^{-2}$ & $3.6 \times 10^{-2}$ & $3.4 \times 10^{-2}$ & $3.3 \times 10^{-2}$ & $3.1 \times 10^{-2}$ \\
+MISE & $7.2 \times 10^{-2}$ & $6.3 \times 10^{-2}$ & $6.2 \times 10^{-2}$ & $6.6 \times 10^{-2}$ & $6.9 \times 10^{-2}$ \\
+RMISE & -  & 0.87  & 0.86 & 0.92 & 0.96 \\ \hline
+Avg. runtime (secs) & 1084 & 208 & 93 & 47 & 30 \\ \hline
 \end{tabular}
-\caption{Mean-integrated squared error, varaince, squared bias, and subsampling variance for $\beta(s)$ given by (1) and (2) respectively.}
+\caption{Mean-integrated squared error, variance, squared bias, and subsampling (SubS) variance for $\beta(s)$ given by (1) and (2) respectively.}
 \label{tab:mise}

### 2023-03-08 simulations/basic_simulation/simulation_summary.R

@@ -1,5 +1,5 @@
 source("./basic_simulation_functions.R")
-setting = "exponential"
+setting = "sine"
 num_simulations = 500
 
 ## BUILD THE DATASET
@@ -78,7 +78,7 @@ for (id in unique_ids) {
   }
 }
 
-table = rbind(unique_rates, subsample_variance/true_variance, variance/true_variance, sqbias/true_variance, mise/true_variance, avg_runtime)
+table = rbind(unique_rates, subsample_variance/true_variance, variance/true_variance, sqbias/true_variance, mise/true_variance, mise/mise[1], round(avg_runtime,0))
 
 round(table, 5)
 mise/mise[1]

### 2023-03-27 write-up/fda-recurrent-supplement.tex
Fixed all simulation tables. now onto adding the aic check on delta and then computation

@@
 \begin{tabular}{c | c c c c c}
-& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
-$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
-\multirow{3}{*}{26} & MISE & $5.9 \times 10^{-2}$ & $6.1 \times 10^{-2}$ & $6.6 \times 10^{-2}$ & $7.4 \times 10^{-2}$ \\
- & Variance & $1.3 \times 10^{-2}$ & $1.6 \times 10^{-2}$ & $12.4 \times 10^{-2}$ & $3.6 \times 10^{-2}$ \\
-  & P-MISE & $7.1 \times 10^{-2}$ & $7.3 \times 10^{-2}$ & $7.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{29} & MISE & $6.9 \times 10^{-2}$ & $7.0 \times 10^{-2}$ & $7.1 \times 10^{-2}$ & $7.8 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.4 \times 10^{-2}$ \\
-  & P-MISE & $7.5 \times 10^{-2}$ & $7.5 \times 10^{-2}$ & $7.7 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{32} & MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $1.9 \times 10^{-2}$ & $3.3 \times 10^{-2}$ \\
-  & P-MISE & $8.1 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.7 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{35} & MISE & $9.0 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $8.8 \times 10^{-2}$ & $9.2 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $2.2 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
-  & P-MISE & $8.5 \times 10^{-2}$ & $8.2 \times 10^{-2}$ & $8.1 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
-\multirow{3}{*}{37} & MISE & $10.3 \times 10^{-2}$ & $10.1 \times 10^{-2}$ & $9.9 \times 10^{-2}$ & $10.1 \times 10^{-2}$ \\
- & Variance & $1.1 \times 10^{-2}$ & $1.4 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.0 \times 10^{-2}$ \\
-  & P-MISE & $9.4 \times 10^{-2}$ & $9.2 \times 10^{-2}$ & $9.0 \times 10^{-2}$ & $9.0 \times 10^{-2}$ \\ \hline
+& & \multicolumn{4}{c}{Sampling rate (per hour)} \\ \cline{3-6}
+$\Delta=$ & & 4.0 & 2.0 & 1.0 & 0.5 \\ \hline
+\multirow{3}{*}{26} & MISE & 0.070 & 0.075 & 0.095 & 0.0117 \\
+ & Variance & $2.1 \times 10^{-2}$ & $2.7 \times 10^{-2}$ & $3.9 \times 10^{-2}$ & $5.8 \times 10^{-2}$ \\
+  & P-MISE & 0.085 & 0.091 & 0.115 & 0.141 \\ \hline
+\multirow{3}{*}{29} & MISE & 0.072 & 0.076 & 0.093 & 0.111 \\
+ & Variance & $1.8 \times 10^{-2}$ & $2.3 \times 10^{-2}$ & $3.4 \times 10^{-2}$ & $5.0 \times 10^{-2}$ \\
+  & P-MISE & 0.079 & 0.083 & 0.101 & 0.121 \\ \hline
+\multirow{3}{*}{32} & MISE & 0.076 & 0.079 & 0.095 & 0.109 \\
+ & Variance & $1.5 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.2 \times 10^{-2}$ & $4.6 \times 10^{-2}$ \\
+  & P-MISE & 0.076 & 0.079 & 0.095 & 0.109 \\\hline
+\multirow{3}{*}{35} & MISE & 0.083 & 0.085 & 0.099 & 0.114 \\
+ & Variance & $1.4 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.2 \times 10^{-2}$ & $4.7 \times 10^{-2}$ \\
+  & P-MISE & 0.081 & 0.083 & 0.096, & 0.110 \\\hline
+\multirow{3}{*}{37} & MISE & 0.091 & 0.092 & 0.106 & 0.117 \\
+ & Variance & $1.5 \times 10^{-2}$ & $2.4 \times 10^{-2}$ & $3.4 \times 10^{-2}$ & $5.0 \times 10^{-2}$ \\
+  & P-MISE & 0.088 & 0.089 & 0.102 & 0.112 \\\hline
 \end{tabular}
-\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
+\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.
+% exponential EXAMPLE
+}
 \label{tab:appmise}

### 2023-03-27 write-up/fda-recurrent.tex
Fixed all simulation tables. now onto adding the aic check on delta and then computation

@@
 \begin{tabular}{c | p{2cm} p{2cm} p{2cm} p{2cm} p{2cm}}
-& \multicolumn{4}{c}{Sampling rate (per)} \\ \cline{2-6}
- & X1 & X2 & X3 & X4 & X5 \\ \hline
+& \multicolumn{4}{c}{Sampling rate (per hour)} \\ \cline{2-6}
+ & 12 & 4.0 & 2.0 & 1.0 & 0.5 \\ \hline
 SubS. Variance & - & $1.2 \times 10^{-5}$ & $2.0 \times 10^{-5}$ & $3.5 \times 10^{-5}$ & $8.6 \times 10^{-5}$  \\
@@
 \begin{tabular}{c | c c c c c}
-& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
-$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
-\multirow{3}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
- & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\
-  & P-MISE & 0.225 & 0.229 & 0.233 & 0.239 \\ \hline
-\multirow{3}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
- & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\
-  & P-MISE & 0.107 & 0.109 & 0.111 & 0.117 \\ \hline
-\multirow{3}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
- & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\
-  & P-MISE & 0.105 & 0.103 & 0.103 & 0.105 \\\hline
-\multirow{3}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
- & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\
-  & P-MISE & 0.273 & 0.270 & 0.269, & 0.269 \\\hline
-\multirow{3}{*}{37} & MISE & 0.589 & 0.588 & 0.589 & 0.591 \\
- & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\
-  & P-MISE & 0.473 & 0.470 & 0.469 & 0.470 \\\hline
+& & \multicolumn{4}{c}{Sampling rate (per hour)} \\ \cline{3-6}
+$\Delta=$ & & 4.0 & 2.0 & 1.0 & 0.5 \\ \hline
+\multirow{3}{*}{26} & MISE & 0.390 & 0.395 & 0.397 & 0.413 \\
+ & Variance & $5.0 \times 10^{-2}$ & $5.3 \times 10^{-2}$ & $5.5 \times 10^{-2}$ & $6.8 \times 10^{-2}$ \\
+  & P-MISE & 0.210 & 0.216 & 0.217 & 0.235 \\ \hline
+\multirow{3}{*}{29} & MISE & 0.220 & 0.225 & 0.224 & 0.238 \\
+ & Variance & $3.7 \times 10^{-2}$ & $4.4 \times 10^{-2}$ & $4.3 \times 10^{-2}$ & $5.5 \times 10^{-2}$ \\
+  & P-MISE & 0.084 & 0.090 & 0.088 & 0.103 \\ \hline
+\multirow{3}{*}{32} & MISE & 0.061 & 0.062 & 0.061 & 0.072 \\
+ & Variance & $2.7 \times 10^{-2}$ & $2.9 \times 10^{-2}$ & $3.0 \times 10^{-2}$ & $4.2 \times 10^{-2}$ \\
+  & P-MISE & 0.061 & 0.062 & 0.061 & 0.072 \\\hline
+\multirow{3}{*}{35} & MISE & 0.246 & 0.247 & 0.249 & 0.255 \\
+ & Variance & $1.9 \times 10^{-2}$ & $2.1 \times 10^{-2}$ & $2.3 \times 10^{-2}$ & $2.9 \times 10^{-2}$ \\
+  & P-MISE & 0.154 & 0.154 & 0.154, & 0.160 \\\hline
+\multirow{3}{*}{37} & MISE & 0.360 & 0.359 & 0.353 & 0.361 \\
+ & Variance & $2.6 \times 10^{-2}$ & $3.3 \times 10^{-2}$ & $3.7 \times 10^{-2}$ & $4.4 \times 10^{-2}$ \\
+  & P-MISE & 0.256 & 0.255 & 0.252 & 0.259 \\\hline
 \end{tabular}
-\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
+\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.
+% SINE EXAMPLE
+}
 \label{tab:mise_delta}
 \end{table}
+
+\subsection{Selection of bandwidth} \label{section:bandwidth}
 

### 2023-03-27 simulations/basic_simulation/simulation_summary.R

@@ -1,5 +1,5 @@
 source("./basic_simulation_functions.R")
-setting = "sine"
+setting = "exponential"
 num_simulations = 500
 
 ## BUILD THE DATASET
@@ -82,3 +82,4 @@ table = rbind(unique_rates, subsample_variance/true_variance, variance/true_vari
 
 round(table, 5)
 mise/mise[1]
+

### 2023-03-27 simulations/delta_incorrect/simulation_summary.R

@@ -1,5 +1,5 @@
 source("./sim_delta_functions.R")
-setting = "sine"
+setting = "exponential"
 num_simulations = 500
 seq_windowlengths = seq(36,52,by=4)
 
@@ -15,45 +15,52 @@ if(setting == "sine") {
   beta_truth = 30*exp(0.2*1:44)/mean(exp(0.2*1:44))
 }
 
-df = df36
-df = df[order(df$ids, df$rates),]
-
-M = 1000 # Num of time points
-times = seq(0,1, length.out = M) # Equally spaced timepoints in [0,1]
-gap = diff(times)[1] # What is assumed gap times
-
-true_variance = mise_calc(beta_truth, beta_truth*0)
-
-unique_rates = unique(df$rates)
-unique_rates = sort(unique_rates, decreasing = T)
-betameans = matrix(nrow = length(unique_rates), ncol = length(8:ncol(df)))
-for (i in 1:length(unique_rates)) {
-  betameans[i,] = colMeans(df[df$rates == unique_rates[i],8:ncol(df)], na.rm = TRUE)
-}
-
-## Simple visualization to show that means are same across subsamples
-## This is a good thing!  Means there is no bias due to subsampling 
-## The bias inherent in even most data case is due to small n and penalization
-par(mar = c(3,3,2,1)+0.1)
-plot(1:length(beta_truth), beta_truth, type = "l", axes = F, xlab = "Time", ylab = "")
-axis(side = 1, labels = F)
-axis(side = 2, labels = F)
-for(i in 1:length(unique_rates)) {
-  lines((length(beta_truth)-length(betameans[1,])+1):length(beta_truth), as.numeric(betameans[i,]), type = "l", col = "darkgrey")
-}
-
-variance = rep(0,length(unique_rates))
-mise = rep(0,length(unique_rates))
-partial_mise = rep(0,length(unique_rates))
-for(i in 1:length(unique_rates)) {
-  subsetdf = df[df$rates == unique_rates[i],8:ncol(df)]
-  for (row in 1:nrow(subsetdf)) {
-    variance[i] = variance[i] + as.numeric(mise_calc(beta_1t = betameans[i,], betaHat.net = subsetdf[row,])/nrow(subsetdf))
-    mise[i] = mise[i] + as.numeric(mise_calc(beta_1t = beta_truth, betaHat.net = subsetdf[row,])/nrow(subsetdf))
[…]

### 2023-03-27 write-up/fda-recurrent.tex
finishing up the simulation section. onto the proofs this afternoon

@@
 \usepackage{tikz}
+\usepackage{diagbox}
 \usetikzlibrary{matrix,arrows}
@@
 Finally, traditional joint modeling often consider the hazard function to depend on the current or cumulative value of the health process~\citep{Li2022,Rizopoulos2010}.
-For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history and use recent history functional linear models~\cite{KIM20111554} to to account for this local dependence.
+For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history and use recent history functional linear models~\cite{KIM20111554} to account for this local dependence.
 We combine these models with advances in longitudinal functional data analysis~\citep{Staicu2010} and penalized functional regression~\citep{Goldsmith2015} to perform scalable inference using off-the-shelf software.
@@
 \subsection{Selection of bandwidth} \label{section:bandwidth}
+
+A critical consideration for recent history functional linear models~\cite{KIM20111554} is the size of the bandwidth, $\Delta$.
+While above we explore the bias-variance trade-off under misspecification of $\Delta$, here we develop a data-adaptive method for bandwidth selection.
+Specifically, given the model complexity does not change as a function of $\Delta$, our proposal is to simple compare AIC across a range of bandwidths and choose that which maximizes the criterion.
+
+Table~\ref{tab:aic_selection} presents AIC-based selection across the 500 simulated datasets analyzed in the prior section.
+We see markedly distinct behavior of the selection criterion in the two settings.
+The AIC-based selection works well for the sinusoidal effect across subsampling rates, but performs poorly in the exponential setting.
+Recall that the sinusoidal effect (choice 2) is significantly different from 0 at $\Delta^\star$ and then 0 for all $s > \Delta^\star$ while the exponential effect (choice 1) decays slowly towards 0 as $s \to^{-} \Delta$.
+Therefore, the selection problem under the sinusoidal effect is much easier than under the exponential effect.
+Moreover, bias, variance, and mean-integrated squared error are not substantially harmed by misspecification of the bandwidth in the exponential setting.
+Thus, we conclude that the AIC-based selection method works well in settings where inference requires estimating the bandwidth accurately, as desired.
+
+\begin{table}[!th]
+\begin{tabular}{c | c c c c | c c c c}
+& \multicolumn{4}{c}{Exponential} & \multicolumn{4}{c}{Sinusoidal} \\ \hline
+\diagbox{$\Delta=$}{Rate} & 4.0 & 2.0 & 1.0 & 0.5 & 4.0 & 2.0 & 1.0 & 0.5 \\ \hline
+26 & {\bf 0.38} & {\bf 0.35} & {\bf 0.30} & 0.26 & 0.00 & 0.00 & 0.00 & 0.00 \\
+29 & 0.22 & 0.20 & 0.17 & 0.12 & 0.09 & 0.14 & 0.15 & 0.18 \\
+32 & 0.11 & 0.12 & 0.13 & 0.18 & {\bf 0.85} & {\bf 0.79} & {\bf 0.76} & {\bf 0.70} \\
+35 & 0.13 & 0.13 & 0.14 & 0.15 & 0.05 & 0.06 & 0.06 & 0.09 \\
+37 & 0.14 & 0.20 & 0.26 & {\bf 0.28} & 0.01 & 0.01 & 0.02 & 0.03 \\ \hline
+\end{tabular}
+\caption{Selection of bandwidth when true $\Delta = 32$ minutes.
+% SINE EXAMPLE
+}
+\label{tab:aic_selection}
+\end{table}
 

### 2023-03-27 simulations/delta_incorrect/choice_of_delta_summary.R

@@ -1,30 +1,38 @@
-setting = "exponential"
-test36 = read.csv(paste("./output_csv/", setting, "_simdelta_results_36.csv", sep = ""))
-test40 = read.csv(paste("./output_csv/", setting, "_simdelta_results_40.csv", sep = ""))
-test44 = read.csv(paste("./output_csv/", setting, "_simdelta_results_44.csv", sep = ""))
-test48 = read.csv(paste("./output_csv/", setting, "_simdelta_results_48.csv", sep = ""))
-test52 = read.csv(paste("./output_csv/", setting, "_simdelta_results_52.csv", sep = ""))
-
-overlapping_ids = Reduce(intersect, list(test36$ids, test40$ids, test44$ids, test48$ids, test52$ids))
-overlapping_ids = overlapping_ids[order(overlapping_ids)]
-whichmin = function(x) {which(x == min(x))}
-result = rep(0,0)
-for(id in overlapping_ids) {
+make_table <- function(setting) {
   
-  temp = cbind(subset(test36, ids == id)[,3],
-               subset(test40, ids == id)[,3],
-               subset(test44, ids == id)[,3],
-               subset(test48, ids == id)[,3],
-               subset(test52, ids == id)[,3])
-  result_id = cbind(id, apply(temp, 1, whichmin), subset(test36, ids == id)$rates)
-  result = rbind(result, result_id)
-
+  test36 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_36.csv", sep = ""))
+  test40 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_40.csv", sep = ""))
+  test44 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_44.csv", sep = ""))
+  test48 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_48.csv", sep = ""))
+  test52 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_52.csv", sep = ""))
+  
+  overlapping_ids = Reduce(intersect, list(test36$ids, test40$ids, test44$ids, test48$ids, test52$ids))
+  overlapping_ids = overlapping_ids[order(overlapping_ids)]
+  whichmin = function(x) {which(x == min(x))}
+  result = rep(0,0)
+  for(id in overlapping_ids) {
+    
+    temp = cbind(subset(test36, ids == id)[,4],
+                 subset(test40, ids == id)[,4],
[…]

### 2023-03-29 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 
+Here we address the time-complexity of solving the approximate score equations relative to maximizing the log-likelihood using Lemma~\ref{lemma:logistic}.
+The complexity of logistic regression with $n$ observations and covariates of dimension $d$ is $O(nd)$; so maximizing the log-likelihood is well approximated by subsampling with a very high rate so that the design-unbiased score functions have minimal approximation error in comparison to directly maximizing the likelihood function.
+Under this assumption, at the final stage we have the relative complexity of $(4 \times 60 \times \tau_i) \cdot n \cdot d$ compared to the expected time-complexity under subsampling with rate $c$ per minute of $(c \times \tau_i) \cdot n \cdot d$.
+In practice, this can lead to large differences in run time; specifically, this means sampling every minute rather than every half hour will be approximately $2^5$ times as long (from 5 minutes to 2.8 hours), while sampling every second will take $2^11$ times as long (from 5 minutes to 180 hours).
+In our first example, the run time was 356 seconds (approximately 5.9 minutes) when sampling every five minutes.
+Thus, if sampling increased to every minute or every second, then run time would increase to approximately 30 minutes and 30 hours respectively.
+In our second example, the run time of 18 minutes would increase to 1.5 hours and 90 hours respectively.
+In both instances, the relative efficiency gain would be negligible suggesting a huge computational increase for minimal relative information gain.
+
 \section{Extensions}

### 2023-03-30 letters/response-ref-1-v1.tex
fix the write-up

@@
 As stated by the second referee, our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
-This motivates Section XX which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
+This motivates Section 5.2 which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
 \vspace{5mm}
@@
 This holds unless we assume the measurement error is set to zero.
-While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
+While we still think it is an important point that should be more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
 Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
@@
 \vspace{5mm}
-We added Section XX to discuss choice of window size; specifically, we consider selection based on Akaike information criteria as part of the simulation study rather than simply the impact of incorrect choice of $\Delta$.
-Figure XX shows that the adaptive tuning of $\Delta$ based on AIC does perofrm well empirically.
+We added Section 5.2 to discuss choice of window size; specifically, we consider selection based on Akaike information criteria as part of the simulation study rather than simply the impact of incorrect choice of $\Delta$.
+Table 4 shows that when inference is sensitive to bandwidth selection, the adaptive tuning of $\Delta$ based on AIC does perform well empirically.
 We now use this approach in selecting~$\Delta$ in the case study, and provide a plot of AIC across several choices of $\Delta$ in Figure XX in Appendix XX.
@@
 
-The smoothing parameter selection is also discussed in Section XX; given our connection to off-the-shelf penalized logistic regression, the smoothing parameter selection occurs as part of this implementation.
-Therefore, we state how these are chosen generally in the software and provide some additional details in Appendix XX.
+The smoothing parameter selection is also discussed in Section 4.6; given our connection to off-the-shelf penalized logistic regression, the smoothing parameter selection occurs as part of this implementation.
+In the R package we are using, `glmnet', the smoothness parameter is chosen via cross-validation.
 \vspace{5mm}

### 2023-03-30 write-up/fda-recurrent.tex
fix the write-up

--- 2023-03-27 write-up/fda-recurrent.tex
@@
 Combining the penalized spline formulation with Lemma~\ref{lemma:logistic} establishes a connection between our approximate score equations and solving a generalized mixed effects logistic regression with offset.
-Given the connection with generalized mixed effects models, the inferential machinery to obtain variance-covariance estimates .
-In this context, we acknowledge another limitation.
-Specificaly, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is oversmoothed.
+Given the connection with generalized mixed effects models, the inferential machinery to obtain variance-covariance estimates.
+\textcolor{blue}{As we leverage the standard R package `glmnet', the smoothness parameter is chosen via cross-validation.} In this context, we acknowledge another limitation.
+Specifically, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is over-smoothed.
 
@@
 We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day.
-We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$.
+We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $ \textcolor{blue}{which decays to 0 as $s$ approaches $\Delta$ from below}, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$ \textcolor{blue}{which is significantly different from 0 as $s$ approaches $\Delta$ from below.}
 
 We generate $1000$ datasets, each consisting of $500$ user-days.
-For a given simulated user-day, we randomly sample non-event times using a Poisson process with rate of a halfhour.
-We use the proposed methodology to construct the estimate $\hat \beta_{i,0.5} (t)$ for the $i$th simulated dataset; we then subsample the sampled non-event times with thinning probability $1/2$, $1/4$ and $1/8$.
-This results in randomly sampled non-event times given by a Poisson process with rates of an hour, two hours, and four hours.
-We can construct the corresponding estimates: $\hat \beta_{i,1} (t)$, $\hat \beta_{i,2} (t)$, and $\hat \beta_{i,4} (t)$ respectively.
+For a given simulated user-day, we randomly sample non-event times using a Poisson process with rate of every five minutes.
+We use the proposed methodology to construct the estimate $\hat \beta_{i,12} (t)$ for the $i$th simulated dataset; we then subsample the sampled non-event times with thinning probability $1/3$, $1/6$, $1/12$, and $1/24$.
+This results in randomly sampled non-event times given by a Poisson process with rates of every fifteen minutes, half-hour, hour, and two hours.
+We can construct the corresponding estimates: $\hat \beta_{i,4} (t)$, $\hat \beta_{i,2} (t)$, $\hat \beta_{i,1} (t)$, and $\hat \beta_{i,0.05} (t)$ respectively.
 Subsampling allows us to compare the variance due to subsampling as compared to the variance due to sampling fewer non-event times.
 
-Since we are primarily interested in accuracy, we report the mean integrated squared error (MISE) defined as~$\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \beta(s))^2 ds$ for $j=0.5,1,2,4$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
+Since we are primarily interested in accuracy, we report the mean integrated squared error (MISE) defined as~$\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \beta(s))^2 ds$ for each~$j$ where~$\beta(s) \equiv 0$ for $s > \Delta$.
 The MISE is defined in this manner to account for settings where~$\Delta$ is unknown.
@@
 Then the squared bias is given by $\int_{0}^{\infty} ( \bar \beta_{j} (s) - \beta (s))^2 ds$ and the variance is given by $\frac{1}{1000} \sum_{j=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \bar \beta_j(s))^2 ds$.
-The subsampling variance is defined as $\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \hat \beta_{i,0.5} (s))^2 ds$.
+The subsampling variance is defined as $\frac{1}{1000} \sum_{i=1}^{1000} \int_{0}^{\infty} ( \hat \beta_{i,j} (s) - \hat \beta_{i,12} (s))^2 ds$.
 Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biased as well as the subsampling variance.
@@
 
-Table~\ref{tab:mise} demonstrates that the variance does increase as the sampling rate decreases.
+Table~\ref{tab:mise} demonstrates that the variance may increase as the sampling rate decreases.
 However, the rate of increase in the MISE is relatively quite low.
-In the second case study, for example, the MISE increases by 6\% while the run time is 3 times faster.
-In the first case study, the MISE remains roughly constant but the run time is 4.7 times faster.
+In the first setting, for example, the RMISE when sampling every hour versus every five minutes is 1.16 while the run time is 14.8 times faster.
+In the second setting, the RMISE remains below or equal to $1$ while the run time is 23.1 times faster when sampling every hour rather than every minute.
 This highlights the efficiency-computation trade-off.
-Extrapolating to a complete data analysis, we can see that the run time will continue to increase significantly with minimal improvement in the mean integrated squared error.
+
+\begin{newmaterial}
+\begin{rmk}[Time-complexity and run-time of maximum likelihood estimation]
+The complexity of logistic regression with $n$ observations and covariates of dimension $d$ is $O(nd)$.
+Maximum likelihood estimation can be well approximated by subsampling at a very high rate; therefore, under a 4Hz sensor the time-complexity of the maximum likelihood estimate is $(4 \times 60 \times \tau_i) \cdot n \cdot d$ compared to the expected time-complexity under subsampling with rate $c$ per minute of $(c \times \tau_i) \cdot n \cdot d$.
+This means using the sensor at its observation frequency of four times per second will take approximately $2^{13}$ times as long as subsampling at a rate of once every half-hour.
+In our first example, using the average run times at each sampling rate to project run time at four times per second yields an approximately 2.0 hour run time.
+In our second example, we project a run time at four times per second of 6.2 hours.
+In both instances, the relative efficiency gain would be negligible suggesting a huge computational increase for minimal relative information gain.
+\end{rmk}
+\end{newmaterial}
+
+% Extrapolating to a complete data analysis, we can see that the run time will continue to increase significantly with minimal improvement in the mean integrated squared error.
 
@@
 When the window length is too large, i.e., $\Delta > \Delta^\star$, then asymptotically the estimation is unbiased as $\beta(t) \equiv 0$ for $t > \Delta$; however, we incur a penalty in finite samples, especially for settings where the function is far from zero near $t = \Delta$.
-We find the MISE increases the absolute error $|\Delta - \Delta^\star|$ increases, while the variance decreases as a function of $\Delta$.
+We find the MISE increases as the absolute error $|\Delta - \Delta^\star|$ increases.
 While the MISE increased for $\Delta < \Delta^\star$, the pointwise estimation error remains low for $t < \Delta$.
@@
 
-A critical consideration for recent history functional linear models~\cite{KIM20111554} is the size of the bandwidth, $\Delta$.
-While above we explore the bias-variance trade-off under misspecification of $\Delta$, here we develop a data-adaptive method for bandwidth selection.
-Specifically, given the model complexity does not change as a function of $\Delta$, our proposal is to simple compare AIC across a range of bandwidths and choose that which maximizes the criterion.
-
+While above we explore the bias-variance trade-off under bandwidth misspecification, here we explore a data-adaptive method for bandwidth selection, a critical consideration for recent history functional linear models~\citep{KIM20111554}.
+Specifically, given the model complexity does not change as a function of $\Delta$, our modest proposal is to compare AIC across a range of bandwidths and choose that which maximizes the criterion.
 Table~\ref{tab:aic_selection} presents AIC-based selection across the 500 simulated datasets analyzed in the prior section.
@@
 The AIC-based selection works well for the sinusoidal effect across subsampling rates, but performs poorly in the exponential setting.
-Recall that the sinusoidal effect (choice 2) is significantly different from 0 at $\Delta^\star$ and then 0 for all $s > \Delta^\star$ while the exponential effect (choice 1) decays slowly towards 0 as $s \to^{-} \Delta$.
+Recall that the sinusoidal effect is significantly different from 0 at $\Delta^\star$ and then 0 for all $s > \Delta^\star$ while the exponential effect decays slowly towards 0 as $s \to^{-} \Delta$.
 Therefore, the selection problem under the sinusoidal effect is much easier than under the exponential effect.
-Moreover, bias, variance, and mean-integrated squared error are not substantially harmed by misspecification of the bandwidth in the exponential setting.
-Thus, we conclude that the AIC-based selection method works well in settings where inference requires estimating the bandwidth accurately, as desired.
+Moreover, bias, variance, and MISE do not vary substantially under bandwidth misspecification in the exponential setting.
+Thus, we conclude that the AIC-based selection method works well in settings where inference depends heavily on estimating the bandwidth accurately.
 
@@
 \end{tabular}
-\caption{Selection of bandwidth when true $\Delta = 32$ minutes.
+\caption{Rate of selection of $\hat \Delta$ across 500 simulations when true $\Delta = 32$ minutes. Bold: most likely bandwidth given a chosen subsampling rate.
 % SINE EXAMPLE

### 2023-03-30 simulations/delta_incorrect/choice_of_delta_summary.R

@@ -1,38 +1,38 @@
-make_table <- function(setting) {
-  
-  test36 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_36.csv", sep = ""))
-  test40 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_40.csv", sep = ""))
-  test44 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_44.csv", sep = ""))
-  test48 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_48.csv", sep = ""))
-  test52 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_52.csv", sep = ""))
-  
-  overlapping_ids = Reduce(intersect, list(test36$ids, test40$ids, test44$ids, test48$ids, test52$ids))
-  overlapping_ids = overlapping_ids[order(overlapping_ids)]
-  whichmin = function(x) {which(x == min(x))}
-  result = rep(0,0)
-  for(id in overlapping_ids) {
+  make_table <- function(setting) {
     
-    temp = cbind(subset(test36, ids == id)[,4],
-                 subset(test40, ids == id)[,4],
-                 subset(test44, ids == id)[,4],
-                 subset(test48, ids == id)[,4],
-                 subset(test52, ids == id)[,4])
-    result_id = cbind(id, apply(temp, 1, whichmin), subset(test36, ids == id)$rates)
-    result = rbind(result, result_id)
+    test36 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_36.csv", sep = ""))
+    test40 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_40.csv", sep = ""))
+    test44 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_44.csv", sep = ""))
+    test48 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_48.csv", sep = ""))
+    test52 = read.csv(paste("./output_csv/", setting, "_simdelta_allresults_wl_52.csv", sep = ""))
     
+    overlapping_ids = Reduce(intersect, list(test36$ids, test40$ids, test44$ids, test48$ids, test52$ids))
+    overlapping_ids = overlapping_ids[order(overlapping_ids)]
+    whichmin = function(x) {which(x == min(x))}
+    result = rep(0,0)
[…]

### 2023-03-31 write-up/fda-recurrent.tex
Update fda-recurrent.tex

@@
 
-The goal of this paper is two-fold: (1) to discuss the appropriate choice of statistical model for joint high-frequency sensor data and the recurrent event data, and (2) to construct a simple, easy-to-implement method for parameter estimation and inference.
-For (1), we discuss an important issue regarding measurement-error models when paired with recurrent event outcomes.
-For (2), we introduce a random subsampling procedure that has several benefits.
+The goal of this paper is to construct a simple, easy-to-implement method for parameter estimation and inference.
+% For (1), we discuss an important issue regarding measurement-error models when paired with recurrent event outcomes.
+To do so, we introduce a random subsampling procedure that has several benefits.
 First, the resulting inference is unbiased; however, there is a computation-efficiency trade-off.
@@
 
-\begin{rmk}[Assessment of computational benefit]
-Cool we have speed ups.
-\end{rmk}
+% \begin{rmk}[Assessment of computational benefit]
+% Cool we have speed ups.
+% \end{rmk}
 
@@
 
+\begin{newmaterial}
 \subsection{Selection of bandwidth} \label{section:bandwidth}
@@
 \end{table}
+
+\end{newmaterial}
 

### 2023-03-31 letters/response-ref-1-v1.tex
editing response to reflect new paper edits

@@
 \vspace{5mm}
-There does not exist a model that can do all so measures of bias and estimation accuracy may be inappropriate.
-To handle this, we simply look at the average or the exponentially weighted where we can conceptualize it as a pre-processing step.
-Then we can apply traditional joint models.
-The issue is that the traditional jiont model cannot scale to this problem.
-Instead, we
-In Section XX, we perform the comparison in a simulation when compared to an unpenalized $K_x = 10$ linear spline model.
-WE see that both are unbiased and our glmnet takes XX minutes while joint modeling took over XX hours.
-To further this comparison, we perform a back-of-the-envelope calculation.
+As far as the author is aware, there does not exist a joint modeling package that can estimate the proposed model directly; one way would be assume you know the functional form of the parameter but not the scale.
+Comparison of measures of bias, estimation accuracy, and numerical speed under this approach to our methodology which learns the functional form are likely inappropriate.
+To address the reviewers concern, we now highlight the extent of the numerical issues in Remark 5.1.
+Specifically, running a traditional joint model will take a similar time to running our approach under very high subsampling frequencies, i.e., at the sensor measurement rate.
+We now subsample at a rate of every five minutes, fifteen minutes, thirty minutes, one hour, and two hours in our simulation section.
+Using a time-complexity analysis (supported by the simulations), we can extrapolate run times when sampling at 4Hz to a standard joint model needing approximately 2 and 6.2 hours to run in our two settings respectively; at 32Hz this would be 16 and 50.1 hours respectively.
+These are extremely high run times when we compare to a run-time of 1.8 and 3.5 minutes when subsampling at rate of every fifteen minutes.
+Moreover, the benefits in terms of variance, squared bias, and MISE appear almost negligible.
 \vspace{5mm}
@@
 As stated by the second referee, our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
-This motivates Section XX which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
+This motivates Section 5.2 which provides an empirical study of choosing the tuning parameter~$\Delta$ using an AIC-based selection criteria.
 \vspace{5mm}
@@
 Then the probability of the observed event times per person follows from $\exp \left( - H_i (\tau_i ; \theta) \right) \prod_{j=1}^{k_i} h \left(T_{i,j};\theta \right)$.
-Thus the log-likelihood of the even process is given by $L_n (\theta)$.
+Thus the log-likelihood of the event process is given by $L_n (\theta)$.
 \vspace{5mm}
@@
 \vspace{5mm}
-The key point raised on page 4 (now in Appendix XX), is that if $\epsilon(t)$ is not equivalently zero, your model assumes risk of an event at time $t$ depends on both the current and future measurements of the process ${\bf x}_i$.
+The key point raised on page 4 of the original submission is that if $\epsilon(t)$ is not equivalently zero, your model assumes risk of an event at time $t$ depends on both the current and future measurements of the process ${\bf x}_i$.
 We argue this is not sensible since current risk should not depend on future risk and thus set $\epsilon(t) = 0$.
@@
 
-\item The ``original'' score equation Page 5, line 9 involves
+\item {\it The ``original'' score equation Page 5, line 9 involves
 $$
@@
 With the basis expansion strategy, you
-would will be estimating the same number of parameters. This type of approach should be added to simulations and evaluated for accuracy and time complexity.
+would will be estimating the same number of parameters. This type of approach should be added to simulations and evaluated for accuracy and time complexity.}
 

### 2023-03-31 letters/response-ref-1-v1.tex
editing response to reflect new paper edits

@@
 \vspace{5mm}
-We have added Remark XX and Section XX in the simulation study to give additional details and empirical evidence.
-First, there are two areas where we have computational improvements.
-First, in the inner integral, we have
-
-Second, even if we have computed $\int_{t-\Delta} x(t-s)\beta(s) ds$
-We explain that if we were to compute it would have to be as a function of $\beta(s)$.
-We implement a naive and demonstrate numerically the computational cost.
-Specifically, each computation in simulation takes XXseconds, and we must compute this XX times to compute the likel
+We have added Remark 5.1 and improved our simulation study to give additional details and empirical evidence.
+There are two areas where we have computational improvements.
+First, in the inner integral, we have to compute $\int_{t-\Delta} X(t,s)\beta(s) ds$ without knowing the functional form of $\beta(s)$; we leverage traditional functional data analysis approaches to convert this into a parametric problem given the Karhunen-Lo{\`e}ve decomposition of $X(t,s)$ and a flexible spline decomposition of~$\beta(s)$.
+Second, even if we have computed $\int_{t-\Delta} x(t-s)\beta(s) ds$, we would still need to integrate over $t$, which is highly non-trivial.
+Performing this integration directly, would then lead to a gradient-based method for maximizing the log-likelihood.
+We now highlight the extent of the numerical issues in Remark 5.1.
+Specifically, running a traditional joint model with such a gradient-based method will take a similar time to running our approach under very high subsampling frequencies, i.e., at the sensor measurement rate.
+We now subsample at a rate of every five minutes, fifteen minutes, thirty minutes, one hour, and two hours in our simulation section.
+Using a time-complexity analysis (supported by the simulations), we can extrapolate run times when sampling at 4Hz to a standard joint model needing approximately 2 and 6.2 hours to run in our two settings respectively; at 32Hz this would be 16 and 50.1 hours respectively.
+These are extremely high run times when we compare to a run-time of 1.8 and 3.5 minutes when subsampling at rate of every fifteen minutes.
+Moreover, the benefits in terms of variance, squared bias, and MISE appear almost negligible.
 \vspace{5mm}

### 2023-05-08 methods/bootstrap_likelihood_estimation.R
Update bootstrap_likelihood_estimation.R

@@ -12,13 +12,16 @@ if(length(args)==0){
   ##supply default values
   current_bootstrap = 1 # CONVERT TO ACCEPT THIS 
   current_mi = 1
+  currrent_Delta = 30
 }else{
   current_bootstrap = as.numeric(args[[1]])
   current_mi = as.numeric(args[[2]])
+  current_Delta = as.numeric(args[[3]])
 }
 
 print(paste("Current bootstrap:", current_bootstrap))
 print(paste("Current mi:", current_mi))
+print(paste("Current Delta:", current_Delta))
 
 ## WINDOWS
 setwd("Z:/SI_data/bootstrap_files_Delta/")
@@ -168,9 +171,6 @@ ridge.fit.cv <- cv.glmnet(all_model.matrix[daytime_obs,], all_Y[daytime_obs],
 runtime = proc.time() - ptm
 
 ridge.fit.lambda <- ridge.fit.cv$lambda.min
-bindeviance[which(c(5,15,30) == Delta)] = deviance(ridge.fit.cv$glmnet.fit)[lambdapath == ridge.fit.lambda]
-lengthY[which(c(5,15,30) == Delta)] = length(all_Y[daytime_obs])
-
 # Extract coefficient values for lambda.1se (without intercept)
 ridge.coef <- (coef(ridge.fit.cv, s = ridge.fit.lambda))[-1]
 intercept <- (coef(ridge.fit.cv, s = ridge.fit.lambda))[1]
@@ -203,15 +203,15 @@ df_acc_summary <- data.frame(sequence = acc_sequence+Delta, estimate = acc_betaH
 # acc_xmax = max(df_acc_summary$sequence[which(df_acc_summary$lowerCI > 0)])
 # acc_xmin = min(df_acc_summary$sequence[which(df_acc_summary$lowerCI > 0)])
 
-png(paste("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/acc_coef_joint_Delta_",Delta,".png", sep = ""),
-    # png(paste("~/Documents/github/fda-recurrentevents/figures/acc_coef_joint_Delta_",Delta,".png", sep = ""),
-    width = 480, height = 480, units = "px", pointsize = 16)
-ggplot(df_acc_summary, aes(x=sequence, y=estimate)) +
-  geom_line(size=1, alpha=0.8) +
-  geom_ribbon(aes(ymin=lowerCI, ymax=upperCI) ,fill="blue", alpha=0.2) +
[…]

### 2023-05-08 methods/bootstrap_likelihood_estimation.R
Update bootstrap_likelihood_estimation.R

@@ -24,10 +24,11 @@ print(paste("Current mi:", current_mi))
 print(paste("Current Delta:", current_Delta))
 
 ## WINDOWS
-setwd("Z:/SI_data/bootstrap_files_Delta/")
+# setwd("Z:/SI_data/bootstrap_files_Delta/")
 # ## LINUX 
-# setwd('/mnt/turbo/SI_data/bootstrap_files')
+# setwd('/mnt/turbo/SI_data/bootstrap_files_Delta')
 ## Turbo
+setwd('/nfs/turbo/sph-wdem/SI_data/bootstrap_files_Delta/')
 
 ## LIBRARIES
 library(ggplot2)

### 2023-05-08 letters/response-ref-1-v1.tex

@@
 All major edits are highlighted in blue in order to help the referees know what has changed as part of the major revision.
+% CHECK
 \vspace{5mm}
@@
 This motivates Section 5.2 which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
+% CHECK
 \vspace{5mm}
@@
 \vspace{5mm}
+\textcolor{red}{
 We added Section 5.2 to discuss choice of window size; specifically, we consider selection based on Akaike information criteria as part of the simulation study rather than simply the impact of incorrect choice of $\Delta$.
 Table 4 shows that when inference is sensitive to bandwidth selection, the adaptive tuning of $\Delta$ based on AIC does perform well empirically.
-We now use this approach in selecting~$\Delta$ in the case study, and provide a plot of AIC across several choices of $\Delta$ in Figure XX in Appendix XX.
+We now use this approach in selecting~$\Delta$ in the case study, and provide a plot of AIC across several choices of $\Delta$ in Figure XX in Appendix XX.}
 \vspace{5mm}
@@
 \vspace{5mm}
-To accommodate the second referee's request without making the discussion too long, we have added Remark XX and XX that provide some intuition.
+\textcolor{red}{To accommodate the second referee's request without making the discussion too long, we have added Remark XX and XX that provide some intuition.
 We have re-written the technical proofs to first provide a sketch proof (with intuition) and then a detailed proof.
-We hope this helps the reader develop intuition on the first read while giving them some helpful guidance in order to facilitate understanding of the detailed theoretical analysis.
+We hope this helps the reader develop intuition on the first read while giving them some helpful guidance in order to facilitate understanding of the detailed theoretical analysis.}
 \vspace{5mm}
@@
 While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
-Therefore, we revert to the step from Park and Staicu (2018) and report the smoothing of the covariance matrix as it comes from their prior work.
+We compare our approach to Park and Staicu (2018) and report the smoothing of the covariance matrix as it comes from their prior work.
 \vspace{5mm}
@@
 \vspace{5mm}
-Both of these were defined in Section 3 but we agree that the definitions were obscured in the writing, which led to confusion.  To address this, we have added a notation glossary as Appendix XX.
+\textcolor{red}{Both of these were defined in Section 3 but we agree that the definitions were obscured in the writing, which led to confusion.
+To address this, we have added a notation glossary as Appendix XX.}
 \vspace{5mm}
@@
 \vspace{5mm}
-The term $\tau$ refers to the censoring time of the observation process. We make this clear in Section XXX, page XX in the revision.
+The term $\tau$ refers to the censoring time of the observation process. We make this clear in page 4 in Section 3 and pages 8--9 in Section 4.1 of the revision.
 \vspace{5mm}
@@
 The subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$; under the assumption that $\pi_i (t) = \lambda$, we can view this as first drawing the number of non-event times from a Poisson distribution, i.e., $| D_i | \sim Poisson(\lambda)$, and then independently selecting $|D_i|$ non-event times uniformly at random over the observation window~$[0, \tau_i]$.
-Given the potential difference between the distributions, we choose to model the mean and covariance functions separately to ensure good approximation when applying th espectral decomposition of the estimated covariance function.
+Given the potential difference between the distributions, we choose to model the mean and covariance functions separately to ensure good approximation when applying the spectral decomposition of the estimated covariance function.
 \vspace{5mm}
 
-Another reason for treating these two distinctly is that the missing data imputation from XXXX requires distinct covariance functions.
-We do think one could marginalize across the two if the mean and covariance functions are not expected to differ greatly and state as such in Section XX (in blue) where we discuss the new Figure XX of the mean and covariance functions for $y=0$ and $y=1$.
+\textcolor{red}{Another reason for treating these two distinctly is that the missing data imputation from XXXX requires distinct covariance functions.
+We do think one could marginalize across the two if the mean and covariance functions are not expected to differ greatly and state as such in Section XX (in blue) where we discuss the new Figure XX of the mean and covariance functions for $y=0$ and $y=1$.}
 \vspace{5mm}
@@
 \vspace{5mm}
-We have added mean and covariance functions to the supplementary materials as suggested.
-The plots demonstrate distinct behavior in the means for $y=0$ and $y=1$; while the covariance functions are XXXX.
+\textcolor{red}{We have added mean and covariance functions to the supplementary materials as suggested.
+The plots demonstrate distinct behavior in the means for $y=0$ and $y=1$; while the covariance functions are XXXX.}
 \vspace{5mm}
@@
 \vspace{5mm}
-We have added Remark XX to provide a simple example where these assumptions are satisfied.
-Specifically, if the functional process is bounded, i.e., $|X(t,s)| < M$ and the other component is also bounded then we are guaranteed to have these assumptions hold.
-This is likely to be true in most practical settings and thus demonstrates the reasonableness of our assumptions for applied settings.
+We have added Remark 4.5 to provide a simple example where these assumptions are satisfied.
+Specifically, if individuals are assumed independent and identically distributed, the functional process is bounded (i.e., $|X(t,s)| < M$), and the subsampling rate is both lower and upper bounded at all risk times then we are guaranteed to have Assumptions 4.3 and 4.4.
+Boundedness is likely to be true in most practical settings and thus demonstrates the reasonableness of our assumptions for applied settings.
 \vspace{5mm}
@@
 Under this assumption, at the final stage we have the relative complexity of $(4 \times 60 \times \tau_i) \cdot n \cdot d$ compared to the expected time-complexity under subsampling with rate $c$ per minute of $(c \times \tau_i) \cdot n \cdot d$.
-In practice, this can lead to large differences in run time; specifically, this means sampling every minute rather than every half hour will be approximately $2^5$ times as long (from 5 minutes to 2.8 hours), while sampling every second will take $2^11$ times as long (from 5 minutes to 180 hours).
-In our first example, the run time was 317 seconds (approximately 5 minutes).
-Sampling every minute or every second would increase run time to approximately 2.8 hours and 180 hours respectively.
-In our second example, the run time of 36 seconds would increase to 19 minutes and 20 hours respectively.
+In practice, this can lead to large differences in run time; specifically, under a 4Hz sensor the time-complexity of the maximum likelihood estimate at its observation frequency will take approximately $2^{13}$ times as long as subsampling at a rate of once every half-hour.
+In our first example, using the average run times at each sampling rate to project run time at four times per second yields an approximately 2.0 hour run time.
+In our second example, we project a run time at four times per second of 6.2 hours.
 In both instances, the relative efficiency gain would be negligible suggesting a huge computational increase for minimal relative information gain.
-We now make this point abundantly clear by amending Section XX to include time complexity discussion and Section XX to include the exact calculations above.
+We now make this point abundantly clear by amending Section 5 with Remark 5.1 to include time complexity discussion.
 
@@
 \vspace{5mm}
-uncongenial comes from Meng. We cite and provide a short formal definition with some intuition.
+Rubin defined conditions for an imputation procedure to be so-called ‘proper’ for a given complete data analysis.
+If in addition the complete data analysis gives frequentist valid inferences, MI using Rubin’s rules yields valid frequentist inferences.
+Meng defined the concept of congeniality between an imputation procedure and an analyst’s complete (and incomplete) analysis procedure.
+The imputation model and the analyst’s complete data procedure are said to be congenial if there exists a unifying Bayesian model which embeds the imputer’s imputation model and the analyst’s complete data procedure.
+A formal definition of congeniality can be found Bartlett and Hughes (2020).
+If an imputation and analysis procedure are congenial, this implies the imputation is proper for the analysis procedure.
+Meng showed that for certain types of uncongeniality, Rubin’s variance estimator is conservative, ensuring the intervals have at least the advertised coverage level.
+In other settings, however, it can be biased downwards, leading to under-coverage of confidence intervals.
+We provide these details as part of our discussion in Section 6.2.1.
 \vspace{5mm}
@@
 Therefore, we keep the plot as is, and note that there is a spike at 30-minutes for EDA while emphasizing that most of the signal appears in the 15-minute range.
-This helps motivate our data adaptive choice of $\Delta$ in Section 7.1 which confirms that $\Delta = 15$ is a more appropriate choice.
+This helps motivate our data adaptive choice of $\Delta$ in Section 7.1 which confirms that $\Delta = 5$ is the most appropriate choice although results are consistent across a range of $\Delta$ choices.
 \vspace{5mm}
@@
 A comparison of AIC provides empirical evidence that the smaller $\Delta$ was an appropriate choice.
-We now present $\Delta = 15$ in the main paper and move the current analysis to be a sensitivity analysis presented in the supplementary materials.
+We now present $\Delta = 5$ in the main paper and move the current analysis to be a sensitivity analysis presented in the supplementary materials.
 \vspace{5mm}

### 2023-05-08 write-up/fda-recurrent.tex

@@
 \label{sec:margcov}
-We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
-The sandwich smoother~\citep{Xiao2013} is used to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
+\textcolor{blue}{
+We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.} The sandwich smoother~\citep{Xiao2013} is used to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
 Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009}, bivariate tensor product splines~\citep{Wood2006}, and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
@@
 \]
-for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((t,s), (t,s^\prime))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$ and $f_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
+for~$0 \leq s,s^\prime \leq \Delta$, where~$c_y((t,s), (t,s^\prime))$ is the covariance function of the windowed covariate history~$X(t,\cdot)$, $\tau$ is the observation window length of the process, and $f_y(T)$ is the intensity function for event ($y=1$) and subsampled ($y=0$) times respectively.
 Estimation of~$\Sigma_y$ occurs in two steps.
@@
 Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
+\begin{rmk}[Practical consequences of Assumptions~\ref{assumption:events} and~\ref{assumption:truncation}]
+Bounded and iid.
+\end{rmk}
+
 With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
@@
 The main disadvantage is that the above framework is uncongenial under many events and/or high sampling rates.
-Again, congeniality ensures good frequentist coverage properties.
-A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
+\textcolor{blue}{\cite{Meng1994} defines congeniality between an imputation procedure and an analyst’s complete (and incomplete) analysis procedure if there exists a unifying Bayesian model which embeds the imputer’s imputation model and the analyst’s complete data procedure.
+A recent discussion of congeniality can be found Bartlett and Hughes (2020).
+Congeniality ensures good frequentist coverage properties.
+In some uncongenial settings, standard multiple imputation can be biased downwards leading to under-coverage of confidence intervals.}    A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
 

### 2023-05-08 methods/bootstrap_likelihood_estimation.R
Update bootstrap_likelihood_estimation.R

@@ -12,16 +12,16 @@ if(length(args)==0){
   ##supply default values
   current_bootstrap = 1 # CONVERT TO ACCEPT THIS 
   current_mi = 1
-  currrent_Delta = 30
+  Delta = 30
 }else{
   current_bootstrap = as.numeric(args[[1]])
   current_mi = as.numeric(args[[2]])
-  current_Delta = as.numeric(args[[3]])
+  Delta = as.numeric(args[[3]])
 }
 
 print(paste("Current bootstrap:", current_bootstrap))
 print(paste("Current mi:", current_mi))
-print(paste("Current Delta:", current_Delta))
+print(paste("Current Delta:", Delta))
 
 ## WINDOWS
 # setwd("Z:/SI_data/bootstrap_files_Delta/")

### 2023-05-08 methods/bootstrap_likelihood_estimation.R
Update bootstrap_likelihood_estimation.R

@@ -236,11 +236,11 @@ print("Onto saving these files")
 
 library(lubridate)
 
-saveRDS(df_acc_summary, paste("acc_bootstrap_Delta_",Delta,
+saveRDS(df_acc_summary, paste("estimates/acc_bootstrap_Delta_",Delta,
                               "_bs_",current_bootstrap, "_mi_", current_mi,
                               "_", today(), ".RDS", sep = ""))
 
-saveRDS(df_eda_summary, paste("eda_bootstrap_Delta_",Delta,
+saveRDS(df_eda_summary, paste("estimates/eda_bootstrap_Delta_",Delta,
                               "_bs_",current_bootstrap, "_mi_", current_mi,
                               "_", today(), ".RDS", sep = ""))
 

### 2023-05-09 write-up/fda-recurrent.tex
updated

@@
 
+\begin{table}
+\begin{tabular}{c c c c}
+Delta & 5 & 15 & 30 \\ \hline
+AIC & - & - & - \\
+BIC & - & - & - \\ \hline
+\end{tabular}
+\end{table}
+
 We analyzed activity index (AI) and electrodermal activity separately as given in equation~~\eqref{eq:hazardlinear} as well as jointly in a multivariate model as in equation~\eqref{eq:multihazardlinear}.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint.png}
+  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint_bootstrap_Delta_15.png}
   \caption{Electrodermal activity (EDA)}
@@
   \centering
-  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint.png}
+  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint_bootstrap_Delta_15.png}
   \caption{Activity Index (AI)
@@
 \end{subfigure}
-\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the complete-case analysis, while highlighted region are the model-based pointwise 95\% confidence intervals.}
+\caption{$\beta (t)$ for AI and EDA with 95\% CI; solid line is point estimate for the complete-case analysis, while highlighted region are the model-based pointwise 95\% confidence intervals.}
 \label{fig:edacc_estimate}
@@
 
-\subsection{Accounting for missing data}
+\subsection{Sensitivity analysis for window length}
 \label{sec:missingdata_heterogeneity}
@@
   \centering
-  \includegraphics[width=.8\linewidth]{../figures/eda_coef_joint_bootstrap.png}
-  \caption{Electrodermal activity (EDA)}
-  \label{fig:eda_estimate_bs}
+  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint_bootstrap_Delta_5.png}
+  \caption{$\Delta = 5$ minutes}
+  \label{fig:acc_estimate_bs_delta5}
 \end{subfigure}%
@@
   \centering
-  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint_bootstrap.png}
-  \caption{Activity Index (AI)
-}  \label{fig:acc_estimate_bs}
+  \includegraphics[width=.8\linewidth]{../figures/acc_coef_joint_bootstrap_Delta_30.png}
+  \caption{$\Delta = 30$ minutes}
+  \label{fig:acc_estimate_bs_delta30}
 \end{subfigure}

### 2023-05-09 methods/bootstrap_combine.R

@@ -5,7 +5,7 @@ number_bootstraps = 200
 number_mi = 2
 seq_of_Delta = c(5,15,30)
 
-current_Delta = 30
+current_Delta = 5
 
 acc_size = length(seq(-current_Delta,0, by = 1/6))
 eda_size = length(seq(-current_Delta,0, by = 1/60))
@@ -118,7 +118,9 @@ ggplot(df_acc_summary, aes(x=sequence, y=estimate)) +
   geom_line(size=1, alpha=0.8) +
   geom_ribbon(aes(ymin=lowerCI, ymax=upperCI) ,fill="blue", alpha=0.2) +
   annotate("rect",xmin=acc_xmin,xmax=acc_xmax,ymin=-Inf,ymax=Inf, alpha=0.1, fill="black") +
-  xlab("Time until Button Press") + ylab(expression(paste(beta, "(s)")))
+  xlab("Time until Button Press") + ylab(expression(paste(beta, "(s)"))) +
+  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=15),
+        axis.title.x = element_text(size=15), axis.title.y = element_text(size=15))
 dev.off()
 
 df_eda_summary <- data.frame(sequence = eda_temp_mi1$sequence,
@@ -131,5 +133,7 @@ png(paste("C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/figures/eda_c
 ggplot(df_eda_summary, aes(x=sequence, y=estimate)) +
   geom_line(size=1, alpha=0.8) +
   geom_ribbon(aes(ymin=lowerCI, ymax=upperCI) ,fill="blue", alpha=0.2) +
-  xlab("Time until Button Press") + ylab(expression(paste(beta, "(s)")))
+  xlab("Time until Button Press") + ylab(expression(paste(beta, "(s)"))) +
+  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=15),
+        axis.title.x = element_text(size=15), axis.title.y = element_text(size=15))
 dev.off()

### 2023-05-22 letters/response-ref-1-v1.tex
Update response-ref-1-v1.tex

@@
 We have addressed the two reviewers' specific comments.
-Responses to their concerns are detailed below and have significantly improved the clarify of this manuscript.
+Responses to their concerns are detailed below and have significantly improved the clarity of this manuscript.
 All major edits are highlighted in blue in order to help the referees know what has changed as part of the major revision.
@@
 We connect and contrast our use of subsampling with the general use of subsampling in handling massive data and its recent use in massive data with functional predictors.
-As stated by the second referee, our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
-This motivates Section 5.2 which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
+As stated by the second referee, our proposed joint model uses a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
+This motivates Section 5.2 which provides an additional empirical study and discussion of the tuning parameter~$\Delta$.
 % CHECK
@@
 \vspace{5mm}
-We have decided that the point we were trying to raise is non-essential to the current manuscript, so we have removed the discussion of \emph{independent evolution} and its implication that $\varepsilon (t) = 0$ for all $t >0$.
+We have decided that the point we were trying to raise is not essential to the main manuscript, so we have moved the discussion of \emph{independent evolution} and its implication that $\varepsilon (t) = 0$ for all $t >0$ to Appendix B.
 The key issue we were trying to clarify is that the traditional joint model formulation in which the observed health process and time-to-event are conditionally independent given the latent process implies that the hazard function conditional on the entire observed health process depends on values after the current time.
@@
 This holds unless we assume the measurement error is set to zero.
-While we still think it is an important point that should be more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
-Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
+While we still think it is an important point that should be more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be left to the appendix.
+Moreover, the empirical impact seems to be limited and therefore it is better to make this point in less central to the current manuscript.
 \vspace{5mm}
@@
 \vspace{5mm}
-\textcolor{red}{
 We added Section 5.2 to discuss choice of window size; specifically, we consider selection based on Akaike information criteria as part of the simulation study rather than simply the impact of incorrect choice of $\Delta$.
 Table 4 shows that when inference is sensitive to bandwidth selection, the adaptive tuning of $\Delta$ based on AIC does perform well empirically.
-We now use this approach in selecting~$\Delta$ in the case study, and provide a plot of AIC across several choices of $\Delta$ in Figure XX in Appendix XX.}
+We now use this approach in selecting~$\Delta$ in the case study, and provide AIC and BIC across several choices of $\Delta$ in Table 6.
 \vspace{5mm}
@@
 \vspace{5mm}
-\textcolor{red}{To accommodate the second referee's request without making the discussion too long, we have added Remark XX and XX that provide some intuition.
-We have re-written the technical proofs to first provide a sketch proof (with intuition) and then a detailed proof.
-We hope this helps the reader develop intuition on the first read while giving them some helpful guidance in order to facilitate understanding of the detailed theoretical analysis.}
+To accommodate the second referee's request without making the discussion too long, we have re-written the technical proofs to include the intuition and integrated this with our detailed proof.
+We point the interested reader to these details in the main paper.
+We hope this helps the reader develop intuition on the first read while giving them some helpful guidance in order to facilitate understanding of the detailed theoretical analysis.
 \vspace{5mm}
@@
 \vspace{5mm}
-As far as the author is aware, there does not exist a joint modeling package that can estimate the proposed model directly; one way would be assume you know the functional form of the parameter but not the scale.
+As far as the author is aware, there does not exist a joint modeling package that can estimate the proposed model directly; one way would be to assume you know the functional form of the parameter but not the scale.
 Comparison of measures of bias, estimation accuracy, and numerical speed under this approach to our methodology which learns the functional form are likely inappropriate.
@@
 \vspace{5mm}
-We apologize for this omission. We have reread everything and addressed any missing notation and equation consistency.
+We apologize for this omission. We have reread everything and addressed any missing notation and equation consistency.  Moreover, we have added a notation glossary in the Appendix.
 \vspace{5mm}
@@
 \vspace{5mm}
-We have removed the discussion of \emph{independent evolution}.
-While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
-We compare our approach to Park and Staicu (2018) and report the smoothing of the covariance matrix as it comes from their prior work.
+We have moved the discussion of \emph{independent evolution} to the Appendix.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted from the main paper.
 \vspace{5mm}
@@
 
-\item {\it The measurement error part can be added to the end with other exten-
-sions. It does not seem to play a major role in developing the algorithm.}
+\item {\it The measurement error part can be added to the end with other extensions. It does not seem to play a major role in developing the algorithm.}
 
@@
 
-\item {\it In equationn (4), page 4, do you mean that equation (2) and the mesurement error part $x_i (t) = \eta_i (t) + \epsilon_i (t)$ imply conditional hazard equation in (3)?
+\item {\it In equation (4), page 4, do you mean that equation (2) and the mesurement error part $x_i (t) = \eta_i (t) + \epsilon_i (t)$ imply conditional hazard equation in (3)?
 If yes, then why is $\int_{t-\Delta}^t \epsilon_i (s) \beta(s) ds = 0$?
@@
 \vspace{5mm}
-We have now omitted this part from the current manuscript.
+We have now moved this part from the current manuscript to the Appendix.
 The key issue we were trying to clarify is that the traditional joint model formulation in which the observed health process and time-to-event are conditionally independent given the latent process implies that the hazard function conditional on the entire observed health process depends on values after the current time.
@@
 In this setting,~$\epsilon_i (t) = 0$ and so $\int_{t-\Delta}^t \epsilon_i (s) \beta(s)ds = 0$ by definition.
-While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
-Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be moved from the main paper to the Appendix.
+Moreover, the empirical impact seems to be limited and therefore it is better to make this point less central to the main manuscript.
 \vspace{5mm}
@@
 \vspace{5mm}
-\textcolor{red}{Both of these were defined in Section 3 but we agree that the definitions were obscured in the writing, which led to confusion.
-To address this, we have added a notation glossary as Appendix XX.}
+Both of these were defined in Section 3 but we agree that the definitions were obscured in the writing, which led to confusion.  To address this, we have added a notation glossary as Appendix A.
 \vspace{5mm}
@@
 We argue this is not sensible since current risk should not depend on future risk and thus set $\epsilon(t) = 0$.
+This discussion is now in Appendix B.
 \vspace{5mm}
@@
 It was used in equation (1) to refer to all the parameters underlying the model to distinguish it from the functional parameter~$\beta (s)$.
-We have added a notation glossary in Appendix XX, and re-iterate its definition in Section 3 to make this clearer.
+We have added a notation glossary in Appendix A, and re-iterate its definition in Section 3 to make this clearer.
 \vspace{5mm}
@@
 \vspace{5mm}
-We thank the referee for this suggestion.  We have now used Remark 2.1 and comment 8 on the computational costs to the end of section 2.2 to motivated our proposed method.
+We thank the referee for this suggestion.  We have now used Remark 2.1 and comment 8 on the computational costs to improve the end of section 2.2 and motivate our proposed method.
 \vspace{5mm}
@@
 Given the potential difference between the distributions, we choose to model the mean and covariance functions separately to ensure good approximation when applying the spectral decomposition of the estimated covariance function.
-\vspace{5mm}
-
-\textcolor{red}{Another reason for treating these two distinctly is that the missing data imputation from XXXX requires distinct covariance functions.
-We do think one could marginalize across the two if the mean and covariance functions are not expected to differ greatly and state as such in Section XX (in blue) where we discuss the new Figure XX of the mean and covariance functions for $y=0$ and $y=1$.}
+To illustrate this, Figure 2 in Appendix D.3 presents mean curves for EDA and AI over the same 30 minute window.
+These figures demonstrate quite distinct behavior.
+Another reason for treating these two distinctly is that the missing data imputation from Petrovich et al.
+(2018) requires distinct covariance functions.
+given the potential difference between the distributions, we choose to model the mean and covariance functions separately to ensure good approximation when applying the spectral decomposition of the estimated covariance function.
+To illustrate this, Figure 2 in Appendix D.3 presents mean curves for EDA and AI over the same 30 minute window.
 \vspace{5mm}
@@
 \vspace{5mm}
-\textcolor{red}{We have added mean and covariance functions to the supplementary materials as suggested.
-The plots demonstrate distinct behavior in the means for $y=0$ and $y=1$; while the covariance functions are XXXX.}
+Figure 2 in Appendix D.3 presents mean curves for EDA and AI over the same 30 minute window.
+The covariance functions did not plot well and given the point regarding need of modeling $y=0$ and $y=1$ separately is answered by the simpelr mean plot, we omitted the covariance function plots.
 \vspace{5mm}
@@
 
-\item {\it Outline the proof of Lemma 3.5 first and then prove the details.  It will be easier to follow.  Alternatively, we }
+\item {\it Outline the proof of Lemma 3.5 first and then prove the details.  It will be easier to follow.}
 
 \vspace{5mm}
-We have added a sketch of the proof with an outline as part of the supplementary details.
+We have added intuition to the detailed proofs where necessary and highlighted these edits in blue.
+We think this will help the reader understand the proof at a high-level and then the interested reader can delve into the technical details.
 \vspace{5mm}
@@
 We have now included the statement of how we define optimality in the main paper to make the theorem statement clear.
-Specifically, we view weights $w^\star$ as optimal if the difference between the asympotic variance~$V (\theta_0; w^\star)$ and the asymptotic variance under any other choice of weights~$w$, $V(\theta_0; w)$ is postivie semi-definite.
+Specifically, we view weights $w^\star$ as optimal if the difference between the asymptotic variance~$V (\theta_0; w^\star)$ and the asymptotic variance under any other choice of weights~$w$, $V(\theta_0; w)$ is positive semi-definite.
 The proof is kept in the supplementary materials for conciseness.
@@
 
-\item {\it Combine section 3.6 and 3.7 in to a single result that establishes a confidence band for $\beta (t)$.
+\item {\it Combine section 3.6 and 3.7 into a single result that establishes a confidence band for $\beta (t)$.
 Also, wont Lemma 3.5 lead to a confidence band for $\hat \beta (t)$?
@@
 We really appreciate this question which has helped clarify an important element of the paper.
-To address this question, we have added Section XX to evaluate the accuracy and time complexity.
+To address this question, we have added Remark 5.1 to evaluate the accuracy and time complexity.
 We recap some of our discussion here for convenience.
@@
 The above calculations hold for even the concurrent model, so that approach is still computationally expensive.
-To show this while ensuring no bias, we have added a simulation in Section XX where we show the computation trade-off and applying the concurrent model without subsampling.
+We discuss this futher in Remark 5.1 where we show the computation trade-off.
 We find the approach computationally demanding with minimal information gain.
@@
 
-\item Page 23, Figure 2. EDA seems to spike at the end. Even AI seems to be
-turning up. Add longer history to both the Figures.
+\item {\it Page 23, Figure 2. EDA seems to spike at the end. Even AI seems to be turning up. Add longer history to both the Figures.}
 \vspace{5mm}
@@
 Therefore, we keep the plot as is, and note that there is a spike at 30-minutes for EDA while emphasizing that most of the signal appears in the 15-minute range.
-This helps motivate our data adaptive choice of $\Delta$ in Section 7.1 which confirms that $\Delta = 5$ is the most appropriate choice although results are consistent across a range of $\Delta$ choices.
+This helps motivate our data adaptive choice of $\Delta$ in Section 7.1 which confirms that $\Delta = 15$ is the most appropriate choice although results are consistent across a range of $\Delta$ choices.
 \vspace{5mm}
@@
 A comparison of AIC provides empirical evidence that the smaller $\Delta$ was an appropriate choice.
-We now present $\Delta = 5$ in the main paper and move the current analysis to be a sensitivity analysis presented in the supplementary materials.
+We now present $\Delta = 15$ in the main paper, and then $\Delta =5$ and $30$ as a sensitivity analysis.
 \vspace{5mm}
@@
 \vspace{5mm}
-As the case study was meant to be illustrative, we only included time-of-day as a covariate.
-We now include the number of button presses in the prior 2 hours as another feature to control for the heterogeneity seen in the .
-These features are described in Section XX.
+As the case study was meant to be illustrative, we controlled for time-of-day by restricting to 9AM to 8PM as users who were asleep cannot press the button and thus were not 'at-risk'.
+We now include the number of button presses in the prior 12 hours as another feature to control for the heterogeneity.
+This is clarified in Section 7.1.
 \vspace{5mm}
@@
 We now state that our proposed joint model includes a recent history functional linear model which we now link with the previous work by Damla Senturk and others.
-This motivates Section XX which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
-We added Section XX to describe the related work and connect our methodology to the existing literature.
+This motivates Section 5.2 which provides additional empirical study and discussion of the tuning parameter~$\Delta$.
+We added Section 1.1 to describe the related work and connect our methodology to the existing literature.
 \vspace{5mm}
@@
 As with other tuning parameters, we take an AIC approach to selection of $\Delta$ in order to be data-adaptive.
-We study the performance of this criteria as part of the simulation study in Section XX and then apply it in the case study to determine the appropriate window-length.
+We study the performance of this criteria as part of the simulation study in Section 5.2 and then apply it in the case study to determine the appropriate window-length.
 \vspace{5mm}
@@
 \vspace{5mm}
-To address this and Referee \#1's comment \#X, we have added Remark XX to clarify settings where Assumption 3.4 will hold.
+To address this and the Referee \#1's comment, we have added Remark 4.5 to clarify settings where Assumption 3.4 will hold.
 Specifically, if the functional process is bounded, i.e., $|X(t,s)| < M$ and the other component is also bounded then we are guaranteed to have these assumptions hold.
@@
 By design, $\beta(s) = 0$ for $s > \Delta$.
-In practical settings we do not anticipate~$\beta(s)$ to be non-smooth and thus chose two scenarios where $\beta$ decays smoothly to 0.
-To address this, we have added a third scenario which has more rapid decay to zero; specifically, the coefficient function mimics taking a temporal difference over two average windows.
-To address the other comment, we have adjusted Table 2 to also report the relative values.
-This allows the reader to understand the MISE increases slowly and is mainly a consequence of the variance increases from reducing the subsampling rate.
+Therefore, we considered two settings for~$\beta(s)$.
+First, an exponential decaying $\beta(s)$ that mimics a weighted averaging over past values that decays to $0$ smoothly as $s \to^- \Delta$.
+Second, a sinusoidal $\beta(s)$ that mimics taking a temporal difference over two average windows but is different from $0$ near $s =\Delta$.
+We clarify these choices in Section 5.
+To address the other comment, we have added a row to Table 2 that reports relative MISE (RMISE).
 \vspace{5mm}
@@
 \vspace{5mm}
-We have decided that the point we were trying to raise is non-essential to the current manuscript, so we have removed the discussion of \emph{independent evolution} and its implication that $\varepsilon (t) = 0$ for all $t >0$.
+We have decided that the point we were trying to raise is not essential to the main manuscript, so we have moved the discussion of \emph{independent evolution} and its implication that $\varepsilon (t) = 0$ for all $t >0$ to the Appendix.
 The key issue we were trying to clarify is that the traditional joint model formulation in which the observed health process and time-to-event are conditionally independent given the latent process implies that the hazard function conditional on the entire observed health process depends on values after the current time.
@@
 The referee is correct that we will only compute conditional probabilities given the historical functional predictor; however, the likelihood factorization means that we implicitly are defining a conditional intensity given the entire functional predictor and the traditional approach has this undesirable dependence.
-While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted.
-Moreover, the empirical impact seems to be limited and therefore it is better to make this point in a different manuscript and case study.
+While the author still thinks it an important point that he wishes was more well appreciated in the joint modeling literature, the current paper has enough content for this discussion to be omitted from the main manuscript.
+Moreover, the empirical impact seems to be limited and therefore it is better to make this point in the Appendix.
 \vspace{5mm}
@@
 Since the event times~$T_i$ depend heavily on the event intensity function~$h_i (t;\theta)$ while the subsampling times~$D_i$ depend on the subsampling intensity function~$\pi_i(t)$, we chose to model the mean and covariance functions separately to ensure good approximation when applying the spectral decomposition of the estimated covariance function.
-For example
-\vspace{5mm}
-
-Another reason for treating these two distinctly is that the missing data imputation from XXXX requires distinct covariance functions.
-We do think one could marginalize across the two if the mean and covariance functions are not expected to differ greatly and state as such in Section XX (in blue) where we discuss the new Figure XX of the mean and covariance functions for $y=0$ and $y=1$.
+Given the potential difference between the distributions, we choose to model the mean and covariance functions separately to ensure good approximation when applying the spectral decomposition of the estimated covariance function.
+To illustrate this, Figure 2 in Appendix D.3 presents mean curves for EDA and AI over the same 30 minute window.
+Another reason for treating these two distinctly is that the missing data imputation from Petrovich et al.
+(2018) requires distinct covariance functions.
 \vspace{5mm}

### 2023-05-22 preprocessing/acc_preprocessing/bp_ai_nonevent.R
Nonevent mean function plot generators

@@ -1,0 +1,103 @@
+## Code for the mean AI plots 30 minuets backward from the button press
+library(tidyverse)
+library(lubridate)
+
+setwd('Z:/SI_data/R21_Study - ACC - AI/')
+
+# read in the button press data and transform the data type of the time stamp
+bp = readRDS('../R21_Study - tags/button_presses.RDS')
+bp = transmute(bp, ID, time=as_datetime(ts))
+
+nonevent_ids = readRDS('../acc_nonevent_complete_case_ids_Delta_30_2023-04-08.RDS')
+nonevent_times = readRDS('../acc_lin_nonevent_complete_case_timestamp_Delta_30_2023-04-08.RDS')
+nonevent_times = as_datetime(nonevent_times)
+
+
+#### Match button presses and the previous AI data ####
+# data frame to store the matched bp and AI
+if(!file.exists("../summary_data/bp_ai_nonevents.rds")) {
+  bp_ai = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, 
+                     Device=NULL, AI_num=NULL, AI_time=NULL, AI=NULL)
+  
+  n_obs = 91 # the largest number of observations (files)
+  for (i in 1:n_obs){
+    print(paste("Current participant", i))
+    id = 1000 + i
+    input_name = paste0('./', id, '_AI.rds')
+    
+    ## check the progress of the loop and save temporary result ##
+    if (i%%10 == 0){
+      print(i)
+      saveRDS(bp_ai, file = './methods/summary_data/bp_ai.rds')
+    }
+    
+    ## match bp and AI for this observation
+    if (file.exists(input_name)){
+      sub_nonevents_times = nonevent_times[nonevent_ids == id]
+      ai = readRDS(input_name)
+      device = unique(ai$Device)
+      
+      # process the data from different devices seperately due to the timestamp overlap 
+      for (d in 1:length(device)){
+        sub_ai =  filter(ai, Device==device[d])
+        for (j in 1:length(sub_nonevents_times)){
+          bp_time = sub_nonevents_times[j]
+          # AI_time in [BP_time-30min, BP_time)
+          ai_per_device_bp = filter(sub_ai, timestamp<bp_time, timestamp>=(bp_time-30*60))
[…]

### 2023-05-22 preprocessing/eda_preprocessing/bp_eda_nonevents.R

@@ -1,0 +1,118 @@
+## Code for the mean EDA plots backward from the button press
+library(tidyverse)
+library(lubridate)
+
+## WINDOWS
+setwd('Z:/SI_data/R21_Study - EDA - scaled/')
+## LINUX
+# setwd("/mnt/turbo/SI_data/R21_Study - EDA - scaled/")
+
+
+# read in the button press data and transform the data type of the timestamp
+bp = readRDS('../R21_Study - tags/button_presses.RDS')
+bp = transmute(bp, ID, time=as_datetime(ts))
+
+nonevent_ids = readRDS('../acc_nonevent_complete_case_ids_Delta_30_2023-04-08.RDS')
+nonevent_times = readRDS('../acc_lin_nonevent_complete_case_timestamp_Delta_30_2023-04-08.RDS')
+nonevent_times = as_datetime(nonevent_times)
+set.seed(1839713)
+## CHOOSING 1000 random times for speed issues
+
+#### Match button presses and the previous EDA (scaled) data ####
+if(!file.exists("../summary_data/bp_eda_nonevents.rds")) {
+  bp_eda = data.frame(ID=NULL, BP_num=NULL, BP_time=NULL, Device=NULL, 
+                      EDA_time=NULL, EDA=NULL, EDA_num=NULL)
+  
+  n_obs = 91 # the largest number of observations (files)
+  for (i in 1:n_obs){
+    print(paste("Current participant", i))
+    id = 1000 + i
+    input_name = paste0('./', id, '_EDA_scaled.rds')
+    
+    ## check the progress of the loop and save temporary result ##
+    if (i%%10 == 0){
+      print(i)
+      saveRDS(bp_eda, file = '../summary_data/bp_eda_nonevents.rds')
+    }
+    
+    ## match bp and EDA for this observation
+    ## observations 1010 and 1012 should be deleted since they do not have ACC data
+    if (file.exists(input_name) & i!=10 & i!=12){ 
+      sub_nonevents_times = nonevent_times[nonevent_ids == id]
+      sub_nonevents_times = sample(sub_nonevents_times, min(1000, length(sub_nonevents_times)))
+      eda = readRDS(input_name)
+      device = unique(eda$Device)
+      
+      # process the data from different devices seperately due to the timestamp overlap 
[…]

### 2023-05-22 write-up/fda-recurrent-supplement.tex
updated main and supp papers

@@
 \newfloatcommand{capbtabbox}{table}[][\FBwidth]
+\newenvironment{newmaterial}{\color{blue}}{\par}
 
@@
 
+\section{Notation Glossary}
+\label{app:notation}
+
+\begin{newmaterial}
+
+\begin{itemize}
+\item[${\bf T}_i$] \quad Observed event times on individual $i$.
+\item[$\tau_i$] \quad Observation window-length for individual $i$
+\item[$N_i(t)$] \quad Counting process associated with ${\bf T}_i$
+\item[${\bf x}_i$] \quad Health process for individual $i$, i.e., $\{ x_i (s) \}_{0 < s <\tau_i} \}$
+\item[$H_{i,t}^{NX}$] \quad The $\sigma$-field generated by all past values $(N_i(s), x_i(s))_{0 \leq s \leq t}$
+\item[$h_i (t| H_{i,t}^{NX}; \theta)$] \quad Fully parametric conditional hazard function
+\item[$h_0 (t; \gamma)$] \quad Parametrized baseline hazard function
+\item[$\Delta$] \quad An unknown, window-length
+\item[$\theta$] \quad All parameters underlying this hazard function including baseline parameters~$\gamma$, event-history and time-in-study parameters~$\alpha$, and the functional parameter $\{ \beta(s) \}_{0 \leq s \leq \Delta}$.
+\item[$H_i (t; \theta)$] \quad Cumulative hazard function
+\item[$U_n (\theta)$] \quad True score equations for log-likelihood of the event process
+\item[${\bf D}_i$] \quad Sampled non-event times on individual $i$.
+\item[$\hat H_i^{(1)} (t; \theta)$] \quad Waagepetersen estimator of the cumulative hazards using all $u \in {\bf T}_i \cup {\bf D}_{i}$
+\item[$\hat U_n (\theta)$] \quad Approximate estimating equations leveraging the Waagepetersen estimator
+\item[$\Sigma_y (s,s')$] \quad Marginal covariance over event or subsampling intensity $f_y (T)$
+\item[$\mu_y (t,s)$] \quad Mean function $s$ time units before time $t$
+\item[$\tilde U_n (\theta)$] \quad Approximate estimating equations leveraging both subsampling and functional data approximations
+\end{itemize}
+
+\end{newmaterial}
+
 \section{Measurement-error models with event processes}
@@
 
-\section{Derivation of the design-unbiased score equations}
+\section{Technical Arguments}
+
+\subsection{Derivation of the design-unbiased score equations}
 \label{app:prooflemma31}
@@
 
-\begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
+\subsection{Proof of Lemma~\ref{lemma:simpleasym}}
+\label{proof:asymp}
+
+\begin{newmaterial}
+We break the proof of Lemma~\ref{lemma:simpleasym} into two sections.
+First, we prove asymptotic consistency, i.e.,~$\hat \theta_n \to \theta_0$ in probability as $n \to \infty$ and $K_x \to \infty$.
+Then, we prove asymptotic normality via Rebolledo's martingale central limit theorem.
+\end{newmaterial}
+\begin{proof}[Proof of asymptotic consistency]
 % Lemma~\ref{lemma:simpleasym} will be proven under regularity conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the following assumptions related to the subsampling rate~$\pi_i (t)$.
@@
 \label{eq:decomp}
-n^{-1} \tilde U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
-  \hat U_n (\theta_0) - U_n (\theta) \right) + n^{-1} \left(\tilde U_n
-  (\theta_0) - \hat U_n (\theta_0) \right)
+n^{-1} \tilde U_n (\theta_0) = \underbrace{n^{-1} U_n (\theta_0 )}_{(I)}  + \underbrace{n^{-1} \left(
+  \hat U_n (\theta_0) - U_n (\theta) \right)}_{(II)} + \underbrace{n^{-1} \left(\tilde U_n
+  (\theta_0) - \hat U_n (\theta_0) \right)}_{(III)}
 \end{equation}
 where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations with $\int_0^\infty X_i(t,s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$ with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$.
+\textcolor{blue}{This decomposes the term into (I) the original score equations which we are known to yield asymptotically consistent estimates, (II) the difference between the original score equations and the design-unbiased score equations, and (III) the design-unbiased score equations with and without the approximation using the Karhunen-Loeve decomposition and~$K_x < \infty$.}
+
 \cite{Andersen1993} (Theorem VI.1.1) show that~$n^{-1} U_n (\theta_0)
@@
 Lenglart's inequality implies the second term converges in probability to zero under Assumptions~\ref{E1} and~\ref{E2}.
-The third term satisfies
+Term (II) satisfies
 \begin{align*}
@@
 expit function.
-The second inequality is due to the Taylor remainder theorem.
-The third is due to~$g^\prime (x) = e^x/(1+e^x)^2 \leq e^0/(1+e^0)^2 = 1/4$, reordering the summation, and rewriting the error term in terms of the difference between~$X(t,s)$ and the approximation with truncation level~$K_x$.
-Letting~$n \to \infty$ allows us to re-write the outside sums in terms of the integrated difference where the integral is with respect to the joint event and sampling distributions~$(f_0(t) + f_1 (t))$.
+\textcolor{blue}{The first inequality is simply due to bringing the metric inside the summand.
+This allows us to upper bound the difference as a sum over differences in the truncated representation and the true representation.}  The second inequality is due to the Taylor remainder theorem \textcolor{blue}{applied to this difference.
+This allows us to upper bound the difference as a product of the impact of the non-linear transformation $g$, and the error due to truncation.} The third is due to~$g^\prime (x) = e^x/(1+e^x)^2 \leq e^0/(1+e^0)^2 = 1/4$, reordering the summation, and rewriting the error term in terms of the difference between~$X(t,s)$ and the approximation with truncation level~$K_x$.
+Letting~$n \to \infty$,  \textcolor{blue}{we can rewrite $n^{-1} \sum_{i=1} \sum_{t \in {\bf T \cup D}}$ as an integral with respect to the joint event and sampling distributions~$(f_0(t) + f_1 (t))$}.
 Finally, under Assumptions~\ref{assumption:truncation}, \cite{Park2018} show
@@
 
+\subsection{Additional Figures}
+\label{app:addtlfigs}
+
+\begin{figure}[!th]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/smoothed_eda_nonevents.png}
+  \caption{Electrodermal activity (EDA)}
+  \label{fig:mean_eda_nonevents}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{../figures/smoothed_ai_nonevents.png}
+  \caption{Activity Index (AI)
+}  \label{fig:mean_acc_nonevents}
+\end{subfigure}
+\caption{Mean EDA and AI at non-event times.
+Note: Due to the data size, non-event times for EDA were further sub-sampled to be maximum 1000 non-event times per individual which leads to the increased variation relative to AI plot}
+\label{fig:mean_edacc_nonevents}
+\end{figure}
+
+\section{Code to Replicate Simulation and Case Study Results}
+The R code used to generate the simulation experiments and case study results in this paper can be obtained at \verb"https://github.com/wdempsey/fda-recurrentevents/".
+
 \newpage

### 2023-05-22 write-up/fda-recurrent.tex
updated main and supp papers

@@
 \title[Recurrent event analysis with functional covariates via random subsampling]{Recurrent event analysis in the presence of real-time high frequency data via random subsampling}
-% \author{Walter Dempsey}
-% \address {Department of Biostatistics, University of Michigan, 1415 Washington Heights, Ann Arbor, MI 48109, USA}
-% \email{wdem@umich.edu}
+\author{Walter Dempsey}
+\address {Department of Biostatistics, University of Michigan, 1415 Washington Heights, Ann Arbor, MI 48109, USA}
+\email{wdem@umich.edu}
 
@@
 Finally, traditional joint modeling often consider the hazard function to depend on the current or cumulative value of the health process~\citep{Li2022,Rizopoulos2010}.
-For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history and use recent history functional linear models~\cite{KIM20111554} to account for this local dependence.
+For high frequency sensor processes, we expect the hazard function to only depend on the recent sensor history and use recent history functional linear models~\citep{KIM20111554} to account for this local dependence.
 We combine these models with advances in longitudinal functional data analysis~\citep{Staicu2010} and penalized functional regression~\citep{Goldsmith2015} to perform scalable inference using off-the-shelf software.
@@
 Let~$N_{i} (t)$ denote the associated counting process of $\bfT_{i}$; that is, $N_i (t) = \sum_{j=1}^{k_i} 1 [ T_{i,j} < t ]$.
-In this section, we assume a single-dimensional health process~$\bfx_i = \{ x_i (s) \}_{0 < s < \tau_i}$ for each participant is measured at a dense grid of time points.
+In this section, we assume a single-dimensional health process~$\bfx_i = \{ x_i (s) \}_{0 \le s < \tau_i}$ for each participant is measured at a dense grid of time points.
 Accelerometer, for example, is measured at a rate of 32Hz (i.e., $32$ times per second).
@@
 Given the high frequency nature of sensor data, this paper assumes the process is measured continuously.
+\textcolor{blue}{See Appendix~\ref{app:notation} for a notation glossary.}
 
@@
 For the sake of notational simplicity, we leave the dependency of the conditional hazard function on~$H_{i,t}^{NX}$ implicit, and write~$h_i (t; \theta)$.
-The component of the log-likelihood related to the event process is then given by
+\textcolor{blue}{In our current setting, we assume the health process $\{ x_i(s)\}_{0 \leq s < \tau_i}$ is directly observed.
+Therefore, we can consider the event process~$\bfT_i$ conditional on the health process~$\bfx_i$, which results in the log-likelihood related to the event process being then given by}
 \[
@@
 where~$H_{i} (\tau_i ; \theta) = \int_{0}^{\tau_i} h_{i} (t; \theta) dt$ is the cumulative hazard function.
-Solving the associated score equations~$U_n (\theta) = {\bf 0}$ yields the maximum likelihood estimator~$\hat \theta$, where
+\textcolor{blue}{See Appendix~\ref{section:memproblems} for additional arguments in favor of our proposed approach.} Solving the associated score equations~$U_n (\theta) = {\bf 0}$ yields the maximum likelihood estimator~$\hat \theta$, where
 \[
@@
 
-\begin{rmk}[Connection to design-based estimation]
-The approximate score equations given by~\eqref{eq:approxscore} arise in design-based inference of point processes.
+% \begin{rmk}[Design-based estimation]
+\textcolor{blue}{
+To circumvent these computational difficulties, we will derive approximate score equations based on design-based inference of point processes.
 Design-based inference is common for spatial point processes~\citep{Waagepetersen2008} where the spatial varying covariate is observed at a random sample of locations.
 It is common in mobile health where ecological momentary assessments~\citep{Rathbun2012,Rathbun2016} are used to randomly sample individuals at various time-points to assess their emotional state.
-In the current setting, we leverage these ideas to form a subsampling protocol that can substantially reduce computationally complexity.
+In the current setting, we will leverage these ideas to form a subsampling protocol that can substantially reduce computationally complexity.
 Therefore, the purpose is quite different.
-Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.
-\end{rmk}
+Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.}
+% \end{rmk}
 
@@
 This estimator depends on the superposition of the event and subsampling processes.
-Proposition~\ref{prop:optimal} shows the estimator for~$\theta$ associated with using~\eqref{eq:WPest} is the most efficient within a suitable class of estimators for the derivative of the cumulative hazard function (including the Horvitz-Thompson estimator).
+Lemma~\ref{prop:optimal} shows the estimator for~$\theta$ associated with using~\eqref{eq:WPest} is the most efficient within a suitable class of estimators for the derivative of the cumulative hazard function (including the Horvitz-Thompson estimator).
 Therefore, we restrict our attention to~\eqref{eq:WPest} for the remainder of this paper.
@@
 \]
-Due to our concern over independent evolution as discussed in section~\ref{section:memproblems}, we do not assume that each observation is observed with white noise and therefore the diagonal elements of~$\hat \Sigma_0$ are not inflated.
+\textcolor{blue}{As we assume the health process~$\bfx_i$ is directly observed}, the diagonal elements of~$\hat \Sigma_0$ are not inflated.
 Second, the estimator~$\hat \Sigma$ is further smoothed again using the sandwich smoother~\citep{Xiao2013}.
@@
 \begin{rmk}[Inference procedure review] \normalfont
+\label{rmk:review}
 Given observed recurrent event and high frequency data~$\{ \bfT_i, \bfx_i \}_{i=1}^n$,
@@
 \item \label{p3} Compute marginal covariance across event times,~$\Sigma_1$, and non-event times,~$\Sigma_0$.
-\item \label{p4} Compute eigendecomposition~$\{ \hat \psi_k^{(y)}, \hat \lambda_k^{()} \}$ of marginal covariance~$\Sigma_y$
+\item \label{p4} Compute eigendecomposition~$\{ \hat \psi_k^{(y)}, \hat \lambda_k^{(y)} \}$ of marginal covariance~$\Sigma_y$
 \item Use the eigendecomposition to construct~$W_{i,t}$ for all $i \in [n]$ and $t \in \bfD_i \cup \bfT_i$
@@
 We state these assumptions and then our main theorems.
-We start by assuming there exists a~$\tau < \infty$ such that all individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
+We start by assuming there exists~$\tau < \infty$ such that all individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
 Moreover, define $R_i (t)$ to be the at-risk indicator for participant~$i$, i.e., $R_i (t) = {\bf 1} [t \in (0,\tau_i)]$.
@@
 Finally, for simplicity, we assume that there exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the true function~$\beta(t)$ sits in the span of the spline basis expansion.
+
+\begin{newmaterial}
 \begin{rmk}[Practical consequences of Assumptions~\ref{assumption:events} and~\ref{assumption:truncation}]
-Bounded and iid.
+Assumptions~\ref{assumption:events} and~\ref{assumption:truncation} contain as a special case the scenario where individuals are independent and identically distributed, the functional process is bounded (i.e., $|X(t,s)| < M$ for some $M <\infty$), and the subsampling rate is both lower and upper bounded at all risk times.
+As such bounds are likely to be true in most practical settings, this demonstrates the reasonableness of our assumptions for applied settings.
 \end{rmk}
-
-With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
-
+\end{newmaterial}
+
+% With the necessary assumptions stated, we can now state Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
 \begin{lemma} \normalfont
@@
 \[
-\sqrt{n} (\hat \theta - \theta) \to N(0, \Xi (\theta)^{-1})
+\sqrt{n} (\hat \theta - \theta) \overset{D}{\to} N(0, \Xi (\theta)^{-1}),
 \]
-where
+where $\overset{D}{\to}$ is convergence is distribution and
 \[
@@
 \end{lemma}
-A design-unbiased estimator for~$\Xi(\theta)$ is
+\textcolor{blue}{Proof of Lemma~\ref{lemma:simpleasym} is presented in Appendix~\ref{proof:asymp}.} A design-unbiased estimator for~$\Xi(\theta)$ is
 \begin{equation}
@@
 \end{lemma}
-
-\noindent
+\noindent \textcolor{blue}{Proof of Lemma~\ref{prop:optimal} is presented in Appendix~\ref{app:optimalweights}.}
 
@@
 
-Under assumption~\ref{assumption:truncation}, we next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
+We next consider the statistical efficiency of our proposed estimator when compared to complete-data maximum likelihood estimation.
 While subsampling introduces additional variation, it may significantly reduce the overall computational burden.
@@
 & 100 & 288 & 144 & 48 & 29 & 14 & 0.990 \\ \hline
-\multirow{3}{*}{32Hz (ACC)}
+\multirow{3}{*}{32Hz (AI)}
 & 5 & 46080 & 23040 & 7680 & 4608 & 2304 & 0.833 \\
@@
 However, in practice, such a choice of~$K_b$ may lead to overfitting the data.
-Following~\cite{Goldsmith2011}, we choose the polynomial spline model and set~$\phi(t) = b_0 + b_1 t + b_2 t^2 + \sum_{j=3}^{K_b} (t-\kappa_j)^3$ where~$\{ \kappa_j \}_{j=3}^{K_b}$ are the chosen knots and assume~$\{ b_j \}_{j=3}^{K_b} \sim N(0, \sigma^2 I)$ to induce smoothness on the spline model.
+Following~\cite{Goldsmith2011}, we choose the linear spline model and set~$\phi(t) = b_0 + b_1 t + \sum_{j=2}^{K_b-1} (t-\kappa_j)_{+}$ where~$\{ \kappa_j \}_{j=2}^{K_b-1}$ are the chosen knots and assume~$\{ b_j \}_{j=2}^{K_b-1} \sim N(0, \sigma^2 I)$ to induce smoothness on the spline model.
 Combining the penalized spline formulation with Lemma~\ref{lemma:logistic} establishes a connection between our approximate score equations and solving a generalized mixed effects logistic regression with offset.
@@
 Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biased as well as the subsampling variance.
-To allow fair comparisons across the two choices of $\beta(s)$, all reported numbers are scaled by the integrated square of the true function $\int_{0}^{\infty} \beta(s)^2 ds$.
-The average runtime (in seconds) is also reported.
+\textcolor{blue}{To allow fair comparisons across the two choices of $\beta(s)$, all reported numbers are scaled by the integrated square of the true function $\int_{0}^{\infty} \beta(s)^2 ds$.
+The relative MISE (RMISE) to the lowest sampling rate and the average runtime (in seconds) are also reported.}
 
@@
 
-While above we explore the bias-variance trade-off under bandwidth misspecification, here we explore a data-adaptive method for bandwidth selection, a critical consideration for recent history functional linear models~\citep{KIM20111554}.
-Specifically, given the model complexity does not change as a function of $\Delta$, our modest proposal is to compare AIC across a range of bandwidths and choose that which maximizes the criterion.
+While above we explore the bias-variance trade-off under bandwidth misspecification, here we explore a data-adaptive method for bandwidth selection.
+This is a critical consideration for recent history functional linear models~\citep{KIM20111554}.
+Given the model complexity does not change as a function of $\Delta$ in our simulations, our proposal is to compare AIC across a range of bandwidths and choose that which maximizes the criterion.
 Table~\ref{tab:aic_selection} presents AIC-based selection across the 500 simulated datasets analyzed in the prior section.
@@
 \begin{table}[!th]
-\begin{tabular}{c | c c c c | c c c c}
+\begin{tabular}{l | c c c c | c c c c}
 & \multicolumn{4}{c}{Exponential} & \multicolumn{4}{c}{Sinusoidal} \\ \hline
@@
 29 & 0.22 & 0.20 & 0.17 & 0.12 & 0.09 & 0.14 & 0.15 & 0.18 \\
-32 & 0.11 & 0.12 & 0.13 & 0.18 & {\bf 0.85} & {\bf 0.79} & {\bf 0.76} & {\bf 0.70} \\
+32$^\star$ & 0.11 & 0.12 & 0.13 & 0.18 & {\bf 0.85} & {\bf 0.79} & {\bf 0.76} & {\bf 0.70} \\
 35 & 0.13 & 0.13 & 0.14 & 0.15 & 0.05 & 0.06 & 0.06 & 0.09 \\
@@
 
-\subsection{Computational comparison to alternative methods} \label{section:simcomparisons}
+% \subsection{Computational comparison to alternative methods} \label{section:simcomparisons}
 
@@
 between~$X(s^\prime,t)$ and $X(s,t)$ for $s^\prime, s \in [0, \Delta]$
-and~$t \in \mathbb{R}_+$. 
+and~$t \in \mathbb{R}_+$. \textcolor{blue}{Note that~\cite{Petrovich2018} requires modeling the mean and covariance functions separately for $y=0$ and $y=1$.}
 %To account for overlap between $\Delta$-windows for $t \in \bfT_i \cup \bfD_i$ - i.e., avoid imputing values at a particular time to different values -- imputation is performed sequentially over this set.
@@
 This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
-To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.
+\textcolor{blue}{To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.}
 % Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
@@
 
-Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{activity index}~\citep{10.1371/journal.pone.0160644}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
+Here, we focus on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{activity index} (AI)~\citep{10.1371/journal.pone.0160644}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.
 Electrodermal activity can be significantly impacted by external factors (e.g., room temperature).
@@
 The individual EDA and AI trajectories are highly variable which tends to obscure patterns and trends.
-In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and AI are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 10 minutes prior to button presses.
+In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and AI are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 30 minutes prior to button presses.
+\textcolor{blue}{In Figure~\ref{fig:mean_edacc_nonevents} in Appendix~\ref{app:addtlfigs}, the mean trajectories of EDA and AI are plotted in reverse time for the sampled non-event times in the 30 minutes prior to the non-event time.
+The distinct mean behavior motivates our desire to model these two processes separately as discussed in Remark~\ref{rmk:review}.}
 
@@
 
-Inspection of Figures~\ref{fig:mean_eda} and~\ref{fig:mean_acc} suggest setting $\Delta = 30$ minutes is adequate to capturing the proximal impact of EDA and AI on the risk of a button press.
+\begin{newmaterial}
+
+Inspection of Figures~\ref{fig:mean_eda} and~\ref{fig:mean_acc} suggest setting $\Delta$ between $5$ and $30$ minutes.
+Here, we investigate three potential window lengths, $\Delta = 5, 15,$ and $30$.
 To ensure minimal loss of efficiency, the subsampling rate was set to once every fifteen minutes.
@@
 After sampling non-event times, complete-case analyses are performed, i.e., sampled times where sensors included in the model have any level of missing data are ignored.
+Table~\ref{tab:aicbic} presents both AIC and BIC criteria on the complete-case data, normalized by the number of observations to ensure fair comparison.
+We find that $\Delta = 15$ is adequate to capturing the proximal impact of EDA and AI on the risk of a button press.
 
@@
 Delta & 5 & 15 & 30 \\ \hline
-AIC & - & - & - \\
-BIC & - & - & - \\ \hline
+AIC & 0.96 & 0.98 & 0.95 \\
+BIC & 0.99 & 1.01 & 0.99 \\ \hline
 \end{tabular}
+\caption{Normalized AIC and BIC for different choices of $\Delta$ using complete-case data.
+For $\Delta = 5$, $K_b = K_x = 31$ due to amount of data in shorter window-length, while $K_b=K_x = 35$ for $\Delta = 15$ and $30$.}
+\label{tab:aicbic}
 \end{table}
 
-We analyzed activity index (AI) and electrodermal activity separately as given in equation~~\eqref{eq:hazardlinear} as well as jointly in a multivariate model as in equation~\eqref{eq:multihazardlinear}.
-Results did not change substantially.
-Figure~\ref{fig:eda_estimate} and~\ref{fig:acc_estimate} presents the estimates from the joint analysis with their associated 95\% confidence intervals.
-We highlight in gray the statistically significant regions.
-Here, we see that standardized EDA is not associated with increased risk of button press, while activity index sees a positive association in the final few minutes prior to a button press.
-In section~\ref{sec:missingdata_heterogeneity}, we investigate whether the results are sensitive to missing data.
- % and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
+In Section~\ref{app:csmissing} of the supplementary materials, Figures~\ref{fig:missing_eda} and~\ref{fig:missing_acc} present the per-patient average fraction with missing EDA and AI in 30-minute windows respectively.
+For EDA, we have a wide range of missingness -- from $0\%$ to over $40$\% across individuals with most between $5\%$ and $30\%$ average fraction missing.
+For AI, the missingness is less pronounced -- from $0\%$ to $15\%$ across individuals with most between $0\%$ and $10\%$ average fraction missing.
 
@@
 
+We analyzed activity index (AI) and electrodermal activity jointly in a multivariate model as in equation~\eqref{eq:multihazardlinear}.
+To account for times when the participants may not be wearing the device and thus not be \emph{at-risk} for a button press, we limited the analysis to data collected between 9AM and 8PM.
+In the analysis, we assumed a constant baseline hazard and included a binary indicator of whether the participant had an event in the past 12 hours to account for patient heterogeneity in the number of events as seen in Figure~\ref{fig:buttons}.
+Figure~\ref{fig:eda_estimate} and~\ref{fig:acc_estimate} presents the estimates from the joint analysis and their associated 95\% confidence intervals using the bootstrap and multiple imputation strategy.
+We highlight in gray the statistically significant regions.
+Here, we see that standardized EDA is not associated with increased risk of button press, while activity index sees a positive association in the final few minutes prior to a button press.
+
+\end{newmaterial}
+ % and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
+
  \subsection{Sensitivity analysis for window length}
@@
 
-In Section~\ref{app:csmissing} of the supplementary materials, Figures~\ref{fig:missing_eda} and~\ref{fig:missing_acc} present the per-patient average fraction with missing EDA and AI in 30-minute windows respectively.
-For EDA, we have a wide range of missingness -- from $0\%$ to over $40$\% across individuals with most between $5\%$ and $30\%$ average fraction missing.
-For AI, the missingness is less pronounced -- from $0\%$ to $15\%$ across individuals with most between $0\%$ and $10\%$ average fraction missing.
-To assess the impact of missing data, we perform the imputation under uncongeniality using the boostrap and multiple imputation strategy described in Section~\ref{section:miuncong}.
-Here we see that the EDA curve remains non-significant while the region where the AI curve is significant widens slightly to be the prior 4 minutes.
+\begin{newmaterial}
+
+ We next investigate whether the results are sensitive to window length.
+Specifically, we re-analyzed activity index (AI) and electrodermal activity jointly in a multivariate model as in equation~\eqref{eq:multihazardlinear} with $\Delta = 5$ and $30$ minutes.
+Figure~\ref{fig:acc_estimate_bs_delta5} and~\ref{fig:acc_estimate_bs_delta30} presents the estimates from the joint analysis for the activity index (AI) and their associated 95\% confidence intervals using the bootstrap and multiple imputation strategy.
+We highlight in gray the statistically significant regions which remain similar across all three choices of window length.
+We do not present results for the standardized EDA as it continues to be not associated with increased risk of button press.
+ \end{newmaterial}
 
@@
 \end{subfigure}
-\caption{$\beta (t)$ for ACC and EDA with 95\% CI; solid line is point estimate for the missing data analysis, while highlighted region are the boostrap and MI based pointwise 95\% confidence intervals.}
+\caption{$\beta (t)$ for AI and EDA with 95\% CI; solid line is point estimate for the missing data analysis, while highlighted region are the boostrap and MI based pointwise 95\% confidence intervals.}
 \label{fig:edacc_estimate_bs}

### 2023-05-22 preprocessing/acc_preprocessing/bp_ai.R
Update bp_ai.R

@@ -62,7 +62,7 @@ bp_ai_summary = bp_ai %>%
 max(bp_ai_summary$AI_entries) # should be no larger than 180 = 30*60/epch
 
 #### Draw the mean plot and the boxplot ####
-bp_ai = bp_ai_read %>%
+bp_ai = bp_ai %>%
   group_by(ID, BP_num, Device) %>%
   mutate(time_to_bp = max(AI_num)+1-AI_num) %>%
   ungroup() %>%

### 2023-05-22 methods/likelihood_estimation.R
small edit to the main estimation functions

@@ -356,4 +356,11 @@ for(Delta in all_Deltas) {
 saveRDS(object = bindeviance, file = "C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/methods/deviance.RDS")
 saveRDS(object = lengthY, file = "C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/methods/numobs.RDS")
 
-(bindeviance + c(31,35,35)*log(lengthY))/lengthY
+
+# bindeviance = readRDS(file = "C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/methods/deviance.RDS")
+# lengthY =readRDS(file = "C:/Users/Balthazar/Documents/GitHub/fda-recurrentevents/methods/numobs.RDS")
+
+
+BIC = (2*bindeviance + c(31,35,35)*log(lengthY))/lengthY
+
+AIC = (2*bindeviance + 2* c(31,35,35))/lengthY

### 2023-10-10 letters/cover letter-fdarecurrent-rd2.tex
Update the files for accepted version

--- 2023-02-02 letters/cover letter-fdarecurrent.tex
@@
 
-\opening{Dear Drs. Jones and Liang:}
+\opening{Dear Dr. Liang:}
 
-Kindly find enclosed the revised version of the manuscript ID JCGS-22-236, "Recurrent event analysis in the presence of real-time high frequency data via random subsampling."   We have revised the article to address the referees’ requests.
-The reports called for a major revision to include related work on modeling event data with functional data predictors as well as historical functional linear models, include empirical evidence and associated discussion of the computational benefits of our proposal, as well as clarify several technical details in the theoretical sections and provide important intuition/sketches of the proofs.
-The new version complies with these requests.
+We thank you for the positive news regarding our manuscript ID JCGS-22-236, "Recurrent event analysis in the presence of real-time high frequency data via random subsampling."   We have revised the article to address the referee's remaining comment.
+The report had a final comment about Section 5.2.
+Specifically, it asked about the suitability of relative MISE as a more useful indicator when selecting Delta rather than AIC.
+The new version complies with this request by adding a sentence regarding this issue in Section 5.2.
 
-We believe the work presents an important advance that gives researchers a simple tool to analyze recurrent event data in the presence of high-frequency sensors data.
-Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.
-
-Thank you for your consideration. We look forward to hearing back from you.
+Thank you again for the positive news regarding our manuscript.  We look forward to seeing our work in the Journal of Computational and Graphical Statistics.
 

### 2023-10-10 write-up/fda-recurrent.tex
Update the files for accepted version

@@
 
-\begin{newmaterial}
+% \begin{newmaterial}
 The use of wearable devices to passively monitor patients has led to a rapid increase in the number of studies with high-frequency sensors that can be conceptualized as studies where measurements are functions.
@@
 
-\end{newmaterial}
+% \end{newmaterial}
 
@@
 Given the high frequency nature of sensor data, this paper assumes the process is measured continuously.
-\textcolor{blue}{See Appendix~\ref{app:notation} for a notation glossary.}
+See Appendix~\ref{app:notation} for a notation glossary.
 
@@
 For the sake of notational simplicity, we leave the dependency of the conditional hazard function on~$H_{i,t}^{NX}$ implicit, and write~$h_i (t; \theta)$.
-\textcolor{blue}{In our current setting, we assume the health process $\{ x_i(s)\}_{0 \leq s < \tau_i}$ is directly observed.
-Therefore, we can consider the event process~$\bfT_i$ conditional on the health process~$\bfx_i$, which results in the log-likelihood related to the event process being then given by}
+In our current setting, we assume the health process $\{ x_i(s)\}_{0 \leq s < \tau_i}$ is directly observed.
+Therefore, we can consider the event process~$\bfT_i$ conditional on the health process~$\bfx_i$, which results in the log-likelihood related to the event process being then given by
 \[
@@
 where~$H_{i} (\tau_i ; \theta) = \int_{0}^{\tau_i} h_{i} (t; \theta) dt$ is the cumulative hazard function.
-\textcolor{blue}{See Appendix~\ref{section:memproblems} for additional arguments in favor of our proposed approach.} Solving the associated score equations~$U_n (\theta) = {\bf 0}$ yields the maximum likelihood estimator~$\hat \theta$, where
+See Appendix~\ref{section:memproblems} for additional arguments in favor of our proposed approach.
+Solving the associated score equations~$U_n (\theta) = {\bf 0}$ yields the maximum likelihood estimator~$\hat \theta$, where
 \[
@@
 % \begin{rmk}[Design-based estimation]
-\textcolor{blue}{
 To circumvent these computational difficulties, we will derive approximate score equations based on design-based inference of point processes.
@@
 Therefore, the purpose is quite different.
-Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.}
+Moreover, the dependence of the intensity function on the recent history of sensor values leads to additional complications that must be addressed.
 % \end{rmk}
@@
 An estimator is design-unbiased if its expectation is equal to that parameter under the probability distribution induced by the sampling design~\citep{Cassel1977}.
-Let~$\bfD_i \subset [0,\textcolor{blue}{\tau_i}]$ denote the random set of subsampled points.
+Let~$\bfD_i \subset [0,\tau_i]$ denote the random set of subsampled points.
 Note, by construction, this random set is distinct from the set of event times with probabilty one, i.e.,~$\pr( \bfT_i \cap \bfD_i = \emptyset) = 1$.
@@
 \label{sec:margcov}
-\textcolor{blue}{
-We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.} The sandwich smoother~\citep{Xiao2013} is used to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
+We start by defining~$X(t,s) = x(t-s)$ to be the sensor measurement~$0 \leq s \leq \Delta$ time units prior to time~$t \in \bfT_i \cup \bfD_i$.
+The sandwich smoother~\citep{Xiao2013} is used to estimate the mean~$\mu_y(t,s) = \E_y [ X(t,s)]$ where the expectation is indexed by whether $t$ is an event ($y=1$) or subsampled ($y=0$) time respectively.
 Alternative bivariate smoothers exist, such as the kernel-based local linear smoother~\citep{Hastie2009}, bivariate tensor product splines~\citep{Wood2006}, and the bivariate penalized spline smoother~\citep{MarxEilers2005}.
@@
 \]
-\textcolor{blue}{As we assume the health process~$\bfx_i$ is directly observed}, the diagonal elements of~$\hat \Sigma_0$ are not inflated.
+As we assume the health process~$\bfx_i$ is directly observed, the diagonal elements of~$\hat \Sigma_0$ are not inflated.
 Second, the estimator~$\hat \Sigma$ is further smoothed again using the sandwich smoother~\citep{Xiao2013}.
@@
 
-\begin{newmaterial}
+% \begin{newmaterial}
 \begin{rmk}[Practical consequences of Assumptions~\ref{assumption:events} and~\ref{assumption:truncation}]
@@
 \end{rmk}
-\end{newmaterial}
+% \end{newmaterial}
 
@@
 \end{lemma}
-\textcolor{blue}{Proof of Lemma~\ref{lemma:simpleasym} is presented in Appendix~\ref{proof:asymp}.} A design-unbiased estimator for~$\Xi(\theta)$ is
+Proof of Lemma~\ref{lemma:simpleasym} is presented in Appendix~\ref{proof:asymp}. A design-unbiased estimator for~$\Xi(\theta)$ is
 \begin{equation}
@@
 This implies that subsampling from an inhomogeneous Poisson process, standard logistic regression software can be used to fit the recurrent event model by specifying an offset equal to~$\log \pi_i (t)$.
-\begin{newmaterial}
+% \begin{newmaterial}
 Based on this, we can leverage existing inferential machinery to obtain variance-covariance estimates of model parameters.
@@
 Here, weights $w^\star$ are considered \emph{optimal} if the difference between the asympotic variance~$V (\theta_0; w^\star)$ and the asymptotic variance under any other choice of weights~$w$, $V(\theta_0; w)$ is positive semi-definite, i.e., any linear contrast has smaller asymptotic variance under weight $w^\star$ than under weight~$w$.
-\end{newmaterial}
+% \end{newmaterial}
 \begin{lemma} \normalfont
@@
 \end{lemma}
-\noindent \textcolor{blue}{Proof of Lemma~\ref{prop:optimal} is presented in Appendix~\ref{app:optimalweights}.}
+Proof of Lemma~\ref{prop:optimal} is presented in Appendix~\ref{app:optimalweights}.
 
@@
 Given the connection with generalized mixed effects models, the inferential machinery to obtain variance-covariance estimates.
-\textcolor{blue}{As we leverage the standard R package `glmnet', the smoothness parameter is chosen via cross-validation.} In this context, we acknowledge another limitation.
+As we leverage the standard R package `glmnet', the smoothness parameter is chosen via cross-validation.
+In this context, we acknowledge another limitation.
 Specifically, penalization may lead to confidence intervals that perform poorly in regions where~$\hat \beta(t)$ is over-smoothed.
@@
 We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day.
-We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $ \textcolor{blue}{which decays to 0 as $s$ approaches $\Delta$ from below}, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$ \textcolor{blue}{which is significantly different from 0 as $s$ approaches $\Delta$ from below.}
+We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $ which decays to 0 as $s$ approaches $\Delta$ from below, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$ which is significantly different from 0 as $s$ approaches $\Delta$ from below.
 
@@
 Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biased as well as the subsampling variance.
-\textcolor{blue}{To allow fair comparisons across the two choices of $\beta(s)$, all reported numbers are scaled by the integrated square of the true function $\int_{0}^{\infty} \beta(s)^2 ds$.
-The relative MISE (RMISE) to the lowest sampling rate and the average runtime (in seconds) are also reported.}
+To allow fair comparisons across the two choices of $\beta(s)$, all reported numbers are scaled by the integrated square of the true function $\int_{0}^{\infty} \beta(s)^2 ds$.
+The relative MISE (RMISE) to the lowest sampling rate and the average runtime (in seconds) are also reported.
 
@@
 
-\begin{newmaterial}
+% \begin{newmaterial}
 \begin{rmk}[Time-complexity and run-time of maximum likelihood estimation]
@@
 
-\begin{newmaterial}
+% \begin{newmaterial}
 \subsection{Selection of bandwidth} \label{section:bandwidth}
@@
 Thus, we conclude that the AIC-based selection method works well in settings where inference depends heavily on estimating the bandwidth accurately.
+Note that when $\Delta$ is inaccurately selected as in the exponential case, the relative MISE when Delta is selected by AIC is a more useful indicator of AIC.
 
@@
 between~$X(s^\prime,t)$ and $X(s,t)$ for $s^\prime, s \in [0, \Delta]$
-and~$t \in \mathbb{R}_+$. \textcolor{blue}{Note that~\cite{Petrovich2018} requires modeling the mean and covariance functions separately for $y=0$ and $y=1$.}
+and~$t \in \mathbb{R}_+$. Note that~\cite{Petrovich2018} requires modeling the mean and covariance functions separately for $y=0$ and $y=1$.
 %To account for overlap between $\Delta$-windows for $t \in \bfT_i \cup \bfD_i$ - i.e., avoid imputing values at a particular time to different values -- imputation is performed sequentially over this set.
@@
 The main disadvantage is that the above framework is uncongenial under many events and/or high sampling rates.
-\textcolor{blue}{\cite{Meng1994} defines congeniality between an imputation procedure and an analyst’s complete (and incomplete) analysis procedure if there exists a unifying Bayesian model which embeds the imputer’s imputation model and the analyst’s complete data procedure.
+\cite{Meng1994} defines congeniality between an imputation procedure and an analyst’s complete (and incomplete) analysis procedure if there exists a unifying Bayesian model which embeds the imputer’s imputation model and the analyst’s complete data procedure.
 A recent discussion of congeniality can be found Bartlett and Hughes (2020).
 Congeniality ensures good frequentist coverage properties.
-In some uncongenial settings, standard multiple imputation can be biased downwards leading to under-coverage of confidence intervals.}    A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
+In some uncongenial settings, standard multiple imputation can be biased downwards leading to under-coverage of confidence intervals.
+A key question is whether we can use the above imputation methods within a general procedure to handle uncongeniality.
 
@@
 This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
-\textcolor{blue}{To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.}
+To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.
 % Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.
@@
 In Figure~\ref{fig:mean_edacc}, the mean trajectories of EDA and AI are plotted in reverse time from button press timestamps, which shows sharp changes in EDA and AI in the 30 minutes prior to button presses.
-\textcolor{blue}{In Figure~\ref{fig:mean_edacc_nonevents} in Appendix~\ref{app:addtlfigs}, the mean trajectories of EDA and AI are plotted in reverse time for the sampled non-event times in the 30 minutes prior to the non-event time.
-The distinct mean behavior motivates our desire to model these two processes separately as discussed in Remark~\ref{rmk:review}.}
+In Figure~\ref{fig:mean_edacc_nonevents} in Appendix~\ref{app:addtlfigs}, the mean trajectories of EDA and AI are plotted in reverse time for the sampled non-event times in the 30 minutes prior to the non-event time.
+The distinct mean behavior motivates our desire to model these two processes separately as discussed in Remark~\ref{rmk:review}.
 
@@
 
-\begin{newmaterial}
+% \begin{newmaterial}
 
@@
 
-\end{newmaterial}
+% \end{newmaterial}
  % and whether individual-level coefficients,~$\beta(t) + b_i (t)$, effects show stronger dependence on EDA and/or AI for certain individuals.
@@
 We do not present results for the standardized EDA as it continues to be not associated with increased risk of button press.
- \end{newmaterial}
+ % \end{newmaterial}
 

### 2023-10-10 letters/cover letter-fdarecurrent-rd2.tex
Update cover letter-fdarecurrent-rd2.tex

@@
 
+The source files for the manuscript and supplementary materials are contained in two zip/tar files as requested.  Below is a list of the files names with a brief summary as requested:\\
+
+{\bf Manuscript zip file:} \emph{fda-recurrent-manuscript.zip}
+\begin{itemize}
+    \item \emph{fda-recurrent.tex} : Manuscript TeX File
+    \item \emph{si-fda-refs.bib} : Manuscript Bib File
+    \item \emph{./figures}: All necessary figures
+\end{itemize}
+{\bf Supplementary materials zip file:} \emph{fda-recurrent-manuscript.zip}
+\begin{itemize}
+    \item \emph{fda-recurrent-supplementary.tex} : Supplementary Materials TeX File
+    \item \emph{si-fda-refs.bib} : Manuscript Bib File
+    \item \emph{./figures}: All necessary figures
+\end{itemize}
+
 Thank you again for the positive news regarding our manuscript.  We look forward to seeing our work in the Journal of Computational and Graphical Statistics.
